### 2021-07-28 psy_peds.Rmd
first version of the R code to run the psy peds trend analysis

---
title: "Hospitalizations for psychiatric conditions among adolescents during the COVID-19 pandemic"
output:
  html_document:
    code_folding: show
    toc: yes
    toc_float: yes
---

```{r libraries, message=FALSE, warning=FALSE, results='hide'}
# Install packages
paket <- function(pak){
  new_pak <- pak[!(pak %in% rownames(installed.packages()))]
  if (length(new_pak)) 
    install.packages(new_pak, dependencies = TRUE,repos="https://cloud.r-project.org/")
  sapply(pak, library, character.only = TRUE)
}

listOfPackages <- c("tidyverse", "RColorBrewer", "knitr", "kableExtra")
paket(listOfPackages)
```

# R session information
```{r}
sessionInfo()
```

```{r}
# options(repr.plot.width = 18, repr.plot.height = 9)
theme_plots <- theme_bw() +
  theme(strip.text = element_text(size = 5),
        axis.text.x = element_text(size = 8), 
        axis.text.y = element_text(size = 6), 
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8),
        title = element_text(size = 10),
        plot.subtitle = element_text(size = 9, face = "italic")) 
theme_set(theme_plots)

# Colorblind palette
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

# Inclusion criteria
The inclusion criteria for this study is:
- Patients inclusion criteria: 
  - Age between 11-17 at the date of data collection (*currently, subject to change in a future version)
- Visits inclusion criteria:
  - Inpatients 
  - Starting date between January 1st, 2019 and July 31st, 2021
  - Comprising least one ICD-10 code pertaining to the specified ICD-10 codes list (mental health related codes)

# Variables that need to be checked/modified by each site 
Change the values of the following variables according to the specificities of your site:
- "folder_4ce_files": folder path where your phase 2.2 data files are located
- "obfuscation": determine the obfuscation threshold (FALSE if no obfuscation, or the numeric value of the obfuscation threshold if any)
- "raceAvailable": set as TRUE or FALSE depending on whether the variable is being collected at your site
- "dateFormat": specify the format of the date at your site (e.g., for "03-AUG-20", the format would be "%d-%b-%y", [see documentation](https://www.stat.berkeley.edu/~s133/dates.html))

```{r message=FALSE, warning=FALSE}
folder_4ce_files <- "/4ceData/Input"
obfuscation =  FALSE
#obfuscation = 3
raceAvailable = TRUE
#raceAvailable = FALSE
dateFormat <- "%d-%b-%y"
```

# Data input
We will use as input the 2.2 data files. Specifically: 
- LocalPatientSummary
- LocalPatientObservation
- LocalPatientClinicalcourse

For sites recording the race, we will also use an additional file:
- LocalPatientRace

## Read csv files

### 4CE phase 2.2 files

```{r message=FALSE, warning=FALSE}
### Read the CSV input files
source("R/readInputFiles.R")

files <- readInputFiles( path      = folder_4ce_files, 
                         separator = ",",
                         skip      = 0, 
                         verbose   = FALSE )

## Create the output folder if it doesn't exist
if (! "output" %in% list.dirs()) dir.create("output")

### Extract the patient summary and observation information. 
demo_raw <- files[[1]]
obs_raw <- files[[2]]
clinical_raw <- files[[3]]

### Read the file containing race information for those sites recording this variable
if( raceAvailable == TRUE ){
  race_raw <- read.delim(paste0( folder_4ce_files, "/LocalPatientRace.csv"), sep = ",", skip = 0)
}
```

### Read the ICD10 codes file
The file containing the ICD10 psychiatric related codes is located in the `public-data` folder of the GitHub repository.

```{r}
icdCodes <- read.csv("public-data/pediatric_psychiatric_ICD10_codes.csv", header = TRUE, colClasses = "character") %>% 
  rename(disorder_group = Mental.Health.Disorder.Group, 
         description = Description, 
         originalCode = ICD10_Code) %>%
  mutate( ICD10_Code= ifelse( nchar( originalCode) > 3, paste0(substr( originalCode, 1, 3), ".", substr(  originalCode, 4, nchar(originalCode))),originalCode)) %>%
  select( disorder_group, ICD10_Code, description)
```

# Data management
We will select only patients that have been diagnosed with at least one of the mental health ICD10 codes at any point during study followup. 
For these patients we will generate a table that will allow us to perform all the analysis required for this study. 
This transformed table will contain 9 columns:
- patient_num: patient internal identifier
- sex (male, female)
- age (in years)
- concept_type (DIAG-ICD10)
- concept_code (ICD-10 full code, e.g, R44.1)
- calendar_date (day-month-year, 30-SEP-20)
- hospitalization status (1, 0)
- n_hospitalization (the hospitalization number)
- len_hospitalization (length of each hospitalization)

The last two columns are not available in the input data files. We will generate those counts using the information available in the clinical raw file.

## Adding columns with hospitalization number and hospitalization length to the clinical raw file
```{r}
count_sequences_hospitalisation <- function(df, ...) {
  count_sequences <- rle(df$in_hospital)
  count_sequences_1 <- lapply(count_sequences, function(x) x[count_sequences$values == 1])
  n_sequences <- seq_along(count_sequences_1$lengths)
  sequences <- rep.int(n_sequences, count_sequences_1$lengths)
  sequences_len <- rep.int(count_sequences_1$lengths, count_sequences_1$lengths)
  df[df$in_hospital == 1, "n_hospitalisation"] <- sequences
  df[df$in_hospital == 1, "len_hospitalisation"] <- sequences_len
  return(df)
}
clinical_raw <- clinical_raw %>%
  group_by(patient_num) %>% 
  group_modify(count_sequences_hospitalisation)
```

## Creating the inputTable
Creation of the inputTable that will be the only table used throughout the rest of the analysis
- selecting only visits with at least one ICD codes pertaining to the mental health categories
- filter visits where patients are between 11 and 17 years old
- add the demographic and hospitalization related information 
- add the ICD code description and the related disorder group

We will also transform the observation dates, and break it by weeks, adding 2 additional columns: "weeks" and "year". Finally, we will dichotomize the study timeframe in two periods, based on cut-off dates, "before_pandemic", and "after_pandemic".

Additionally, we will also add the information about race to the table for sites collecting this information.

### Check that we have patients with the ICD10 mental codes
```{r}
### check that the ICD codes follow the regular expression [A-Z][0-9][0-9AB]\.?[0-9A-TV-Z]{0,4}
codesToReview <- obs_raw %>% dplyr::filter(concept_type == "DIAG-ICD10",
                                           ! grepl( "[A-Z][0-9][0-9AB]\\.?[0-9A-TV-Z]{0,4}", concept_code))
print(codesToReview)

### Estimate how many of the mental codes are in the dataset
mentalHealthCodes <- unique( obs_raw %>%
  dplyr::filter( concept_type == "DIAG-ICD10", 
                 concept_code %in% icdCodes$ICD10_Code ) %>%
  dplyr::mutate( concept_code = as.character( concept_code ), 
                 concept_length = nchar( concept_code ), 
                 concept_check = ifelse( concept_length == 4 | concept_length > 8, "Check", "OK")) %>%
  dplyr::select( concept_code, concept_length, concept_check) )

summary(as.factor( mentalHealthCodes$concept_check ) )

print( paste0("There are ", length(mentalHealthCodes$concept_code), " ICD10 mental codes out of the total ",length(unique(icdCodes$ICD10_Code)), " in the dataset"))
```

### Generating the table
```{r message=FALSE, warning=FALSE}
# Getting site name, will be used later in the analysis
site <- unique(as.character(demo_raw$siteid))

# cut-off dates used in the analysis 
history_start_date <- "2019-01-01"
history_end_date <- "2021-05-31"
pandemic_start_date <- "2020-03-15"
#early_pandemic_start_date <- "2020-03-15"
#late_pandemic_start_date <- "2021-01-01"

input_data <- obs_raw %>% 
  full_join( clinical_raw,  by = c("patient_num", "days_since_admission")) %>%
  full_join( demo_raw, by = "patient_num") %>%
  dplyr::mutate( date = as.Date( calendar_date, format = dateFormat ),
                 weeks = cut( date, breaks= "week"),
                 year = format( date, "%Y")) %>% 
  dplyr::filter( concept_type == "DIAG-ICD10", 
                 in_hospital == 1,
                 concept_code %in% icdCodes$ICD10_Code,
                 age < 18 & age > 10,
                 date >= history_start_date & date <= history_end_date) %>%
  # dplyr::mutate(period = ifelse( date < early_pandemic_start_date,
  #                                 "before_pandemic",
  #                                 ifelse(date < late_pandemic_start_date,
  #                                        "early_pandemic", 
  #                                        "late_pandemic"))) %>%
  dplyr::mutate(period = ifelse( date < pandemic_start_date,
                                  "before_pandemic","after_pandemic")) %>%
  left_join( icdCodes, by = c("concept_code" = "ICD10_Code")) %>%
  dplyr::select( patient_num, sex, age, concept_type, concept_code, calendar_date, in_hospital, n_hospitalisation, len_hospitalisation, 
                 date, weeks, year, period, disorder_group, description )
if( raceAvailable == TRUE ){
  input_data <- input_data %>%
    left_join( race_raw, by = "patient_num" )
}
```

# Data analysis

## Patients having presented history of mental health related issues before pandemic
We estimate the number of patients that presented a mental health disorder (grouped by mental health disorder groups - icdCodes$disorder_group):
- only before the pandemic
- only after the pandemic
- both before and after pandemic

The output is a barplot showing in the x-axis the number of patients, and in the y-axis the distinct disorder groups. Each disorder group has 3 bars associated based on the three previous options.

```{r}
patient_count_barplot_period <- input_data %>% 
  dplyr::group_by( disorder_group, patient_num ) %>% 
  dplyr::mutate( 
    when = ifelse( "before_pandemic" %in% period & "after_pandemic" %in% period,
                   "Before and After Pandemic",
                   ifelse( "before_pandemic" %in% period & ! "after_pandemic" %in% period,
                           "Only Before Pandemic",
                           "Only After Pandemic"
                   )
    ) 
    %>% factor(levels = c("Only Before Pandemic", "Before and After Pandemic", "Only After Pandemic"))
  ) %>%
  dplyr::ungroup() %>% 
  dplyr::select( patient_num, disorder_group, when ) %>%
  dplyr::group_by( disorder_group, when ) %>%
  dplyr::summarise( count = n_distinct( patient_num ) ) %>% 
  dplyr::mutate(count = ifelse(count > obfuscation | isFALSE(obfuscation),
                               count, 
                               0.5), 
                siteid = site)

patient_count_barplot_period %>%
  ggplot(aes(x = as.factor( disorder_group), y = count, fill = when)) +
  geom_bar(stat="identity", position = "dodge") +
  scale_fill_manual(values = cbPalette, name = "Period", guide = guide_legend(reverse = TRUE)) +
  labs(y = "Unique patients count",
       x = "Disorder group",
       title = "Patients with at least one visit", 
       subtitle = "According to visit period and pathology group", 
       legend = "Period")+
  theme(legend.position = "bottom", 
        legend.direction = "vertical",
        title = element_text(size=10),
        plot.subtitle = element_text(size = 8),
        legend.text = element_text(size = 8),
        axis.text.x = element_text(size =7, angle = 0, hjust = 0.5, vjust = 0)) + 
  coord_flip()

ggsave("output/disorder_before_after_pandemic_all.png", width = 18, height = 9, units="cm")
```

## Diagnoses count per week
Here we calculate the number of patients hospitalized per week (exactly, the number of patients being assigned a diagnosis on that week, hence usually corresponding to the end of hospitalization, ie not taking into account the length of the encounter). 
The barplot shows the weeks on the x-axis, and the patient counts on the y-axis. The horizontal lines is the (quadratic) temporal trend of these patient counts.
```{r}
patient_count_per_week <- input_data[, c("patient_num", "weeks")] %>%
  unique() %>%
  group_by(weeks) %>%
  tally() %>%
  mutate(n = ifelse( n < obfuscation | isFALSE(obfuscation),
                     n,
                     0.5),
         siteid = site )

#estimate the number of patients per week
patient_count_per_week %>%
  dplyr::mutate(weeks = as.Date(weeks)) %>% 
  dplyr::mutate(period = ifelse( weeks < pandemic_start_date,
                                  "before_pandemic",
                                  #ifelse(weeks < late_pandemic_start_date,
                                  #       "early_pandemic", 
                                  #       "late_pandemic")
                                 "after_pandemic")) %>% 
  ggplot(aes(x = weeks, y = n, fill = period)) +
  geom_bar(stat="identity")+
  scale_fill_manual(values = cbPalette) +
  labs(y = "Counts",
       x = "Calendar Weeks",
       title = "Weekly patient counts with mental health related ICD codes") + 
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)

ggsave("output/overall_weekly_count_all.png", width = 18, height = 9, units="cm")
```

## Temporal trend of hospitalized patients per day

```{r, }
# Creating encounter table with observation assigned to encounters
# 1. Creating encounters identifiers based on clinical_raw information about start and end date of an encounter --> n_hospitalisation
# 2. Assigning ICD codes to encounters (because they are assigned to the first or last day but usually are meaningful for the entire encounter)
encounters_obs <- clinical_raw %>% 
  filter(in_hospital == 1) %>% 
  select(patient_num, days_since_admission, n_hospitalisation) %>% 
  left_join(obs_raw[obs_raw$concept_type == "DIAG-ICD10",], on = c("patient_num", "days_since_admission")) %>% 
  drop_na(concept_type) %>% 
  distinct(patient_num, n_hospitalisation, concept_type, concept_code)

obs_by_day <- encounters_obs %>% 
  filter(concept_code %in% icdCodes$ICD10_Code) %>% 
  left_join(clinical_raw[clinical_raw$in_hospital == 1, c("patient_num", "n_hospitalisation", "calendar_date")], 
            on = c("patient_num", "n_hospitalisation")) %>% 
  group_by(calendar_date) %>% 
  summarise(patient_count = n_distinct(patient_num)) %>% 
  mutate(calendar_date = as.Date(calendar_date, format=dateFormat), 
         period = if_else(calendar_date < pandemic_start_date,
                          "Before Pandemic",
                          #if_else(calendar_date < late_pandemic_start_date,
                          #        "Early Pandemic", 
                          #        "Late Pandemic")
                          "After Pandemic"
         ), 
         patient_count = ifelse(patient_count > obfuscation | isFALSE(obfuscation), 
                                 patient_count, 
                                 0.5)
  )

ggplot(obs_by_day, aes(x = calendar_date,
                       y = patient_count,
                       color = period))  +
  geom_line() +
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  #geom_vline(xintercept = as.Date(late_pandemic_start_date),
  #           linetype = "dashed") +
  scale_color_manual(values = cbPalette) +
  labs(x = "Calendar date",
       y = "Patients count", 
       title = "Daily Hospitalized Patient Counts", 
       subtitle = "Hospitalisation encounters with at least one mental health diagnosis",
       legend = "Period")
ggsave("output/hospitalized_patients_all.png", width = 18, height = 9, units="cm")
```

### Patients count per week aggregated by sex
Same graphics as "Patient Count per Week", dichotomized by patients sex.
```{r}
patient_count_per_week_sex <- input_data[, c("patient_num", "sex","weeks")] %>%
  unique() %>%
  group_by(weeks, sex) %>%
  tally() %>%
  mutate(n = ifelse( n > obfuscation | isFALSE(obfuscation),
                     n, 
                     0.5),
         siteid = site)

patient_count_per_week_sex %>%
  ggplot(aes(x = as.Date(weeks), y = n, fill = sex)) +
  geom_bar(stat="identity", position = "dodge")+
  scale_fill_manual(values = cbPalette) +
  labs(y = "Counts",
       x = "Week",
       title = "Weekly patient counts with mental health related ICD codes aggregated by sex",
       subtitle = "Per Sex") + 
  theme(legend.position = "bottom") + 
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=sex), colour= "black", size = 0.5)
```

### Patients count per week aggregated by age group 
This graphic will only make sense if we include a wide range of age.

```{r}
#estimate the number of patients per week by age group
input_data[, c("patient_num", "age","weeks")] %>%
  unique() %>%
  mutate( age_group = ifelse( age < 15, "early adolescence (11-14 years)", "middle adolescence (15-17 years)")) %>%
  group_by(weeks, age_group) %>%
  tally() %>%
  mutate(n = ifelse( n > obfuscation, n, obfuscation * 0.5))  %>%
    ggplot(aes(x = as.Date(weeks), y = n, fill = age_group)) +
    geom_bar(stat="identity", position = "dodge")+
    labs(y = "Counts",
         x = "Week",
         title = "Weekly patient counts with mental health related ICD codes aggregated by age group",
         subtitle = "Per age group") + 
    theme(legend.position = "bottom") + 
  scale_fill_manual(values = cbPalette) +
    geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=sex), colour= "black", size = 0.5)

```

## Patients counts per week grouping by ICD subgroup category 
Here we estimate the number of patients hospitalized per week but we break it based on the distinct disorder groups. 
The output is a barplot showing in the x-axis the week, and in the y-axis the patient counts. Additionally a temporal trend line is added on top of the bars.
Each barplot represent a disorder group.

```{r}

patient_count_per_week_disorder_group <- input_data[input_data$in_hospital == 1,c("patient_num", "weeks", "disorder_group")] %>%
  unique() %>%
  group_by(weeks, disorder_group) %>% tally() %>%
  mutate(n = ifelse( n > obfuscation | isFALSE(obfuscation), n, 0.5), siteid = site)

patient_count_per_week_disorder_group %>%
  mutate(weeks = as.Date(weeks), 
         period = if_else(weeks < pandemic_start_date,
                          "Before Pandemic",
                          #if_else(weeks < late_pandemic_start_date,
                          #        "Early Pandemic", 
                          #        "Late Pandemic")
                          "After Pandemic"
         )
  ) %>% 
  ggplot(aes(x = as.Date(weeks), y= n, fill = period, group = disorder_group )) +
  geom_bar(stat = "identity")+
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  #geom_vline(xintercept = as.Date(late_pandemic_start_date),
  #           linetype = "dashed") +
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se= TRUE, aes(colour=period), colour= "black", size = 0.5) + 
  labs(x = "Week",
       y = "Patient Counts",
       title = "Weekly patient counts with mental health related ICD codes", 
       subtitle = "Grouped by mental health ICD subcategories") + 
  ylim(0, NA) + 
  facet_wrap(. ~ disorder_group, scales = "free_y") +
  theme(strip.text.x = element_text(size = 4),
        axis.text.x = element_text(size = 5, angle = 90), 
        axis.text.y = element_text(size = 6), 
        axis.title.x = element_text(size = 10),
        axis.title.y = element_text(size = 10),
        title = element_text(size = 12),
        plot.subtitle = element_text(size = 11),
        legend.position = "none") +
  scale_fill_manual(values = cbPalette[-1])
ggsave("output/categorical_counts_inpatient_teens_all.png", width = 18, height = 9, units="cm")
```

# Summary table 1
Here we will estimate the counts to populate table 1, with the counts for early and late pandemic, and a statistical test comparing both. 
This table one will include:
- sex
- age
- race (when available)
- hospitalization length
- diagnosis codes 
- diagnosis groups

```{r}
t.test2 <- function(m1, m2, s1, s2, n1, n2) {
  # t.test computed from sample statistics, ie using mean, standard deviation and count value of the samples
  # useful in case we can only retrieve aggregated stats from the sites
  # ref: https://stats.stackexchange.com/questions/30394/how-to-perform-two-sample-t-tests-in-r-by-inputting-sample-statistics-rather-tha 
  # and https://en.wikipedia.org/wiki/Student%27s_t-test?oldformat=true#Equal_or_unequal_sample_sizes,_unequal_variances_(sX1_%3E_2sX2_or_sX2_%3E_2sX1)
  se <- sqrt( (s1^2/n1) + (s2^2/n2) )
  # welch-satterthwaite df
  df <- ( (s1^2/n1 + s2^2/n2)^2 )/( (s1^2/n1)^2/(n1-1) + (s2^2/n2)^2/(n2-1) )
  t <- (m1-m2)/se 
  dat <- c(m1-m2, se, t, 2*pt(-abs(t),df))    
  names(dat) <- c("Difference of means", "Std Error", "t", "p-value")
  return(dat)
}
```

## Age
```{r}
age_stats <- input_data %>%
  dplyr::distinct(patient_num, period, age) %>%
  filter(age != 0) %>% 
  dplyr::group_by( period ) %>% 
  dplyr::summarise( mean = mean( age ), 
                    sd = sd( age ), 
                    count = n()) %>% 
  dplyr::mutate(count = ifelse( count > obfuscation, count, obfuscation * 0.5)) %>%
  dplyr::ungroup() %>% 
  dplyr::bind_rows(
    dplyr::summarise(input_data, period="Total", mean=mean(age), sd=sd(age), count = n())
  )

age_test <- t.test2(m1 = age_stats[[1, "mean"]], 
                    m2 = age_stats[[2, "mean"]], 
                    s1 = age_stats[[1, "sd"]], 
                    s2 = age_stats[[2, "sd"]], 
                    n1 = age_stats[[1, "count"]], 
                    n2 = age_stats[[2, "count"]]) %>% 
  `[[`("p-value")
```

## Sex
```{r}
sex <- input_data %>% 
  dplyr::group_by( sex ) %>%
  dplyr::summarise( count = n_distinct(patient_num) ) %>%
  dplyr::mutate( count = ifelse( count > obfuscation, count, obfuscation * 0.5), 
                 total = sum( count ), percentage = round( count/total*100, 2)  )

sex_by_period <- input_data %>% 
  group_by(period, sex ) %>% 
  summarise(count = n_distinct(patient_num)) %>%
  group_by(period, .add=TRUE) %>% 
  mutate( count = ifelse( count > obfuscation, count, obfuscation * 0.5),
          percentage=round(count/sum( count )*100, 2)) %>% 
  replace_na(list(male = 0, female = 0))

sex_test <- sex_by_period %>% 
  select(-percentage) %>% 
  pivot_wider(names_from = sex, values_from = count) %>% 
  ungroup() %>% 
  select(male, female) %>% 
  chisq.test() %>% 
  `[[`("p.value")
```

## Race
```{r}
if( raceAvailable == TRUE ){
  race <- input_data %>%
    dplyr::group_by( race_4ce ) %>%
    dplyr::summarise( count = n_distinct(patient_num) ) %>%
    dplyr::mutate( count = ifelse( count > obfuscation, count, obfuscation * 0.5), 
                   total = sum( count ), percentage = round( count/total*100, 2)  )

  race_by_period <- input_data %>%
    group_by(period, race_4ce ) %>% 
    summarise(count = n_distinct(patient_num)) %>%
    group_by(period, .add=TRUE) %>% 
    mutate( count = ifelse( count > obfuscation, count, obfuscation * 0.5),
            percentage=round(count/sum( count )*100, 2))

summary_race_test <- race_by_period %>% 
  select(-percentage) %>% 
  tidyr::pivot_wider(names_from = period,
                     values_from = count) %>%
  select(after_pandemic) %>% 
  tidyr::replace_na(list(after_pandemic = 0)) %>% 
  chisq.test() %>% 
  `[[`("p.value")
}
```

## Diagnosis

### By ICD code
```{r}
summary_diagnosis <- input_data %>%
  dplyr::distinct( patient_num, description, period, weeks, concept_code ) %>% 
  dplyr::group_by( concept_code, description, period ) %>%
  dplyr::count( patient_num ) %>% 
  dplyr::tally(name = "count") %>%
  dplyr::mutate(count = ifelse( count > obfuscation, count, obfuscation * 0.5)) %>%
  dplyr::arrange( desc(count) ) %>% 
  dplyr::ungroup()

summary_diagnosis_test <- summary_diagnosis %>% 
  tidyr::pivot_wider(names_from = period,
                     values_from = count) %>%
  #select(late_pandemic, early_pandemic) %>% 
  select(after_pandemic) %>%
  #tidyr::replace_na(list(late_pandemic = 0, early_pandemic = 0)) %>% 
  tidyr::replace_na(list(after_pandemic = 0)) %>% 
  chisq.test() %>% 
  `[[`("p.value")

```

### By disorder group
```{r}
summary_disorder_group <- input_data %>%
  dplyr::group_by( disorder_group ) %>%
  dplyr::count( patient_num ) %>% 
  dplyr::tally(name = "count") %>%
  dplyr::mutate(count = ifelse( count > obfuscation, count, obfuscation * 0.5)) %>%
  dplyr::arrange( desc( count ))

summary_disorder_group_by_period <- input_data %>% 
  dplyr::distinct( patient_num, disorder_group, period, weeks ) %>% 
  dplyr::group_by( period, disorder_group ) %>%
  summarise(count = n_distinct(patient_num)) %>%
  group_by(period, add=TRUE) %>%
  mutate(count = ifelse( count > obfuscation, count, obfuscation * 0.5), 
         percentage=round(count/sum( count )*100, 2))

summary_disorder_group_by_period_test <- summary_disorder_group_by_period %>% 
  pivot_wider(id_cols = disorder_group, names_from = period, values_from = c(count, percentage)) %>% 
  mutate(across(where(is.numeric), ~replace_na(.x, replace = 0))) %>% 
  #select(count_early_pandemic, count_late_pandemic) %>% 
  select(count_after_pandemic) %>%
  chisq.test() %>% 
  `[[`("p.value")

```

## Hospitalization

### Length of hospitalization 
```{r}
## Calculating also mean and sd in case we can only retrieve aggregated data from the sites
hospit_length <- input_data %>%
  dplyr::distinct( patient_num, n_hospitalisation, len_hospitalisation, period ) %>%
  dplyr::group_by( period ) %>% 
  dplyr::summarise( n_hospitalisations = n(),
                    length_hospt_median = median(len_hospitalisation), 
                    IQR = IQR( len_hospitalisation), 
                    length_hospt_mean = round(mean(len_hospitalisation), 2),
                    length_hospt_sd = round(sd(len_hospitalisation), 2))

hospit_length_mean_test <- t.test2(
  m1 = hospit_length[[1, "length_hospt_mean"]],
  m2 = hospit_length[[2, "length_hospt_mean"]],
  s1 = hospit_length[[1, "length_hospt_sd"]],
  s2 = hospit_length[[2, "length_hospt_sd"]],
  n1 = hospit_length[[1, "n_hospitalisations"]],
  n2 = hospit_length[[2, "n_hospitalisations"]]
)

```

### Repeat hospitalizations by patient
```{r}
hospitalisation_count <- input_data %>% 
  dplyr::distinct( patient_num, period, n_hospitalisation ) %>% 
  dplyr::group_by( period, patient_num ) %>%
  dplyr::filter(n_hospitalisation == max(n_hospitalisation)) %>%
  dplyr::group_by( period, add = FALSE) %>% 
  dplyr::summarise( mean_count_hospitalizations = mean(n_hospitalisation), 
                    max_count_hospitalizations = max(n_hospitalisation),
                    median_count_hospitalizations = median(n_hospitalisation), 
                    sd_hospitalisation_count = sd(n_hospitalisation),
                    IQR_hospitalisation_count = IQR( n_hospitalisation ))

repeated_hospitalisation <- input_data %>% 
  dplyr::distinct( patient_num, period, n_hospitalisation ) %>% 
  dplyr::group_by(patient_num, period) %>% 
  dplyr::filter(n_hospitalisation == max(n_hospitalisation)) %>% 
  dplyr::mutate(repeated_hospitalisation = if_else(n_hospitalisation > 1, "repeated", "non_repeated")) %>% 
  dplyr::group_by( period, add = FALSE ) %>% 
  dplyr::count( repeated_hospitalisation ) %>% 
  tidyr::pivot_wider(names_from = repeated_hospitalisation, values_from = n) %>% 
  replace_na(list("non_repeated" = 0, "repeated" = 0))
repeated_hospitalisation_test <- chisq.test(repeated_hospitalisation[, c("non_repeated", "repeated")]) %>% 
  `[[`("p.value")
```

## Pulling together Table 1 statistics
```{r}
format_counts <- function(count, percentage) {
  glue::glue("{count} ({percentage} %)", count = count, percentage = percentage)
}

sex_table1 <- sex_by_period %>% 
  bind_rows(
    sex_by_period %>%
      group_by(period) %>% 
      dplyr::summarise(sex = "Total", count = sum(count), percentage = sum(percentage))
  ) %>% 
  mutate(result = format_counts(count, percentage)) %>% 
  select(-c(count, percentage)) %>% 
  pivot_wider( names_from = period, values_from = result ) %>% 
  rename(name = sex)

if( raceAvailable == TRUE ){
  race_table1 <- race_by_period %>% 
  bind_rows(
    race_by_period %>%
      group_by(period) %>% 
      dplyr::summarise(race_4ce = "Total", count = sum(count), percentage = sum(percentage))
  ) %>% 
  mutate(result = format_counts(count, percentage)) %>% 
  select(-c(count, percentage)) %>% 
  pivot_wider( names_from = period, values_from = result ) %>% 
  rename(name = race_4ce)
}

diagnosis_table1 <- summary_diagnosis %>% 
    bind_rows(
    sex_by_period %>%
      group_by(period) %>% 
      dplyr::summarise(concept_code = "Total", count = sum(count))
  ) %>%
  pivot_wider( names_from = period, values_from = count) %>% 
  mutate(diag = paste(concept_code, description, sep = ": ")) %>% 
  select(diag, after_pandemic) %>% 
  rename(name = diag)

disorder_table1 <- summary_disorder_group_by_period %>% 
  bind_rows(
    dplyr::summarise(., disorder_group="Total", across(where(is.numeric), sum))
  ) %>% 
  mutate(result = format_counts(count, percentage)) %>% 
  select(-c(count, percentage)) %>% 
  pivot_wider( names_from = period, values_from = result ) %>% 
  # replace_na( list(early_pandemic = "0 (0 %)", 
  #                  late_pandemic = "0 (0 %)")) %>% 
    replace_na( list(after_pandemic = "0 (0 %)")) %>% 
  rename(name = disorder_group)

hospit_table1 <- hospit_length %>% 
  mutate(median = glue::glue("{median} ({IQR})", median = length_hospt_median, IQR = IQR), 
         mean = glue::glue("{mean} ({sd})", mean = length_hospt_mean, sd = length_hospt_sd), 
         n_hospitalisations = as.character(n_hospitalisations)) %>% 
  select(median, mean, n_hospitalisations, period) %>% 
  pivot_longer(cols = c("n_hospitalisations", "median", "mean")) %>% 
  pivot_wider(names_from = period, values_from = value)

repeated_hospit_table1 <- repeated_hospitalisation %>% 
  pivot_longer(cols = c("non_repeated", "repeated")) %>% 
  pivot_wider(names_from = period, values_from = value)

list_tests <- c("Sex" = sex_test, 
             "Hospitalisations length" = repeated_hospitalisation_test, 
             "Repeated hospitalisations" = repeated_hospitalisation_test,
             "Diagnoses" = summary_diagnosis_test, 
             "Disorder groups" = summary_disorder_group_by_period_test)

list_table1 <- list(
    "Sex" = sex_table1, 
    "Hospitalisations length" = hospit_table1,
    "Repeated hospitalisations" = repeated_hospit_table1,
    "Disorder groups" = disorder_table1, 
    "Diagnoses" = diagnosis_table1)
if (raceAvailable) {
  list_tests["Race"] = summary_race_test
  list_table1 = c(list("Race" = race_table1), 
          list_table1
  )
}

tests <- list_tests %>% 
  data.frame() %>% 
  round(2)
names(tests) <- "pvalue"
tests["Category"] <- row.names(tests)

table1 <- list_table1 %>% 
  lapply(function(df) mutate(df, across(.fns = as.character))) %>% 
  dplyr::bind_rows(.id = "Category") %>% 
  dplyr::left_join(tests, by = "Category")

table1 %>% 
  group_by(Category) %>%
  mutate(pvalue = ifelse(duplicated(pvalue, fromLast = TRUE), "", as.character(pvalue))) %>%
  ungroup() %>%
  mutate(Category = "") %>%
  mutate(across(.cols = c(before_pandemic, after_pandemic), .fns = replace_na, "0")) %>% 
  rename_with(str_to_title) %>%
  rename_with(sub, pattern = "_", replacement = " ") %>%
  mutate(across(.fns = str_to_title)) %>% 
  mutate(across(.fns = sub, pattern = "_", replacement = " ")) %>% 
  kable( booktabs = T) %>% 
  kableExtra::kable_styling(latex_options = "striped") %>% 
  kableExtra::group_rows(index = setNames(rle(table1$Category)[[1]], rle(table1$Category)[[2]]))
```

# Save the output in a RData file. 
```{r}

length_hospitalisation_values <- input_data %>% 
  dplyr::distinct( patient_num, n_hospitalisation, len_hospitalisation ) %>%
  dplyr::mutate( siteid = site ) %>%
  dplyr::select( len_hospitalisation, siteid )

save( patient_count_barplot_period,
      patient_count_per_week,
      patient_count_per_week_sex,
      patient_count_per_week_disorder_group,
      table1,
      length_hospitalisation_values,
      obs_by_day,
      file = paste0("output/peds_psy_aggResults_", unique(as.character(demo_raw$siteid)), ".RData"))
```

### 2021-07-28 R/readInputFiles.R

@@ -1,0 +1,92 @@
+#' Function to read the Phase 2.0 input files with patient-level data
+#'
+#' Given the \code{path} where the files are located it will generates
+#' a \code{list} with the \code{data.frame} object of the different files.
+#'
+#' @param path A path were the input files are located.
+#' @param separator The separator between columns (by default ",").
+#' @param skip In case the file does not start with the column names, add the number of lines that should be skipped at the beggining.
+#' @param verbose By default \code{FALSE}. Change it to \code{TRUE} to get an on-time log from the function.
+#' @return An object of class \code{list} with the \code{data.frames}.
+#' @examples
+#'
+#' dataSet <- readInputFiles(
+#'               path      = "./",
+#'               separator = ",",
+#'               skip      = 1,
+#'              )
+#' @export readInputFiles
+
+readInputFiles <- function( path, separator = ",", skip = 0, verbose = FALSE ){
+  
+  #set the working directory
+  setwd(path)
+  
+  #check that the input files needed are in the path 
+  if( verbose == TRUE){
+    print('Checking if the files are located in the directory provided')
+  }
+  
+  filesInDirectory <- list.files( path = "./")
+  
+  #for this analysis we need two of the files, patientSummary and patientObservation
+  checkPatientSummary <- ifelse( "localpatientsummary.csv" %in% tolower( filesInDirectory ), "OK", "File missing")
+  
+  if(checkPatientSummary == "OK"){
+    if(verbose== TRUE){
+      print("LocalPatientSummary.csv file in directory")
+    }
+  }else{
+    print("LocalPatientSummary.csv file not present in the directory. Please check if the file name and the 
+                                       directory are correct")
+    stop()
+  }
+  
+  checkPatientObservations <- ifelse( "localpatientobservations.csv" %in% tolower( filesInDirectory ), "OK", "File missing")
+  
+  if(checkPatientObservations == "OK"){
+    if(verbose== TRUE){
+      print("LocalPatientObservations.csv file in directory")
+    }
+  }else{
+    print("LocalPatientObservations.csv file not present in the directory. Please check if the file name and the 
+                                       directory are correct")
+    stop()
+  }
+  
+  checkPatientClinicalCourse <- ifelse( "localpatientclinicalcourse.csv" %in% tolower( filesInDirectory ), "OK", "File missing")
+  
[…]

### 2021-07-28 psy_peds.Rmd
correcting cohorts issue, and adding comparison plot psy/non-psy diagnoses

@@
 
+```{r}
+# cut-off dates used in the analysis, these should not be changed
+history_start_date <- "2019-01-01"
+history_end_date <- "2021-05-31"
+pandemic_start_date <- "2020-03-15"
+```
+
 # Data input
@@
 ### Extract the patient summary and observation information. 
-demo_raw <- files[[1]]
-obs_raw <- files[[2]]
-clinical_raw <- files[[3]]
+demo_raw <- files[["patientSummary"]]
+obs_raw <- files[["patientObservations"]]
+clinical_raw <- files[["patientClinicalCourse"]]
 
@@
 print( paste0("There are ", length(mentalHealthCodes$concept_code), " ICD10 mental codes out of the total ",length(unique(icdCodes$ICD10_Code)), " in the dataset"))
+
+### Joining the icd code description to the table 
+obs_raw <- obs_raw %>%
+  left_join(icdCodes, by = c("concept_code" = "ICD10_Code"))
+```
+
+### Data management for the patientClinicalCourse table
+```{r}
+clinical_raw <- clinical_raw %>% 
+  dplyr::mutate( date = as.Date( calendar_date, format = dateFormat ),
+                 weeks = as.Date(cut( date, breaks= "week")),
+                 year = format( date, "%Y"), 
+                 period = ifelse( date < pandemic_start_date,
+                                  "before_pandemic","after_pandemic")) %>% 
+  dplyr::filter(cohort  %in% c("AllAdm", "AllNotAdm"))
 ```
@@
 
-# cut-off dates used in the analysis 
-history_start_date <- "2019-01-01"
-history_end_date <- "2021-05-31"
-pandemic_start_date <- "2020-03-15"
 #early_pandemic_start_date <- "2020-03-15"
@@
 
-input_data <- obs_raw %>% 
-  full_join( clinical_raw,  by = c("patient_num", "days_since_admission")) %>%
-  full_join( demo_raw, by = "patient_num") %>%
-  dplyr::mutate( date = as.Date( calendar_date, format = dateFormat ),
-                 weeks = cut( date, breaks= "week"),
-                 year = format( date, "%Y")) %>% 
+all_codes <- obs_raw %>% 
+  full_join( clinical_raw,  by = c("patient_num", "cohort", "days_since_admission")) %>%
+  full_join( demo_raw, by = c("patient_num", "cohort")) %>%
   dplyr::filter( concept_type == "DIAG-ICD10", 
                  in_hospital == 1,
-                 concept_code %in% icdCodes$ICD10_Code,
                  age < 18 & age > 10,
                  date >= history_start_date & date <= history_end_date) %>%
-  # dplyr::mutate(period = ifelse( date < early_pandemic_start_date,
-  #                                 "before_pandemic",
-  #                                 ifelse(date < late_pandemic_start_date,
-  #                                        "early_pandemic", 
-  #                                        "late_pandemic"))) %>%
-  dplyr::mutate(period = ifelse( date < pandemic_start_date,
-                                  "before_pandemic","after_pandemic")) %>%
-  left_join( icdCodes, by = c("concept_code" = "ICD10_Code")) %>%
-  dplyr::select( patient_num, sex, age, concept_type, concept_code, calendar_date, in_hospital, n_hospitalisation, len_hospitalisation, 
-                 date, weeks, year, period, disorder_group, description )
+  dplyr::select( patient_num, cohort, sex, age, concept_type, concept_code, calendar_date,
+                 in_hospital, n_hospitalisation, len_hospitalisation, date, weeks, year,
+                 period, disorder_group, description )
+
+input_data <- all_codes %>% 
+  filter(concept_code %in% icdCodes$ICD10_Code)
+
 if( raceAvailable == TRUE ){
   input_data <- input_data %>%
-    left_join( race_raw, by = "patient_num" )
+    left_join( race_raw, by = c("patient_num", "cohort") )
 }
@@
 patient_count_barplot_period <- input_data %>% 
-  dplyr::group_by( disorder_group, patient_num ) %>% 
+  dplyr::group_by( disorder_group, patient_num, cohort) %>% 
   dplyr::mutate( 
@@
   dplyr::ungroup() %>% 
-  dplyr::select( patient_num, disorder_group, when ) %>%
+  dplyr::select( patient_num, cohort, disorder_group, when ) %>%
   dplyr::group_by( disorder_group, when ) %>%
-  dplyr::summarise( count = n_distinct( patient_num ) ) %>% 
+  dplyr::summarise( count = n_distinct( patient_num, cohort ) ) %>% 
   dplyr::mutate(count = ifelse(count > obfuscation | isFALSE(obfuscation),
@@
 ```{r}
-patient_count_per_week <- input_data[, c("patient_num", "weeks")] %>%
+patient_count_per_week <- input_data[, c("patient_num", "cohort", "weeks")] %>%
   unique() %>%
@@
                                   "before_pandemic",
-                                  #ifelse(weeks < late_pandemic_start_date,
-                                  #       "early_pandemic", 
-                                  #       "late_pandemic")
                                  "after_pandemic")) %>% 
@@
 
+```{r}
+patient_count_per_week <- all_codes %>% 
+  mutate(icd_code_category = ifelse(concept_code %in% icdCodes$ICD10_Code, 
+                               "psy", 
+                               "somatic")
+  ) %>% 
+  select(patient_num, cohort, weeks, icd_code_category, period) %>% 
+  unique() %>%
+  group_by(weeks, icd_code_category, period) %>%
+  tally() %>%
+  mutate(n = ifelse( n < obfuscation | isFALSE(obfuscation),
+                     n,
+                     0.5),
+         siteid = site )
+
+#estimate the number of patients per week
+patient_count_per_week %>%
+  ggplot(aes(x = weeks, y = n, fill = icd_code_category, color = icd_code_category)) +
+  geom_point() +
+  geom_line() +
+  geom_vline(xintercept = as.Date(pandemic_start_date),
+             linetype = "dashed") +
+  scale_fill_manual(values = cbPalette) +
+  scale_color_manual(values = cbPalette) +
+  labs(y = "Counts",
+       x = "Calendar Weeks",
+       title = "Weekly patient counts with mental health related ICD codes") + 
+  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)
+
+ggsave("output/overall_weekly_count_all_comparison.png", width = 18, height = 9, units="cm")
+```
+
 ## Temporal trend of hospitalized patients per day
@@
   filter(in_hospital == 1) %>% 
-  select(patient_num, days_since_admission, n_hospitalisation) %>% 
-  left_join(obs_raw[obs_raw$concept_type == "DIAG-ICD10",], on = c("patient_num", "days_since_admission")) %>% 
+  select(patient_num, cohort, days_since_admission, n_hospitalisation) %>% 
+  left_join(obs_raw[obs_raw$concept_type == "DIAG-ICD10",], on = c("patient_num", "cohort", "days_since_admission")) %>% 
   drop_na(concept_type) %>% 
-  distinct(patient_num, n_hospitalisation, concept_type, concept_code)
+  distinct(patient_num, cohort, n_hospitalisation, concept_type, concept_code)
 
@@
   filter(concept_code %in% icdCodes$ICD10_Code) %>% 
-  left_join(clinical_raw[clinical_raw$in_hospital == 1, c("patient_num", "n_hospitalisation", "calendar_date")], 
+  left_join(clinical_raw[clinical_raw$in_hospital == 1, c("patient_num", "cohort", "n_hospitalisation", "calendar_date")], 
             on = c("patient_num", "n_hospitalisation")) %>% 
   group_by(calendar_date) %>% 
-  summarise(patient_count = n_distinct(patient_num)) %>% 
+  summarise(patient_count = n_distinct(patient_num, cohort)) %>% 
   mutate(calendar_date = as.Date(calendar_date, format=dateFormat), 
@@
                           "Before Pandemic",
-                          #if_else(calendar_date < late_pandemic_start_date,
-                          #        "Early Pandemic", 
-                          #        "Late Pandemic")
                           "After Pandemic"
@@
 ```{r}
-patient_count_per_week_sex <- input_data[, c("patient_num", "sex","weeks")] %>%
+patient_count_per_week_sex <- input_data[, c("patient_num", "cohort", "sex","weeks")] %>%
   unique() %>%
@@
   geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=sex), colour= "black", size = 0.5)
+ggsave("output/overall_weekly_count_all.png", width = 18, height = 9, units="cm")
+
 ```
@@
 #estimate the number of patients per week by age group
-input_data[, c("patient_num", "age","weeks")] %>%
+input_data[, c("patient_num", "cohort", "age","weeks")] %>%
   unique() %>%
@@
     geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=sex), colour= "black", size = 0.5)
+ggsave("output/overall_weekly_count_age.png", width = 18, height = 9, units="cm")
 
@@
 
-patient_count_per_week_disorder_group <- input_data[input_data$in_hospital == 1,c("patient_num", "weeks", "disorder_group")] %>%
+patient_count_per_week_disorder_group <- input_data[input_data$in_hospital == 1, c("patient_num", "cohort", "weeks", "disorder_group")] %>%
   unique() %>%
@@
                           "Before Pandemic",
-                          #if_else(weeks < late_pandemic_start_date,
-                          #        "Early Pandemic", 
-                          #        "Late Pandemic")
                           "After Pandemic"
@@
 age_stats <- input_data %>%
-  dplyr::distinct(patient_num, period, age) %>%
+  dplyr::distinct(patient_num, cohort, period, age) %>%
   filter(age != 0) %>% 
@@
   dplyr::group_by( sex ) %>%
-  dplyr::summarise( count = n_distinct(patient_num) ) %>%
+  dplyr::summarise( count = n_distinct(patient_num, cohort) ) %>%
   dplyr::mutate( count = ifelse( count > obfuscation, count, obfuscation * 0.5), 
@@
   group_by(period, sex ) %>% 
-  summarise(count = n_distinct(patient_num)) %>%
+  summarise(count = n_distinct(patient_num, cohort)) %>%
   group_by(period, .add=TRUE) %>% 
@@
     dplyr::group_by( race_4ce ) %>%
-    dplyr::summarise( count = n_distinct(patient_num) ) %>%
+    dplyr::summarise( count = n_distinct(patient_num, cohort) ) %>%
     dplyr::mutate( count = ifelse( count > obfuscation, count, obfuscation * 0.5), 
@@
     group_by(period, race_4ce ) %>% 
-    summarise(count = n_distinct(patient_num)) %>%
+    summarise(count = n_distinct(patient_num, cohort)) %>%
     group_by(period, .add=TRUE) %>% 
@@
 summary_diagnosis <- input_data %>%
-  dplyr::distinct( patient_num, description, period, weeks, concept_code ) %>% 
+  dplyr::distinct( patient_num, cohort, description, period, weeks, concept_code ) %>% 
   dplyr::group_by( concept_code, description, period ) %>%
-  dplyr::count( patient_num ) %>% 
+  dplyr::count( patient_num, cohort ) %>% 
   dplyr::tally(name = "count") %>%
@@
   dplyr::group_by( disorder_group ) %>%
-  dplyr::count( patient_num ) %>% 
+  dplyr::count( patient_num, cohort ) %>% 
   dplyr::tally(name = "count") %>%
@@
 hospit_length <- input_data %>%
-  dplyr::distinct( patient_num, n_hospitalisation, len_hospitalisation, period ) %>%
+  dplyr::distinct( patient_num, cohort, n_hospitalisation, len_hospitalisation, period ) %>%
   dplyr::group_by( period ) %>% 
@@
 hospitalisation_count <- input_data %>% 
-  dplyr::distinct( patient_num, period, n_hospitalisation ) %>% 
-  dplyr::group_by( period, patient_num ) %>%
+  dplyr::distinct( patient_num, cohort, period, n_hospitalisation ) %>% 
+  dplyr::group_by( period, patient_num, cohort ) %>%
   dplyr::filter(n_hospitalisation == max(n_hospitalisation)) %>%
@@
 repeated_hospitalisation <- input_data %>% 
-  dplyr::distinct( patient_num, period, n_hospitalisation ) %>% 
-  dplyr::group_by(patient_num, period) %>% 
+  dplyr::distinct( patient_num, cohort, period, n_hospitalisation ) %>% 
+  dplyr::group_by(patient_num, cohort, period) %>% 
   dplyr::filter(n_hospitalisation == max(n_hospitalisation)) %>% 

### 2021-07-28 R/readInputFiles.R

@@ -17,63 +17,42 @@
 #'              )
 #' @export readInputFiles
 
-readInputFiles <- function( path, separator = ",", skip = 0, verbose = FALSE ){
-  
-  #set the working directory
-  setwd(path)
+readInputFiles <- function( path, separator = ",", skip = 0, verbose = FALSE, ... ){
   
   #check that the input files needed are in the path 
   if( verbose == TRUE){
     print('Checking if the files are located in the directory provided')
   }
   
-  filesInDirectory <- list.files( path = "./")
-  
-  #for this analysis we need two of the files, patientSummary and patientObservation
-  checkPatientSummary <- ifelse( "localpatientsummary.csv" %in% tolower( filesInDirectory ), "OK", "File missing")
-  
-  if(checkPatientSummary == "OK"){
-    if(verbose== TRUE){
-      print("LocalPatientSummary.csv file in directory")
+  filesInDirectory <- list.files(path = path)
+  fourcefile_names <- c(
+    "LocalPatientSummary.csv",
+    "LocalPatientObservations.csv",
+    "LocalPatientClinicalCourse.csv"
+  )
+  check_files <- fourcefile_names %in% filesInDirectory
+  if (all(check_files)) {
+    if (verbose) {
+      cat("All of ", fourcefile_names, " are in the directory")
     }
-  }else{
-    print("LocalPatientSummary.csv file not present in the directory. Please check if the file name and the 
-                                       directory are correct")
+  } else {
+    cat("Following files not found in the file directory:\n")
+    cat(fourcefile_names[!check_files])
+    cat("Please check if the file name and the directory are correct")
     stop()
   }
-  
-  checkPatientObservations <- ifelse( "localpatientobservations.csv" %in% tolower( filesInDirectory ), "OK", "File missing")
-  
-  if(checkPatientObservations == "OK"){
-    if(verbose== TRUE){
-      print("LocalPatientObservations.csv file in directory")
-    }
-  }else{
-    print("LocalPatientObservations.csv file not present in the directory. Please check if the file name and the 
-                                       directory are correct")
-    stop()
-  }
-  
-  checkPatientClinicalCourse <- ifelse( "localpatientclinicalcourse.csv" %in% tolower( filesInDirectory ), "OK", "File missing")
-  
-  if(checkPatientClinicalCourse == "OK"){
-    if(verbose== TRUE){
-      print("LocalPatientClinicalCourse.csv file in directory")
-    }
-  }else{
[…]

### 2021-08-04 psy_peds.Rmd
after the pandemic changed into during the pandermic; first graphic sorted in descending order

@@
 We will also transform the observation dates, and break it by weeks, adding 2 additional columns: "weeks" and "year".
-Finally, we will dichotomize the study timeframe in two periods, based on cut-off dates, "before_pandemic", and "after_pandemic".
+Finally, we will dichotomize the study timeframe in two periods, based on cut-off dates, "before_pandemic", and "during_pandemic".
 
@@
                  period = ifelse( date < pandemic_start_date,
-                                  "before_pandemic","after_pandemic")) %>% 
+                                  "before_pandemic","during_pandemic")) %>% 
   dplyr::filter(cohort  %in% c("AllAdm", "AllNotAdm"))
@@
 - only before the pandemic
-- only after the pandemic
-- both before and after pandemic
+- only during the pandemic
+- both before and during pandemic
 
@@
   dplyr::mutate( 
-    when = ifelse( "before_pandemic" %in% period & "after_pandemic" %in% period,
-                   "Before and After Pandemic",
-                   ifelse( "before_pandemic" %in% period & ! "after_pandemic" %in% period,
-                           "Only Before Pandemic",
-                           "Only After Pandemic"
+    when = ifelse( "before_pandemic" %in% period & "during_pandemic" %in% period,
+                   "Before and During Pandemic",
+                   ifelse( "before_pandemic" %in% period & ! "during_pandemic" %in% period,
+                           "Before the Pandemic",
+                           "During the Pandemic"
                    )
     ) 
-    %>% factor(levels = c("Only Before Pandemic", "Before and After Pandemic", "Only After Pandemic"))
+    %>% factor(levels = c("Before the Pandemic", "Before and During Pandemic", "During the Pandemic"))
   ) %>%
@@
 
+# create a vector with the disorder groups sorted in decreasing order 
+order_disorders <- patient_count_barplot_period %>% 
+  filter( when == "During the Pandemic") %>%
+  arrange( count ) %>%
+  select( disorder_group )
+
+patient_count_barplot_period$orderDisorder <- factor(patient_count_barplot_period$disorder_group,levels = as.factor( order_disorders$disorder_group ))
+
 patient_count_barplot_period %>%
-  ggplot(aes(x = as.factor( disorder_group), y = count, fill = when)) +
+  ggplot(aes(x = as.factor( orderDisorder), y = count, fill = when)) +
   geom_bar(stat="identity", position = "dodge") +
@@
 
-ggsave("output/disorder_before_after_pandemic_all.png", width = 18, height = 9, units="cm")
+ggsave("output/disorder_before_during_pandemic_all.png", width = 18, height = 9, units="cm")
+```
+
+### Relative to the total number of patients in each period
+
+```{r}
+patients_per_period <- input_data %>%
+  dplyr::group_by( period ) %>%
+  dplyr::summarise( count = n_distinct( patient_num ) ) %>% 
+  dplyr::mutate(count = ifelse(count > obfuscation | isFALSE(obfuscation),
+                               count, 
+                               0.5), 
+                siteid = site)
+
 ```
@@
                                   "before_pandemic",
-                                 "after_pandemic")) %>% 
+                                 "during_pandemic")) %>% 
   ggplot(aes(x = weeks, y = n, fill = period)) +
@@
          period = if_else(calendar_date < pandemic_start_date,
-                          "Before Pandemic",
-                          "After Pandemic"
+                          "Before the Start of the Pandemic",
+                          "After the Start of the Pandemic"
          ), 
@@
          period = if_else(weeks < pandemic_start_date,
-                          "Before Pandemic",
-                          "After Pandemic"
+                          "Before the Start of the Pandemic",
+                          "After the Start of the Pandemic"
          )
@@
                      values_from = count) %>%
-  select(after_pandemic) %>% 
-  tidyr::replace_na(list(after_pandemic = 0)) %>% 
+  select(during_pandemic) %>% 
+  tidyr::replace_na(list(during_pandemic = 0)) %>% 
   chisq.test() %>% 
@@
   #select(late_pandemic, early_pandemic) %>% 
-  select(after_pandemic) %>%
+  select(during_pandemic) %>%
   #tidyr::replace_na(list(late_pandemic = 0, early_pandemic = 0)) %>% 
-  tidyr::replace_na(list(after_pandemic = 0)) %>% 
+  tidyr::replace_na(list(during_pandemic = 0)) %>% 
   chisq.test() %>% 
@@
   #select(count_early_pandemic, count_late_pandemic) %>% 
-  select(count_after_pandemic) %>%
+  select(count_during_pandemic) %>%
   chisq.test() %>% 
@@
   mutate(diag = paste(concept_code, description, sep = ": ")) %>% 
-  select(diag, after_pandemic) %>% 
+  select(diag, during_pandemic) %>% 
   rename(name = diag)
@@
   #                  late_pandemic = "0 (0 %)")) %>% 
-    replace_na( list(after_pandemic = "0 (0 %)")) %>% 
+    replace_na( list(during_pandemic = "0 (0 %)")) %>% 
   rename(name = disorder_group)
@@
   mutate(Category = "") %>%
-  mutate(across(.cols = c(before_pandemic, after_pandemic), .fns = replace_na, "0")) %>% 
+  mutate(across(.cols = c(before_pandemic, during_pandemic), .fns = replace_na, "0")) %>% 
   rename_with(str_to_title) %>%

### 2021-08-18 psy_peds.Rmd
weekly count for distinct diseaasee groups distinct than mental

@@
 # cut-off dates used in the analysis, these should not be changed
-history_start_date <- "2019-01-01"
+history_start_date <- "2017-01-01"
 history_end_date <- "2021-05-31"
@@
 ggsave("output/overall_weekly_count_all_comparison.png", width = 18, height = 9, units="cm")
+```
+```{r}
+icdCategory <- read.delim("public-data/icd10Codes.txt", header = FALSE, colClasses = "character")
+colnames(icdCategory) <- c("icd_code_category", "icd_description_category")
+patient_count_per_week_byCategory <- all_codes %>% 
+  mutate( icd_code_category = ifelse(substr(concept_code, 1, 1) %in% c("D", "H"), substr(concept_code, 1, 2), substr(concept_code, 1, 1) ))
+
+patient_count_per_week_byCategory <- left_join( patient_count_per_week_byCategory, icdCategory) %>%
+  select(patient_num, cohort, weeks, icd_description_category, period) %>% 
+  unique() %>%
+  group_by(weeks, icd_description_category, period) %>%
+  tally() %>%
+  mutate(n = ifelse( n < obfuscation | isFALSE(obfuscation),
+                     n,
+                     0.5),
+         siteid = site )
+
+patient_count_per_week_byCategory %>%
+  filter( icd_description_category %in% c("Neoplams","Diseases of the respiratory system", "Diseases of the nervous system", "Diseases of the digestive system", "Diseases of the blood", "Congenital malformations, deformations and chromosomal abnormalities" )) %>%
+  ggplot(aes(x = weeks, y = n, fill = icd_description_category, color = icd_description_category)) +
+  geom_point() +
+  geom_line() +
+  geom_vline(xintercept = as.Date(pandemic_start_date),
+             linetype = "dashed") +
+  scale_fill_manual(values = cbPalette) +
+  scale_color_manual(values = cbPalette) +
+  labs(y = "Counts",
+       x = "Calendar Weeks",
+       title = "Weekly patient counts") + 
+  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)
+
+ggsave("output/overall_weekly_count_all_comparison_other_categories.png", width = 18, height = 9, units="cm")
+
 ```

### 2021-08-18 psy_peds.Rmd
weekly count for distinct diseaasee groups distinct than mental

@@
 ggsave("output/overall_weekly_count_all_comparison_other_categories.png", width = 18, height = 9, units="cm")
+
+patient_count_per_week_byCategory %>%
+  filter( icd_description_category %in% c("Neoplams","Diseases of the respiratory system", "Diseases of the nervous system", "Diseases of the digestive system", "Diseases of the blood", "Congenital malformations, deformations and chromosomal abnormalities" )) %>%
+  ggplot(aes(weeks, n, color = factor(icd_description_category))) +
+  geom_line() +
+  geom_vline(xintercept = as.Date(pandemic_start_date),
+             linetype = "dashed") +
+  geom_point() + 
+  theme(axis.text.x = element_text(angle=45, vjust = 1, hjust=1), legend.position = "none" ) +
+  labs(title = "Weekly patient counts",
+       subtitle = "(by ICD category)",
+       y = "Counts", x = "Calendar weeks") + 
+  facet_wrap(~ icd_description_category)
+
+ggsave("output/overall_weekly_count_all_comparison_other_categories_facet_wrap.png", width = 18, height = 9, units="cm")
 

### 2021-08-18 psy_peds.Rmd
adding percentage patients trends & icd_categories comparison

@@
   dplyr::mutate( date = as.Date( calendar_date, format = dateFormat ),
-                 weeks = as.Date(cut( date, breaks= "week")),
+                 weeks = as.Date(cut( date, breaks = "week")),
+                 month = as.Date(cut( date, breaks = "month")),
                  year = format( date, "%Y"), 
@@
   full_join( demo_raw, by = c("patient_num", "cohort")) %>% 
+  mutate(icd_code_category = ifelse(concept_code %in% icdCodes$ICD10_Code, 
+                               "psy", 
+                               "somatic")) %>% 
   dplyr::filter( concept_type == "DIAG-ICD10", 
@@
   dplyr::select( patient_num, cohort, sex, age, concept_type, concept_code, calendar_date,
-                 in_hospital, n_hospitalisation, len_hospitalisation, date, weeks, year,
-                 period, disorder_group, description )
+                 in_hospital, n_hospitalisation, len_hospitalisation, date, weeks, month, year,
+                 period, disorder_group, description, icd_code_category )
 
 input_data <- all_codes %>% 
-  filter(concept_code %in% icdCodes$ICD10_Code)
+  filter(concept_code %in% icdCodes$ICD10_Code) %>%
 
@@
 
-patient_count_barplot_period$orderDisorder <- factor(patient_count_barplot_period$disorder_group,levels = as.factor( order_disorders$disorder_group ))
+patient_count_barplot_period$orderDisorder <- factor(patient_count_barplot_period$disorder_group,
+                                                     levels = as.factor( order_disorders$disorder_group ))
 
@@
 The barplot shows the weeks on the x-axis, and the patient counts on the y-axis. The horizontal lines is the (quadratic) temporal trend of these patient counts.
+
 ```{r}
-patient_count_per_week <- input_data[, c("patient_num", "cohort", "weeks")] %>%
+all_codes[, c("patient_num", "cohort", "month")] 
+```
+
+```{r}
+patient_count_per_week <- all_codes[, c("patient_num", "cohort", "weeks", "icd_code_category")] %>%
   unique() %>%
-  group_by(weeks) %>%
+  group_by(weeks, icd_code_category) %>%
   tally() %>%
@@
 
+patient_count_per_month <- all_codes[, c("patient_num", "cohort", "month", "icd_code_category")] %>%
+  unique() %>%
+  group_by(month, icd_code_category) %>%
+  tally() %>%
+  mutate(n = ifelse( n < obfuscation | isFALSE(obfuscation),
+                     n,
+                     0.5),
+         siteid = site )
+
+```
+
+```{r}
+#estimate the number of patients per week
+patient_count_per_week %>%
+  tidyr::pivot_wider(names_from = icd_code_category, 
+                     values_from = n) %>% 
+  dplyr::mutate(weeks = as.Date(weeks)) %>% 
+  dplyr::mutate(period = ifelse( weeks < pandemic_start_date,
+                                  "before_pandemic",
+                                 "during_pandemic"), 
+                percentage = psy * 100 / (somatic + psy)) %>% 
+  ggplot(aes(x = weeks, y = percentage, fill = period)) +
+  # geom_bar(stat="identity")+
+  scale_fill_manual(values = cbPalette) +
+  labs(y = "Counts",
+       x = "Calendar Weeks",
+       title = "Percentage of patients with mental health related ICD codes (weekly)") + 
+  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)
+
+ggsave("output/overall_weekly_count_all.png", width = 18, height = 9, units="cm")
+```
+
+```{r}
 #estimate the number of patients per week
@@
                                  "during_pandemic")) %>% 
-  ggplot(aes(x = weeks, y = n, fill = period)) +
-  geom_bar(stat="identity")+
+  ggplot(aes(x = weeks, y = n, fill = icd_code_category)) +
+  # geom_bar(stat="identity")+
   scale_fill_manual(values = cbPalette) +
@@
        title = "Weekly patient counts with mental health related ICD codes") + 
-  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)
+  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=icd_code_category), colour= "black", size = 0.5)
 
@@
 patient_count_per_week <- all_codes %>% 
-  mutate(icd_code_category = ifelse(concept_code %in% icdCodes$ICD10_Code, 
-                               "psy", 
-                               "somatic")
-  ) %>% 
   select(patient_num, cohort, weeks, icd_code_category, period) %>% 
@@
 patient_count_per_week_byCategory <- left_join( patient_count_per_week_byCategory, icdCategory) %>%
+  mutate( icd_description_category = ifelse( concept_code %in% icdCodes$ICD10_Code, "psy", icd_description_category)) %>%
   select(patient_num, cohort, weeks, icd_description_category, period) %>% 
@@
 patient_count_per_week_byCategory %>%
-  filter( icd_description_category %in% c("Neoplams","Diseases of the respiratory system", "Diseases of the nervous system", "Diseases of the digestive system", "Diseases of the blood", "Congenital malformations, deformations and chromosomal abnormalities" )) %>%
+  filter( icd_description_category %in% c("psy","Neoplams","Diseases of the respiratory system", "Diseases of the nervous system", "Diseases of the digestive system", "Diseases of the blood", "Congenital malformations, deformations and chromosomal abnormalities" )) %>%
   ggplot(aes(x = weeks, y = n, fill = icd_description_category, color = icd_description_category)) +
-  geom_point() +
+  #geom_point(alpha=0.2) +
   geom_line() +
@@
        title = "Weekly patient counts") +   
+  theme(axis.text.x = element_text(angle=45, vjust = 1, hjust=1), legend.position = "bottom", legend.text = element_text(size=7), legend.title = element_text(size=7) ) +
   geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)

### 2021-08-18 psy_peds.Rmd
adding percentage plots

--- 2021-08-18 psy_peds.Rmd
@@
 
+icdCategory <- read.delim("public-data/icd10Codes.txt", header = FALSE, colClasses = "character")
+colnames(icdCategory) <- c("icd_code_letter", "icd_description_category")
+
 all_codes <- obs_raw %>% 
@@
   full_join( demo_raw, by = c("patient_num", "cohort")) %>% 
+  mutate(icd_code_category = ifelse(concept_code %in% icdCodes$ICD10_Code, 
+                               "psy", 
+                               "others")) %>% 
   dplyr::filter( concept_type == "DIAG-ICD10", 
@@
   dplyr::select( patient_num, cohort, sex, age, concept_type, concept_code, calendar_date,
-                 in_hospital, n_hospitalisation, len_hospitalisation, date, weeks, year,
-                 period, disorder_group, description )
+                 in_hospital, n_hospitalisation, len_hospitalisation, date, weeks, month, year,
+                 period, disorder_group, description, icd_code_category )
+
+all_codes <- all_codes %>% 
+  mutate( icd_code_letter = ifelse(substr(concept_code, 1, 1) %in% c("D", "H"), substr(concept_code, 1, 2), substr(concept_code, 1, 1) ))
+
+all_codes <- left_join( all_codes, icdCategory) %>%
+  mutate( icd_description_category = replace_na( ifelse( concept_code %in% icdCodes$ICD10_Code, "Psy", icd_description_category), "Others"))
 
@@
 The barplot shows the weeks on the x-axis, and the patient counts on the y-axis. The horizontal lines is the (quadratic) temporal trend of these patient counts.
+
 ```{r}
-patient_count_per_week <- input_data[, c("patient_num", "cohort", "weeks")] %>%
+patient_count_per_week_category <- all_codes[, c("patient_num", "cohort", "weeks", "icd_description_category")] %>%
   unique() %>%
-  group_by(weeks) %>%
+  group_by(weeks, icd_description_category) %>%
   tally() %>%
@@
 
-#estimate the number of patients per week
-patient_count_per_week %>%
-  dplyr::mutate(weeks = as.Date(weeks)) %>% 
-  dplyr::mutate(period = ifelse( weeks < pandemic_start_date,
-                                  "before_pandemic",
-                                 "during_pandemic")) %>% 
-  ggplot(aes(x = weeks, y = n, fill = period)) +
-  geom_bar(stat="identity")+
-  scale_fill_manual(values = cbPalette) +
-  labs(y = "Counts",
-       x = "Calendar Weeks",
-       title = "Weekly patient counts with mental health related ICD codes") + 
-  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)
+patient_count_per_week <- all_codes[, c("patient_num", "cohort", "weeks")] %>%
+  unique() %>%
+  group_by(weeks) %>%
+  tally( name = "total_patients") %>%
+  mutate(total_patients = ifelse( total_patients < obfuscation | isFALSE(obfuscation),
+                     total_patients,
+                     0.5),
+         siteid = site )
 
-ggsave("output/overall_weekly_count_all.png", width = 18, height = 9, units="cm")
+patient_count_per_month_category <- all_codes[, c("patient_num", "cohort", "month", "icd_description_category")] %>%
+  unique() %>%
+  group_by(month, icd_description_category) %>%
+  tally( name = "patients_per_category" ) %>%
+  mutate(patients_per_category = ifelse( patients_per_category < obfuscation | isFALSE(obfuscation),
+                     patients_per_category,
+                     0.5),
+         siteid = site )
+
+patient_count_per_month <- all_codes[, c("patient_num", "cohort", "month")] %>%
+  unique() %>%
+  group_by(month) %>%
+  tally( name = "total_patients") %>%
+  mutate(total_patients = ifelse( total_patients < obfuscation | isFALSE(obfuscation),
+                     total_patients,
+                     0.5),
+         siteid = site )
+
+patient_count_monthly <- merge( patient_count_per_month_category, patient_count_per_month)
 ```
@@
 ```{r}
-patient_count_per_week <- all_codes %>% 
-  mutate(icd_code_category = ifelse(concept_code %in% icdCodes$ICD10_Code, 
-                               "psy", 
-                               "somatic")
-  ) %>% 
-  select(patient_num, cohort, weeks, icd_code_category, period) %>% 
-  unique() %>%
-  group_by(weeks, icd_code_category, period) %>%
-  tally() %>%
-  mutate(n = ifelse( n < obfuscation | isFALSE(obfuscation),
-                     n,
-                     0.5),
-         siteid = site )
+#estimate the number of patients per week
+patient_count_monthly %>%
+  dplyr::filter(icd_description_category == "Psy") %>% 
+  dplyr::mutate(month = as.Date(month)) %>% 
+  dplyr::mutate(period = ifelse( month < pandemic_start_date,
+                                  "before_pandemic",
+                                 "during_pandemic"), 
+                percentage = patients_per_category * 100 / total_patients) %>% 
+  ggplot(aes(x = month, y = percentage, fill = period)) +
+  #geom_line()+
+  scale_fill_manual(values = cbPalette[c(3,4)]) +
+  labs(y = "Percentage",
+       x = "Calendar Months",
+       title = "Percentage of patients with mental health related ICD codes (monthly)") + 
+  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)
 
-#estimate the number of patients per week
-patient_count_per_week %>%
-  ggplot(aes(x = weeks, y = n, fill = icd_code_category, color = icd_code_category)) +
+ggsave("output/overall_monthly_count_all.png", width = 18, height = 9, units="cm")
+```
+
+```{r}
+patient_count_monthly %>%
+  filter( icd_description_category %in% c("Psy")) %>%
+  pivot_longer(names_to = "patients", cols=c(patients_per_category, total_patients)) %>%
+  mutate( patients = ifelse( patients == "patients_per_category", "Psy", "Total")) %>%
+  ggplot(aes(x = month, y = value, fill = patients, color = patients)) +
   geom_point() +
@@
   labs(y = "Counts",
-       x = "Calendar Weeks",
-       title = "Weekly patient counts with mental health related ICD codes") + 
+       x = "Calendar Months",
+       title = "Monthly patient counts with mental health related ICD codes") + 
   geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)
@@
 ```
+
 ```{r}
-icdCategory <- read.delim("public-data/icd10Codes.txt", header = FALSE, colClasses = "character")
-colnames(icdCategory) <- c("icd_code_category", "icd_description_category")
-patient_count_per_week_byCategory <- all_codes %>% 
-  mutate( icd_code_category = ifelse(substr(concept_code, 1, 1) %in% c("D", "H"), substr(concept_code, 1, 2), substr(concept_code, 1, 1) ))
-
-patient_count_per_week_byCategory <- left_join( patient_count_per_week_byCategory, icdCategory) %>%
-  select(patient_num, cohort, weeks, icd_description_category, period) %>% 
-  unique() %>%
-  group_by(weeks, icd_description_category, period) %>%
-  tally() %>%
-  mutate(n = ifelse( n < obfuscation | isFALSE(obfuscation),
-                     n,
-                     0.5),
-         siteid = site )
-
-patient_count_per_week_byCategory %>%
-  filter( icd_description_category %in% c("Neoplams","Diseases of the respiratory system", "Diseases of the nervous system", "Diseases of the digestive system", "Diseases of the blood", "Congenital malformations, deformations and chromosomal abnormalities" )) %>%
-  ggplot(aes(x = weeks, y = n, fill = icd_description_category, color = icd_description_category)) +
-  geom_point() +
+patient_count_monthly %>%
+  filter( icd_description_category %in% c("Psy","Neoplams","Diseases of the respiratory system","Diseases of the digestive system")) %>%
+  ggplot(aes(x = month, y = patients_per_category, fill = icd_description_category, color = icd_description_category)) +
+  #geom_point(alpha=0.2) +
   geom_line() +
@@
   labs(y = "Counts",
-       x = "Calendar Weeks",
-       title = "Weekly patient counts") + 
+       x = "Calendar Months",
+       title = "Months patient counts") +   
+  theme(axis.text.x = element_text(angle=45, vjust = 1, hjust=1), legend.position = "bottom", legend.text = element_text(size=7), legend.title = element_text(size=7) ) +
   geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)
 
-ggsave("output/overall_weekly_count_all_comparison_other_categories.png", width = 18, height = 9, units="cm")
-
+ggsave("output/overall_monthly_count_all_comparison_other_categories.png", width = 18, height = 9, units="cm")
 ```
@@
 
+###percentage
+patient_count_per_week_sex <- merge( patient_count_per_week_sex, patient_count_per_week )
+
+patient_count_per_week_sex %>%
+  mutate(percentage = n * 100 / total_patients ) %>%
+  ggplot(aes(x = as.Date(weeks), y = percentage, fill = sex)) +
+  geom_bar(stat="identity", position = "dodge")+
+  scale_fill_manual(values = cbPalette) +
+  labs(y = "Percentage",
+       x = "Week",
+       title = "Weekly patient percentage with mental health related ICD codes aggregated by sex",
+       subtitle = "Per Sex") + 
+  theme(legend.position = "bottom") + 
+  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=sex), colour= "black", size = 0.5)
+ggsave("output/overall_weekly_count_all.png", width = 18, height = 9, units="cm")
+
 ```
@@
 
-patient_count_per_week_disorder_group <- input_data[input_data$in_hospital == 1, c("patient_num", "cohort", "weeks", "disorder_group")] %>%
+patient_count_per_week_disorder_group <- input_data[, c("patient_num", "cohort", "weeks", "disorder_group")] %>%
   unique() %>%
-  group_by(weeks, disorder_group) %>% tally() %>%
-  mutate(n = ifelse( n > obfuscation | isFALSE(obfuscation), n, 0.5), siteid = site)
+  group_by(weeks, disorder_group) %>% tally( name = "psychiatric_patient_count") %>%
+  mutate(psychiatric_patient_count = ifelse( psychiatric_patient_count > obfuscation | isFALSE(obfuscation), psychiatric_patient_count, 0.5), siteid = site)
+
+patient_count_per_week_disorder_group <- merge( patient_count_per_week_disorder_group, patient_count_per_week)
 
@@
                           "After the Start of the Pandemic"
-         )
+         ), 
+         percentage = psychiatric_patient_count * 100 / total_patients
   ) %>% 
-  ggplot(aes(x = as.Date(weeks), y= n, fill = period, group = disorder_group )) +
+  ggplot(aes(x = as.Date(weeks), y= percentage, fill = period, group = disorder_group )) +
   geom_bar(stat = "identity")+
@@
   labs(x = "Week",
-       y = "Patient Counts",
-       title = "Weekly patient counts with mental health related ICD codes", 
+       y = "Patient Percentage",
+       title = "Weekly patient percentage with mental health related ICD codes", 
        subtitle = "Grouped by mental health ICD subcategories") + 

### 2021-08-28 psy_peds.Rmd
fixing hospitalization length count

@@
 count_sequences_hospitalisation <- function(df, ...) {
-  count_sequences <- rle(df$in_hospital)
+  # TODO: To be optimized to make it faster 
+  total_span <- data.frame(total_span = seq(min(df$days_since_admission),
+                    max(df$days_since_admission))
+  )
+  total_span <- left_join(total_span, df[,c("in_hospital", "days_since_admission")],
+            by = c("total_span" = "days_since_admission")
+  ) %>% replace_na(list(in_hospital = 0))
+  count_sequences <- rle(total_span$in_hospital)
   count_sequences_1 <- lapply(count_sequences, function(x) x[count_sequences$values == 1])
@@
 clinical_raw <- clinical_raw %>%
-  group_by(patient_num) %>% 
+  group_by(patient_num, cohort) %>% 
   group_modify(count_sequences_hospitalisation)
@@
                                   "before_pandemic","during_pandemic")) %>% 
-  dplyr::filter(cohort  %in% c("AllAdm", "AllNotAdm"))
+  dplyr::filter(cohort  %in% c("AllAdm"))
 ```

### 2021-09-08 psy_peds.Rmd
adding time series analysis

@@
 ```{r message=FALSE, warning=FALSE}
-folder_4ce_files <- "/4ceData/Input"
+folder_4ce_files <- "Input"
 obfuscation =  FALSE
@@
 
+```{r}
+# Adding filtering on in_hospital == 1 here to reduce subsequent compute time
+clinical_raw <- filter(clinical_raw, in_hospital == 1)
+```
+
 ### Read the ICD10 codes file
@@
 count_sequences_hospitalisation <- function(df, ...) {
-  # TODO: To be optimized to make it faster 
+  # TODO: To be optimized
   total_span <- data.frame(total_span = seq(min(df$days_since_admission),
                                             max(df$days_since_admission))
-  )
-  total_span <- left_join(total_span, df[,c("in_hospital", "days_since_admission")],
+  ) %>% 
+    left_join(df[,c("in_hospital", "days_since_admission")],
               by = c("total_span" = "days_since_admission")
@@
 
-patient_count_monthly <- merge( patient_count_per_month_category, patient_count_per_month)
+patient_count_monthly <- merge( patient_count_per_month_category, patient_count_per_month) %>%
+  dplyr::mutate(month = as.Date(month)) %>% 
+  dplyr::mutate(period = ifelse( month < pandemic_start_date,
+                                  "before_pandemic",
+                                 "during_pandemic"), 
+                percentage = patients_per_category * 100 / total_patients) 
 ```
@@
 patient_count_monthly %>% 
-  dplyr::filter(icd_description_category == "Psy") %>% 
-  dplyr::mutate(month = as.Date(month)) %>% 
-  dplyr::mutate(period = ifelse( month < pandemic_start_date,
-                                  "before_pandemic",
-                                 "during_pandemic"), 
-                percentage = patients_per_category * 100 / total_patients) %>% 
+  filter(icd_description_category == "Psy") %>% 
   ggplot(aes(x = month, y = percentage, fill = period)) +
@@
 ggsave("output/overall_monthly_count_all_comparison_other_categories.png", width = 18, height = 9, units="cm")
+```
+
+## Interrupted Time-Serie Analysis
+
+```{r} 
+# 
+patient_count_monthly_psy <- patient_count_monthly %>% 
+  filter(icd_description_category == "Psy", 
+         !month %in% c(as.Date("2019-01-01"), as.Date("2019-02-01")) & month != max(month)) %>% 
+  mutate(time = seq_along(month), 
+         month_year = format(month, "%m"))
+```
+
+```{r}
+# Model without seasonality
+time_serie_analysis_plotting <- function(model, patient_count_monthly_psy) {
+  summary(model)
+  output_model <- broom::tidy(model)
+  coeff_slope <- round(output_model[output_model$term == "periodduring_pandemic:time", "estimate"][[1]], 3)
+  pvalue_slope <- round(output_model[output_model$term == "periodduring_pandemic:time", "p.value"][[1]], 3)
+  annotation <- paste0("Lockdown slope: ", coeff_slope, " (pvalue: ", pvalue_slope, ")")
+  pred_model <- predict(model, patient_count_monthly_psy, type = "response", interval = "confidence") %>% as.data.frame()
+  pred_model_b <- predict(model, mutate(patient_count_monthly_psy, period = "before_pandemic"), type = "response", interval = "confidence") %>%
+    as.data.frame()
+  names(pred_model_b) <- paste0(names(pred_model_b), "_wo_intervention")
+
+  pred_model <- bind_cols(patient_count_monthly_psy, pred_model, pred_model_b)
+  plot <- ggplot(pred_model, aes(x = month, y = percentage)) + 
+    geom_point() +
+    geom_line(aes(x = month, y = fit, colour = "Mean pandemic")) +
+    geom_line(aes(x = month, y = fit_wo_intervention, colour = "Expected mean w/o pandemic"), linetype = "dashed") + 
+    scale_color_manual(name = "Time Series projection", values = c("Mean pandemic" = "black", "Expected mean w/o pandemic" = "red")) +
+    geom_ribbon(aes(ymin = lwr, ymax = upr), alpha = 0.3) +
+    # geom_ribbon(aes(ymin = lwr, ymax = upr), alpha = 0.3) +
+    geom_vline(xintercept = as.Date(pandemic_start_date), linetype = "dashed") +
+    annotate("text", x = as.Date("2019-10-01"), y = max(pred_model$percentage) + 5, label = annotation) +
+    theme(legend.position = "bottom")
+  return(plot)
+}
+```
+
+```{r}
+# Time series analysis of percentage of psy codes per encounter, linear (without seasonality)
+ts_linear_percent <- lm(percentage ~ period * time, patient_count_monthly_psy)
+time_serie_analysis_plotting(ts_linear_percent, patient_count_monthly_psy) + 
+    labs(title = "Interrupted Time-Serie Analysis: Percentage of encounters with psychiatric ICD codes", 
+         subtitle = "Without seasonality effect",
+         x = "Percentage", 
+         y = "Month")
+ggsave("output/time_series_percent_linear.png", width = 18, height = 9, units="cm")
+
+```
+
+```{r}
+# Time series analysis of percentage of psy codes per encounter, with seasonality
+ts_seasonal_percent <- lm(percentage ~ period*time + harmonic(month_year, 2, 12), patient_count_monthly_psy)
+time_serie_analysis_plotting(ts_seasonal_percent, patient_count_monthly_psy) +
+  labs(title = "Interrupted Time-Serie Analysis: Percentage of encounters with psychiatric ICD codes", 
+       subtitle = "With seasonality effect",
+       x = "Percentage", 
+       y = "Month")
+ggsave("output/time_series_percent_seasonal.png", width = 18, height = 9, units="cm")
+
+```
+
+```{r}
+# Time series analysis of percentage of icd code count, with seasonality
+ts_count_seasonal <- glm(patients_per_category ~ period*time + harmonic(month_year, 1, 12),
+                       family = quasipoisson,
+                       patient_count_monthly_psy)
+
+output_ts_seasonal_count <- broom::tidy(ts_count_seasonal)
+coeff_slope <- round(output_ts_seasonal_count[output_ts_seasonal_count$term == "periodduring_pandemic:time", "estimate"][[1]], 3)
+pvalue_slope <- round(output_ts_seasonal_count[output_ts_seasonal_count$term == "periodduring_pandemic:time", "p.value"][[1]], 3)
+annotation <- paste0("Lockdown slope: ", coeff_slope, " (pvalue: ", pvalue_slope, ")")
+pred_ts_seasonal_count <- predict(ts_count_seasonal, patient_count_monthly_psy, type = "response") %>% data.frame(fit = .)
+pred_ts_seasonal_count_b <- predict(ts_count_seasonal, mutate(patient_count_monthly_psy, period = "before_pandemic"), type = "response")%>% data.frame(fit = .)
+names(pred_ts_seasonal_count_b) <- paste0(names(pred_ts_seasonal_count_b), "_wo_intervention")
+
+pred_ts_seasonal_count <- bind_cols(patient_count_monthly_psy, pred_ts_seasonal_count, pred_ts_seasonal_count_b)
+ggplot(pred_ts_seasonal_count, aes(x = month, y = patients_per_category)) + 
+  geom_point() +
+  geom_line(aes(x = month, y = fit, colour = "Mean pandemic")) +
+  geom_line(aes(x = month, y = fit_wo_intervention, colour = "Expected mean w/o pandemic"), linetype = "dashed") + 
+  scale_color_manual(name = "Time Series projection", values = c("Mean pandemic" = "black", "Expected mean w/o pandemic" = "red")) +
+  geom_vline(xintercept = as.Date(pandemic_start_date), linetype = "dashed") +
+  annotate("text", x = as.Date("2019-10-01"), y = max(pred_ts_seasonal_count$patients_per_category) + 5, label = annotation) +
+  labs(title = "Interrupted Time-Serie Analysis: Count of encounters with psychiatric ICD codes", 
+       subtitle = "With seasonality effect",
+       x = "Percentage", 
+       y = "Month") + 
+  theme(legend.position = "bottom")
+ggsave("output/time_series_count_seasonal.png", width = 18, height = 9, units="cm")
 ```

### 2021-09-08 R/readInputFiles.R

@@ -18,12 +18,12 @@
 #' @export readInputFiles
 
 readInputFiles <- function( path, separator = ",", skip = 0, verbose = FALSE, ... ){
-  
-  #check that the input files needed are in the path 
+
+  #check that the input files needed are in the path
   if( verbose == TRUE){
     print('Checking if the files are located in the directory provided')
   }
-  
+
   filesInDirectory <- list.files(path = path)
   fourcefile_names <- c(
     "LocalPatientSummary.csv",
@@ -38,37 +38,38 @@ readInputFiles <- function( path, separator = ",", skip = 0, verbose = FALSE, ..
   } else {
     cat("Following files not found in the file directory:\n")
     cat(fourcefile_names[!check_files])
+    cat("\n")
     cat("Please check if the file name and the directory are correct")
     stop()
   }
 
-  #read the files 
+  #read the files
   if( verbose == TRUE){
     cat( 'Reading \n')
     cat(fourcefile_names)
     cat(" files")
   }
-  
+
   read_delim_4ce <- function(file_name, ...) read.delim(file.path(path, file_name), sep = separator, skip = skip, ...)
   patientSummary <- read_delim_4ce("LocalPatientSummary.csv")
   patientObservations <- read_delim_4ce("LocalPatientObservations.csv")
   patientClinicalCourse <- read_delim_4ce("LocalPatientClinicalCourse.csv")
-  
+
   if( verbose == TRUE){
     print( paste0( "LocalPatientsummary file contains: ", nrow( patientSummary ), " rows and ", ncol( patientSummary ), " columns."))
     print( paste0( "LocalPatientobservation file contains: ", nrow( patientObservations ), " rows and ", ncol( patientObservations ), " columns."))
     print( paste0( "LocalPatientClinicalCourse file contains: ", nrow( patientClinicalCourse ), " rows and ", ncol( patientClinicalCourse ), " columns."))
   }
-  
+
   #return it as a list
   files <- list("patientSummary"        = patientSummary,
                 "patientObservations"   = patientObservations,
                 "patientClinicalCourse" = patientClinicalCourse
   )
-  
+
   if( verbose == TRUE){
     print( "A list with the three data.frames read is being generated")
   }
-  
+
   return( files )
 }

### 2021-09-09 psy_peds.Rmd
adjusted age

@@
 
-listOfPackages <- c("tidyverse", "RColorBrewer", "knitr", "kableExtra")
+listOfPackages <- c("tidyverse", "RColorBrewer", "knitr", "kableExtra", "tsModel")
 paket(listOfPackages)
@@
 The last two columns are not available in the input data files. We will generate those counts using the information available in the clinical raw file.
+
+```{r}
+# Adjusting patient age to the time of visit, assuming the age used is the patients' age at the time of extraction (last discharge date available in the data)
+last_data_date <- max(as.Date(demo_raw$last_discharge_date, "%Y-%m-%d"))
+clinical_raw <- left_join(clinical_raw, demo_raw[c("patient_num", "cohort", "age")], by = c("patient_num", "cohort"))
+clinical_raw$age_time_visit <- clinical_raw$age - floor(as.numeric(last_data_date - as.Date(clinical_raw$calendar_date, dateFormat)) / 365.25)
+
+```
 
@@
 
-### Estimate how many of the mental codes are in the dataset
+### Calculate how many of the mental codes are in the dataset
 mentalHealthCodes <- unique( obs_raw %>%

### 2021-09-09 psy_peds.Rmd
optimizing hospitalisation sequences count

@@
 ```{r}
-# Adjusting patient age to the time of visit, assuming the age used is the patients' age at the time of extraction (last discharge date available in the data)
+# Adjusting patient age at the time of visit, assuming the age used is the patients' age at the time of extraction (last discharge date available in the data)
 last_data_date <- max(as.Date(demo_raw$last_discharge_date, "%Y-%m-%d"))
@@
 clinical_raw$age_time_visit <- clinical_raw$age - floor(as.numeric(last_data_date - as.Date(clinical_raw$calendar_date, dateFormat)) / 365.25)
-
 ```
@@
 count_sequences_hospitalisation <- function(df, ...) {
-  # TODO: To be optimized
-  total_span <- data.frame(total_span = seq(min(df$days_since_admission),
+  seq_hospitalisation_df <- data.frame(total_span = seq(min(df$days_since_admission),
                                                         max(df$days_since_admission))
   ) %>%
-    left_join(df[,c("in_hospital", "days_since_admission")],
-              by = c("total_span" = "days_since_admission")
-    ) %>% replace_na(list(in_hospital = 0))
-  count_sequences <- rle(total_span$in_hospital)
+    left_join(df, by = c("total_span" = "days_since_admission")) %>%
+    replace_na(list(in_hospital = 0))
+  count_sequences <- rle(seq_hospitalisation_df$in_hospital)
   count_sequences_1 <- lapply(count_sequences, function(x) x[count_sequences$values == 1])
@@
   sequences_len <- rep.int(count_sequences_1$lengths, count_sequences_1$lengths)
-  df[df$in_hospital == 1, "n_hospitalisation"] <- sequences
-  df[df$in_hospital == 1, "len_hospitalisation"] <- sequences_len
-  return(df)
+  stopifnot(length(df$days_since_admission) == length(sequences))
+  data.frame(days_since_admission = df$days_since_admission,
+             n_hospitalisation = sequences,
+             len_hospitalisation = sequences_len)
 }
-clinical_raw <- clinical_raw %>%
+stopifnot(all(clinical_raw$in_hospital == 1))
+hospitalisations_seq_df <- clinical_raw %>%
+  distinct(patient_num, cohort, days_since_admission, in_hospital) %>%
   group_by(patient_num, cohort) %>%
   group_modify(count_sequences_hospitalisation)
+clinical_raw <- left_join(clinical_raw,
+                          hospitalisations_seq_df,
+                          by = c("patient_num", "cohort", "days_since_admission"))
 ```
@@
   mutate(diag = paste(concept_code, description, sep = ": ")) %>% 
-  select(diag, during_pandemic) %>% 
+  select(diag, before_pandemic, during_pandemic) %>% 
   rename(name = diag)
@@
   pivot_wider( names_from = period, values_from = result ) %>% 
-  # replace_na( list(early_pandemic = "0 (0 %)", 
-  #                  late_pandemic = "0 (0 %)")) %>% 
   replace_na( list(during_pandemic = "0 (0 %)")) %>% 

### 2021-09-13 psy_peds.Rmd
fixing age bug

@@
 clinical_raw$age_time_visit <- clinical_raw$age - floor(as.numeric(last_data_date - as.Date(clinical_raw$calendar_date, dateFormat)) / 365.25)
+clinical_raw <- select(clinical_raw, -age)
 ```

### 2021-09-13 psy_peds.Rmd
read the most updated ICD file codes

@@
 ```{r}
-icdCodes <- read.csv("public-data/pediatric_psychiatric_ICD10_codes.csv", header = TRUE, colClasses = "character") %>% 
-  rename(disorder_group = Mental.Health.Disorder.Group, 
-         description = Description, 
-         originalCode = ICD10_Code) %>%
-  mutate( ICD10_Code= ifelse( nchar( originalCode) > 3, paste0(substr( originalCode, 1, 3), ".", substr(  originalCode, 4, nchar(originalCode))),originalCode)) %>%
-  select( disorder_group, ICD10_Code, description)
+#icdCodes <- read.csv("public-data/pediatric_psychiatric_ICD10_codes.csv", header = TRUE, colClasses = "character") %>% 
+#  rename(disorder_group = Mental.Health.Disorder.Group, 
+#         description = Description, 
+#         originalCode = ICD10_Code) %>%
+#  mutate( ICD10_Code= ifelse( nchar( originalCode) > 3, paste0(substr( originalCode, 1, 3), ".", substr(  originalCode, 4, nchar(originalCode))),originalCode)) %>%
+#  select( disorder_group, ICD10_Code, description)
+
+icdCodes <- read.csv("public-data/pediatric_psychiatric_ICD10_codesV2.csv", header = TRUE, colClasses = "character") %>%
+   filter( flag %in% c("originalCodes", "FrenchSpecificCodes")) %>%
+   select( disorder_group, ICD10_Code, description )
 ```

### 2021-09-14 psy_peds.Rmd
Correcting history start date

@@
 # cut-off dates used in the analysis, these should not be changed
-history_start_date <- "2017-01-01"
+history_start_date <- "2019-01-01"
 history_end_date <- "2021-05-31"

### 2021-09-14 psy_peds.Rmd
Update psy_peds.Rmd

@@
 history_start_date <- "2019-01-01"
-history_end_date <- "2021-05-31"
+history_end_date <- "2021-08-31"
 pandemic_start_date <- "2020-03-15"

### 2021-09-15 psy_peds.Rmd
correcting stats based on age at the time of visits

@@
                  in_hospital == 1,
-                 age < 18 & age > 10,
+                 age_time_visit < 18 & age_time_visit > 10,
                  date >= history_start_date & date <= history_end_date) %>%
-  dplyr::select( patient_num, cohort, sex, age, concept_type, concept_code, calendar_date,
+  dplyr::select( patient_num, cohort, sex, age, age_time_visit, concept_type, concept_code, calendar_date,
                  in_hospital, n_hospitalisation, len_hospitalisation, date, weeks, month, year,
@@
 #estimate the number of patients per week by age group
-input_data[, c("patient_num", "cohort", "age","weeks")] %>%
+input_data[, c("patient_num", "cohort", "age_time_visit","weeks")] %>%
   unique() %>%
-  mutate( age_group = ifelse( age < 15, "early adolescence (11-14 years)", "middle adolescence (15-17 years)")) %>%
+  mutate( age_group = ifelse( age_time_visit < 15, "early adolescence (11-14 years)", "middle adolescence (15-17 years)")) %>%
   group_by(weeks, age_group) %>%
@@
 age_stats <- input_data %>%
-  dplyr::distinct(patient_num, cohort, period, age) %>%
-  filter(age != 0) %>% 
+  dplyr::group_by(patient_num, cohort, period) %>% 
+  dplyr::filter(age_time_visit == max(age_time_visit)) %>%
+  dplyr::ungroup() %>% 
+  dplyr::distinct(patient_num, cohort, period, age_time_visit) %>% 
+  dplyr::filter(age_time_visit != 0) %>% 
   dplyr::group_by( period ) %>% 
-  dplyr::summarise( mean = mean( age ), 
-                    sd = sd( age ), 
+  dplyr::summarise( mean = mean( age_time_visit ), 
+                    sd = sd( age_time_visit ), 
                     count = n()) %>% 
@@
   dplyr::bind_rows(
-    dplyr::summarise(input_data, period="Total", mean=mean(age), sd=sd(age), count = n())
+    dplyr::summarise(input_data,
+                     period="Total",
+                     mean = mean(age_time_visit),
+                     sd = sd(age_time_visit),
+                     count = n())
   )

### 2021-09-15 psy_peds.Rmd
correcting time series analysis plots

@@
          subtitle = "Without seasonality effect",
-         x = "Percentage", 
-         y = "Month")
+         x = "Month", 
+         y = "Percentage")
 ggsave("output/time_series_percent_linear.png", width = 18, height = 9, units="cm")
@@
 # Time series analysis of percentage of psy codes per encounter, with seasonality
-ts_seasonal_percent <- lm(percentage ~ period*time + harmonic(month_year, 2, 12), patient_count_monthly_psy)
+ts_seasonal_percent <- lm(percentage ~ period*time + harmonic(month_year, 1, 12), patient_count_monthly_psy)
 time_serie_analysis_plotting(ts_seasonal_percent, patient_count_monthly_psy) +
@@
        subtitle = "With seasonality effect",
-       x = "Percentage", 
-       y = "Month")
+       x = "Month", 
+       y = "Percentage")
 ggsave("output/time_series_percent_seasonal.png", width = 18, height = 9, units="cm")
@@
        subtitle = "With seasonality effect",
-       x = "Percentage", 
-       y = "Month") + 
+       x = "Month", 
+       y = "Count") + 
   theme(legend.position = "bottom")

### 2021-09-28 psy_peds.Rmd
new statistical analysis markdown

--- 2021-09-14 psy_peds.Rmd
@@
 ```{r message=FALSE, warning=FALSE}
-folder_4ce_files <- "Input"
+folder_4ce_files <- "/Input/09212021_2/"
 obfuscation =  FALSE
@@
 history_start_date <- "2019-01-01"
-history_end_date <- "2021-08-31"
+history_end_date <- "2021-05-31"
 pandemic_start_date <- "2020-03-15"
@@
 
-files <- readInputFiles( path      = folder_4ce_files, 
+files <- readInputFiles( path      = "../Input/09212021_2/", 
                          separator = ",",
@@
 ## Create the output folder if it doesn't exist
-if (! "output" %in% list.dirs()) dir.create("output")
+#if (! "output" %in% list.dirs()) dir.create("output")
 
@@
 if( raceAvailable == TRUE ){
-  race_raw <- read.delim(paste0( folder_4ce_files, "/LocalPatientRace.csv"), sep = ",", skip = 0)
+  race_raw <- read.delim(paste0("../", folder_4ce_files, "/LocalPatientRace.csv"), sep = ",", skip = 0)
 }
@@
                           by = c("patient_num", "cohort", "days_since_admission"))
+
+new_hospitalizations_per_day <- clinical_raw %>%
+group_by(patient_num, n_hospitalisation) %>% slice(which.min(as.Date(calendar_date, dateFormat)))
+
+newhosp <- new_hospitalizations_per_day %>%
+    filter( in_hospital == 1, 
+          as.Date( calendar_date, dateFormat ) < '2021-05-01' &
+           as.Date( calendar_date, dateFormat ) > '2019-03-01' ) %>%
+    group_by( calendar_date) %>%
+  summarise(count = n_distinct(patient_num))
+
+ggplot(newhosp, aes(x = as.Date(calendar_date, dateFormat),
+                       y = count))  +
+  geom_point() +
+  scale_x_date(date_breaks = "1 month", date_labels =  "%b %Y") +
+  theme(axis.text.x=element_text(angle=60, hjust=1))+
+  labs(x = "Calendar date",
+       y = "Patients count", 
+       title = "New daily admissions (2.2 data)")
+
+xx <- clinical_raw[ , .SD[which.min(as.Date(calendar_date, dateFormat))], by = c(n_hospitalization, patient_num)]
+
 ```

### 2021-09-28 psy_peds_2.Rmd
new statistical analysis markdown

---
title: "Hospitalizations for psychiatric conditions among adolescents during the COVID-19 pandemic"
output:
  html_document:
    code_folding: show
    toc: yes
    toc_float: yes
---

```{r libraries, message=FALSE, warning=FALSE, results='hide'}
# Install packages
paket <- function(pak){
  new_pak <- pak[!(pak %in% rownames(installed.packages()))]
  if (length(new_pak)) 
    install.packages(new_pak, dependencies = TRUE,repos="https://cloud.r-project.org/")
  sapply(pak, library, character.only = TRUE)
}

listOfPackages <- c("tidyverse", "RColorBrewer", "knitr", "kableExtra", "tsModel")
paket(listOfPackages)
```

# R session information
```{r}
sessionInfo()
```

```{r}
# options(repr.plot.width = 18, repr.plot.height = 9)
theme_plots <- theme_bw() +
  theme(strip.text = element_text(size = 5),
        axis.text.x = element_text(size = 8), 
        axis.text.y = element_text(size = 6), 
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8),
        title = element_text(size = 10),
        plot.subtitle = element_text(size = 9, face = "italic")) 
theme_set(theme_plots)

# Colorblind palette
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

# Inclusion criteria
The inclusion criteria for this study is:
- Patients inclusion criteria: 
  - Age between 11-17 at the date of visit
- Visits inclusion criteria:
  - Inpatients (at least one hospitalization > 1 day)
  - Starting date between January 1st, 2019 and July 31st, 2021
  - Comprising least one ICD-10 code pertaining to the specified ICD-10 codes list (mental health related codes)

# Variables that need to be checked/modified by each site 
Change the values of the following variables according to the specificities of your site:
- "folder_4ce_files": folder path where your phase 2.2 data files are located
- "obfuscation": determine the obfuscation threshold (FALSE if no obfuscation, or the numeric value of the obfuscation threshold if any)
- "raceAvailable": set as TRUE or FALSE depending on whether the variable is being collected at your site
- "dateFormat": specify the format of the date at your site (e.g., for "03-AUG-20", the format would be "%d-%b-%y", [see documentation](https://www.stat.berkeley.edu/~s133/dates.html))

```{r message=FALSE, warning=FALSE}
folder_4ce_files <- "/Input/09212021_2/"
obfuscation =  FALSE
#obfuscation = 3
raceAvailable = TRUE
#raceAvailable = FALSE
dateFormat <- "%d-%b-%y"
min_hosp_days <- 2
```

```{r}
# cut-off dates used in the analysis, these should not be changed
history_start_date <- "2019-01-01"
history_end_date <- "2021-05-31"
pandemic_start_date <- "2020-03-15"
```

# Data input
We will use as input the 2.2 data files. Specifically: 
- LocalPatientSummary
- LocalPatientObservation
- LocalPatientClinicalcourse

For sites recording the race, we will also use an additional file:
- LocalPatientRace

## Read csv files

### 4CE phase 2.2 files

```{r message=FALSE, warning=FALSE}
### Read the CSV input files
source("R/readInputFiles.R")

files <- readInputFiles( path      = "../Input/09212021_2/", 
                         separator = ",",
                         skip      = 0, 
                         verbose   = FALSE )

## Create the output folder if it doesn't exist
#if (! "output" %in% list.dirs()) dir.create("output")

### Extract the patient summary and observation information. 
demo_raw <- files[["patientSummary"]]
obs_raw <- files[["patientObservations"]]
clinical_raw <- files[["patientClinicalCourse"]]

### Read the file containing race information for those sites recording this variable
if( raceAvailable == TRUE ){
  race_raw <- read.delim(paste0("../", folder_4ce_files, "/LocalPatientRace.csv"), sep = ",", skip = 0)
}
```

```{r}
# Adding filtering on in_hospital == 1 here to reduce subsequent compute time
clinical_raw <- filter(clinical_raw, in_hospital == 1)
```

### Adding columns with hospitalization number and hospitalization length to the clinical raw file
```{r}
count_sequences_hospitalisation <- function(df, ...) {
  seq_hospitalisation_df <- data.frame(total_span = seq(min(df$days_since_admission),
                                                        max(df$days_since_admission))
  ) %>%
    left_join(df, by = c("total_span" = "days_since_admission")) %>%
    replace_na(list(in_hospital = 0))
  count_sequences <- rle(seq_hospitalisation_df$in_hospital)
  count_sequences_1 <- lapply(count_sequences, function(x) x[count_sequences$values == 1])
  n_sequences <- seq_along(count_sequences_1$lengths)
  sequences <- rep.int(n_sequences, count_sequences_1$lengths)
  sequences_len <- rep.int(count_sequences_1$lengths, count_sequences_1$lengths)
  stopifnot(length(df$days_since_admission) == length(sequences))
  data.frame(days_since_admission = df$days_since_admission,
             n_hospitalisation = sequences,
             len_hospitalisation = sequences_len)
}
stopifnot(all(clinical_raw$in_hospital == 1))
hospitalisations_seq_df <- clinical_raw %>%
  distinct(patient_num, cohort, days_since_admission, in_hospital) %>%
  group_by(patient_num, cohort) %>%
  group_modify(count_sequences_hospitalisation)

clinical_raw <- left_join(clinical_raw,
                          hospitalisations_seq_df,
                          by = c("patient_num", "cohort", "days_since_admission"))
```

### Read the ICD10 codes file
The file containing the ICD10 psychiatric related codes is located in the `public-data` folder of the GitHub repository.

```{r}
#icdCodes <- read.csv("public-data/pediatric_psychiatric_ICD10_codes.csv", header = TRUE, colClasses = "character") %>% 
#  rename(disorder_group = Mental.Health.Disorder.Group, 
#         description = Description, 
#         originalCode = ICD10_Code) %>%
#  mutate( ICD10_Code= ifelse( nchar( originalCode) > 3, paste0(substr( originalCode, 1, 3), ".", substr(  originalCode, 4, nchar(originalCode))),originalCode)) %>%
#  select( disorder_group, ICD10_Code, description)

icdCodes <- read.csv("public-data/pediatric_psychiatric_ICD10_codesV2.csv", header = TRUE, colClasses = "character") %>%
   filter( flag %in% c("originalCodes", "FrenchSpecificCodes")) %>%
   select( disorder_group, ICD10_Code, description )
```

## All cohort
We are selecting encounters based on the following criteria:
- associated with at least 1 ICD codes
- with a duration of hospitalization spanning over at least two consecutive days
- Starting on or after 2019/01/01

```{r}
# Adjusting patient age at the time of visit, assuming the age used is the patients' age at the time of extraction (last discharge date available in the data)
last_data_date <- max(as.Date(demo_raw$last_discharge_date, "%Y-%m-%d"))
clinical_raw <- left_join(clinical_raw, demo_raw[c("patient_num", "cohort", "age", "sex")], by = c("patient_num", "cohort"))
clinical_raw$age_time_visit <- clinical_raw$age - floor(as.numeric(last_data_date - as.Date(clinical_raw$calendar_date, dateFormat)) / 365.25)
clinical_raw <- select(clinical_raw, -age)

clinical_raw_filter <- clinical_raw %>%
  filter( len_hospitalisation > min_hosp_days) %>%
  dplyr::mutate( date = as.Date( calendar_date, format = dateFormat ),
                 weeks = as.Date(cut( date, breaks = "week")),
                 month = as.Date(cut( date, breaks = "month")),
                 year = format( date, "%Y"), 
                 period = ifelse( date < pandemic_start_date,
                                  "before_pandemic","during_pandemic")) %>% 
  dplyr::filter(cohort  %in% c("AllAdm"))
stopifnot(unique(all_codes$in_hospital) == 1)
stopifnot(unique(all_codes$len_hospitalisation > min_hosp_days))
```

```{r}
all_codes <- clinical_raw_filter %>% 
  left_join( obs_raw,  by = c("patient_num", "cohort", "days_since_admission")) %>%
  filter( concept_type == 'DIAG-ICD10')
```

## First estimation: total number of patients without filtering by any disease subtype

### New hospitalizations
```{r, fig.width=12, fig.height=12}
new_hospitalizations_per_day <- all_codes %>%
  group_by(patient_num, n_hospitalisation) %>% slice(which.min(date)) %>%
  filter( in_hospital == 1, 
         date < '2021-05-01' &
         date >= '2019-03-01' ) %>%
    group_by( date, weeks, month) %>%
  summarise(count = n_distinct(patient_num)) %>% 
  ungroup()

new_hospitalizations_per_week <- new_hospitalizations_per_day %>%
  filter(weeks >= '2019-03-01') %>% 
  group_by( weeks) %>%
  summarise(count = sum(count))

new_hospitalizations_per_month3 <- new_hospitalizations_per_day %>%
  group_by( month) %>%
  summarise(count = sum(count))

nh_day <- ggplot(new_hospitalizations_per_day,
                 aes(x = date, y = count))  +
  geom_point() +
  scale_x_date(date_breaks = "1 month", date_labels =  "%b %Y") +
  theme(axis.text.x=element_text(angle=60, hjust=1))+
  labs(x = "Calendar date",
       y = "Patients count", 
       title = "New admissions (per day)")

nh_week <- ggplot(new_hospitalizations_per_week, aes(x = weeks,
                       y = count))  +
  geom_point() +
  scale_x_date(date_breaks = "1 month", date_labels =  "%b %Y") +
  theme(axis.text.x=element_text(angle=60, hjust=1))+
  labs(x = "Calendar date",
       y = "Patients count", 
       title = "New admissions (per week)")

nh_month <- ggplot(new_hospitalizations_per_month, aes(x = month,
                       y = count))  +
  geom_point() +
  scale_x_date(date_breaks = "1 month", date_labels =  "%b %Y") +
  theme(axis.text.x=element_text(angle=60, hjust=1))+
  labs(x = "Calendar date",
       y = "Patients count", 
       title = "New admissions (per month)")

require(gridExtra)
grid.arrange(nh_day,
             nh_week,
             nh_month,
             ncol=1)
```

### Total hospitalize patients
```{r, fig.height=12, fig.width=12}
total_hospitalizations_per_day <- all_codes %>%
  group_by(patient_num, date ) %>%
  filter(date < '2021-05-01' &
         date >= '2019-03-01' ) %>%
    group_by( date, weeks, month) %>%
  summarise(count = n_distinct(patient_num)) %>% 
  ungroup()

total_hospitalizations_per_week <- total_hospitalizations_per_day %>% 
  filter(weeks > '2019-03-01') %>% 
    group_by( weeks ) %>%
  summarise(count = sum(count))

total_hospitalizations_per_month <- total_hospitalizations_per_day %>%
  group_by( month ) %>%
  summarise(count = sum(count))

th_day <- ggplot(total_hospitalizations_per_day, aes(x = date,
                       y = count))  +
  geom_line() +
  scale_x_date(date_breaks = "1 month", date_labels =  "%b %Y") +
  theme(axis.text.x=element_text(angle=60, hjust=1))+
  labs(x = "Calendar date",
       y = "Patients count", 
       title = "Total admissions (per day)")

th_week <- ggplot(total_hospitalizations_per_week, aes(x = weeks,
                       y = count))  +
  geom_line() +
  scale_x_date(date_breaks = "1 month", date_labels =  "%b %Y") +
  theme(axis.text.x=element_text(angle=60, hjust=1))+
  labs(x = "Calendar date",
       y = "Patients count", 
       title = "Total admissions (per week)")

th_month <- ggplot(total_hospitalizations_per_month, aes(x = month,
                       y = count))  +
  geom_line() +
  scale_x_date(date_breaks = "1 month", date_labels =  "%b %Y") +
  theme(axis.text.x=element_text(angle=60, hjust=1))+
  labs(x = "Calendar date",
       y = "Patients count", 
       title = "Currently hospitalized patients (per month)")

require(gridExtra)
grid.arrange(th_day, th_week, th_month, ncol=1)
```

## Second estimation: total number of patients filtering by psychiatric disorders
## Creating the inputTable
Creation of the inputTable that will be the only table used throughout the rest of the analysis
- selecting only visits with at least one ICD codes pertaining to the mental health categories
- filter visits where patients are between 11 and 17 years old
- add the demographic and hospitalization related information 
- add the ICD code description and the related disorder group

### Check that we have patients with the ICD10 mental codes
```{r}
### check that the ICD codes follow the regular expression [A-Z][0-9][0-9AB]\.?[0-9A-TV-Z]{0,4}
codesToReview <- obs_raw %>% dplyr::filter(concept_type == "DIAG-ICD10",
                                           ! grepl( "[A-Z][0-9][0-9AB]\\.?[0-9A-TV-Z]{0,4}", concept_code))
print(codesToReview)

### Calculate how many of the mental codes are in the dataset
mentalHealthCodes <- unique( all_codes %>%
  dplyr::filter( concept_type == "DIAG-ICD10", 
                 concept_code %in% icdCodes$ICD10_Code ) %>%
  dplyr::mutate( concept_code = as.character( concept_code ), 
                 concept_length = nchar( concept_code ), 
                 concept_check = ifelse( concept_length == 4 | concept_length > 8, "Check", "OK")) %>%
  dplyr::select( concept_code, concept_length, concept_check) )

summary(as.factor( mentalHealthCodes$concept_check ) )

print( paste0("There are ", length(mentalHealthCodes$concept_code), " ICD10 mental codes out of the total ",length(unique(icdCodes$ICD10_Code)), " in the dataset"))
```

### Add the mental ICD code description and filter 
```{r message=FALSE, warning=FALSE}
# Getting site name, will be used later in the analysis
site <- unique(as.character(demo_raw$siteid))

icdCategory <- read.delim("public-data/icd10Codes.txt", header = FALSE, colClasses = "character")
colnames(icdCategory) <- c("icd_code_letter", "icd_description_category")

all_codes_desc <- all_codes %>% 
    left_join( icdCodes, by=c("concept_code" = "ICD10_Code") ) %>%
  mutate(icd_code_category = ifelse(concept_code %in% icdCodes$ICD10_Code, 
                               "psy", 
                               "others")) %>%
  dplyr::select( patient_num, cohort, sex, age_time_visit, concept_type, concept_code, calendar_date,
                 in_hospital, n_hospitalisation, len_hospitalisation, date, weeks, month, year,
                 period, disorder_group, description, icd_code_category )

all_codes_desc <- all_codes_desc %>% 
  mutate( icd_code_letter = ifelse(substr(concept_code, 1, 1) %in% c("D", "H"), substr(concept_code, 1, 2), substr(concept_code, 1, 1) ))

all_codes_desc <- left_join( all_codes_desc, icdCategory) %>%
  mutate( icd_description_category = replace_na( ifelse( concept_code %in% icdCodes$ICD10_Code, "Psy", icd_description_category), "Others"))

input_data <- all_codes_desc %>% 
  filter(concept_code %in% icdCodes$ICD10_Code)

if( raceAvailable == TRUE ){
  input_data <- input_data %>%
    left_join( race_raw, by = c("patient_num", "cohort") )
}
```

### 2021-09-28 psy_peds_2.Rmd
updating currently hospitalized patient counts

@@
 total_hospitalizations_per_day <- all_codes %>%
-  group_by(patient_num, date ) %>%
   filter(date < '2021-05-01' &
          date >= '2019-03-01' ) %>%
-    group_by( date, weeks, month) %>%
-  summarise(count = n_distinct(patient_num)) %>% 
-  ungroup()
+    group_by( date ) %>%
+  summarise(count = n_distinct(patient_num))
 
-total_hospitalizations_per_week <- total_hospitalizations_per_day %>% 
-  filter(weeks > '2019-03-01') %>% 
+total_hospitalizations_per_week <- all_codes %>%
+  filter(date < '2021-05-01' &
+         date >= '2019-03-01' ) %>%
     group_by( weeks ) %>%
-  summarise(count = sum(count))
+  summarise(count = n_distinct(patient_num))
 
-total_hospitalizations_per_month <- total_hospitalizations_per_day %>%
+total_hospitalizations_per_month <- all_codes %>%
+  filter(date < '2021-05-01' &
+         date >= '2019-03-01' ) %>%
     group_by( month ) %>%
-  summarise(count = sum(count))
+  summarise(count = n_distinct(patient_num))
 

### 2021-09-28 psy_peds_2.Rmd
new daily graphics added

@@
   dplyr::filter(cohort  %in% c("AllAdm"))
-stopifnot(unique(all_codes$in_hospital) == 1)
-stopifnot(unique(all_codes$len_hospitalisation > min_hosp_days))
+
 ```
@@
   filter( concept_type == 'DIAG-ICD10')
+stopifnot(unique(all_codes$in_hospital) == 1)
+stopifnot(unique(all_codes$len_hospitalisation > min_hosp_days))
 ```
@@
 
-new_hospitalizations_per_month3 <- new_hospitalizations_per_day %>%
+new_hospitalizations_per_month <- new_hospitalizations_per_day %>%
   group_by( month) %>%
@@
 ```
+
+#### Psychiatric related conditions vs. any other condition
+
+##### New admissions
+
+##### Total hospitalizations
+We have estimated the total number of hospitalizations in the previous steps:
+- total_hospitalization_per_day
+- total_hospitalization_per_week
+- total_hospitalization_per_month
+
+We will use these counts to estimate the percentage and to compare total counts, vs psychiatric related conditions.
+
+```{r}
+total_hospitalizations_per_day_psy <- input_data %>%
+  filter(date < '2021-05-01' &
+         date >= '2019-03-01' ) %>%
+    group_by( date ) %>%
+  summarise(count_psy = n_distinct(patient_num)) %>%
+  left_join( total_hospitalizations_per_day ) %>%
+  mutate( period = ifelse( date < pandemic_start_date,
+                                  "before_pandemic",
+                                 "during_pandemic"), 
+                percentage = count_psy * 100 / count)
+
+thpd_psyPer <- total_hospitalizations_per_day_psy %>%
+  ggplot(aes(x = date, y = percentage, fill = period)) +
+  #geom_line()+
+  scale_fill_manual(values = cbPalette[c(3,4)]) +
+  labs(y = "Percentage",
+       x = "Date (by day)",
+       title = "Percentage of patients with mental health related ICD codes (daily)") + 
+  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)
+
+thpd_psyCount <- total_hospitalizations_per_day_psy %>%
+  pivot_longer(names_to = "counts", cols=c(count_psy, count)) %>%
+  mutate( counts = ifelse( counts == "count_psy", "Psy", "Total")) %>%
+  ggplot(aes(x = date, y = value, fill = counts, color = counts)) +
+  geom_point() +
+  geom_line() +
+  geom_vline(xintercept = as.Date(pandemic_start_date),
+             linetype = "dashed") +
+  scale_fill_manual(values = cbPalette) +
+  scale_color_manual(values = cbPalette) +
+  labs(y = "Counts",
+       x = "Date (by day)",
+       title = "Daily patient counts with mental health related ICD codes") + 
+  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)
+
+```
+
+We will compare the psychiatric conditions with some other disorders (e.g, digestive, respiratory, etc.)
+We will use these counts to estimate the percentage and to compare total counts, vs psychiatric related conditions.
+
+```{r}
+total_hospitalizations_per_day_disorder_group <- all_codes_desc %>%
+  filter(date < '2021-05-01' &
+         date >= '2019-03-01' & 
+        icd_description_category %in% c("Psy","Neoplams","Diseases of the respiratory system","Diseases of the digestive system" ))%>%
+    group_by( date, icd_description_category ) %>%
+  summarise(counts = n_distinct(patient_num))
+
+thpd_disorderGroup <- total_hospitalizations_per_day_disorder_group %>%
+  ggplot(aes(x = date, y = counts, fill = icd_description_category, color = icd_description_category)) +
+  #geom_point(alpha=0.2) +
+  geom_line() +
+  geom_vline(xintercept = as.Date(pandemic_start_date),
+             linetype = "dashed") +
+  scale_fill_manual(values = cbPalette) +
+  scale_color_manual(values = cbPalette) +
+  labs(y = "patient number",
+       x = "days",
+       title = "Total number of hospitalized patients (daily)") +   
+  theme(axis.text.x = element_text(angle=45, vjust = 1, hjust=1), legend.position = "bottom", legend.text = element_text(size=7), legend.title = element_text(size=7) ) +
+  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)

### 2021-09-28 psy_peds_2.Rmd
reformating

--- 2021-09-28 psy_peds_2.Rmd
@@
 ```{r message=FALSE, warning=FALSE}
-folder_4ce_files <- "/Input/09212021_2/"
+folder_4ce_files <- "Input/09212021_2/"
 obfuscation =  FALSE
@@
 pandemic_start_date <- "2020-03-15"
+start_date_plots <- "2019-03-01"
+end_date_plots <- "2021-05-01"
 ```
@@
 
-files <- readInputFiles( path      = "../Input/09212021_2/", 
+files <- readInputFiles( path      = folder_4ce_files, 
                          separator = ",",
@@
 if( raceAvailable == TRUE ){
-  race_raw <- read.delim(paste0("../", folder_4ce_files, "/LocalPatientRace.csv"), sep = ",", skip = 0)
+  race_raw <- read.delim(file.path(folder_4ce_files, "/LocalPatientRace.csv"), sep = ",", skip = 0)
 }
@@
 clinical_raw_filter <- clinical_raw %>%
-  filter( len_hospitalisation > min_hosp_days) %>%
+  filter( len_hospitalisation > min_hosp_days, 
+          cohort  %in% c("AllAdm")) %>%
   dplyr::mutate( date = as.Date( calendar_date, format = dateFormat ),
@@
                  period = ifelse( date < pandemic_start_date,
-                                  "before_pandemic","during_pandemic")) %>% 
-  dplyr::filter(cohort  %in% c("AllAdm"))
-stopifnot(unique(all_codes$in_hospital) == 1)
-stopifnot(unique(all_codes$len_hospitalisation > min_hosp_days))
+                                  "before_pandemic", "during_pandemic"))
 ```
@@
 ```{r}
+stopifnot(unique(clinical_raw_filter$cohort) == "AllAdm")
 all_codes <- clinical_raw_filter %>% 
-  left_join( obs_raw,  by = c("patient_num", "cohort", "days_since_admission")) %>%
+  left_join( obs_raw,  by = c("patient_num", "days_since_admission")) %>%
   filter( concept_type == 'DIAG-ICD10')
+stopifnot(unique(all_codes$in_hospital) == 1)
+stopifnot(unique(all_codes$len_hospitalisation > min_hosp_days))
 ```
@@
   filter( in_hospital == 1, 
-         date < '2021-05-01' &
-         date >= '2019-03-01' ) %>%
+         date < end_date_plots &
+         date >= start_date_plots ) %>%
     group_by( date, weeks, month) %>%
@@
 new_hospitalizations_per_week <- new_hospitalizations_per_day %>%
-  filter(weeks >= '2019-03-01') %>% 
+  filter(weeks >= start_date_plots) %>% 
   group_by( weeks) %>%
@@
 total_hospitalizations_per_day <- all_codes %>%
-  group_by(patient_num, date ) %>%
-  filter(date < '2021-05-01' &
-         date >= '2019-03-01' ) %>%
-    group_by( date, weeks, month) %>%
-  summarise(count = n_distinct(patient_num)) %>% 
-  ungroup()
+  filter(date < end_date_plots &
+         date >= start_date_plots ) %>%
+    group_by( date ) %>%
+  summarise(count = n_distinct(patient_num))
 
-total_hospitalizations_per_week <- total_hospitalizations_per_day %>% 
-  filter(weeks > '2019-03-01') %>% 
+total_hospitalizations_per_week <- all_codes %>%
+  filter(date < end_date_plots &
+         date >= start_date_plots ) %>%
     group_by( weeks ) %>%
-  summarise(count = sum(count))
+  summarise(count = n_distinct(patient_num))
 
-total_hospitalizations_per_month <- total_hospitalizations_per_day %>%
+total_hospitalizations_per_month <- all_codes %>%
+  filter(date < end_date_plots &
+         date >= start_date_plots ) %>%
     group_by( month ) %>%
-  summarise(count = sum(count))
+  summarise(count = n_distinct(patient_num))
 
@@
 ```
+
+#### Psychiatric related conditions vs. any other condition
+
+##### New admissions
+
+##### Total hospitalizations
+We have estimated the total number of hospitalizations in the previous steps:
+- total_hospitalization_per_day
+- total_hospitalization_per_week
+- total_hospitalization_per_month
+
+We will use these counts to estimate the percentage and to compare total counts, vs psychiatric related conditions.
+
+```{r}
+total_hospitalizations_per_day_psy <- input_data %>%
+  filter(date < end_date_plots &
+         date >= start_date_plots ) %>%
+    group_by( date ) %>%
+  summarise(count_psy = n_distinct(patient_num)) %>%
+  left_join( total_hospitalizations_per_day ) %>%
+  mutate( period = ifelse( date < pandemic_start_date,
+                                  "before_pandemic",
+                                 "during_pandemic"), 
+                percentage = count_psy * 100 / count)
+
+thpd_psyPer <- total_hospitalizations_per_day_psy %>%
+  ggplot(aes(x = date, y = percentage, fill = period)) +
+  #geom_line()+
+  scale_fill_manual(values = cbPalette[c(3,4)]) +
+  labs(y = "Percentage",
+       x = "Date (by day)",
+       title = "Percentage of patients with mental health related ICD codes (daily)") + 
+  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)
+thpd_psyPer
+
+thpd_psyCount <- total_hospitalizations_per_day_psy %>%
+  pivot_longer(names_to = "counts", cols=c(count_psy, count)) %>%
+  mutate( counts = ifelse( counts == "count_psy", "Psy", "Total")) %>%
+  ggplot(aes(x = date, y = value, fill = counts, color = counts)) +
+  geom_point() +
+  geom_line() +
+  geom_vline(xintercept = as.Date(pandemic_start_date),
+             linetype = "dashed") +
+  scale_fill_manual(values = cbPalette) +
+  scale_color_manual(values = cbPalette) +
+  labs(y = "Counts",
+       x = "Date (by day)",
+       title = "Daily patient counts with mental health related ICD codes") + 
+  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)
+thpd_psyCount
+
+```
+
+We will compare the psychiatric conditions with some other disorders (e.g, digestive, respiratory, etc.)
+We will use these counts to estimate the percentage and to compare total counts, vs psychiatric related conditions.
+
+```{r}
+total_hospitalizations_per_day_disorder_group <- all_codes_desc %>%
+  filter(date < end_date_plots &
+         date >= start_date_plots & 
+        icd_description_category %in% c("Psy","Neoplams","Diseases of the respiratory system","Diseases of the digestive system" ))%>%
+    group_by( date, icd_description_category ) %>%
+  summarise(counts = n_distinct(patient_num))
+
+thpd_disorderGroup <- total_hospitalizations_per_day_disorder_group %>%
+  ggplot(aes(x = date, y = counts, fill = icd_description_category, color = icd_description_category)) +
+  #geom_point(alpha=0.2) +
+  geom_line() +
+  geom_vline(xintercept = as.Date(pandemic_start_date),
+             linetype = "dashed") +
+  scale_fill_manual(values = cbPalette) +
+  scale_color_manual(values = cbPalette) +
+  labs(y = "patient number",
+       x = "days",
+       title = "Total number of hospitalized patients (daily)") +   
+  theme(axis.text.x = element_text(angle=45, vjust = 1, hjust=1), legend.position = "bottom", legend.text = element_text(size=7), legend.title = element_text(size=7) ) +
+  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)
+thpd_disorderGroup
+```

### 2021-09-29 psy_peds_3.Rmd
new version

--- 2021-09-28 psy_peds_2.Rmd
@@
 
-listOfPackages <- c("tidyverse", "RColorBrewer", "knitr", "kableExtra", "tsModel")
+listOfPackages <- c("tidyverse", "RColorBrewer", "knitr", "kableExtra", "tsModel", "gridExtra")
 paket(listOfPackages)
@@
 ### Extract the patient summary and observation information. 
-demo_raw <- files[["patientSummary"]]
-obs_raw <- files[["patientObservations"]]
-clinical_raw <- files[["patientClinicalCourse"]]
+demo_raw <- files[["patientSummary"]] %>% filter(cohort == "AllAdm")
+obs_raw <- files[["patientObservations"]] %>% filter(cohort == "AllAdm")
+clinical_raw <- files[["patientClinicalCourse"]] %>% filter(cohort == "AllAdm")
 
@@
 ```{r}
-# Adding filtering on in_hospital == 1 here to reduce subsequent compute time
+# Filtering on inpatients here (in_hospital == 1), to reduce subsequent compute time of the encounter length calculation 
 clinical_raw <- filter(clinical_raw, in_hospital == 1)
@@
 ```{r}
-#icdCodes <- read.csv("public-data/pediatric_psychiatric_ICD10_codes.csv", header = TRUE, colClasses = "character") %>% 
-#  rename(disorder_group = Mental.Health.Disorder.Group, 
-#         description = Description, 
-#         originalCode = ICD10_Code) %>%
-#  mutate( ICD10_Code= ifelse( nchar( originalCode) > 3, paste0(substr( originalCode, 1, 3), ".", substr(  originalCode, 4, nchar(originalCode))),originalCode)) %>%
-#  select( disorder_group, ICD10_Code, description)
-
 icdCodes <- read.csv("public-data/pediatric_psychiatric_ICD10_codesV2.csv", header = TRUE, colClasses = "character") %>%
@@
    select( disorder_group, ICD10_Code, description )
+```
+
+### Check that we have patients with the ICD10 mental codes
+```{r}
+### check that the ICD codes follow the regular expression [A-Z][0-9][0-9AB]\.?[0-9A-TV-Z]{0,4}
+codesToReview <- obs_raw %>% dplyr::filter(concept_type == "DIAG-ICD10",
+                                           ! grepl( "[A-Z][0-9][0-9AB]\\.?[0-9A-TV-Z]{0,4}", concept_code))
+print(codesToReview)
+
+### Calculate how many of the mental codes are in the dataset
+mentalHealthCodes <- unique( obs_raw %>%
+  dplyr::filter( concept_type == "DIAG-ICD10", 
+                 concept_code %in% icdCodes$ICD10_Code ) %>%
+  dplyr::mutate( concept_code = as.character( concept_code ), 
+                 concept_length = nchar( concept_code ), 
+                 concept_check = ifelse( concept_length == 4 | concept_length > 8, "Check", "OK")) %>%
+  dplyr::select( concept_code, concept_length, concept_check) )
+
+summary(as.factor( mentalHealthCodes$concept_check ) )
+
+print( paste0("There are ", length(mentalHealthCodes$concept_code), " ICD10 mental codes out of the total ",length(unique(icdCodes$ICD10_Code)), " in the dataset"))
 ```
@@
 
-clinical_raw_filter <- clinical_raw %>%
+obs_filtered <- obs_raw %>% 
+  filter(concept_type == 'DIAG-ICD10', 
+         days_since_admission >= 0)
+
+clinical_filtered <- clinical_raw %>%
   filter( len_hospitalisation > min_hosp_days, 
@@
                                   "before_pandemic", "during_pandemic"))
-```
 
-```{r}
-stopifnot(unique(clinical_raw_filter$cohort) == "AllAdm")
-all_codes <- clinical_raw_filter %>% 
-  left_join( obs_raw,  by = c("patient_num", "days_since_admission")) %>%
-  filter( concept_type == 'DIAG-ICD10')
+# Steps necessary to filter:
+  # - encounters without ICD
+  # - ICD diagnosed outside of encounters
+  # - patients from the patients table that don't have any encounters matching the criteria
+clinical_table <- clinical_filtered %>% 
+  inner_join( distinct(obs_filtered, patient_num, days_since_admission),  by = c("patient_num", "days_since_admission")) %>% 
+  distinct(patient_num, n_hospitalisation) %>% 
+  inner_join(clinical_filtered, by = c("patient_num", "n_hospitalisation"))
+observation_table <- obs_filtered %>% 
+  inner_join(distinct(clinical_table, patient_num, days_since_admission), by = c("patient_num", "days_since_admission"))
+demographic_table <- demo_raw %>% 
+  inner_join(distinct(clinical_table, patient_num), by = "patient_num")
+all_codes <- observation_table %>% 
+  inner_join(clinical_table, on = c("patient_num", "days_since_admission")) %>% 
+  inner_join(select(demographic_table, -days_since_admission), on = c("patient_num"))
+stopifnot(nrow(all_codes) == nrow(observation_table))
+stopifnot(unique(all_codes$cohort) == "AllAdm")
 stopifnot(unique(all_codes$in_hospital) == 1)
 stopifnot(unique(all_codes$len_hospitalisation > min_hosp_days))
+
 ```
@@
 ```{r, fig.width=12, fig.height=12}
-new_hospitalizations_per_day <- all_codes %>%
-  group_by(patient_num, n_hospitalisation) %>% slice(which.min(date)) %>%
+
+new_hospitalizations_per_day <- clinical_filtered %>%
+  group_by(patient_num, n_hospitalisation) %>%
+  slice(which.min(date)) %>%
   filter( in_hospital == 1, 
@@
 
-nh_day <- ggplot(new_hospitalizations_per_day,
-                 aes(x = date, y = count))  +
+admission_plot_count <- function(df,
+                               x,
+                               y, 
+                               title = "New admission counts",
+                               color = NULL,
+                               size = NULL) {
+  plot <- ggplot(df, aes_string(x = x, y = y, color = color, size = size))  +
     geom_point() +
@@
          y = "Patients count", 
+       title = title)
+  return(plot)
+}
+
+nh_day <- admission_plot_count(df = new_hospitalizations_per_day, 
+                     x = "date",
+                     y = "count", 
                      title = "New admissions (per day)")
 
-nh_week <- ggplot(new_hospitalizations_per_week, aes(x = weeks,
-                       y = count))  +
-  geom_point() +
-  scale_x_date(date_breaks = "1 month", date_labels =  "%b %Y") +
-  theme(axis.text.x=element_text(angle=60, hjust=1))+
-  labs(x = "Calendar date",
-       y = "Patients count", 
+nh_week <- admission_plot_count(df = new_hospitalizations_per_week, 
+                     x = "weeks",
+                     y = "count", 
                      title = "New admissions (per week)")
 
-nh_month <- ggplot(new_hospitalizations_per_month, aes(x = month,
-                       y = count))  +
-  geom_point() +
-  scale_x_date(date_breaks = "1 month", date_labels =  "%b %Y") +
-  theme(axis.text.x=element_text(angle=60, hjust=1))+
-  labs(x = "Calendar date",
-       y = "Patients count", 
+nh_month <- admission_plot_count(df = new_hospitalizations_per_month, 
+                     x = "month",
+                     y = "count", 
                      title = "New admissions (per month)")
 
-require(gridExtra)
 grid.arrange(nh_day,
@@
 ```{r, fig.height=12, fig.width=12}
-total_hospitalizations_per_day <- all_codes %>%
+total_hospitalizations_per_day <- clinical_filtered %>%
   filter(date < end_date_plots &
@@
 
-total_hospitalizations_per_week <- all_codes %>%
+unique_patients_hospitalized_per_week <- clinical_filtered %>%
   filter(date < end_date_plots &
@@
 
-total_hospitalizations_per_month <- all_codes %>%
+unique_patients_hospitalized_per_month <- clinical_filtered %>%
   filter(date < end_date_plots &
@@
 
-th_week <- ggplot(total_hospitalizations_per_week, aes(x = weeks,
+th_week <- ggplot(unique_patients_hospitalized_per_week, aes(x = weeks,
                        y = count))  +
@@
 
-th_month <- ggplot(total_hospitalizations_per_month, aes(x = month,
+th_month <- ggplot(unique_patients_hospitalized_per_month, aes(x = month,
                        y = count))  +
@@
 
-### Check that we have patients with the ICD10 mental codes
-```{r}
-### check that the ICD codes follow the regular expression [A-Z][0-9][0-9AB]\.?[0-9A-TV-Z]{0,4}
-codesToReview <- obs_raw %>% dplyr::filter(concept_type == "DIAG-ICD10",
-                                           ! grepl( "[A-Z][0-9][0-9AB]\\.?[0-9A-TV-Z]{0,4}", concept_code))
-print(codesToReview)
-
-### Calculate how many of the mental codes are in the dataset
-mentalHealthCodes <- unique( all_codes %>%
-  dplyr::filter( concept_type == "DIAG-ICD10", 
-                 concept_code %in% icdCodes$ICD10_Code ) %>%
-  dplyr::mutate( concept_code = as.character( concept_code ), 
-                 concept_length = nchar( concept_code ), 
-                 concept_check = ifelse( concept_length == 4 | concept_length > 8, "Check", "OK")) %>%
-  dplyr::select( concept_code, concept_length, concept_check) )
-
-summary(as.factor( mentalHealthCodes$concept_check ) )
-
-print( paste0("There are ", length(mentalHealthCodes$concept_code), " ICD10 mental codes out of the total ",length(unique(icdCodes$ICD10_Code)), " in the dataset"))
-```
-
-### Add the mental ICD code description and filter 
+### Add the psychiatric/mental health ICD code description and filter 
 ```{r message=FALSE, warning=FALSE}
@@
 
-icdCategory <- read.delim("public-data/icd10Codes.txt", header = FALSE, colClasses = "character")
+icdCategory <- read.delim("public-data/icd10Codes.txt",
+                          header = FALSE,
+                          colClasses = "character")
 colnames(icdCategory) <- c("icd_code_letter", "icd_description_category")
@@
   mutate( icd_description_category = replace_na( ifelse( concept_code %in% icdCodes$ICD10_Code, "Psy", icd_description_category), "Others")) 
-
-input_data <- all_codes_desc %>% 
+all_codes_psy <- all_codes_desc %>% 
   filter(concept_code %in% icdCodes$ICD10_Code)
@@
 if( raceAvailable == TRUE ){
-  input_data <- input_data %>%
+  all_codes_psy <- all_codes_psy %>%
     left_join( race_raw, by = c("patient_num", "cohort") )
@@
 ```{r}
-total_hospitalizations_per_day_psy <- input_data %>%
+count_icd_per_day_psy <- all_codes_psy %>%
   filter(date < end_date_plots &
@@
   summarise(count_psy = n_distinct(patient_num)) %>%
-  left_join( total_hospitalizations_per_day ) %>%
   mutate( period = ifelse( date < pandemic_start_date,
                            "before_pandemic",
-                                 "during_pandemic"), 
-                percentage = count_psy * 100 / count)
+                           "during_pandemic")
+  )
 
-thpd_psyPer <- total_hospitalizations_per_day_psy %>%
-  ggplot(aes(x = date, y = percentage, fill = period)) +
-  #geom_line()+
+count_icd_per_day <- all_codes_desc %>%
+  filter(date < end_date_plots &
+         date >= start_date_plots ) %>%
+    group_by( date ) %>%
+  summarise(count_icd = n_distinct(patient_num)) %>%
+  mutate( period = ifelse( date < pandemic_start_date,
+                           "before_pandemic",
+                           "during_pandemic")
+  )
+
+count_icd_per_day <- count_icd_per_day %>%
+  left_join(count_icd_per_day_psy, 
+            by = c("date", "period")) %>% 
+  replace_na(list(count_psy = 0))
+
+count_icd_per_day$percentage_psy <- count_icd_per_day$count_psy / count_icd_per_day$count_icd
+
+thpd_psyPer <- count_icd_per_day %>%
+  ggplot(aes(x = date, y = percentage_psy, fill = period)) +
   scale_fill_manual(values = cbPalette[c(3,4)]) +
@@
 thpd_psyPer
+```
 
-thpd_psyCount <- total_hospitalizations_per_day_psy %>%
-  pivot_longer(names_to = "counts", cols=c(count_psy, count)) %>%
+```{r}
+thpd_psyCount <- count_icd_per_day %>%
+  pivot_longer(names_to = "counts", cols=c(count_psy, count_icd)) %>%
   mutate( counts = ifelse( counts == "count_psy", "Psy", "Total")) %>%
@@
   geom_point() +
-  geom_line() +
+  # geom_line() +
   geom_vline(xintercept = as.Date(pandemic_start_date),
@@
 thpd_psyCount
-
 ```

### 2021-09-29 psy_peds_3.Rmd
graphic outputs added to the new version of the analysis

@@
 
-listOfPackages <- c("tidyverse", "RColorBrewer", "knitr", "kableExtra", "tsModel", "gridExtra")
+listOfPackages <- c("tidyverse", "RColorBrewer", "knitr", "kableExtra", "tsModel", "gridExtra", "dplyr")
 paket(listOfPackages)
@@
 ```{r message=FALSE, warning=FALSE}
-folder_4ce_files <- "Input/09212021_2/"
+folder_4ce_files <- "../Input/09212021_2/"
 obfuscation =  FALSE
@@
    filter( flag %in% c("originalCodes", "FrenchSpecificCodes")) %>%
-   select( disorder_group, ICD10_Code, description )
+   dplyr::select( disorder_group, ICD10_Code, description )
 ```
@@
 clinical_raw$age_time_visit <- clinical_raw$age - floor(as.numeric(last_data_date - as.Date(clinical_raw$calendar_date, dateFormat)) / 365.25)
-clinical_raw <- select(clinical_raw, -age)
+clinical_raw <- dplyr::select(clinical_raw, -age)
 
@@
 
-# Steps necessary to filter:
+# Steps necessary to filter rows:
   # - encounters without ICD
@@
   inner_join(clinical_filtered, by = c("patient_num", "n_hospitalisation"))
+# Adding new columns
 observation_table <- obs_filtered %>% 
-  inner_join(distinct(clinical_table, patient_num, days_since_admission), by = c("patient_num", "days_since_admission"))
+  inner_join(clinical_table[, c("patient_num", "days_since_admission")], by = c("patient_num", "days_since_admission"))
 demographic_table <- demo_raw %>% 
@@
 all_codes <- observation_table %>% 
-  inner_join(clinical_table, on = c("patient_num", "days_since_admission")) %>% 
-  inner_join(select(demographic_table, -days_since_admission), on = c("patient_num"))
-stopifnot(nrow(all_codes) == nrow(observation_table))
+   inner_join(clinical_table, by = c("patient_num", "days_since_admission")) 
+#%>% 
+#   inner_join(dplyr::select(demographic_table, -days_since_admission), by = c("patient_num"))
+
 stopifnot(unique(all_codes$cohort) == "AllAdm")
@@
 stopifnot(unique(all_codes$len_hospitalisation > min_hosp_days))
-
+stopifnot(nrow(all_codes) == nrow(observation_table))
+stopifnot(length(unique(all_codes$patient_num)) == length(unique(demographic_table$patient_num)))
+stopifnot(length(unique(all_codes$patient_num)) == length(unique(observation_table$patient_num)))
+stopifnot(length(unique(all_codes$patient_num)) == length(unique(clinical_table$patient_num)))
+stopifnot(nrow(distinct(all_codes, patient_num, n_hospitalisation)) == nrow(distinct(clinical_table, patient_num, n_hospitalisation)))
+stopifnot(nrow(distinct(all_codes, patient_num, days_since_admission)) == nrow(distinct(observation_table, patient_num, days_since_admission)))
 ```
@@
        y = "Patients count", 
-       title = "Total admissions (per day)")
+       title = "Currently hospitalized patients (per day)")
 
@@
        y = "Patients count", 
-       title = "Total admissions (per week)")
+       title = "Currently hospitalized patients (per week)")
 
@@
 grid.arrange(th_day, th_week, th_month, ncol=1)
+```
+
+#### Group by sex
+
+#### Per week
+```{r}
+# counts per week and sex
+patients_hospitalized_per_week_agg_sex <- clinical_filtered %>%
+  filter(date < end_date_plots &
+         date >= start_date_plots & 
+           sex %in% c("female", "male")) %>%
+    group_by( weeks, sex ) %>%
+  summarise(count_sex = n_distinct(patient_num))
+
+# the total counts per week was estimated before: unique_patients_hospitalized_per_week
+patients_hospitalized_per_week_agg_sex_perc <- unique_patients_hospitalized_per_week %>%
+  left_join( patients_hospitalized_per_week_agg_sex, by = c("weeks")) %>% 
+  replace_na(list(count_sex = 0))
+
+patients_hospitalized_per_week_agg_sex_perc$percentage <- 100*patients_hospitalized_per_week_agg_sex_perc$count_sex/patients_hospitalized_per_week_agg_sex_perc$count
+
+patients_hospitalized_per_week_agg_sex %>%
+  ggplot(aes(x = weeks, y = count_sex, fill = sex)) +
+  geom_bar(stat="identity", position = "dodge")+
+  scale_fill_manual(values = cbPalette) +
+  labs(y = "Counts",
+       x = "Week",
+       title = "Weekly patient counts aggregated by sex",
+       subtitle = "Per Sex") + 
+  theme(legend.position = "bottom") + 
+  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=sex), colour= "black", size = 0.5)
+
+patients_hospitalized_per_week_agg_sex_perc %>%
+  ggplot(aes(x = weeks, y = percentage, fill = sex)) +
+  geom_bar(stat="identity", position = "dodge")+
+  scale_fill_manual(values = cbPalette) +
+  labs(y = "Percentage",
+       x = "Week",
+       title = "Weekly patient (%) aggregated by sex",
+       subtitle = "Per Sex") + 
+  theme(legend.position = "bottom") + 
+  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=sex), colour= "black", size = 0.5)
 ```
@@
                                "others")) %>%
-  dplyr::select( patient_num, cohort, sex, age_time_visit, concept_type, concept_code, calendar_date,
+  dplyr::select( patient_num, sex, age_time_visit, concept_type, concept_code, calendar_date,
                  in_hospital, n_hospitalisation, len_hospitalisation, date, weeks, month, year,
@@
   all_codes_psy <- all_codes_psy %>%
-    left_join( race_raw, by = c("patient_num", "cohort") )
+    left_join( race_raw, by = c("patient_num") )
 }
+```
+
+### Aggregated by sex
+```{r}
+# counts per week and sex
+patients_hospitalized_per_week_agg_sex <- all_codes_psy %>%
+  filter(date < end_date_plots &
+         date >= start_date_plots & 
+           sex %in% c("female", "male")) %>%
+    group_by( weeks, sex ) %>%
+  summarise(count_sex = n_distinct(patient_num))
+
+patients_hospitalized_per_week_agg_sex %>%
+  ggplot(aes(x = weeks, y = count_sex, fill = sex)) +
+  geom_bar(stat="identity", position = "dodge")+
+  scale_fill_manual(values = cbPalette) +
+  labs(y = "Counts",
+       x = "Week",
+       title = "Weekly patient counts aggregated by sex",
+       subtitle = "Per Sex") + 
+  theme(legend.position = "bottom") + 
+  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=sex), colour= "black", size = 0.5)
+
+##percentage
+total_patients_psy_per_week <- all_codes_psy %>%
+  filter(date < end_date_plots &
+         date >= start_date_plots ) %>%
+    group_by( weeks ) %>%
+  summarise(count = n_distinct(patient_num))
+
+patients_hospitalized_per_week_agg_sex_perc <- total_patients_psy_per_week %>%
+  left_join( patients_hospitalized_per_week_agg_sex, by = c("weeks")) %>% 
+  replace_na(list(count_sex = 0))
+
+patients_hospitalized_per_week_agg_sex_perc$percentage <- 100*patients_hospitalized_per_week_agg_sex_perc$count_sex/patients_hospitalized_per_week_agg_sex_perc$count
+
+patients_hospitalized_per_week_agg_sex_perc %>%
+  ggplot(aes(x = weeks, y = percentage, fill = sex)) +
+  geom_bar(stat="identity", position = "dodge")+
+  scale_fill_manual(values = cbPalette) +
+  labs(y = "Percentage ",
+       x = "Week",
+       title = "Weekly patient (%) aggregated by sex",
+       subtitle = "Per Sex") + 
+  theme(legend.position = "bottom") + 
+  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=sex), colour= "black", size = 0.5)
+
 ```
@@
 
-count_icd_per_day$percentage_psy <- count_icd_per_day$count_psy / count_icd_per_day$count_icd
+count_icd_per_day$percentage_psy <- 100*count_icd_per_day$count_psy / count_icd_per_day$count_icd
 
@@
        title = "Percentage of patients with mental health related ICD codes (daily)") + 
-  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)
+  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)+
+  geom_point()
 thpd_psyPer
+
+count_icd_per_day %>%
+  ggplot(aes(x = date, y = percentage_psy, fill = period)) +
+  scale_fill_manual(values = cbPalette[c(3,4)]) +
+  labs(y = "Percentage",
+       x = "Date (by day)",
+       title = "Percentage of patients with mental health related ICD codes (daily)") + 
+  geom_point()
+  #geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)
 ```
 
+By week
+
+```{r}
+count_icd_per_week_psy <- all_codes_psy %>%
+  filter(date < end_date_plots &
+         date >= start_date_plots ) %>%
+    group_by( weeks ) %>%
+  summarise(count_psy = n_distinct(patient_num)) %>%
+  mutate( period = ifelse( weeks < pandemic_start_date,
+                           "before_pandemic",
+                           "during_pandemic")
+  )
+
+count_icd_per_week <- all_codes_desc %>%
+  filter(date < end_date_plots &
+         date >= start_date_plots ) %>%
+    group_by( weeks ) %>%
+  summarise(count_icd = n_distinct(patient_num)) %>%
+  mutate( period = ifelse( weeks < pandemic_start_date,
+                           "before_pandemic",
+                           "during_pandemic")
+  )
+
+count_icd_per_week <- count_icd_per_week %>%
+  left_join(count_icd_per_week_psy, 
+            by = c("weeks", "period")) %>% 
+  replace_na(list(count_psy = 0))
+
+count_icd_per_week$percentage_psy <- 100*count_icd_per_week$count_psy / count_icd_per_week$count_icd
+
+thpd_psyPer <- count_icd_per_week %>%
+  ggplot(aes(x = weeks, y = percentage_psy, fill=period)) +
+  scale_fill_manual(values = cbPalette[c(3,4)]) +
+  labs(y = "Percentage",
+       x = "Date (by week)",
+       title = "Percentage of patients with mental health related ICD codes (weekly)") + 
+  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)+
+  geom_point()
+thpd_psyPer
+
+count_icd_per_week %>%
+  ggplot(aes(x = weeks, y = percentage_psy, fill = period)) +
+  scale_fill_manual(values = cbPalette[c(3,4)]) +
+  labs(y = "Percentage",
+       x = "Date (by week)",
+       title = "Percentage of patients with mental health related ICD codes (weekly)") + 
+  geom_point()
+  #geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)
+```
+
+# Per day
 ```{r}
@@
 thpd_psyCount
+
+```
+
+# Per week
+```{r}
+thpd_psyCount <- count_icd_per_week %>%
+  pivot_longer(names_to = "counts", cols=c(count_psy, count_icd)) %>%
+  mutate( counts = ifelse( counts == "count_psy", "Psy", "Total")) %>%
+  ggplot(aes(x = weeks, y = value, fill = counts, color = counts)) +
+  geom_point() +
+  # geom_line() +
+  geom_vline(xintercept = as.Date(pandemic_start_date),
+             linetype = "dashed") +
+  scale_fill_manual(values = cbPalette) +
+  scale_color_manual(values = cbPalette) +
+  labs(y = "Counts",
+       x = "Date (by week)",
+       title = "Weekly patient counts with mental health related ICD codes") + 
+  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)
+thpd_psyCount
+
 ```
@@
 ```
+
+#### Psychiatric conditions aggregating by suptype
+
+##### New admissions
+
+##### Total hospitalizations
+
+We estimate and plot the counts per disorder group.
+
+Per day
+```{r}
+count_disorder_group_per_day <- all_codes_psy %>%
+  filter(date < end_date_plots &
+         date >= start_date_plots ) %>%
+    group_by( date, disorder_group ) %>%
+  summarise(count_dg = n_distinct(patient_num)) %>%
+  mutate( period = ifelse( date < pandemic_start_date,
+                           "before_pandemic",
+                           "during_pandemic")
+  )
+
+dg_per_day <- count_disorder_group_per_day %>%  
+  ggplot(aes(x = date, y= count_dg, fill = period, group = disorder_group )) +
+  geom_bar(stat = "identity")+
+  geom_vline(xintercept = as.Date(pandemic_start_date),
+             linetype = "dashed") +
+  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se= TRUE, aes(colour=period), colour= "black", size = 0.5) + 
+  labs(x = "Days",
+       y = "Patient Count",
+       title = "Daily patient counts with mental health related ICD codes", 
+       subtitle = "Grouped by mental health ICD subcategories") + 
+  ylim(0, NA) + 
+  facet_wrap(. ~ disorder_group, scales = "free_y") +
+  theme(strip.text.x = element_text(size = 4),
+        axis.text.x = element_text(size = 5, angle = 90), 
+        axis.text.y = element_text(size = 6), 
+        axis.title.x = element_text(size = 10),
+        axis.title.y = element_text(size = 10),
+        title = element_text(size = 12),
+        plot.subtitle = element_text(size = 11),
+        legend.position = "none") +
+  scale_fill_manual(values = cbPalette[-1])
+
+dg_per_day
+```
+
+Per week
+```{r}
+count_disorder_group_per_week <- all_codes_psy %>%
+  filter(date < end_date_plots &
+         date >= start_date_plots ) %>%
+    group_by( weeks, disorder_group ) %>%
+  summarise(count_dg = n_distinct(patient_num)) %>%
+  mutate( period = ifelse( weeks < pandemic_start_date,
+                           "before_pandemic",
+                           "during_pandemic")
+  )
+
+dg_per_week <- count_disorder_group_per_week %>%  
+  ggplot(aes(x = weeks, y= count_dg, fill = period, group = disorder_group )) +
+  geom_bar(stat = "identity")+
+  geom_vline(xintercept = as.Date(pandemic_start_date),
+             linetype = "dashed") +
+  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se= TRUE, aes(colour=period), colour= "black", size = 0.5) + 
+  labs(x = "Weeks",
+       y = "Patient Count",
+       title = "Weekly patient counts with mental health related ICD codes", 
+       subtitle = "Grouped by mental health ICD subcategories") + 
+  ylim(0, NA) + 
+  facet_wrap(. ~ disorder_group, scales = "free_y") +
+  theme(strip.text.x = element_text(size = 4),
+        axis.text.x = element_text(size = 5, angle = 90), 
+        axis.text.y = element_text(size = 6), 
+        axis.title.x = element_text(size = 10),
+        axis.title.y = element_text(size = 10),
+        title = element_text(size = 12),
+        plot.subtitle = element_text(size = 11),
+        legend.position = "none") +
+  scale_fill_manual(values = cbPalette[-1])
+
+dg_per_week
+```
+
+Per month
+```{r}
+count_disorder_group_per_month <- all_codes_psy %>%
+  filter(date < end_date_plots &
+         date >= start_date_plots ) %>%
+    group_by( month, disorder_group ) %>%
+  summarise(count_dg = n_distinct(patient_num)) %>%
+  mutate( period = ifelse( month < pandemic_start_date,
+                           "before_pandemic",
+                           "during_pandemic")
+  )
+
+dg_per_month <- count_disorder_group_per_month %>%  
+  ggplot(aes(x = month, y= count_dg, fill = period, group = disorder_group )) +
+  geom_bar(stat = "identity")+
+  geom_vline(xintercept = as.Date(pandemic_start_date),
+             linetype = "dashed") +
+  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se= TRUE, aes(colour=period), colour= "black", size = 0.5) + 
+  labs(x = "Months",
+       y = "Patient Count",
+       title = "Monthly patient counts with mental health related ICD codes", 
+       subtitle = "Grouped by mental health ICD subcategories") + 
+  ylim(0, NA) + 
+  facet_wrap(. ~ disorder_group, scales = "free_y") +
+  theme(strip.text.x = element_text(size = 4),
+        axis.text.x = element_text(size = 5, angle = 90), 
+        axis.text.y = element_text(size = 6), 
+        axis.title.x = element_text(size = 10),
+        axis.title.y = element_text(size = 10),
+        title = element_text(size = 12),
+        plot.subtitle = element_text(size = 11),
+        legend.position = "none") +
+  scale_fill_manual(values = cbPalette[-1])
+
+dg_per_month
+```
+
+Percentage using as denominator total number of patients per day with a psy disorder.
+
+Per day
+- Total counts previously estimated (count_icd_per_day_psy).
+- Counts per disorder group previously estimated (count_disorder_group_per_day)
+
+```{r}
+# join the total with the disorder group count to estimate the percentage
+perc_disorder_group_per_day <- count_icd_per_day_psy %>%
+  left_join(count_disorder_group_per_day, 
+            by = c("date", "period")) %>% 
+  replace_na(list(count_dg = 0))
+
+perc_disorder_group_per_day$percentage_dg <- 100*perc_disorder_group_per_day$count_dg / perc_disorder_group_per_day$count_psy
+
+dgPerc_per_day <- perc_disorder_group_per_day %>%  
+  ggplot(aes(x = date, y= percentage_dg, fill = period, group = disorder_group )) +
+  geom_bar(stat = "identity")+
+  geom_vline(xintercept = as.Date(pandemic_start_date),
+             linetype = "dashed") +
+  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se= TRUE, aes(colour=period), colour= "black", size = 0.5) + 
+  labs(x = "Days",
+       y = "Patient (%)",
+       title = "Percentage of daily patients with mental health related ICD codes", 
+       subtitle = "Grouped by mental health ICD subcategories") + 
+  ylim(0, NA) + 
+  facet_wrap(. ~ disorder_group, scales = "free_y") +
+  theme(strip.text.x = element_text(size = 4),
+        axis.text.x = element_text(size = 5, angle = 90), 
+        axis.text.y = element_text(size = 6), 
+        axis.title.x = element_text(size = 10),
+        axis.title.y = element_text(size = 10),
+        title = element_text(size = 12),
+        plot.subtitle = element_text(size = 11),
+        legend.position = "none") +
+  scale_fill_manual(values = cbPalette[-1])
+
+dgPerc_per_day
+```
+
+Per week
+- Counts per disorder group previously estimated (count_disorder_group_per_week)
+
+```{r}
+# estimate the weekly counts
+count_icd_per_week_psy <- all_codes_psy %>%
+  filter(date < end_date_plots &
+         date >= start_date_plots ) %>%
+    group_by( weeks ) %>%
+  summarise(count_psy = n_distinct(patient_num)) %>%
+  mutate( period = ifelse( weeks < pandemic_start_date,
+                           "before_pandemic",
+                           "during_pandemic")
+  )
+
+# join the total with the disorder group count to estimate the percentage
+perc_disorder_group_per_week <- count_icd_per_week_psy %>%
+  left_join(count_disorder_group_per_week, 
+            by = c("weeks", "period")) %>% 
+  replace_na(list(count_dg = 0))
+
+perc_disorder_group_per_week$percentage_dg <- 100*perc_disorder_group_per_week$count_dg / perc_disorder_group_per_week$count_psy
+
+dgPerc_per_week <- perc_disorder_group_per_week %>%  
+  ggplot(aes(x = weeks, y= percentage_dg, fill = period, group = disorder_group )) +
+  geom_bar(stat = "identity")+
+  geom_vline(xintercept = as.Date(pandemic_start_date),
+             linetype = "dashed") +
+  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se= TRUE, aes(colour=period), colour= "black", size = 0.5) + 
+  labs(x = "Weeks",
+       y = "Patient (%)",
+       title = "Percentage of weekly patients with mental health related ICD codes", 
+       subtitle = "Grouped by mental health ICD subcategories") + 
+  ylim(0, NA) + 
+  facet_wrap(. ~ disorder_group, scales = "free_y") +
+  theme(strip.text.x = element_text(size = 4),
+        axis.text.x = element_text(size = 5, angle = 90), 
+        axis.text.y = element_text(size = 6), 
+        axis.title.x = element_text(size = 10),
+        axis.title.y = element_text(size = 10),
+        title = element_text(size = 12),
+        plot.subtitle = element_text(size = 11),
+        legend.position = "none") +
+  scale_fill_manual(values = cbPalette[-1])
+
+dgPerc_per_week
+```
+
+Months 
+```{r}
+# estimate the monthly counts
+count_icd_per_month_psy <- all_codes_psy %>%
+  filter(date < end_date_plots &
+         date >= start_date_plots ) %>%
+    group_by( month ) %>%
+  summarise(count_psy = n_distinct(patient_num)) %>%
+  mutate( period = ifelse( month < pandemic_start_date,
+                           "before_pandemic",
+                           "during_pandemic")
+  )
+
+# join the total with the disorder group count to estimate the percentage
+perc_disorder_group_per_month <- count_icd_per_month_psy %>%
+  left_join(count_disorder_group_per_month, 
+            by = c("month", "period")) %>% 
+  replace_na(list(count_dg = 0))
+
+perc_disorder_group_per_month$percentage_dg <- 100*perc_disorder_group_per_month$count_dg / perc_disorder_group_per_month$count_psy
+
+dgPerc_per_month <- perc_disorder_group_per_month %>%  
+  ggplot(aes(x = month, y= percentage_dg, fill = period, group = disorder_group )) +
+  geom_bar(stat = "identity")+
+  geom_vline(xintercept = as.Date(pandemic_start_date),
+             linetype = "dashed") +
+  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se= TRUE, aes(colour=period), colour= "black", size = 0.5) + 
+  labs(x = "Months",
+       y = "Patient (%)",
+       title = "Percentage of monthly patients with mental health related ICD codes", 
+       subtitle = "Grouped by mental health ICD subcategories") + 
+  ylim(0, NA) + 
+  facet_wrap(. ~ disorder_group, scales = "free_y") +
+  theme(strip.text.x = element_text(size = 4),
+        axis.text.x = element_text(size = 5, angle = 90), 
+        axis.text.y = element_text(size = 6), 
+        axis.title.x = element_text(size = 10),
+        axis.title.y = element_text(size = 10),
+        title = element_text(size = 12),
+        plot.subtitle = element_text(size = 11),
+        legend.position = "none") +
+  scale_fill_manual(values = cbPalette[-1])
+
+dgPerc_per_month
+```
+
+Percentage using as denominator total number of patients
+
+Per day
+- Counts per disorder group previously estimated (count_disorder_group_per_day)
+
+```{r}
+# estimate the total number per day 
+count_icd_per_day <- all_codes_desc %>%
+  filter(date < end_date_plots &
+         date >= start_date_plots ) %>%
+    group_by( date ) %>%
+  summarise(count_icd = n_distinct(patient_num)) %>%
+  mutate( period = ifelse( date < pandemic_start_date,
+                           "before_pandemic",
+                           "during_pandemic")
+  )
+
+# join the total with the disorder group count to estimate the percentage
+perc_disorder_group_per_day <- count_icd_per_day %>%
+  left_join(count_disorder_group_per_day, 
+            by = c("date", "period")) %>% 
+  replace_na(list(count_dg = 0))
+
+perc_disorder_group_per_day$percentage_dg <- 100*perc_disorder_group_per_day$count_dg / perc_disorder_group_per_day$count_icd
+
+dgPerc_per_day <- perc_disorder_group_per_day %>%  
+  ggplot(aes(x = date, y= percentage_dg, fill = period, group = disorder_group )) +
+  geom_bar(stat = "identity")+
+  geom_vline(xintercept = as.Date(pandemic_start_date),
+             linetype = "dashed") +
+  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se= TRUE, aes(colour=period), colour= "black", size = 0.5) + 
+  labs(x = "Days",
+       y = "Patient (%)",
+       title = "Percentage of daily patients with mental health related ICD codes", 
+       subtitle = "Grouped by mental health ICD subcategories") + 
+  ylim(0, NA) + 
+  facet_wrap(. ~ disorder_group, scales = "free_y") +
+  theme(strip.text.x = element_text(size = 4),
+        axis.text.x = element_text(size = 5, angle = 90), 
+        axis.text.y = element_text(size = 6), 
+        axis.title.x = element_text(size = 10),
+        axis.title.y = element_text(size = 10),
+        title = element_text(size = 12),
+        plot.subtitle = element_text(size = 11),
+        legend.position = "none") +
+  scale_fill_manual(values = cbPalette[-1])
+
+dgPerc_per_day
+```
+
+Per week
+- Counts per disorder group previously estimated (count_disorder_group_per_week)
+
+```{r}
+# estimate the weekly counts
+count_icd_per_week <- all_codes_desc %>%
+  filter(date < end_date_plots &
+         date >= start_date_plots ) %>%
+    group_by( weeks ) %>%
+  summarise(count_icd = n_distinct(patient_num)) %>%
+  mutate( period = ifelse( weeks < pandemic_start_date,
+                           "before_pandemic",
+                           "during_pandemic")
+  )
+
+# join the total with the disorder group count to estimate the percentage
+perc_disorder_group_per_week <- count_icd_per_week %>%
+  left_join(count_disorder_group_per_week, 
+            by = c("weeks", "period")) %>% 
+  replace_na(list(count_dg = 0))
+
+perc_disorder_group_per_week$percentage_dg <- 100*perc_disorder_group_per_week$count_dg / perc_disorder_group_per_week$count_icd
+
+dgPerc_per_week <- perc_disorder_group_per_week %>%  
+  ggplot(aes(x = weeks, y= percentage_dg, fill = period, group = disorder_group )) +
+  geom_bar(stat = "identity")+
+  geom_vline(xintercept = as.Date(pandemic_start_date),
+             linetype = "dashed") +
+  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se= TRUE, aes(colour=period), colour= "black", size = 0.5) + 
+  labs(x = "Weeks",
+       y = "Patient (%)",
+       title = "Percentage of weekly patients with mental health related ICD codes", 
+       subtitle = "Grouped by mental health ICD subcategories") + 
+  ylim(0, NA) + 
+  facet_wrap(. ~ disorder_group, scales = "free_y") +
+  theme(strip.text.x = element_text(size = 4),
+        axis.text.x = element_text(size = 5, angle = 90), 
+        axis.text.y = element_text(size = 6), 
+        axis.title.x = element_text(size = 10),
+        axis.title.y = element_text(size = 10),
+        title = element_text(size = 12),
+        plot.subtitle = element_text(size = 11),
+        legend.position = "none") +
+  scale_fill_manual(values = cbPalette[-1])
+
+dgPerc_per_week
+```
+
+Months 
+```{r}
+# estimate the monthly counts
+count_icd_per_month <- all_codes_desc %>%
+  filter(date < end_date_plots &
+         date >= start_date_plots ) %>%
+    group_by( month ) %>%
+  summarise(count_icd = n_distinct(patient_num)) %>%
+  mutate( period = ifelse( month < pandemic_start_date,
+                           "before_pandemic",
+                           "during_pandemic")
+  )
+
+# join the total with the disorder group count to estimate the percentage
+perc_disorder_group_per_month <- count_icd_per_month %>%
+  left_join(count_disorder_group_per_month, 
+            by = c("month", "period")) %>% 
+  replace_na(list(count_dg = 0))
+
+perc_disorder_group_per_month$percentage_dg <- 100*perc_disorder_group_per_month$count_dg / perc_disorder_group_per_month$count_icd
+
+dgPerc_per_month <- perc_disorder_group_per_month %>%  
+  ggplot(aes(x = month, y= percentage_dg, fill = period, group = disorder_group )) +
+  geom_bar(stat = "identity")+
+  geom_vline(xintercept = as.Date(pandemic_start_date),
+             linetype = "dashed") +
+  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se= TRUE, aes(colour=period), colour= "black", size = 0.5) + 
+  labs(x = "Months",
+       y = "Patient (%)",
+       title = "Percentage of monthly patients with mental health related ICD codes", 
+       subtitle = "Grouped by mental health ICD subcategories") + 
+  ylim(0, NA) + 
+  facet_wrap(. ~ disorder_group, scales = "free_y") +
+  theme(strip.text.x = element_text(size = 4),
+        axis.text.x = element_text(size = 5, angle = 90), 
+        axis.text.y = element_text(size = 6), 
+        axis.title.x = element_text(size = 10),
+        axis.title.y = element_text(size = 10),
+        title = element_text(size = 12),
+        plot.subtitle = element_text(size = 11),
+        legend.position = "none") +
+  scale_fill_manual(values = cbPalette[-1])
+
+dgPerc_per_month
+```

### 2021-09-29 psy_peds_3.Rmd
data wrangling

@@
     toc_float: yes
+editor_options: 
+  markdown: 
+    wrap: 72
 ---
@@
 # R session information
+
 ```{r}
@@
 # Inclusion criteria
+
 The inclusion criteria for this study is:
+
 -   Patients inclusion criteria:
+
     -   Age between 11-17 at the date of visit
+
 -   Visits inclusion criteria:
-  - Inpatients (at least one hospitalization > 1 day)
-  - Starting date between January 1st, 2019 and July 31st, 2021
-  - Comprising least one ICD-10 code pertaining to the specified ICD-10 codes list (mental health related codes)
+
+    -   Inpatients (at least one hospitalization \> 1 day)
+
+    -   Start date: January 1st, 2019
+
+    -   End date: May 31st, 2021
+
+    -   Comprising at least one ICD-10 code during hospitalization
 
 # Variables that need to be checked/modified by each site
-Change the values of the following variables according to the specificities of your site:
-- "folder_4ce_files": folder path where your phase 2.2 data files are located
-- "obfuscation": determine the obfuscation threshold (FALSE if no obfuscation, or the numeric value of the obfuscation threshold if any)
-- "raceAvailable": set as TRUE or FALSE depending on whether the variable is being collected at your site
-- "dateFormat": specify the format of the date at your site (e.g., for "03-AUG-20", the format would be "%d-%b-%y", [see documentation](https://www.stat.berkeley.edu/~s133/dates.html))
+
+Change the values of the following variables according to the
+specificities of your site:
+
+1.  "folder_4ce_files": folder path where your phase 2.2 data files are
+    located
+
+2.   "obfuscation": determine the obfuscation threshold (FALSE if no
+    obfuscation, or the numeric value of the obfuscation threshold if
+    any)
+
+3.   "raceAvailable": set as TRUE or FALSE depending on whether the
+    variable is being collected at your site
+
+4.   "dateFormat": specify the format of the date at your site (e.g., for
+    "03-AUG-20", the format would be "%d-%b-%y", [see
+    documentation](https://www.stat.berkeley.edu/~s133/dates.html))
 
 ```{r message=FALSE, warning=FALSE}
-folder_4ce_files <- "../Input/09212021_2/"
+folder_4ce_files <- "Input/09212021_2/"
 obfuscation =  FALSE
@@
 min_hosp_days <- 2
+data_update_date <- "2021-06-05"
 ```
@@
 pandemic_start_date <- "2020-03-15"
-start_date_plots <- "2019-03-01"
+start_date_plots <- "2019-01-15"
 end_date_plots <- "2021-05-01"
@@
 # Data input
+
 We will use as input the 2.2 data files. Specifically:
+
 - LocalPatientSummary
+
 - LocalPatientObservation
+
 - LocalPatientClinicalcourse
@@
 For sites recording the race, we will also use an additional file:
+
 - LocalPatientRace
@@
 
+### Read the ICD10 codes file
+
+The file containing the ICD10 psychiatric related codes is located in
+the `public-data` folder of the GitHub repository.
+
+```{r}
+icdCodes <- read.csv("public-data/pediatric_psychiatric_ICD10_codesV2.csv", header = TRUE, colClasses = "character") %>%
+   filter( flag %in% c("originalCodes", "FrenchSpecificCodes")) %>%
+   dplyr::select( disorder_group, ICD10_Code, description )
+```
+
+## Data-management
+
+### Filtering on "inpatient" status
+
 ```{r}
@@
 
-### Adding columns with hospitalization number and hospitalization length to the clinical raw file
+### Calculating duration of individual hospitalizations
+
+Adding columns with hospitalization number and hospitalization length to the clinical raw file
+
 ```{r}
@@
 
-### Read the ICD10 codes file
-The file containing the ICD10 psychiatric related codes is located in the `public-data` folder of the GitHub repository.
-
-```{r}
-icdCodes <- read.csv("public-data/pediatric_psychiatric_ICD10_codesV2.csv", header = TRUE, colClasses = "character") %>%
-   filter( flag %in% c("originalCodes", "FrenchSpecificCodes")) %>%
-   dplyr::select( disorder_group, ICD10_Code, description )
-```
-
-### Check that we have patients with the ICD10 mental codes
+### Counting ICD10 mental health codes in the obs_raw table
+
 ```{r}
@@
 
-## All cohort
-We are selecting encounters based on the following criteria:
-- associated with at least 1 ICD codes
-- with a duration of hospitalization spanning over at least two consecutive days
+## Applying inclusion criteria to filter and format the data
+
+We are selecting hospitalizations based on the following criteria:
+
+- Associated with at least 1 ICD code
+
+- With a duration of hospitalization spanning over at least two consecutive days
+
 - Starting on or after 2019/01/01
 
-```{r}
-# Adjusting patient age at the time of visit, assuming the age used is the patients' age at the time of extraction (last discharge date available in the data)
-last_data_date <- max(as.Date(demo_raw$last_discharge_date, "%Y-%m-%d"))
-clinical_raw <- left_join(clinical_raw, demo_raw[c("patient_num", "cohort", "age", "sex")], by = c("patient_num", "cohort"))
-clinical_raw$age_time_visit <- clinical_raw$age - floor(as.numeric(last_data_date - as.Date(clinical_raw$calendar_date, dateFormat)) / 365.25)
-clinical_raw <- dplyr::select(clinical_raw, -age)
-
-obs_filtered <- obs_raw %>% 
+### Calculating age at hospitalization time
+```{r}
+# Calculating patient age at the time of visit, assuming the age used is the patients' age at the time of extraction (last discharge date available in the data)
+obs_raw_age <- left_join(obs_raw, demo_raw[c("patient_num", "age")], by = c("patient_num")) %>% 
+  left_join(clinical_raw[, c("patient_num", "days_since_admission", "calendar_date")], 
+            by = c("patient_num", "days_since_admission"))
+stopifnot(nrow(obs_raw_age) == nrow(obs_raw))
+obs_raw_age$age_time_diagnosis <- obs_raw_age$age - floor(as.numeric(as.Date(data_update_date) - as.Date(obs_raw_age$calendar_date, dateFormat)) / 365.25)
+```
+
+``` {r}
+obs_filtered_age <- obs_raw_age %>% 
   filter(concept_type == 'DIAG-ICD10', 
@@
                                   "before_pandemic", "during_pandemic"))
+```
+
+```{r}
+# Steps necessary to filter rows:
+  # - encounters without ICD
+  # - ICD diagnosed outside of encounters
+  # - patients from the patients table that don't have any encounters matching the criteria
+clinical_table <- clinical_filtered %>% 
+  inner_join( distinct(obs_filtered_age, patient_num, days_since_admission),  by = c("patient_num", "days_since_admission")) %>% 
+  distinct(patient_num, n_hospitalisation) %>% 
+  inner_join(clinical_filtered, by = c("patient_num", "n_hospitalisation"))
+# Adding new columns
+observation_table <- obs_filtered_age %>% 
+  inner_join(clinical_table[, c("patient_num", "days_since_admission")], by = c("patient_num", "days_since_admission"))
+demographic_table <- demo_raw %>% 
+  inner_join(distinct(clinical_table, patient_num), by = "patient_num")
+all_codes <- observation_table %>% 
+   inner_join(clinical_table, by = c("patient_num", "days_since_admission")) %>% 
+   inner_join(demographic_table[, c("patient_num", "sex")], by = "patient_num")
+```
+
+```{r}
+stopifnot(unique(all_codes$cohort) == "AllAdm")
+stopifnot(unique(all_codes$in_hospital) == 1)
+stopifnot(unique(all_codes$len_hospitalisation > min_hosp_days))
+stopifnot(nrow(all_codes) == nrow(observation_table))
+stopifnot(length(unique(all_codes$patient_num)) == length(unique(demographic_table$patient_num)))
+stopifnot(length(unique(all_codes$patient_num)) == length(unique(observation_table$patient_num)))
+stopifnot(length(unique(all_codes$patient_num)) == length(unique(clinical_table$patient_num)))
+stopifnot(nrow(distinct(all_codes, patient_num, n_hospitalisation)) == nrow(distinct(clinical_table, patient_num, n_hospitalisation)))
+stopifnot(nrow(distinct(all_codes, patient_num, days_since_admission)) == nrow(distinct(observation_table, patient_num, days_since_admission)))
+```
+
+## First estimation: total number of patients without filtering by any disease subtype
+
+```{r}
+admission_plot_count <- function(df,
+                               x,
+                               y, 
+                               title,
+                               y_title,
+                               size = NULL, 
+                               color = NULL){
+  plot <- ggplot(df, aes_string(x = x, y = y, color = color, size = size))  +
+    geom_point() +
+    scale_x_date(date_breaks = "1 month", date_labels =  "%b %Y", limits = c(start_date_plots, end_date_plots), expand = expansion(mult = 0.01, add = 1)) +
+    theme(axis.text.x=element_text(angle=60, hjust=1))+
+    labs(x = "Calendar date",
+         y = y_title, 
+       title = title) 
+  return(plot)
+}
+
+```
+
+### New hospitalizations
+
+```{r}
+
+new_hospitalizations_per_week <- clinical_table %>%
+  group_by(patient_num, n_hospitalisation) %>%
+  slice(which.min(weeks)) %>%
+  filter( in_hospital == 1 ) %>%
+    group_by( weeks ) %>%
+  summarise(count = n_distinct(patient_num)) %>% 
+  ungroup()
+
+start_date_plots <- as.Date("2019-01-01")
+end_date_plots <- as.Date("2021-05-24")
+
+nh_week <- admission_plot_count(df = new_hospitalizations_per_week, 
+                     x = "weeks",
+                     y = "count", 
+                     title = "New admissions (per week)", 
+                     y_title = "Patient Count")
+
+nh_week
+```
+
+### Filtering on age at hospitalization time
+
+Keeping only patients with age between 11 and 17
+
+```{r}
+
+# Filtering hospitalizations where patient age is in inclusion criteria
+observation_table_ado <- filter(observation_table, age_time_diagnosis >= 11 & age_time_diagnosis <= 17)
+
+# Filtering observation tables which patient_num and days_since_admission are still in clinical table after filtering on age
+
+clinical_test <- clinical_table[, c("patient_num", "n_hospitalisation", "days_since_admission")] %>% 
+  inner_join(distinct(observation_table_ado, patient_num, days_since_admission, age_time_diagnosis), 
+            by = c("patient_num", "days_since_admission"))
+
+clinical_table_ado <- clinical_table %>% 
+  inner_join(clinical_test[, c("patient_num", "n_hospitalisation", "age_time_diagnosis")],
+             by = c("patient_num", "n_hospitalisation"))
+
+observation_table_ado <- clinical_table %>% 
+  inner_join(observation_table[, c("patient_num", "days_since_admission")], 
+             by = c("patient_num", "days_since_admission"))
+
+# Filtering observation tables which patient_num are still in clinical table after filtering on age
+demographic_table_ado <- demographic_table %>% 
+  inner_join(distinct(clinical_table_ado, patient_num), by = "patient_num")
+
+all_codes_ado <- all_codes %>% 
+   filter(age_time_visit >= 11 & age_time_visit <= 17)
+
+head(observation_table_ado)
+head(clinical_table_ado)
+head(all_codes_ado)
+```
+
+```{r}
+stopifnot(unique(all_codes_ado$cohort) == "AllAdm")
+stopifnot(unique(all_codes_ado$in_hospital) == 1)
+stopifnot(unique(all_codes_ado$len_hospitalisation > min_hosp_days))
+stopifnot(nrow(all_codes_ado) == nrow(observation_table_ado))
+stopifnot(length(unique(all_codes_ado$patient_num)) == length(unique(demographic_table_ado$patient_num)))
+stopifnot(length(unique(all_codes_ado$patient_num)) == length(unique(observation_table_ado$patient_num)))
+stopifnot(length(unique(all_codes_ado$patient_num)) == length(unique(clinical_table_ado$patient_num)))
+stopifnot(nrow(distinct(all_codes_ado, patient_num, n_hospitalisation)) == nrow(distinct(clinical_table_ado, patient_num, n_hospitalisation)))
+stopifnot(nrow(distinct(all_codes_ado, patient_num, days_since_admission)) == nrow(distinct(observation_table_ado, patient_num, days_since_admission)))
+```
+
+```{r}
 
@@
 
-stopifnot(unique(all_codes$cohort) == "AllAdm")
-stopifnot(unique(all_codes$in_hospital) == 1)
-stopifnot(unique(all_codes$len_hospitalisation > min_hosp_days))
-stopifnot(nrow(all_codes) == nrow(observation_table))
-stopifnot(length(unique(all_codes$patient_num)) == length(unique(demographic_table$patient_num)))
-stopifnot(length(unique(all_codes$patient_num)) == length(unique(observation_table$patient_num)))
-stopifnot(length(unique(all_codes$patient_num)) == length(unique(clinical_table$patient_num)))
-stopifnot(nrow(distinct(all_codes, patient_num, n_hospitalisation)) == nrow(distinct(clinical_table, patient_num, n_hospitalisation)))
-stopifnot(nrow(distinct(all_codes, patient_num, days_since_admission)) == nrow(distinct(observation_table, patient_num, days_since_admission)))
-```
-
-## First estimation: total number of patients without filtering by any disease subtype
-
-### New hospitalizations
-```{r, fig.width=12, fig.height=12}
-
-new_hospitalizations_per_day <- clinical_filtered %>%
-  group_by(patient_num, n_hospitalisation) %>%
-  slice(which.min(date)) %>%
-  filter( in_hospital == 1, 
-         date < end_date_plots &
-         date >= start_date_plots ) %>%
-    group_by( date, weeks, month) %>%
-  summarise(count = n_distinct(patient_num)) %>% 
-  ungroup()
-
-new_hospitalizations_per_week <- new_hospitalizations_per_day %>%
-  filter(weeks >= start_date_plots) %>% 
-  group_by( weeks) %>%
-  summarise(count = sum(count))
-
-new_hospitalizations_per_month <- new_hospitalizations_per_day %>%
-  group_by( month) %>%
-  summarise(count = sum(count))
-
-admission_plot_count <- function(df,
-                               x,
-                               y, 
-                               title = "New admission counts",
-                               color = NULL,
-                               size = NULL) {
-  plot <- ggplot(df, aes_string(x = x, y = y, color = color, size = size))  +
-    geom_point() +
-    scale_x_date(date_breaks = "1 month", date_labels =  "%b %Y") +
-    theme(axis.text.x=element_text(angle=60, hjust=1))+
-    labs(x = "Calendar date",
-         y = "Patients count", 
-       title = title)
-  return(plot)
-}
-
-nh_day <- admission_plot_count(df = new_hospitalizations_per_day, 
-                     x = "date",
-                     y = "count", 
-                     title = "New admissions (per day)")
-
-nh_week <- admission_plot_count(df = new_hospitalizations_per_week, 
-                     x = "weeks",
-                     y = "count", 
-                     title = "New admissions (per week)")
-
-nh_month <- admission_plot_count(df = new_hospitalizations_per_month, 
-                     x = "month",
-                     y = "count", 
-                     title = "New admissions (per month)")
-
-grid.arrange(nh_day,
-             nh_week,
-             nh_month,
-             ncol=1)
 ```
@@
 ### Total hospitalize patients
+
 ```{r, fig.height=12, fig.width=12}
@@
 #### Per week
+
 ```{r}
@@
 ## Second estimation: total number of patients filtering by psychiatric disorders
+
 ## Creating the inputTable
-Creation of the inputTable that will be the only table used throughout the rest of the analysis
-- selecting only visits with at least one ICD codes pertaining to the mental health categories
-- filter visits where patients are between 11 and 17 years old
-- add the demographic and hospitalization related information 
-- add the ICD code description and the related disorder group
+
+Creation of the inputTable that will be the only table used throughout
+the rest of the analysis - selecting only visits with at least one ICD
+codes pertaining to the mental health categories - filter visits where
+patients are between 11 and 17 years old - add the demographic and
+hospitalization related information - add the ICD code description and
+the related disorder group
 
 ### Add the psychiatric/mental health ICD code description and filter
+
 ```{r message=FALSE, warning=FALSE}
@@
 ### Aggregated by sex
+
 ```{r}
@@
 ##### Total hospitalizations
-We have estimated the total number of hospitalizations in the previous steps:
-- total_hospitalization_per_day
-- total_hospitalization_per_week
+
+We have estimated the total number of hospitalizations in the previous
+steps: - total_hospitalization_per_day - total_hospitalization_per_week
 - total_hospitalization_per_month
 
-We will use these counts to estimate the percentage and to compare total counts, vs psychiatric related conditions.
+We will use these counts to estimate the percentage and to compare total
+counts, vs psychiatric related conditions.
 
@@
 # Per day
+
 ```{r}
@@
 # Per week
+
 ```{r}
@@
 
-We will compare the psychiatric conditions with some other disorders (e.g, digestive, respiratory, etc.)
-We will use these counts to estimate the percentage and to compare total counts, vs psychiatric related conditions.
+We will compare the psychiatric conditions with some other disorders
+(e.g, digestive, respiratory, etc.) We will use these counts to estimate
+the percentage and to compare total counts, vs psychiatric related
+conditions.
 
@@
 Per day
+
 ```{r}
@@
 Per week
+
 ```{r}
@@
 Per month
+
 ```{r}
@@
 
-Percentage using as denominator total number of patients per day with a psy disorder.
-
-Per day
-- Total counts previously estimated (count_icd_per_day_psy).
-- Counts per disorder group previously estimated (count_disorder_group_per_day)
+Percentage using as denominator total number of patients per day with a
+psy disorder.
+
+Per day - Total counts previously estimated (count_icd_per_day_psy). -
+Counts per disorder group previously estimated
+(count_disorder_group_per_day)
 
@@
 
-Per week
-- Counts per disorder group previously estimated (count_disorder_group_per_week)
+Per week - Counts per disorder group previously estimated
+(count_disorder_group_per_week)
 
@@
 Months
+
 ```{r}
@@
 
-Per day
-- Counts per disorder group previously estimated (count_disorder_group_per_day)
+Per day - Counts per disorder group previously estimated
+(count_disorder_group_per_day)
 

### 2021-09-29 psy_peds_3.Rmd
age at diagnosis

@@
   inner_join(distinct(observation_table_ado, patient_num, days_since_admission, age_time_diagnosis), 
-            by = c("patient_num", "days_since_admission"))
+            by = c("patient_num", "days_since_admission")) %>% 
+  select(patient_num, n_hospitalisation, age_time_diagnosis)
 
@@
 
-observation_table_ado <- clinical_table %>% 
-  inner_join(observation_table[, c("patient_num", "days_since_admission")], 
-             by = c("patient_num", "days_since_admission"))
-
 # Filtering observation tables which patient_num are still in clinical table after filtering on age
@@
 all_codes_ado <- all_codes %>% 
-   filter(age_time_visit >= 11 & age_time_visit <= 17)
+   filter(age_time_diagnosis >= 11 & age_time_diagnosis <= 17)
 

### 2021-09-29 psy_peds_3.Rmd
age at diagnosis

@@
 
-```{r}
-
-# Steps necessary to filter rows:
-  # - encounters without ICD
-  # - ICD diagnosed outside of encounters
-  # - patients from the patients table that don't have any encounters matching the criteria
-clinical_table <- clinical_filtered %>% 
-  inner_join( distinct(obs_filtered, patient_num, days_since_admission),  by = c("patient_num", "days_since_admission")) %>% 
-  distinct(patient_num, n_hospitalisation) %>% 
-  inner_join(clinical_filtered, by = c("patient_num", "n_hospitalisation"))
-# Adding new columns
-observation_table <- obs_filtered %>% 
-  inner_join(clinical_table[, c("patient_num", "days_since_admission")], by = c("patient_num", "days_since_admission"))
-demographic_table <- demo_raw %>% 
-  inner_join(distinct(clinical_table, patient_num), by = "patient_num")
-all_codes <- observation_table %>% 
-   inner_join(clinical_table, by = c("patient_num", "days_since_admission")) 
-#%>% 
-#   inner_join(dplyr::select(demographic_table, -days_since_admission), by = c("patient_num"))
-
-```
-
 ### Total hospitalize patients

### 2021-09-29 psy_peds_3.Rmd
graphics sorted and filtered by week only

@@
 all_codes <- observation_table %>% 
-   inner_join(clinical_table, by = c("patient_num", "days_since_admission")) %>% 
+   inner_join(select(clinical_table, -calendar_date), by = c("patient_num", "days_since_admission")) %>% 
    inner_join(demographic_table[, c("patient_num", "sex")], by = "patient_num")
@@
 ```{r}
-
 # Filtering hospitalizations where patient age is in inclusion criteria
@@
             by = c("patient_num", "days_since_admission")) %>% 
-  select(patient_num, n_hospitalisation, age_time_diagnosis)
+  distinct(patient_num, n_hospitalisation, age_time_diagnosis)
 
 clinical_table_ado <- clinical_table %>% 
-  inner_join(clinical_test[, c("patient_num", "n_hospitalisation", "age_time_diagnosis")],
+  inner_join(clinical_test,
              by = c("patient_num", "n_hospitalisation"))
@@
    filter(age_time_diagnosis >= 11 & age_time_diagnosis <= 17)
-
-head(observation_table_ado)
-head(clinical_table_ado)
-head(all_codes_ado)
 ```
@@
 
-### Total hospitalize patients
+### Total hospitalize patients age range 11-17
 
-```{r, fig.height=12, fig.width=12}
-total_hospitalizations_per_day <- clinical_filtered %>%
-  filter(date < end_date_plots &
-         date >= start_date_plots ) %>%
-    group_by( date ) %>%
-  summarise(count = n_distinct(patient_num))
-
-unique_patients_hospitalized_per_week <- clinical_filtered %>%
+```{r}
+unique_patients_hospitalized_per_week <- clinical_table_ado %>%
   filter(date < end_date_plots &
@@
 
-unique_patients_hospitalized_per_month <- clinical_filtered %>%
-  filter(date < end_date_plots &
-         date >= start_date_plots ) %>%
-    group_by( month ) %>%
-  summarise(count = n_distinct(patient_num))
-
-th_day <- ggplot(total_hospitalizations_per_day, aes(x = date,
-                       y = count))  +
-  geom_line() +
-  scale_x_date(date_breaks = "1 month", date_labels =  "%b %Y") +
-  theme(axis.text.x=element_text(angle=60, hjust=1))+
-  labs(x = "Calendar date",
-       y = "Patients count", 
-       title = "Currently hospitalized patients (per day)")
-
-th_week <- ggplot(unique_patients_hospitalized_per_week, aes(x = weeks,
-                       y = count))  +
-  geom_line() +
-  scale_x_date(date_breaks = "1 month", date_labels =  "%b %Y") +
-  theme(axis.text.x=element_text(angle=60, hjust=1))+
-  labs(x = "Calendar date",
-       y = "Patients count", 
-       title = "Currently hospitalized patients (per week)")
-
-th_month <- ggplot(unique_patients_hospitalized_per_month, aes(x = month,
-                       y = count))  +
-  geom_line() +
-  scale_x_date(date_breaks = "1 month", date_labels =  "%b %Y") +
-  theme(axis.text.x=element_text(angle=60, hjust=1))+
-  labs(x = "Calendar date",
-       y = "Patients count", 
-       title = "Currently hospitalized patients (per month)")
-
-require(gridExtra)
-grid.arrange(th_day, th_week, th_month, ncol=1)
+th_week <- admission_plot_count(df = unique_patients_hospitalized_per_week, 
+                     x = "weeks",
+                     y = "count", 
+                     title = "Total hospitalization (per week)", 
+                     y_title = "Patient Count") +
+  labs( subtitle = "(age range 11-17)")
+th_week
 ```
 
-#### Group by sex
-
-#### Per week
-
-```{r}
-# counts per week and sex
-patients_hospitalized_per_week_agg_sex <- clinical_filtered %>%
-  filter(date < end_date_plots &
-         date >= start_date_plots & 
-           sex %in% c("female", "male")) %>%
-    group_by( weeks, sex ) %>%
-  summarise(count_sex = n_distinct(patient_num))
-
-# the total counts per week was estimated before: unique_patients_hospitalized_per_week
-patients_hospitalized_per_week_agg_sex_perc <- unique_patients_hospitalized_per_week %>%
-  left_join( patients_hospitalized_per_week_agg_sex, by = c("weeks")) %>% 
-  replace_na(list(count_sex = 0))
-
-patients_hospitalized_per_week_agg_sex_perc$percentage <- 100*patients_hospitalized_per_week_agg_sex_perc$count_sex/patients_hospitalized_per_week_agg_sex_perc$count
-
-patients_hospitalized_per_week_agg_sex %>%
-  ggplot(aes(x = weeks, y = count_sex, fill = sex)) +
-  geom_bar(stat="identity", position = "dodge")+
-  scale_fill_manual(values = cbPalette) +
-  labs(y = "Counts",
-       x = "Week",
-       title = "Weekly patient counts aggregated by sex",
-       subtitle = "Per Sex") + 
-  theme(legend.position = "bottom") + 
-  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=sex), colour= "black", size = 0.5)
-
-patients_hospitalized_per_week_agg_sex_perc %>%
-  ggplot(aes(x = weeks, y = percentage, fill = sex)) +
-  geom_bar(stat="identity", position = "dodge")+
-  scale_fill_manual(values = cbPalette) +
-  labs(y = "Percentage",
-       x = "Week",
-       title = "Weekly patient (%) aggregated by sex",
-       subtitle = "Per Sex") + 
-  theme(legend.position = "bottom") + 
-  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=sex), colour= "black", size = 0.5)
-```
-
-## Second estimation: total number of patients filtering by psychiatric disorders
+## Total number of patients filtering by psychiatric disorders
 
@@
 
-Creation of the inputTable that will be the only table used throughout
-the rest of the analysis - selecting only visits with at least one ICD
-codes pertaining to the mental health categories - filter visits where
-patients are between 11 and 17 years old - add the demographic and
-hospitalization related information - add the ICD code description and
-the related disorder group
+Creation of the inputTable that will be the only table used throughout the rest of the analysis - selecting only visits with at least one ICD
+codes pertaining to the mental health categories - filter visits where patients are between 11 and 17 years old - add the demographic and
+hospitalization related information - add the ICD code description and the related disorder group
 
@@
 
-all_codes_desc <- all_codes %>% 
+all_codes_desc <- all_codes_ado %>% 
     left_join( icdCodes, by=c("concept_code" = "ICD10_Code") ) %>%
@@
                                "others")) %>%
-  dplyr::select( patient_num, sex, age_time_visit, concept_type, concept_code, calendar_date,
+  dplyr::select( patient_num, sex, age_time_diagnosis, concept_type, concept_code, calendar_date,
                  in_hospital, n_hospitalisation, len_hospitalisation, date, weeks, month, year,
@@
 }
+```
+
+### Total patients hospitalized comparing psychiatric disorders vs other
+
+```{r}
+count_icd_per_week_psy <- all_codes_psy %>%
+  filter(date < end_date_plots &
+         date >= start_date_plots ) %>%
+    group_by( weeks ) %>%
+  summarise(count_psy = n_distinct(patient_num)) %>%
+  mutate( period = ifelse( weeks < pandemic_start_date,
+                           "before_pandemic",
+                           "during_pandemic")
+  )
+
+count_icd_per_week <- all_codes_desc %>%
+  filter(date < end_date_plots &
+         date >= start_date_plots ) %>%
+    group_by( weeks ) %>%
+  summarise(count_icd = n_distinct(patient_num)) %>%
+  mutate( period = ifelse( weeks < pandemic_start_date,
+                           "before_pandemic",
+                           "during_pandemic")
+  )
+
+count_icd_per_week <- count_icd_per_week %>%
+  left_join(count_icd_per_week_psy, 
+            by = c("weeks", "period")) %>% 
+  replace_na(list(count_psy = 0))
+
+count_icd_per_week$percentage_psy <- 100*count_icd_per_week$count_psy / count_icd_per_week$count_icd
+
+thpd_psyCount <- count_icd_per_week %>%
+  pivot_longer(names_to = "counts", cols=c(count_psy, count_icd)) %>%
+  mutate( counts = ifelse( counts == "count_psy", "Psy", "Total")) %>%
+  ggplot(aes(x = weeks, y = value, fill = counts, color = counts)) +
+  geom_point() +
+  # geom_line() +
+  geom_vline(xintercept = as.Date(pandemic_start_date),
+             linetype = "dashed") +
+  scale_fill_manual(values = cbPalette) +
+  scale_color_manual(values = cbPalette) +
+  labs(y = "Counts",
+       x = "Date (by week)",
+       title = "Weekly patient counts with mental health related ICD codes") + 
+  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)
+thpd_psyCount
+
+```
+
+```{r}
+thpd_psyPer <- count_icd_per_week %>%
+  ggplot(aes(x = weeks, y = percentage_psy, fill=period)) +
+  scale_fill_manual(values = cbPalette[c(3,4)]) +
+  geom_vline(xintercept = as.Date(pandemic_start_date),
+             linetype = "dashed") +
+  labs(y = "Percentage",
+       x = "Date (by week)",
+       title = "Percentage of patients with mental health related ICD codes (weekly)") + 
+  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)+
+  geom_point()
+thpd_psyPer
+```
+
+#### Psychiatric conditions aggregating by suptype
+
+##### Total hospitalizations Per week
+
+```{r}
+count_disorder_group_per_week <- all_codes_psy %>%
+  filter(date < end_date_plots &
+         date >= start_date_plots ) %>%
+    group_by( weeks, disorder_group ) %>%
+  summarise(count_dg = n_distinct(patient_num)) %>%
+  mutate( period = ifelse( weeks < pandemic_start_date,
+                           "before_pandemic",
+                           "during_pandemic")
+  )
+
+dg_per_week <- count_disorder_group_per_week %>%  
+  ggplot(aes(x = weeks, y= count_dg, fill = period, group = disorder_group )) +
+  geom_bar(stat = "identity")+
+  geom_vline(xintercept = as.Date(pandemic_start_date),
+             linetype = "dashed") +
+  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se= TRUE, aes(colour=period), colour= "black", size = 0.5) + 
+  labs(x = "Weeks",
+       y = "Patient Count",
+       title = "Weekly patient counts with mental health related ICD codes", 
+       subtitle = "Grouped by mental health ICD subcategories") + 
+  ylim(0, NA) + 
+  facet_wrap(. ~ disorder_group, scales = "free_y") +
+  theme(strip.text.x = element_text(size = 4),
+        axis.text.x = element_text(size = 5, angle = 90), 
+        axis.text.y = element_text(size = 6), 
+        axis.title.x = element_text(size = 10),
+        axis.title.y = element_text(size = 10),
+        title = element_text(size = 12),
+        plot.subtitle = element_text(size = 11),
+        legend.position = "none") +
+  scale_fill_manual(values = cbPalette[-1])
+
+dg_per_week
+```
+
+Percentage using as denominator total number of patients
+
+```{r}
+# estimate the weekly counts
+count_icd_per_week <- all_codes_desc %>%
+  filter(date < end_date_plots &
+         date >= start_date_plots ) %>%
+    group_by( weeks ) %>%
+  summarise(count_icd = n_distinct(patient_num)) %>%
+  mutate( period = ifelse( weeks < pandemic_start_date,
+                           "before_pandemic",
+                           "during_pandemic")
+  )
+
+# join the total with the disorder group count to estimate the percentage
+perc_disorder_group_per_week <- count_icd_per_week %>%
+  left_join(count_disorder_group_per_week, 
+            by = c("weeks", "period")) %>% 
+  replace_na(list(count_dg = 0))
+
+perc_disorder_group_per_week$percentage_dg <- 100*perc_disorder_group_per_week$count_dg / perc_disorder_group_per_week$count_icd
+
+dgPerc_per_week <- perc_disorder_group_per_week %>%  
+  ggplot(aes(x = weeks, y= percentage_dg, fill = period, group = disorder_group )) +
+  geom_bar(stat = "identity")+
+  geom_vline(xintercept = as.Date(pandemic_start_date),
+             linetype = "dashed") +
+  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se= TRUE, aes(colour=period), colour= "black", size = 0.5) + 
+  labs(x = "Weeks",
+       y = "Patient (%)",
+       title = "Percentage of weekly patients with mental health related ICD codes", 
+       subtitle = "Grouped by mental health ICD subcategories") + 
+  ylim(0, NA) + 
+  facet_wrap(. ~ disorder_group, scales = "free_y") +
+  theme(strip.text.x = element_text(size = 4),
+        axis.text.x = element_text(size = 5, angle = 90), 
+        axis.text.y = element_text(size = 6), 
+        axis.title.x = element_text(size = 10),
+        axis.title.y = element_text(size = 10),
+        title = element_text(size = 12),
+        plot.subtitle = element_text(size = 11),
+        legend.position = "none") +
+  scale_fill_manual(values = cbPalette[-1])
+
+dgPerc_per_week
 ```
@@
   geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=sex), colour= "black", size = 0.5)
-
 ```
-
-#### Psychiatric related conditions vs. any other condition
-
-##### New admissions
-
-##### Total hospitalizations
-
-We have estimated the total number of hospitalizations in the previous
-steps: - total_hospitalization_per_day - total_hospitalization_per_week
-- total_hospitalization_per_month
-
-We will use these counts to estimate the percentage and to compare total
-counts, vs psychiatric related conditions.
-
-```{r}
-count_icd_per_day_psy <- all_codes_psy %>%
-  filter(date < end_date_plots &
-         date >= start_date_plots ) %>%
-    group_by( date ) %>%
-  summarise(count_psy = n_distinct(patient_num)) %>%
-  mutate( period = ifelse( date < pandemic_start_date,
-                           "before_pandemic",
-                           "during_pandemic")
-  )
-
-count_icd_per_day <- all_codes_desc %>%
-  filter(date < end_date_plots &
-         date >= start_date_plots ) %>%
-    group_by( date ) %>%
-  summarise(count_icd = n_distinct(patient_num)) %>%
-  mutate( period = ifelse( date < pandemic_start_date,
-                           "before_pandemic",
-                           "during_pandemic")
-  )
-
-count_icd_per_day <- count_icd_per_day %>%
-  left_join(count_icd_per_day_psy, 
-            by = c("date", "period")) %>% 
-  replace_na(list(count_psy = 0))
-
-count_icd_per_day$percentage_psy <- 100*count_icd_per_day$count_psy / count_icd_per_day$count_icd
-
-thpd_psyPer <- count_icd_per_day %>%
-  ggplot(aes(x = date, y = percentage_psy, fill = period)) +
-  scale_fill_manual(values = cbPalette[c(3,4)]) +
-  labs(y = "Percentage",
-       x = "Date (by day)",
-       title = "Percentage of patients with mental health related ICD codes (daily)") + 
-  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)+
-  geom_point()
-thpd_psyPer
-
-count_icd_per_day %>%
-  ggplot(aes(x = date, y = percentage_psy, fill = period)) +
-  scale_fill_manual(values = cbPalette[c(3,4)]) +
-  labs(y = "Percentage",
-       x = "Date (by day)",
-       title = "Percentage of patients with mental health related ICD codes (daily)") + 
-  geom_point()
-  #geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)
-```
-
-By week
-
-```{r}
-count_icd_per_week_psy <- all_codes_psy %>%
-  filter(date < end_date_plots &
-         date >= start_date_plots ) %>%
-    group_by( weeks ) %>%
-  summarise(count_psy = n_distinct(patient_num)) %>%
-  mutate( period = ifelse( weeks < pandemic_start_date,
-                           "before_pandemic",
-                           "during_pandemic")
-  )
-
-count_icd_per_week <- all_codes_desc %>%
-  filter(date < end_date_plots &
-         date >= start_date_plots ) %>%
-    group_by( weeks ) %>%
-  summarise(count_icd = n_distinct(patient_num)) %>%
-  mutate( period = ifelse( weeks < pandemic_start_date,
-                           "before_pandemic",
-                           "during_pandemic")
-  )
-
-count_icd_per_week <- count_icd_per_week %>%
-  left_join(count_icd_per_week_psy, 
-            by = c("weeks", "period")) %>% 
-  replace_na(list(count_psy = 0))
-
-count_icd_per_week$percentage_psy <- 100*count_icd_per_week$count_psy / count_icd_per_week$count_icd
-
-thpd_psyPer <- count_icd_per_week %>%
-  ggplot(aes(x = weeks, y = percentage_psy, fill=period)) +
-  scale_fill_manual(values = cbPalette[c(3,4)]) +
-  labs(y = "Percentage",
-       x = "Date (by week)",
-       title = "Percentage of patients with mental health related ICD codes (weekly)") + 
-  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)+
-  geom_point()
-thpd_psyPer
-
-count_icd_per_week %>%
-  ggplot(aes(x = weeks, y = percentage_psy, fill = period)) +
-  scale_fill_manual(values = cbPalette[c(3,4)]) +
-  labs(y = "Percentage",
-       x = "Date (by week)",
-       title = "Percentage of patients with mental health related ICD codes (weekly)") + 
-  geom_point()
-  #geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)
-```
-
-# Per day
-
-```{r}
-thpd_psyCount <- count_icd_per_day %>%
-  pivot_longer(names_to = "counts", cols=c(count_psy, count_icd)) %>%
-  mutate( counts = ifelse( counts == "count_psy", "Psy", "Total")) %>%
-  ggplot(aes(x = date, y = value, fill = counts, color = counts)) +
-  geom_point() +
-  # geom_line() +
-  geom_vline(xintercept = as.Date(pandemic_start_date),
-             linetype = "dashed") +
-  scale_fill_manual(values = cbPalette) +
-  scale_color_manual(values = cbPalette) +
-  labs(y = "Counts",
-       x = "Date (by day)",
-       title = "Daily patient counts with mental health related ICD codes") + 
-  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)
-thpd_psyCount
-
-```
-
-# Per week
-
-```{r}
-thpd_psyCount <- count_icd_per_week %>%
-  pivot_longer(names_to = "counts", cols=c(count_psy, count_icd)) %>%
-  mutate( counts = ifelse( counts == "count_psy", "Psy", "Total")) %>%
-  ggplot(aes(x = weeks, y = value, fill = counts, color = counts)) +
-  geom_point() +
-  # geom_line() +
-  geom_vline(xintercept = as.Date(pandemic_start_date),
-             linetype = "dashed") +
-  scale_fill_manual(values = cbPalette) +
-  scale_color_manual(values = cbPalette) +
-  labs(y = "Counts",
-       x = "Date (by week)",
-       title = "Weekly patient counts with mental health related ICD codes") + 
-  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)
-thpd_psyCount
-
-```
-
-We will compare the psychiatric conditions with some other disorders
-(e.g, digestive, respiratory, etc.) We will use these counts to estimate
-the percentage and to compare total counts, vs psychiatric related
-conditions.
-
-```{r}
-total_hospitalizations_per_day_disorder_group <- all_codes_desc %>%
-  filter(date < end_date_plots &
-         date >= start_date_plots & 
-        icd_description_category %in% c("Psy","Neoplams","Diseases of the respiratory system","Diseases of the digestive system" ))%>%
-    group_by( date, icd_description_category ) %>%
-  summarise(counts = n_distinct(patient_num))
-
-thpd_disorderGroup <- total_hospitalizations_per_day_disorder_group %>%
-  ggplot(aes(x = date, y = counts, fill = icd_description_category, color = icd_description_category)) +
-  #geom_point(alpha=0.2) +
-  geom_line() +
-  geom_vline(xintercept = as.Date(pandemic_start_date),
-             linetype = "dashed") +
-  scale_fill_manual(values = cbPalette) +
-  scale_color_manual(values = cbPalette) +
-  labs(y = "patient number",
-       x = "days",
-       title = "Total number of hospitalized patients (daily)") +   
-  theme(axis.text.x = element_text(angle=45, vjust = 1, hjust=1), legend.position = "bottom", legend.text = element_text(size=7), legend.title = element_text(size=7) ) +
-  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)
-thpd_disorderGroup
-```
-
-#### Psychiatric conditions aggregating by suptype
-
-##### New admissions
-
-##### Total hospitalizations
-
-We estimate and plot the counts per disorder group.
-
-Per day
-
-```{r}
-count_disorder_group_per_day <- all_codes_psy %>%
-  filter(date < end_date_plots &
-         date >= start_date_plots ) %>%
-    group_by( date, disorder_group ) %>%
-  summarise(count_dg = n_distinct(patient_num)) %>%
-  mutate( period = ifelse( date < pandemic_start_date,
-                           "before_pandemic",
-                           "during_pandemic")
-  )
-
-dg_per_day <- count_disorder_group_per_day %>%  
-  ggplot(aes(x = date, y= count_dg, fill = period, group = disorder_group )) +
-  geom_bar(stat = "identity")+
-  geom_vline(xintercept = as.Date(pandemic_start_date),
-             linetype = "dashed") +
-  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se= TRUE, aes(colour=period), colour= "black", size = 0.5) + 
-  labs(x = "Days",
-       y = "Patient Count",
-       title = "Daily patient counts with mental health related ICD codes", 
-       subtitle = "Grouped by mental health ICD subcategories") + 
-  ylim(0, NA) + 
-  facet_wrap(. ~ disorder_group, scales = "free_y") +
-  theme(strip.text.x = element_text(size = 4),
-        axis.text.x = element_text(size = 5, angle = 90), 
-        axis.text.y = element_text(size = 6), 
-        axis.title.x = element_text(size = 10),
-        axis.title.y = element_text(size = 10),
-        title = element_text(size = 12),
-        plot.subtitle = element_text(size = 11),
-        legend.position = "none") +
-  scale_fill_manual(values = cbPalette[-1])
-
-dg_per_day
-```
-
-Per week
-
-```{r}
-count_disorder_group_per_week <- all_codes_psy %>%
-  filter(date < end_date_plots &
-         date >= start_date_plots ) %>%
-    group_by( weeks, disorder_group ) %>%
-  summarise(count_dg = n_distinct(patient_num)) %>%
-  mutate( period = ifelse( weeks < pandemic_start_date,
-                           "before_pandemic",
-                           "during_pandemic")
-  )
-
-dg_per_week <- count_disorder_group_per_week %>%  
-  ggplot(aes(x = weeks, y= count_dg, fill = period, group = disorder_group )) +
-  geom_bar(stat = "identity")+
-  geom_vline(xintercept = as.Date(pandemic_start_date),
-             linetype = "dashed") +
-  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se= TRUE, aes(colour=period), colour= "black", size = 0.5) + 
-  labs(x = "Weeks",
-       y = "Patient Count",
-       title = "Weekly patient counts with mental health related ICD codes", 
-       subtitle = "Grouped by mental health ICD subcategories") + 
-  ylim(0, NA) + 
-  facet_wrap(. ~ disorder_group, scales = "free_y") +
-  theme(strip.text.x = element_text(size = 4),
-        axis.text.x = element_text(size = 5, angle = 90), 
-        axis.text.y = element_text(size = 6), 
-        axis.title.x = element_text(size = 10),
-        axis.title.y = element_text(size = 10),
-        title = element_text(size = 12),
-        plot.subtitle = element_text(size = 11),
-        legend.position = "none") +
-  scale_fill_manual(values = cbPalette[-1])
-
-dg_per_week
-```
-
-Per month
-
-```{r}
-count_disorder_group_per_month <- all_codes_psy %>%
-  filter(date < end_date_plots &
-         date >= start_date_plots ) %>%
-    group_by( month, disorder_group ) %>%
-  summarise(count_dg = n_distinct(patient_num)) %>%
-  mutate( period = ifelse( month < pandemic_start_date,
-                           "before_pandemic",
-                           "during_pandemic")
-  )
-
-dg_per_month <- count_disorder_group_per_month %>%  
-  ggplot(aes(x = month, y= count_dg, fill = period, group = disorder_group )) +
-  geom_bar(stat = "identity")+
-  geom_vline(xintercept = as.Date(pandemic_start_date),
-             linetype = "dashed") +
-  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se= TRUE, aes(colour=period), colour= "black", size = 0.5) + 
-  labs(x = "Months",
-       y = "Patient Count",
-       title = "Monthly patient counts with mental health related ICD codes", 
-       subtitle = "Grouped by mental health ICD subcategories") + 
-  ylim(0, NA) + 
-  facet_wrap(. ~ disorder_group, scales = "free_y") +
-  theme(strip.text.x = element_text(size = 4),
-        axis.text.x = element_text(size = 5, angle = 90), 
-        axis.text.y = element_text(size = 6), 
-        axis.title.x = element_text(size = 10),
-        axis.title.y = element_text(size = 10),
-        title = element_text(size = 12),
-        plot.subtitle = element_text(size = 11),
-        legend.position = "none") +
-  scale_fill_manual(values = cbPalette[-1])
-
-dg_per_month
-```
-
-Percentage using as denominator total number of patients per day with a
-psy disorder.
-
-Per day - Total counts previously estimated (count_icd_per_day_psy). -
-Counts per disorder group previously estimated
-(count_disorder_group_per_day)
-
-```{r}
-# join the total with the disorder group count to estimate the percentage
-perc_disorder_group_per_day <- count_icd_per_day_psy %>%
-  left_join(count_disorder_group_per_day, 
-            by = c("date", "period")) %>% 
-  replace_na(list(count_dg = 0))
-
-perc_disorder_group_per_day$percentage_dg <- 100*perc_disorder_group_per_day$count_dg / perc_disorder_group_per_day$count_psy
-
-dgPerc_per_day <- perc_disorder_group_per_day %>%  
-  ggplot(aes(x = date, y= percentage_dg, fill = period, group = disorder_group )) +
-  geom_bar(stat = "identity")+
-  geom_vline(xintercept = as.Date(pandemic_start_date),
-             linetype = "dashed") +
-  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se= TRUE, aes(colour=period), colour= "black", size = 0.5) + 
-  labs(x = "Days",
-       y = "Patient (%)",
-       title = "Percentage of daily patients with mental health related ICD codes", 
-       subtitle = "Grouped by mental health ICD subcategories") + 
-  ylim(0, NA) + 
-  facet_wrap(. ~ disorder_group, scales = "free_y") +
-  theme(strip.text.x = element_text(size = 4),
-        axis.text.x = element_text(size = 5, angle = 90), 
-        axis.text.y = element_text(size = 6), 
-        axis.title.x = element_text(size = 10),
-        axis.title.y = element_text(size = 10),
-        title = element_text(size = 12),
-        plot.subtitle = element_text(size = 11),
-        legend.position = "none") +
-  scale_fill_manual(values = cbPalette[-1])
-
-dgPerc_per_day
-```
-
-Per week - Counts per disorder group previously estimated
-(count_disorder_group_per_week)
-
-```{r}
-# estimate the weekly counts
-count_icd_per_week_psy <- all_codes_psy %>%
-  filter(date < end_date_plots &
-         date >= start_date_plots ) %>%
-    group_by( weeks ) %>%
-  summarise(count_psy = n_distinct(patient_num)) %>%
-  mutate( period = ifelse( weeks < pandemic_start_date,
-                           "before_pandemic",
-                           "during_pandemic")
-  )
-
-# join the total with the disorder group count to estimate the percentage
-perc_disorder_group_per_week <- count_icd_per_week_psy %>%
-  left_join(count_disorder_group_per_week, 
-            by = c("weeks", "period")) %>% 
-  replace_na(list(count_dg = 0))
-
-perc_disorder_group_per_week$percentage_dg <- 100*perc_disorder_group_per_week$count_dg / perc_disorder_group_per_week$count_psy
-
-dgPerc_per_week <- perc_disorder_group_per_week %>%  
-  ggplot(aes(x = weeks, y= percentage_dg, fill = period, group = disorder_group )) +
-  geom_bar(stat = "identity")+
-  geom_vline(xintercept = as.Date(pandemic_start_date),
-             linetype = "dashed") +
-  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se= TRUE, aes(colour=period), colour= "black", size = 0.5) + 
-  labs(x = "Weeks",
-       y = "Patient (%)",
-       title = "Percentage of weekly patients with mental health related ICD codes", 
-       subtitle = "Grouped by mental health ICD subcategories") + 
-  ylim(0, NA) + 
-  facet_wrap(. ~ disorder_group, scales = "free_y") +
-  theme(strip.text.x = element_text(size = 4),
-        axis.text.x = element_text(size = 5, angle = 90), 
-        axis.text.y = element_text(size = 6), 
-        axis.title.x = element_text(size = 10),
-        axis.title.y = element_text(size = 10),
-        title = element_text(size = 12),
-        plot.subtitle = element_text(size = 11),
-        legend.position = "none") +
-  scale_fill_manual(values = cbPalette[-1])
-
-dgPerc_per_week
-```
-
-Months
-
-```{r}
-# estimate the monthly counts
-count_icd_per_month_psy <- all_codes_psy %>%
-  filter(date < end_date_plots &
-         date >= start_date_plots ) %>%
-    group_by( month ) %>%
-  summarise(count_psy = n_distinct(patient_num)) %>%
-  mutate( period = ifelse( month < pandemic_start_date,
-                           "before_pandemic",
-                           "during_pandemic")
-  )
-
-# join the total with the disorder group count to estimate the percentage
-perc_disorder_group_per_month <- count_icd_per_month_psy %>%
-  left_join(count_disorder_group_per_month, 
-            by = c("month", "period")) %>% 
-  replace_na(list(count_dg = 0))
-
-perc_disorder_group_per_month$percentage_dg <- 100*perc_disorder_group_per_month$count_dg / perc_disorder_group_per_month$count_psy
-
-dgPerc_per_month <- perc_disorder_group_per_month %>%  
-  ggplot(aes(x = month, y= percentage_dg, fill = period, group = disorder_group )) +
-  geom_bar(stat = "identity")+
-  geom_vline(xintercept = as.Date(pandemic_start_date),
-             linetype = "dashed") +
-  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se= TRUE, aes(colour=period), colour= "black", size = 0.5) + 
-  labs(x = "Months",
-       y = "Patient (%)",
-       title = "Percentage of monthly patients with mental health related ICD codes", 
-       subtitle = "Grouped by mental health ICD subcategories") + 
-  ylim(0, NA) + 
-  facet_wrap(. ~ disorder_group, scales = "free_y") +
-  theme(strip.text.x = element_text(size = 4),
-        axis.text.x = element_text(size = 5, angle = 90), 
-        axis.text.y = element_text(size = 6), 
-        axis.title.x = element_text(size = 10),
-        axis.title.y = element_text(size = 10),
-        title = element_text(size = 12),
-        plot.subtitle = element_text(size = 11),
-        legend.position = "none") +
-  scale_fill_manual(values = cbPalette[-1])
-
-dgPerc_per_month
-```
-
-Percentage using as denominator total number of patients
-
-Per day - Counts per disorder group previously estimated
-(count_disorder_group_per_day)
-
-```{r}
-# estimate the total number per day 
-count_icd_per_day <- all_codes_desc %>%
-  filter(date < end_date_plots &
-         date >= start_date_plots ) %>%
-    group_by( date ) %>%
-  summarise(count_icd = n_distinct(patient_num)) %>%
-  mutate( period = ifelse( date < pandemic_start_date,
-                           "before_pandemic",
-                           "during_pandemic")
-  )
-
-# join the total with the disorder group count to estimate the percentage
-perc_disorder_group_per_day <- count_icd_per_day %>%
-  left_join(count_disorder_group_per_day, 
-            by = c("date", "period")) %>% 
-  replace_na(list(count_dg = 0))
-
-perc_disorder_group_per_day$percentage_dg <- 100*perc_disorder_group_per_day$count_dg / perc_disorder_group_per_day$count_icd
-
-dgPerc_per_day <- perc_disorder_group_per_day %>%  
-  ggplot(aes(x = date, y= percentage_dg, fill = period, group = disorder_group )) +
-  geom_bar(stat = "identity")+
-  geom_vline(xintercept = as.Date(pandemic_start_date),
-             linetype = "dashed") +
-  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se= TRUE, aes(colour=period), colour= "black", size = 0.5) + 
-  labs(x = "Days",
-       y = "Patient (%)",
-       title = "Percentage of daily patients with mental health related ICD codes", 
-       subtitle = "Grouped by mental health ICD subcategories") + 
-  ylim(0, NA) + 
-  facet_wrap(. ~ disorder_group, scales = "free_y") +
-  theme(strip.text.x = element_text(size = 4),
-        axis.text.x = element_text(size = 5, angle = 90), 
-        axis.text.y = element_text(size = 6), 
-        axis.title.x = element_text(size = 10),
-        axis.title.y = element_text(size = 10),
-        title = element_text(size = 12),
-        plot.subtitle = element_text(size = 11),
-        legend.position = "none") +
-  scale_fill_manual(values = cbPalette[-1])
-
-dgPerc_per_day
-```
-
-Per week - Counts per disorder group previously estimated
-(count_disorder_group_per_week)
-
-```{r}
-# estimate the weekly counts
-count_icd_per_week <- all_codes_desc %>%
-  filter(date < end_date_plots &
-         date >= start_date_plots ) %>%
-    group_by( weeks ) %>%
-  summarise(count_icd = n_distinct(patient_num)) %>%
-  mutate( period = ifelse( weeks < pandemic_start_date,
-                           "before_pandemic",
-                           "during_pandemic")
-  )
-
-# join the total with the disorder group count to estimate the percentage
-perc_disorder_group_per_week <- count_icd_per_week %>%
-  left_join(count_disorder_group_per_week, 
-            by = c("weeks", "period")) %>% 
-  replace_na(list(count_dg = 0))
-
-perc_disorder_group_per_week$percentage_dg <- 100*perc_disorder_group_per_week$count_dg / perc_disorder_group_per_week$count_icd
-
-dgPerc_per_week <- perc_disorder_group_per_week %>%  
-  ggplot(aes(x = weeks, y= percentage_dg, fill = period, group = disorder_group )) +
-  geom_bar(stat = "identity")+
-  geom_vline(xintercept = as.Date(pandemic_start_date),
-             linetype = "dashed") +
-  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se= TRUE, aes(colour=period), colour= "black", size = 0.5) + 
-  labs(x = "Weeks",
-       y = "Patient (%)",
-       title = "Percentage of weekly patients with mental health related ICD codes", 
-       subtitle = "Grouped by mental health ICD subcategories") + 
-  ylim(0, NA) + 
-  facet_wrap(. ~ disorder_group, scales = "free_y") +
-  theme(strip.text.x = element_text(size = 4),
-        axis.text.x = element_text(size = 5, angle = 90), 
-        axis.text.y = element_text(size = 6), 
-        axis.title.x = element_text(size = 10),
-        axis.title.y = element_text(size = 10),
-        title = element_text(size = 12),
-        plot.subtitle = element_text(size = 11),
-        legend.position = "none") +
-  scale_fill_manual(values = cbPalette[-1])
-
-dgPerc_per_week
-```
-
-Months
-
-```{r}
-# estimate the monthly counts
-count_icd_per_month <- all_codes_desc %>%
-  filter(date < end_date_plots &
-         date >= start_date_plots ) %>%
-    group_by( month ) %>%
-  summarise(count_icd = n_distinct(patient_num)) %>%
-  mutate( period = ifelse( month < pandemic_start_date,
-                           "before_pandemic",
-                           "during_pandemic")
-  )
-
-# join the total with the disorder group count to estimate the percentage
-perc_disorder_group_per_month <- count_icd_per_month %>%
-  left_join(count_disorder_group_per_month, 
-            by = c("month", "period")) %>% 
-  replace_na(list(count_dg = 0))
-
-perc_disorder_group_per_month$percentage_dg <- 100*perc_disorder_group_per_month$count_dg / perc_disorder_group_per_month$count_icd
-
-dgPerc_per_month <- perc_disorder_group_per_month %>%  
-  ggplot(aes(x = month, y= percentage_dg, fill = period, group = disorder_group )) +
-  geom_bar(stat = "identity")+
-  geom_vline(xintercept = as.Date(pandemic_start_date),
-             linetype = "dashed") +
-  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se= TRUE, aes(colour=period), colour= "black", size = 0.5) + 
-  labs(x = "Months",
-       y = "Patient (%)",
-       title = "Percentage of monthly patients with mental health related ICD codes", 
-       subtitle = "Grouped by mental health ICD subcategories") + 
-  ylim(0, NA) + 
-  facet_wrap(. ~ disorder_group, scales = "free_y") +
-  theme(strip.text.x = element_text(size = 4),
-        axis.text.x = element_text(size = 5, angle = 90), 
-        axis.text.y = element_text(size = 6), 
-        axis.title.x = element_text(size = 10),
-        axis.title.y = element_text(size = 10),
-        title = element_text(size = 12),
-        plot.subtitle = element_text(size = 11),
-        legend.position = "none") +
-  scale_fill_manual(values = cbPalette[-1])
-
-dgPerc_per_month
-```

### 2021-10-01 psy_peds_3.Rmd
obfuscation added and time period as a variable

@@
 data_update_date <- "2021-06-05"
+time_period <-"weeks" #possible values: days, weeks, months
 ```
@@
                                   "before_pandemic", "during_pandemic"))
+
+clinical_filtered$time_p <- NA
+if( time_period == "weeks"){
+  clinical_filtered$time_p <- clinical_filtered$weeks
+}else if(time_period == "days"){
+  clinical_filtered$time_p <- clinical_filtered$date
+}else if(time_period == "months"){
+  clinical_filtered$time_p <- clinical_filtered$month
+  }
 ```
@@
 ```{r}
-
-new_hospitalizations_per_week <- clinical_table %>%
+new_hospitalizations <- clinical_table %>%
   group_by(patient_num, n_hospitalisation) %>%
-  slice(which.min(weeks)) %>%
+  slice(which.min(time_p)) %>%
   filter( in_hospital == 1 ) %>%
-    group_by( weeks ) %>%
-  summarise(count = n_distinct(patient_num)) %>% 
+    group_by( time_p ) %>%
+  summarise(count = ifelse(n_distinct(patient_num) > obfuscation | isFALSE(obfuscation),
+                               n_distinct(patient_num), 
+                               0.5)) %>% 
   ungroup()
@@
 
-nh_week <- admission_plot_count(df = new_hospitalizations_per_week, 
-                     x = "weeks",
+nh <- admission_plot_count(df = new_hospitalizations, 
+                     x = "time_p",
                      y = "count", 
-                     title = "New admissions (per week)", 
+                     title = paste0("New admissions (per ", time_period,")"), 
                      y_title = "Patient Count")
 
-nh_week
+nh
 ```
@@
 
+### New admissions age range 11-17
+
+```{r}
+new_hospitalizations_ado <- clinical_table_ado %>%
+  group_by(patient_num, n_hospitalisation) %>%
+  slice(which.min(time_p)) %>%
+  filter( in_hospital == 1 ) %>%
+    group_by( time_p ) %>%
+  summarise(count = ifelse(n_distinct(patient_num) > obfuscation | isFALSE(obfuscation),
+                               n_distinct(patient_num), 
+                               0.5)) %>% 
+  ungroup()
+
+start_date_plots <- as.Date("2019-01-01")
+end_date_plots <- as.Date("2021-05-24")
+
+nh_ado <- admission_plot_count(df = new_hospitalizations_ado, 
+                     x = "time_p",
+                     y = "count", 
+                     title = paste0("New admissions (per ", time_period, ")"), 
+                     y_title = "Patient Count")+
+    labs( subtitle = "(age range 11-17)")
+
+nh_ado
+```
+
 ### Total hospitalize patients age range 11-17
@@
 ```{r}
-unique_patients_hospitalized_per_week <- clinical_table_ado %>%
+unique_patients_hospitalized <- clinical_table_ado %>%
   filter(date < end_date_plots &
          date >= start_date_plots ) %>%
-    group_by( weeks ) %>%
-  summarise(count = n_distinct(patient_num))
+    group_by( time_p ) %>%
+  summarise(count = ifelse(n_distinct(patient_num) > obfuscation | isFALSE(obfuscation),
+                               n_distinct(patient_num), 
+                               0.5))
 
-th_week <- admission_plot_count(df = unique_patients_hospitalized_per_week, 
-                     x = "weeks",
+th <- admission_plot_count(df = unique_patients_hospitalized, 
+                     x = "time_p",
                      y = "count", 
-                     title = "Total hospitalization (per week)", 
+                     title = paste0("Total hospitalization (per ", time_period, ")"), 
                      y_title = "Patient Count") +
   labs( subtitle = "(age range 11-17)")
-th_week
+th
 ```
@@
                  in_hospital, n_hospitalisation, len_hospitalisation, date, weeks, month, year,
-                 period, disorder_group, description, icd_code_category )
+                 period, time_p, disorder_group, description, icd_code_category )
 
@@
 ```{r}
-count_icd_per_week_psy <- all_codes_psy %>%
+count_icd_psy <- all_codes_psy %>%
   filter(date < end_date_plots &
          date >= start_date_plots ) %>%
-    group_by( weeks ) %>%
-  summarise(count_psy = n_distinct(patient_num)) %>%
-  mutate( period = ifelse( weeks < pandemic_start_date,
+    group_by( time_p ) %>%
+  summarise(count_psy =  ifelse(n_distinct(patient_num) > obfuscation | isFALSE(obfuscation),
+                               n_distinct(patient_num), 
+                               0.5)) %>%
+  mutate( period = ifelse( time_p < pandemic_start_date,
                            "before_pandemic",
@@
 
-count_icd_per_week <- all_codes_desc %>%
+count_icd <- all_codes_desc %>%
   filter(date < end_date_plots &
          date >= start_date_plots ) %>%
-    group_by( weeks ) %>%
-  summarise(count_icd = n_distinct(patient_num)) %>%
-  mutate( period = ifelse( weeks < pandemic_start_date,
+    group_by( time_p ) %>%
+  summarise(count_icd =  ifelse(n_distinct(patient_num) > obfuscation | isFALSE(obfuscation),
+                               n_distinct(patient_num), 
+                               0.5)) %>%
+  mutate( period = ifelse( time_p < pandemic_start_date,
                            "before_pandemic",
@@
 
-count_icd_per_week <- count_icd_per_week %>%
-  left_join(count_icd_per_week_psy, 
-            by = c("weeks", "period")) %>% 
+count_icd <- count_icd %>%
+  left_join(count_icd_psy, 
+            by = c("time_p", "period")) %>% 
   replace_na(list(count_psy = 0))
 
-count_icd_per_week$percentage_psy <- 100*count_icd_per_week$count_psy / count_icd_per_week$count_icd
+count_icd$percentage_psy <- 100*count_icd$count_psy / count_icd$count_icd
 
-thpd_psyCount <- count_icd_per_week %>%
+thpd_psyCount <- count_icd %>%
   pivot_longer(names_to = "counts", cols=c(count_psy, count_icd)) %>%
   mutate( counts = ifelse( counts == "count_psy", "Psy", "Total")) %>%
-  ggplot(aes(x = weeks, y = value, fill = counts, color = counts)) +
+  ggplot(aes(x = time_p, y = value, fill = counts, color = counts)) +
   geom_point() +
@@
   labs(y = "Counts",
-       x = "Date (by week)",
-       title = "Weekly patient counts with mental health related ICD codes") + 
+       x = paste0("Date (by ", time_period,")"),
+       title = paste0("Per ", time_period , " patient counts with mental health related ICD codes")) + 
   geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)
@@
 ```{r}
-thpd_psyPer <- count_icd_per_week %>%
-  ggplot(aes(x = weeks, y = percentage_psy, fill=period)) +
+thpd_psyPer <- count_icd %>%
+  ggplot(aes(x = time_p, y = percentage_psy, fill=period)) +
   scale_fill_manual(values = cbPalette[c(3,4)]) +
@@
   labs(y = "Percentage",
-       x = "Date (by week)",
-       title = "Percentage of patients with mental health related ICD codes (weekly)") + 
+       x = paste0("Date (by ", time_period,")"),
+       title = paste0("Percentage of patients with mental health related ICD codes (per ", time_period,")")) + 
   geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)+
@@
 
-##### Total hospitalizations Per week
+##### Total hospitalizations
 
 ```{r}
-count_disorder_group_per_week <- all_codes_psy %>%
+count_disorder_group <- all_codes_psy %>%
   filter(date < end_date_plots &
          date >= start_date_plots ) %>%
-    group_by( weeks, disorder_group ) %>%
-  summarise(count_dg = n_distinct(patient_num)) %>%
-  mutate( period = ifelse( weeks < pandemic_start_date,
+    group_by( time_p, disorder_group ) %>%
+  summarise(count_dg =  ifelse(n_distinct(patient_num) > obfuscation | isFALSE(obfuscation),
+                               n_distinct(patient_num), 
+                               0.5)) %>%
+  mutate( period = ifelse( time_p < pandemic_start_date,
                            "before_pandemic",
@@
 
-dg_per_week <- count_disorder_group_per_week %>%  
-  ggplot(aes(x = weeks, y= count_dg, fill = period, group = disorder_group )) +
+dg <- count_disorder_group %>%  
+  ggplot(aes(x = time_p, y= count_dg, fill = period, group = disorder_group )) +
   geom_bar(stat = "identity")+
@@
   geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se= TRUE, aes(colour=period), colour= "black", size = 0.5) + 
-  labs(x = "Weeks",
+  labs(x = time_period,
        y = "Patient Count",
-       title = "Weekly patient counts with mental health related ICD codes", 
+       title = paste0("Per ", time_period, " patient counts with mental health related ICD codes"), 
        subtitle = "Grouped by mental health ICD subcategories") + 
@@
 
-dg_per_week
+dg
 ```
@@
 ```{r}
-# estimate the weekly counts
-count_icd_per_week <- all_codes_desc %>%
+# estimate the counts
+count_icd <- all_codes_desc %>%
   filter(date < end_date_plots &
          date >= start_date_plots ) %>%
-    group_by( weeks ) %>%
-  summarise(count_icd = n_distinct(patient_num)) %>%
-  mutate( period = ifelse( weeks < pandemic_start_date,
+    group_by( time_p ) %>%
+  summarise(count_icd =  ifelse(n_distinct(patient_num) > obfuscation | isFALSE(obfuscation),
+                               n_distinct(patient_num), 
+                               0.5)) %>%
+  mutate( period = ifelse( time_p < pandemic_start_date,
                            "before_pandemic",
@@
 # join the total with the disorder group count to estimate the percentage
-perc_disorder_group_per_week <- count_icd_per_week %>%
-  left_join(count_disorder_group_per_week, 
-            by = c("weeks", "period")) %>% 
+perc_disorder_group <- count_icd %>%
+  left_join(count_disorder_group, 
+            by = c("time_p", "period")) %>% 
   replace_na(list(count_dg = 0))
 
-perc_disorder_group_per_week$percentage_dg <- 100*perc_disorder_group_per_week$count_dg / perc_disorder_group_per_week$count_icd
+perc_disorder_group$percentage_dg <- 100*perc_disorder_group$count_dg / perc_disorder_group$count_icd
 
-dgPerc_per_week <- perc_disorder_group_per_week %>%  
-  ggplot(aes(x = weeks, y= percentage_dg, fill = period, group = disorder_group )) +
+dgPerc <- perc_disorder_group %>%  
+  ggplot(aes(x = time_p, y= percentage_dg, fill = period, group = disorder_group )) +
   geom_bar(stat = "identity")+
@@
   geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se= TRUE, aes(colour=period), colour= "black", size = 0.5) + 
-  labs(x = "Weeks",
+  labs(x = time_period,
        y = "Patient (%)",
-       title = "Percentage of weekly patients with mental health related ICD codes", 
+       title = paste0("Percentage of patients per ", time_period,"with mental health related ICD codes"), 
        subtitle = "Grouped by mental health ICD subcategories") + 
@@
 
-dgPerc_per_week
+dgPerc
 ```
@@
 ```{r}
-# counts per week and sex
-patients_hospitalized_per_week_agg_sex <- all_codes_psy %>%
+# counts per time period and sex
+patients_hospitalized_agg_sex <- all_codes_psy %>%
   filter(date < end_date_plots &
@@
            sex %in% c("female", "male")) %>%
-    group_by( weeks, sex ) %>%
-  summarise(count_sex = n_distinct(patient_num))
+    group_by( time_p, sex ) %>%
+  summarise(count_sex =  ifelse(n_distinct(patient_num) > obfuscation | isFALSE(obfuscation),
+                               n_distinct(patient_num), 
+                               0.5))
 
-patients_hospitalized_per_week_agg_sex %>%
-  ggplot(aes(x = weeks, y = count_sex, fill = sex)) +
+patients_hospitalized_agg_sex %>%
+  ggplot(aes(x = time_p, y = count_sex, fill = sex)) +
   geom_bar(stat="identity", position = "dodge")+
@@
   labs(y = "Counts",
-       x = "Week",
-       title = "Weekly patient counts aggregated by sex",
+       x = time_period,
+       title = paste0("Patient counts per ", time_period, " aggregated by sex"),
        subtitle = "Per Sex") + 
@@
 ##percentage
-total_patients_psy_per_week <- all_codes_psy %>%
+total_patients_psy <- all_codes_psy %>%
   filter(date < end_date_plots &
          date >= start_date_plots ) %>%
-    group_by( weeks ) %>%
-  summarise(count = n_distinct(patient_num))
+    group_by( time_p ) %>%
+  summarise(count =  ifelse(n_distinct(patient_num) > obfuscation | isFALSE(obfuscation),
+                               n_distinct(patient_num), 
+                               0.5))
 
-patients_hospitalized_per_week_agg_sex_perc <- total_patients_psy_per_week %>%
-  left_join( patients_hospitalized_per_week_agg_sex, by = c("weeks")) %>% 
+patients_hospitalized_agg_sex_perc <- total_patients_psy %>%
+  left_join( patients_hospitalized_agg_sex, by = c("time_p")) %>% 
   replace_na(list(count_sex = 0))
 
-patients_hospitalized_per_week_agg_sex_perc$percentage <- 100*patients_hospitalized_per_week_agg_sex_perc$count_sex/patients_hospitalized_per_week_agg_sex_perc$count
+patients_hospitalized_agg_sex_perc$percentage <- 100*patients_hospitalized_agg_sex_perc$count_sex/patients_hospitalized_agg_sex_perc$count
 
-patients_hospitalized_per_week_agg_sex_perc %>%
-  ggplot(aes(x = weeks, y = percentage, fill = sex)) +
+patients_hospitalized_agg_sex_perc %>%
+  ggplot(aes(x = time_p, y = percentage, fill = sex)) +
   geom_bar(stat="identity", position = "dodge")+
@@
   labs(y = "Percentage ",
-       x = "Week",
-       title = "Weekly patient (%) aggregated by sex",
+       x = time_period,
+       title = paste0("Patient (%) per ", time_period," aggregated by sex"),
        subtitle = "Per Sex") + 

### 2021-10-01 psy_peds_3.Rmd
ratio estimation and visualization added

@@
 data_update_date <- "2021-06-05"
-time_period <-"weeks" #possible values: days, weeks, months
+time_period <-"months" #possible values: days, weeks, months
 ```
@@
 
+##### Percentage 
 ```{r}
@@
 thpd_psyPer
+```
+##### Ratio: patients with psychiatric condition / patients without psychiatric condition
+
+### Total patients hospitalized comparing psychiatric disorders vs other
+
+```{r}
+all_patients <- unique( all_codes_desc[, c("patient_num", "time_p")])
+patients_with_psy_conditions <- unique( all_codes_psy[, c("patient_num", "time_p")] )
+patients_without_psy_conditions <- setdiff(all_patients,patients_with_psy_conditions)
+
+patients_without_psy_conditions_counts <- patients_without_psy_conditions %>%
+  group_by( time_p ) %>%
+  summarise(count_no_psy_patients =  ifelse(n_distinct(patient_num) > obfuscation | isFALSE(obfuscation),
+                               n_distinct(patient_num), 
+                               0.5)) %>%
+  mutate( period = ifelse( time_p < pandemic_start_date,
+                           "before_pandemic",
+                           "during_pandemic"))
+
+patients_with_psy_conditions_counts <-  patients_with_psy_conditions %>%
+  group_by( time_p ) %>%
+  summarise(count_psy_patients =  ifelse(n_distinct(patient_num) > obfuscation | isFALSE(obfuscation),
+                               n_distinct(patient_num), 
+                               0.5)) %>%
+  mutate( period = ifelse( time_p < pandemic_start_date,
+                           "before_pandemic",
+                           "during_pandemic"))
+
+ratios_patients_with_without_psy <- patients_without_psy_conditions_counts %>%
+  full_join( patients_with_psy_conditions_counts, by=c("time_p", "period")) %>%
+  mutate( ratio = count_psy_patients/count_no_psy_patients)
+
+ratio_psy_non_psy <- ratios_patients_with_without_psy %>%
+  ggplot(aes(x = time_p, y = ratio, fill = period, color = period)) +
+  geom_point() +
+  geom_vline(xintercept = as.Date(pandemic_start_date),
+             linetype = "dashed") +
+  scale_fill_manual(values = cbPalette) +
+  scale_color_manual(values = cbPalette) +
+  labs(y = "Ratio (with psy conditions/without psy conditions)",
+       x = paste0("Date (by ", time_period,")"),
+       title = paste0("Ratio patient with vs without psychiatric conditions ( per ", time_period,")"))
+
+ratio_psy_non_psy
+
 ```

### 2021-10-01 archive_versions/psy_peds_2.Rmd
archive folder for the previous versions of the Rmd

--- 2021-09-28 psy_peds_2.Rmd
@@
 ### Read the CSV input files
-source("R/readInputFiles.R")
+source("./R/readInputFiles.R")
 

### 2021-10-04 psy_peds_3.Rmd
adding time series analysis

@@
 ```{r message=FALSE, warning=FALSE}
-folder_4ce_files <- "../Input/09212021_2/"
+folder_4ce_files <- "Input/09212021_2/"
 obfuscation =  FALSE
@@
 
-clinical_filtered$time_p <- NA
 if( time_period == "weeks"){
@@
 }
+
 ```
@@
 
+## Interrupted Time-Series Analysis
+
+```{r}
+# Model without seasonality
+time_series_analysis_plotting <- function(model, df_count_psy, count_col, clearance_period = FALSE) {
+  output_model <- broom::tidy(model)
+  coeff_slope <- round(output_model[output_model$term == "periodduring_pandemic:time", "estimate"][[1]], 3)
+  pvalue_slope <- round(output_model[output_model$term == "periodduring_pandemic:time", "p.value"][[1]], 3)
+  annotation <- paste0("Lockdown slope: ", coeff_slope, " (pvalue: ", pvalue_slope, ")")
+  pred_model <- predict(model, patient_count_monthly_psy, type = "response", interval = "confidence") %>% as.data.frame()
+  pred_model_b <- predict(model, mutate(patient_count_monthly_psy, period = "before_pandemic"), type = "response", interval = "confidence") %>%
+    as.data.frame()
+  names(pred_model_b) <- paste0(names(pred_model_b), "_wo_intervention")
+  pred_model <- bind_cols(df_count_psy, pred_model, pred_model_b)
+  if (clearance_period) {
+    pred_model$clearance_period <- ifelse(pred_model$time_p < as.Date(cut(start_clear_period, breaks = time_period)) ,
+                                          "before_pandemic",
+                                          ifelse(pred_model$time_p > as.Date(cut(end_clear_period, breaks = time_period)),
+                                                 "during_pandemic",
+                                                 "during_clearance")
+    )
+    pred_model$lwr <- ifelse(pred_model$clearance_period == "during_clearance", NA, pred_model$lwr)
+    pred_model$upr <- ifelse(pred_model$clearance_period == "during_clearance", NA, pred_model$upr)
+    pred_model$fit_wo_intervention <- ifelse(pred_model$time_p >= max(pred_model$time_p[pred_model$clearance_period == "before pandemic"]),
+                                             pred_model$fit_wo_intervention,
+                                             NA)
+  } else {
+    pred_model$clearance_period <- pred_model$period
+  }
+  print(pred_model$clearance_period)
+  plot <- ggplot(pred_model, aes_string(x = "time_p", y = count_col)) +
+    geom_point(aes(colour = clearance_period)) +
+    geom_line(aes(x = time_p, y = fit, colour = clearance_period)) +
+    geom_line(aes(x = time_p, y = fit_wo_intervention, colour = "Expected mean w/o pandemic"), linetype = "dashed") +
+    geom_ribbon(aes(ymin = lwr, ymax = upr, colour = clearance_period), alpha = 0.3, colour = NA) +
+    scale_color_manual(name = "Time Series projection", values = c("before_pandemic" = "black",
+                                                                   "Expected mean w/o pandemic" = "red",
+                                                                   "during_pandemic" = "black",
+                                                                   "during_clearance" = "white")) +
+    annotate("text", x = as.Date("2019-10-01"), y = max(pred_model[[count_col]]) + 5, label = annotation) +
+    theme(legend.position = "bottom")
+
+  if (clearance_period) {
+    plot +
+      geom_vline(xintercept = as.Date(max(pred_model$time_p[pred_model$clearance_period == "before_pandemic"])), linetype = "dashed") +
+      geom_vline(xintercept = as.Date(min(pred_model$time_p[pred_model$clearance_period == "during_pandemic"])), linetype = "dashed")
+  } else {
+    plot + geom_vline(xintercept = as.Date(pandemic_start_date), linetype = "dashed")
+  }
+}
+```
+
+```{r}
+# Time series analysis of percentage of psy codes per encounter, linear (without seasonality)
+patient_count_monthly_psy <- count_icd_psy %>%
+  filter(time_p > as.Date(cut( start_date_plots, breaks = time_period)) &
+           time_p < as.Date(cut( end_date_plots, breaks = time_period))) %>%
+  mutate(time = seq_along(time_p),
+         month_year = format(time_p, "%m")
+  )
+start_clear_period <- as.Date("2020-03-15")
+end_clear_period <- as.Date("2020-05-15")
+patient_count_monthly_psy_clear <- patient_count_monthly_psy %>%
+  mutate( count_psy = ifelse(time_p < as.Date(cut( start_clear_period, breaks = time_period)) |
+                               time_p > as.Date(cut( end_clear_period, breaks = time_period)),
+                             count_psy,
+                             NA)
+  )
+```
+
+```{r}
+ts_linear_count <- lm(count_psy ~ period * time, patient_count_monthly_psy)
+ts_linear_count_clear <- lm(count_psy ~ period * time, patient_count_monthly_psy_clear)
+ts_seasonal_count <- lm(count_psy ~ period*time + harmonic(month_year, 1, 12), patient_count_monthly_psy)
+ts_seasonal_count_clear <- lm(count_psy ~ period*time + harmonic(month_year, 1, 12), patient_count_monthly_psy_clear)
+```
+
+```{r}
+time_series_analysis_plotting(ts_linear_count,
+                              patient_count_monthly_psy, "count_psy",
+                              clearance_period = FALSE) +
+  labs(title = "Interrupted Time-Series Analysis: Percentage of encounters with psychiatric ICD codes",
+       subtitle = "Without seasonality effect",
+       x = time_period,
+       y = "Count")
+```
+
+```{r}
+time_series_analysis_plotting(ts_linear_count_clear,
+                              patient_count_monthly_psy, "count_psy",
+                              clearance_period = TRUE) +
+  labs(title = "Interrupted Time-Series Analysis: Percentage of encounters with psychiatric ICD codes",
+       subtitle = "Without seasonality effect",
+       x = time_period,
+       y = "Count")
+```
+
+```{r}
+time_series_analysis_plotting(ts_seasonal_count,
+                              patient_count_monthly_psy, "count_psy",
+                              clearance_period = FALSE) +
+  labs(title = "Interrupted Time-Series Analysis: Percentage of encounters with psychiatric ICD codes",
+       subtitle = "With seasonality effect",
+       x = time_period,
+       y = "Count")
+```
+
+```{r}
+time_series_analysis_plotting(ts_seasonal_count_clear,
+                              patient_count_monthly_psy, "count_psy",
+                              clearance_period = TRUE) +
+  labs(title = "Interrupted Time-Series Analysis: Count of encounters with psychiatric ICD codes",
+       subtitle = "With seasonality effect",
+       x = time_period,
+       y = "Count")
+```
+
 #### Psychiatric conditions aggregating by suptype

### 2021-10-06 psy_peds_3.Rmd
table 1 + updating plots

@@
 # cut-off dates used in the analysis, these should not be changed
-history_start_date <- "2019-01-01"
-history_end_date <- "2021-05-31"
-pandemic_start_date <- "2020-03-15"
-start_date_plots <- "2019-01-15"
-end_date_plots <- "2021-05-01"
+history_start_date <- as.Date("2019-01-01")
+history_end_date <- as.Date("2021-05-31")
+pandemic_start_date <- as.Date("2020-03-15")
+start_date_plots <- as.Date("2019-01-15")
+end_date_plots <- as.Date("2021-05-01")
 ```
@@
     geom_point(aes(colour = clearance_period)) +
+    geom_line(aes(x = time_p, y = fit_wo_intervention, colour = "Expected mean w/o pandemic"), linetype = "dashed") +
     geom_line(aes(x = time_p, y = fit, colour = clearance_period)) +
-    geom_line(aes(x = time_p, y = fit_wo_intervention, colour = "Expected mean w/o pandemic"), linetype = "dashed") +
     geom_ribbon(aes(ymin = lwr, ymax = upr, colour = clearance_period), alpha = 0.3, colour = NA) +
@@
   )
+
 start_clear_period <- as.Date("2020-03-15")
@@
 ```
+
+## Table 1
+
+```{r}
+
+t.test2 <- function(m1, m2, s1, s2, n1, n2) {
+  # t.test computed from sample statistics, ie using mean, standard deviation and count value of the samples
+  # useful in case we can only retrieve aggregated stats from the sites
+  # ref: https://stats.stackexchange.com/questions/30394/how-to-perform-two-sample-t-tests-in-r-by-inputting-sample-statistics-rather-tha
+  # and https://en.wikipedia.org/wiki/Student%27s_t-test?oldformat=true#Equal_or_unequal_sample_sizes,_unequal_variances_(sX1_%3E_2sX2_or_sX2_%3E_2sX1)
+  se <- sqrt( (s1^2/n1) + (s2^2/n2) )
+  # welch-satterthwaite df
+  df <- ( (s1^2/n1 + s2^2/n2)^2 )/( (s1^2/n1)^2/(n1-1) + (s2^2/n2)^2/(n2-1) )
+  t <- (m1-m2)/se
+  dat <- c(m1-m2, se, t, 2*pt(-abs(t),df))
+  names(dat) <- c("Difference of means", "Std Error", "t", "p-value")
+  return(dat)
+}
+
+age_by_period <- all_codes_psy %>%
+  distinct(patient_num, age_time_diagnosis, period) %>% 
+  dplyr::group_by(period) %>%
+  dplyr::summarise( mean = mean( age_time_diagnosis ),
+                    sd = sd( age_time_diagnosis ),
+                    count = n_distinct(patient_num))
+
+age_test <- t.test2(m1 = age_by_period[[1, "mean"]],
+                    m2 = age_by_period[[2, "mean"]],
+                    s1 = age_by_period[[1, "sd"]],
+                    s2 = age_by_period[[2, "sd"]],
+                    n1 = age_by_period[[1, "count"]],
+                    n2 = age_by_period[[2, "count"]]) %>%
+  `[[`("p-value")
+```
+
+## Sex
+```{r}
+sex <- all_codes_psy %>%
+  dplyr::distinct(patient_num, sex) %>% 
+  dplyr::group_by( sex ) %>%
+  dplyr::summarise( count = n_distinct(patient_num) ) %>%
+  dplyr::mutate( count = ifelse( count > obfuscation, count, obfuscation * 0.5),
+                 total = sum( count ), percentage = round( count/total*100, 2)  )
+
+sex_by_period <- all_codes_psy %>%
+  dplyr::distinct(patient_num, period, sex) %>% 
+  group_by(period, sex ) %>%
+  summarise(count = n_distinct(patient_num)) %>%
+  group_by(period, .add=TRUE) %>%
+  mutate( count = ifelse( count > obfuscation, count, obfuscation * 0.5),
+          percentage=round(count/sum( count )*100, 2)) %>%
+  replace_na(list(male = 0, female = 0))
+
+sex_test <- sex_by_period %>%
+  select(-percentage) %>%
+  pivot_wider(names_from = sex, values_from = count) %>%
+  ungroup() %>%
+  select(male, female) %>%
+  chisq.test() %>%
+  `[[`("p.value")
+```
+
+## Diagnosis
+
+### By ICD code
+```{r}
+summary_diagnosis <- all_codes_psy %>%
+  dplyr::distinct( patient_num, description, period, weeks, concept_code ) %>%
+  dplyr::group_by( concept_code, description, period ) %>%
+  dplyr::count( patient_num ) %>%
+  dplyr::tally(name = "count") %>%
+  dplyr::mutate(count = ifelse( count > obfuscation, count, obfuscation * 0.5)) %>%
+  dplyr::arrange( desc(count) ) %>%
+  dplyr::ungroup()
+
+summary_diagnosis_test <- summary_diagnosis %>%
+  tidyr::pivot_wider(names_from = period,
+                     values_from = count) %>%
+  #select(late_pandemic, early_pandemic) %>%
+  select(during_pandemic) %>%
+  #tidyr::replace_na(list(late_pandemic = 0, early_pandemic = 0)) %>%
+  tidyr::replace_na(list(during_pandemic = 0)) %>%
+  chisq.test() %>%
+  `[[`("p.value")
+
+```
+
+### By disorder group
+```{r}
+summary_disorder_group <- all_codes_psy %>%
+  dplyr::group_by( disorder_group ) %>%
+  dplyr::count( patient_num ) %>%
+  dplyr::tally(name = "count") %>%
+  dplyr::mutate(count = ifelse( count > obfuscation, count, obfuscation * 0.5)) %>%
+  dplyr::arrange( desc( count ))
+
+summary_disorder_group_by_period <- all_codes_psy %>%
+  dplyr::distinct( patient_num, disorder_group, period, weeks ) %>%
+  dplyr::group_by( period, disorder_group ) %>%
+  summarise(count = n_distinct(patient_num)) %>%
+  group_by(period, add=TRUE) %>%
+  mutate(count = ifelse( count > obfuscation, count, obfuscation * 0.5),
+         percentage = count/sum( count )*100)
+
+summary_disorder_group_by_period_test <- summary_disorder_group_by_period %>%
+  pivot_wider(id_cols = disorder_group, names_from = period, values_from = c(count, percentage)) %>%
+  mutate(across(where(is.numeric), ~replace_na(.x, replace = 0))) %>%
+  #select(count_early_pandemic, count_late_pandemic) %>%
+  select(count_during_pandemic) %>%
+  chisq.test() %>%
+  `[[`("p.value")
+
+```
+
+```{r}
+format_counts <- function(count, percentage) {
+  glue::glue("{count} ({percentage} %)", count = count, percentage = percentage)
+}
+
+age_table1 <- age_by_period %>% 
+  mutate(result = glue::glue("{mean} (sd: {sd})", mean = round(mean, 2), sd = round(sd, 2))) %>% 
+  select(result, period) %>% 
+  pivot_wider( names_from = period, values_from = result ) %>%
+  mutate(name = "Age")
+
+sex_table1 <- sex_by_period %>%
+  bind_rows(
+    sex_by_period %>%
+      group_by(period) %>%
+      dplyr::summarise(sex = "Total", count = sum(count), percentage = sum(percentage))
+  ) %>%
+  mutate(result = format_counts(count, percentage)) %>%
+  select(-c(count, percentage)) %>%
+  pivot_wider( names_from = period, values_from = result ) %>%
+  rename(name = sex)
+
+disorder_table1 <- summary_disorder_group_by_period %>%
+  arrange(desc(count)) %>% 
+  bind_rows(
+    dplyr::summarise(., disorder_group="Total", across(where(is.numeric), sum))
+  ) %>%
+  mutate(percentage = round(percentage, 2)) %>% 
+  mutate(result = format_counts(count, percentage)) %>%
+  select(-c(count, percentage)) %>%
+  pivot_wider( names_from = period, values_from = result ) %>%
+  replace_na( list(during_pandemic = "0 (0 %)")) %>%
+  rename(name = disorder_group)
+```
+
+```{r}
+list_tests <- c("Sex" = sex_test,
+                "Age" = age_test,
+                "Diagnoses" = summary_diagnosis_test,
+                "Disorder groups" = summary_disorder_group_by_period_test
+)
+
+list_table1 <- list(
+  "Sex" = sex_table1,
+  "Age" = age_table1,
+  "Disorder groups" = disorder_table1)
+
+tests <- list_tests %>%
+  data.frame() %>%
+  round(2)
+names(tests) <- "pvalue"
+tests["Category"] <- row.names(tests)
+
+table1 <- list_table1 %>%
+  lapply(function(df) mutate(df, across(.fns = as.character))) %>%
+  dplyr::bind_rows(.id = "Category") %>%
+  dplyr::left_join(tests, by = "Category")
+
+table1 %>%
+  group_by(Category) %>%
+  mutate(pvalue = ifelse(duplicated(pvalue, fromLast = TRUE), "", as.character(pvalue))) %>%
+  ungroup() %>%
+  mutate(Category = "") %>%
+  mutate(across(.cols = c(before_pandemic, during_pandemic), .fns = replace_na, "0")) %>%
+  rename_with(str_to_title) %>%
+  rename_with(sub, pattern = "_", replacement = " ") %>%
+  mutate(across(.fns = str_to_title)) %>%
+  mutate(across(.fns = sub, pattern = "_", replacement = " ")) %>%
+  kable( booktabs = T) %>%
+  kableExtra::kable_styling(latex_options = "striped") %>%
+  kableExtra::group_rows(index = setNames(rle(table1$Category)[[1]], rle(table1$Category)[[2]]))
+```

### 2021-10-06 psy_peds_3.Rmd
updating document layout

@@
   html_document:
-    code_folding: show
+    code_folding: hide
     toc: yes
@@
 ---
+
+# Setting up the R environment
+
+### Installing packages and loading the library
 
@@
 
-# R session information
+### R session information
 
@@
 ```
+
+### Plots aesthetics information
 
@@
 
-# Inclusion criteria
-
+# Data retrieval
+
+## Inclusion criteria
 The inclusion criteria for this study is:
@@
 
-# Variables that need to be checked/modified by each site
+## Variables that need to be checked/modified by each site
 
@@
     documentation](https://www.stat.berkeley.edu/~s133/dates.html))
+
+5. "data_update_date": date at which the data has been updated in the local 
+datawarehouse. Used to estimate patient age at time of visit, since patients age
+in the 4CE demographic file is expected the age at data update.
+
+6. "time_period": Time unit for the different graphs of the plots: months (default), weeks, days
 
@@
 
-# Data input
+## Data loading
 
@@
 
-## Read csv files
-
-### 4CE phase 2.2 files
+### Reading 4CE phase 2.2 files
 
@@
 
-## Applying inclusion criteria to filter and format the data
-
-We are selecting hospitalizations based on the following criteria:
-
-- Associated with at least 1 ICD code
-
-- With a duration of hospitalization spanning over at least two consecutive days
-
-- Starting on or after 2019/01/01
-
 ### Calculating age at hospitalization time
@@
 ```
+
+### Applying inclusion criteria to filter and format the data
+
+We are selecting hospitalizations based on the following criteria:
+
+- Associated with at least 1 ICD code
+
+- With a duration of hospitalization spanning over at least two consecutive days
+
+- Starting on or after 2019/01/01
 
@@
 
-## First estimation: total number of patients without filtering by any disease subtype
+### Filtering on age 11-17
+
+```{r}
+# Filtering hospitalizations where patient age is in inclusion criteria
+observation_table_ado <- filter(observation_table, age_time_diagnosis >= 11 & age_time_diagnosis <= 17)
+
+# Filtering observation tables which patient_num and days_since_admission are still in clinical table after filtering on age
+
+clinical_test <- clinical_table[, c("patient_num", "n_hospitalisation", "days_since_admission")] %>% 
+  inner_join(distinct(observation_table_ado, patient_num, days_since_admission, age_time_diagnosis), 
+            by = c("patient_num", "days_since_admission")) %>% 
+  distinct(patient_num, n_hospitalisation, age_time_diagnosis)
+
+clinical_table_ado <- clinical_table %>% 
+  inner_join(clinical_test,
+             by = c("patient_num", "n_hospitalisation"))
+
+# Filtering observation tables which patient_num are still in clinical table after filtering on age
+demographic_table_ado <- demographic_table %>% 
+  inner_join(distinct(clinical_table_ado, patient_num), by = "patient_num")
+
+all_codes_ado <- all_codes %>% 
+   filter(age_time_diagnosis >= 11 & age_time_diagnosis <= 17)
+```
+
+```{r}
+stopifnot(unique(all_codes_ado$cohort) == "AllAdm")
+stopifnot(unique(all_codes_ado$in_hospital) == 1)
+stopifnot(unique(all_codes_ado$len_hospitalisation > min_hosp_days))
+stopifnot(nrow(all_codes_ado) == nrow(observation_table_ado))
+stopifnot(length(unique(all_codes_ado$patient_num)) == length(unique(demographic_table_ado$patient_num)))
+stopifnot(length(unique(all_codes_ado$patient_num)) == length(unique(observation_table_ado$patient_num)))
+stopifnot(length(unique(all_codes_ado$patient_num)) == length(unique(clinical_table_ado$patient_num)))
+stopifnot(nrow(distinct(all_codes_ado, patient_num, n_hospitalisation)) == nrow(distinct(clinical_table_ado, patient_num, n_hospitalisation)))
+stopifnot(nrow(distinct(all_codes_ado, patient_num, days_since_admission)) == nrow(distinct(observation_table_ado, patient_num, days_since_admission)))
+```
+
+### Add the psychiatric/mental health ICD code description and filter
+
+```{r message=FALSE, warning=FALSE}
+# Getting site name, will be used later in the analysis
+site <- unique(as.character(demo_raw$siteid))
+
+icdCategory <- read.delim("public-data/icd10Codes.txt",
+                          header = FALSE,
+                          colClasses = "character")
+colnames(icdCategory) <- c("icd_code_letter", "icd_description_category")
+
+all_codes_desc <- all_codes_ado %>% 
+    left_join( icdCodes, by=c("concept_code" = "ICD10_Code") ) %>%
+  mutate(icd_code_category = ifelse(concept_code %in% icdCodes$ICD10_Code, 
+                               "psy", 
+                               "others")) %>%
+  dplyr::select( patient_num, sex, age_time_diagnosis, concept_type, concept_code, calendar_date,
+                 in_hospital, n_hospitalisation, len_hospitalisation, date, weeks, month, year,
+                 period, time_p, disorder_group, description, icd_code_category )
+
+all_codes_desc <- all_codes_desc %>% 
+  mutate( icd_code_letter = ifelse(substr(concept_code, 1, 1) %in% c("D", "H"), substr(concept_code, 1, 2), substr(concept_code, 1, 1) ))
+
+all_codes_desc <- left_join( all_codes_desc, icdCategory) %>%
+  mutate( icd_description_category = replace_na( ifelse( concept_code %in% icdCodes$ICD10_Code, "Psy", icd_description_category), "Others")) 
+all_codes_psy <- all_codes_desc %>% 
+  filter(concept_code %in% icdCodes$ICD10_Code)
+
+if( raceAvailable == TRUE ){
+  all_codes_psy <- all_codes_psy %>%
+    left_join( race_raw, by = c("patient_num") )
+}
+```
+
+# Results
+
+## Admission
+
+First calculation: total number of patients, matching the inclusion criteria, admitted on one day (without filtering by any disease subtype).
+Summed up over one week, or one month depending on the time unit chosen.
+Not filtered on qge.
 
@@
 
-### New hospitalizations
-
 ```{r}
@@
 
-### Filtering on age at hospitalization time
-
-Keeping only patients with age between 11 and 17
-
-```{r}
-# Filtering hospitalizations where patient age is in inclusion criteria
-observation_table_ado <- filter(observation_table, age_time_diagnosis >= 11 & age_time_diagnosis <= 17)
-
-# Filtering observation tables which patient_num and days_since_admission are still in clinical table after filtering on age
-
-clinical_test <- clinical_table[, c("patient_num", "n_hospitalisation", "days_since_admission")] %>% 
-  inner_join(distinct(observation_table_ado, patient_num, days_since_admission, age_time_diagnosis), 
-            by = c("patient_num", "days_since_admission")) %>% 
-  distinct(patient_num, n_hospitalisation, age_time_diagnosis)
-
-clinical_table_ado <- clinical_table %>% 
-  inner_join(clinical_test,
-             by = c("patient_num", "n_hospitalisation"))
-
-# Filtering observation tables which patient_num are still in clinical table after filtering on age
-demographic_table_ado <- demographic_table %>% 
-  inner_join(distinct(clinical_table_ado, patient_num), by = "patient_num")
-
-all_codes_ado <- all_codes %>% 
-   filter(age_time_diagnosis >= 11 & age_time_diagnosis <= 17)
-```
-
-```{r}
-stopifnot(unique(all_codes_ado$cohort) == "AllAdm")
-stopifnot(unique(all_codes_ado$in_hospital) == 1)
-stopifnot(unique(all_codes_ado$len_hospitalisation > min_hosp_days))
-stopifnot(nrow(all_codes_ado) == nrow(observation_table_ado))
-stopifnot(length(unique(all_codes_ado$patient_num)) == length(unique(demographic_table_ado$patient_num)))
-stopifnot(length(unique(all_codes_ado$patient_num)) == length(unique(observation_table_ado$patient_num)))
-stopifnot(length(unique(all_codes_ado$patient_num)) == length(unique(clinical_table_ado$patient_num)))
-stopifnot(nrow(distinct(all_codes_ado, patient_num, n_hospitalisation)) == nrow(distinct(clinical_table_ado, patient_num, n_hospitalisation)))
-stopifnot(nrow(distinct(all_codes_ado, patient_num, days_since_admission)) == nrow(distinct(observation_table_ado, patient_num, days_since_admission)))
-```
-
-### New admissions age range 11-17
+## New admissions age range 11-17
 
@@
 
-### Total hospitalize patients age range 11-17
+## Total hospitalize patients age range 11-17
+
+Calculating the total number of patients hospitalized on a given day (taking into account the hospitalization length).
 
@@
 
-## Total number of patients filtering by psychiatric disorders
-
-## Creating the inputTable
-
-Creation of the inputTable that will be the only table used throughout the rest of the analysis - selecting only visits with at least one ICD
-codes pertaining to the mental health categories - filter visits where patients are between 11 and 17 years old - add the demographic and
-hospitalization related information - add the ICD code description and the related disorder group
-
-### Add the psychiatric/mental health ICD code description and filter
-
-```{r message=FALSE, warning=FALSE}
-# Getting site name, will be used later in the analysis
-site <- unique(as.character(demo_raw$siteid))
-
-icdCategory <- read.delim("public-data/icd10Codes.txt",
-                          header = FALSE,
-                          colClasses = "character")
-colnames(icdCategory) <- c("icd_code_letter", "icd_description_category")
-
-all_codes_desc <- all_codes_ado %>% 
-    left_join( icdCodes, by=c("concept_code" = "ICD10_Code") ) %>%
-  mutate(icd_code_category = ifelse(concept_code %in% icdCodes$ICD10_Code, 
-                               "psy", 
-                               "others")) %>%
-  dplyr::select( patient_num, sex, age_time_diagnosis, concept_type, concept_code, calendar_date,
-                 in_hospital, n_hospitalisation, len_hospitalisation, date, weeks, month, year,
-                 period, time_p, disorder_group, description, icd_code_category )
-
-all_codes_desc <- all_codes_desc %>% 
-  mutate( icd_code_letter = ifelse(substr(concept_code, 1, 1) %in% c("D", "H"), substr(concept_code, 1, 2), substr(concept_code, 1, 1) ))
-
-all_codes_desc <- left_join( all_codes_desc, icdCategory) %>%
-  mutate( icd_description_category = replace_na( ifelse( concept_code %in% icdCodes$ICD10_Code, "Psy", icd_description_category), "Others")) 
-all_codes_psy <- all_codes_desc %>% 
-  filter(concept_code %in% icdCodes$ICD10_Code)
-
-if( raceAvailable == TRUE ){
-  all_codes_psy <- all_codes_psy %>%
-    left_join( race_raw, by = c("patient_num") )
-}
-```
-
-### Total patients hospitalized comparing psychiatric disorders vs other
+## Patient count filtered on psychiatric disorders
+
+### New admissions
 
@@
 
-##### Percentage 
+### Percentage 
 ```{r}
@@
 ```
-##### Ratio: patients with psychiatric condition / patients without psychiatric condition
-
-### Total patients hospitalized comparing psychiatric disorders vs other
+### Ratio: patients with psychiatric condition / patients without psychiatric condition
 
@@
 
-## Interrupted Time-Series Analysis
+## Interrupted Time-Series Analyses
 
@@
   }
-  print(pred_model$clearance_period)
   plot <- ggplot(pred_model, aes_string(x = "time_p", y = count_col)) +
@@
 
+### ITS: no clearance period, no seasonality effect
+
 ```{r}
@@
                               clearance_period = FALSE) +
-  labs(title = "Interrupted Time-Series Analysis: Percentage of encounters with psychiatric ICD codes",
+  labs(title = "Interrupted Time-Series Analysis: Count of encounters with psychiatric ICD codes",
        subtitle = "Without seasonality effect",
@@
 ```
+
+### ITS: clearance period, no seasonality effect
 
@@
 
+### ITS: no clearance period, with seasonality effect
+
 ```{r}
@@
 
+### ITS: with clearance period, with seasonality effect
+
 ```{r}
@@
 
-#### Psychiatric conditions aggregating by suptype
-
-##### Total hospitalizations
+## Psychiatric conditions aggregating by subcategories
+
+### New admissions: count
 
@@
 
-Percentage using as denominator total number of patients
+### New admissions: percentage
 
@@
 
-### Aggregated by sex
+## Psychiatric conditions, dichotomized on patients' sex
+
+### Count
 
@@
   geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=sex), colour= "black", size = 0.5)
-
+```
+
+### Percentage
+
+```{r}
 ##percentage
@@
 
-## Sex
 ```{r}
@@
 
-## Diagnosis
-
-### By ICD code
 ```{r}
@@
 
-### By disorder group
 ```{r}
@@
   summarise(count = n_distinct(patient_num)) %>%
-  group_by(period, add=TRUE) %>%
+  group_by(period, .add=TRUE) %>%
   mutate(count = ifelse( count > obfuscation, count, obfuscation * 0.5),

### 2021-10-06 psy_peds_3.Rmd
adding percentage count, and its with percentage

@@
 
-### Percentage 
-```{r}
-thpd_psyPer <- count_icd %>%
-  ggplot(aes(x = time_p, y = percentage_psy, fill=period)) +
-  scale_fill_manual(values = cbPalette[c(3,4)]) +
-  geom_vline(xintercept = as.Date(pandemic_start_date),
-             linetype = "dashed") +
-  labs(y = "Percentage",
-       x = paste0("Date (by ", time_period,")"),
-       title = paste0("Percentage of patients with mental health related ICD codes (per ", time_period,")")) + 
-  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)+
-  geom_point()
-thpd_psyPer
-```
 ### Ratio: patients with psychiatric condition / patients without psychiatric condition
@@
   full_join( patients_with_psy_conditions_counts, by=c("time_p", "period")) %>%
-  mutate( ratio = count_psy_patients/count_no_psy_patients)
-
+  mutate( ratio = count_psy_patients/count_no_psy_patients, 
+          percentage = 100 * count_psy_patients/(count_no_psy_patients + count_psy_patients))
+
+```
+
+```{r}
+ratios_patients_with_without_psy %>%
+  pivot_longer(cols = c("count_no_psy_patients", "count_psy_patients"), 
+               values_to = "count", 
+               names_to = "count_type") %>% 
+  ggplot(aes(x = time_p, y = count, color = count_type)) +
+  geom_point() +
+  geom_vline(xintercept = as.Date(pandemic_start_date),
+             linetype = "dashed") +
+  scale_fill_manual(values = cbPalette) +
+  scale_color_manual(values = cbPalette) +
+  labs(y = "Count",
+       x = paste0("Date (by ", time_period,")"),
+       title = paste0("Count comparison: patients with vs without psychiatric conditions (per ", time_period,")")) +
+  guides(color = guide_legend(title="Patient diagnoses types"))
+```
+
+```{r}
 ratio_psy_non_psy <- ratios_patients_with_without_psy %>%
@@
 
+```{r}
+ratios_patients_with_without_psy %>%
+  ggplot(aes(x = time_p, y = percentage, color = period)) +
+  geom_point() +
+  geom_vline(xintercept = as.Date(pandemic_start_date),
+             linetype = "dashed") +
+  scale_fill_manual(values = cbPalette) +
+  scale_color_manual(values = cbPalette) +
+  labs(y = "Percentage",
+       x = paste0("Date (by ", time_period,")"),
+       title = paste0("Percentage patient with vs total population ( per ", time_period,")")) +
+  guides(color = guide_legend(title="Patient diagnoses types"))
+```
+
 ## Interrupted Time-Series Analyses
@@
   annotation <- paste0("Lockdown slope: ", coeff_slope, " (pvalue: ", pvalue_slope, ")")
-  pred_model <- predict(model, patient_count_monthly_psy, type = "response", interval = "confidence") %>% as.data.frame()
-  pred_model_b <- predict(model, mutate(patient_count_monthly_psy, period = "before_pandemic"), type = "response", interval = "confidence") %>%
+  pred_model <- predict(model, df_count_psy, type = "response", interval = "confidence") %>% as.data.frame()
+  pred_model_b <- predict(model, mutate(df_count_psy, period = "before_pandemic"), type = "response", interval = "confidence") %>%
     as.data.frame()
@@
 # Time series analysis of percentage of psy codes per encounter, linear (without seasonality)
-patient_count_monthly_psy <- count_icd_psy %>%
+patient_count_psy_period_psy <- ratios_patients_with_without_psy %>%
   filter(time_p > as.Date(cut( start_date_plots, breaks = time_period)) &
@@
 end_clear_period <- as.Date("2020-05-15")
-patient_count_monthly_psy_clear <- patient_count_monthly_psy %>%
-  mutate( count_psy = ifelse(time_p < as.Date(cut( start_clear_period, breaks = time_period)) |
+patient_count_psy_period_psy_clear <- patient_count_psy_period_psy %>%
+  mutate( count_no_psy_patients = ifelse(time_p < as.Date(cut( start_clear_period, breaks = time_period)) |
                                time_p > as.Date(cut( end_clear_period, breaks = time_period)),
-                             count_psy,
+                             count_no_psy_patients,
+                             NA), 
+          count_no_psy_patients = ifelse(time_p < as.Date(cut( start_clear_period, breaks = time_period)) |
+                               time_p > as.Date(cut( end_clear_period, breaks = time_period)),
+                             count_no_psy_patients,
+                             NA), 
+          ratio = ifelse(time_p < as.Date(cut( start_clear_period, breaks = time_period)) |
+                           time_p > as.Date(cut( end_clear_period, breaks = time_period)),
+                         ratio,
+                         NA),
+          percentage = ifelse(time_p < as.Date(cut( start_clear_period, breaks = time_period)) |
+                           time_p > as.Date(cut( end_clear_period, breaks = time_period)),
+                         percentage,
                          NA)
@@
 
-```{r}
-ts_linear_count <- lm(count_psy ~ period * time, patient_count_monthly_psy)
-ts_linear_count_clear <- lm(count_psy ~ period * time, patient_count_monthly_psy_clear)
-ts_seasonal_count <- lm(count_psy ~ period*time + harmonic(month_year, 1, 12), patient_count_monthly_psy)
-ts_seasonal_count_clear <- lm(count_psy ~ period*time + harmonic(month_year, 1, 12), patient_count_monthly_psy_clear)
-```
-
-### ITS: no clearance period, no seasonality effect
+```{r, its count based}
+ts_linear_count <- lm(count_psy_patients ~ period * time, patient_count_psy_period_psy)
+ts_linear_count_clear <- lm(count_psy_patients ~ period * time, patient_count_psy_period_psy_clear)
+ts_seasonal_count <- lm(count_psy_patients ~ period*time + harmonic(month_year, 1, 12), patient_count_psy_period_psy)
+ts_seasonal_count_clear <- lm(count_psy_patients ~ period*time + harmonic(month_year, 1, 12), patient_count_psy_period_psy_clear)
+```
+
+```{r, its percentage based}
+ts_linear_percent <- lm(percentage ~ period * time, patient_count_psy_period_psy)
+ts_linear_percent_clear <- lm(percentage ~ period * time, patient_count_psy_period_psy_clear)
+ts_seasonal_percent <- lm(percentage ~ period*time + harmonic(month_year, 1, 12), patient_count_psy_period_psy)
+ts_seasonal_percent_clear <- lm(percentage ~ period*time + harmonic(month_year, 1, 12), patient_count_psy_period_psy_clear)
+```
+
+### ITS: patient counts, no clearance period, no seasonality effect
 
@@
 time_series_analysis_plotting(ts_linear_count,
-                              patient_count_monthly_psy, "count_psy",
+                              patient_count_psy_period_psy,
+                              "count_psy_patients",
                               clearance_period = FALSE) +
@@
 
-### ITS: clearance period, no seasonality effect
+### ITS: patient counts, clearance period, no seasonality effect
 
@@
 time_series_analysis_plotting(ts_linear_count_clear,
-                              patient_count_monthly_psy, "count_psy",
+                              patient_count_psy_period_psy,
+                              "count_psy_patients",
                               clearance_period = TRUE) +
@@
 
-### ITS: no clearance period, with seasonality effect
+### ITS: patient counts, no clearance period, with seasonality effect
 
@@
 time_series_analysis_plotting(ts_seasonal_count,
-                              patient_count_monthly_psy, "count_psy",
+                              patient_count_psy_period_psy, "count_psy",
                               clearance_period = FALSE) +
@@
 
-### ITS: with clearance period, with seasonality effect
+### ITS: patient counts, with clearance period, with seasonality effect
 
@@
 time_series_analysis_plotting(ts_seasonal_count_clear,
-                              patient_count_monthly_psy, "count_psy",
+                              patient_count_psy_period_psy, "count_psy",
                               clearance_period = TRUE) +
@@
        y = "Count")
+```
+
+### ITS: psy percentage, no clearance period, no seasonality effect
+
+```{r}
+time_series_analysis_plotting(ts_linear_percent,
+                              patient_count_psy_period_psy,
+                              "percentage",
+                              clearance_period = FALSE) +
+  labs(title = "Interrupted Time-Series Analysis: Percentage of encounters with psychiatric ICD codes",
+       subtitle = "Without seasonality effect",
+       x = time_period,
+       y = "Percentage")
+```
+
+### ITS: psy percentage, clearance period, no seasonality effect
+
+```{r}
+time_series_analysis_plotting(ts_linear_percent_clear,
+                              patient_count_psy_period_psy,
+                              "percentage",
+                              clearance_period = TRUE) +
+  labs(title = "Interrupted Time-Series Analysis: Percentage of encounters with psychiatric ICD codes",
+       subtitle = "Without seasonality effect",
+       x = time_period,
+       y = "Percentage")
+```
+
+### ITS: psy percentage, no clearance period, with seasonality effect
+
+```{r}
+time_series_analysis_plotting(ts_seasonal_percent,
+                              patient_count_psy_period_psy,
+                              "percentage",
+                              clearance_period = FALSE) +
+  labs(title = "Interrupted Time-Series Analysis: Percentage of encounters with psychiatric ICD codes",
+       subtitle = "Without seasonality effect",
+       x = time_period,
+       y = "Percentage")
+```
+
+### ITS: psy percentage, with clearance period, with seasonality effect
+
+```{r}
+time_series_analysis_plotting(ts_seasonal_percent_clear,
+                              patient_count_psy_period_psy,
+                              "percentage",
+                              clearance_period = TRUE) +
+  labs(title = "Interrupted Time-Series Analysis: Percentage of encounters with psychiatric ICD codes",
+       subtitle = "With seasonality effect",
+       x = time_period,
+       y = "Percentage")
 ```

### 2021-10-06 psy_peds_3.Rmd
fixing bug

@@
 time_series_analysis_plotting(ts_seasonal_count,
-                              patient_count_psy_period_psy, "count_psy",
+                              patient_count_psy_period_psy,
+                              "count_psy_patients",
                               clearance_period = FALSE) +
@@
 time_series_analysis_plotting(ts_seasonal_count_clear,
-                              patient_count_psy_period_psy, "count_psy",
+                              patient_count_psy_period_psy,
+                              "count_psy_patients",
                               clearance_period = TRUE) +

### 2021-10-17 psy_peds_3.Rmd
update Rmd code to plot specific disorder subgroups

@@
 
+#### Subset: the most prevalent disorder group
+```{r}
+dg_subset <- count_disorder_group %>%
+  filter( disorder_group %in% c("Anxiety Disorders", "Depressive Disorders", "Suicide or Self-Injury" )) %>%
+  ggplot(aes(x = time_p, y= count_dg, fill = period, group = disorder_group )) +
+  geom_bar(stat = "identity")+
+  geom_vline(xintercept = as.Date(pandemic_start_date),
+             linetype = "dashed") +
+  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se= TRUE, aes(colour=period), colour= "black", size = 0.5) + 
+  labs(x = time_period,
+       y = "Patient Count",
+       title = paste0("Per ", time_period, " patient counts with mental health related ICD codes"), 
+       subtitle = "Grouped by mental health ICD subcategories") + 
+  ylim(0, NA) + 
+  facet_wrap(. ~ disorder_group, scales = "free_y") +
+  theme(strip.text.x = element_text(size = 12),
+        axis.text.x = element_text(size = 12, angle = 90), 
+        axis.text.y = element_text(size = 11), 
+        axis.title.x = element_text(size = 12),
+        axis.title.y = element_text(size = 11),
+        title = element_text(size = 12),
+        plot.subtitle = element_text(size = 11),
+        legend.position = "none") +
+  scale_fill_manual(values = cbPalette[-1])
+
+dg_subset
+```
+
 ### New admissions: percentage
@@
 ```
+```{r}
+dgPerc_subset <- perc_disorder_group %>%  
+  filter( disorder_group %in% c("Anxiety Disorders", "Depressive Disorders", "Suicide or Self-Injury" )) %>%
+  ggplot(aes(x = time_p, y= percentage_dg, fill = period, group = disorder_group )) +
+  geom_bar(stat = "identity")+
+  geom_vline(xintercept = as.Date(pandemic_start_date),
+             linetype = "dashed") +
+  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se= TRUE, aes(colour=period), colour= "black", size = 0.5) + 
+  labs(x = time_period,
+       y = "Patient (%)",
+       title = paste0("Percentage of patients per ", time_period,"with mental health related ICD codes"), 
+       subtitle = "Grouped by mental health ICD subcategories") + 
+  ylim(0, NA) + 
+  facet_wrap(. ~ disorder_group, scales = "free_y") +
+  theme(strip.text.x = element_text(size = 12),
+        axis.text.x = element_text(size = 12, angle = 90), 
+        axis.text.y = element_text(size = 12), 
+        axis.title.x = element_text(size = 12),
+        axis.title.y = element_text(size = 12),
+        title = element_text(size = 12),
+        plot.subtitle = element_text(size = 11),
+        legend.position = "none") +
+  scale_fill_manual(values = cbPalette[-1])
+
+dgPerc_subset
+```
 

### 2021-10-20 psy_peds_3.Rmd
bootstrap added

@@
 
+### ITS bootstrap
+
+```{r, eval=FALSE}
+### list to save the bootstrap results 
+list_results <- list()
+list_predict_results <- list()
+number_of_iterations <- 1000
+
+set.seed( seed = 1 )
+
+for( i in 1:number_of_iterations){
+#  print(i)
+  sample_size = nrow(all_codes_desc)
+  sample_size
+  row_indices <- 1:nrow(all_codes_desc)
+
+  sample_df <- all_codes_desc[sample(row_indices, size = sample_size, replace = TRUE),]
+  dim(sample_df)
+
+  sample_df_psy <- sample_df %>% 
+    filter(concept_code %in% icdCodes$ICD10_Code)
+
+  all_patients_bootstrap <- unique( sample_df[, c("patient_num", "time_p")])
+  patients_with_psy_conditions_bootstrap <- unique( sample_df_psy[, c("patient_num", "time_p")] )
+  patients_without_psy_conditions_bootstrap <- setdiff(all_patients_bootstrap,patients_with_psy_conditions_bootstrap)
+
+  patients_without_psy_conditions_counts_bootstrap <- patients_without_psy_conditions_bootstrap %>%
+    group_by( time_p ) %>%
+    summarise(count_no_psy_patients =  ifelse(n_distinct(patient_num) > obfuscation | isFALSE(obfuscation),
+                                              n_distinct(patient_num), 
+                                              0.5)) %>%
+    mutate( period = ifelse( time_p < pandemic_start_date,
+                             "before_pandemic",
+                             "during_pandemic"))
+
+  patients_with_psy_conditions_counts_bootstrap <-  patients_with_psy_conditions_bootstrap %>%
+    group_by( time_p ) %>%
+    summarise(count_psy_patients =  ifelse(n_distinct(patient_num) > obfuscation | isFALSE(obfuscation),
+                                           n_distinct(patient_num), 
+                                           0.5)) %>%
+    mutate( period = ifelse( time_p < pandemic_start_date,
+                             "before_pandemic",
+                             "during_pandemic"))
+
+  ratios_patients_with_without_psy_bootstrap <- patients_without_psy_conditions_counts_bootstrap %>%
+    full_join( patients_with_psy_conditions_counts_bootstrap, by=c("time_p", "period")) %>%
+    mutate( ratio = count_psy_patients/count_no_psy_patients, 
+            percentage = 100 * count_psy_patients/(count_no_psy_patients + count_psy_patients))
+
+  patient_count_psy_period_psy_bootstrap <- ratios_patients_with_without_psy_bootstrap %>%
+    filter(time_p > as.Date(cut( start_date_plots, breaks = time_period)) &
+             time_p < as.Date(cut( end_date_plots, breaks = time_period))) %>%
+    mutate(time = seq_along(time_p),
+           month_year = format(time_p, "%m")
+    )
+
+  ts_linear_percent_bootstrap <- lm(percentage ~ period * time, patient_count_psy_period_psy_bootstrap)
+  output_model_bootstrap <- broom::tidy( ts_linear_percent_bootstrap )
+  bootstrap_results <- output_model_bootstrap["estimate"]
+  list_results[[i]] <- bootstrap_results
+
+  pred_model_bootstrap <- predict(ts_linear_percent_bootstrap, patient_count_psy_period_psy_bootstrap, type = "response") %>% as.data.frame()
+  pred_model_bootstrap <- bind_cols(patient_count_psy_period_psy_bootstrap, pred_model_bootstrap)
+  list_predict_results[[i]] <- pred_model_bootstrap
+
+}
+```
+
+```{r }
+load("../ooutput_bootsrap1000iterations.RData")
+load("../output_bootstrap_predict_results1000iterations.RData")
+
+ggplot(data = output_bootstrap_predict_results, aes(x = time_p, 
+                                                    y = fittedValue)) + 
+  geom_line(aes(group = iteration,
+                color = period), 
+            data = output_bootstrap_predict_results[output_bootstrap_predict_results$period == "before_pandemic", ], 
+            alpha = 0.05) +
+  geom_line(aes(group = iteration,
+                color = period), 
+            data = output_bootstrap_predict_results[output_bootstrap_predict_results$period == "during_pandemic", ], 
+            alpha = 0.05) +
+  geom_smooth(formula = y ~ x, data = output_bootstrap_predict_results[output_bootstrap_predict_results$period == "during_pandemic", ], 
+              color = "black") +
+  geom_smooth(formula = y ~ x, data = output_bootstrap_predict_results[output_bootstrap_predict_results$period == "before_pandemic", ], 
+              color = "black")
+```
+
+```{r }
+
+```
+
 ## Psychiatric conditions aggregating by subcategories

### 2021-10-20 psy_peds_3.Rmd
adding parametric bootstrap

@@
 
+#### Parametric bootstrap
+
+```{r}
+bootstrap_size <- 1000
+binom_proportion <- function(n, prob, size) {
+  (rbinom(n, size, prob) / size) * 100
+}
+bootstrapped_timepoints <- mapply(binom_proportion,
+                                  n = bootstrap_size,
+                                  size = patient_count_psy_period_psy$count_no_psy_patients +
+                                    patient_count_psy_period_psy$count_psy_patients,
+                                  prob = patient_count_psy_period_psy$percentage/100)
+
+bootstrapped_sample_list <- vector(mode = "list", length = bootstrap_size)
+bootstrapped_coefficients_list <- vector(mode = "list", length = bootstrap_size)
+for (iteration in 1:bootstrap_size) {
+  df_iteration <- data.frame(period = patient_count_psy_period_psy$period,
+                             time = patient_count_psy_period_psy$time,
+                             percentage = bootstrapped_timepoints[iteration, ])
+  ts_linear_percent <- lm(percentage ~ period * time, df_iteration)
+  linear_model <- broom::tidy(ts_linear_percent) %>%
+    filter(term %in% c("time", "periodduring_pandemic:time"))
+  coefficients <- linear_model$estimate
+  names(coefficients) <- linear_model$term
+  bootstrapped_coefficients_list[[iteration]] <- coefficients
+  df_iteration$predicted <- predict(ts_linear_percent)
+  df_iteration$time_p <- patient_count_psy_period_psy$time_p
+  bootstrapped_sample_list[[iteration]] <- df_iteration
+}
+
+quantile_bootstrapped_fitted <- bootstrapped_fitted_df %>%
+  group_by(time_p, period) %>%
+  summarise("2.5%" = quantile(predicted, 0.025),
+            "50%" = quantile(predicted, 0.5),
+            "97.5%" = quantile(predicted, 0.975)) %>%
+  pivot_longer(cols = c("2.5%", "50%", "97.5%"),
+               names_to = "quantiles")
+
+bootstrapped_fitted_df <- bind_rows(bootstrapped_sample_list, .id = "iteration")
+bootstrapped_coefficients_df <- bind_rows(bootstrapped_coefficients_list, .id = "iteration")
+mean(bootstrapped_coefficients_df$time)
+mean(bootstrapped_coefficients_df$time + bootstrapped_coefficients_df[["periodduring_pandemic:time"]])
+pvalue_coeff_interaction <- round(t.test(bootstrapped_coefficients_df[["periodduring_pandemic:time"]])$p.value, 4)
+pvalue_coeff_interaction <- ifelse(pvalue_coeff_interaction == 0,
+                                   "< 1e-4",
+                                   str(pvalue_coeff_interaction))
+plot_annotation <- paste0("Diff. in slope coeffs before and after pandemic: ",
+                     round(mean(bootstrapped_coefficients_df[["periodduring_pandemic:time"]]), 2),
+                     " (pvalue: ",
+                     pvalue_coeff_interaction,
+                     ")")
+
+ggplot(bootstrapped_fitted_df, aes(x = time_p, y = predicted, group = interaction(iteration, period))) +
+  geom_line(aes(color = period),
+            alpha = 0.03) +
+  geom_line(data = quantile_bootstrapped_fitted,
+            aes(y = value, group = interaction(quantiles, period)),
+            linetype = "dashed") +
+  geom_line(data = quantile_bootstrapped_fitted[quantile_bootstrapped_fitted$quantiles == "50%", ],
+            aes(y = value, group = period)) +
+  geom_point(data = patient_count_psy_period_psy,
+             aes(y = percentage),
+             alpha = 0.5) +
+  scale_colour_manual(values = cbPalette) +
+  guides(color = guide_legend(override.aes = list(alpha = 1))) +
+  annotate("text",
+           x = as.Date("2019-10-01"),
+           y = max(patient_count_psy_period_psy[["percentage"]]) + 5,
+           label = annotation) +
+  labs(title = "Interrupted Time Series Analysis: Percentage of patients with psychiatric conditions",
+       subtitle = "Parametric bootstrap estimation: (1000 iterations)",
+       x = "Calendar dates",
+       y = "Percentage",
+       color = "Period")
+
+```
+
+#### Resampling bootstrap
 ```{r, eval=FALSE}

### 2021-10-25 psy_peds_3.Rmd
adding the code to save the outputs and remove the resampling bootstrap code

--- 2021-10-17 psy_peds_3.Rmd
@@
 
+### ITS bootstrap
+
+#### Parametric bootstrap
+
+```{r}
+bootstrap_size <- 1000
+binom_proportion <- function(n, prob, size) {
+  (rbinom(n, size, prob) / size) * 100
+}
+bootstrapped_timepoints <- mapply(binom_proportion,
+                                  n = bootstrap_size,
+                                  size = patient_count_psy_period_psy$count_no_psy_patients +
+                                    patient_count_psy_period_psy$count_psy_patients,
+                                  prob = patient_count_psy_period_psy$percentage/100)
+
+bootstrapped_sample_list <- vector(mode = "list", length = bootstrap_size)
+bootstrapped_coefficients_list <- vector(mode = "list", length = bootstrap_size)
+for (iteration in 1:bootstrap_size) {
+  df_iteration <- data.frame(period = patient_count_psy_period_psy$period,
+                             time = patient_count_psy_period_psy$time,
+                             percentage = bootstrapped_timepoints[iteration, ])
+  ts_linear_percent <- lm(percentage ~ period * time, df_iteration)
+  linear_model <- broom::tidy(ts_linear_percent) %>%
+    filter(term %in% c("time", "periodduring_pandemic:time"))
+  coefficients <- linear_model$estimate
+  names(coefficients) <- linear_model$term
+  bootstrapped_coefficients_list[[iteration]] <- coefficients
+  df_iteration$predicted <- predict(ts_linear_percent)
+  df_iteration$time_p <- patient_count_psy_period_psy$time_p
+  bootstrapped_sample_list[[iteration]] <- df_iteration
+}
+
+bootstrapped_fitted_df <- bind_rows(bootstrapped_sample_list, .id = "iteration")
+
+quantile_bootstrapped_fitted <- bootstrapped_fitted_df %>%
+  group_by(time_p, period) %>%
+  summarise("2.5%" = quantile(predicted, 0.025),
+            "50%" = quantile(predicted, 0.5),
+            "97.5%" = quantile(predicted, 0.975)) %>%
+  pivot_longer(cols = c("2.5%", "50%", "97.5%"),
+               names_to = "quantiles")
+
+bootstrapped_coefficients_df <- bind_rows(bootstrapped_coefficients_list, .id = "iteration")
+mean(bootstrapped_coefficients_df$time)
+mean(bootstrapped_coefficients_df$time + bootstrapped_coefficients_df[["periodduring_pandemic:time"]])
+pvalue_coeff_interaction <- round(t.test(bootstrapped_coefficients_df[["periodduring_pandemic:time"]])$p.value, 4)
+pvalue_coeff_interaction <- ifelse(pvalue_coeff_interaction == 0,
+                                   "< 1e-4",
+                                   str(pvalue_coeff_interaction))
+plot_annotation <- paste0("Diff. in slope coeffs before and after pandemic: ",
+                     round(mean(bootstrapped_coefficients_df[["periodduring_pandemic:time"]]), 2),
+                     " (pvalue: ",
+                     pvalue_coeff_interaction,
+                     ")")
+
+# ggplot(bootstrapped_fitted_df, aes(x = time_p, y = predicted, group = interaction(iteration, period))) +
+#   geom_line(aes(color = period),
+#             alpha = 0.03) +
+#   geom_line(data = quantile_bootstrapped_fitted,
+#             aes(y = value, group = interaction(quantiles, period)),
+#             linetype = "dashed") +
+#   geom_line(data = quantile_bootstrapped_fitted[quantile_bootstrapped_fitted$quantiles == "50%", ],
+#             aes(y = value, group = period)) +
+#   geom_point(data = patient_count_psy_period_psy,
+#              aes(y = percentage),
+#              alpha = 0.5) +
+#   scale_colour_manual(values = cbPalette) +
+#   guides(color = guide_legend(override.aes = list(alpha = 1))) +
+#   annotate("text",
+#            x = as.Date("2019-10-01"),
+#            y = max(patient_count_psy_period_psy[["percentage"]]) + 5,
+#            label = annotation) +
+#   labs(title = "Interrupted Time Series Analysis: Percentage of patients with psychiatric conditions",
+#        subtitle = "Parametric bootstrap estimation: (1000 iterations)",
+#        x = "Calendar dates",
+#        y = "Percentage",
+#        color = "Period")
+
+```
+
 ## Psychiatric conditions aggregating by subcategories
@@
 ```
+
+# Save the output in a RData file. 
+```{r}
+#length_hospitalisation_values <- input_data %>% 
+#  dplyr::distinct( patient_num, n_hospitalisation, len_hospitalisation ) %>%
+#  dplyr::mutate( siteid = site ) %>%
+#  dplyr::select( len_hospitalisation, siteid )
+save( count_icd,
+      patient_count_psy_period_psy,
+      patient_count_psy_period_psy_clear,
+      bootstrapped_coefficients_df,
+      bootstrapped_fitted_df,
+      perc_disorder_group,
+      patients_hospitalized_agg_sex_perc,
+      table1,
+#      length_hospitalisation_values,
+#      obs_by_day,
+      file = paste0("output/peds_psy_aggResults_", unique(as.character(demo_raw$siteid)), ".RData"))
+```

### 2021-10-25 psy_peds_3.Rmd
code added to save the output aggregate counts

@@
 clinical_filtered <- clinical_raw %>%
-  filter( len_hospitalisation > min_hosp_days, 
+  filter( len_hospitalisation >= min_hosp_days, 
           cohort  %in% c("AllAdm")) %>%
@@
 stopifnot(unique(all_codes$in_hospital) == 1)
-stopifnot(unique(all_codes$len_hospitalisation > min_hosp_days))
+stopifnot(unique(all_codes$len_hospitalisation >= min_hosp_days))
 stopifnot(nrow(all_codes) == nrow(observation_table))
@@
 stopifnot(unique(all_codes_ado$in_hospital) == 1)
-stopifnot(unique(all_codes_ado$len_hospitalisation > min_hosp_days))
+stopifnot(unique(all_codes_ado$len_hospitalisation >= min_hosp_days))
 stopifnot(nrow(all_codes_ado) == nrow(observation_table_ado))
@@
 
-# ggplot(bootstrapped_fitted_df, aes(x = time_p, y = predicted, group = interaction(iteration, period))) +
-#   geom_line(aes(color = period),
-#             alpha = 0.03) +
-#   geom_line(data = quantile_bootstrapped_fitted,
-#             aes(y = value, group = interaction(quantiles, period)),
-#             linetype = "dashed") +
-#   geom_line(data = quantile_bootstrapped_fitted[quantile_bootstrapped_fitted$quantiles == "50%", ],
-#             aes(y = value, group = period)) +
-#   geom_point(data = patient_count_psy_period_psy,
-#              aes(y = percentage),
-#              alpha = 0.5) +
-#   scale_colour_manual(values = cbPalette) +
-#   guides(color = guide_legend(override.aes = list(alpha = 1))) +
-#   annotate("text",
-#            x = as.Date("2019-10-01"),
-#            y = max(patient_count_psy_period_psy[["percentage"]]) + 5,
-#            label = annotation) +
-#   labs(title = "Interrupted Time Series Analysis: Percentage of patients with psychiatric conditions",
-#        subtitle = "Parametric bootstrap estimation: (1000 iterations)",
-#        x = "Calendar dates",
-#        y = "Percentage",
-#        color = "Period")
+ggplot(bootstrapped_fitted_df, aes(x = time_p, y = predicted, group = interaction(iteration, period))) +
+  geom_line(aes(color = period),
+            alpha = 0.03) +
+  geom_line(data = quantile_bootstrapped_fitted,
+            aes(y = value, group = interaction(quantiles, period)),
+            linetype = "dashed") +
+  geom_line(data = quantile_bootstrapped_fitted[quantile_bootstrapped_fitted$quantiles == "50%", ],
+            aes(y = value, group = period)) +
+  geom_point(data = patient_count_psy_period_psy,
+             aes(y = percentage),
+             alpha = 0.5) +
+  scale_colour_manual(values = cbPalette) +
+  guides(color = guide_legend(override.aes = list(alpha = 1))) +
+  annotate("text",
+           x = as.Date("2019-10-01"),
+           y = max(patient_count_psy_period_psy[["percentage"]]) + 5,
+           label = plot_annotation) +
+  labs(title = "Interrupted Time Series Analysis: Percentage of patients with psychiatric conditions",
+       subtitle = "Parametric bootstrap estimation: (1000 iterations)",
+       x = "Calendar dates",
+       y = "Percentage",
+       color = "Period")
 
@@
 ```{r}
-#length_hospitalisation_values <- input_data %>% 
-#  dplyr::distinct( patient_num, n_hospitalisation, len_hospitalisation ) %>%
-#  dplyr::mutate( siteid = site ) %>%
-#  dplyr::select( len_hospitalisation, siteid )
+length_hospitalisation_values <- all_codes_psy %>% 
+  dplyr::distinct( patient_num, n_hospitalisation, len_hospitalisation, period ) %>%
+  dplyr::mutate( siteid = site ) %>%
+  dplyr::select( len_hospitalisation, siteid, period )
+
+# add the site id as column for each of the output data frames
+count_icd$siteid = site 
+patient_count_psy_period_psy$siteid = site 
+patient_count_psy_period_psy_clear$siteid = site 
+bootstrapped_coefficients_df$siteid = site 
+bootstrapped_fitted_df$siteid = site 
+perc_disorder_group$siteid = site 
+patients_hospitalized_agg_sex_perc$siteid = site 
+table1$siteid = site
+
 save( count_icd,
@@
       table1,
-#      length_hospitalisation_values,
-#      obs_by_day,
-      file = paste0("output/peds_psy_aggResults_", unique(as.character(demo_raw$siteid)), ".RData"))
-```
+      length_hospitalisation_values,
+      file = paste0("output/peds_psy_aggResults_", site, ".RData"))
+```

### 2021-10-26 psy_peds_3.Rmd
update the saved outputs

@@
 # estimate the counts
-count_icd <- all_codes_desc %>%
+count_icd_all <- all_codes_desc %>%
   filter(date < end_date_plots &
@@
 # join the total with the disorder group count to estimate the percentage
-perc_disorder_group <- count_icd %>%
+perc_disorder_group <- count_icd_all %>%
   left_join(count_disorder_group, 

### 2021-10-26 psy_peds_3.Rmd
add one additional output

@@
 table1$siteid = site 
+ratios_patients_with_without_psy$siteid = site
 
 save( count_icd,
+      ratios_patients_with_without_psy,
       patient_count_psy_period_psy,

### 2021-10-27 metanalysis_visualization/pedsResults.Rmd
new folder added with code for metanalysis and visualization putting together the results from different participating sites

---
title: "pediatricResults"
author: "aGutierrezSacristan"
date: "10/26/2021"
output: html_document
---
```{r libraries, message=FALSE, warning=FALSE, results='hide'}
# Install packages
paket <- function(pak){
  new_pak <- pak[!(pak %in% rownames(installed.packages()))]
  if (length(new_pak)) 
    install.packages(new_pak, dependencies = TRUE,repos="https://cloud.r-project.org/")
  sapply(pak, library, character.only = TRUE)
}
listOfPackages <- c("tidyverse", "RColorBrewer", "knitr", "kableExtra", "tsModel", "gridExtra", "dplyr", "metafor", "meta")
paket(listOfPackages)
```

```{r}
# options(repr.plot.width = 18, repr.plot.height = 9)
theme_plots <- theme_bw() +
  theme(strip.text = element_text(size = 5),
        axis.text.x = element_text(size = 8), 
        axis.text.y = element_text(size = 6), 
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8),
        title = element_text(size = 10),
        plot.subtitle = element_text(size = 9, face = "italic")) 
theme_set(theme_plots)

# Colorblind palette
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

```{r}
# cut-off dates used in the analysis, these should not be changed
time_period <-"months" #possible values: days, weeks, months
history_start_date <- as.Date("2019-01-01")
history_end_date <- as.Date("2021-05-31")
pandemic_start_date <- as.Date("2020-03-15")
start_date_plots <- as.Date("2019-01-15")
end_date_plots <- as.Date("2021-05-01")
```

## Read the input file by site
Each site shared the aggregated counts results. Putting all the results together.

```{r}
outputDir <- "./output/"
outputFiles <- list.files( path = outputDir, pattern = ".RData")
```

Put all the aggregate counts from the different sites together
```{r}
for( i in 1:length( outputFiles )){
  load(paste0( outputDir, outputFiles[i]))
  if( i == 1 ){
    count_icd_all <- count_icd
    patients_hospitalized_agg_sex_perc_all <- patients_hospitalized_agg_sex_perc
    patient_count_psy_period_psy_all <- patient_count_psy_period_psy
    perc_disorder_group_all <- perc_disorder_group
    ratios_patients_with_without_psy_all <- ratios_patients_with_without_psy
  }else{
    count_icd_all <- rbind(count_icd_all, count_icd)
    patients_hospitalized_agg_sex_perc_all <- rbind( patients_hospitalized_agg_sex_perc_all, patients_hospitalized_agg_sex_perc)
    patient_count_psy_period_psy_all <- rbind( patient_count_psy_period_psy_all,patient_count_psy_period_psy)
    perc_disorder_group_all <- rbind(perc_disorder_group_all, perc_disorder_group)
    ratios_patients_with_without_psy_all <- rbind( ratios_patients_with_without_psy_all, ratios_patients_with_without_psy)
  }
  rm(count_icd,patients_hospitalized_agg_sex_perc,patient_count_psy_period_psy,
     perc_disorder_group,ratios_patients_with_without_psy,
     patient_count_psy_period_psy_clear, table1, bootstrapped_coefficients_df,
     bootstrapped_fitted_df, length_hospitalisation_values)
}
```

## Patient Count with Mental Health Related ICD codes

```{r}
thpd_psyCount <- count_icd_all %>%
  pivot_longer(names_to = "counts", cols=c(count_psy, count_icd)) %>%
  mutate( counts = ifelse( counts == "count_psy", "Psy", "Total")) %>%
  ggplot(aes(x = time_p, y = value, fill = counts, color = counts)) +
  facet_grid(siteid~.) +
  geom_point() +
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  scale_fill_manual(values = cbPalette) +
  scale_color_manual(values = cbPalette) +
  labs(y = "Counts",
       x = paste0("Date (by ", time_period,")"),
       title = paste0("Per ", time_period , " patient counts with mental health related ICD codes")) + 
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)
thpd_psyCount
```

#### Both results together

```{r}
count_icd_all_combined <-  count_icd_all %>%
  group_by(time_p, period) %>% 
  summarise(count_psy_all = sum(count_psy), 
            count_icd_all = sum( count_icd ))

thpd_psyCount_combined <- count_icd_all_combined %>%
  pivot_longer(names_to = "counts", cols=c(count_psy_all, count_icd_all)) %>%
  mutate( counts = ifelse( counts == "count_psy_all", "Psy", "Total")) %>%
  ggplot(aes(x = time_p, y = value, fill = counts, color = counts)) +
  geom_point() +
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  scale_fill_manual(values = cbPalette) +
  scale_color_manual(values = cbPalette) +
  labs(y = "Counts",
       x = paste0("Date (by ", time_period,")"),
       title = paste0("Per ", time_period , " patient counts with mental health related ICD codes")) + 
  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)
thpd_psyCount_combined
```

## Ratio patients with vs without psychiatric conditions
```{r}
ratio_psy_non_psy <- ratios_patients_with_without_psy_all %>%
  ggplot(aes(x = time_p, y = ratio, fill = period, color = period)) +
  geom_point() +
  facet_grid(siteid~.) +
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  scale_fill_manual(values = cbPalette) +
  scale_color_manual(values = cbPalette) +
  labs(y = "Ratio (with psy conditions/without psy conditions)",
       x = paste0("Date (by ", time_period,")"),
       title = paste0("Ratio patient with vs without psychiatric conditions ( per ", time_period,")"))

ratio_psy_non_psy
```

## Percenge patients with vs without psychiatric conditions
```{r}
percentage_psy_non_psy <- ratios_patients_with_without_psy_all %>%
  ggplot(aes(x = time_p, y = percentage, color = period)) +
  facet_grid(siteid~.) +
  geom_point() +
  geom_vline(xintercept = as.Date(pandemic_start_date),
             linetype = "dashed") +
  scale_fill_manual(values = cbPalette) +
  scale_color_manual(values = cbPalette) +
  labs(y = "Percentage",
       x = paste0("Date (by ", time_period,")"),
       title = paste0("Percentage patient with vs total population ( per ", time_period,")")) +
  guides(color = guide_legend(title="Patient diagnoses types"))

percentage_psy_non_psy
```

### Test meta-analysis percentage

```{r}
#split this into before and after pandemic and run it twice (one for each subset)

#estimate the summary effect size
ratios_patients_with_without_psy_all$total <- ratios_patients_with_without_psy_all$count_no_psy_patients + ratios_patients_with_without_psy_all$count_psy_patients

ratios_bp <- ratios_patients_with_without_psy_all %>%
  filter( period == "before_pandemic")

ratios_dp <- ratios_patients_with_without_psy_all %>%
  filter( period == "during_pandemic")

metainputdata <- ratios_dp
#RR no transformation
#PLO the logit transformation
#PFT the double arcsine transformation
ies=escalc(xi=count_psy_patients, ni=total, data=metainputdata, measure="PR")
summary(ies$vi)

# pool the individual effect size
# DL random effects using the DerSimonian-Laird estimator
#REML random effects using the restricted maximum-likelihood estimator
pes = rma(yi, vi, data=ies, method="REML")
print(pes)
confint(pes)

pes.summary=metaprop(count_psy_patients, total, as.character(time_p),
                     data=metainputdata, 
                     sm="PRAW")
forest(pes.summary,layout = "JAMA")
```

### Boxplot showing ratio per month

```{r}
ratios_patients_with_without_psy_all %>%
  ggplot( aes(x=as.factor(time_p), y=ratio)) +
    geom_boxplot() +
    scale_fill_viridis(discrete = TRUE, alpha=0.6) +
    geom_jitter(color="black", size=0.4, alpha=0.9,
                aes(shape = siteid)) +
    theme(
      legend.position="bottom",
      plot.title = element_text(size=11)
    ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6)) +
    ggtitle("Ratio boxplot") +
    xlab("")
```

#### Overlap lines: one per site
```{r}
ratios_patients_with_without_psy_all %>%
  ggplot( aes(x=as.factor(time_p), y=ratio, 
              group = siteid, color = siteid))+
  geom_point() + 
  geom_line() +
  scale_fill_viridis(discrete = TRUE, alpha=0.6) +
    theme(
      legend.position="bottom",
      plot.title = element_text(size=11)
    ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6)) +
    ggtitle("Dot and line plot") +
    xlab("period (months)")

```

### 2021-10-28 metanalysis_visualization/pedsResults.Rmd
adding trends meta-analysis

@@
 ```
+
+#### Meta-analysis of temporal trends
+
+```{r}
+# Fixed effect
+
+load("output/peds_psy_aggResults_BCH.RData")
+df_meta <- bootstrapped_fitted_df %>%
+  filter(iteration %in% 1:4) %>%
+  mutate(siteid = as.factor(iteration))
+
+cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
+theme_plots <- theme_bw() +
+  theme(strip.text = element_text(size = 5),
+        axis.text.x = element_text(size = 8),
+        axis.text.y = element_text(size = 6),
+        axis.title.x = element_text(size = 8),
+        axis.title.y = element_text(size = 8),
+        title = element_text(size = 10),
+        plot.subtitle = element_text(size = 9, face = "italic"))
+theme_set(theme_plots)
+
+# Colorblind palette
+cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
+
+## Fixed Effect model
+
+fixed_model <- lm(
+  percentage ~ time*period + time*siteid,
+  data = df_meta)
+
+result_fixed_model <- broom::tidy(fixed_model)
+
+df_meta$fitted_fixed <- fitted(fixed_model)
+
+pvalue <- result_fixed_model[result_fixed_model$term == "time:periodduring_pandemic",
+                             "p.value"][[1]] %>%
+  round(4)
+
+annotation_plot <- paste0(
+  "Difference of trend before/during pandemic: p",
+  ifelse(pvalue == 0,
+         "< 1e^-4",
+         paste0("=", as.character(pvalue)))
+)
+
+ggplot(data = df_meta, aes(x = time_p,
+                           y = percentage,
+                           color = siteid)) +
+  geom_point() +
+  geom_line(aes(y=fitted_fixed), data = df_meta[df_meta$period == "before_pandemic", ]) +
+  geom_line(aes(y=fitted_fixed), data = df_meta[df_meta$period == "during_pandemic", ]) +
+  geom_line(aes(y=predicted, color=NULL), data = df_meta[df_meta$period == "before_pandemic", ]) +
+  geom_line(aes(y=predicted, color=NULL), data = df_meta[df_meta$period == "during_pandemic", ]) +
+  scale_fill_manual(values = cbPalette) +
+  scale_color_manual(values = cbPalette) +
+  labs(title = "Meta-analysis: percentage of visits with Mental Health ICD code across sites",
+       subtitle = "Sites effect treated as fixed effect",
+       color = "Sites",
+       x = "Calendar date",
+       y = "Percentage"
+       ) +
+  annotate("text",
+           x = as.Date("2019-10-01"),
+           y = max(df_meta[["percentage"]]),
+           label = annotation_plot,
+           )
+
+# Same but with random effect for the sites effect on trends
+library(lme4)
+mixed_model <- lmer(percentage ~ time*period + (time|siteid),
+                    data = df_meta)
+
+```

### 2021-11-04 metanalysis_visualization/pedsResults.Rmd
adding general trend meta analysis

--- 2021-10-27 metanalysis_visualization/pedsResults.Rmd
@@
 }
-listOfPackages <- c("tidyverse", "RColorBrewer", "knitr", "kableExtra", "tsModel", "gridExtra", "dplyr", "metafor", "meta")
+listOfPackages <- c("tidyverse", "RColorBrewer", "knitr", "kableExtra", "tsModel", "gridExtra", "dplyr", "metafor", "meta", "viridis")
 paket(listOfPackages)
@@
 ```{r}
-outputDir <- "./output/"
-outputFiles <- list.files( path = outputDir, pattern = ".RData")
+outputDir <- "../output/"
+outputFiles <- list.files( path = outputDir, pattern = "\\.RData")
+getwd()
 ```
 
-Put all the aggregate counts from the different sites together
+Put all the aggregated counts from the different sites together
 ```{r}
@@
 ```
+
+#### Meta-analysis of temporal trends
+
+```{r}
+## Estimate sampling variances
+meta_df <- ratios_patients_with_without_psy_all
+meta_df$time <- as.numeric(as.factor(meta_df$time_p))
+##### Sampling model 
+trend_model <- "percentage ~ period * time"
+trends <- meta_df %>% 
+  group_by(siteid) %>% 
+  group_modify( ~ broom::tidy(lm(as.formula(trend_model), data = .x))) %>%
+  filter(term == c("periodduring_pandemic:time")) %>% 
+  mutate(variance = std.error * std.error) %>% 
+  select(siteid, estimate, variance)
+
+```
+
+```{r}
+meta_df$sampling_variances
+## Fixed Effect model
+
+model_meta <- metafor::rma(yi = estimate, 
+             vi = variance,
+             measure = "GEN",
+  data = trends)
+
+model_meta %>% 
+  broom::tidy()
+
+predict(model_meta)
+fitted(model_meta)
+
+metafor::forest(model_meta,
+                atransf=exp,
+                slab = trends$siteid,
+                header="Sites",
+                mlab="Overall temporal trend")
+```

### 2021-11-09 metanalysis_visualization/visualization.Rmd
new markdown for the visualization added

--- 2021-11-04 metanalysis_visualization/pedsResults.Rmd
@@
 ---
-title: "pediatricResults"
-author: "aGutierrezSacristan"
-date: "10/26/2021"
-output: html_document
+title: "Results visualization across sites"
+output:
+  html_document:
+    code_folding: hide
+    toc: yes
+    toc_float: yes
+editor_options: 
+  markdown: 
+    wrap: 72
 ---
+
+## Setting up the R environment
+
+### Installing packages and loading the library
+
 ```{r libraries, message=FALSE, warning=FALSE, results='hide'}
@@
 paket(listOfPackages)
+```
+
+### R session information
+
+```{r, results='hide'}
+sessionInfo()
 ```
@@
 ```{r}
-outputDir <- "../output/"
+outputDir <- "../../output/"
 outputFiles <- list.files( path = outputDir, pattern = "\\.RData")
+country_map <- read.delim( file = paste0( outputDir, "country_mapping.txt"))
 getwd()
@@
 
-## Patient Count with Mental Health Related ICD codes
+## Percentage: patients with psychiatric conditions / total number of patients
+
+### By site
 
 ```{r}
-thpd_psyCount <- count_icd_all %>%
-  pivot_longer(names_to = "counts", cols=c(count_psy, count_icd)) %>%
-  mutate( counts = ifelse( counts == "count_psy", "Psy", "Total")) %>%
-  ggplot(aes(x = time_p, y = value, fill = counts, color = counts)) +
-  facet_grid(siteid~.) +
-  geom_point() +
+ratios_patients_with_without_psy_all %>%
+  filter(time_p <= history_end_date &
+         time_p >= history_start_date ) %>%
+  ggplot( aes(x=time_p, y=percentage, 
+              group = siteid, color = siteid))+
+  geom_point( aes(shape=siteid, color=siteid) )+
+    geom_line() +
+  geom_vline(xintercept = as.Date(pandemic_start_date),
+             linetype = "dashed") +
+  scale_fill_viridis(discrete = TRUE, alpha=0.6) +
+    theme(
+      legend.position="bottom",
+      plot.title = element_text(size=11)
+    ) +
+  scale_x_date(date_labels = "%b-%d-%Y", breaks = "month") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6)) +
+    ggtitle("Percentage: patients with psychiatric conditions / total number of patients") +
+    xlab("period (months)")
+```
+
+### By Country
+
+```{r}
+ratios_patients_with_without_psy_all_country <- left_join( ratios_patients_with_without_psy_all, country_map )
+ratios_patients_with_without_psy_all_country %>%
+    filter(time_p <= history_end_date &
+         time_p >= history_start_date )  %>%
+  ggplot(aes(x = time_p, y = percentage, fill = siteid, color = siteid)) +
+  geom_point( aes(shape=siteid, color=siteid)  ) +
+  geom_line() +
+  facet_grid(country~.) +
   geom_vline(xintercept = as.Date(pandemic_start_date),
@@
   scale_color_manual(values = cbPalette) +
-  labs(y = "Counts",
+  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6)) +
+  scale_x_date(date_labels = "%b-%d-%Y", breaks = "month", limits = c( history_start_date, history_end_date  )) +
+  labs(y = "Ratio (with psy conditions/without psy conditions)",
        x = paste0("Date (by ", time_period,")"),
-       title = paste0("Per ", time_period , " patient counts with mental health related ICD codes")) + 
-  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)
-thpd_psyCount
+       title = paste0("Percentage: patients with psychiatric conditions / total number of patients ( per ", time_period,")"))
 ```
 
-#### Both results together
+## Ratio: patients with vs. patients without psychiatric conditions
+
+### By site
+```{r}
+ratios_patients_with_without_psy_all %>%
+  filter(time_p <= history_end_date &
+         time_p >= history_start_date ) %>%
+  ggplot( aes(x=time_p, y=ratio, 
+              group = siteid, color = siteid))+
+  geom_point( aes(shape=siteid, color=siteid) ) + 
+  geom_line() +
+  geom_vline(xintercept = as.Date(pandemic_start_date),
+             linetype = "dashed") +
+  scale_fill_viridis(discrete = TRUE, alpha=0.6) +
+    theme(
+      legend.position="bottom",
+      plot.title = element_text(size=11)
+    ) +
+  scale_x_date(date_labels = "%b-%d-%Y", breaks = "month") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6)) +
+    ggtitle("Ratio: patients with vs. patients without psychiatric conditions") +
+    xlab("period (months)")
+
+```
+
+### By Country
 
 ```{r}
-count_icd_all_combined <-  count_icd_all %>%
-  group_by(time_p, period) %>% 
-  summarise(count_psy_all = sum(count_psy), 
-            count_icd_all = sum( count_icd ))
-
-thpd_psyCount_combined <- count_icd_all_combined %>%
-  pivot_longer(names_to = "counts", cols=c(count_psy_all, count_icd_all)) %>%
-  mutate( counts = ifelse( counts == "count_psy_all", "Psy", "Total")) %>%
-  ggplot(aes(x = time_p, y = value, fill = counts, color = counts)) +
-  geom_point() +
+ratios_patients_with_without_psy_all_country %>%
+    filter(time_p <= history_end_date &
+         time_p >= history_start_date ) %>%
+  ggplot(aes(x = time_p, y = ratio, fill = siteid, color = siteid)) +
+  geom_point( aes(shape=siteid, color=siteid)  ) +
+  geom_line() +
+  facet_grid(country~.) +
   geom_vline(xintercept = as.Date(pandemic_start_date),
@@
   scale_color_manual(values = cbPalette) +
-  labs(y = "Counts",
-       x = paste0("Date (by ", time_period,")"),
-       title = paste0("Per ", time_period , " patient counts with mental health related ICD codes")) + 
-  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)
-thpd_psyCount_combined
-```
-
-## Ratio patients with vs without psychiatric conditions
-```{r}
-ratio_psy_non_psy <- ratios_patients_with_without_psy_all %>%
-  ggplot(aes(x = time_p, y = ratio, fill = period, color = period)) +
-  geom_point() +
-  facet_grid(siteid~.) +
-  geom_vline(xintercept = as.Date(pandemic_start_date),
-             linetype = "dashed") +
-  scale_fill_manual(values = cbPalette) +
-  scale_color_manual(values = cbPalette) +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6)) +
+  scale_x_date(date_labels = "%b-%d-%Y", breaks = "month", limits = c( history_start_date, history_end_date  )) +
   labs(y = "Ratio (with psy conditions/without psy conditions)",
@@
        title = paste0("Ratio patient with vs without psychiatric conditions ( per ", time_period,")"))
-
-ratio_psy_non_psy
 ```
 
-## Percenge patients with vs without psychiatric conditions
-```{r}
-percentage_psy_non_psy <- ratios_patients_with_without_psy_all %>%
-  ggplot(aes(x = time_p, y = percentage, color = period)) +
-  facet_grid(siteid~.) +
-  geom_point() +
-  geom_vline(xintercept = as.Date(pandemic_start_date),
-             linetype = "dashed") +
-  scale_fill_manual(values = cbPalette) +
-  scale_color_manual(values = cbPalette) +
-  labs(y = "Percentage",
-       x = paste0("Date (by ", time_period,")"),
-       title = paste0("Percentage patient with vs total population ( per ", time_period,")")) +
-  guides(color = guide_legend(title="Patient diagnoses types"))
+## Meta-analysis percentage
 
-percentage_psy_non_psy
-```
-
-### Test meta-analysis percentage
-
-```{r}
+```{r, results='hide', fig.show='hide'}
 #split this into before and after pandemic and run it twice (one for each subset)
@@
 
-#### Overlap lines: one per site
-```{r}
-ratios_patients_with_without_psy_all %>%
-  ggplot( aes(x=as.factor(time_p), y=ratio, 
-              group = siteid, color = siteid))+
-  geom_point() + 
-  geom_line() +
-  scale_fill_viridis(discrete = TRUE, alpha=0.6) +
-    theme(
-      legend.position="bottom",
-      plot.title = element_text(size=11)
-    ) +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6)) +
-    ggtitle("Dot and line plot") +
-    xlab("period (months)")
-
-```
-
-#### Meta-analysis of temporal trends
+## Meta-analysis of temporal trends
 

### 2021-11-09 psy_peds_3.Rmd
qc added to check which are the codes used in each site

--- 2021-10-26 psy_peds_3.Rmd
@@
 
-listOfPackages <- c("tidyverse", "RColorBrewer", "knitr", "kableExtra", "tsModel", "gridExtra", "dplyr")
+listOfPackages <- c("tidyverse", "RColorBrewer", "knitr", "kableExtra", "tsModel", "gridExtra", "dplyr", "lubridate")
 paket(listOfPackages)
@@
 stopifnot(nrow(distinct(all_codes_ado, patient_num, days_since_admission)) == nrow(distinct(observation_table_ado, patient_num, days_since_admission)))
+```
+
+#### QC step: do a summary of the mental health ICD codes present in each site
+As output we will generate a table with: the number of observations per code, number of patients diagnosed with each code and the date range in which the code was used.
+
+```{r}
+#filter the whole dataset to only mental codes
+mental_codes_qc <- all_codes_ado %>% 
+  filter( concept_code %in% mentalHealthCodes$concept_code )
+
+mental_codes_qc_output <- mental_codes_qc %>%
+  group_by( concept_code ) %>%
+  dplyr::summarise( distinct_patients = ifelse(length(unique(patient_num)) > obfuscation | isFALSE(obfuscation),
+                               length(unique(patient_num)), 
+                               0.5),
+                    distinct_observations = ifelse(length(patient_num) > obfuscation | isFALSE(obfuscation),
+                               length(patient_num), 
+                               0.5),                           
+                    min_date = min( dmy( calendar_date ) ), 
+                    max_date = max( dmy( calendar_date ) ))%>% 
+  arrange(desc(distinct_observations))
+
+mental_codes_qc_output <- left_join( mental_codes_qc_output, icdCodes, by=c("concept_code"= "ICD10_Code")) 
+mental_codes_qc_output  <- mental_codes_qc_output[, c("disorder_group", "description", "concept_code", "distinct_patients", 
+                                                    "distinct_observations", "min_date", "max_date")]
+
+mental_codes_qc_output[1:10, ]
 ```
@@
 # add the site id as column for each of the output data frames
+mental_codes_qc_output$siteid = site 
 count_icd$siteid = site 
@@
 table1$siteid = site 
+ratios_patients_with_without_psy$siteid = site
 
 save( count_icd,
+      mental_codes_qc_output,
+      ratios_patients_with_without_psy,
       patient_count_psy_period_psy,

### 2021-11-16 metanalysis_visualization/visualization.Rmd
adding country meta analysis, and fixing pvalue and left_join(obs_raw ...) bug)

@@
 ## Estimate sampling variances
-meta_df <- ratios_patients_with_without_psy_all
+meta_df <- ratios_patients_with_without_psy_all_country
 meta_df$time <- as.numeric(as.factor(meta_df$time_p))
-##### Sampling model 
+```
+
+```{r}
 trend_model <- "percentage ~ period * time"
-trends <- meta_df %>% 
+```
+
+```{r}
+trends_siteid <- meta_df %>% 
   group_by(siteid) %>% 
@@
 
+model_meta_siteid <- metafor::rma(yi = estimate, 
+             vi = variance,
+             measure = "GEN",
+             method = "DL",
+  data = trends_siteid) 
+metafor::forest(model_meta_siteid,
+                slab = trends_siteid$siteid,
+                header="Sites",
+                mlab="Overall temporal trend")
 ```
@@
 ```{r}
-meta_df$sampling_variances
-## Fixed Effect model
+trends_country <- meta_df %>% 
+  group_by(country) %>% 
+  group_modify( ~ broom::tidy(lm(as.formula(trend_model), data = .x))) %>%
+  filter(term == c("periodduring_pandemic:time")) %>% 
+  mutate(variance = std.error * std.error) %>% 
+  select(country, estimate, variance)
 
-model_meta <- metafor::rma(yi = estimate, 
+model_meta_country <- metafor::rma(yi = estimate, 
              vi = variance,
              measure = "GEN",
-  data = trends)
-
-model_meta %>% 
-  broom::tidy()
-
-predict(model_meta)
-fitted(model_meta)
-
-metafor::forest(model_meta,
-                atransf=exp,
-                slab = trends$siteid,
+             method = "DL",
+  data = trends_country) 
+metafor::forest(model_meta_country,
+                slab = trends_country$country,
                 header="Sites",

### 2021-11-16 psy_peds_3.Rmd
adding country meta analysis, and fixing pvalue and left_join(obs_raw ...) bug)

@@
 obs_raw_age <- left_join(obs_raw, demo_raw[c("patient_num", "age")], by = c("patient_num")) %>% 
-  left_join(clinical_raw[, c("patient_num", "days_since_admission", "calendar_date")], 
+  inner_join(clinical_raw[, c("patient_num", "days_since_admission", "calendar_date")], 
             by = c("patient_num", "days_since_admission"))
-stopifnot(nrow(obs_raw_age) == nrow(obs_raw))
 obs_raw_age$age_time_diagnosis <- obs_raw_age$age - floor(as.numeric(as.Date(data_update_date) - as.Date(obs_raw_age$calendar_date, dateFormat)) / 365.25)
@@
 pvalue_coeff_interaction <- round(t.test(bootstrapped_coefficients_df[["periodduring_pandemic:time"]])$p.value, 4)
-pvalue_coeff_interaction <- ifelse(pvalue_coeff_interaction == 0,
-                                   "< 1e-4",
-                                   str(pvalue_coeff_interaction))
 plot_annotation <- paste0("Diff. in slope coeffs before and after pandemic: ",

### 2021-11-17 metanalysis_visualization/visualization.Rmd
added visualization aggregated by country

@@
 
+#### Showing each independent site 
 ```{r}
@@
        title = paste0("Percentage: patients with psychiatric conditions / total number of patients ( per ", time_period,")"))
+```
+
+#### Aggregating the sites in each country
+
+```{r}
+ratios_patients_with_without_psy_all_country <- left_join( ratios_patients_with_without_psy_all, country_map )
+
+ratios_patients_with_without_psy_aggregated <- ratios_patients_with_without_psy_all_country %>%
+  group_by( time_p, country ) %>%
+  mutate( no_psy = sum( count_no_psy_patients ), 
+          psy = sum( count_psy_patients ), 
+          total = sum(count_no_psy_patients) + sum(count_psy_patients), 
+          ratio_by_country = psy / no_psy, 
+          perc_by_country = 100*(psy / total) ) %>%
+  select( time_p, no_psy, psy, total, ratio_by_country, perc_by_country, country)
+
+ratios_patients_with_without_psy_aggregated %>%
+  filter(time_p <= history_end_date &
+         time_p >= history_start_date ) %>%
+  ggplot( aes(x=time_p, y=perc_by_country, 
+              group = country, color = country))+
+  geom_point( aes(shape=country, color=country) )+
+    geom_line() +
+  geom_vline(xintercept = as.Date(pandemic_start_date),
+             linetype = "dashed") +
+  scale_colour_viridis_d() +
+    theme(
+      legend.position="bottom",
+      plot.title = element_text(size=11)
+    ) +
+  scale_x_date(date_labels = "%b-%d-%Y", breaks = "month") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6)) +
+    ggtitle(paste0("Percentage: patients with psychiatric conditions / total number of patients ( per ", time_period,") \n (aggregated by country)")) +
+    xlab("period (months)") +
+    ylab("percentage (%)")
 ```
@@
 
+#### Showing each site per country 
 ```{r}
@@
        title = paste0("Ratio patient with vs without psychiatric conditions ( per ", time_period,")"))
+```
+
+#### Aggregating the values of the sites per country
+
+```{r}
+ratios_patients_with_without_psy_aggregated %>%
+  filter(time_p <= history_end_date &
+         time_p >= history_start_date ) %>%
+  ggplot( aes(x=time_p, y=ratio_by_country, 
+              group = country, color = country))+
+  geom_point( aes(shape=country, color=country) )+
+    geom_line() +
+  geom_vline(xintercept = as.Date(pandemic_start_date),
+             linetype = "dashed") +
+  scale_colour_viridis_d() +
+    theme(
+      legend.position="bottom",
+      plot.title = element_text(size=11)
+    ) +
+  scale_x_date(date_labels = "%b-%d-%Y", breaks = "month") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6)) +
+    ggtitle("Ratios: patients with psychiatric conditions / without psychiatric conditions \n (aggregated by country)") +
+    xlab("period (months)") +
+    ylab("ratio")
 ```

### 2021-11-17 metanalysis_visualization/visualization.Rmd
added visualization for disorder group and sex ratio

@@
 ```
+
+## Percentage of patients by disorder group
+
+```{r}
+perc_disorder_group_all %>%
+    filter(time_p <= history_end_date &
+         time_p >= history_start_date &
+           disorder_group %in% c("Anxiety Disorders", "Depressive Disorders", "Suicide or Self-Injury" )) %>%
+  ggplot(aes(x = time_p, y = percentage_dg, fill = siteid, color = siteid)) +
+  geom_point( aes(shape=siteid, color=siteid)  ) +
+  geom_line() +
+  facet_grid(. ~ disorder_group) +
+  geom_vline(xintercept = as.Date(pandemic_start_date),
+             linetype = "dashed") +
+  scale_fill_manual(values = cbPalette) +
+  scale_color_manual(values = cbPalette) +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6)) +
+  scale_x_date(date_labels = "%b-%d-%Y", breaks = "month", limits = c( history_start_date, history_end_date  )) +
+  labs(y = "Percentage of patients (%)",
+       x = paste0("Date (by ", time_period,")"),
+       title = paste0("Percentage of patients by disorder group \n ( per ", time_period,")"))
+
+perc_disorder_group_all %>%
+    filter(time_p <= history_end_date &
+         time_p >= history_start_date &
+          #disorder_group %in% c("Anxiety Disorders", "Depressive Disorders", "Suicide or Self-Injury" )) %>%
+          disorder_group %in% c("Suicide or Self-Injury" )) %>%
+  ggplot(aes(x = time_p, y = percentage_dg, fill = siteid, color = siteid)) +
+  geom_point( aes(shape=siteid, color=siteid)  ) +
+  geom_line() +
+  geom_vline(xintercept = as.Date(pandemic_start_date),
+             linetype = "dashed") +
+  scale_fill_manual(values = cbPalette) +
+  scale_color_manual(values = cbPalette) +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6)) +
+  scale_x_date(date_labels = "%b-%d-%Y", breaks = "month", limits = c( history_start_date, history_end_date  )) +
+  labs(y = "Percentage of patients (%)",
+       x = paste0("Date (by ", time_period,")"),
+       title = paste0("Percentage of patients with Suicide or Self-Injury \n ( per ", time_period,")"))
+```
+
+## Sex ratio per site
+
+```{r}
+sex_ratio_by_site <- patients_hospitalized_agg_sex_perc_all %>%
+  filter( sex == "female") %>%
+  group_by( time_p, siteid ) %>%
+  mutate( female_count = count_sex, 
+          male_count = count - count_sex, 
+          sex_ratio_by_site = female_count / male_count ) %>%
+  select( time_p, female_count, male_count, sex_ratio_by_site, siteid)
+
+sex_ratio_by_site %>%
+    filter(time_p <= history_end_date &
+         time_p >= history_start_date) %>%
+  ggplot(aes(x = time_p, y = sex_ratio_by_site, fill = siteid, color = siteid)) +
+  geom_point( aes(shape=siteid, color=siteid)  ) +
+  geom_line() +
+  geom_vline(xintercept = as.Date(pandemic_start_date),
+             linetype = "dashed") +
+   geom_hline(yintercept = 1,
+             linetype = "dashed", color = "red") +
+  scale_fill_manual(values = cbPalette) +
+  scale_color_manual(values = cbPalette) +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6)) +
+  scale_x_date(date_labels = "%b-%d-%Y", breaks = "month", limits = c( history_start_date, history_end_date  )) +
+  labs(y = "sex ratio (female/male)",
+       x = paste0("Date (by ", time_period,")"),
+       title = paste0("Sex ratio of patients with mental disorders \n ( per ", time_period,")"))
+```

### 2021-11-17 metanalysis_visualization/visualization.Rmd
added visualization for total number of patients per site per month

@@
 ```
+
+### Total number of patients per month an site
+
+```{r}
+patients_hospitalized_agg_sex_perc_all %>%
+  select( time_p, count, siteid) %>%
+  unique() %>%
+    filter(time_p <= history_end_date &
+         time_p >= history_start_date) %>%
+  ggplot(aes(x = time_p, y = count, fill = siteid, color = siteid)) +
+  geom_point( aes(shape=siteid, color=siteid)  ) +
+  geom_line() +
+  geom_vline(xintercept = as.Date(pandemic_start_date),
+             linetype = "dashed") +
+  scale_fill_manual(values = cbPalette) +
+  scale_color_manual(values = cbPalette) +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6)) +
+  scale_x_date(date_labels = "%b-%d-%Y", breaks = "month", limits = c( history_start_date, history_end_date  )) +
+  labs(y = "number of patients",
+       x = paste0("Date (by ", time_period,")"),
+       title = paste0("Total number of patients with mental disorders per site \n ( per ", time_period,")"))
+```
+```

### 2021-11-22 metanalysis_visualization/visualization.Rmd
R code to further analyze the overlap based on ICD codes

--- 2021-11-17 metanalysis_visualization/visualization.Rmd
@@
 ```{r}
-outputDir <- "../../output/"
+outputDir <- "../../output/latest_output/"
 outputFiles <- list.files( path = outputDir, pattern = "\\.RData")
@@
 for( i in 1:length( outputFiles )){
+  print(i)
   load(paste0( outputDir, outputFiles[i]))
@@
     ratios_patients_with_without_psy_all <- ratios_patients_with_without_psy
+    mental_codes_qc_output_all <- mental_codes_qc_output %>%
+      select( disorder_group, description, concept_code, distinct_patients, distinct_observations, siteid )
   }else{
@@
     ratios_patients_with_without_psy_all <- rbind( ratios_patients_with_without_psy_all, ratios_patients_with_without_psy)
+   mental_codes_qc_output <- mental_codes_qc_output %>%
+      select( disorder_group, description, concept_code, distinct_patients, distinct_observations, siteid )
+    mental_codes_qc_output_all <- rbind( mental_codes_qc_output_all, mental_codes_qc_output)
   }
@@
      patient_count_psy_period_psy_clear, table1, bootstrapped_coefficients_df,
-     bootstrapped_fitted_df, length_hospitalisation_values)
+     bootstrapped_fitted_df, mental_codes_qc_output)
 }
@@
 ```
+
+### Summary of the codes used per site 
+```{r}
+library(tidyr)
+summary_codes <- mental_codes_qc_output_all %>%
+  select( disorder_group, description, concept_code, distinct_patients, siteid ) %>%
+  spread( siteid, distinct_patients)
+
+# extract all codes that are available across all sites
+codes_in_all_sites <- summary_codes[ complete.cases( summary_codes ), ]
+codes_in_all_sites <- codes_in_all_sites[
+  with(codes_in_all_sites, order(BCH, FRBDX, Lurie, UMICH, decreasing = TRUE)),
+]
+print( paste0("There are in total ", nrow( codes_in_all_sites), " codes in common across all sites"))
+
+#### upSetR 
+library(devtools)
+install_github("jokergoo/ComplexHeatmap")
+library( ComplexHeatmap)
+BCH_concepts = mental_codes_qc_output_all[ mental_codes_qc_output_all$siteid == "BCH", "concept_code"]
+FRBDX_concepts = mental_codes_qc_output_all[ mental_codes_qc_output_all$siteid == "FRBDX", "concept_code"]
+Lurie_concepts = mental_codes_qc_output_all[ mental_codes_qc_output_all$siteid == "Lurie", "concept_code"]
+UMICH_concepts = mental_codes_qc_output_all[ mental_codes_qc_output_all$siteid == "UMICH", "concept_code"]
+
+concepts = list( 
+          BCH = BCH_concepts$concept_code,
+          FRBDX = FRBDX_concepts$concept_code, 
+          Lurie = Lurie_concepts$concept_code, 
+          UMICH = UMICH_concepts$concept_code
+          )
+
+upset(fromList(concepts), order.by = "freq")
+
+all_other_sites <- unique( c( FRBDX_concepts$concept_code, Lurie$concept_code, UMICH$concept_code ))
+concepts_in_bch_not_in_other_sites <- BCH_concepts[! BCH_concepts$concept_code %in% all_other_sites, ]
+codesToReview <- mental_codes_qc_output_all[ mental_codes_qc_output_all$concept_code %in% concepts_in_bch_not_in_other_sites$concept_code, ]
+
+write.table( codesToReview, file="icdCodesOnlyInBCH.txt", col.names = TRUE, row.names = FALSE, sep = "\t", quote = FALSE )
+write.table( codes_in_all_sites, file="icdCodesPresentInAllSites.txt", col.names = TRUE, row.names = FALSE, sep = "\t", quote = FALSE )
+```

### 2021-11-23 metanalysis_visualization/visualization.Rmd
pvalues bootstrap calculation

--- 2021-11-17 metanalysis_visualization/visualization.Rmd
@@
 ```{r}
-outputDir <- "../../output/"
+outputDir <- "../output/"
 outputFiles <- list.files( path = outputDir, pattern = "\\.RData")
 country_map <- read.delim( file = paste0( outputDir, "country_mapping.txt"))
-getwd()
 ```
@@
   if( i == 1 ){
+  print(i)
     count_icd_all <- count_icd
@@
     ratios_patients_with_without_psy_all <- ratios_patients_with_without_psy
+    bootstrapped_fitted_all <- bootstrapped_fitted_df
+    bootstrapped_coefficients_all <- bootstrapped_coefficients_df
   }else{
@@
     ratios_patients_with_without_psy_all <- rbind( ratios_patients_with_without_psy_all, ratios_patients_with_without_psy)
+    bootstrapped_fitted_all <-  rbind( bootstrapped_fitted_all, bootstrapped_fitted_df)
+    bootstrapped_coefficients_all <-  rbind( bootstrapped_coefficients_all, bootstrapped_coefficients_df)
   }
@@
 ```
+
+### Total number of patients per month an site
+
+```{r}
+patients_hospitalized_agg_sex_perc_all %>%
+  select( time_p, count, siteid) %>%
+  unique() %>%
+    filter(time_p <= history_end_date &
+         time_p >= history_start_date) %>%
+  ggplot(aes(x = time_p, y = count, fill = siteid, color = siteid)) +
+  geom_point( aes(shape=siteid, color=siteid)  ) +
+  geom_line() +
+  geom_vline(xintercept = as.Date(pandemic_start_date),
+             linetype = "dashed") +
+  scale_fill_manual(values = cbPalette) +
+  scale_color_manual(values = cbPalette) +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6)) +
+  scale_x_date(date_labels = "%b-%d-%Y", breaks = "month", limits = c( history_start_date, history_end_date  )) +
+  labs(y = "number of patients",
+       x = paste0("Date (by ", time_period,")"),
+       title = paste0("Total number of patients with mental disorders per site \n ( per ", time_period,")"))
+```
+
+## Boostrap vizualisation
+```{r}
+# bootstrap_size <- 1000
+# binom_proportion <- function(n, prob, size) {
+#   (rbinom(n, size, prob) / size) * 100
+# }
+# 
+# # View(patient_count_psy_period_psy_all)
+# 
+# bootstrap_function <- function(df, ...) {
+#   bootstrapped_timepoints <- mapply(binom_proportion,
+#          n = bootstrap_size,
+#          size = df$count_no_psy_patients +
+#            df$count_psy_patients,
+#          prob = df$percentage/100)
+#   
+#   bootstrapped_sample_list <- vector(mode = "list", length = bootstrap_size)
+#   bootstrapped_coefficients_list <- vector(mode = "list", length = bootstrap_size)
+#   for (iteration in 1:bootstrap_size) {
+#   df_iteration <- data.frame(period = df$period,
+#                              time = df$time,
+#                              percentage = bootstrapped_timepoints[iteration, ])
+#   ts_linear_percent <- lm(percentage ~ period * time, df_iteration)
+#   linear_model <- broom::tidy(ts_linear_percent) %>%
+#     filter(term %in% c("time", "periodduring_pandemic:time"))
+#   coefficients <- linear_model$estimate
+#   names(coefficients) <- linear_model$term
+#   bootstrapped_coefficients_list[[iteration]] <- coefficients
+#   df_iteration$predicted <- predict(ts_linear_percent)
+#   df_iteration$time_p <- df$time_p
+#   bootstrapped_sample_list[[iteration]] <- df_iteration
+#   }
+#   bootstrapped_fitted_df <- bind_rows(bootstrapped_sample_list, .id = "iteration")
+#   bootstrapped_coefficients_df <- bind_rows(bootstrapped_coefficients_list, .id = "iteration")
+#   list("bootstrapped_fitted_df" = bootstrapped_fitted_df,
+#        "bootstrapped_coefficients_df" = bootstrapped_coefficients_df)
+# }
+# 
+# 
+# bootstrapped_all <- patient_count_psy_period_psy_all %>% 
+#   group_by(siteid) %>% 
+#   group_map(bootstrap_function) %>% 
+#   set_names(nm = names_site_id)
+# 
+# bootstrapped_fitted_all <- purrr::map(bootstrapped_all, "bootstrapped_fitted_df") %>% 
+#   bind_rows(.id = "siteid")
+# bootstrapped_coefficients_all <- purrr::map(bootstrapped_all, "bootstrapped_coefficients_df") %>% 
+#   bind_rows(.id = "siteid")
+
+names_site_id <- unique(patient_count_psy_period_psy_all$siteid)
+
+quantile_bootstrapped_fitted_all <- bootstrapped_fitted_all %>%
+  group_by(time_p, period, siteid) %>%
+  summarise("2.5%" = quantile(predicted, 0.025),
+            "50%" = quantile(predicted, 0.5),
+            "97.5%" = quantile(predicted, 0.975)) %>%
+  pivot_longer(cols = c("2.5%", "50%", "97.5%"),
+               names_to = "quantiles")
+
+annotation_function <- function(df, ...) {
+  pvalue_coeff_interaction <- round(
+    (1- max(
+      sum(df[["periodduring_pandemic:time"]] > 0,
+          df[["periodduring_pandemic:time"]] < 0)
+    )/nrow(df)) * 2
+    , 3) 
+  paste0("Coeffs differences: ",
+                      round(mean(df[["periodduring_pandemic:time"]]), 2),
+                      " (pvalue: ",
+                      pvalue_coeff_interaction,
+                      ")")
+}
+
+plot_annotations <- bootstrapped_coefficients_all %>% 
+  group_by(siteid) %>% 
+  group_map( annotation_function ) %>% 
+  set_names(names_site_id) %>% 
+  unlist()
+plot_annotations_df <- data.frame(
+  siteid = names(plot_annotations), 
+  label = plot_annotations
+)
+
+ggplot(bootstrapped_fitted_all, aes(x = time_p, y = predicted, group = interaction(iteration, period))) +
+  facet_wrap("siteid", scales = "free_y") +
+  geom_line(aes(color = period),
+            alpha = 0.03) +
+  geom_line(data = quantile_bootstrapped_fitted_all,
+            aes(y = value, group = interaction(quantiles, period)),
+            linetype = "dashed") +
+  geom_line(data = quantile_bootstrapped_fitted_all[quantile_bootstrapped_fitted_all$quantiles == "50%", ],
+            aes(y = value, group = period)) +
+  geom_point(data = patient_count_psy_period_psy_all,
+             aes(y = percentage, group=NULL),
+             alpha = 0.5) +
+  scale_colour_manual(values = cbPalette) +
+  guides(color = guide_legend(override.aes = list(alpha = 1))) +
+  geom_label(data = plot_annotations_df, 
+             aes(label=label), 
+            x = -Inf, y = +Inf, hjust=0, vjust=1,
+            size = 4,
+            inherit.aes = FALSE) +
+  labs(title = "Interrupted Time Series Analysis: Percentage of patients with psychiatric conditions",
+       subtitle = "Parametric bootstrap estimation: (1000 iterations)",
+       x = "Calendar dates",
+       y = "Percentage",
+       color = "Period") +
+  theme(strip.text = element_text(size = 10),
+        axis.text.x = element_text(size = 8), 
+        axis.text.y = element_text(size = 8), 
+        axis.title.x = element_text(size = 8),
+        axis.title.y = element_text(size = 8),
+        title = element_text(size = 14),
+        plot.subtitle = element_text(size = 12, face = "italic")) 
+ggsave("boostrapped_all.png", height = 20, width = 32, units = "cm")
+```

### 2021-11-23 psy_peds_3.Rmd
pvalues bootstrap calculation

@@
 mean(bootstrapped_coefficients_df$time + bootstrapped_coefficients_df[["periodduring_pandemic:time"]])
-pvalue_coeff_interaction <- round(t.test(bootstrapped_coefficients_df[["periodduring_pandemic:time"]])$p.value, 4)
-plot_annotation <- paste0("Diff. in slope coeffs before and after pandemic: ",
+
+pvalue_coeff_interaction <- round(
+  (1- max(
+    sum(bootstrapped_coefficients_df[["periodduring_pandemic:time"]] > 0,
+        bootstrapped_coefficients_df[["periodduring_pandemic:time"]] < 0)
+  )/nrow(df)) * 2
+  , 3)
+
+plot_annotation <- paste0("Coeffs differences: ",
        round(mean(bootstrapped_coefficients_df[["periodduring_pandemic:time"]]), 2),

### 2021-11-23 metanalysis_visualization/visualization.Rmd

--- 2021-11-22 metanalysis_visualization/visualization.Rmd
@@
 ```{r}
-outputDir <- "../../output/latest_output/"
+outputDir <- "../output/latest_output/"
 outputFiles <- list.files( path = outputDir, pattern = "\\.RData")
 country_map <- read.delim( file = paste0( outputDir, "country_mapping.txt"))
-getwd()
 ```
@@
 
+## Boostrap vizualisation
+```{r}
+names_site_id <- unique(patient_count_psy_period_psy_all$siteid)
+
+quantile_bootstrapped_fitted_all <- bootstrapped_fitted_all %>%
+  group_by(time_p, period, siteid) %>%
+  summarise("2.5%" = quantile(predicted, 0.025),
+            "50%" = quantile(predicted, 0.5),
+            "97.5%" = quantile(predicted, 0.975)) %>%
+  pivot_longer(cols = c("2.5%", "50%", "97.5%"),
+               names_to = "quantiles")
+
+annotation_function <- function(df, ...) {
+  pvalue_coeff_interaction <- round(
+    (1- max(
+      sum(df[["periodduring_pandemic:time"]] > 0),
+      sum(df[["periodduring_pandemic:time"]] < 0)
+    )/nrow(df)) * 2
+    , 3) 
+  paste0("Coeffs differences: ",
+                      round(mean(df[["periodduring_pandemic:time"]]), 2),
+                      " (pvalue: ",
+                      pvalue_coeff_interaction,
+                      ")")
+}
+
+plot_annotations <- bootstrapped_coefficients_all %>% 
+  group_by(siteid) %>% 
+  group_map( annotation_function ) %>% 
+  set_names(names_site_id) %>% 
+  unlist()
+plot_annotations_df <- data.frame(
+  siteid = names(plot_annotations), 
+  label = plot_annotations
+)
+
+ggplot(bootstrapped_fitted_all, aes(x = time_p, y = predicted, group = interaction(iteration, period))) +
+  facet_wrap("siteid", scales = "free_y") +
+  geom_line(aes(color = period),
+            alpha = 0.03) +
+  geom_line(data = quantile_bootstrapped_fitted_all,
+            aes(y = value, group = interaction(quantiles, period)),
+            linetype = "dashed") +
+  geom_line(data = quantile_bootstrapped_fitted_all[quantile_bootstrapped_fitted_all$quantiles == "50%", ],
+            aes(y = value, group = period)) +
+  geom_point(data = patient_count_psy_period_psy_all,
+             aes(y = percentage, group=NULL),
+             alpha = 0.5) +
+  scale_colour_manual(values = cbPalette) +
+  guides(color = guide_legend(override.aes = list(alpha = 1))) +
+  geom_label(data = plot_annotations_df, 
+             aes(label=label), 
+            x = -Inf, y = +Inf, hjust=0, vjust=1,
+            size = 4,
+            inherit.aes = FALSE) +
+  labs(title = "Interrupted Time Series Analysis: Percentage of patients with psychiatric conditions",
+       subtitle = "Parametric bootstrap estimation: (1000 iterations)",
+       x = "Calendar dates",
+       y = "Percentage",
+       color = "Period") +
+  theme(strip.text = element_text(size = 10),
+        axis.text.x = element_text(size = 8), 
+        axis.text.y = element_text(size = 8), 
+        axis.title.x = element_text(size = 8),
+        axis.title.y = element_text(size = 8),
+        title = element_text(size = 14),
+        plot.subtitle = element_text(size = 12, face = "italic")) 
+ggsave("boostrapped_all.png", height = 20, width = 32, units = "cm")
+```
+
 ## Meta-analysis of temporal trends

### 2021-11-23 psy_peds_3.Rmd
add bootstrap pvalues

@@
 
+getwd()
 files <- readInputFiles( path      = folder_4ce_files, 
@@
   (1- max(
-    sum(bootstrapped_coefficients_df[["periodduring_pandemic:time"]] > 0,
-        bootstrapped_coefficients_df[["periodduring_pandemic:time"]] < 0)
-  )/nrow(df)) * 2
+    sum(bootstrapped_coefficients_df[["periodduring_pandemic:time"]] > 0),
+        sum(bootstrapped_coefficients_df[["periodduring_pandemic:time"]] < 0)
+  )/nrow(bootstrapped_coefficients_df)) * 2
   , 3)

### 2021-11-30 metanalysis_visualization/visualization.Rmd
adding lines to read bootstrapped_coefficients_all

@@
     ratios_patients_with_without_psy_all <- ratios_patients_with_without_psy
+    bootstrapped_fitted_all <- bootstrapped_fitted_df
+    bootstrapped_coefficients_all <- bootstrapped_coefficients_df
     mental_codes_qc_output_all <- mental_codes_qc_output %>%
       select( disorder_group, description, concept_code, distinct_patients, distinct_observations, siteid )
-  }else{
+  } else {
     count_icd_all <- rbind(count_icd_all, count_icd)
@@
     ratios_patients_with_without_psy_all <- rbind( ratios_patients_with_without_psy_all, ratios_patients_with_without_psy)
+    bootstrapped_fitted_all <- rbind(bootstrapped_fitted_all, bootstrapped_fitted_df)
+    bootstrapped_coefficients_all <- rbind(bootstrapped_coefficients_all, bootstrapped_coefficients_df)
    mental_codes_qc_output <- mental_codes_qc_output %>%
@@
 library(tidyr)
+
 summary_codes <- mental_codes_qc_output_all %>%

### 2021-11-30 metanalysis_visualization/visualization.Rmd
qc codes adapted for one more site

@@
 }
-listOfPackages <- c("tidyverse", "RColorBrewer", "knitr", "kableExtra", "tsModel", "gridExtra", "dplyr", "metafor", "meta", "viridis")
+listOfPackages <- c("tidyverse", "tidyr","devtools","RColorBrewer", "knitr", "kableExtra", "tsModel", "gridExtra", "dplyr", "metafor", "meta", "viridis", "UpSetR")
 paket(listOfPackages)
+
+install_github("jokergoo/ComplexHeatmap")
+library( ComplexHeatmap)
 ```
@@
 ```{r}
-outputDir <- "../output/latest_output/"
+outputDir <- "../../output/latest_output/"
 outputFiles <- list.files( path = outputDir, pattern = "\\.RData")
@@
 ```{r}
-library(tidyr)
+mental_codes_qc_output_all <- mental_codes_qc_output_all %>%
+    mutate( concept_code = gsub("[.]", "", concept_code ) )
 
@@
 codes_in_all_sites <- codes_in_all_sites[
-  with(codes_in_all_sites, order(BCH, FRBDX, Lurie, UMICH, decreasing = TRUE)),
+  with(codes_in_all_sites, order(BCH, FRBDX, Lurie, UMICH, APHP, decreasing = TRUE)),
 ]
@@
 #### upSetR
-library(devtools)
-install_github("jokergoo/ComplexHeatmap")
-library( ComplexHeatmap)
+
 BCH_concepts = mental_codes_qc_output_all[ mental_codes_qc_output_all$siteid == "BCH", "concept_code"]
@@
 UMICH_concepts = mental_codes_qc_output_all[ mental_codes_qc_output_all$siteid == "UMICH", "concept_code"]
+APHP_concepts = mental_codes_qc_output_all[ mental_codes_qc_output_all$siteid == "APHP", "concept_code"]
 
@@
           Lurie = Lurie_concepts$concept_code, 
-          UMICH = UMICH_concepts$concept_code
+          UMICH = UMICH_concepts$concept_code, 
+          APHP  = APHP_concepts$concept_code
           )
@@
 
-all_other_sites <- unique( c( FRBDX_concepts$concept_code, Lurie$concept_code, UMICH$concept_code ))
+all_other_sites <- unique( c( FRBDX_concepts$concept_code, APHP_concepts$concept_code,
+                              Lurie_concepts$concept_code, UMICH_concepts$concept_code ))
 concepts_in_bch_not_in_other_sites <- BCH_concepts[! BCH_concepts$concept_code %in% all_other_sites, ]

### 2021-11-30 metanalysis_visualization/visualization.Rmd
dates correction

--- 2021-11-30 metanalysis_visualization/visualization.Rmd
@@
 trends_siteid <- meta_df %>% 
+  filter((time_p >= as.Date("2019-02-01")) & (time_p <= as.Date("2021-05-01"))) %>% 
+  filter(siteid != "APHP") %>%
   group_by(siteid) %>% 
@@
   group_by(country) %>% 
+  filter((time_p >= as.Date("2019-02-01")) & (time_p <= as.Date("2021-05-01"))) %>% 
   group_modify( ~ broom::tidy(lm(as.formula(trend_model), data = .x))) %>%

### 2021-11-30 metanalysis_visualization/visualization.Rmd

--- 2021-11-30 metanalysis_visualization/visualization.Rmd
@@
 trends_siteid <- meta_df %>% 
+  filter((time_p >= as.Date("2019-02-01")) & (time_p <= as.Date("2021-05-01"))) %>% 
+  filter(siteid != "APHP") %>% 
   group_by(siteid) %>% 
@@
 trends_country <- meta_df %>% 
+  filter((time_p >= as.Date("2019-02-01")) & (time_p <= as.Date("2021-05-01"))) %>% 
   group_by(country) %>% 

### 2021-11-30 metanalysis_visualization/visualization.Rmd
ICD QC code updated adding upset for each disorder group and excel sheet with summary

@@
 
-# extract all codes that are available across all sites
-codes_in_all_sites <- summary_codes[ complete.cases( summary_codes ), ]
-codes_in_all_sites <- codes_in_all_sites[
-  with(codes_in_all_sites, order(BCH, FRBDX, Lurie, UMICH, APHP, decreasing = TRUE)),
+## add a column with the sites combination 
+getSitesWithEachCode <- function(row) {
+  sites <- c( row['BCH'],  row['Lurie'], row['UMICH'], row['FRBDX'], row['APHP'])
+  sites <- sites[!is.na( sites ) ]
+  siteVal <- paste0( names( sites ), collapse = " - ")
+  return( siteVal)
+}
+summary_codes <- summary_codes[, c("disorder_group", "description", "concept_code", 
+                                   "BCH", "Lurie", "UMICH",  "FRBDX", "APHP") ]                        
+summary_codes$sitesCombination <- apply(summary_codes, 1, FUN = getSitesWithEachCode)
+summary_codes <- summary_codes[
+  with(summary_codes, order(BCH, FRBDX, Lurie, UMICH, APHP, decreasing = TRUE)),
 ]
-print( paste0("There are in total ", nrow( codes_in_all_sites), " codes in common across all sites"))
+print( paste0("There are in total ", nrow( summary_codes[ summary_codes$sitesCombination == "BCH - Lurie - UMICH - FRBDX - APHP", ]), " codes in common across all sites"))
+
+write.table( summary_codes, file="summary_codes.txt", col.names = TRUE, row.names = FALSE, sep = "\t", quote = FALSE )
 
 #### upSetR 
+upsetForDisorderGroup <- function( df, disorderGroup = "ALL"){
+  if( disorderGroup != "ALL"){
+     df <- df %>%
+       filter( disorder_group == disorderGroup ) 
+  }
 
-BCH_concepts = mental_codes_qc_output_all[ mental_codes_qc_output_all$siteid == "BCH", "concept_code"]
-FRBDX_concepts = mental_codes_qc_output_all[ mental_codes_qc_output_all$siteid == "FRBDX", "concept_code"]
-Lurie_concepts = mental_codes_qc_output_all[ mental_codes_qc_output_all$siteid == "Lurie", "concept_code"]
-UMICH_concepts = mental_codes_qc_output_all[ mental_codes_qc_output_all$siteid == "UMICH", "concept_code"]
-APHP_concepts = mental_codes_qc_output_all[ mental_codes_qc_output_all$siteid == "APHP", "concept_code"]
+  BCH_concepts = df[ df$siteid == "BCH", "concept_code"]
+  FRBDX_concepts = df[ df$siteid == "FRBDX", "concept_code"]
+  Lurie_concepts = df[ df$siteid == "Lurie", "concept_code"]
+  UMICH_concepts = df[ df$siteid == "UMICH", "concept_code"]
+  APHP_concepts = df[ df$siteid == "APHP", "concept_code"]
 
@@
           )
+  output <- upset(fromList(concepts), order.by = "freq")
+  return( output )
+}
 
-upset(fromList(concepts), order.by = "freq")
+upsetForDisorderGroup( df = mental_codes_qc_output_all, disorderGroup = "ALL")
+```
 
-all_other_sites <- unique( c( FRBDX_concepts$concept_code, APHP_concepts$concept_code,
-                              Lurie_concepts$concept_code, UMICH_concepts$concept_code ))
-concepts_in_bch_not_in_other_sites <- BCH_concepts[! BCH_concepts$concept_code %in% all_other_sites, ]
-codesToReview <- mental_codes_qc_output_all[ mental_codes_qc_output_all$concept_code %in% concepts_in_bch_not_in_other_sites$concept_code, ]
-
-write.table( codesToReview, file="icdCodesOnlyInBCH.txt", col.names = TRUE, row.names = FALSE, sep = "\t", quote = FALSE )
-write.table( codes_in_all_sites, file="icdCodesPresentInAllSites.txt", col.names = TRUE, row.names = FALSE, sep = "\t", quote = FALSE )
+### UpSet R for each disorder group
+```{r}
+disorders <- unique( mental_codes_qc_output_all$disorder_group)
+for( i in 1:length(disorders) ){
+  print( disorders[i] )
+  upsetForDisorderGroup( df = mental_codes_qc_output_all, disorderGroup = disorders[i])
+}
 ```

### 2021-12-02 metanalysis_visualization/visualization.Rmd
visualization code updated: new plot for apendix for the psychiatric ICD codes dictionary QC

@@
 mental_codes_qc_output_all <- mental_codes_qc_output_all %>%
-    mutate( concept_code = gsub("[.]", "", concept_code ) )
+    mutate( concept_code = gsub("[.]", "", concept_code ) ) %>%
+  unique()
 
@@
 
-### UpSet R for each disorder group
+### Estimate number of ICD codes per site and disorder group
+Represent this in 16 pannels, if we put all together, due to the different scales we will not be able to really see the differences between results. 
 ```{r}
-disorders <- unique( mental_codes_qc_output_all$disorder_group)
-for( i in 1:length(disorders) ){
-  print( disorders[i] )
-  upsetForDisorderGroup( df = mental_codes_qc_output_all, disorderGroup = disorders[i])
+numberOfCodesPerDisorderGroup <- mental_codes_qc_output_all %>%
+  group_by( disorder_group, siteid ) %>%
+  mutate( icd_codes_count = n( )) %>%
+  select( disorder_group, siteid, icd_codes_count ) %>%
+  unique()
+
+swr = function(string, nwrap=30) {
+  paste(strwrap(string, width=nwrap), collapse="\n")
 }
+swr = Vectorize(swr)
+
+numberOfCodesPerDisorderGroup %>%
+  mutate( disorder_group = swr( disorder_group ) ) %>%
+  ggplot( aes( x= reorder( siteid, icd_codes_count), y = icd_codes_count)) +
+  geom_bar( stat = "identity") +
+  facet_wrap( ~ disorder_group, scales = "free" ) +
+  labs(x ="site identifier", y = "ICD dictionary (number of ICD codes)") +
+  theme(axis.text.x=element_text(angle=45,hjust=1, size = 6), 
+        strip.text = element_text(size = 7), 
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) 
 ```

### 2021-12-08 metanalysis_visualization/visualization.Rmd
fixing pvalues and meta analysis

--- 2021-11-30 metanalysis_visualization/visualization.Rmd
@@
 ```{r}
-outputDir <- "../../output/latest_output/"
+outputDir <- "../output/latest_output/"
 outputFiles <- list.files( path = outputDir, pattern = "\\.RData")
@@
   summarise("2.5%" = quantile(predicted, 0.025),
-            "50%" = quantile(predicted, 0.5),
+            "50%" = median(predicted, 0.5),
             "97.5%" = quantile(predicted, 0.975)) %>%
@@
 annotation_function <- function(df, ...) {
-  pvalue_coeff_interaction <- round(
-    (1- max(
-      sum(df[["periodduring_pandemic:time"]] > 0),
-      sum(df[["periodduring_pandemic:time"]] < 0)
-    )/nrow(df)) * 2
-    , 3) 
+  # pvalue_coeff_interaction <- round(
+  #   (1- max(
+  #     sum(df[["periodduring_pandemic:time"]] > 0),
+  #     sum(df[["periodduring_pandemic:time"]] < 0)
+  #   )/nrow(df)) * 2
+  #   , 3)
+
+  z <-  mean(df[["periodduring_pandemic:time"]]) / sd(df[["periodduring_pandemic:time"]])
+
+  pvalue_coeff_interaction = round(pnorm(abs(z), lower.tail = FALSE) * 2, 3)
   paste0("Coeffs differences: ",
@@
 ```{r}
-trend_model <- "percentage ~ period * time"
+trends_siteid <- bootstrapped_coefficients_all %>% 
+  group_by(siteid) %>% 
+  summarise(mean_boot = mean(`periodduring_pandemic:time`),
+         var_boot = var(`periodduring_pandemic:time`), 
+         z = mean_boot / sqrt(var_boot), 
+         pvalue = pnorm(z, lower.tail = FALSE)*2 )
 ```
@@
 ```{r}
-trends_siteid <- meta_df %>% 
-  filter((time_p >= as.Date("2019-02-01")) & (time_p <= as.Date("2021-05-01"))) %>% 
-  filter(siteid != "APHP") %>% 
-  group_by(siteid) %>% 
-  group_modify( ~ broom::tidy(lm(as.formula(trend_model), data = .x))) %>%
-  filter(term == c("periodduring_pandemic:time")) %>% 
-  mutate(variance = std.error * std.error) %>% 
-  select(siteid, estimate, variance)
+# Fixed effects
+model_meta_siteid <- metafor::rma(yi = mean_boot, 
+             vi = var_boot,
+             measure = "MN",
+             method = "FE",
+  data = trends_siteid) 
+metafor::forest(model_meta_siteid,
+                slab = trends_siteid$siteid,
+                header="Sites",
+                mlab="Overall temporal trend")
+```
 
-model_meta_siteid <- metafor::rma(yi = estimate, 
-             vi = variance,
-             measure = "GEN",
+```{r}
+# Random effects
+model_meta_siteid <- metafor::rma(yi = mean_boot, 
+                                  vi = var_boot,
+                                  measure = "MN",
                                   method = "DL",
@@
 ```{r}
-trends_country <- meta_df %>% 
-  filter((time_p >= as.Date("2019-02-01")) & (time_p <= as.Date("2021-05-01"))) %>% 
+trends_country <- bootstrapped_coefficients_all %>% 
+  left_join(country_map, by = "siteid") %>% 
   group_by(country) %>% 
-  group_modify( ~ broom::tidy(lm(as.formula(trend_model), data = .x))) %>%
-  filter(term == c("periodduring_pandemic:time")) %>% 
-  mutate(variance = std.error * std.error) %>% 
-  select(country, estimate, variance)
+  summarise(mean_boot = mean(`periodduring_pandemic:time`),
+         var_boot = var(`periodduring_pandemic:time`), 
+         z = mean_boot / sqrt(var_boot), 
+         pvalue = pnorm(z, lower.tail = FALSE)*2 )
 
-model_meta_country <- metafor::rma(yi = estimate, 
-             vi = variance,
-             measure = "GEN",
-             method = "DL",
+model_meta_country <- metafor::rma(yi = mean_boot, 
+             vi = var_boot,
+             measure = "MN",
+             method = "FE",
   data = trends_country) 

### 2021-12-08 psy_peds_3.Rmd
dates correction
fixing pvalues and meta analysis

@@
 
-start_date_plots <- as.Date("2019-01-01")
-end_date_plots <- as.Date("2021-05-24")
-
 nh <- admission_plot_count(df = new_hospitalizations, 
@@
   ungroup()
-
-start_date_plots <- as.Date("2019-01-01")
-end_date_plots <- as.Date("2021-05-24")
 
@@
 
-bootstrapped_coefficients_df <- bind_rows(bootstrapped_coefficients_list, .id = "iteration")
-mean(bootstrapped_coefficients_df$time)
-mean(bootstrapped_coefficients_df$time + bootstrapped_coefficients_df[["periodduring_pandemic:time"]])
-
-pvalue_coeff_interaction <- round(
-  (1- max(
-    sum(bootstrapped_coefficients_df[["periodduring_pandemic:time"]] > 0),
-        sum(bootstrapped_coefficients_df[["periodduring_pandemic:time"]] < 0)
-  )/nrow(bootstrapped_coefficients_df)) * 2
-  , 3)
+bootstrapped_coefficients_df <- bind_rows(bootstrapped_coefficients_list,
+                                          .id = "iteration")
+
+z <-  mean(bootstrapped_coefficients_df[["periodduring_pandemic:time"]]) / sd(bootstrapped_coefficients_df[["periodduring_pandemic:time"]])
+
+pvalue_coeff_interaction <- round(pnorm(abs(z), lower.tail = FALSE) * 2, 3)
 

### 2021-12-08 metanalysis_visualization/visualization.Rmd
update pvalue computation

@@
 ```{r}
+# Recovering
+coefficient_slope_calculation <- function(df, ...) {
+  ts_linear_percent <- lm(percentage ~ period * time, df)
+  linear_model <- broom::tidy(ts_linear_percent) %>%
+    filter(term %in% c("periodduring_pandemic:time"))
+}
+
+coefficients_df <- patient_count_psy_period_psy_all %>% 
+  group_by(siteid) %>% 
+  group_modify(coefficient_slope_calculation) %>% 
+  rename(observed_diff = estimate)
+```
+
+```{r}
 names_site_id <- unique(patient_count_psy_period_psy_all$siteid)
@@
 annotation_function <- function(df, ...) {
-  # pvalue_coeff_interaction <- round(
-  #   (1- max(
-  #     sum(df[["periodduring_pandemic:time"]] > 0),
-  #     sum(df[["periodduring_pandemic:time"]] < 0)
-  #   )/nrow(df)) * 2
-  #   , 3)
-
-  z <-  mean(df[["periodduring_pandemic:time"]]) / sd(df[["periodduring_pandemic:time"]])
-
+  z <- unique(df[["observed_diff"]]) / sd(df[["periodduring_pandemic:time"]])
   pvalue_coeff_interaction = round(pnorm(abs(z), lower.tail = FALSE) * 2, 3)
   paste0("Coeffs differences: ",
-         round(mean(df[["periodduring_pandemic:time"]]), 2),
+         round(unique(df[["observed_diff"]]), 2),
          " (pvalue: ",
@@
 
-plot_annotations <- bootstrapped_coefficients_all %>% 
+df_coefficients_boot <- bootstrapped_coefficients_all %>% 
+  left_join(coefficients_df, by = "siteid") %>% 
+  filter(siteid != "APHP")
+
+plot_annotations <- df_coefficients_boot %>% 
   group_by(siteid) %>% 
@@
 ## Estimate sampling variances
-meta_df <- ratios_patients_with_without_psy_all_country
-meta_df$time <- as.numeric(as.factor(meta_df$time_p))
 ```
@@
 ```{r}
-trends_siteid <- bootstrapped_coefficients_all %>% 
+  trends_siteid <- df_coefficients_boot %>% 
   group_by(siteid) %>% 
   summarise(mean_boot = mean(`periodduring_pandemic:time`),
+            observed = unique(observed_diff),
          var_boot = var(`periodduring_pandemic:time`), 
@@
 ```{r}
-# Fixed effects
-model_meta_siteid <- metafor::rma(yi = mean_boot, 
-             vi = var_boot,
-             measure = "MN",
-             method = "FE",
-  data = trends_siteid) 
-metafor::forest(model_meta_siteid,
-                slab = trends_siteid$siteid,
-                header="Sites",
-                mlab="Overall temporal trend")
-```
-
-```{r}
-# Random effects
-model_meta_siteid <- metafor::rma(yi = mean_boot, 
+# Random effect effects
+model_meta_siteid <- metafor::rma(yi = observed, 
              vi = var_boot,
@@
 ```{r}
-trends_country <- bootstrapped_coefficients_all %>% 
+trends_country <- df_coefficients_boot %>% 
   left_join(country_map, by = "siteid") %>% 

### 2021-12-15 metanalysis_visualization/visualization.Rmd
update pvalue computation
adding the adapted code to incorporate new site results

@@
 ```{r}
+### add CHOP 
+ratios_patients_with_without_psy_chop <- read.delim ( file = "/Users/alba/Desktop/ratios_patients_with_without_psy_chop.txt")
+
+ratios_patients_with_without_psy_all <- rbind( ratios_patients_with_without_psy_all, ratios_patients_with_without_psy_chop)
+
 ratios_patients_with_without_psy_all %>%
@@
 ```{r}
-names_site_id <- unique(patient_count_psy_period_psy_all$siteid)
 
@@
   filter(siteid != "APHP")
+
+names_site_id <- unique(df_coefficients_boot$siteid)
 
@@
 ### Summary of the codes used per site 
-```{r}
+```{r, eval=FALSE}
 mental_codes_qc_output_all <- mental_codes_qc_output_all %>%
@@
 Represent this in 16 pannels, if we put all together, due to the different scales we will not be able to really see the differences between results. 
-```{r}
+```{r, eval=FALSE}
 numberOfCodesPerDisorderGroup <- mental_codes_qc_output_all %>%

### 2021-12-15 psy_peds_3_CHOP.Rmd
adding the adapted code to incorporate new site results

--- 2021-12-08 psy_peds_3.Rmd
@@
 ---
-title: "Hospitalizations for psychiatric conditions among adolescents during the COVID-19 pandemic"
+title: "Hospitalizations for psychiatric conditions among adolescents during the COVID-19 pandemic: adapted to CHOP input format"
 output:
@@
 ```{r message=FALSE, warning=FALSE}
-folder_4ce_files <- "Input/09212021_2/"
-obfuscation =  FALSE
-#obfuscation = 3
 raceAvailable = TRUE
 #raceAvailable = FALSE
-dateFormat <- "%d-%b-%y"
-min_hosp_days <- 2
-data_update_date <- "2021-06-05"
+#data_update_date <- "2021-06-05"
 time_period <-"months" #possible values: days, weeks, months
@@
 
-We will use as input the 2.2 data files. Specifically:
+We will use as input the files from CHOP with their counts.
 
-- LocalPatientSummary
-
-- LocalPatientObservation
-
-- LocalPatientClinicalcourse
-
-For sites recording the race, we will also use an additional file:
-
-- LocalPatientRace
-
-### Reading 4CE phase 2.2 files
+### Reading the files with total counts per month and patients with psychiatric conditions per month.
 
 ```{r message=FALSE, warning=FALSE}
-### Read the CSV input files
-source("R/readInputFiles.R")
+outputDir <- "../output/CHOP/"
+allCounts <- read.csv( file = paste0( outputDir, "All-Counts.csv"))
+psyCounts <- read.csv( file = paste0( outputDir, "Psyc-Counts.csv"))
 
-getwd()
-files <- readInputFiles( path      = folder_4ce_files, 
-                         separator = ",",
-                         skip      = 0, 
-                         verbose   = FALSE )
+allCounts <- allCounts %>%
+  mutate( time_p = ymd(paste( allCounts$yr1, allCounts$mo1, "01", sep="-")), 
+          period = ifelse( time_p < pandemic_start_date,
+                           "before_pandemic",
+                           "during_pandemic"), 
+          siteid = "CHOP") %>%
+  select( time_p, count_all = ct1, period, siteid )
 
-## Create the output folder if it doesn't exist
-#if (! "output" %in% list.dirs()) dir.create("output")
-
-### Extract the patient summary and observation information. 
-demo_raw <- files[["patientSummary"]] %>% filter(cohort == "AllAdm")
-obs_raw <- files[["patientObservations"]] %>% filter(cohort == "AllAdm")
-clinical_raw <- files[["patientClinicalCourse"]] %>% filter(cohort == "AllAdm")
-
-### Read the file containing race information for those sites recording this variable
-if( raceAvailable == TRUE ){
-  race_raw <- read.delim(file.path(folder_4ce_files, "/LocalPatientRace.csv"), sep = ",", skip = 0)
-}
-```
-
-### Read the ICD10 codes file
-
-The file containing the ICD10 psychiatric related codes is located in
-the `public-data` folder of the GitHub repository.
-
-```{r}
-icdCodes <- read.csv("public-data/pediatric_psychiatric_ICD10_codesV2.csv", header = TRUE, colClasses = "character") %>%
-   filter( flag %in% c("originalCodes", "FrenchSpecificCodes")) %>%
-   dplyr::select( disorder_group, ICD10_Code, description )
-```
-
-## Data-management
-
-### Filtering on "inpatient" status
-
-```{r}
-# Filtering on inpatients here (in_hospital == 1), to reduce subsequent compute time of the encounter length calculation 
-clinical_raw <- filter(clinical_raw, in_hospital == 1)
-```
-
-### Calculating duration of individual hospitalizations
-
-Adding columns with hospitalization number and hospitalization length to the clinical raw file
-
-```{r}
-count_sequences_hospitalisation <- function(df, ...) {
-  seq_hospitalisation_df <- data.frame(total_span = seq(min(df$days_since_admission),
-                                                        max(df$days_since_admission))
-  ) %>%
-    left_join(df, by = c("total_span" = "days_since_admission")) %>%
-    replace_na(list(in_hospital = 0))
-  count_sequences <- rle(seq_hospitalisation_df$in_hospital)
-  count_sequences_1 <- lapply(count_sequences, function(x) x[count_sequences$values == 1])
-  n_sequences <- seq_along(count_sequences_1$lengths)
-  sequences <- rep.int(n_sequences, count_sequences_1$lengths)
-  sequences_len <- rep.int(count_sequences_1$lengths, count_sequences_1$lengths)
-  stopifnot(length(df$days_since_admission) == length(sequences))
-  data.frame(days_since_admission = df$days_since_admission,
-             n_hospitalisation = sequences,
-             len_hospitalisation = sequences_len)
-}
-stopifnot(all(clinical_raw$in_hospital == 1))
-hospitalisations_seq_df <- clinical_raw %>%
-  distinct(patient_num, cohort, days_since_admission, in_hospital) %>%
-  group_by(patient_num, cohort) %>%
-  group_modify(count_sequences_hospitalisation)
-
-clinical_raw <- left_join(clinical_raw,
-                          hospitalisations_seq_df,
-                          by = c("patient_num", "cohort", "days_since_admission"))
-```
-
-### Counting ICD10 mental health codes in the obs_raw table
-
-```{r}
-### check that the ICD codes follow the regular expression [A-Z][0-9][0-9AB]\.?[0-9A-TV-Z]{0,4}
-codesToReview <- obs_raw %>% dplyr::filter(concept_type == "DIAG-ICD10",
-                                           ! grepl( "[A-Z][0-9][0-9AB]\\.?[0-9A-TV-Z]{0,4}", concept_code))
-print(codesToReview)
-
-### Calculate how many of the mental codes are in the dataset
-mentalHealthCodes <- unique( obs_raw %>%
-  dplyr::filter( concept_type == "DIAG-ICD10", 
-                 concept_code %in% icdCodes$ICD10_Code ) %>%
-  dplyr::mutate( concept_code = as.character( concept_code ), 
-                 concept_length = nchar( concept_code ), 
-                 concept_check = ifelse( concept_length == 4 | concept_length > 8, "Check", "OK")) %>%
-  dplyr::select( concept_code, concept_length, concept_check) )
-
-summary(as.factor( mentalHealthCodes$concept_check ) )
-
-print( paste0("There are ", length(mentalHealthCodes$concept_code), " ICD10 mental codes out of the total ",length(unique(icdCodes$ICD10_Code)), " in the dataset"))
-```
-
-### Calculating age at hospitalization time
-```{r}
-# Calculating patient age at the time of visit, assuming the age used is the patients' age at the time of extraction (last discharge date available in the data)
-obs_raw_age <- left_join(obs_raw, demo_raw[c("patient_num", "age")], by = c("patient_num")) %>% 
-  inner_join(clinical_raw[, c("patient_num", "days_since_admission", "calendar_date")], 
-            by = c("patient_num", "days_since_admission"))
-obs_raw_age$age_time_diagnosis <- obs_raw_age$age - floor(as.numeric(as.Date(data_update_date) - as.Date(obs_raw_age$calendar_date, dateFormat)) / 365.25)
-```
-
-### Applying inclusion criteria to filter and format the data
-
-We are selecting hospitalizations based on the following criteria:
-
-- Associated with at least 1 ICD code
-
-- With a duration of hospitalization spanning over at least two consecutive days
-
-- Starting on or after 2019/01/01
-
-``` {r}
-obs_filtered_age <- obs_raw_age %>% 
-  filter(concept_type == 'DIAG-ICD10', 
-         days_since_admission >= 0)
-
-clinical_filtered <- clinical_raw %>%
-  filter( len_hospitalisation >= min_hosp_days, 
-          cohort  %in% c("AllAdm")) %>%
-  dplyr::mutate( date = as.Date( calendar_date, format = dateFormat ),
-                 weeks = as.Date(cut( date, breaks = "week")),
-                 month = as.Date(cut( date, breaks = "month")),
-                 year = format( date, "%Y"), 
-                 period = ifelse( date < pandemic_start_date,
-                                  "before_pandemic", "during_pandemic"))
-
-if( time_period == "weeks"){
-  clinical_filtered$time_p <- clinical_filtered$weeks
-}else if(time_period == "days"){
-  clinical_filtered$time_p <- clinical_filtered$date
-}else if(time_period == "months"){
-  clinical_filtered$time_p <- clinical_filtered$month
-}
-
-```
-
-```{r}
-# Steps necessary to filter rows:
-  # - encounters without ICD
-  # - ICD diagnosed outside of encounters
-  # - patients from the patients table that don't have any encounters matching the criteria
-clinical_table <- clinical_filtered %>% 
-  inner_join( distinct(obs_filtered_age, patient_num, days_since_admission),  by = c("patient_num", "days_since_admission")) %>% 
-  distinct(patient_num, n_hospitalisation) %>% 
-  inner_join(clinical_filtered, by = c("patient_num", "n_hospitalisation"))
-# Adding new columns
-observation_table <- obs_filtered_age %>% 
-  inner_join(clinical_table[, c("patient_num", "days_since_admission")], by = c("patient_num", "days_since_admission"))
-demographic_table <- demo_raw %>% 
-  inner_join(distinct(clinical_table, patient_num), by = "patient_num")
-all_codes <- observation_table %>% 
-   inner_join(select(clinical_table, -calendar_date), by = c("patient_num", "days_since_admission")) %>% 
-   inner_join(demographic_table[, c("patient_num", "sex")], by = "patient_num")
-```
-
-```{r}
-stopifnot(unique(all_codes$cohort) == "AllAdm")
-stopifnot(unique(all_codes$in_hospital) == 1)
-stopifnot(unique(all_codes$len_hospitalisation >= min_hosp_days))
-stopifnot(nrow(all_codes) == nrow(observation_table))
-stopifnot(length(unique(all_codes$patient_num)) == length(unique(demographic_table$patient_num)))
-stopifnot(length(unique(all_codes$patient_num)) == length(unique(observation_table$patient_num)))
-stopifnot(length(unique(all_codes$patient_num)) == length(unique(clinical_table$patient_num)))
-stopifnot(nrow(distinct(all_codes, patient_num, n_hospitalisation)) == nrow(distinct(clinical_table, patient_num, n_hospitalisation)))
-stopifnot(nrow(distinct(all_codes, patient_num, days_since_admission)) == nrow(distinct(observation_table, patient_num, days_since_admission)))
-```
-
-### Filtering on age 11-17
-
-```{r}
-# Filtering hospitalizations where patient age is in inclusion criteria
-observation_table_ado <- filter(observation_table, age_time_diagnosis >= 11 & age_time_diagnosis <= 17)
-
-# Filtering observation tables which patient_num and days_since_admission are still in clinical table after filtering on age
-
-clinical_test <- clinical_table[, c("patient_num", "n_hospitalisation", "days_since_admission")] %>% 
-  inner_join(distinct(observation_table_ado, patient_num, days_since_admission, age_time_diagnosis), 
-            by = c("patient_num", "days_since_admission")) %>% 
-  distinct(patient_num, n_hospitalisation, age_time_diagnosis)
-
-clinical_table_ado <- clinical_table %>% 
-  inner_join(clinical_test,
-             by = c("patient_num", "n_hospitalisation"))
-
-# Filtering observation tables which patient_num are still in clinical table after filtering on age
-demographic_table_ado <- demographic_table %>% 
-  inner_join(distinct(clinical_table_ado, patient_num), by = "patient_num")
-
-all_codes_ado <- all_codes %>% 
-   filter(age_time_diagnosis >= 11 & age_time_diagnosis <= 17)
-```
-
-```{r}
-stopifnot(unique(all_codes_ado$cohort) == "AllAdm")
-stopifnot(unique(all_codes_ado$in_hospital) == 1)
-stopifnot(unique(all_codes_ado$len_hospitalisation >= min_hosp_days))
-stopifnot(nrow(all_codes_ado) == nrow(observation_table_ado))
-stopifnot(length(unique(all_codes_ado$patient_num)) == length(unique(demographic_table_ado$patient_num)))
-stopifnot(length(unique(all_codes_ado$patient_num)) == length(unique(observation_table_ado$patient_num)))
-stopifnot(length(unique(all_codes_ado$patient_num)) == length(unique(clinical_table_ado$patient_num)))
-stopifnot(nrow(distinct(all_codes_ado, patient_num, n_hospitalisation)) == nrow(distinct(clinical_table_ado, patient_num, n_hospitalisation)))
-stopifnot(nrow(distinct(all_codes_ado, patient_num, days_since_admission)) == nrow(distinct(observation_table_ado, patient_num, days_since_admission)))
-```
-
-#### QC step: do a summary of the mental health ICD codes present in each site
-As output we will generate a table with: the number of observations per code, number of patients diagnosed with each code and the date range in which the code was used.
-
-```{r}
-#filter the whole dataset to only mental codes
-mental_codes_qc <- all_codes_ado %>% 
-  filter( concept_code %in% mentalHealthCodes$concept_code )
-
-mental_codes_qc_output <- mental_codes_qc %>%
-  group_by( concept_code ) %>%
-  dplyr::summarise( distinct_patients = ifelse(length(unique(patient_num)) > obfuscation | isFALSE(obfuscation),
-                               length(unique(patient_num)), 
-                               0.5),
-                    distinct_observations = ifelse(length(patient_num) > obfuscation | isFALSE(obfuscation),
-                               length(patient_num), 
-                               0.5),                           
-                    min_date = min( dmy( calendar_date ) ), 
-                    max_date = max( dmy( calendar_date ) ))%>% 
-  arrange(desc(distinct_observations))
-
-mental_codes_qc_output <- left_join( mental_codes_qc_output, icdCodes, by=c("concept_code"= "ICD10_Code")) 
-mental_codes_qc_output  <- mental_codes_qc_output[, c("disorder_group", "description", "concept_code", "distinct_patients", 
-                                                    "distinct_observations", "min_date", "max_date")]
-
-mental_codes_qc_output[1:10, ]
-```
-
-### Add the psychiatric/mental health ICD code description and filter
-
-```{r message=FALSE, warning=FALSE}
-# Getting site name, will be used later in the analysis
-site <- unique(as.character(demo_raw$siteid))
-
-icdCategory <- read.delim("public-data/icd10Codes.txt",
-                          header = FALSE,
-                          colClasses = "character")
-colnames(icdCategory) <- c("icd_code_letter", "icd_description_category")
-
-all_codes_desc <- all_codes_ado %>% 
-    left_join( icdCodes, by=c("concept_code" = "ICD10_Code") ) %>%
-  mutate(icd_code_category = ifelse(concept_code %in% icdCodes$ICD10_Code, 
-                               "psy", 
-                               "others")) %>%
-  dplyr::select( patient_num, sex, age_time_diagnosis, concept_type, concept_code, calendar_date,
-                 in_hospital, n_hospitalisation, len_hospitalisation, date, weeks, month, year,
-                 period, time_p, disorder_group, description, icd_code_category )
-
-all_codes_desc <- all_codes_desc %>% 
-  mutate( icd_code_letter = ifelse(substr(concept_code, 1, 1) %in% c("D", "H"), substr(concept_code, 1, 2), substr(concept_code, 1, 1) ))
-
-all_codes_desc <- left_join( all_codes_desc, icdCategory) %>%
-  mutate( icd_description_category = replace_na( ifelse( concept_code %in% icdCodes$ICD10_Code, "Psy", icd_description_category), "Others")) 
-all_codes_psy <- all_codes_desc %>% 
-  filter(concept_code %in% icdCodes$ICD10_Code)
-
-if( raceAvailable == TRUE ){
-  all_codes_psy <- all_codes_psy %>%
-    left_join( race_raw, by = c("patient_num") )
-}
+psyCounts <- psyCounts %>%
+  mutate( time_p = ymd(paste( psyCounts$yr2, psyCounts$mo2, "01", sep="-")), 
+          period = ifelse( time_p < pandemic_start_date,
+                           "before_pandemic",
+                           "during_pandemic"), 
+          siteid = "CHOP" ) %>%
+  select( time_p, count_psy_patients = ct2, period, siteid )
 ```
@@
 
-```{r}
-new_hospitalizations <- clinical_table %>%
-  group_by(patient_num, n_hospitalisation) %>%
-  slice(which.min(time_p)) %>%
-  filter( in_hospital == 1 ) %>%
-    group_by( time_p ) %>%
-  summarise(count = ifelse(n_distinct(patient_num) > obfuscation | isFALSE(obfuscation),
-                               n_distinct(patient_num), 
-                               0.5)) %>% 
-  ungroup()
-
-nh <- admission_plot_count(df = new_hospitalizations, 
-                     x = "time_p",
-                     y = "count", 
-                     title = paste0("New admissions (per ", time_period,")"), 
-                     y_title = "Patient Count")
-
-nh
-```
-
-## New admissions age range 11-17
-
-```{r}
-new_hospitalizations_ado <- clinical_table_ado %>%
-  group_by(patient_num, n_hospitalisation) %>%
-  slice(which.min(time_p)) %>%
-  filter( in_hospital == 1 ) %>%
-    group_by( time_p ) %>%
-  summarise(count = ifelse(n_distinct(patient_num) > obfuscation | isFALSE(obfuscation),
-                               n_distinct(patient_num), 
-                               0.5)) %>% 
-  ungroup()
-
-nh_ado <- admission_plot_count(df = new_hospitalizations_ado, 
-                     x = "time_p",
-                     y = "count", 
-                     title = paste0("New admissions (per ", time_period, ")"), 
-                     y_title = "Patient Count")+
-    labs( subtitle = "(age range 11-17)")
-
-nh_ado
-```
-
 ## Total hospitalize patients age range 11-17
@@
 ```{r}
-unique_patients_hospitalized <- clinical_table_ado %>%
-  filter(date < end_date_plots &
-         date >= start_date_plots ) %>%
-    group_by( time_p ) %>%
-  summarise(count = ifelse(n_distinct(patient_num) > obfuscation | isFALSE(obfuscation),
-                               n_distinct(patient_num), 
-                               0.5))
-
-th <- admission_plot_count(df = unique_patients_hospitalized, 
+th <- admission_plot_count(df = allCounts, 
                      x = "time_p",
-                     y = "count", 
+                     y = "count_all", 
                      title = paste0("Total hospitalization (per ", time_period, ")"), 
@@
 ```{r}
-count_icd_psy <- all_codes_psy %>%
-  filter(date < end_date_plots &
-         date >= start_date_plots ) %>%
-    group_by( time_p ) %>%
-  summarise(count_psy =  ifelse(n_distinct(patient_num) > obfuscation | isFALSE(obfuscation),
-                               n_distinct(patient_num), 
-                               0.5)) %>%
-  mutate( period = ifelse( time_p < pandemic_start_date,
-                           "before_pandemic",
-                           "during_pandemic")
-  )
-
-count_icd <- all_codes_desc %>%
-  filter(date < end_date_plots &
-         date >= start_date_plots ) %>%
-    group_by( time_p ) %>%
-  summarise(count_icd =  ifelse(n_distinct(patient_num) > obfuscation | isFALSE(obfuscation),
-                               n_distinct(patient_num), 
-                               0.5)) %>%
-  mutate( period = ifelse( time_p < pandemic_start_date,
-                           "before_pandemic",
-                           "during_pandemic")
-  )
-
-count_icd <- count_icd %>%
-  left_join(count_icd_psy, 
-            by = c("time_p", "period")) %>% 
-  replace_na(list(count_psy = 0))
-
-count_icd$percentage_psy <- 100*count_icd$count_psy / count_icd$count_icd
-
-thpd_psyCount <- count_icd %>%
-  pivot_longer(names_to = "counts", cols=c(count_psy, count_icd)) %>%
-  mutate( counts = ifelse( counts == "count_psy", "Psy", "Total")) %>%
-  ggplot(aes(x = time_p, y = value, fill = counts, color = counts)) +
+thpd_psyCount <- psyCounts %>%
+  ggplot(aes(x = time_p, y = count_psy_patients )) +
   geom_point() +
-  # geom_line() +
   geom_vline(xintercept = as.Date(pandemic_start_date),
@@
 ```{r}
-all_patients <- unique( all_codes_desc[, c("patient_num", "time_p")])
-patients_with_psy_conditions <- unique( all_codes_psy[, c("patient_num", "time_p")] )
-patients_without_psy_conditions <- setdiff(all_patients,patients_with_psy_conditions)
+ratios_patients_with_without_psy <- inner_join( allCounts, psyCounts )
 
-patients_without_psy_conditions_counts <- patients_without_psy_conditions %>%
-  group_by( time_p ) %>%
-  summarise(count_no_psy_patients =  ifelse(n_distinct(patient_num) > obfuscation | isFALSE(obfuscation),
-                               n_distinct(patient_num), 
-                               0.5)) %>%
-  mutate( period = ifelse( time_p < pandemic_start_date,
-                           "before_pandemic",
-                           "during_pandemic"))
-
-patients_with_psy_conditions_counts <-  patients_with_psy_conditions %>%
-  group_by( time_p ) %>%
-  summarise(count_psy_patients =  ifelse(n_distinct(patient_num) > obfuscation | isFALSE(obfuscation),
-                               n_distinct(patient_num), 
-                               0.5)) %>%
-  mutate( period = ifelse( time_p < pandemic_start_date,
-                           "before_pandemic",
-                           "during_pandemic"))
-
-ratios_patients_with_without_psy <- patients_without_psy_conditions_counts %>%
-  full_join( patients_with_psy_conditions_counts, by=c("time_p", "period")) %>%
-  mutate( ratio = count_psy_patients/count_no_psy_patients, 
-          percentage = 100 * count_psy_patients/(count_no_psy_patients + count_psy_patients))
+ratios_patients_with_without_psy <- ratios_patients_with_without_psy %>%
+  group_by( time_p, period, siteid ) %>%
+  mutate( count_no_psy_patients = count_all - count_psy_patients, 
+          ratio = round( count_psy_patients / count_no_psy_patients, 3), 
+          percentage = round( 100*count_psy_patients / count_all, 1) ) %>%
+  select( time_p, count_no_psy_patients, period, count_psy_patients, ratio, percentage, siteid )
 
@@
 ```
-
-## Psychiatric conditions aggregating by subcategories
-
-### New admissions: count
-
-```{r}
-count_disorder_group <- all_codes_psy %>%
-  filter(date < end_date_plots &
-         date >= start_date_plots ) %>%
-    group_by( time_p, disorder_group ) %>%
-  summarise(count_dg =  ifelse(n_distinct(patient_num) > obfuscation | isFALSE(obfuscation),
-                               n_distinct(patient_num), 
-                               0.5)) %>%
-  mutate( period = ifelse( time_p < pandemic_start_date,
-                           "before_pandemic",
-                           "during_pandemic")
-  )
-
-dg <- count_disorder_group %>%  
-  ggplot(aes(x = time_p, y= count_dg, fill = period, group = disorder_group )) +
-  geom_bar(stat = "identity")+
-  geom_vline(xintercept = as.Date(pandemic_start_date),
-             linetype = "dashed") +
-  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se= TRUE, aes(colour=period), colour= "black", size = 0.5) + 
-  labs(x = time_period,
-       y = "Patient Count",
-       title = paste0("Per ", time_period, " patient counts with mental health related ICD codes"), 
-       subtitle = "Grouped by mental health ICD subcategories") + 
-  ylim(0, NA) + 
-  facet_wrap(. ~ disorder_group, scales = "free_y") +
-  theme(strip.text.x = element_text(size = 4),
-        axis.text.x = element_text(size = 5, angle = 90), 
-        axis.text.y = element_text(size = 6), 
-        axis.title.x = element_text(size = 10),
-        axis.title.y = element_text(size = 10),
-        title = element_text(size = 12),
-        plot.subtitle = element_text(size = 11),
-        legend.position = "none") +
-  scale_fill_manual(values = cbPalette[-1])
-
-dg
-```
-
-#### Subset: the most prevalent disorder group
-```{r}
-dg_subset <- count_disorder_group %>%
-  filter( disorder_group %in% c("Anxiety Disorders", "Depressive Disorders", "Suicide or Self-Injury" )) %>%
-  ggplot(aes(x = time_p, y= count_dg, fill = period, group = disorder_group )) +
-  geom_bar(stat = "identity")+
-  geom_vline(xintercept = as.Date(pandemic_start_date),
-             linetype = "dashed") +
-  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se= TRUE, aes(colour=period), colour= "black", size = 0.5) + 
-  labs(x = time_period,
-       y = "Patient Count",
-       title = paste0("Per ", time_period, " patient counts with mental health related ICD codes"), 
-       subtitle = "Grouped by mental health ICD subcategories") + 
-  ylim(0, NA) + 
-  facet_wrap(. ~ disorder_group, scales = "free_y") +
-  theme(strip.text.x = element_text(size = 12),
-        axis.text.x = element_text(size = 12, angle = 90), 
-        axis.text.y = element_text(size = 11), 
-        axis.title.x = element_text(size = 12),
-        axis.title.y = element_text(size = 11),
-        title = element_text(size = 12),
-        plot.subtitle = element_text(size = 11),
-        legend.position = "none") +
-  scale_fill_manual(values = cbPalette[-1])
-
-dg_subset
-```
-
-### New admissions: percentage
-
-```{r}
-# estimate the counts
-count_icd_all <- all_codes_desc %>%
-  filter(date < end_date_plots &
-         date >= start_date_plots ) %>%
-    group_by( time_p ) %>%
-  summarise(count_icd =  ifelse(n_distinct(patient_num) > obfuscation | isFALSE(obfuscation),
-                               n_distinct(patient_num), 
-                               0.5)) %>%
-  mutate( period = ifelse( time_p < pandemic_start_date,
-                           "before_pandemic",
-                           "during_pandemic")
-  )
-
-# join the total with the disorder group count to estimate the percentage
-perc_disorder_group <- count_icd_all %>%
-  left_join(count_disorder_group, 
-            by = c("time_p", "period")) %>% 
-  replace_na(list(count_dg = 0))
-
-perc_disorder_group$percentage_dg <- 100*perc_disorder_group$count_dg / perc_disorder_group$count_icd
-
-dgPerc <- perc_disorder_group %>%  
-  ggplot(aes(x = time_p, y= percentage_dg, fill = period, group = disorder_group )) +
-  geom_bar(stat = "identity")+
-  geom_vline(xintercept = as.Date(pandemic_start_date),
-             linetype = "dashed") +
-  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se= TRUE, aes(colour=period), colour= "black", size = 0.5) + 
-  labs(x = time_period,
-       y = "Patient (%)",
-       title = paste0("Percentage of patients per ", time_period,"with mental health related ICD codes"), 
-       subtitle = "Grouped by mental health ICD subcategories") + 
-  ylim(0, NA) + 
-  facet_wrap(. ~ disorder_group, scales = "free_y") +
-  theme(strip.text.x = element_text(size = 4),
-        axis.text.x = element_text(size = 5, angle = 90), 
-        axis.text.y = element_text(size = 6), 
-        axis.title.x = element_text(size = 10),
-        axis.title.y = element_text(size = 10),
-        title = element_text(size = 12),
-        plot.subtitle = element_text(size = 11),
-        legend.position = "none") +
-  scale_fill_manual(values = cbPalette[-1])
-
-dgPerc
-```
-```{r}
-dgPerc_subset <- perc_disorder_group %>%  
-  filter( disorder_group %in% c("Anxiety Disorders", "Depressive Disorders", "Suicide or Self-Injury" )) %>%
-  ggplot(aes(x = time_p, y= percentage_dg, fill = period, group = disorder_group )) +
-  geom_bar(stat = "identity")+
-  geom_vline(xintercept = as.Date(pandemic_start_date),
-             linetype = "dashed") +
-  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se= TRUE, aes(colour=period), colour= "black", size = 0.5) + 
-  labs(x = time_period,
-       y = "Patient (%)",
-       title = paste0("Percentage of patients per ", time_period,"with mental health related ICD codes"), 
-       subtitle = "Grouped by mental health ICD subcategories") + 
-  ylim(0, NA) + 
-  facet_wrap(. ~ disorder_group, scales = "free_y") +
-  theme(strip.text.x = element_text(size = 12),
-        axis.text.x = element_text(size = 12, angle = 90), 
-        axis.text.y = element_text(size = 12), 
-        axis.title.x = element_text(size = 12),
-        axis.title.y = element_text(size = 12),
-        title = element_text(size = 12),
-        plot.subtitle = element_text(size = 11),
-        legend.position = "none") +
-  scale_fill_manual(values = cbPalette[-1])
-
-dgPerc_subset
-```
-
-## Psychiatric conditions, dichotomized on patients' sex
-
-### Count
-
-```{r}
-# counts per time period and sex
-patients_hospitalized_agg_sex <- all_codes_psy %>%
-  filter(date < end_date_plots &
-         date >= start_date_plots & 
-           sex %in% c("female", "male")) %>%
-    group_by( time_p, sex ) %>%
-  summarise(count_sex =  ifelse(n_distinct(patient_num) > obfuscation | isFALSE(obfuscation),
-                               n_distinct(patient_num), 
-                               0.5))
-
-patients_hospitalized_agg_sex %>%
-  ggplot(aes(x = time_p, y = count_sex, fill = sex)) +
-  geom_bar(stat="identity", position = "dodge")+
-  scale_fill_manual(values = cbPalette) +
-  labs(y = "Counts",
-       x = time_period,
-       title = paste0("Patient counts per ", time_period, " aggregated by sex"),
-       subtitle = "Per Sex") + 
-  theme(legend.position = "bottom") + 
-  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=sex), colour= "black", size = 0.5)
-```
-
-### Percentage
-
-```{r}
-##percentage
-total_patients_psy <- all_codes_psy %>%
-  filter(date < end_date_plots &
-         date >= start_date_plots ) %>%
-    group_by( time_p ) %>%
-  summarise(count =  ifelse(n_distinct(patient_num) > obfuscation | isFALSE(obfuscation),
-                               n_distinct(patient_num), 
-                               0.5))
-
-patients_hospitalized_agg_sex_perc <- total_patients_psy %>%
-  left_join( patients_hospitalized_agg_sex, by = c("time_p")) %>% 
-  replace_na(list(count_sex = 0))
-
-patients_hospitalized_agg_sex_perc$percentage <- 100*patients_hospitalized_agg_sex_perc$count_sex/patients_hospitalized_agg_sex_perc$count
-
-patients_hospitalized_agg_sex_perc %>%
-  ggplot(aes(x = time_p, y = percentage, fill = sex)) +
-  geom_bar(stat="identity", position = "dodge")+
-  scale_fill_manual(values = cbPalette) +
-  labs(y = "Percentage ",
-       x = time_period,
-       title = paste0("Patient (%) per ", time_period," aggregated by sex"),
-       subtitle = "Per Sex") + 
-  theme(legend.position = "bottom") + 
-  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=sex), colour= "black", size = 0.5)
-```
-
-## Table 1
-
-```{r}
-
-t.test2 <- function(m1, m2, s1, s2, n1, n2) {
-  # t.test computed from sample statistics, ie using mean, standard deviation and count value of the samples
-  # useful in case we can only retrieve aggregated stats from the sites
-  # ref: https://stats.stackexchange.com/questions/30394/how-to-perform-two-sample-t-tests-in-r-by-inputting-sample-statistics-rather-tha
-  # and https://en.wikipedia.org/wiki/Student%27s_t-test?oldformat=true#Equal_or_unequal_sample_sizes,_unequal_variances_(sX1_%3E_2sX2_or_sX2_%3E_2sX1)
-  se <- sqrt( (s1^2/n1) + (s2^2/n2) )
-  # welch-satterthwaite df
-  df <- ( (s1^2/n1 + s2^2/n2)^2 )/( (s1^2/n1)^2/(n1-1) + (s2^2/n2)^2/(n2-1) )
-  t <- (m1-m2)/se
-  dat <- c(m1-m2, se, t, 2*pt(-abs(t),df))
-  names(dat) <- c("Difference of means", "Std Error", "t", "p-value")
-  return(dat)
-}
-
-age_by_period <- all_codes_psy %>%
-  distinct(patient_num, age_time_diagnosis, period) %>% 
-  dplyr::group_by(period) %>%
-  dplyr::summarise( mean = mean( age_time_diagnosis ),
-                    sd = sd( age_time_diagnosis ),
-                    count = n_distinct(patient_num))
-
-age_test <- t.test2(m1 = age_by_period[[1, "mean"]],
-                    m2 = age_by_period[[2, "mean"]],
-                    s1 = age_by_period[[1, "sd"]],
-                    s2 = age_by_period[[2, "sd"]],
-                    n1 = age_by_period[[1, "count"]],
-                    n2 = age_by_period[[2, "count"]]) %>%
-  `[[`("p-value")
-```
-
-```{r}
-sex <- all_codes_psy %>%
-  dplyr::distinct(patient_num, sex) %>% 
-  dplyr::group_by( sex ) %>%
-  dplyr::summarise( count = n_distinct(patient_num) ) %>%
-  dplyr::mutate( count = ifelse( count > obfuscation, count, obfuscation * 0.5),
-                 total = sum( count ), percentage = round( count/total*100, 2)  )
-
-sex_by_period <- all_codes_psy %>%
-  dplyr::distinct(patient_num, period, sex) %>% 
-  group_by(period, sex ) %>%
-  summarise(count = n_distinct(patient_num)) %>%
-  group_by(period, .add=TRUE) %>%
-  mutate( count = ifelse( count > obfuscation, count, obfuscation * 0.5),
-          percentage=round(count/sum( count )*100, 2)) %>%
-  replace_na(list(male = 0, female = 0))
-
-sex_test <- sex_by_period %>%
-  select(-percentage) %>%
-  pivot_wider(names_from = sex, values_from = count) %>%
-  ungroup() %>%
-  select(male, female) %>%
-  chisq.test() %>%
-  `[[`("p.value")
-```
-
-```{r}
-summary_diagnosis <- all_codes_psy %>%
-  dplyr::distinct( patient_num, description, period, weeks, concept_code ) %>%
-  dplyr::group_by( concept_code, description, period ) %>%
-  dplyr::count( patient_num ) %>%
-  dplyr::tally(name = "count") %>%
-  dplyr::mutate(count = ifelse( count > obfuscation, count, obfuscation * 0.5)) %>%
-  dplyr::arrange( desc(count) ) %>%
-  dplyr::ungroup()
-
-summary_diagnosis_test <- summary_diagnosis %>%
-  tidyr::pivot_wider(names_from = period,
-                     values_from = count) %>%
-  #select(late_pandemic, early_pandemic) %>%
-  select(during_pandemic) %>%
-  #tidyr::replace_na(list(late_pandemic = 0, early_pandemic = 0)) %>%
-  tidyr::replace_na(list(during_pandemic = 0)) %>%
-  chisq.test() %>%
-  `[[`("p.value")
-
-```
-
-```{r}
-summary_disorder_group <- all_codes_psy %>%
-  dplyr::group_by( disorder_group ) %>%
-  dplyr::count( patient_num ) %>%
-  dplyr::tally(name = "count") %>%
-  dplyr::mutate(count = ifelse( count > obfuscation, count, obfuscation * 0.5)) %>%
-  dplyr::arrange( desc( count ))
-
-summary_disorder_group_by_period <- all_codes_psy %>%
-  dplyr::distinct( patient_num, disorder_group, period, weeks ) %>%
-  dplyr::group_by( period, disorder_group ) %>%
-  summarise(count = n_distinct(patient_num)) %>%
-  group_by(period, .add=TRUE) %>%
-  mutate(count = ifelse( count > obfuscation, count, obfuscation * 0.5),
-         percentage = count/sum( count )*100)
-
-summary_disorder_group_by_period_test <- summary_disorder_group_by_period %>%
-  pivot_wider(id_cols = disorder_group, names_from = period, values_from = c(count, percentage)) %>%
-  mutate(across(where(is.numeric), ~replace_na(.x, replace = 0))) %>%
-  #select(count_early_pandemic, count_late_pandemic) %>%
-  select(count_during_pandemic) %>%
-  chisq.test() %>%
-  `[[`("p.value")
-
-```
-
-```{r}
-format_counts <- function(count, percentage) {
-  glue::glue("{count} ({percentage} %)", count = count, percentage = percentage)
-}
-
-age_table1 <- age_by_period %>% 
-  mutate(result = glue::glue("{mean} (sd: {sd})", mean = round(mean, 2), sd = round(sd, 2))) %>% 
-  select(result, period) %>% 
-  pivot_wider( names_from = period, values_from = result ) %>%
-  mutate(name = "Age")
-
-sex_table1 <- sex_by_period %>%
-  bind_rows(
-    sex_by_period %>%
-      group_by(period) %>%
-      dplyr::summarise(sex = "Total", count = sum(count), percentage = sum(percentage))
-  ) %>%
-  mutate(result = format_counts(count, percentage)) %>%
-  select(-c(count, percentage)) %>%
-  pivot_wider( names_from = period, values_from = result ) %>%
-  rename(name = sex)
-
-disorder_table1 <- summary_disorder_group_by_period %>%
-  arrange(desc(count)) %>% 
-  bind_rows(
-    dplyr::summarise(., disorder_group="Total", across(where(is.numeric), sum))
-  ) %>%
-  mutate(percentage = round(percentage, 2)) %>% 
-  mutate(result = format_counts(count, percentage)) %>%
-  select(-c(count, percentage)) %>%
-  pivot_wider( names_from = period, values_from = result ) %>%
-  replace_na( list(during_pandemic = "0 (0 %)")) %>%
-  rename(name = disorder_group)
-```
-
-```{r}
-list_tests <- c("Sex" = sex_test,
-                "Age" = age_test,
-                "Diagnoses" = summary_diagnosis_test,
-                "Disorder groups" = summary_disorder_group_by_period_test
-)
-
-list_table1 <- list(
-  "Sex" = sex_table1,
-  "Age" = age_table1,
-  "Disorder groups" = disorder_table1)
-
-tests <- list_tests %>%
-  data.frame() %>%
-  round(2)
-names(tests) <- "pvalue"
-tests["Category"] <- row.names(tests)
-
-table1 <- list_table1 %>%
-  lapply(function(df) mutate(df, across(.fns = as.character))) %>%
-  dplyr::bind_rows(.id = "Category") %>%
-  dplyr::left_join(tests, by = "Category")
-
-table1 %>%
-  group_by(Category) %>%
-  mutate(pvalue = ifelse(duplicated(pvalue, fromLast = TRUE), "", as.character(pvalue))) %>%
-  ungroup() %>%
-  mutate(Category = "") %>%
-  mutate(across(.cols = c(before_pandemic, during_pandemic), .fns = replace_na, "0")) %>%
-  rename_with(str_to_title) %>%
-  rename_with(sub, pattern = "_", replacement = " ") %>%
-  mutate(across(.fns = str_to_title)) %>%
-  mutate(across(.fns = sub, pattern = "_", replacement = " ")) %>%
-  kable( booktabs = T) %>%
-  kableExtra::kable_styling(latex_options = "striped") %>%
-  kableExtra::group_rows(index = setNames(rle(table1$Category)[[1]], rle(table1$Category)[[2]]))
-```
-
-# Save the output in a RData file. 
-```{r}
-length_hospitalisation_values <- all_codes_psy %>% 
-  dplyr::distinct( patient_num, n_hospitalisation, len_hospitalisation, period ) %>%
-  dplyr::mutate( siteid = site ) %>%
-  dplyr::select( len_hospitalisation, siteid, period )
-
-# add the site id as column for each of the output data frames
-mental_codes_qc_output$siteid = site 
-count_icd$siteid = site 
-patient_count_psy_period_psy$siteid = site 
-patient_count_psy_period_psy_clear$siteid = site 
-bootstrapped_coefficients_df$siteid = site 
-bootstrapped_fitted_df$siteid = site 
-perc_disorder_group$siteid = site 
-patients_hospitalized_agg_sex_perc$siteid = site 
-table1$siteid = site 
-ratios_patients_with_without_psy$siteid = site
-
-save( count_icd,
-      mental_codes_qc_output,
-      ratios_patients_with_without_psy,
-      patient_count_psy_period_psy,
-      patient_count_psy_period_psy_clear,
-      bootstrapped_coefficients_df,
-      bootstrapped_fitted_df,
-      perc_disorder_group,
-      patients_hospitalized_agg_sex_perc,
-      table1,
-      length_hospitalisation_values,
-      file = paste0("output/peds_psy_aggResults_", site, ".RData"))
-```

### 2021-12-16 psy_peds_3_CHOP.Rmd
select the correct column to do the lm, time_p instead of time

@@
 ```{r, its count based}
-ts_linear_count <- lm(count_psy_patients ~ period * time, patient_count_psy_period_psy)
-ts_linear_count_clear <- lm(count_psy_patients ~ period * time, patient_count_psy_period_psy_clear)
-ts_seasonal_count <- lm(count_psy_patients ~ period*time + harmonic(month_year, 1, 12), patient_count_psy_period_psy)
-ts_seasonal_count_clear <- lm(count_psy_patients ~ period*time + harmonic(month_year, 1, 12), patient_count_psy_period_psy_clear)
+ts_linear_count <- lm(count_psy_patients ~ period * time_p, patient_count_psy_period_psy)
+ts_linear_count_clear <- lm(count_psy_patients ~ period * time_p, patient_count_psy_period_psy_clear)
+ts_seasonal_count <- lm(count_psy_patients ~ period*time_p + harmonic(month_year, 1, 12), patient_count_psy_period_psy)
+ts_seasonal_count_clear <- lm(count_psy_patients ~ period*time_p + harmonic(month_year, 1, 12), patient_count_psy_period_psy_clear)
 ```
@@
 ```{r, its percentage based}
-ts_linear_percent <- lm(percentage ~ period * time, patient_count_psy_period_psy)
-ts_linear_percent_clear <- lm(percentage ~ period * time, patient_count_psy_period_psy_clear)
-ts_seasonal_percent <- lm(percentage ~ period*time + harmonic(month_year, 1, 12), patient_count_psy_period_psy)
-ts_seasonal_percent_clear <- lm(percentage ~ period*time + harmonic(month_year, 1, 12), patient_count_psy_period_psy_clear)
+ts_linear_percent <- lm(percentage ~ period * time_p, patient_count_psy_period_psy)
+ts_linear_percent_clear <- lm(percentage ~ period * time_p, patient_count_psy_period_psy_clear)
+ts_seasonal_percent <- lm(percentage ~ period*time_p + harmonic(month_year, 1, 12), patient_count_psy_period_psy)
+ts_seasonal_percent_clear <- lm(percentage ~ period*time_p + harmonic(month_year, 1, 12), patient_count_psy_period_psy_clear)
 ```
@@
   df_iteration <- data.frame(period = patient_count_psy_period_psy$period,
-                             time = patient_count_psy_period_psy$time,
+                             time = patient_count_psy_period_psy$time_p,
                              percentage = bootstrapped_timepoints[iteration, ])

### 2022-01-21 psy_peds_3_CHOP.Rmd
Integrate additional analysis to the CHOP pipeline. Analysis by disorder group

--- 2021-12-15 psy_peds_3_CHOP.Rmd
@@
 
-listOfPackages <- c("tidyverse", "RColorBrewer", "knitr", "kableExtra", "tsModel", "gridExtra", "dplyr", "lubridate")
+listOfPackages <- c("tidyverse", "RColorBrewer", "knitr", "kableExtra", "tsModel", "gridExtra", "dplyr", "lubridate", "magrittr")
 paket(listOfPackages)
@@
 ```{r message=FALSE, warning=FALSE}
-outputDir <- "../output/CHOP/"
-allCounts <- read.csv( file = paste0( outputDir, "All-Counts.csv"))
-psyCounts <- read.csv( file = paste0( outputDir, "Psyc-Counts.csv"))
+allCounts <- read.csv( "../output/CHOP/All-Counts.csv")
+psyCounts <- read.csv( "../output/CHOP/Psyc-Counts.csv")
 
@@
 # Time series analysis of percentage of psy codes per encounter, linear (without seasonality)
-patient_count_psy_period_psy <- ratios_patients_with_without_psy %>%
-  filter(time_p > as.Date(cut( start_date_plots, breaks = time_period)) &
-           time_p < as.Date(cut( end_date_plots, breaks = time_period))) %>%
-  mutate(time = seq_along(time_p),
-         month_year = format(time_p, "%m")
-  )
+patient_count_psy_period_psy <- left_join(allCounts,
+                                          psyCounts) %>%
+  mutate(count_no_psy_patients = count_all - count_psy_patients,
+         time = seq_along(time_p),
+         period = ifelse(time_p < as.Date("2020-04-01"),
+                         "before_pandemic",
+                         "during_pandemic"),
+         ratio = count_psy_patients / count_no_psy_patients,
+         percentage = count_psy_patients * 100 / (count_psy_patients + count_no_psy_patients),
+         month_year = format.Date(time_p, "%m"),
+         siteid = "CHOP") %>%
+  select(time_p, count_no_psy_patients, period, count_psy_patients,
+         ratio, percentage, time, month_year, siteid)
 
@@
 ```
+
+## Psychiatric conditions aggregating by subcategories
+```{r}
+# Generate CHOP tables aggregated by sex and by mental disorder category
+psy_female_counts <- read_csv("../output/CHOP/Psy-Female-11-17.csv")
+psy_male_counts <- read_csv("../output/CHOP/Psy-Male-11-17.csv")
+
+psy_tot <- psy_female_counts %>% 
+  select(c(-Month, -Year)) %>%
+  add( psy_male_counts %>% select(c(-Month, -Year))) %>%
+  mutate(time_p = as.Date(paste(psy_female_counts$Year, psy_female_counts$Month, "01", sep = "-")))
+```
+
+```{r}
+count_disorder_group <- psy_tot %>%
+  pivot_longer(!time_p, names_to = "disorder_group", values_to = "count_dg") %>%
+  inner_join( patient_count_psy_period_psy %>%
+                mutate( total = count_no_psy_patients + count_psy_patients) %>%
+                select( time_p, count_no_psy_patients, count_psy_patients, total, period ))
+
+dg <- count_disorder_group %>%  
+  ggplot(aes(x = time_p, y= count_dg, fill = period, group = disorder_group )) +
+  geom_bar(stat = "identity")+
+  geom_vline(xintercept = as.Date(pandemic_start_date),
+             linetype = "dashed") +
+  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se= TRUE, aes(colour=period), colour= "black", size = 0.5) + 
+  labs(x = time_period,
+       y = "Patient Count",
+       title = paste0("Per ", time_period, " patient counts with mental health related ICD codes"), 
+       subtitle = "Grouped by mental health ICD subcategories") + 
+  ylim(0, NA) + 
+  facet_wrap(. ~ disorder_group, scales = "free_y") +
+  theme(strip.text.x = element_text(size = 4),
+        axis.text.x = element_text(size = 5, angle = 90), 
+        axis.text.y = element_text(size = 6), 
+        axis.title.x = element_text(size = 10),
+        axis.title.y = element_text(size = 10),
+        title = element_text(size = 12),
+        plot.subtitle = element_text(size = 11),
+        legend.position = "none") +
+  scale_fill_manual(values = cbPalette[-1])
+dg
+```
+
+#### Subset: the most prevalent disorder group
+```{r}
+dg_subset <- count_disorder_group %>%
+  filter( disorder_group %in% c("Anxiety Disorders", "Depressive Disorders", "Suicide or Self-Injury" )) %>%
+  ggplot(aes(x = time_p, y= count_dg, fill = period, group = disorder_group )) +
+  geom_bar(stat = "identity")+
+  geom_vline(xintercept = as.Date(pandemic_start_date),
+             linetype = "dashed") +
+  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se= TRUE, aes(colour=period), colour= "black", size = 0.5) + 
+  labs(x = time_period,
+       y = "Patient Count",
+       title = paste0("Per ", time_period, " patient counts with mental health related ICD codes"), 
+       subtitle = "Grouped by mental health ICD subcategories") + 
+  ylim(0, NA) + 
+  facet_wrap(. ~ disorder_group, scales = "free_y") +
+  theme(strip.text.x = element_text(size = 12),
+        axis.text.x = element_text(size = 12, angle = 90), 
+        axis.text.y = element_text(size = 11), 
+        axis.title.x = element_text(size = 12),
+        axis.title.y = element_text(size = 11),
+        title = element_text(size = 12),
+        plot.subtitle = element_text(size = 11),
+        legend.position = "none") +
+  scale_fill_manual(values = cbPalette[-1])
+dg_subset
+```
+
+### New admissions: percentage
+
+```{r}
+# estimate the percentage using as denominator the total number of patients per month with any psychiatric condition
+perc_disorder_group <- count_disorder_group %>%
+  mutate( percentage_dg = 100*count_dg / count_psy_patients )
+
+dgPerc <- perc_disorder_group %>%  
+  ggplot(aes(x = time_p, y= percentage_dg, fill = period, group = disorder_group )) +
+  geom_bar(stat = "identity")+
+  geom_vline(xintercept = as.Date(pandemic_start_date),
+             linetype = "dashed") +
+  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se= TRUE, aes(colour=period), colour= "black", size = 0.5) + 
+  labs(x = time_period,
+       y = "Patient (%)",
+       title = paste0("Percentage of patients per ", time_period,"with mental health related ICD codes"), 
+       subtitle = "Grouped by mental health ICD subcategories") + 
+  ylim(0, NA) + 
+  facet_wrap(. ~ disorder_group, scales = "free_y") +
+  theme(strip.text.x = element_text(size = 4),
+        axis.text.x = element_text(size = 5, angle = 90), 
+        axis.text.y = element_text(size = 6), 
+        axis.title.x = element_text(size = 10),
+        axis.title.y = element_text(size = 10),
+        title = element_text(size = 12),
+        plot.subtitle = element_text(size = 11),
+        legend.position = "none") +
+  scale_fill_manual(values = cbPalette[-1])
+dgPerc
+```
+
+```{r}
+dgPerc_subset <- perc_disorder_group %>%  
+  filter( disorder_group %in% c("Anxiety Disorders", "Depressive Disorders", "Suicide or Self-Injury" )) %>%
+  ggplot(aes(x = time_p, y= percentage_dg, fill = period, group = disorder_group )) +
+  geom_bar(stat = "identity")+
+  geom_vline(xintercept = as.Date(pandemic_start_date),
+             linetype = "dashed") +
+  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se= TRUE, aes(colour=period), colour= "black", size = 0.5) + 
+  labs(x = time_period,
+       y = "Patient (%)",
+       title = paste0("Percentage of patients per ", time_period,"with mental health related ICD codes"), 
+       subtitle = "Grouped by mental health ICD subcategories") + 
+  ylim(0, NA) + 
+  facet_wrap(. ~ disorder_group, scales = "free_y") +
+  theme(strip.text.x = element_text(size = 12),
+        axis.text.x = element_text(size = 12, angle = 90), 
+        axis.text.y = element_text(size = 12), 
+        axis.title.x = element_text(size = 12),
+        axis.title.y = element_text(size = 12),
+        title = element_text(size = 12),
+        plot.subtitle = element_text(size = 11),
+        legend.position = "none") +
+  scale_fill_manual(values = cbPalette[-1])
+dgPerc_subset
+```
+
+# Save the output in a RData file. 
+```{r}
+site = "CHOP"
+patient_count_psy_period_psy$siteid = site 
+patient_count_psy_period_psy_clear$siteid = site 
+bootstrapped_coefficients_df$siteid = site 
+bootstrapped_fitted_df$siteid = site 
+ratios_patients_with_without_psy$siteid = site 
+quantile_bootstrapped_fitted$siteid = site
+perc_disorder_group$siteid = site
+
+save( ratios_patients_with_without_psy,
+      patient_count_psy_period_psy,
+      patient_count_psy_period_psy_clear,
+      bootstrapped_coefficients_df,
+      bootstrapped_fitted_df,
+      quantile_bootstrapped_fitted,
+      perc_disorder_group,
+      file = paste0("../output/CHOP/peds_psy_aggResults_", site, ".RData"))
+
+```

### 2022-01-22 psy_peds_3_CHOP.Rmd
add to the CHOP pipeline the analysis aggregated by sex

@@
 
+## Psychiatric conditions, dichotomized on patients' sex
+
+### Count
+```{r}
+psy_female_counts <- read_csv("../output/CHOP/Psy-Counts-Female.csv")
+psy_male_counts <- read_csv("../output/CHOP/Psy-Counts-Male.csv")
+
+psy_female_counts <- psy_female_counts %>%
+  mutate(female = Counts, 
+         time_p = as.Date(paste(Year, Month, "01", sep = "-"))) %>%
+  select( time_p, female)
+
+psy_male_counts <- psy_male_counts %>%
+  mutate(male = Counts, 
+         time_p = as.Date(paste(Year, Month, "01", sep = "-"))) %>%
+  select( time_p, male)
+
+psy_total_counts <- psy_female_counts %>%
+  inner_join( psy_male_counts)
+
+patients_hospitalized_agg_sex <- psy_total_counts %>%
+  pivot_longer(!time_p, names_to = "sex", values_to = "count_sex")
+
+patients_hospitalized_agg_sex %>%
+  ggplot(aes(x = time_p, y = count_sex, fill = sex)) +
+  geom_bar(stat="identity", position = "dodge")+
+  scale_fill_manual(values = cbPalette) +
+  labs(y = "Counts",
+       x = time_period,
+       title = paste0("Patient counts per ", time_period, " aggregated by sex"),
+       subtitle = "Per Sex") + 
+  theme(legend.position = "bottom") + 
+  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=sex), colour= "black", size = 0.5)
+
+```
+
+### Percentage
+
+```{r}
+##percentage
+patients_hospitalized_agg_sex_perc <- psyCounts %>%
+  left_join( patients_hospitalized_agg_sex, by = c("time_p")) %>% 
+  replace_na(list(count_sex = 0))
+patients_hospitalized_agg_sex_perc$percentage <- 100*patients_hospitalized_agg_sex_perc$count_sex/patients_hospitalized_agg_sex_perc$count_psy_patients
+patients_hospitalized_agg_sex_perc %>%
+  ggplot(aes(x = time_p, y = percentage, fill = sex)) +
+  geom_bar(stat="identity", position = "dodge")+
+  scale_fill_manual(values = cbPalette) +
+  labs(y = "Percentage ",
+       x = time_period,
+       title = paste0("Patient (%) per ", time_period," aggregated by sex"),
+       subtitle = "Per Sex") + 
+  theme(legend.position = "bottom") + 
+  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=sex), colour= "black", size = 0.5)
+```
+
 # Save the output in a RData file. 
@@
 perc_disorder_group$siteid = site 
+patients_hospitalized_agg_sex_perc$siteid = site
 
@@
       perc_disorder_group,
+      patients_hospitalized_agg_sex_perc,
       file = paste0("../output/CHOP/peds_psy_aggResults_", site, ".RData"))

### 2022-01-25 metanalysis_visualization/results_all_sites.Rmd
fix the colnames of the CHOP outputs to match with the rest of sites. create a markdown to visualize the current outputs together

--- 2021-12-08 metanalysis_visualization/visualization.Rmd
@@
 for( i in 1:length( outputFiles )){
-  print(i)
+  print(outputFiles[i])
   load(paste0( outputDir, outputFiles[i]))
   if( i == 1 ){
-    count_icd_all <- count_icd
     patients_hospitalized_agg_sex_perc_all <- patients_hospitalized_agg_sex_perc
@@
     bootstrapped_coefficients_all <- bootstrapped_coefficients_df
-    mental_codes_qc_output_all <- mental_codes_qc_output %>%
-      select( disorder_group, description, concept_code, distinct_patients, distinct_observations, siteid )
   } else {
-    count_icd_all <- rbind(count_icd_all, count_icd)
-    patients_hospitalized_agg_sex_perc_all <- rbind( patients_hospitalized_agg_sex_perc_all, patients_hospitalized_agg_sex_perc)
+    patients_hospitalized_agg_sex_perc_all <- rbind( patients_hospitalized_agg_sex_perc_all, patients_hospitalized_agg_sex_perc )
     patient_count_psy_period_psy_all <- rbind( patient_count_psy_period_psy_all,patient_count_psy_period_psy)
@@
     bootstrapped_coefficients_all <- rbind(bootstrapped_coefficients_all, bootstrapped_coefficients_df)
-   mental_codes_qc_output <- mental_codes_qc_output %>%
-      select( disorder_group, description, concept_code, distinct_patients, distinct_observations, siteid )
-    mental_codes_qc_output_all <- rbind( mental_codes_qc_output_all, mental_codes_qc_output)
   }
-  rm(count_icd,patients_hospitalized_agg_sex_perc,patient_count_psy_period_psy,
+  rm(patients_hospitalized_agg_sex_perc,patient_count_psy_period_psy,
      perc_disorder_group,ratios_patients_with_without_psy,
-     patient_count_psy_period_psy_clear, table1, bootstrapped_coefficients_df,
-     bootstrapped_fitted_df, mental_codes_qc_output)
+     patient_count_psy_period_psy_clear, bootstrapped_coefficients_df,
+     bootstrapped_fitted_df)
 }
 ```
-
-## Percentage: patients with psychiatric conditions / total number of patients
 
@@
 ```
-
-### Summary of the codes used per site 
-```{r}
-mental_codes_qc_output_all <- mental_codes_qc_output_all %>%
-    mutate( concept_code = gsub("[.]", "", concept_code ) ) %>%
-  unique()
-
-summary_codes <- mental_codes_qc_output_all %>%
-  select( disorder_group, description, concept_code, distinct_patients, siteid ) %>%
-  spread( siteid, distinct_patients)
-
-## add a column with the sites combination 
-getSitesWithEachCode <- function(row) {
-  sites <- c( row['BCH'],  row['Lurie'], row['UMICH'], row['FRBDX'], row['APHP'])
-  sites <- sites[!is.na( sites ) ]
-  siteVal <- paste0( names( sites ), collapse = " - ")
-  return( siteVal)
-}
-summary_codes <- summary_codes[, c("disorder_group", "description", "concept_code", 
-                                   "BCH", "Lurie", "UMICH",  "FRBDX", "APHP") ]                        
-summary_codes$sitesCombination <- apply(summary_codes, 1, FUN = getSitesWithEachCode)
-summary_codes <- summary_codes[
-  with(summary_codes, order(BCH, FRBDX, Lurie, UMICH, APHP, decreasing = TRUE)),
-]
-print( paste0("There are in total ", nrow( summary_codes[ summary_codes$sitesCombination == "BCH - Lurie - UMICH - FRBDX - APHP", ]), " codes in common across all sites"))
-
-write.table( summary_codes, file="summary_codes.txt", col.names = TRUE, row.names = FALSE, sep = "\t", quote = FALSE )
-
-#### upSetR 
-upsetForDisorderGroup <- function( df, disorderGroup = "ALL"){
-  if( disorderGroup != "ALL"){
-     df <- df %>%
-       filter( disorder_group == disorderGroup ) 
-  }
-
-  BCH_concepts = df[ df$siteid == "BCH", "concept_code"]
-  FRBDX_concepts = df[ df$siteid == "FRBDX", "concept_code"]
-  Lurie_concepts = df[ df$siteid == "Lurie", "concept_code"]
-  UMICH_concepts = df[ df$siteid == "UMICH", "concept_code"]
-  APHP_concepts = df[ df$siteid == "APHP", "concept_code"]
-
-  concepts = list( 
-          BCH = BCH_concepts$concept_code,
-          FRBDX = FRBDX_concepts$concept_code, 
-          Lurie = Lurie_concepts$concept_code, 
-          UMICH = UMICH_concepts$concept_code, 
-          APHP  = APHP_concepts$concept_code
-          )
-  output <- upset(fromList(concepts), order.by = "freq")
-  return( output )
-}
-
-upsetForDisorderGroup( df = mental_codes_qc_output_all, disorderGroup = "ALL")
-```
-
-### Estimate number of ICD codes per site and disorder group
-Represent this in 16 pannels, if we put all together, due to the different scales we will not be able to really see the differences between results. 
-```{r}
-numberOfCodesPerDisorderGroup <- mental_codes_qc_output_all %>%
-  group_by( disorder_group, siteid ) %>%
-  mutate( icd_codes_count = n( )) %>%
-  select( disorder_group, siteid, icd_codes_count ) %>%
-  unique()
-
-swr = function(string, nwrap=30) {
-  paste(strwrap(string, width=nwrap), collapse="\n")
-}
-swr = Vectorize(swr)
-
-numberOfCodesPerDisorderGroup %>%
-  mutate( disorder_group = swr( disorder_group ) ) %>%
-  ggplot( aes( x= reorder( siteid, icd_codes_count), y = icd_codes_count)) +
-  geom_bar( stat = "identity") +
-  facet_wrap( ~ disorder_group, scales = "free" ) +
-  labs(x ="site identifier", y = "ICD dictionary (number of ICD codes)") +
-  theme(axis.text.x=element_text(angle=45,hjust=1, size = 6), 
-        strip.text = element_text(size = 7), 
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) 
-```

### 2022-01-25 psy_peds_3_CHOP.Rmd
fix the colnames of the CHOP outputs to match with the rest of sites. create a markdown to visualize the current outputs together

@@
 perc_disorder_group <- count_disorder_group %>%
-  mutate( percentage_dg = 100*count_dg / count_psy_patients )
+  mutate( percentage_dg = 100*count_dg / count_psy_patients ) %>%
+  select( time_p, count_icd = count_psy_patients, period, disorder_group, count_dg, percentage_dg )
 
@@
 patients_hospitalized_agg_sex_perc$percentage <- 100*patients_hospitalized_agg_sex_perc$count_sex/patients_hospitalized_agg_sex_perc$count_psy_patients
+patients_hospitalized_agg_sex_perc <- patients_hospitalized_agg_sex_perc %>% 
+  select( time_p, count= count_psy_patients, sex, count_sex, percentage )
 patients_hospitalized_agg_sex_perc %>%

### 2022-01-26 metanalysis_visualization/visualization.Rmd
add additional outputs from CHOP input data. integrate all the visualizations in one notebook and add new visualizations

@@
      perc_disorder_group,ratios_patients_with_without_psy,
-     patient_count_psy_period_psy_clear, table1, bootstrapped_coefficients_df,
+     patient_count_psy_period_psy_clear, bootstrapped_coefficients_df,
      bootstrapped_fitted_df, mental_codes_qc_output)
@@
 ```{r}
-### add CHOP 
-ratios_patients_with_without_psy_chop <- read.delim ( file = "/Users/alba/Desktop/ratios_patients_with_without_psy_chop.txt")
-
-ratios_patients_with_without_psy_all <- rbind( ratios_patients_with_without_psy_all, ratios_patients_with_without_psy_chop)
-
 ratios_patients_with_without_psy_all %>%
@@
     xlab("period (months)")
+
+ratios_patients_with_without_psy_all %>%
+  filter(time_p <= history_end_date &
+         time_p >= history_start_date ) %>%
+  ggplot( aes(x=time_p, y=ratio, 
+              group = siteid, color = siteid))+
+    facet_wrap("siteid") +
+  geom_point( aes(shape=siteid, color=siteid) ) + 
+  geom_line() +
+  geom_vline(xintercept = as.Date(pandemic_start_date),
+             linetype = "dashed") +
+  scale_fill_viridis(discrete = TRUE, alpha=0.6) +
+    theme(
+      legend.position="bottom",
+      plot.title = element_text(size=11)
+    ) +
+  scale_x_date(date_labels = "%b-%d-%Y", breaks = "month") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6)) +
+  guides(color = guide_legend(override.aes = list(alpha = 1))) +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6)) +
+    ggtitle("Ratio: patients with vs. patients without psychiatric conditions") +
+    xlab("period (months)") +
+  theme(strip.text = element_text(size = 10),
+        axis.text.x = element_text(size = 8), 
+        axis.text.y = element_text(size = 8), 
+        axis.title.x = element_text(size = 8),
+        axis.title.y = element_text(size = 8))
 ```
@@
 forest(pes.summary,layout = "JAMA")
-```
-
-### Boxplot showing ratio per month
-
-```{r}
-ratios_patients_with_without_psy_all %>%
-  ggplot( aes(x=as.factor(time_p), y=ratio)) +
-    geom_boxplot() +
-    scale_fill_viridis(discrete = TRUE, alpha=0.6) +
-    geom_jitter(color="black", size=0.4, alpha=0.9,
-                aes(shape = siteid)) +
-    theme(
-      legend.position="bottom",
-      plot.title = element_text(size=11)
-    ) +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6)) +
-    ggtitle("Ratio boxplot") +
-    xlab("")
 ```

### 2022-01-26 psy_peds_3_CHOP.Rmd
add additional outputs from CHOP input data. integrate all the visualizations in one notebook and add new visualizations

@@
   select( time_p, count_psy_patients = ct2, period, siteid )
+```
+
+### Read the ICD10 codes file
+
+The file containing the ICD10 psychiatric related codes is located in
+the `public-data` folder of the GitHub repository.
+
+```{r}
+icdCodes <- read.csv("public-data/pediatric_psychiatric_ICD10_codesV2.csv", header = TRUE, colClasses = "character") %>%
+   filter( flag %in% c("originalCodes", "FrenchSpecificCodes")) %>%
+   dplyr::select( disorder_group, concept_code = ICD10_Code, description )
+```
+
+#### QC step: do a summary of the mental health ICD codes present in each site
+As output we will generate a table with: the number of observations per code, number of patients diagnosed with each code and the date range in which the code was used.
+
+```{r}
+icdCounts_females <- read.csv( "../output/CHOP/Psy-Female-All Ages_ICDlevel.csv") %>%
+  pivot_longer(!c(Month, Year), names_to = "concept_code", values_to = "patients") %>%
+  mutate(  time_p = ymd(paste( Year, Month, "01", sep="-") ) ) %>%
+  select( time_p, concept_code, patients ) %>%
+    filter( patients > 0 ) %>%
+  group_by( concept_code ) %>%
+  summarise( total_females = sum( patients ), 
+             min_date = min( as.Date( time_p )) , 
+             max_date = max( as.Date( time_p )))
+
+icdCounts_males <- read.csv( "../output/CHOP/Psy-Male-All Ages_ICDlevel.csv") %>%
+  pivot_longer(!c(Month, Year), names_to = "concept_code", values_to = "patients") %>%
+  mutate(  time_p = ymd(paste( Year, Month, "01", sep="-") ) ) %>%
+  select( time_p, concept_code, patients ) %>%
+  filter( patients > 0 ) %>%
+  group_by( concept_code ) %>%
+  summarise( total_males = sum( patients ), 
+             min_date = min( as.Date( time_p )) , 
+             max_date = max( as.Date( time_p )))
+
+all_codes_ado <- full_join( icdCounts_females, icdCounts_males, by = "concept_code") %>%
+  rowwise() %>% 
+  mutate( distinct_patients = sum(total_females, total_males, na.rm = TRUE ), 
+          distinct_observations = sum(total_females, total_males, na.rm = TRUE ),
+          min_date = min( min_date.x, min_date.y, na.rm = TRUE), 
+          max_date = max( max_date.x, max_date.y, na.rm = TRUE)) %>%
+  select( concept_code, distinct_patients, distinct_observations, min_date, max_date ) %>%
+  inner_join( icdCodes )
+
+mental_codes_qc_output  <- all_codes_ado[, c("disorder_group", "description", "concept_code", "distinct_patients", 
+                                                    "distinct_observations", "min_date", "max_date")]
+mental_codes_qc_output[1:10, ]
 ```
@@
 }
+```
+
+```{r}
+count_icd <- left_join(allCounts, psyCounts) %>%
+  mutate(period = ifelse(time_p < as.Date("2020-04-01"),
+                         "before_pandemic",
+                         "during_pandemic"),
+         percentage_psy = count_psy_patients * 100 /count_all,
+         siteid = "CHOP") %>%
+  select(time_p, count_icd = count_all, period, count_psy = count_psy_patients,
+         percentage_psy, siteid)
 ```
@@
 patients_hospitalized_agg_sex_perc$siteid = site 
+mental_codes_qc_output$siteid = site
 
-save( ratios_patients_with_without_psy,
+save( count_icd, 
+      ratios_patients_with_without_psy,
       patient_count_psy_period_psy,
@@
       patients_hospitalized_agg_sex_perc,
+      mental_codes_qc_output,
       file = paste0("../output/CHOP/peds_psy_aggResults_", site, ".RData"))

### 2022-01-26 metanalysis_visualization/visualization.Rmd
integrate visualization notebooks and add more visualizations

@@
 
+## Patient Count with Mental Health Related ICD codes
+
+```{r}
+thpd_psyCount <- count_icd_all %>%
+  pivot_longer(names_to = "counts", cols=c(count_psy, count_icd)) %>%
+  mutate( counts = ifelse( counts == "count_psy", "Psy", "Total")) %>%
+  ggplot(aes(x = time_p, y = value, fill = counts, color = counts)) +
+  facet_grid(siteid~.) +
+  geom_point() +
+  geom_vline(xintercept = as.Date(pandemic_start_date),
+             linetype = "dashed") +
+  scale_fill_manual(values = cbPalette) +
+  scale_color_manual(values = cbPalette) +
+  labs(y = "Counts",
+       x = paste0("Date (by ", time_period,")"),
+       title = paste0("Per ", time_period , " patient counts with mental health related ICD codes")) + 
+  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)
+thpd_psyCount
+```
+
+#### Both results together
+
+```{r}
+count_icd_all_combined <-  count_icd_all %>%
+  group_by(time_p, period) %>% 
+  summarise(count_psy_all = sum(count_psy), 
+            count_icd_all = sum( count_icd ))
+
+thpd_psyCount_combined <- count_icd_all_combined %>%
+  pivot_longer(names_to = "counts", cols=c(count_psy_all, count_icd_all)) %>%
+  mutate( counts = ifelse( counts == "count_psy_all", "Psy", "Total")) %>%
+  ggplot(aes(x = time_p, y = value, fill = counts, color = counts)) +
+  geom_point() +
+  geom_vline(xintercept = as.Date(pandemic_start_date),
+             linetype = "dashed") +
+  scale_fill_manual(values = cbPalette) +
+  scale_color_manual(values = cbPalette) +
+  labs(y = "Counts",
+       x = paste0("Date (by ", time_period,")"),
+       title = paste0("Per ", time_period , " patient counts with mental health related ICD codes")) + 
+  geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)
+thpd_psyCount_combined
+```
+
+## Ratio patients with vs without psychiatric conditions
+```{r}
+ratio_psy_non_psy <- ratios_patients_with_without_psy_all %>%
+  ggplot(aes(x = time_p, y = ratio, fill = period, color = period)) +
+  geom_point() +
+  facet_grid(siteid~.) +
+  geom_vline(xintercept = as.Date(pandemic_start_date),
+             linetype = "dashed") +
+  scale_fill_manual(values = cbPalette) +
+  scale_color_manual(values = cbPalette) +
+  labs(y = "Ratio (with psy conditions/without psy conditions)",
+       x = paste0("Date (by ", time_period,")"),
+       title = paste0("Ratio patient with vs without psychiatric conditions ( per ", time_period,")"))
+
+ratio_psy_non_psy
+```
+
 ## Percentage: patients with psychiatric conditions / total number of patients
@@
 ### By site
+
+```{r}
+percentage_psy_non_psy <- ratios_patients_with_without_psy_all %>%
+  ggplot(aes(x = time_p, y = percentage, color = period)) +
+  facet_grid(siteid~.) +
+  geom_point() +
+  geom_vline(xintercept = as.Date(pandemic_start_date),
+             linetype = "dashed") +
+  scale_fill_manual(values = cbPalette) +
+  scale_color_manual(values = cbPalette) +
+  labs(y = "Percentage",
+       x = paste0("Date (by ", time_period,")"),
+       title = paste0("Percentage patient with vs total population ( per ", time_period,")")) +
+  guides(color = guide_legend(title="Patient diagnoses types"))
+
+percentage_psy_non_psy
+```
 
@@
 ### Summary of the codes used per site 
-```{r, eval=FALSE}
+```{r}
 mental_codes_qc_output_all <- mental_codes_qc_output_all %>%
@@
 getSitesWithEachCode <- function(row) {
-  sites <- c( row['BCH'],  row['Lurie'], row['UMICH'], row['FRBDX'], row['APHP'])
+  sites <- c( row['BCH'],  row['Lurie'], row['UMICH'], row['FRBDX'], row['CHOP'])
   sites <- sites[!is.na( sites ) ]
@@
 summary_codes <- summary_codes[, c("disorder_group", "description", "concept_code", 
-                                   "BCH", "Lurie", "UMICH",  "FRBDX", "APHP") ]                        
+                                   "BCH", "CHOP","Lurie", "UMICH",  "FRBDX") ]                        
 summary_codes$sitesCombination <- apply(summary_codes, 1, FUN = getSitesWithEachCode)
 summary_codes <- summary_codes[
-  with(summary_codes, order(BCH, FRBDX, Lurie, UMICH, APHP, decreasing = TRUE)),
+  with(summary_codes, order(BCH, CHOP, FRBDX, Lurie, UMICH, decreasing = TRUE)),
 ]
-print( paste0("There are in total ", nrow( summary_codes[ summary_codes$sitesCombination == "BCH - Lurie - UMICH - FRBDX - APHP", ]), " codes in common across all sites"))
+print( paste0("There are in total ", nrow( summary_codes[ summary_codes$sitesCombination == "BCH - CHOP - Lurie - UMICH - FRBDX", ]), " codes in common across all sites"))
 
@@
   UMICH_concepts = df[ df$siteid == "UMICH", "concept_code"]
-  APHP_concepts = df[ df$siteid == "APHP", "concept_code"]
+  CHOP_concepts = df[ df$siteid == "CHOP", "concept_code"]
 
@@
           UMICH = UMICH_concepts$concept_code, 
-          APHP  = APHP_concepts$concept_code
+          CHOP  = CHOP_concepts$concept_code
           )
@@
 Represent this in 16 pannels, if we put all together, due to the different scales we will not be able to really see the differences between results. 
-```{r, eval=FALSE}
+```{r}
+mental_codes_qc_output_all <- mental_codes_qc_output_all %>%
+    mutate( concept_code = gsub("[.]", "", concept_code ) ) %>%
+  unique()
+
 numberOfCodesPerDisorderGroup <- mental_codes_qc_output_all %>%
@@
 
+positions <- c("BCH", "CHOP", "UMICH", "Lurie", "FRBDX")
+
 numberOfCodesPerDisorderGroup %>%
   mutate( disorder_group = swr( disorder_group ) ) %>%
-  ggplot( aes( x= reorder( siteid, icd_codes_count), y = icd_codes_count)) +
+  inner_join( country_map ) %>%
+  ggplot( aes( x= siteid, y = icd_codes_count, fill = country )) +
   geom_bar( stat = "identity") +
+  scale_fill_manual("Country of each site", values = c("USA" = "navy", "France" = "orange")) +
   facet_wrap( ~ disorder_group, scales = "free" ) +
   labs(x ="site identifier", y = "ICD dictionary (number of ICD codes)") +
+  scale_x_discrete(limits = positions) +
   theme(axis.text.x=element_text(angle=45,hjust=1, size = 6), 

### 2022-01-27 metanalysis_visualization/manuscript_Figures.Rmd
new notebook with potential figures for the manuscript

--- 2022-01-26 metanalysis_visualization/visualization.Rmd
@@
 ---
-title: "Results visualization across sites"
+title: "Visualizations for the manuscript"
 output:
@@
 ---
-
-## Setting up the R environment
 
@@
 }
-listOfPackages <- c("tidyverse", "tidyr","devtools","RColorBrewer", "knitr", "kableExtra", "tsModel", "gridExtra", "dplyr", "metafor", "meta", "viridis", "UpSetR")
+listOfPackages <- c("tidyverse", "tidyr","devtools","RColorBrewer", "knitr", "kableExtra", "tsModel", "gridExtra", "dplyr", "metafor", "meta", "viridis", "UpSetR", "scales")
 paket(listOfPackages)
@@
 install_github("jokergoo/ComplexHeatmap")
+install_github("teunbrand/ggh4x")
+
 library( ComplexHeatmap)
-```
-
-### R session information
-
-```{r, results='hide'}
-sessionInfo()
+library( ggh4x )
 ```
@@
 
-## Percentage: patients with psychiatric conditions / total number of patients
-
-### By site
+### Table 1
 
 ```{r}
-ratios_patients_with_without_psy_all %>%
-  filter(time_p <= history_end_date &
-         time_p >= history_start_date ) %>%
-  ggplot( aes(x=time_p, y=percentage, 
-              group = siteid, color = siteid))+
-  geom_point( aes(shape=siteid, color=siteid) )+
-    geom_line() +
-  geom_vline(xintercept = as.Date(pandemic_start_date),
-             linetype = "dashed") +
-  scale_fill_viridis(discrete = TRUE, alpha=0.6) +
-    theme(
-      legend.position="bottom",
-      plot.title = element_text(size=11)
-    ) +
-  scale_x_date(date_labels = "%b-%d-%Y", breaks = "month") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6)) +
-    ggtitle("Percentage: patients with psychiatric conditions / total number of patients") +
-    xlab("period (months)")
+sitesWithTable1 <- outputFiles[c(1,3:5)]
 
-ratios_patients_with_without_psy_all %>%
-  filter(time_p <= history_end_date &
-         time_p >= history_start_date ) %>%
-  ggplot( aes(x=time_p, y=ratio, 
-              group = siteid, color = siteid))+
-    facet_wrap("siteid") +
-  geom_point( aes(shape=siteid, color=siteid) ) + 
-  geom_line() +
-  geom_vline(xintercept = as.Date(pandemic_start_date),
-             linetype = "dashed") +
-  scale_fill_viridis(discrete = TRUE, alpha=0.6) +
-    theme(
-      legend.position="bottom",
-      plot.title = element_text(size=11)
-    ) +
-  scale_x_date(date_labels = "%b-%d-%Y", breaks = "month") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6)) +
-  guides(color = guide_legend(override.aes = list(alpha = 1))) +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6)) +
-    ggtitle("Ratio: patients with vs. patients without psychiatric conditions") +
-    xlab("period (months)") +
-  theme(strip.text = element_text(size = 10),
-        axis.text.x = element_text(size = 8), 
-        axis.text.y = element_text(size = 8), 
-        axis.title.x = element_text(size = 8),
-        axis.title.y = element_text(size = 8))
+for( i in 1:length( sitesWithTable1 )){
+  print(i)
+  load(paste0( outputDir, sitesWithTable1[i]))
+  if( i == 1 ){
+    table1_all <- table1
+  } else {
+    table1_all <- rbind( table1_all, table1)
+  }
+  rm( table1)
+}
 ```
 
-### By Country
+### Sex Summary
+```{r}
+sexInfo_per_country <- table1_all %>%
+  filter( Category == "Sex") %>%
+  mutate ( counts_before = as.numeric(sapply(strsplit( as.character(before_pandemic), " "), '[', 1)), 
+           counts_during = as.numeric(sapply(strsplit( as.character(during_pandemic), " "), '[', 1))) %>%
+  select( name, siteid, counts_before, counts_during ) %>%
+  inner_join( country_map ) %>%
+  group_by( name, country ) %>%
+  summarise( before = sum( counts_before), 
+             during = sum( counts_during )) %>%
+  pivot_longer(names_to = "period", cols=c(before, during))
 
-#### Showing each independent site 
-```{r}
-ratios_patients_with_without_psy_all_country <- left_join( ratios_patients_with_without_psy_all, country_map )
-ratios_patients_with_without_psy_all_country %>%
-    filter(time_p <= history_end_date &
-         time_p >= history_start_date )  %>%
-  ggplot(aes(x = time_p, y = percentage, fill = siteid, color = siteid)) +
-  geom_point( aes(shape=siteid, color=siteid)  ) +
-  geom_line() +
-  facet_grid(country~.) +
-  geom_vline(xintercept = as.Date(pandemic_start_date),
-             linetype = "dashed") +
-  scale_fill_manual(values = cbPalette) +
-  scale_color_manual(values = cbPalette) +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6)) +
-  scale_x_date(date_labels = "%b-%d-%Y", breaks = "month", limits = c( history_start_date, history_end_date  )) +
-  labs(y = "Ratio (with psy conditions/without psy conditions)",
-       x = paste0("Date (by ", time_period,")"),
-       title = paste0("Percentage: patients with psychiatric conditions / total number of patients ( per ", time_period,")"))
-```
+sexInfo_all <- table1_all %>%
+  filter( Category == "Sex") %>%
+  mutate ( counts_before = as.numeric(sapply(strsplit( as.character(before_pandemic), " "), '[', 1)), 
+           counts_during = as.numeric(sapply(strsplit( as.character(during_pandemic), " "), '[', 1))) %>%
+  select( name, siteid, counts_before, counts_during ) %>%
+  inner_join( country_map ) %>%
+  group_by( name ) %>%
+  summarise( before = sum( counts_before), 
+             during = sum( counts_during )) %>%
+  pivot_longer(names_to = "period", cols=c(before, during)) %>%
+  mutate( country = 'ALL' ) %>%
+  select( name, country, period, value ) %>%
+  rbind( sexInfo_per_country)
 
-#### Aggregating the sites in each country
-
-```{r}
-ratios_patients_with_without_psy_all_country <- left_join( ratios_patients_with_without_psy_all, country_map )
-
-ratios_patients_with_without_psy_aggregated <- ratios_patients_with_without_psy_all_country %>%
-  group_by( time_p, country ) %>%
-  mutate( no_psy = sum( count_no_psy_patients ), 
-          psy = sum( count_psy_patients ), 
-          total = sum(count_no_psy_patients) + sum(count_psy_patients), 
-          ratio_by_country = psy / no_psy, 
-          perc_by_country = 100*(psy / total) ) %>%
-  select( time_p, no_psy, psy, total, ratio_by_country, perc_by_country, country)
-
-ratios_patients_with_without_psy_aggregated %>%
-  filter(time_p <= history_end_date &
-         time_p >= history_start_date ) %>%
-  ggplot( aes(x=time_p, y=perc_by_country, 
-              group = country, color = country))+
-  geom_point( aes(shape=country, color=country) )+
-    geom_line() +
-  geom_vline(xintercept = as.Date(pandemic_start_date),
-             linetype = "dashed") +
-  scale_colour_viridis_d() +
-    theme(
-      legend.position="bottom",
-      plot.title = element_text(size=11)
-    ) +
-  scale_x_date(date_labels = "%b-%d-%Y", breaks = "month") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6)) +
-    ggtitle(paste0("Percentage: patients with psychiatric conditions / total number of patients ( per ", time_period,") \n (aggregated by country)")) +
-    xlab("period (months)") +
-    ylab("percentage (%)")
-```
-
-## Ratio: patients with vs. patients without psychiatric conditions
-
-### By site
-```{r}
-ratios_patients_with_without_psy_all %>%
-  filter(time_p <= history_end_date &
-         time_p >= history_start_date ) %>%
-  ggplot( aes(x=time_p, y=ratio, 
-              group = siteid, color = siteid))+
-  geom_point( aes(shape=siteid, color=siteid) ) + 
-  geom_line() +
-  geom_vline(xintercept = as.Date(pandemic_start_date),
-             linetype = "dashed") +
-  scale_fill_viridis(discrete = TRUE, alpha=0.6) +
-    theme(
-      legend.position="bottom",
-      plot.title = element_text(size=11)
-    ) +
-  scale_x_date(date_labels = "%b-%d-%Y", breaks = "month") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6)) +
-    ggtitle("Ratio: patients with vs. patients without psychiatric conditions") +
-    xlab("period (months)")
+sexInfo_all %>%
+  ggplot( aes( x= name, y = value, fill = period )) +
+  geom_bar( stat = "identity", position = "dodge") +
+  facet_wrap("country", scales = "free_y") + 
+  labs(x ="", y = "number of patients") +
+  theme(axis.text.x=element_text(hjust=1, size = 8), 
+        strip.text = element_text(size = 10, face = "bold"),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), strip.background = element_blank(),
+        panel.border = element_rect(colour = "black", fill = NA))+ 
+  scale_fill_manual("Pandemic \n period", values = c("before" = "grey", "during" = "orange"))
 
@@
 
-### By Country
+```{r}
+patients_hospitalized_agg_sex_total <- patients_hospitalized_agg_sex_perc_all %>%
+  select( - percentage ) %>%
+  group_by( time_p, sex ) %>%
+  mutate( total_cases = sum( count ), 
+          sex_cases = sum( count_sex ), 
+          percentage = 100* sex_cases / total_cases, 
+          country = "ALL")
 
-#### Showing each site per country 
-```{r}
-ratios_patients_with_without_psy_all_country %>%
-    filter(time_p <= history_end_date &
-         time_p >= history_start_date ) %>%
-  ggplot(aes(x = time_p, y = ratio, fill = siteid, color = siteid)) +
-  geom_point( aes(shape=siteid, color=siteid)  ) +
-  geom_line() +
-  facet_grid(country~.) +
-  geom_vline(xintercept = as.Date(pandemic_start_date),
-             linetype = "dashed") +
-  scale_fill_manual(values = cbPalette) +
-  scale_color_manual(values = cbPalette) +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6)) +
-  scale_x_date(date_labels = "%b-%d-%Y", breaks = "month", limits = c( history_start_date, history_end_date  )) +
-  labs(y = "Ratio (with psy conditions/without psy conditions)",
-       x = paste0("Date (by ", time_period,")"),
-       title = paste0("Ratio patient with vs without psychiatric conditions ( per ", time_period,")"))
+patients_hospitalized_agg_sex_total %>%
+  ggplot( aes( x= as.Date( time_p ), y = percentage, group = sex )) +
+  geom_line( aes( linetype = sex , color = sex)) +
+  geom_point( aes( shape = sex, color = sex )) +
+  scale_colour_manual(values=c("#FF6600","#339999")) +
+  scale_fill_manual(values=c("#FF6600","#339999"))+
+  geom_vline(xintercept = as.Date(pandemic_start_date), linetype = "dashed")+
+  labs(x ="", y = "percentage of patients (%)")+
+  scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0))+
+  theme(axis.text.x=element_text(angle = 45, hjust=1, size = 6),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.border = element_rect(colour = "black", fill = NA))
+
+patients_hospitalized_agg_sex_country <- patients_hospitalized_agg_sex_perc_all %>%
+  select( - percentage ) %>%
+  inner_join( country_map ) %>%
+  group_by( time_p, sex, country ) %>%
+  mutate( total_cases = sum( count ), 
+          sex_cases = sum( count_sex ), 
+          percentage = 100* sex_cases / total_cases) %>%
+  rbind( patients_hospitalized_agg_sex_total)
+
+patients_hospitalized_agg_sex_country %>%
+  ggplot( aes( x= as.Date( time_p ), y = percentage, group = sex )) +
+  geom_line( aes(linetype = sex, color = sex )) +
+  geom_point( aes( shape= sex, color = sex )) +
+  scale_colour_manual(values=c("#FF6600","#339999")) +
+  scale_fill_manual(values=c("#FF6600","#339999"))+
+  #scale_y_continuous(breaks = seq(0, 30, by = 1)) +
+  facet_wrap2("country", axes = "all") + 
+  geom_vline(xintercept = as.Date(pandemic_start_date), linetype = "dashed") +
+  labs(x ="", y = "percentage of patients (%)")+
+  scale_x_date(breaks = "2 month", labels=date_format("%b-%y"), expand=c(0.03,0)) +
+  theme(axis.text.x=element_text(angle = 45, hjust=1, size = 6), 
+        strip.text = element_text(size = 10, face = "bold"),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), strip.background = element_blank(),
+        panel.border = element_rect(colour = "black", fill = NA), 
+        legend.position="bottom")
 ```
 
-#### Aggregating the values of the sites per country
-
-```{r}
-ratios_patients_with_without_psy_aggregated %>%
-  filter(time_p <= history_end_date &
-         time_p >= history_start_date ) %>%
-  ggplot( aes(x=time_p, y=ratio_by_country, 
-              group = country, color = country))+
-  geom_point( aes(shape=country, color=country) )+
-    geom_line() +
-  geom_vline(xintercept = as.Date(pandemic_start_date),
-             linetype = "dashed") +
-  scale_colour_viridis_d() +
-    theme(
-      legend.position="bottom",
-      plot.title = element_text(size=11)
-    ) +
-  scale_x_date(date_labels = "%b-%d-%Y", breaks = "month") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6)) +
-    ggtitle("Ratios: patients with psychiatric conditions / without psychiatric conditions \n (aggregated by country)") +
-    xlab("period (months)") +
-    ylab("ratio")
-```
-
-## Meta-analysis percentage
-
-```{r, results='hide', fig.show='hide'}
-#split this into before and after pandemic and run it twice (one for each subset)
-
-#estimate the summary effect size
-ratios_patients_with_without_psy_all$total <- ratios_patients_with_without_psy_all$count_no_psy_patients + ratios_patients_with_without_psy_all$count_psy_patients
-
-ratios_bp <- ratios_patients_with_without_psy_all %>%
-  filter( period == "before_pandemic")
-
-ratios_dp <- ratios_patients_with_without_psy_all %>%
-  filter( period == "during_pandemic")
-
-metainputdata <- ratios_dp
-#RR no transformation
-#PLO the logit transformation
-#PFT the double arcsine transformation
-ies=escalc(xi=count_psy_patients, ni=total, data=metainputdata, measure="PR")
-summary(ies$vi)
-
-# pool the individual effect size
-# DL random effects using the DerSimonian-Laird estimator
-#REML random effects using the restricted maximum-likelihood estimator
-pes = rma(yi, vi, data=ies, method="REML")
-print(pes)
-confint(pes)
-
-pes.summary=metaprop(count_psy_patients, total, as.character(time_p),
-                     data=metainputdata, 
-                     sm="PRAW")
-forest(pes.summary,layout = "JAMA")
-```
+## Interrupted Time Series Analysis
 
@@
 
-## Percentage of patients by disorder group
+### Disorder group 
+```{r}
+perc_disorder_group_total <- perc_disorder_group_all %>%
+  select( - percentage_dg ) %>%
+  group_by( time_p, disorder_group ) %>%
+  mutate( total_cases = sum( count_icd ), 
+          dg_cases = sum( count_dg ), 
+          percentage = 100* dg_cases / total_cases, 
+          country = "ALL")
 
-```{r}
-perc_disorder_group_all %>%
-    filter(time_p <= history_end_date &
-         time_p >= history_start_date &
-           disorder_group %in% c("Anxiety Disorders", "Depressive Disorders", "Suicide or Self-Injury" )) %>%
-  ggplot(aes(x = time_p, y = percentage_dg, fill = siteid, color = siteid)) +
-  geom_point( aes(shape=siteid, color=siteid)  ) +
-  geom_line() +
-  facet_grid(. ~ disorder_group) +
-  geom_vline(xintercept = as.Date(pandemic_start_date),
-             linetype = "dashed") +
-  scale_fill_manual(values = cbPalette) +
-  scale_color_manual(values = cbPalette) +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6)) +
-  scale_x_date(date_labels = "%b-%d-%Y", breaks = "month", limits = c( history_start_date, history_end_date  )) +
-  labs(y = "Percentage of patients (%)",
-       x = paste0("Date (by ", time_period,")"),
-       title = paste0("Percentage of patients by disorder group \n ( per ", time_period,")"))
+perc_disorder_group_total %>%
+  ggplot( aes( x= as.Date( time_p ), y = percentage, group = disorder_group )) +
+  geom_line( aes(color = disorder_group )) +
+  geom_point( aes( color = disorder_group )) +
+  geom_vline(xintercept = as.Date(pandemic_start_date), linetype = "dashed")+
+  labs(x ="", y = "percentage of patients (%)")+
+  scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0))+
+  theme(axis.text.x=element_text(angle = 45, hjust=1, size = 6),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.border = element_rect(colour = "black", fill = NA))
 
-perc_disorder_group_all %>%
-    filter(time_p <= history_end_date &
-         time_p >= history_start_date &
-          #disorder_group %in% c("Anxiety Disorders", "Depressive Disorders", "Suicide or Self-Injury" )) %>%
-          disorder_group %in% c("Suicide or Self-Injury" )) %>%
-  ggplot(aes(x = time_p, y = percentage_dg, fill = siteid, color = siteid)) +
-  geom_point( aes(shape=siteid, color=siteid)  ) +
-  geom_line() +
-  geom_vline(xintercept = as.Date(pandemic_start_date),
-             linetype = "dashed") +
-  scale_fill_manual(values = cbPalette) +
-  scale_color_manual(values = cbPalette) +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6)) +
-  scale_x_date(date_labels = "%b-%d-%Y", breaks = "month", limits = c( history_start_date, history_end_date  )) +
-  labs(y = "Percentage of patients (%)",
-       x = paste0("Date (by ", time_period,")"),
-       title = paste0("Percentage of patients with Suicide or Self-Injury \n ( per ", time_period,")"))
+perc_disorder_group_country <- perc_disorder_group_all %>%
+  select( - percentage_dg ) %>%
+  inner_join( country_map ) %>%
+  group_by( time_p, disorder_group, country ) %>%
+  mutate( total_cases = sum( count_icd ), 
+          dg_cases = sum( count_dg ), 
+          percentage = 100* dg_cases / total_cases ) %>%
+  rbind( perc_disorder_group_total)
+
+perc_disorder_group_country %>%
+  filter( disorder_group %in% c("Anxiety Disorders", "Depressive Disorders", "Suicide or Self-Injury", 
+  "Mental Health Symptoms", "Obsessive-Compulsive and Related Disorders" )) %>%
+  ggplot( aes( x= as.Date( time_p ), y = percentage, group = disorder_group )) +
+  geom_line( aes(linetype = disorder_group, color = disorder_group )) +
+  geom_point( aes( shape= disorder_group, color = disorder_group )) +
+  scale_y_continuous(breaks = seq(0, 30, by = 1)) +
+  facet_wrap2("country", axes = "all") + 
+  geom_vline(xintercept = as.Date(pandemic_start_date), linetype = "dashed") +
+  labs(x ="", y = "percentage of patients (%)")+
+  scale_x_date(breaks = "2 month", labels=date_format("%b-%y"), expand=c(0.03,0)) +
+  theme(axis.text.x=element_text(angle = 45, hjust=1, size = 6), 
+        strip.text = element_text(size = 10, face = "bold"),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), strip.background = element_blank(),
+        panel.border = element_rect(colour = "black", fill = NA), 
+        legend.position="bottom")
+
+perc_disorder_group_country %>%
+  filter( disorder_group %in% c("Anxiety Disorders", "Depressive Disorders", "Suicide or Self-Injury", 
+  "Mental Health Symptoms", "Obsessive-Compulsive and Related Disorders" )) %>%
+  ggplot( aes( x= as.Date( time_p ), y = percentage, group = country )) +
+  geom_line( aes(linetype = country, color = country )) +
+  geom_point( aes( shape= country, color = country )) +
+  scale_y_continuous(breaks = seq(0, 30, by = 5)) +
+  facet_wrap2("disorder_group", axes = "all") + 
+  geom_vline(xintercept = as.Date(pandemic_start_date), linetype = "dashed") +
+  labs(x ="", y = "percentage of patients (%)")+
+  scale_x_date(breaks = "2 month", labels=date_format("%b-%y"), expand=c(0.03,0)) +
+  scale_color_manual( values = cbPalette ) + 
+  theme(axis.text.x=element_text(angle = 45, hjust=1, size = 6), 
+        strip.text = element_text(size = 10, face = "bold"),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), strip.background = element_blank(),
+        panel.border = element_rect(colour = "black", fill = NA), 
+        legend.position="bottom")
+
 ```
 
-## Sex ratio per site
-
+### Estimate number of ICD codes per site and disorder group
+Represent this in 16 pannels, if we put all together, due to the different scales we will not be able to really see the differences between results. 
 ```{r}
-sex_ratio_by_site <- patients_hospitalized_agg_sex_perc_all %>%
-  filter( sex == "female") %>%
-  group_by( time_p, siteid ) %>%
-  mutate( female_count = count_sex, 
-          male_count = count - count_sex, 
-          sex_ratio_by_site = female_count / male_count ) %>%
-  select( time_p, female_count, male_count, sex_ratio_by_site, siteid)
-
-sex_ratio_by_site %>%
-    filter(time_p <= history_end_date &
-         time_p >= history_start_date) %>%
-  ggplot(aes(x = time_p, y = sex_ratio_by_site, fill = siteid, color = siteid)) +
-  geom_point( aes(shape=siteid, color=siteid)  ) +
-  geom_line() +
-  geom_vline(xintercept = as.Date(pandemic_start_date),
-             linetype = "dashed") +
-   geom_hline(yintercept = 1,
-             linetype = "dashed", color = "red") +
-  scale_fill_manual(values = cbPalette) +
-  scale_color_manual(values = cbPalette) +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6)) +
-  scale_x_date(date_labels = "%b-%d-%Y", breaks = "month", limits = c( history_start_date, history_end_date  )) +
-  labs(y = "sex ratio (female/male)",
-       x = paste0("Date (by ", time_period,")"),
-       title = paste0("Sex ratio of patients with mental disorders \n ( per ", time_period,")"))
-```
-
-### Summary of the codes used per site 
-```{r, eval=FALSE}
 mental_codes_qc_output_all <- mental_codes_qc_output_all %>%
@@
 
-summary_codes <- mental_codes_qc_output_all %>%
-  select( disorder_group, description, concept_code, distinct_patients, siteid ) %>%
-  spread( siteid, distinct_patients)
-
-## add a column with the sites combination 
-getSitesWithEachCode <- function(row) {
-  sites <- c( row['BCH'],  row['Lurie'], row['UMICH'], row['FRBDX'], row['APHP'])
-  sites <- sites[!is.na( sites ) ]
-  siteVal <- paste0( names( sites ), collapse = " - ")
-  return( siteVal)
-}
-summary_codes <- summary_codes[, c("disorder_group", "description", "concept_code", 
-                                   "BCH", "Lurie", "UMICH",  "FRBDX", "APHP") ]                        
-summary_codes$sitesCombination <- apply(summary_codes, 1, FUN = getSitesWithEachCode)
-summary_codes <- summary_codes[
-  with(summary_codes, order(BCH, FRBDX, Lurie, UMICH, APHP, decreasing = TRUE)),
-]
-print( paste0("There are in total ", nrow( summary_codes[ summary_codes$sitesCombination == "BCH - Lurie - UMICH - FRBDX - APHP", ]), " codes in common across all sites"))
-
-write.table( summary_codes, file="summary_codes.txt", col.names = TRUE, row.names = FALSE, sep = "\t", quote = FALSE )
-
-#### upSetR 
-upsetForDisorderGroup <- function( df, disorderGroup = "ALL"){
-  if( disorderGroup != "ALL"){
-     df <- df %>%
-       filter( disorder_group == disorderGroup ) 
-  }
-
-  BCH_concepts = df[ df$siteid == "BCH", "concept_code"]
-  FRBDX_concepts = df[ df$siteid == "FRBDX", "concept_code"]
-  Lurie_concepts = df[ df$siteid == "Lurie", "concept_code"]
-  UMICH_concepts = df[ df$siteid == "UMICH", "concept_code"]
-  APHP_concepts = df[ df$siteid == "APHP", "concept_code"]
-
-  concepts = list( 
-          BCH = BCH_concepts$concept_code,
-          FRBDX = FRBDX_concepts$concept_code, 
-          Lurie = Lurie_concepts$concept_code, 
-          UMICH = UMICH_concepts$concept_code, 
-          APHP  = APHP_concepts$concept_code
-          )
-  output <- upset(fromList(concepts), order.by = "freq")
-  return( output )
-}
-
-upsetForDisorderGroup( df = mental_codes_qc_output_all, disorderGroup = "ALL")
-```
-
-### Estimate number of ICD codes per site and disorder group
-Represent this in 16 pannels, if we put all together, due to the different scales we will not be able to really see the differences between results. 
-```{r, eval=FALSE}
 numberOfCodesPerDisorderGroup <- mental_codes_qc_output_all %>%
@@
 
+positions <- c("BCH", "CHOP", "UMICH", "Lurie", "FRBDX")
+
 numberOfCodesPerDisorderGroup %>%
   mutate( disorder_group = swr( disorder_group ) ) %>%
-  ggplot( aes( x= reorder( siteid, icd_codes_count), y = icd_codes_count)) +
+  inner_join( country_map ) %>%
+  ggplot( aes( x= siteid, y = icd_codes_count, fill = country )) +
   geom_bar( stat = "identity") +
+  scale_fill_manual("Country of each site", values = c("USA" = "navy", "France" = "orange")) +
   facet_wrap( ~ disorder_group, scales = "free" ) +
   labs(x ="site identifier", y = "ICD dictionary (number of ICD codes)") +
+  scale_x_discrete(limits = positions) +
   theme(axis.text.x=element_text(angle=45,hjust=1, size = 6), 
         strip.text = element_text(size = 7), 
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) 
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
+        legend.position="bottom") 
 ```
+
+### R session information
+
+```{r, results='hide'}
+sessionInfo()
+```

### 2022-02-08 psy_peds_3_CHOP.Rmd
the table 1 was added to CHOP pipeline

@@
 ```
+## Table 1
+
+```{r}
+table1 <- read.csv(file = "../output/CHOP/Table 1.csv")
+table1 %>% 
+   kable( booktabs = T) %>%
+  kableExtra::kable_styling(latex_options = "striped") 
+```
 
@@
 mental_codes_qc_output$siteid = site
+table1$siteid = site
 
@@
       mental_codes_qc_output,
+      table1,
       file = paste0("../output/CHOP/peds_psy_aggResults_", site, ".RData"))

### 2022-02-08 metanalysis_visualization/manuscript_Figures.Rmd
integration of table 1

@@
       select( disorder_group, description, concept_code, distinct_patients, distinct_observations, siteid )
+    table1_all <- table1
   } else {
@@
     mental_codes_qc_output_all <- rbind( mental_codes_qc_output_all, mental_codes_qc_output)
+    table1_all <- rbind( table1_all, table1)
     }
@@
      patient_count_psy_period_psy_clear, bootstrapped_coefficients_df,
-     bootstrapped_fitted_df, mental_codes_qc_output)
+     bootstrapped_fitted_df, mental_codes_qc_output, table1)
 }
@@
 ```{r}
-sitesWithTable1 <- outputFiles[c(1,3:5)]
+sex_table1_all <- table1_all %>%
+  filter( Category == "Sex" & name != "Total") %>%
+  mutate( before_counts = sapply(strsplit( before_pandemic, " "), '[', 1), 
+          during_counts = sapply(strsplit( during_pandemic, " "), '[', 1)) %>%
+  group_by( name ) %>%
+  summarise( before_pandemic_count = sum( as.numeric( before_counts )), 
+             during_pandemic_count = sum( as.numeric( during_counts ))) %>%
+  mutate( Category = "Sex") %>%
+  select( Category, name, before_pandemic_count, during_pandemic_count )
 
-for( i in 1:length( sitesWithTable1 )){
-  print(i)
-  load(paste0( outputDir, sitesWithTable1[i]))
-  if( i == 1 ){
-    table1_all <- table1
-  } else {
-    table1_all <- rbind( table1_all, table1)
-  }
-  rm( table1)
-}
+sex_table1_all <- rbind( sex_table1_all, c("Total", "",sum( sex_table1_all$before_pandemic_count), sum( sex_table1_all$during_pandemic_count)))
+
+dg_table1_all <- table1_all %>%
+  filter( Category == "Disorder groups" & name != "Total") %>%
+  mutate( before_counts = sapply(strsplit( before_pandemic, " "), '[', 1), 
+          during_counts = sapply(strsplit( during_pandemic, " "), '[', 1)) %>%
+  group_by( name ) %>%
+  summarise( before_pandemic_count = sum( as.numeric( before_counts ), na.rm = TRUE), 
+             during_pandemic_count = sum( as.numeric( during_counts ), na.rm = TRUE)) %>%
+  mutate( Category = "Disorder groups") %>%
+  select( Category, name, before_pandemic_count, during_pandemic_count )
+
+table1 <- rbind( sex_table1_all, dg_table1_all )
+
+table1 %>%
+  kable( booktabs = T) %>%
+  kableExtra::kable_styling(latex_options = "striped") %>%
+  kableExtra::group_rows(index = setNames(rle(table1$Category)[[1]], rle(table1$Category)[[2]]))
 ```

### 2022-02-08 psy_peds_3_CHOP.Rmd
integration of table 1

@@
 table1 <- read.csv(file = "../output/CHOP/Table 1.csv")
+
+sex <- table1[ c(2:4), ] %>%
+  mutate( Category = "Sex", 
+          name = c("female", "male", "Total"),           
+          before_pandemic = paste0( Before.Pandemic, " (", Before.Pandemic.1, " %)"), 
+          during_pandemic = paste0( During.Pandemic, " (", During.Pandemic.1, " %)"), 
+          pvalue = NA) %>%
+  select( Category, name, before_pandemic, during_pandemic, pvalue)
+
+age <- table1[ c(6), ] %>%
+  mutate( Category = "Age", 
+          name = "Age", 
+          before_pandemic = paste0( Before.Pandemic, " (sd: ", Before.Pandemic.1, ")"), 
+          during_pandemic = paste0( During.Pandemic, " (sd: ", During.Pandemic.1, ")"), 
+          pvalue = NA
+          ) %>%
+  select( Category, name, before_pandemic, during_pandemic, pvalue)
+
+t.test2 <- function(m1, m2, s1, s2, n1, n2) {
+  # t.test computed from sample statistics, ie using mean, standard deviation and count value of the samples
+  # useful in case we can only retrieve aggregated stats from the sites
+  # ref: https://stats.stackexchange.com/questions/30394/how-to-perform-two-sample-t-tests-in-r-by-inputting-sample-statistics-rather-tha
+  # and https://en.wikipedia.org/wiki/Student%27s_t-test?oldformat=true#Equal_or_unequal_sample_sizes,_unequal_variances_(sX1_%3E_2sX2_or_sX2_%3E_2sX1)
+  se <- sqrt( (s1^2/n1) + (s2^2/n2) )
+  # welch-satterthwaite df
+  df <- ( (s1^2/n1 + s2^2/n2)^2 )/( (s1^2/n1)^2/(n1-1) + (s2^2/n2)^2/(n2-1) )
+  t <- (m1-m2)/se
+  dat <- c(m1-m2, se, t, 2*pt(-abs(t),df))
+  names(dat) <- c("Difference of means", "Std Error", "t", "p-value")
+  return(dat)
+}
+
+age$pvalue <- t.test2(m1 = 14.51,
+                    m2 = 14.56,
+                    s1 = 1.89,
+                    s2 = 1.84,
+                    n1 = 2592,
+                    n2 = 2419) %>%
+  `[[`("p-value")
+
+disorder_groups_t1 <- table1[ c(8:24), ] %>%
+  mutate( Category = "Disorder groups", 
+          name = X,           
+          before_pandemic = paste0( Before.Pandemic, " (", Before.Pandemic.1, " %)"), 
+          during_pandemic = paste0( During.Pandemic, " (", During.Pandemic.1, " %)"), 
+          pvalue = NA
+          ) %>%
+  select( Category, name, before_pandemic, during_pandemic, pvalue)
+
+table1 <- rbind( sex, age, disorder_groups_t1 )
+
 table1 %>% 
   kable( booktabs = T) %>%
-  kableExtra::kable_styling(latex_options = "striped") 
+  kableExtra::kable_styling(latex_options = "striped") %>%
+  kableExtra::group_rows(index = setNames(rle(table1$Category)[[1]], rle(table1$Category)[[2]]))
 ```

### 2022-02-27 metanalysis_visualization/manuscript_Figures.Rmd
CCHMC added
add APHP and CCHMC to the visualization notebook

@@
 df_coefficients_boot <- bootstrapped_coefficients_all %>% 
-  left_join(coefficients_df, by = "siteid") %>% 
-  filter(siteid != "APHP")
+  left_join(coefficients_df, by = "siteid")
 
@@
 
-positions <- c("BCH", "CHOP", "UMICH", "Lurie", "FRBDX")
+positions <- c("BCH", "CCHMC","CHOP", "UMICH", "Lurie", "FRBDX", "APHP")
 

### 2022-03-04 metanalysis_visualization/manuscript_Figures.Rmd
code added to get ITS per country. pvalue needs to be added

@@
 
+Aggregated by Country
+
+```{r}
+patient_count_psy_period_psy_country <- patient_count_psy_period_psy_all %>% 
+  left_join(country_map, by = "siteid") %>%
+  group_by(country, time_p) %>% 
+  mutate(total_patients = count_no_psy_patients +count_psy_patients) %>%
+  summarise(psy_cases = sum(count_psy_patients),
+            total_cases = sum(total_patients), 
+            percentages = round(100*psy_cases/total_cases,2), 
+            time = time, 
+            period = period )
+
+source( "time_series_analysis_plotting.R")
+patient_count_psy_period_psy_country_france <- patient_count_psy_period_psy_country %>%
+  filter( country == "France")
+ts_linear_percent_france <- lm(percentages ~ period * time, patient_count_psy_period_psy_country_france)
+
+time_series_analysis_plotting(ts_linear_percent_france,
+                              patient_count_psy_period_psy_country_france,
+                              "percentages",
+                              clearance_period = FALSE) +
+  labs(title = "Interrupted Time-Series Analysis: Percentage of encounters with psychiatric ICD codes (France)",
+       subtitle = "Without seasonality effect",
+       x = time_period,
+       y = "Percentage")
+
+patient_count_psy_period_psy_country_usa <- patient_count_psy_period_psy_country %>%
+  filter( country == "USA")
+ts_linear_percent_usa <- lm(percentages ~ period * time, patient_count_psy_period_psy_country_usa)
+
+time_series_analysis_plotting(ts_linear_percent_usa,
+                              patient_count_psy_period_psy_country_usa,
+                              "percentages",
+                              clearance_period = FALSE) +
+  labs(title = "Interrupted Time-Series Analysis: Percentage of encounters with psychiatric ICD codes (USA)",
+       subtitle = "Without seasonality effect",
+       x = time_period,
+       y = "Percentage")
+
+```
+```{r}
+bootstrap_size <- 1000
+
+binom_proportion <- function(n, prob, size) {
+  (rbinom(n, size, prob) / size) * 100
+}
+bootstrapped_timepoints_france <- mapply(binom_proportion,
+                                  n = bootstrap_size,
+                                  size = patient_count_psy_period_psy_country_france$total_cases,
+                                  prob = patient_count_psy_period_psy_country_france$percentages/100)
+bootstrapped_sample_list <- vector(mode = "list", length = bootstrap_size)
+bootstrapped_coefficients_list <- vector(mode = "list", length = bootstrap_size)
+for (iteration in 1:bootstrap_size) {
+  df_iteration <- data.frame(period = patient_count_psy_period_psy_country_france$period,
+                             time = patient_count_psy_period_psy_country_france$time,
+                             percentage = bootstrapped_timepoints_france[iteration, ])
+  ts_linear_percent <- lm(percentage ~ period * time, df_iteration)
+  linear_model <- broom::tidy(ts_linear_percent) %>%
+    filter(term %in% c("time", "periodduring_pandemic:time"))
+  coefficients <- linear_model$estimate
+  names(coefficients) <- linear_model$term
+  bootstrapped_coefficients_list[[iteration]] <- coefficients
+  df_iteration$predicted <- predict(ts_linear_percent)
+  df_iteration$time_p <- patient_count_psy_period_psy_country_france$time_p
+  bootstrapped_sample_list[[iteration]] <- df_iteration
+}
+bootstrapped_fitted_df_france <- bind_rows(bootstrapped_sample_list, .id = "iteration")
+bootstrapped_fitted_df_france$country <- "France"
+
+bootstrapped_timepoints_usa <- mapply(binom_proportion,
+                                  n = bootstrap_size,
+                                  size = patient_count_psy_period_psy_country_usa$total_cases,
+                                  prob = patient_count_psy_period_psy_country_usa$percentages/100)
+bootstrapped_sample_list_usa <- vector(mode = "list", length = bootstrap_size)
+bootstrapped_coefficients_list <- vector(mode = "list", length = bootstrap_size)
+for (iteration in 1:bootstrap_size) {
+  df_iteration <- data.frame(period = patient_count_psy_period_psy_country_usa$period,
+                             time = patient_count_psy_period_psy_country_usa$time,
+                             percentage = bootstrapped_timepoints_usa[iteration, ])
+  ts_linear_percent <- lm(percentage ~ period * time, df_iteration)
+  linear_model <- broom::tidy(ts_linear_percent) %>%
+    filter(term %in% c("time", "periodduring_pandemic:time"))
+  coefficients <- linear_model$estimate
+  names(coefficients) <- linear_model$term
+  bootstrapped_coefficients_list[[iteration]] <- coefficients
+  df_iteration$predicted <- predict(ts_linear_percent)
+  df_iteration$time_p <- patient_count_psy_period_psy_country_usa$time_p
+  bootstrapped_sample_list_usa[[iteration]] <- df_iteration
+}
+bootstrapped_fitted_df_usa <- bind_rows(bootstrapped_sample_list_usa, .id = "iteration")
+bootstrapped_fitted_df_usa$country <- "USA"
+
+bootstrapped_fitted_all_by_country <- rbind( bootstrapped_fitted_df_france, bootstrapped_fitted_df_usa)
+
+quantile_bootstrapped_fitted_country <- bootstrapped_fitted_all_by_country %>%
+  group_by(time_p, period, country) %>%
+  summarise("2.5%" = quantile(predicted, 0.025),
+            "50%" = median(predicted, 0.5),
+            "97.5%" = quantile(predicted, 0.975)) %>%
+  pivot_longer(cols = c("2.5%", "50%", "97.5%"),
+               names_to = "quantiles")
+
+annotation_function <- function(df, ...) {
+  z <- unique(df[["observed_diff"]]) / sd(df[["periodduring_pandemic:time"]])
+  pvalue_coeff_interaction = round(pnorm(abs(z), lower.tail = FALSE) * 2, 3)
+  paste0("Coeffs differences: ",
+         round(unique(df[["observed_diff"]]), 2),
+         " (pvalue: ",
+         pvalue_coeff_interaction,
+         ")")
+}
+
+coefficient_slope_calculation <- function(df, ...) {
+  ts_linear_percent <- lm(percentages ~ period * time, df)
+  linear_model <- broom::tidy(ts_linear_percent) %>%
+    filter(term %in% c("periodduring_pandemic:time"))
+}
+
+coefficients_df_by_country <- patient_count_psy_period_psy_country %>% 
+  group_by(country) %>% 
+  group_modify(coefficient_slope_calculation) %>% 
+  rename(observed_diff = estimate)
+
+df_coefficients_boot_by_country <- bootstrapped_fitted_all_by_country %>% 
+  left_join(coefficients_df_by_country, by = "country")
+
+names_countries <- unique(df_coefficients_boot_by_country$country)
+
+plot_annotations_country <- df_coefficients_boot_by_country %>% 
+  group_by(country) %>% 
+  group_map( annotation_function ) %>% 
+  set_names(names_countries) %>% 
+  unlist()
+plot_annotations_df_country <- data.frame(
+  country = names(plot_annotations_country), 
+  label_countr = plot_annotations_country
+)
+
+ggplot(bootstrapped_fitted_all_by_country, aes(x = time_p, y = predicted, group = interaction(iteration, period))) +
+  facet_wrap("country", scales = "free_y") +
+  geom_line(aes(color = period),
+            alpha = 0.03) +
+  geom_line(data = quantile_bootstrapped_fitted_country,
+            aes(y = value, group = interaction(quantiles, period)),
+            linetype = "dashed") +
+  geom_line(data = quantile_bootstrapped_fitted_country[quantile_bootstrapped_fitted_country$quantiles == "50%", ],
+            aes(y = value, group = period)) +
+  geom_point(data = patient_count_psy_period_psy_country,
+             aes(y = percentages, group=NULL),
+             alpha = 0.5) +
+  scale_colour_manual(values = cbPalette) +
+  guides(color = guide_legend(override.aes = list(alpha = 1))) +
+  geom_label(data = plot_annotations_df_country, 
+             aes(label=label_countr), 
+            x = -Inf, y = +Inf, hjust=0, vjust=1,
+            size = 4,
+            inherit.aes = FALSE) +
+  labs(title = "Interrupted Time Series Analysis: Percentage of patients with psychiatric conditions",
+       subtitle = "Parametric bootstrap estimation: (1000 iterations)",
+       x = "Calendar dates",
+       y = "Percentage",
+       color = "Period") +
+  theme(strip.text = element_text(size = 10),
+        axis.text.x = element_text(size = 8), 
+        axis.text.y = element_text(size = 8), 
+        axis.title.x = element_text(size = 8),
+        axis.title.y = element_text(size = 8),
+        title = element_text(size = 14),
+        plot.subtitle = element_text(size = 12, face = "italic")) 
+ggsave("boostrapped_country.png", height = 20, width = 32, units = "cm")
+
+```
+
 ## Meta-analysis of temporal trends

### 2022-03-04 metanalysis_visualization/time_series_analysis_plotting.R

@@ -1,0 +1,44 @@
+time_series_analysis_plotting <- function(model, df_count_psy, count_col, clearance_period = FALSE) {
+  output_model <- broom::tidy(model)
+  coeff_slope <- round(output_model[output_model$term == "periodduring_pandemic:time", "estimate"][[1]], 3)
+  pvalue_slope <- round(output_model[output_model$term == "periodduring_pandemic:time", "p.value"][[1]], 3)
+  annotation <- paste0("Lockdown slope: ", coeff_slope, " (pvalue: ", pvalue_slope, ")")
+  pred_model <- predict(model, df_count_psy, type = "response", interval = "confidence") %>% as.data.frame()
+  pred_model_b <- predict(model, mutate(df_count_psy, period = "before_pandemic"), type = "response", interval = "confidence") %>%
+    as.data.frame()
+  names(pred_model_b) <- paste0(names(pred_model_b), "_wo_intervention")
+  pred_model <- bind_cols(df_count_psy, pred_model, pred_model_b)
+  if (clearance_period) {
+    pred_model$clearance_period <- ifelse(pred_model$time_p < as.Date(cut(start_clear_period, breaks = time_period)) ,
+                                          "before_pandemic",
+                                          ifelse(pred_model$time_p > as.Date(cut(end_clear_period, breaks = time_period)),
+                                                 "during_pandemic",
+                                                 "during_clearance")
+    )
+    pred_model$lwr <- ifelse(pred_model$clearance_period == "during_clearance", NA, pred_model$lwr)
+    pred_model$upr <- ifelse(pred_model$clearance_period == "during_clearance", NA, pred_model$upr)
+    pred_model$fit_wo_intervention <- ifelse(pred_model$time_p >= max(pred_model$time_p[pred_model$clearance_period == "before pandemic"]),
+                                             pred_model$fit_wo_intervention,
+                                             NA)
+  } else {
+    pred_model$clearance_period <- pred_model$period
+  }
+  plot <- ggplot(pred_model, aes_string(x = "time_p", y = count_col)) +
+    geom_point(aes(colour = clearance_period)) +
+    geom_line(aes(x = time_p, y = fit_wo_intervention, colour = "Expected mean w/o pandemic"), linetype = "dashed") +
+    geom_line(aes(x = time_p, y = fit, colour = clearance_period)) +
+    geom_ribbon(aes(ymin = lwr, ymax = upr, colour = clearance_period), alpha = 0.3, colour = NA) +
[…]

### 2022-03-09 metanalysis_visualization/manuscript_Figures.Rmd
code to generate a pannel with the ITS with seasonality by site

@@
 
+## ITS with seasonality
+
+### By site
+```{r}
+head(patient_count_psy_period_psy_all)
+sites <- unique( patient_count_psy_period_psy_all$siteid )
+
+p <- list()
+
+for( i in 1:length(sites)){
+  print(i)
+  subset_site <- patient_count_psy_period_psy_all %>%
+    filter( siteid == sites[i])
+  ts_seasonal_percent_per_site <- lm(percentage ~ period*time + harmonic(month_year, 1, 12), subset_site)
+  p[[i]] <- time_series_analysis_plotting(ts_seasonal_percent_per_site,
+                              subset_site,
+                              "percentage",
+                              clearance_period = FALSE) +
+  labs(title = sites[i],
+       subtitle = "ITS with seasonality effect",
+       x = "Calendar dates",
+       y = "Percentage") +
+     theme(strip.text = element_text(size = 10),
+        axis.text.x = element_text(size = 8), 
+        axis.text.y = element_text(size = 8), 
+        axis.title.x = element_text(size = 8),
+        axis.title.y = element_text(size = 8),
+        title = element_text(size = 14),
+        plot.subtitle = element_text(size = 12, face = "italic"))
+
+}
+
+g <- do.call(grid.arrange,p)
+ggsave("its_seasonality_site.png",g, height = 20, width = 32, units = "cm")
+
+```
+
+### By country
+```{r}
+
+```
+
 ### R session information

### 2022-03-12 metanalysis_visualization/manuscript_Figures.Rmd
figures updated for the manuscript

--- 2022-02-27 metanalysis_visualization/manuscript_Figures.Rmd
@@
 pandemic_start_date <- as.Date("2020-03-15")
-start_date_plots <- as.Date("2019-01-15")
-end_date_plots <- as.Date("2021-05-01")
+start_date_plots <- as.Date("2019-01-01")
+end_date_plots <- as.Date("2021-05-31")
 ```
@@
 
-patients_hospitalized_agg_sex_country <- patients_hospitalized_agg_sex_perc_all %>%
-  select( - percentage ) %>%
-  inner_join( country_map ) %>%
-  group_by( time_p, sex, country ) %>%
-  mutate( total_cases = sum( count ), 
-          sex_cases = sum( count_sex ), 
-          percentage = 100* sex_cases / total_cases) %>%
-  rbind( patients_hospitalized_agg_sex_total)
+#patients_hospitalized_agg_sex_country <- patients_hospitalized_agg_sex_perc_all %>%
+#  select( - percentage ) %>%
+#  inner_join( country_map ) %>%
+#  group_by( time_p, sex, country ) %>%
+#  mutate( total_cases = sum( count ), 
+#          sex_cases = sum( count_sex ), 
+#          percentage = 100* sex_cases / total_cases) %>%
+#  rbind( patients_hospitalized_agg_sex_total)
 
-patients_hospitalized_agg_sex_country %>%
-  ggplot( aes( x= as.Date( time_p ), y = percentage, group = sex )) +
-  geom_line( aes(linetype = sex, color = sex )) +
-  geom_point( aes( shape= sex, color = sex )) +
-  scale_colour_manual(values=c("#FF6600","#339999")) +
-  scale_fill_manual(values=c("#FF6600","#339999"))+
-  #scale_y_continuous(breaks = seq(0, 30, by = 1)) +
-  facet_wrap2("country", axes = "all") + 
-  geom_vline(xintercept = as.Date(pandemic_start_date), linetype = "dashed") +
-  labs(x ="", y = "percentage of patients (%)")+
-  scale_x_date(breaks = "2 month", labels=date_format("%b-%y"), expand=c(0.03,0)) +
-  theme(axis.text.x=element_text(angle = 45, hjust=1, size = 6), 
-        strip.text = element_text(size = 10, face = "bold"),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), strip.background = element_blank(),
-        panel.border = element_rect(colour = "black", fill = NA), 
-        legend.position="bottom")
+# patients_hospitalized_agg_sex_country %>%
+#   ggplot( aes( x= as.Date( time_p ), y = percentage, group = sex )) +
+#   geom_line( aes(linetype = sex, color = sex )) +
+#   geom_point( aes( shape= sex, color = sex )) +
+#   scale_colour_manual(values=c("#FF6600","#339999")) +
+#   scale_fill_manual(values=c("#FF6600","#339999"))+
+#   #scale_y_continuous(breaks = seq(0, 30, by = 1)) +
+#   facet_wrap2("country", axes = "all") + 
+#   geom_vline(xintercept = as.Date(pandemic_start_date), linetype = "dashed") +
+#   labs(x ="", y = "percentage of patients (%)")+
+#   scale_x_date(breaks = "2 month", labels=date_format("%b-%y"), expand=c(0.03,0)) +
+#   theme(axis.text.x=element_text(angle = 45, hjust=1, size = 6), 
+#         strip.text = element_text(size = 10, face = "bold"),
+#         panel.grid.major = element_blank(), panel.grid.minor = element_blank(), strip.background = element_blank(),
+#         panel.border = element_rect(colour = "black", fill = NA), 
+#         legend.position="bottom")
 ```
@@
         panel.border = element_rect(colour = "black", fill = NA))
+#   
+# perc_disorder_group_country <- perc_disorder_group_all %>%
+#   select( - percentage_dg ) %>%
+#   inner_join( country_map ) %>%
+#   group_by( time_p, disorder_group, country ) %>%
+#   mutate( total_cases = sum( count_icd ), 
+#           dg_cases = sum( count_dg ), 
+#           percentage = 100* dg_cases / total_cases ) %>%
+#   rbind( perc_disorder_group_total)
 
-perc_disorder_group_country <- perc_disorder_group_all %>%
-  select( - percentage_dg ) %>%
-  inner_join( country_map ) %>%
-  group_by( time_p, disorder_group, country ) %>%
-  mutate( total_cases = sum( count_icd ), 
-          dg_cases = sum( count_dg ), 
-          percentage = 100* dg_cases / total_cases ) %>%
-  rbind( perc_disorder_group_total)
-
-perc_disorder_group_country %>%
+perc_disorder_group_total %>%
   filter( disorder_group %in% c("Anxiety Disorders", "Depressive Disorders", "Suicide or Self-Injury", 
-  "Mental Health Symptoms", "Obsessive-Compulsive and Related Disorders" )) %>%
-  ggplot( aes( x= as.Date( time_p ), y = percentage, group = disorder_group )) +
-  geom_line( aes(linetype = disorder_group, color = disorder_group )) +
-  geom_point( aes( shape= disorder_group, color = disorder_group )) +
-  scale_y_continuous(breaks = seq(0, 30, by = 1)) +
-  facet_wrap2("country", axes = "all") + 
-  geom_vline(xintercept = as.Date(pandemic_start_date), linetype = "dashed") +
-  labs(x ="", y = "percentage of patients (%)")+
-  scale_x_date(breaks = "2 month", labels=date_format("%b-%y"), expand=c(0.03,0)) +
-  theme(axis.text.x=element_text(angle = 45, hjust=1, size = 6), 
-        strip.text = element_text(size = 10, face = "bold"),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), strip.background = element_blank(),
-        panel.border = element_rect(colour = "black", fill = NA), 
-        legend.position="bottom")
-
-perc_disorder_group_country %>%
-  filter( disorder_group %in% c("Anxiety Disorders", "Depressive Disorders", "Suicide or Self-Injury", 
-  "Mental Health Symptoms", "Obsessive-Compulsive and Related Disorders" )) %>%
+  "Mental Health Symptoms", "Feeding and Eating Disorders" )) %>%
   ggplot( aes( x= as.Date( time_p ), y = percentage, group = country )) +
@@
   geom_point( aes( shape= country, color = country )) +
-  scale_y_continuous(breaks = seq(0, 30, by = 5)) +
-  facet_wrap2("disorder_group", axes = "all") + 
+  #scale_y_continuous(breaks = seq(0, 30, by = 5)) +
+  facet_wrap2("disorder_group", axes = "all", scales ="free") + 
   geom_vline(xintercept = as.Date(pandemic_start_date), linetype = "dashed") +
   labs(x ="", y = "percentage of patients (%)")+
-  scale_x_date(breaks = "2 month", labels=date_format("%b-%y"), expand=c(0.03,0)) +
+  scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0)) +
   scale_color_manual( values = cbPalette ) + 
@@
 
+## ITS with seasonality
+
+### By site
+```{r}
+head(patient_count_psy_period_psy_all)
+sites <- unique( patient_count_psy_period_psy_all$siteid )
+source("../metanalysis_visualization/time_series_analysis_plotting.R")
+p <- list()
+
+for( i in 1:length(sites)){
+  print(i)
+  subset_site <- patient_count_psy_period_psy_all %>%
+    filter( siteid == sites[i])
+  ts_seasonal_percent_per_site <- lm(percentage ~ period*time + harmonic(month_year, 1, 12), subset_site)
+  p[[i]] <- time_series_analysis_plotting(ts_seasonal_percent_per_site,
+                              subset_site,
+                              "percentage",
+                              clearance_period = FALSE) +
+  labs(title = sites[i],
+       x = "Calendar dates",
+       y = "Percentage") +
+     theme(strip.text = element_text(size = 10),
+        axis.text.x = element_text(size = 8), 
+        axis.text.y = element_text(size = 8), 
+        axis.title.x = element_text(size = 8),
+        axis.title.y = element_text(size = 8),
+        title = element_text(size = 14),
+        plot.subtitle = element_text(size = 12, face = "italic"))
+
+}
+
+g <- do.call(grid.arrange,p)
+ggsave("its_seasonality_site.png",g, height = 20, width = 32, units = "cm")
+
+```
+
 ### R session information

### 2022-03-14 metanalysis_visualization/manuscript_Figures.Rmd
Minor changes

--- 2022-03-09 metanalysis_visualization/manuscript_Figures.Rmd
@@
 
-install_github("jokergoo/ComplexHeatmap")
-install_github("teunbrand/ggh4x")
-
+if (! "ComplexHeatmap" %in% rownames(installed.packages())) install_github("jokergoo/ComplexHeatmap")
+if (! "ggh4x" %in% rownames(installed.packages())) install_github("teunbrand/ggh4x")
 library( ComplexHeatmap)
@@
   } else {
+    if (outputFiles[i] == "peds_psy_aggResults_CHOP.RData") {
+      count_icd_all <- tryCatch(
+        rbind(count_icd_all, count_icd),
+        error = function(e) count_icd_all
+      )
+      patients_hospitalized_agg_sex_perc_all <- tryCatch( 
+        rbind( patients_hospitalized_agg_sex_perc_all, patients_hospitalized_agg_sex_perc),
+        error = function(e) patients_hospitalized_agg_sex_perc_all
+      )
+      perc_disorder_group_all <- tryCatch(
+        rbind(perc_disorder_group_all, perc_disorder_group),
+        error = function(e) perc_disorder_group_all
+      )
+      mental_codes_qc_output_all <- tryCatch(
+        rbind( mental_codes_qc_output_all,
+               mental_codes_qc_output %>%
+                 select( disorder_group, description, concept_code, distinct_patients, distinct_observations, siteid )
+        ),
+        error = function(e) mental_codes_qc_output_all
+      )
+      table1_all <- table1_all
+    } else {
       count_icd_all <- rbind(count_icd_all, count_icd)
+      perc_disorder_group_all <- rbind(perc_disorder_group_all, perc_disorder_group)
       patients_hospitalized_agg_sex_perc_all <- rbind( patients_hospitalized_agg_sex_perc_all, patients_hospitalized_agg_sex_perc)
+      mental_codes_qc_output_all <- rbind( mental_codes_qc_output_all,
+                                           mental_codes_qc_output %>%
+                                             select( disorder_group, description, concept_code, distinct_patients, distinct_observations, siteid )
+      )
+      table1_all <- rbind( table1_all, table1)
+    }
     patient_count_psy_period_psy_all <- rbind( patient_count_psy_period_psy_all,patient_count_psy_period_psy)
-    perc_disorder_group_all <- rbind(perc_disorder_group_all, perc_disorder_group)
     ratios_patients_with_without_psy_all <- rbind( ratios_patients_with_without_psy_all, ratios_patients_with_without_psy)
@@
     bootstrapped_coefficients_all <- rbind(bootstrapped_coefficients_all, bootstrapped_coefficients_df)
-   mental_codes_qc_output <- mental_codes_qc_output %>%
-      select( disorder_group, description, concept_code, distinct_patients, distinct_observations, siteid )
-    mental_codes_qc_output_all <- rbind( mental_codes_qc_output_all, mental_codes_qc_output)
-    table1_all <- rbind( table1_all, table1)
-    }
     rm(count_icd,patients_hospitalized_agg_sex_perc,patient_count_psy_period_psy,
@@
   }
+}
+
 ```
@@
 sessionInfo()
-```

### 2022-03-14 metanalysis_visualization/manuscript_Figures.Rmd
reverting chunk aggregating tables

--- 2022-03-12 metanalysis_visualization/manuscript_Figures.Rmd
@@
     count_icd_all <- rbind(count_icd_all, count_icd)
+    perc_disorder_group_all <- rbind(perc_disorder_group_all, perc_disorder_group)
     patients_hospitalized_agg_sex_perc_all <- rbind( patients_hospitalized_agg_sex_perc_all, patients_hospitalized_agg_sex_perc)
+    mental_codes_qc_output_all <- rbind( mental_codes_qc_output_all,
+                                         mental_codes_qc_output %>%
+                                           select( disorder_group, description, concept_code, distinct_patients, distinct_observations, siteid )
+    )
+    table1_all <- rbind( table1_all, table1)
     patient_count_psy_period_psy_all <- rbind( patient_count_psy_period_psy_all,patient_count_psy_period_psy)
-    perc_disorder_group_all <- rbind(perc_disorder_group_all, perc_disorder_group)
     ratios_patients_with_without_psy_all <- rbind( ratios_patients_with_without_psy_all, ratios_patients_with_without_psy)

### 2022-03-15 metanalysis_visualization/manuscript_Figures.Rmd
Adding mean age calculation

@@
 
+```{r Extracting age information}
+
+age_table1_all <- table1_all %>% 
+  filter(name == "Age" | (name == "Total" & Category == "Sex")) %>% 
+  select(name, siteid, before_pandemic, during_pandemic) %>% 
+  pivot_longer(cols = c(before_pandemic, during_pandemic), 
+               names_to = "period",
+               values_to = "values") %>% 
+  pivot_wider(names_from = name, 
+              values_from = values, 
+              id_cols = c(siteid, period)) %>% 
+  mutate(mean_age = as.numeric(sub("\\s\\(.*", "", x = Age, perl = TRUE)),
+         sd_age = as.numeric(str_extract(pattern = "(?<=\\(sd:\\s)[0-9]{1,2}\\.[0-9]{1,2}(?=\\))", 
+                                         string = Age)),
+         Total = as.numeric(sub("\\s\\(.*", "", x = Total, perl = TRUE))
+  ) %>% 
+  group_by(period) %>% 
+  summarise(mean_age = round(sum(Total * mean_age) / sum(Total), 2), 
+            sd_age = round(sum(Total * sd_age) / sum(Total), 2)) %>% 
+  mutate(Age = paste0(as.character(mean_age), " (sd: ", as.character(sd_age), ")")) %>% 
+  select(period, Age) %>% 
+  pivot_wider(names_from = period, values_from = Age) %>% 
+  mutate(Category = "Age", 
+         name = "mean (sd)") %>% 
+  select(Category, name, before_pandemic, during_pandemic) %>% 
+  rename(before_pandemic_count = before_pandemic, 
+         during_pandemic_count = during_pandemic)
+```
+
 ```{r}
@@
 
-table1 <- rbind( sex_table1_all, dg_table1_all )
+table1 <- rbind( sex_table1_all, age_table1_all, dg_table1_all )
 
@@
 
-## Boostrap vizualisation
 ```{r}
@@
 ```
+
+### Primary analysis
 
@@
              aes(label=label), 
-            x = -Inf, y = +Inf, hjust=0, vjust=1,
+            x = -Inf,
+            y = +Inf,
+            hjust = 0,
+            vjust = 1,
             size = 4,

### 2022-03-15 metanalysis_visualization/manuscript_Figures.Rmd
count plots added

@@
 
+### Total counts with psychiatric conditions and total counts per month
+
+#### At site level
+```{r}
+library(reshape2)
+
+counts <- patient_count_psy_period_psy_all %>%
+  mutate( total_cases = count_no_psy_patients + count_psy_patients) %>%
+  select( siteid, time_p, time, month_year, count_psy_patients, total_cases)
+
+counts2 <- reshape2::melt(counts ,  id.vars = c('siteid', 'time_p', 'time','month_year'), variable.name = 'group')
+counts2$group <- gsub("count_psy_patients", "patients with psychiatric conditions", counts2$group )
+counts2$group <- gsub("total_cases", "total number of patients", counts2$group )
+
+counts2 %>% ggplot( aes( x= as.Date( time_p ), y = value, group= group )) +
+  geom_point( aes( shape= group, color = group )) +
+  facet_wrap2("siteid", axes = "all", scales ="free") + 
+  geom_vline(xintercept = as.Date(pandemic_start_date), linetype = "dashed") +
+  labs(x ="", y = "number of patients")+
+  scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0)) +
+  scale_color_manual( values = cbPalette ) + 
+  theme(axis.text.x=element_text(angle = 45, hjust=1, size = 6), 
+        strip.text = element_text(size = 10, face = "bold"),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), strip.background = element_blank(),
+        panel.border = element_rect(colour = "black", fill = NA), 
+        legend.position="bottom")
+```
+
+#### All together
+```{r}
+counts_agg <- patient_count_psy_period_psy_all %>%
+  group_by( time_p ) %>%
+  mutate( psy_cases = sum( count_psy_patients), 
+    total_cases = sum(count_no_psy_patients) + sum(count_psy_patients)) %>%
+  select( time_p, time, month_year, psy_cases, total_cases)
+
+counts_agg2 <- reshape2::melt(counts_agg ,  id.vars = c('time_p', 'time','month_year'), variable.name = 'group')
+counts_agg2$group <- gsub("count_psy_patients", "patients with psychiatric conditions", counts_agg2$group )
+counts_agg2$group <- gsub("total_cases", "total number of patients", counts_agg2$group )
+
+counts_agg2 %>% ggplot( aes( x= as.Date( time_p ), y = value, group= group )) +
+  geom_point( aes( shape= group, color = group )) +
+  geom_vline(xintercept = as.Date(pandemic_start_date), linetype = "dashed") +
+  labs(x ="", y = "number of patients")+
+  scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0)) +
+  scale_color_manual( values = cbPalette ) + 
+  theme(axis.text.x=element_text(angle = 45, hjust=1, size = 6), 
+        strip.text = element_text(size = 10, face = "bold"),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), strip.background = element_blank(),
+        panel.border = element_rect(colour = "black", fill = NA), 
+        legend.position="bottom")
+```
+
 ### R session information

### 2022-03-15 metanalysis_visualization/manuscript_Figures.Rmd
updating length of stay and bootstrap

--- 2022-03-15 metanalysis_visualization/manuscript_Figures.Rmd
@@
     bootstrapped_coefficients_all <- bootstrapped_coefficients_df
+    length_hospitalisation_values_all <- length_hospitalisation_values
     mental_codes_qc_output_all <- mental_codes_qc_output %>%
@@
     bootstrapped_coefficients_all <- rbind(bootstrapped_coefficients_all, bootstrapped_coefficients_df)
+    length_hospitalisation_values_all <- tryCatch(
+      rbind(length_hospitalisation_values_all, length_hospitalisation_values),
+      error = function(e) length_hospitalisation_values_all
+    )
     rm(count_icd,patients_hospitalized_agg_sex_perc,patient_count_psy_period_psy,
@@
        patient_count_psy_period_psy_clear, bootstrapped_coefficients_df,
-       bootstrapped_fitted_df, mental_codes_qc_output, table1)
+       bootstrapped_fitted_df, mental_codes_qc_output, table1, length_hospitalisation_values)
   }
@@
 ```{r}
+median_table1_all <- length_hospitalisation_values_all %>% 
+  filter(siteid != "CCHMC") %>% # temporary
+  group_by(period) %>% 
+  summarise(median_hospitalisation = median(len_hospitalisation)) %>% 
+  pivot_wider(names_from = period, values_from = median_hospitalisation) %>% 
+  mutate(Category = "Length of stays", 
+         name = "median") %>% 
+  select(Category, name, before_pandemic, during_pandemic) %>% 
+  rename(before_pandemic_count = before_pandemic, 
+         during_pandemic_count = during_pandemic)
+```
+
+```{r}
 sex_table1_all <- table1_all %>%
@@
 
-table1 <- rbind( sex_table1_all, age_table1_all, dg_table1_all )
+table1 <- rbind( sex_table1_all, age_table1_all, median_table1_all, dg_table1_all )
 
@@
 
+## Recompute the bootstrap
+
+```{r}
+# Call function from another file : bootstrap.R
+bootstrap_size <- 1000
+binom_proportion <- function(n, prob, size) {
+  (rbinom(n, size, prob) / size) * 100
+}
+
+bootstrapping_slopes <- function(patient_count_psy_period_psy, 
+                                    bootstrap_size) {
+  bootstrapped_timepoints <- mapply(binom_proportion,
+                                    n = bootstrap_size,
+                                    size = patient_count_psy_period_psy$count_no_psy_patients +
+                                      patient_count_psy_period_psy$count_psy_patients,
+                                    prob = patient_count_psy_period_psy$percentage/100)
+  bootstrapped_sample_list <- vector(mode = "list", length = bootstrap_size)
+  bootstrapped_coefficients_list <- vector(mode = "list", length = bootstrap_size)
+  for (iteration in 1:bootstrap_size) {
+    df_iteration <- data.frame(period = patient_count_psy_period_psy$period,
+                               time = patient_count_psy_period_psy$time,
+                               percentage = bootstrapped_timepoints[iteration, ])
+    ts_linear_percent <- lm(percentage ~ period * time, df_iteration)
+    linear_model <- broom::tidy(ts_linear_percent) %>%
+      filter(term %in% c("time", "periodduring_pandemic:time"))
+    coefficients <- linear_model$estimate
+    names(coefficients) <- linear_model$term
+    bootstrapped_coefficients_list[[iteration]] <- coefficients
+    df_iteration$predicted <- predict(ts_linear_percent)
+    df_iteration$time_p <- patient_count_psy_period_psy$time_p
+    bootstrapped_sample_list[[iteration]] <- df_iteration
+  }
+  bootstrapped_fitted_df <- bind_rows(bootstrapped_sample_list, .id = "iteration")
+  bootstrapped_coefficients_df <- bind_rows(bootstrapped_coefficients_list,
+                                          .id = "iteration")
+  return(list(bootstrapped_coefficients_df, 
+              bootstrapped_fitted_df))
+}
+
+# Get count_icd_all in the same format that patient_count_psy_period_psy_all, because first n last months missing ... 
+count_icd_all_boot <- count_icd_all %>% 
+  group_by(siteid) %>% # necessary to generate time correctly 
+  rename(count_psy_patients = count_psy, 
+         percentage = percentage_psy) %>% 
+  mutate(count_no_psy_patients = count_icd - count_psy_patients, 
+         month_year = format(time_p, "%m"),
+         time = seq_along(time_p))
+
+bootstrapped <- count_icd_all %>% 
+  group_by(siteid) %>% 
+  group_map( ~bootstrapping_slopes(.x, bootstrap_size = bootstrap_size))
+names(test) <- unique(count_icd_all$siteid)
+
+bootstrapped_coefficients_all <- lapply(bootstrapped, `[[`, 1) %>% bind_rows(.id = "siteid")
+bootstrapped_fitted_all <- lapply(bootstrapped, `[[`, 2) %>% bind_rows(.id = "siteid")
+```
+
 ### Primary analysis
@@
 ```{r}
-
 quantile_bootstrapped_fitted_all <- bootstrapped_fitted_all %>%
@@
 )
+```
 
+```{r}
+### January 19-May 21
+ggplot(bootstrapped_fitted_all, aes(x = time_p, y = predicted, group = interaction(iteration, period))) +
+  facet_wrap("siteid", scales = "free_y") +
+  geom_line(aes(color = period),
+            alpha = 0.03) +
+  geom_line(data = quantile_bootstrapped_fitted_all,
+            aes(y = value, group = interaction(quantiles, period)),
+            linetype = "dashed") +
+  geom_line(data = quantile_bootstrapped_fitted_all[quantile_bootstrapped_fitted_all$quantiles == "50%", ],
+            aes(y = value, group = period)) +
+  geom_point(data = count_icd_all_boot,
+             aes(y = percentage, group=NULL),
+             alpha = 0.5) +
+  scale_colour_manual(values = cbPalette) +
+  guides(color = guide_legend(override.aes = list(alpha = 1))) +
+  geom_label(data = plot_annotations_df, 
+             aes(label=label), 
+            x = -Inf,
+            y = +Inf,
+            hjust = 0,
+            vjust = 1,
+            size = 4,
+            inherit.aes = FALSE) +
+  labs(title = "Interrupted Time Series Analysis: Percentage of patients with psychiatric conditions",
+       subtitle = "Parametric bootstrap estimation: (1000 iterations)",
+       x = "Calendar dates",
+       y = "Percentage",
+       color = "Period") +
+  theme(strip.text = element_text(size = 10),
+        axis.text.x = element_text(size = 8), 
+        axis.text.y = element_text(size = 8), 
+        axis.title.x = element_text(size = 8),
+        axis.title.y = element_text(size = 8),
+        title = element_text(size = 14),
+        plot.subtitle = element_text(size = 12, face = "italic")) 
+ggsave("boostrapped_all_extended.png", height = 20, width = 32, units = "cm")
+```
+
+```{r}
+### Feb 19-April 21
 ggplot(bootstrapped_fitted_all, aes(x = time_p, y = predicted, group = interaction(iteration, period))) +

### 2022-03-16 metanalysis_visualization/manuscript_Figures.Rmd
fixing bootstrap

@@
 
-bootstrapped <- count_icd_all %>% 
+bootstrapped <- count_icd_all_boot %>% 
   group_by(siteid) %>% 
   group_map( ~bootstrapping_slopes(.x, bootstrap_size = bootstrap_size))
-names(test) <- unique(count_icd_all$siteid)
+names(bootstrapped) <- unique(count_icd_all$siteid)
 
@@
 ```
-
-### R session information
-
-```{r, results='hide'}
-sessionInfo()

### 2022-03-22 metanalysis_visualization/manuscript_Figures.Rmd
update supplementary figure 2 as a heatmap

@@
 }
-listOfPackages <- c("tidyverse", "tidyr","devtools","RColorBrewer", "knitr", "kableExtra", "tsModel", "gridExtra", "dplyr", "metafor", "meta", "viridis", "UpSetR", "scales")
+listOfPackages <- c("tidyverse", "tidyr","devtools","RColorBrewer", "knitr", "kableExtra", "tsModel", "gridExtra", "dplyr", "metafor", "meta", "viridis", "UpSetR", "scales", "EHRtemporalVariability")
 paket(listOfPackages)
@@
   ggplot( aes( x= as.Date( time_p ), y = percentage, group = disorder_group )) +
+  facet_wrap( ~ disorder_group, scales = "free" ) +
   geom_line( aes(color = disorder_group )) +
@@
         panel.border = element_rect(colour = "black", fill = NA))
-#   
-# perc_disorder_group_country <- perc_disorder_group_all %>%
-#   select( - percentage_dg ) %>%
-#   inner_join( country_map ) %>%
-#   group_by( time_p, disorder_group, country ) %>%
-#   mutate( total_cases = sum( count_icd ), 
-#           dg_cases = sum( count_dg ), 
-#           percentage = 100* dg_cases / total_cases ) %>%
-#   rbind( perc_disorder_group_total)
+
+percentages_htm <- perc_disorder_group_total %>%
+  select( time_p, disorder_group, percentage ) %>%
+  unique() %>%
+  pivot_wider( names_from =  disorder_group, 
+               values_from = percentage )
+
+dates <- percentages_htm$time_p
+
+percentages_htm <- percentages_htm[  , -1]
+
+counts_htm <- perc_disorder_group_total %>%
+  select( time_p, disorder_group, dg_cases ) %>%
+  unique() %>%
+  pivot_wider( names_from =  disorder_group, 
+               values_from = dg_cases )
+counts_htm <- counts_htm[  , -1]
+
+tempVariabilityInput <- new('DataTemporalMap', probabilityMap = as.matrix(percentages_htm), 
+         countsMap = as.matrix(counts_htm), dates = dates, support = as.data.frame(names(percentages_htm)), 
+         variableName = "example", variableType = "character", period = "month")
+plotDataTemporalMap(tempVariabilityInput, absolute = TRUE)
+
+#### facet wrap in 3 groups based on %
+### most prevalent
+perc_disorder_group_total %>%
+  filter(disorder_group %in% c("Anxiety Disorders", "Depressive Disorders", "Suicide or Self-Injury" ))  %>%
+  ggplot( aes( x= as.Date( time_p ), y = percentage, group = country )) +
+  facet_wrap( ~ disorder_group ) +
+  geom_line( aes(color = country )) +
+  geom_point( aes( color = country )) +
+  geom_vline(xintercept = as.Date(pandemic_start_date), linetype = "dashed")+
+  labs(x ="", y = "percentage of patients (%)")+
+  scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0))+
+  theme(axis.text.x=element_text(angle = 45, hjust=1, size = 6),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.border = element_rect(colour = "black", fill = NA)) +
+  scale_color_manual( values = cbPalette )
+
+#medium prevalent
+perc_disorder_group_total %>%
+  filter(disorder_group %in% c("Trauma and Stressor-Related Disorders", "Mental Health Symptom", "Disruptive, Impulse Control and Conduct Disorders", 
+                               "Feeding and Eating Disorders"))  %>%
+  ggplot( aes( x= as.Date( time_p ), y = percentage, group = country )) +
+  facet_wrap( ~ disorder_group ) +
+  geom_line( aes(color = country )) +
+  geom_point( aes( color = country )) +
+  geom_vline(xintercept = as.Date(pandemic_start_date), linetype = "dashed")+
+  labs(x ="", y = "percentage of patients (%)")+
+  scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0))+
+  theme(axis.text.x=element_text(angle = 45, hjust=1, size = 6),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.border = element_rect(colour = "black", fill = NA)) +
+  scale_color_manual( values = cbPalette )
+
+#low prevalent
+perc_disorder_group_total %>%
+  filter(disorder_group %in% c("Miscellaneous", "Substance-Related and Addictive Disorders", 
+                               "Personality Disorders", "Bipolar and Related Disorders", "Sleep-Wake Disorders", 
+                               "Schizophrenia Spectrum and Other Psychotic Disorders", "Somatic Symptom and Related Disorders", 
+                               "Obsessive-Compulsive and Related Disorders", "Dissociative Disorders" ))  %>%
+  ggplot( aes( x= as.Date( time_p ), y = percentage, group = country )) +
+  facet_wrap( ~ disorder_group ) +
+  geom_line( aes(color = country )) +
+  geom_point( aes( color = country )) +
+  geom_vline(xintercept = as.Date(pandemic_start_date), linetype = "dashed")+
+  labs(x ="", y = "percentage of patients (%)")+
+  scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0))+
+  theme(axis.text.x=element_text(angle = 45, hjust=1, size = 6),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.border = element_rect(colour = "black", fill = NA)) +
+  scale_color_manual( values = cbPalette )
+
+perc_disorder_group_total %>%
+  filter(disorder_group %in% c("Anxiety Disorders", "Depressive Disorders", "Suicide or Self-Injury" ))  %>%
+  ggplot( aes( x= as.Date( time_p ), y = percentage, group = country )) +
+  facet_wrap( ~ disorder_group ) +
+  geom_line( aes(color = country )) +
+  geom_point( aes( color = country )) +
+  geom_vline(xintercept = as.Date(pandemic_start_date), linetype = "dashed") +
+  labs(x ="", y = "percentage of patients (%)")+
+  scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0)) +
+  scale_color_manual( values = cbPalette ) + 
+  theme(axis.text.x=element_text(angle = 45, hjust=1, size = 6), 
+        strip.text = element_text(size = 10, face = "bold"),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), strip.background = element_blank(),
+        panel.border = element_rect(colour = "black", fill = NA), 
+        legend.position="bottom")
 

### 2022-03-23 metanalysis_visualization/manuscript_Figures.Rmd
fixing meta analysis by country

--- 2022-03-16 metanalysis_visualization/manuscript_Figures.Rmd
@@
 ```{r}
-## Estimate sampling variances
-```
-
-```{r}
 trends_siteid <- df_coefficients_boot %>% 
@@
                                   data = trends_siteid)
+
 metafor::forest(model_meta_siteid,
@@
                 header="Sites",
-                mlab="Overall temporal trend")
+                mlab="Overall periods' slope differences")
 ```
@@
 ```{r}
-trends_country <- df_coefficients_boot %>% 
+trends_country <- trends_siteid %>% 
   left_join(country_map, by = "siteid") %>% 
   group_by(country) %>% 
-  summarise(mean_boot = mean(`periodduring_pandemic:time`),
-         var_boot = var(`periodduring_pandemic:time`), 
-         z = mean_boot / sqrt(var_boot), 
-         pvalue = pnorm(z, lower.tail = FALSE)*2 )
-
-model_meta_country <- metafor::rma(yi = mean_boot, 
+  group_modify(.f = function(x, ...) {
+    as.data.frame(
+      predict(rma(yi = observed, 
                 vi = var_boot,
                 measure = "MN",
-             method = "FE",
-  data = trends_country) 
-metafor::forest(model_meta_country,
+                method = "DL",
+                data = x))
+    )
+  },
+  .keep = TRUE)
+
+model_meta_country <- rma(data = trends_country, 
+                          yi = pred, 
+                          vi = se**2, 
+                          measure = "MN", 
+                          method = "DL")
+
+forest(model_meta_country, 
        slab = trends_country$country,
-                header="Sites",
-                mlab="Overall temporal trend")
+       header = "Countries",
+       mlab = "Overall periods' slope differences")
 ```

### 2022-03-23 metanalysis_visualization/manuscript_Figures.Rmd
forest plot: reordering and changing square sizes proportions

@@
                 slab = trends_siteid$siteid,
+                plim = c(1, NA),
+                order = "obs",
+                showweights = TRUE,
                 header="Sites",
@@
        slab = trends_country$country,
+       order = "obs", 
+       showweights = TRUE,
        header = "Countries",

### 2022-03-23 metanalysis_visualization/manuscript_Figures.Rmd
forest plot: reordering and changing square sizes proportions

@@
        slab = trends_country$country,
+       plim = c(1, NA),
        order = "obs", 

### 2022-03-23 psy_peds_3.Rmd
ICD french codes V3

@@
 ```{r}
-icdCodes <- read.csv("public-data/pediatric_psychiatric_ICD10_codesV2.csv", header = TRUE, colClasses = "character") %>%
+icdCodes <- read.csv("public-data/pediatric_psychiatric_ICD10_codesV3.csv", header = TRUE, colClasses = "character") %>%
    filter( flag %in% c("originalCodes", "FrenchSpecificCodes")) %>%

### 2022-03-28 metanalysis_visualization/manuscript_Figures.Rmd
counts by site updated by using count_icd_all as input that contains all months for all sites

@@
 
-counts <- patient_count_psy_period_psy_all %>%
-  mutate( total_cases = count_no_psy_patients + count_psy_patients) %>%
-  select( siteid, time_p, time, month_year, count_psy_patients, total_cases)
+counts <- count_icd_all %>%
+  mutate( time = seq_along(time_p),
+         month_year = format(time_p, "%m")) %>%
+  select( siteid, time_p, time, month_year, count_psy_patients = count_psy, total_cases = count_icd )
 
@@
 ```{r}
-counts_agg <- patient_count_psy_period_psy_all %>%
+counts_agg <- counts %>%
   group_by( time_p ) %>%
   mutate( psy_cases = sum( count_psy_patients), 
-    total_cases = sum(count_no_psy_patients) + sum(count_psy_patients)) %>%
+    total_cases = sum(total_cases)) %>%
   select( time_p, time, month_year, psy_cases, total_cases)

### 2022-03-28 metanalysis_visualization/manuscript_Figures.Rmd
heatmap added the vline and monthly X axis

@@
          variableName = "example", variableType = "character", period = "month")
-plotDataTemporalMap(tempVariabilityInput, absolute = TRUE)
+
+vline <- function(x = 0, color = "green") {
+  list(
+    type = "line",
+    y0 = 0,
+    y1 = 1,
+    yref = "paper",
+    x0 = x,
+    x1 = x,
+    line = list(color = color, dash="dot"))
+}
+
+fig = plotDataTemporalMap(x)
+
+fig <- fig %>%
+  layout(
+    title = " ",
+    xaxis = list(
+      type = 'date',
+      tickformat = "%b-%y", 
+      tickangle=-45, 
+      tickvals=dates), 
+    shapes = list(vline(as.Date("2020-03-15"))))
+
+fig
 

### 2022-03-29 metanalysis_visualization/manuscript_Figures.Rmd
beginning working on ITS

--- 2022-03-23 metanalysis_visualization/manuscript_Figures.Rmd
@@
 }
+```
 
+```{r}
 # Get count_icd_all in the same format that patient_count_psy_period_psy_all, because first n last months missing ... 

### 2022-03-29 metanalysis_visualization/manuscript_Figures.Rmd
acronyms in plots changed by long version

--- 2022-03-28 metanalysis_visualization/manuscript_Figures.Rmd
@@
 }
-listOfPackages <- c("tidyverse", "tidyr","devtools","RColorBrewer", "knitr", "kableExtra", "tsModel", "gridExtra", "dplyr", "metafor", "meta", "viridis", "UpSetR", "scales", "EHRtemporalVariability")
+listOfPackages <- c("tidyverse", "tidyr","devtools","RColorBrewer", "knitr", "kableExtra", "tsModel", "gridExtra", "dplyr", "metafor", "meta", "viridis", "UpSetR", "scales", "EHRtemporalVariability", "plotly")
 paket(listOfPackages)
@@
 
-names_site_id <- unique(df_coefficients_boot$siteid)
+df_coefficients_boot <- df_coefficients_boot %>%
+  left_join( country_map )
+
+names_site_id <- unique(df_coefficients_boot$plotsLabel)
 
 plot_annotations <- df_coefficients_boot %>% 
-  group_by(siteid) %>% 
+  group_by(plotsLabel) %>% 
   group_map( annotation_function ) %>% 
@@
 plot_annotations_df <- data.frame(
-  siteid = names(plot_annotations), 
+  plotsLabel = names(plot_annotations), 
   label = plot_annotations
@@
 ### January 19-May 21
+bootstrapped_fitted_all <- bootstrapped_fitted_all %>%
+  left_join( country_map )
+
+quantile_bootstrapped_fitted_all <- quantile_bootstrapped_fitted_all %>%
+  left_join( country_map )
+
+count_icd_all_boot <- count_icd_all_boot %>%
+  left_join( country_map )
+
 ggplot(bootstrapped_fitted_all, aes(x = time_p, y = predicted, group = interaction(iteration, period))) +
-  facet_wrap("siteid", scales = "free_y") +
+  facet_wrap("plotsLabel", scales = "free_y", labeller = label_wrap_gen(width = 35,multi_line = TRUE)) +
   geom_line(aes(color = period),
@@
 ### Feb 19-April 21
+patient_count_psy_period_psy_all <- patient_count_psy_period_psy_all %>%
+  left_join( country_map )
+
 ggplot(bootstrapped_fitted_all, aes(x = time_p, y = predicted, group = interaction(iteration, period))) +
-  facet_wrap("siteid", scales = "free_y") +
+  facet_wrap("plotsLabel", scales = "free_y", labeller = label_wrap_gen(width = 35,multi_line = TRUE)) +
   geom_line(aes(color = period),
@@
 trends_siteid <- df_coefficients_boot %>% 
-  group_by(siteid) %>% 
+  group_by(plotsLabel) %>% 
   summarise(mean_boot = mean(`periodduring_pandemic:time`),
@@
 metafor::forest(model_meta_siteid,
-                slab = trends_siteid$siteid,
+                slab = trends_siteid$plotsLabel,
                 plim = c(1, NA),
                 order = "obs",
+                cex=0.7,
                 showweights = TRUE,
@@
 trends_country <- trends_siteid %>% 
-  left_join(country_map, by = "siteid") %>% 
+  left_join(country_map, by = "plotsLabel") %>% 
   group_by(country) %>% 
@@
 
-fig = plotDataTemporalMap(x)
+fig = plotDataTemporalMap(tempVariabilityInput, absolute = TRUE)
 
@@
 library(reshape2)
+count_icd_all <- count_icd_all %>%
+  left_join( country_map )
 
@@
          month_year = format(time_p, "%m")) %>%
-  select( siteid, time_p, time, month_year, count_psy_patients = count_psy, total_cases = count_icd )
+  select( plotsLabel, time_p, time, month_year, count_psy_patients = count_psy, total_cases = count_icd )
 
-counts2 <- reshape2::melt(counts ,  id.vars = c('siteid', 'time_p', 'time','month_year'), variable.name = 'group')
+counts2 <- reshape2::melt(counts ,  id.vars = c('plotsLabel', 'time_p', 'time','month_year'), variable.name = 'group')
 counts2$group <- gsub("count_psy_patients", "patients with psychiatric conditions", counts2$group )
@@
   geom_point( aes( shape= group, color = group )) +
-  facet_wrap2("siteid", axes = "all", scales ="free") + 
+  facet_wrap2("plotsLabel", axes = "all", scales ="free", labeller = label_wrap_gen(width = 35,multi_line = TRUE)) + 
   geom_vline(xintercept = as.Date(pandemic_start_date), linetype = "dashed") +

### 2022-03-31 metanalysis_visualization/manuscript_Figures.Rmd
ongoing ITS changes

--- 2022-03-29 metanalysis_visualization/manuscript_Figures.Rmd
@@
 
-### Sex Summary
-```{r}
-sexInfo_per_country <- table1_all %>%
-  filter( Category == "Sex") %>%
-  mutate ( counts_before = as.numeric(sapply(strsplit( as.character(before_pandemic), " "), '[', 1)), 
-           counts_during = as.numeric(sapply(strsplit( as.character(during_pandemic), " "), '[', 1))) %>%
-  select( name, siteid, counts_before, counts_during ) %>%
-  inner_join( country_map ) %>%
-  group_by( name, country ) %>%
-  summarise( before = sum( counts_before), 
-             during = sum( counts_during )) %>%
-  pivot_longer(names_to = "period", cols=c(before, during))
-
-sexInfo_all <- table1_all %>%
-  filter( Category == "Sex") %>%
-  mutate ( counts_before = as.numeric(sapply(strsplit( as.character(before_pandemic), " "), '[', 1)), 
-           counts_during = as.numeric(sapply(strsplit( as.character(during_pandemic), " "), '[', 1))) %>%
-  select( name, siteid, counts_before, counts_during ) %>%
-  inner_join( country_map ) %>%
-  group_by( name ) %>%
-  summarise( before = sum( counts_before), 
-             during = sum( counts_during )) %>%
-  pivot_longer(names_to = "period", cols=c(before, during)) %>%
-  mutate( country = 'ALL' ) %>%
-  select( name, country, period, value ) %>%
-  rbind( sexInfo_per_country)
-
-sexInfo_all %>%
-  ggplot( aes( x= name, y = value, fill = period )) +
-  geom_bar( stat = "identity", position = "dodge") +
-  facet_wrap("country", scales = "free_y") + 
-  labs(x ="", y = "number of patients") +
-  theme(axis.text.x=element_text(hjust=1, size = 8), 
-        strip.text = element_text(size = 10, face = "bold"),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), strip.background = element_blank(),
-        panel.border = element_rect(colour = "black", fill = NA))+ 
-  scale_fill_manual("Pandemic \n period", values = c("before" = "grey", "during" = "orange"))
-
-```
-
-```{r}
-patients_hospitalized_agg_sex_total <- patients_hospitalized_agg_sex_perc_all %>%
-  select( - percentage ) %>%
-  group_by( time_p, sex ) %>%
-  mutate( total_cases = sum( count ), 
-          sex_cases = sum( count_sex ), 
-          percentage = 100* sex_cases / total_cases, 
-          country = "ALL")
-
-patients_hospitalized_agg_sex_total %>%
-  ggplot( aes( x= as.Date( time_p ), y = percentage, group = sex )) +
-  geom_line( aes( linetype = sex , color = sex)) +
-  geom_point( aes( shape = sex, color = sex )) +
-  scale_colour_manual(values=c("#FF6600","#339999")) +
-  scale_fill_manual(values=c("#FF6600","#339999"))+
-  geom_vline(xintercept = as.Date(pandemic_start_date), linetype = "dashed")+
-  labs(x ="", y = "percentage of patients (%)")+
-  scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0))+
-  theme(axis.text.x=element_text(angle = 45, hjust=1, size = 6),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.border = element_rect(colour = "black", fill = NA))
-
-#patients_hospitalized_agg_sex_country <- patients_hospitalized_agg_sex_perc_all %>%
-#  select( - percentage ) %>%
-#  inner_join( country_map ) %>%
-#  group_by( time_p, sex, country ) %>%
-#  mutate( total_cases = sum( count ), 
-#          sex_cases = sum( count_sex ), 
-#          percentage = 100* sex_cases / total_cases) %>%
-#  rbind( patients_hospitalized_agg_sex_total)
-
-# patients_hospitalized_agg_sex_country %>%
-#   ggplot( aes( x= as.Date( time_p ), y = percentage, group = sex )) +
-#   geom_line( aes(linetype = sex, color = sex )) +
-#   geom_point( aes( shape= sex, color = sex )) +
-#   scale_colour_manual(values=c("#FF6600","#339999")) +
-#   scale_fill_manual(values=c("#FF6600","#339999"))+
-#   #scale_y_continuous(breaks = seq(0, 30, by = 1)) +
-#   facet_wrap2("country", axes = "all") + 
-#   geom_vline(xintercept = as.Date(pandemic_start_date), linetype = "dashed") +
-#   labs(x ="", y = "percentage of patients (%)")+
-#   scale_x_date(breaks = "2 month", labels=date_format("%b-%y"), expand=c(0.03,0)) +
-#   theme(axis.text.x=element_text(angle = 45, hjust=1, size = 6), 
-#         strip.text = element_text(size = 10, face = "bold"),
-#         panel.grid.major = element_blank(), panel.grid.minor = element_blank(), strip.background = element_blank(),
-#         panel.border = element_rect(colour = "black", fill = NA), 
-#         legend.position="bottom")
-```
-
 ## Interrupted Time Series Analysis
@@
 
-## Recompute the bootstrap
+## Define bootstrap functions
 
@@
 
-bootstrapping_slopes <- function(patient_count_psy_period_psy, 
-                                    bootstrap_size) {
-  bootstrapped_timepoints <- mapply(binom_proportion,
+bootstrapping_slopes <- function(df, 
+                                 bootstrap_size, 
+                                 count = "count_icd", 
+                                 percentage = "percentage", 
+                                 group_name = .y) {
+  bootstrapped_timepoints <- mapply(
+    binom_proportion,
     n = bootstrap_size,
-                                    size = patient_count_psy_period_psy$count_no_psy_patients +
-                                      patient_count_psy_period_psy$count_psy_patients,
-                                    prob = patient_count_psy_period_psy$percentage/100)
+    size = df[[count]],
+    prob = df[[percentage]]/100)
   bootstrapped_sample_list <- vector(mode = "list", length = bootstrap_size)
@@
   for (iteration in 1:bootstrap_size) {
-    df_iteration <- data.frame(period = patient_count_psy_period_psy$period,
-                               time = patient_count_psy_period_psy$time,
+    df_iteration <- data.frame(period = df$period,
+                               time = df$time,
                                percentage = bootstrapped_timepoints[iteration, ])
-    ts_linear_percent <- lm(percentage ~ period * time, df_iteration)
+    ts_linear_percent <- lm(as.formula(paste0(percentage, "~ period * time")), df_iteration)
     linear_model <- broom::tidy(ts_linear_percent) %>%
@@
     df_iteration$predicted <- predict(ts_linear_percent)
-    df_iteration$time_p <- patient_count_psy_period_psy$time_p
+    df_iteration$time_p <- df$time_p
     bootstrapped_sample_list[[iteration]] <- df_iteration
   }
-  bootstrapped_fitted_df <- bind_rows(bootstrapped_sample_list, .id = "iteration")
+  bootstrapped_fitted_df <- bind_rows(bootstrapped_sample_list, .id = "iteration") %>% 
+    mutate(group = group_name)
   bootstrapped_coefficients_df <- bind_rows(bootstrapped_coefficients_list,
-                                          .id = "iteration")
+                                          .id = "iteration") %>% 
+    mutate(group = group_name)
   return(list(bootstrapped_coefficients_df, 
@@
   group_by(siteid) %>% # necessary to generate time correctly 
-  rename(count_psy_patients = count_psy, 
-         percentage = percentage_psy) %>% 
-  mutate(count_no_psy_patients = count_icd - count_psy_patients, 
-         month_year = format(time_p, "%m"),
+  rename(percentage = percentage_psy) %>% 
+  mutate(month_year = format(time_p, "%m"),
          time = seq_along(time_p))
@@
   group_map( ~bootstrapping_slopes(.x, bootstrap_size = bootstrap_size))
-names(bootstrapped) <- unique(count_icd_all$siteid)
+names(bootstrapped) <- unique(count_icd_all_boot$siteid)
 
@@
 ### Primary analysis
+
+```{r}
+annotation_function <- function(df, ...) {
+  z <- unique(df[["observed_diff"]]) / sd(df[["periodduring_pandemic:time"]])
+  pvalue_coeff_interaction = round(pnorm(abs(z), lower.tail = FALSE) * 2, 3)
+  paste0("Coeffs differences: ",
+         round(unique(df[["observed_diff"]]), 2),
+         " (pvalue: ",
+         pvalue_coeff_interaction,
+         ")")
+}
+```
 
@@
                names_to = "quantiles")
+```
 
-annotation_function <- function(df, ...) {
-  z <- unique(df[["observed_diff"]]) / sd(df[["periodduring_pandemic:time"]])
-  pvalue_coeff_interaction = round(pnorm(abs(z), lower.tail = FALSE) * 2, 3)
-  paste0("Coeffs differences: ",
-         round(unique(df[["observed_diff"]]), 2),
-         " (pvalue: ",
-         pvalue_coeff_interaction,
-         ")")
-}
-
+```{r}
 df_coefficients_boot <- bootstrapped_coefficients_all %>% 
@@
 ```{r}
-# Random effect effects
+# Random effect
 model_meta_siteid <- metafor::rma(yi = observed, 
@@
        mlab = "Overall periods' slope differences")
+```
+
+### Sex Summary
+
+```{r}
+sexInfo_per_country <- table1_all %>%
+  filter( Category == "Sex") %>%
+  mutate ( counts_before = as.numeric(sapply(strsplit( as.character(before_pandemic), " "), '[', 1)), 
+           counts_during = as.numeric(sapply(strsplit( as.character(during_pandemic), " "), '[', 1))) %>%
+  select( name, siteid, counts_before, counts_during ) %>%
+  inner_join( country_map ) %>%
+  group_by( name, country ) %>%
+  summarise( before = sum( counts_before), 
+             during = sum( counts_during )) %>%
+  pivot_longer(names_to = "period", cols=c(before, during))
+
+sexInfo_all <- table1_all %>%
+  filter( Category == "Sex") %>%
+  mutate ( counts_before = as.numeric(sapply(strsplit( as.character(before_pandemic), " "), '[', 1)), 
+           counts_during = as.numeric(sapply(strsplit( as.character(during_pandemic), " "), '[', 1))) %>%
+  select( name, siteid, counts_before, counts_during ) %>%
+  inner_join( country_map ) %>%
+  group_by( name ) %>%
+  summarise( before = sum( counts_before), 
+             during = sum( counts_during )) %>%
+  pivot_longer(names_to = "period", cols=c(before, during)) %>%
+  mutate( country = 'ALL' ) %>%
+  select( name, country, period, value ) %>%
+  rbind( sexInfo_per_country)
+
+sexInfo_all %>%
+  ggplot( aes( x= name, y = value, fill = period )) +
+  geom_bar( stat = "identity", position = "dodge") +
+  facet_wrap("country", scales = "free_y") + 
+  labs(x ="", y = "number of patients") +
+  theme(axis.text.x=element_text(hjust=1, size = 8), 
+        strip.text = element_text(size = 10, face = "bold"),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), strip.background = element_blank(),
+        panel.border = element_rect(colour = "black", fill = NA))+ 
+  scale_fill_manual("Pandemic \n period", values = c("before" = "grey", "during" = "orange"))
+
+```
+
+```{r}
+patients_hospitalized_agg_sex_total <- patients_hospitalized_agg_sex_perc_all %>%
+  select( - percentage ) %>%
+  group_by( time_p, sex ) %>%
+  mutate( total_cases = sum( count ), 
+          sex_cases = sum( count_sex ), 
+          percentage = 100* sex_cases / total_cases, 
+          country = "ALL")
+
+patients_hospitalized_agg_sex_total %>%
+  ggplot( aes( x= as.Date( time_p ), y = percentage, group = sex )) +
+  geom_line( aes( linetype = sex , color = sex)) +
+  geom_point( aes( shape = sex, color = sex )) +
+  scale_colour_manual(values=c("#FF6600","#339999")) +
+  scale_fill_manual(values=c("#FF6600","#339999"))+
+  geom_vline(xintercept = as.Date(pandemic_start_date), linetype = "dashed")+
+  labs(x ="", y = "percentage of patients (%)")+
+  scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0))+
+  theme(axis.text.x=element_text(angle = 45, hjust=1, size = 6),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.border = element_rect(colour = "black", fill = NA))
+
+```
+
+```{r Interrupted time series by sex}
+
+  # pivot_wider(names_from =  sex, values_from = c(count, percentage))
+  # group_by(siteid) %>% # necessary to generate time correctly 
+  # rename(count_psy_patients = count_psy, 
+  #        percentage = percentage_psy) %>% 
+  # mutate(count_no_psy_patients = count_icd - count_psy_patients, 
+  #        month_year = format(time_p, "%m"),
+  #        time = seq_along(time_p))
+
+bootstrap_size = 10
+
+patients_hospitalized_agg_sex_perc_all <- patients_hospitalized_agg_sex_perc_all %>% 
+  mutate(period = ifelse(time_p <= as.Date("2020-03-15"), "before_pandemic", "during_pandemic")) %>% 
+  select(time_p, period, sex, count, percentage, siteid)
+
+bootstrapped_sex <- patients_hospitalized_agg_sex_perc_all %>%
+  group_by(siteid, sex) %>% 
+  mutate(month_year = format(time_p, "%m"),
+         time = seq_along(time_p))  %>% 
+ group_map( ~bootstrapping_slopes(.x, bootstrap_size = bootstrap_size, count = "count"))
+names(bootstrapped_sex) <- unique(patients_hospitalized_agg_sex_perc_all$siteid)
+
+bootstrapped_coefficients_sex <- lapply(bootstrapped_sex, `[[`, 1) %>% bind_rows(.id = "siteid")
+bootstrapped_fitted_sex <- lapply(bootstrapped_sex, `[[`, 2) %>% bind_rows(.id = "siteid")
+
+quantile_bootstrapped_fitted_sex <- bootstrapped_fitted_sex %>%
+  group_by(time_p, period, siteid) %>%
+  summarise("2.5%" = quantile(predicted, 0.025),
+            "50%" = median(predicted, 0.5),
+            "97.5%" = quantile(predicted, 0.975)) %>%
+  pivot_longer(cols = c("2.5%", "50%", "97.5%"),
+               names_to = "quantiles")
+
+```
+
+```{r}
+coefficients_df_sex <- patients_hospitalized_agg_sex_perc_all %>% 
+  group_by(siteid, sex) %>% 
+    mutate(month_year = format(time_p, "%m"),
+         time = seq_along(time_p))  %>% 
+  group_modify(coefficient_slope_calculation) %>% 
+  rename(observed_diff = estimate)
+
+```
+
+```{r}
+df_coefficients_boot_sex <- bootstrapped_coefficients_sex %>% 
+  left_join(coefficients_df_sex, by = c("siteid", "sex"))
+```
+
+```{r}
+names_site_id <- unique(df_coefficients_boot_sex$siteid)
+
+plot_annotations <- df_coefficients_boot_sex %>% 
+  group_by(siteid, sex) %>% 
+  group_map( annotation_function ) %>% 
+  set_names(names_site_id) %>% 
+  unlist()
+plot_annotations_df <- data.frame(
+  siteid = names(plot_annotations), 
+  label = plot_annotations
+)
 ```

### 2022-03-31 psy_peds_3.Rmd
correcting study dates

@@
 # cut-off dates used in the analysis, these should not be changed
-history_start_date <- as.Date("2019-01-01")
-history_end_date <- as.Date("2021-05-31")
+# history_start_date <- as.Date("2019-01-01")
+# history_end_date <- as.Date("2021-05-31")
 pandemic_start_date <- as.Date("2020-03-15")
-start_date_plots <- as.Date("2019-01-15")
-end_date_plots <- as.Date("2021-05-01")
+start_date_plots <- as.Date("2019-01-01")
+end_date_plots <- as.Date("2021-06-01")
 ```

### 2022-03-31 psy_peds_3.Rmd
correcting study dates

--- 2022-03-23 psy_peds_3.Rmd
@@
 # cut-off dates used in the analysis, these should not be changed
-history_start_date <- as.Date("2019-01-01")
-history_end_date <- as.Date("2021-05-31")
+# history_start_date <- as.Date("2019-01-01")
+# history_end_date <- as.Date("2021-05-31")
 pandemic_start_date <- as.Date("2020-03-15")
-start_date_plots <- as.Date("2019-01-15")
-end_date_plots <- as.Date("2021-05-01")
+start_date_plots <- as.Date("2018-12-01")
+end_date_plots <- as.Date("2021-07-01")
 ```

### 2022-03-31 metanalysis_visualization/manuscript_Figures.Rmd
determine end date as may 2021 for plots and add consistent x axis for all the plots

@@
 time_period <-"months" #possible values: days, weeks, months
-history_start_date <- as.Date("2019-01-01")
-history_end_date <- as.Date("2021-05-31")
+#history_start_date <- as.Date("2019-01-01")
+#history_end_date <- as.Date("2021-05-31")
 pandemic_start_date <- as.Date("2020-03-15")
@@
 for( i in 1:length( outputFiles )){
-  print(i)
+  print(outputFiles[i])
   load(paste0( outputDir, outputFiles[i]))
@@
 
+# Filter by dates
+```{r}
+start_date_plots <- as.Date("2019-01-01")
+end_date_plots <- as.Date("2021-05-31")
+
+bootstrapped_fitted_all <-  bootstrapped_fitted_all %>%
+  filter( time_p <= end_date_plots &
+          time_p >= start_date_plots)
+
+count_icd_all <-  count_icd_all %>%
+  filter( time_p <= end_date_plots &
+          time_p >= start_date_plots)
+
+patient_count_psy_period_psy_all <- patient_count_psy_period_psy_all %>%
+  filter( time_p <= end_date_plots &
+          time_p >= start_date_plots)
+
+patients_hospitalized_agg_sex_perc_all <- patients_hospitalized_agg_sex_perc_all %>%
+  filter( time_p <= end_date_plots &
+          time_p >= start_date_plots)
+
+perc_disorder_group_all <- perc_disorder_group_all %>%
+  filter( time_p <= end_date_plots &
+          time_p >= start_date_plots)
+
+ratios_patients_with_without_psy_all <- ratios_patients_with_without_psy_all %>%  
+  filter( time_p <= end_date_plots &
+          time_p >= start_date_plots)
+```
+
 ### Table 1
@@
 ```{r Extracting age information}
-
 age_table1_all <- table1_all %>% 
@@
             inherit.aes = FALSE) +
+  scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0))+
   labs(title = "Interrupted Time Series Analysis: Percentage of patients with psychiatric conditions",
@@
   theme(strip.text = element_text(size = 10),
-        axis.text.x = element_text(size = 8), 
+        axis.text.x=element_text(angle = 45, hjust=1, size = 8),
         axis.text.y = element_text(size = 8), 
@@
 ggsave("boostrapped_all_extended.png", height = 20, width = 32, units = "cm")
-```
-
-```{r}
-### Feb 19-April 21
-patient_count_psy_period_psy_all <- patient_count_psy_period_psy_all %>%
-  left_join( country_map )
-
-ggplot(bootstrapped_fitted_all, aes(x = time_p, y = predicted, group = interaction(iteration, period))) +
-  facet_wrap("plotsLabel", scales = "free_y", labeller = label_wrap_gen(width = 35,multi_line = TRUE)) +
-  geom_line(aes(color = period),
-            alpha = 0.03) +
-  geom_line(data = quantile_bootstrapped_fitted_all,
-            aes(y = value, group = interaction(quantiles, period)),
-            linetype = "dashed") +
-  geom_line(data = quantile_bootstrapped_fitted_all[quantile_bootstrapped_fitted_all$quantiles == "50%", ],
-            aes(y = value, group = period)) +
-  geom_point(data = patient_count_psy_period_psy_all,
-             aes(y = percentage, group=NULL),
-             alpha = 0.5) +
-  scale_colour_manual(values = cbPalette) +
-  guides(color = guide_legend(override.aes = list(alpha = 1))) +
-  geom_label(data = plot_annotations_df, 
-             aes(label=label), 
-            x = -Inf,
-            y = +Inf,
-            hjust = 0,
-            vjust = 1,
-            size = 4,
-            inherit.aes = FALSE) +
-  labs(title = "Interrupted Time Series Analysis: Percentage of patients with psychiatric conditions",
-       subtitle = "Parametric bootstrap estimation: (1000 iterations)",
-       x = "Calendar dates",
-       y = "Percentage",
-       color = "Period") +
-  theme(strip.text = element_text(size = 10),
-        axis.text.x = element_text(size = 8), 
-        axis.text.y = element_text(size = 8), 
-        axis.title.x = element_text(size = 8),
-        axis.title.y = element_text(size = 8),
-        title = element_text(size = 14),
-        plot.subtitle = element_text(size = 12, face = "italic")) 
-ggsave("boostrapped_all.png", height = 20, width = 32, units = "cm")
 ```
@@
        y = "Percentage") +
+      scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0)) +
      theme(strip.text = element_text(size = 10),
-        axis.text.x = element_text(size = 8), 
+        axis.text.x = element_text(angle = 45, hjust=1, size = 8), 
         axis.text.y = element_text(size = 8), 
@@
 
-counts2 %>% ggplot( aes( x= as.Date( time_p ), y = value, group= group )) +
+counts_plot <- counts2 %>% ggplot( aes( x= as.Date( time_p ), y = value, group= group )) +
   geom_point( aes( shape= group, color = group )) +
@@
         legend.position="bottom")
+counts_plot
+ggsave("counts_plot.png",counts_plot, height = 20, width = 32, units = "cm")
+
 ```

### 2022-04-01 metanalysis_visualization/manuscript_Figures.Rmd
ITS sex implementation

--- 2022-03-31 metanalysis_visualization/manuscript_Figures.Rmd
@@
 outputFiles <- list.files( path = outputDir, pattern = "\\.RData")
-country_map <- read.delim( file = paste0( outputDir, "country_mapping.txt"))
+country_map <- read.delim( file = "../public-data/country_mapping.txt")
 ```
@@
 bootstrapping_slopes <- function(df,
+                                 group_name,
                                  bootstrap_size,
                                  count = "count_icd",
-                                 percentage = "percentage", 
-                                 group_name = .y) {
+                                 percentage = "percentage") {
   bootstrapped_timepoints <- mapply(
@@
   bootstrapped_fitted_df <- bind_rows(bootstrapped_sample_list, .id = "iteration") %>%
-    mutate(group = group_name)
+    bind_cols(group_name)
   bootstrapped_coefficients_df <- bind_rows(bootstrapped_coefficients_list,
                                             .id = "iteration") %>%
-    mutate(group = group_name)
+    bind_cols(group_name)
   return(list(bootstrapped_coefficients_df,
@@
   group_by(siteid) %>% 
-  group_map( ~bootstrapping_slopes(.x, bootstrap_size = bootstrap_size))
+  group_map( ~bootstrapping_slopes(.x, group_name = .y, bootstrap_size = bootstrap_size))
 names(bootstrapped) <- unique(count_icd_all_boot$siteid)
 
-bootstrapped_coefficients_all <- lapply(bootstrapped, `[[`, 1) %>% bind_rows(.id = "siteid")
-bootstrapped_fitted_all <- lapply(bootstrapped, `[[`, 2) %>% bind_rows(.id = "siteid")
+bootstrapped_coefficients_all <- lapply(bootstrapped, `[[`, 1) %>% bind_rows()
+bootstrapped_fitted_all <- lapply(bootstrapped, `[[`, 2) %>% bind_rows()
 ```
@@
 ```{r}
-### Feb 19-April 21
 patient_count_psy_period_psy_all <- patient_count_psy_period_psy_all %>%
@@
 ```{r Interrupted time series by sex}
-
-  # pivot_wider(names_from =  sex, values_from = c(count, percentage))
-  # group_by(siteid) %>% # necessary to generate time correctly 
-  # rename(count_psy_patients = count_psy, 
-  #        percentage = percentage_psy) %>% 
-  # mutate(count_no_psy_patients = count_icd - count_psy_patients, 
-  #        month_year = format(time_p, "%m"),
-  #        time = seq_along(time_p))
-
 bootstrap_size = 10
@@
          time = seq_along(time_p))  %>% 
- group_map( ~bootstrapping_slopes(.x, bootstrap_size = bootstrap_size, count = "count"))
-names(bootstrapped_sex) <- unique(patients_hospitalized_agg_sex_perc_all$siteid)
+ group_map( ~bootstrapping_slopes(.x, group_name = .y, bootstrap_size = bootstrap_size, count = "count"))
 
-bootstrapped_coefficients_sex <- lapply(bootstrapped_sex, `[[`, 1) %>% bind_rows(.id = "siteid")
-bootstrapped_fitted_sex <- lapply(bootstrapped_sex, `[[`, 2) %>% bind_rows(.id = "siteid")
+bootstrapped_coefficients_sex <- lapply(bootstrapped_sex, `[[`, 1) %>% bind_rows()
+bootstrapped_fitted_sex <- lapply(bootstrapped_sex, `[[`, 2) %>% bind_rows()
+```
 
+```{r Interrupted time series by sex}
 quantile_bootstrapped_fitted_sex <- bootstrapped_fitted_sex %>%
-  group_by(time_p, period, siteid) %>%
+  group_by(time_p, period, siteid, sex) %>%
   summarise("2.5%" = quantile(predicted, 0.025),
@@
                names_to = "quantiles")
-
 ```
@@
 ```{r}
-names_site_id <- unique(df_coefficients_boot_sex$siteid)
+ggplot(bootstrapped_fitted_all, aes(x = time_p, y = predicted, group = interaction(iteration, period))) +
+  facet_wrap("plotsLabel", scales = "free_y", labeller = label_wrap_gen(width = 35,multi_line = TRUE)) +
+  geom_line(aes(color = period),
+            alpha = 0.03) +
+  geom_line(data = quantile_bootstrapped_fitted_all,
+            aes(y = value, group = interaction(quantiles, period)),
+            linetype = "dashed") +
+  geom_line(data = quantile_bootstrapped_fitted_all[quantile_bootstrapped_fitted_all$quantiles == "50%", ],
+            aes(y = value, group = period)) +
+  geom_point(data = patient_count_psy_period_psy_all,
+             aes(y = percentage, group=NULL),
+             alpha = 0.5) +
+  scale_colour_manual(values = cbPalette) +
+  guides(color = guide_legend(override.aes = list(alpha = 1))) +
+  geom_label(data = plot_annotations_df, 
+             aes(label=label), 
+            x = -Inf,
+            y = +Inf,
+            hjust = 0,
+            vjust = 1,
+            size = 4,
+            inherit.aes = FALSE) +
+  labs(title = "Interrupted Time Series Analysis: Percentage of patients with psychiatric conditions",
+       subtitle = "Parametric bootstrap estimation: (1000 iterations)",
+       x = "Calendar dates",
+       y = "Percentage",
+       color = "Period") +
+  theme(strip.text = element_text(size = 10),
+        axis.text.x = element_text(size = 8), 
+        axis.text.y = element_text(size = 8), 
+        axis.title.x = element_text(size = 8),
+        axis.title.y = element_text(size = 8),
+        title = element_text(size = 14),
+        plot.subtitle = element_text(size = 12, face = "italic")) 
+```
 
-plot_annotations <- df_coefficients_boot_sex %>% 
-  group_by(siteid, sex) %>% 
-  group_map( annotation_function ) %>% 
-  set_names(names_site_id) %>% 
-  unlist()
+```{r}
+# expand.grid(
+#   unique(df_coefficients_boot_sex$siteid),
+#   unique(df_coefficients_boot_sex$sex)
+# ) %>% 
+#   t() %>% 
+#   paste(unique(.$siteid),
+#         unique(.$sex), 
+#         collapse = " "
+#   )
+# 
+# names_plot_annotations_sex <- c()
+# for (site in unique(df_coefficients_boot_sex$siteid)) {
+#   for (sex in unique(df_coefficients_boot_sex$sex)) {
+#     names_plot_annotations_sex <- c(names_plot_annotations_sex, paste(site, sex))
+#   }
+# }
+
+# plot_annotations_sex <- df_coefficients_boot_sex %>% 
+#   group_by(siteid, sex) %>% 
+#   group_map( annotation_function ) %>% 
+#   set_names(names_site_id) %>% 
+#   unlist()
 plot_annotations_df <- data.frame(
   siteid = names(plot_annotations), 
-  label = plot_annotations
+  # label = plot_annotations
 )

### 2022-04-01 metanalysis_visualization/manuscript_Figures.Rmd

--- 2022-03-31 metanalysis_visualization/manuscript_Figures.Rmd
@@
 ggsave("boostrapped_all_extended.png", height = 20, width = 32, units = "cm")
+```
+
+```{r}
+patient_count_psy_period_psy_all <- patient_count_psy_period_psy_all %>%
+  left_join( country_map )
+
+ggplot(bootstrapped_fitted_all, aes(x = time_p, y = predicted, group = interaction(iteration, period))) +
+  facet_wrap("plotsLabel", scales = "free_y", labeller = label_wrap_gen(width = 35,multi_line = TRUE)) +
+  geom_line(aes(color = period),
+            alpha = 0.03) +
+  geom_line(data = quantile_bootstrapped_fitted_all,
+            aes(y = value, group = interaction(quantiles, period)),
+            linetype = "dashed") +
+  geom_line(data = quantile_bootstrapped_fitted_all[quantile_bootstrapped_fitted_all$quantiles == "50%", ],
+            aes(y = value, group = period)) +
+  geom_point(data = patient_count_psy_period_psy_all,
+             aes(y = percentage, group=NULL),
+             alpha = 0.5) +
+  scale_colour_manual(values = cbPalette) +
+  guides(color = guide_legend(override.aes = list(alpha = 1))) +
+  geom_label(data = plot_annotations_df, 
+             aes(label=label), 
+            x = -Inf,
+            y = +Inf,
+            hjust = 0,
+            vjust = 1,
+            size = 4,
+            inherit.aes = FALSE) +
+  labs(title = "Interrupted Time Series Analysis: Percentage of patients with psychiatric conditions",
+       subtitle = "Parametric bootstrap estimation: (1000 iterations)",
+       x = "Calendar dates",
+       y = "Percentage",
+       color = "Period") +
+  theme(strip.text = element_text(size = 10),
+        axis.text.x = element_text(size = 8), 
+        axis.text.y = element_text(size = 8), 
+        axis.title.x = element_text(size = 8),
+        axis.title.y = element_text(size = 8),
+        title = element_text(size = 14),
+        plot.subtitle = element_text(size = 12, face = "italic")) 
+ggsave("boostrapped_all.png", height = 20, width = 32, units = "cm")
 ```

### 2022-04-01 psy_peds_3_CHOP.Rmd

@@
 start_date_plots <- as.Date("2019-01-15")
-end_date_plots <- as.Date("2021-05-01")
+end_date_plots <- as.Date("2021-06-01")
 ```
@@
 
+# Add the length of hospitalization data
+```{r}
+length_hospitalisation_values <- read.delim("../output/CHOP/length_hospitalization_formated.txt") %>%
+  mutate( period = tolower( Period )) %>%
+  select(  len_hospitalisation = Hospitalization.Length,  siteid, period)
+```
+
 # Save the output in a RData file. 
@@
       mental_codes_qc_output,
+      length_hospitalisation_values,
       table1,

### 2022-04-04 metanalysis_visualization/manuscript_Figures.Rmd
sex difference bootstrap

@@
 
-bootstrapped_fitted_all <-  bootstrapped_fitted_all %>%
-  filter( time_p <= end_date_plots &
-          time_p >= start_date_plots)
+filter_by_dates <- function(df_,
+                            start_date_plot,
+                            end_date_plots) {
+  df_[df_[["time_p"]] <= end_date_plots & df_[["time_p"]] >= start_date_plots, ]
+}
 
-count_icd_all <-  count_icd_all %>%
-  filter( time_p <= end_date_plots &
-          time_p >= start_date_plots)
+bootstrapped_fitted_all <- filter_by_dates(bootstrapped_fitted_all, start_date_plots, end_date_plots)
+count_icd_all <- filter_by_dates(count_icd_all, start_date_plots, end_date_plots)
+patient_count_psy_period_psy_all <- filter_by_dates(patient_count_psy_period_psy_all, start_date_plots, end_date_plots)
+patients_hospitalized_agg_sex_perc_all <- filter_by_dates(patients_hospitalized_agg_sex_perc_all, start_date_plots, end_date_plots)
+perc_disorder_group_all <- filter_by_dates(perc_disorder_group_all, start_date_plots, end_date_plots)
+ratios_patients_with_without_psy_all <- filter_by_dates(ratios_patients_with_without_psy_all, start_date_plots, end_date_plots)
+```
 
-patient_count_psy_period_psy_all <- patient_count_psy_period_psy_all %>%
-  filter( time_p <= end_date_plots &
-          time_p >= start_date_plots)
+# Add site complete name information
 
-patients_hospitalized_agg_sex_perc_all <- patients_hospitalized_agg_sex_perc_all %>%
-  filter( time_p <= end_date_plots &
-          time_p >= start_date_plots)
+```{r}
+bootstrapped_fitted_all <- left_join(bootstrapped_fitted_all, country_map, on = "site_id")
+count_icd_all <- left_join(count_icd_all, country_map, on = "site_id")
+patient_count_psy_period_psy_all <- left_join(patient_count_psy_period_psy_all, country_map, on = "site_id")
+patients_hospitalized_agg_sex_perc_all <- left_join(patients_hospitalized_agg_sex_perc_all, country_map, on = "site_id")
+perc_disorder_group_all <- left_join(perc_disorder_group_all, country_map, on = "site_id")
+ratios_patients_with_without_psy_all <- left_join(ratios_patients_with_without_psy_all, country_map, on = "site_id")
 
-perc_disorder_group_all <- perc_disorder_group_all %>%
-  filter( time_p <= end_date_plots &
-          time_p >= start_date_plots)
-
-ratios_patients_with_without_psy_all <- ratios_patients_with_without_psy_all %>%  
-  filter( time_p <= end_date_plots &
-          time_p >= start_date_plots)
 ```
@@
   z <- unique(df[["observed_diff"]]) / sd(df[["periodduring_pandemic:time"]])
-  pvalue_coeff_interaction = round(pnorm(abs(z), lower.tail = FALSE) * 2, 3)
+  pvalue_coeff_interaction <- round(pnorm(abs(z), lower.tail = FALSE) * 2, 3)
   paste0("Coeffs differences: ",
@@
          " (pvalue: ",
-         pvalue_coeff_interaction,
+         format.pval(pvalue_coeff_interaction),
          ")")
@@
 df_coefficients_boot <- bootstrapped_coefficients_all %>% 
-  left_join(coefficients_df, by = "siteid")
-
-df_coefficients_boot <- df_coefficients_boot %>%
-  left_join( country_map )
+  left_join(coefficients_df, by = "siteid") %>% 
+  left_join(country_map, by = "siteid")
 
@@
 ```{r}
-### January 19-May 21
-bootstrapped_fitted_all <- bootstrapped_fitted_all %>%
-  left_join( country_map )
+bootstrapped_fitted_all <- bootstrapped_fitted_all %>% left_join(country_map, by = "siteid")
 
-quantile_bootstrapped_fitted_all <- quantile_bootstrapped_fitted_all %>%
-  left_join( country_map )
-
-count_icd_all_boot <- count_icd_all_boot %>%
-  left_join( country_map )
-
-ggplot(bootstrapped_fitted_all, aes(x = time_p, y = predicted, group = interaction(iteration, period))) +
-  facet_wrap("plotsLabel", scales = "free_y", labeller = label_wrap_gen(width = 35,multi_line = TRUE)) +
-  geom_line(aes(color = period),
-            alpha = 0.03) +
-  geom_line(data = quantile_bootstrapped_fitted_all,
-            aes(y = value, group = interaction(quantiles, period)),
-            linetype = "dashed") +
-  geom_line(data = quantile_bootstrapped_fitted_all[quantile_bootstrapped_fitted_all$quantiles == "50%", ],
-            aes(y = value, group = period)) +
-  geom_point(data = count_icd_all_boot,
-             aes(y = percentage, group=NULL),
-             alpha = 0.5) +
-  scale_colour_manual(values = cbPalette) +
-  guides(color = guide_legend(override.aes = list(alpha = 1))) +
-  geom_label(data = plot_annotations_df, 
-             aes(label=label), 
-            x = -Inf,
-            y = +Inf,
-            hjust = 0,
-            vjust = 1,
-            size = 4,
-            inherit.aes = FALSE) +
-  scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0))+
-  labs(title = "Interrupted Time Series Analysis: Percentage of patients with psychiatric conditions",
-       subtitle = "Parametric bootstrap estimation: (1000 iterations)",
-       x = "Calendar dates",
-       y = "Percentage",
-       color = "Period") +
-  theme(strip.text = element_text(size = 10),
-        axis.text.x=element_text(angle = 45, hjust=1, size = 8),
-        axis.text.y = element_text(size = 8), 
-        axis.title.x = element_text(size = 8),
-        axis.title.y = element_text(size = 8),
-        title = element_text(size = 14),
-        plot.subtitle = element_text(size = 12, face = "italic")) 
-ggsave("boostrapped_all_extended.png", height = 20, width = 32, units = "cm")
+quantile_bootstrapped_fitted_all <- quantile_bootstrapped_fitted_all %>% left_join(country_map, by = "siteid")
 ```
@@
 ```{r}
-patient_count_psy_period_psy_all <- patient_count_psy_period_psy_all %>%
-  left_join( country_map )
-
 ggplot(bootstrapped_fitted_all, aes(x = time_p, y = predicted, group = interaction(iteration, period))) +
@@
 trends_country <- trends_siteid %>% 
-  left_join(country_map, by = "plotsLabel") %>% 
-  group_by(country) %>% 
+  group_by(plotsLabel) %>% 
   group_modify(.f = function(x, ...) {
@@
 forest(model_meta_country, 
-       slab = trends_country$country,
+       slab = trends_country$plotsLabel,
        plim = c(1, NA),
@@
   pivot_longer(names_to = "period", cols=c(before, during)) 
-
 sexInfo_all <- table1_all %>%
@@
   rbind( sexInfo_per_country)
-
 sexInfo_all %>%
@@
   group_by( time_p, sex ) %>%
-  mutate( total_cases = sum( count ), 
-          sex_cases = sum( count_sex ), 
-          percentage = 100* sex_cases / total_cases, 
-          country = "ALL")
+  summarise( total = sum( count ), 
+             count_sex = sum( count_sex ), 
+             percentage = 100* count_sex / total, 
+             siteid = "ALL")
 
@@
 ```{r Interrupted time series by sex}
-bootstrap_size = 10
 
-patients_hospitalized_agg_sex_perc_all <- patients_hospitalized_agg_sex_perc_all %>% 
+patients_hospitalized_agg_sex_total <- patients_hospitalized_agg_sex_total %>% 
   mutate(period = ifelse(time_p <= as.Date("2020-03-15"), "before_pandemic", "during_pandemic")) %>% 
-  select(time_p, period, sex, count, percentage, siteid)
+  select(time_p, period, sex, count_sex, percentage, siteid)
+# %>% 
+#   inner_join(country_map)
 
-bootstrapped_sex <- patients_hospitalized_agg_sex_perc_all %>%
-  group_by(siteid, sex) %>% 
+bootstrapped_sex <- patients_hospitalized_agg_sex_total %>%
+  group_by(sex) %>% 
+  # group_by(plotsLabel, sex) %>% 
   mutate(month_year = format(time_p, "%m"),
          time = seq_along(time_p))  %>% 
- group_map( ~bootstrapping_slopes(.x, group_name = .y, bootstrap_size = bootstrap_size, count = "count"))
+ group_map(~ bootstrapping_slopes(.x,
+                                  group_name = .y,
+                                  bootstrap_size = bootstrap_size,
+                                  count = "count_sex", 
+                                  percentage = "percentage"))
 
@@
 quantile_bootstrapped_fitted_sex <- bootstrapped_fitted_sex %>%
-  group_by(time_p, period, siteid, sex) %>%
+  # group_by(time_p, period, plotsLabel, sex) %>%
+  group_by(time_p, period,  sex) %>%
   summarise("2.5%" = quantile(predicted, 0.025),
@@
 ```{r}
-coefficients_df_sex <- patients_hospitalized_agg_sex_perc_all %>% 
-  group_by(siteid, sex) %>% 
+coefficients_df_sex <- patients_hospitalized_agg_sex_total %>% 
+  group_by(sex) %>% 
+  # group_by(plotsLabel, sex) %>% 
     mutate(month_year = format(time_p, "%m"),
@@
 df_coefficients_boot_sex <- bootstrapped_coefficients_sex %>% 
-  left_join(coefficients_df_sex, by = c("siteid", "sex"))
+  # left_join(coefficients_df_sex, by = c("plotsLabel", "sex"))
+  left_join(coefficients_df_sex, by = c("sex"))
 ```
@@
 ```{r}
-ggplot(bootstrapped_fitted_all, aes(x = time_p, y = predicted, group = interaction(iteration, period))) +
-  facet_wrap("plotsLabel", scales = "free_y", labeller = label_wrap_gen(width = 35,multi_line = TRUE)) +
-  geom_line(aes(color = period),
+names_sex <- unique(df_coefficients_boot_sex$sex)
+
+## Using only male, but female and male necessarily yield equal p-values
+plot_annotations_sex_df <- annotation_function(
+  df_coefficients_boot_sex[df_coefficients_boot_sex$sex == "female", ]
+) %>% data.frame(label = ., 
+                 plotsLabel = "Female")
+
+# plot_annotations_sex_df <- df_coefficients_boot_sex %>% 
+#   group_by(sex) %>% 
+#   group_map(annotation_function) %>%
+#   paste(str_to_title(names_sex), ., sep = ": ") %>% 
+#   as_tibble() %>% 
+#   rename(label = value)
+# bind_cols(group_names) %>% 
+#                rename(label = value) %>% 
+#                mutate(label = paste(stringr::str_to_title(sex), label, sep = ": ")) %>% 
+#                mutate(plotsLabel = names_site_id)
+#              , .keep = TRUE) %>% 
+  # group_map( function(df, group_names) group_by(df, plotsLabel) %>%
+  #              annotation_function() %>%
+               # as_tibble() %>%
+               # bind_cols(group_names) %>%
+               # rename(label = value) %>%
+               # mutate(label = paste(stringr::str_to_title(sex), label, sep = ": ")) %>%
+  #              mutate(plotsLabel = names_site_id)
+  #            , .keep = TRUE) %>% 
+  # set_names(unique(df_coefficients_boot_sex$sex))
+
+```
+
+```{r bootstrap_plot}
+ggplot(bootstrapped_fitted_sex, aes(x = time_p, y = predicted, group = interaction(iteration, period, sex))) +
+  # facet_wrap("plotsLabel", scales = "free_y", labeller = label_wrap_gen(width = 35,multi_line = TRUE)) +
+  geom_line(aes(color = interaction(period, sex)),
             alpha = 0.03) +
-  geom_line(data = quantile_bootstrapped_fitted_all,
-            aes(y = value, group = interaction(quantiles, period)),
+  geom_line(data = quantile_bootstrapped_fitted_sex,
+            aes(y = value, group = interaction(quantiles, period, sex)),
             linetype = "dashed") +
-  geom_line(data = quantile_bootstrapped_fitted_all[quantile_bootstrapped_fitted_all$quantiles == "50%", ],
-            aes(y = value, group = period)) +
-  geom_point(data = patient_count_psy_period_psy_all,
-             aes(y = percentage, group=NULL),
-             alpha = 0.5) +
-  scale_colour_manual(values = cbPalette) +
+  geom_line(data = quantile_bootstrapped_fitted_sex[quantile_bootstrapped_fitted_sex$quantiles == "50%", ],
+            aes(y = value, group = interaction(period, sex))) +
+  geom_point(data = patients_hospitalized_agg_sex_total,
+             aes(y = percentage, group=sex, color = interaction(period, sex)),
+             alpha = 1) +
+  scale_color_brewer(palette = "Paired") +
+  # scale_colour_manual(values = c("#A6CEE3", "#1F78B4", "#B2DF8A", "#33A02C")) +
   guides(color = guide_legend(override.aes = list(alpha = 1))) +
-  geom_label(data = plot_annotations_df, 
+  geom_label(data = plot_annotations_sex_df,
              aes(label=label),
@@
             inherit.aes = FALSE) +
+  # geom_label(data = plot_annotations_sex_df[["male"]],
+  #            aes(label=label),
+  #            x = -Inf,
+  #            y = 15,
+  #            hjust = 0,
+  #            vjust = 0,
+  #            size = 4,
+  #           inherit.aes = FALSE) +
+  # ylim(10, 90) +
   labs(title = "Interrupted Time Series Analysis: Percentage of patients with psychiatric conditions",
@@
 ```
+### Disorder group fix
 
 ```{r}
-# expand.grid(
-#   unique(df_coefficients_boot_sex$siteid),
-#   unique(df_coefficients_boot_sex$sex)
-# ) %>% 
-#   t() %>% 
-#   paste(unique(.$siteid),
-#         unique(.$sex), 
-#         collapse = " "
-#   )
-# 
-# names_plot_annotations_sex <- c()
-# for (site in unique(df_coefficients_boot_sex$siteid)) {
-#   for (sex in unique(df_coefficients_boot_sex$sex)) {
-#     names_plot_annotations_sex <- c(names_plot_annotations_sex, paste(site, sex))
-#   }
-# }
 
-# plot_annotations_sex <- df_coefficients_boot_sex %>% 
-#   group_by(siteid, sex) %>% 
-#   group_map( annotation_function ) %>% 
-#   set_names(names_site_id) %>% 
-#   unlist()
-plot_annotations_df <- data.frame(
-  siteid = names(plot_annotations), 
-  # label = plot_annotations
-)
+perc_disorder_group_total_fix <- perc_disorder_group_all %>%
+  group_by( time_p, disorder_group ) %>%
+  summarise( total_diagnoses = sum(count_icd, na.rm = TRUE),
+             percentage = mean(percentage_dg * count_icd / total_diagnoses, na.rm = TRUE), 
+             country = "ALL")
+
 ```
@@
 ```{r}
-perc_disorder_group_total <- perc_disorder_group_all %>%
+perc_disorder_group_total_first <- perc_disorder_group_all %>%
   select( - percentage_dg ) %>%
@@
 
-perc_disorder_group_total %>%
+perc_disorder_group_total_fix %>%
   ggplot( aes( x= as.Date( time_p ), y = percentage, group = disorder_group )) +
@@
   mutate( disorder_group = swr( disorder_group ) )  %>%
-  inner_join( country_map ) %>%
   ggplot( aes( x= siteid, y = icd_codes_count, fill = country )) +
@@
 library(reshape2)
-count_icd_all <- count_icd_all %>%
-  left_join( country_map )
 

### 2022-04-04 metanalysis_visualization/manuscript_Figures.Rmd
fixing subgroup percentages count

@@
 ```{r}
+table_iterations_disorder <- with(perc_disorder_group_all,
+  expand.grid(time_p = unique(time_p),
+              disorder_group = unique(disorder_group), 
+              siteid = unique(siteid)
+  )
+)
 
-perc_disorder_group_total_fix <- perc_disorder_group_all %>%
+perc_disorder_group_complete <- perc_disorder_group_all %>% 
+  full_join(table_iterations_disorder) %>% 
+  arrange(siteid, time_p, disorder_group) %>% 
+  group_by(siteid, time_p) %>% 
+  fill(count_icd, 
+       period, 
+       country , 
+       plotsLabel, 
+       .direction = "downup") %>% 
+  ungroup() %>% 
+  replace_na(list(count_dg = 0, percentage_dg = 0))
+
+perc_disorder_group_total_fix <- perc_disorder_group_complete %>%
   group_by( time_p, disorder_group ) %>%
   summarise( total_diagnoses = sum(count_icd, na.rm = TRUE),
-             percentage = mean(percentage_dg * count_icd / total_diagnoses, na.rm = TRUE), 
+             sum_dg_psy = sum(count_dg),
+             percentage = (sum_dg_psy / total_diagnoses) * 100, 
              country = "ALL")

### 2022-04-04 metanalysis_visualization/manuscript_Figures.Rmd
perc_disorder_group_all changed to perc_disorder_group_total_first and country added to one of the plots

@@
 
-```{r Interrupted time series by sex}
+```{r Interrupted time series by sex 2}
 quantile_bootstrapped_fitted_sex <- bootstrapped_fitted_sex %>%
@@
   group_by( time_p, disorder_group ) %>%
-  mutate( total_cases = sum( count_icd ), 
+  mutate( total_cases = sum( count_icd, na.rm = TRUE ), 
           dg_cases = sum( count_dg ), 
@@
 
-perc_disorder_group_total_fix %>%
+perc_disorder_group_total_first %>%
   ggplot( aes( x= as.Date( time_p ), y = percentage, group = disorder_group )) +
@@
 
-percentages_htm <- perc_disorder_group_total %>%
+percentages_htm <- perc_disorder_group_total_first %>%
   select( time_p, disorder_group, percentage ) %>%
@@
 
-counts_htm <- perc_disorder_group_total %>%
+counts_htm <- perc_disorder_group_total_first %>%
   select( time_p, disorder_group, dg_cases ) %>%
@@
 ### most prevalent
-perc_disorder_group_total %>%
+perc_disorder_group_total_first %>%
   filter(disorder_group %in% c("Anxiety Disorders", "Depressive Disorders", "Suicide or Self-Injury" ))  %>%
@@
 #medium prevalent
-perc_disorder_group_total %>%
+perc_disorder_group_total_first %>%
   filter(disorder_group %in% c("Trauma and Stressor-Related Disorders", "Mental Health Symptom", "Disruptive, Impulse Control and Conduct Disorders", 
@@
 #low prevalent
-perc_disorder_group_total %>%
+perc_disorder_group_total_first %>%
   filter(disorder_group %in% c("Miscellaneous", "Substance-Related and Addictive Disorders", 
@@
 
-perc_disorder_group_total %>%
+perc_disorder_group_total_first %>%
   filter(disorder_group %in% c("Anxiety Disorders", "Depressive Disorders", "Suicide or Self-Injury" ))  %>%
@@
 
-perc_disorder_group_total %>%
+perc_disorder_group_total_first %>%
   filter( disorder_group %in% c("Anxiety Disorders", "Depressive Disorders", "Suicide or Self-Injury", 
@@
 numberOfCodesPerDisorderGroup %>%
+  left_join( country_map ) %>%
   mutate( disorder_group = swr( disorder_group ) )  %>%

### 2022-04-04 metanalysis_visualization/manuscript_Figures.Rmd
roll back to previous forest plot by siteid label. change axis of plots to have all the months

@@
 ```{r}
-outputDir <- "../../output/latest_output/"
+outputDir <- "../../output/previous_output/"
 outputFiles <- list.files( path = outputDir, pattern = "\\.RData")
@@
 df_coefficients_boot <- bootstrapped_coefficients_all %>% 
-  left_join(coefficients_df, by = "siteid") %>% 
-  left_join(country_map, by = "siteid")
-
-names_site_id <- unique(df_coefficients_boot$plotsLabel)
-
+  left_join(coefficients_df, by = "siteid")
+names_site_id <- unique(df_coefficients_boot$siteid)
 plot_annotations <- df_coefficients_boot %>% 
-  group_by(plotsLabel) %>% 
+  group_by(siteid) %>% 
   group_map( annotation_function ) %>% 
@@
 plot_annotations_df <- data.frame(
-  plotsLabel = names(plot_annotations), 
+  siteid = names(plot_annotations), 
   label = plot_annotations
@@
 ggplot(bootstrapped_fitted_all, aes(x = time_p, y = predicted, group = interaction(iteration, period))) +
-  facet_wrap("plotsLabel", scales = "free_y", labeller = label_wrap_gen(width = 35,multi_line = TRUE)) +
+  facet_wrap("siteid", scales = "free_y", labeller = label_wrap_gen(width = 35,multi_line = TRUE)) +
   geom_line(aes(color = period),
@@
   scale_colour_manual(values = cbPalette) +
+  scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0))+
   guides(color = guide_legend(override.aes = list(alpha = 1))) +
@@
         plot.subtitle = element_text(size = 12, face = "italic")) 
-ggsave("boostrapped_all.png", height = 20, width = 32, units = "cm")
+ggsave("boostrapped_all.png", height = 20, width = 45, units = "cm")
 ```
@@
 trends_siteid <- df_coefficients_boot %>% 
-  group_by(plotsLabel) %>% 
+  group_by(siteid) %>% 
   summarise(mean_boot = mean(`periodduring_pandemic:time`),
@@
 metafor::forest(model_meta_siteid,
-                slab = trends_siteid$plotsLabel,
+                slab = trends_siteid$siteid,
                 plim = c(1, NA),
@@
 trends_country <- trends_siteid %>% 
-  group_by(plotsLabel) %>% 
+  left_join(country_map, by = "siteid") %>% 
+  group_by(country) %>% 
   group_modify(.f = function(x, ...) {
@@
 forest(model_meta_country, 
-       slab = trends_country$plotsLabel,
+       slab = trends_country$country,
        plim = c(1, NA),
@@
   scale_color_brewer(palette = "Paired") +
+  scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0))+
   # scale_colour_manual(values = c("#A6CEE3", "#1F78B4", "#B2DF8A", "#33A02C")) +
@@
   theme(strip.text = element_text(size = 10),
-        axis.text.x = element_text(size = 8),
+        axis.text.x=element_text(angle = 45, hjust=1, size = 8),
         axis.text.y = element_text(size = 8),

### 2022-04-04 metanalysis_visualization/manuscript_Figures.Rmd
ITS and forest plot long lables fixed

@@
 ```{r}
+plot_annotations_df <-plot_annotations_df %>%
+  left_join( country_map )
+
 ggplot(bootstrapped_fitted_all, aes(x = time_p, y = predicted, group = interaction(iteration, period))) +
-  facet_wrap("siteid", scales = "free_y", labeller = label_wrap_gen(width = 35,multi_line = TRUE)) +
+  facet_wrap("plotsLabel", scales = "free_y", labeller = label_wrap_gen(width = 35,multi_line = TRUE)) +
   geom_line(aes(color = period),
@@
         plot.subtitle = element_text(size = 12, face = "italic")) 
-ggsave("boostrapped_all.png", height = 20, width = 45, units = "cm")
+ggsave("boostrapped_all_long.png", height = 20, width = 45, units = "cm")
 ```
@@
 # Random effect
+trends_siteid <- trends_siteid %>%
+  left_join( country_map )
 model_meta_siteid <- metafor::rma(yi = observed, 
@@
 metafor::forest(model_meta_siteid,
-                slab = trends_siteid$siteid,
+                slab = trends_siteid$plotsLabel,
                 plim = c(1, NA),
@@
 trends_country <- trends_siteid %>% 
-  left_join(country_map, by = "siteid") %>% 
   group_by(country) %>% 

### 2022-04-05 metanalysis_visualization/manuscript_Figures.Rmd
Subcases count fix

@@
 
-perc_disorder_group_total_fix <- perc_disorder_group_complete %>%
+perc_disorder_group_total <- perc_disorder_group_complete %>%
   group_by( time_p, disorder_group ) %>%
   summarise( total_diagnoses = sum(count_icd, na.rm = TRUE),
-             sum_dg_psy = sum(count_dg),
-             percentage = (sum_dg_psy / total_diagnoses) * 100, 
+             dg_cases = sum(count_dg),
+             percentage = (dg_cases / total_diagnoses) * 100, 
              country = "ALL")
@@
 ```{r}
-perc_disorder_group_total_first <- perc_disorder_group_all %>%
-  select( - percentage_dg ) %>%
-  group_by( time_p, disorder_group ) %>%
-  mutate( total_cases = sum( count_icd, na.rm = TRUE ), 
-          dg_cases = sum( count_dg ), 
-          percentage = 100* dg_cases / total_cases, 
-          country = "ALL")
-
-perc_disorder_group_total_first %>%
+perc_disorder_group_total %>%
   ggplot( aes( x= as.Date( time_p ), y = percentage, group = disorder_group )) +
@@
 
-percentages_htm <- perc_disorder_group_total_first %>%
+percentages_htm <- perc_disorder_group_total %>%
   select( time_p, disorder_group, percentage ) %>%
@@
 
-counts_htm <- perc_disorder_group_total_first %>%
+counts_htm <- perc_disorder_group_total %>%
   select( time_p, disorder_group, dg_cases ) %>%
@@
 counts_htm <- counts_htm[  , -1]
-
 tempVariabilityInput <- new('DataTemporalMap', probabilityMap = as.matrix(percentages_htm), 
@@
 ### most prevalent
-perc_disorder_group_total_first %>%
+perc_disorder_group_total %>%
   filter(disorder_group %in% c("Anxiety Disorders", "Depressive Disorders", "Suicide or Self-Injury" ))  %>%
@@
 #medium prevalent
-perc_disorder_group_total_first %>%
+perc_disorder_group_total %>%
   filter(disorder_group %in% c("Trauma and Stressor-Related Disorders", "Mental Health Symptom", "Disruptive, Impulse Control and Conduct Disorders", 
@@
 #low prevalent
-perc_disorder_group_total_first %>%
+perc_disorder_group_total %>%
   filter(disorder_group %in% c("Miscellaneous", "Substance-Related and Addictive Disorders", 
@@
 
-perc_disorder_group_total_first %>%
+perc_disorder_group_total %>%
   filter(disorder_group %in% c("Anxiety Disorders", "Depressive Disorders", "Suicide or Self-Injury" ))  %>%
@@
 
-perc_disorder_group_total_first %>%
+perc_disorder_group_total %>%
   filter( disorder_group %in% c("Anxiety Disorders", "Depressive Disorders", "Suicide or Self-Injury", 
@@
 Represent this in 16 pannels, if we put all together, due to the different scales we will not be able to really see the differences between results.
+
 ```{r}

### 2022-04-05 metanalysis_visualization/manuscript_Figures.Rmd
subgroup analysis

@@
 
+#### Bootstrapping results by sex
+
 ```{r Interrupted time series by sex}
-
 patients_hospitalized_agg_sex_total <- patients_hospitalized_agg_sex_total %>% 
@@
 #   inner_join(country_map)
+```
 
+```{r Interrupted time series by sex}
 bootstrapped_sex <- patients_hospitalized_agg_sex_total %>%
@@
 bootstrapped_fitted_sex <- lapply(bootstrapped_sex, `[[`, 2) %>% bind_rows()
-```
-
-```{r Interrupted time series by sex 2}
 quantile_bootstrapped_fitted_sex <- bootstrapped_fitted_sex %>%
@@
                names_to = "quantiles")
-```
-
-```{r}
 coefficients_df_sex <- patients_hospitalized_agg_sex_total %>% 
@@
 
-```
-
-```{r}
 df_coefficients_boot_sex <- bootstrapped_coefficients_sex %>% 
@@
 
-```{r}
+```{r annotation bootstrap disorder sex}
 names_sex <- unique(df_coefficients_boot_sex$sex)
@@
 ```
-### Disorder group fix
+
+### Disorder group
 
@@
              percentage = (dg_cases / total_diagnoses) * 100, 
-             country = "ALL")
+             country = "ALL", 
+             siteid = "ALL") %>% 
+  mutate(period = ifelse(time_p < pandemic_start_date,
+                         "before_pandemic", 
+                         "during_pandemic"))
+
+```
+
+#### Bootstrapping results by disorder groups
+
+```{r}
+disorders_to_select <- c(
+  "Anxiety Disorders", 
+  "Depressive Disorders", 
+  "Feeding and Eating Disorders", 
+  "Suicide or Self-Injury"
+)
+
+bootstrapped_disorders <- perc_disorder_group_total %>%
+  filter(disorder_group %in% disorders_to_select) %>% 
+  group_by(disorder_group) %>% 
+  mutate(month_year = format(time_p, "%m"),
+         time = seq_along(time_p))  %>% 
+ group_map(~ bootstrapping_slopes(.x,
+                                  group_name = .y,
+                                  bootstrap_size = bootstrap_size,
+                                  count = "dg_cases", 
+                                  percentage = "percentage"))
+
+bootstrapped_coefficients_disorder <- lapply(bootstrapped_disorders, `[[`, 1) %>% bind_rows()
+bootstrapped_fitted_disorder <- lapply(bootstrapped_disorders, `[[`, 2) %>% bind_rows()
+quantile_bootstrapped_fitted_disorder <- bootstrapped_fitted_disorder %>%
+  # group_by(time_p, period, plotsLabel, sex) %>%
+  group_by(time_p, period,  disorder_group) %>%
+  summarise("2.5%" = quantile(predicted, 0.025),
+            "50%" = median(predicted, 0.5),
+            "97.5%" = quantile(predicted, 0.975)) %>%
+  pivot_longer(cols = c("2.5%", "50%", "97.5%"),
+               names_to = "quantiles")
+coefficients_df_disorder <- perc_disorder_group_total %>% 
+  group_by(disorder_group) %>% 
+  mutate(month_year = format(time_p, "%m"),
+         time = seq_along(time_p))  %>% 
+  group_modify(coefficient_slope_calculation) %>% 
+  rename(observed_diff = estimate)
+
+df_coefficients_boot_disorder <- bootstrapped_coefficients_disorder %>% 
+  left_join(coefficients_df_disorder, by = c("disorder_group"))
+```
+
+```{r annotation boostrap disorders}
+names_disorder <- unique(df_coefficients_boot_disorder$disorder_group)
+
+plot_annotations_disorder <- df_coefficients_boot_disorder %>% 
+  group_by(disorder_group) %>% 
+  group_map( annotation_function ) %>% 
+  set_names(names_disorder) %>% 
+  unlist()
+plot_annotations_disorder_df <- data.frame(
+  disorder_group = names(plot_annotations_disorder), 
+  label = plot_annotations_disorder
+)
+```
+
+```{r}
+# plot_annotations_df <- plot_annotations_df %>%
+#   left_join( country_map )
+
+ggplot(bootstrapped_fitted_disorder,
+       aes(x = time_p, y = predicted, group = interaction(iteration, period))) +
+  facet_wrap("disorder_group",
+             scales = "free_y",
+             labeller = label_wrap_gen(width = 35,multi_line = TRUE)) +
+  geom_line(aes(color = period),
+            alpha = 0.03) +
+  geom_line(data = quantile_bootstrapped_fitted_disorder,
+            aes(y = value, group = interaction(quantiles, period)),
+            linetype = "dashed") +
+  geom_line(data = quantile_bootstrapped_fitted_disorder[quantile_bootstrapped_fitted_disorder$quantiles == "50%", ],
+            aes(y = value, group = period)) +
+  geom_point(data = filter(perc_disorder_group_total,
+                           disorder_group %in% disorders_to_select),
+             aes(y = percentage, group=NULL),
+             alpha = 0.5) +
+  scale_colour_manual(values = cbPalette) +
+  scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0))+
+  guides(color = guide_legend(override.aes = list(alpha = 1))) +
+  geom_label(data = plot_annotations_disorder_df, 
+             aes(label=label), 
+            x = -Inf,
+            y = +Inf,
+            hjust = 0,
+            vjust = 1,
+            size = 4,
+            inherit.aes = FALSE) +
+  labs(title = "Interrupted Time Series Analysis: Percentage of patients with psychiatric conditions",
+       subtitle = "Parametric bootstrap estimation: (1000 iterations)",
+       x = "Calendar dates",
+       y = "Percentage",
+       color = "Period") +
+  theme(strip.text = element_text(size = 10),
+        axis.text.x=element_text(angle = 45, hjust=1, size = 8),
+        axis.text.y = element_text(size = 8), 
+        axis.title.x = element_text(size = 8),
+        axis.title.y = element_text(size = 8),
+        title = element_text(size = 14),
+        plot.subtitle = element_text(size = 12, face = "italic")) 
+# ggsave("boostrapped_disorders.png", height = 20, width = 45, units = "cm")
 

### 2022-04-05 psy_peds_3_umich.Rmd
added the rmd run by umich

--- 2022-03-31 psy_peds_3.Rmd
@@
 ```{r message=FALSE, warning=FALSE}
-folder_4ce_files <- "Input/09212021_2/"
+folder_4ce_files <- "/Users/hanauer/Desktop/DONT_INDEX/4ceData/Input/phase22_11NOV2021/"
 obfuscation =  FALSE
@@
 min_hosp_days <- 2
-data_update_date <- "2021-06-05"
+data_update_date <- "2021-11-04"
 time_period <-"months" #possible values: days, weeks, months
@@
 
-getwd()
 files <- readInputFiles( path      = folder_4ce_files, 
@@
 ```{r}
-icdCodes <- read.csv("public-data/pediatric_psychiatric_ICD10_codesV3.csv", header = TRUE, colClasses = "character") %>%
+icdCodes <- read.csv("public-data/pediatric_psychiatric_ICD10_codesV2.csv", header = TRUE, colClasses = "character") %>%
    filter( flag %in% c("originalCodes", "FrenchSpecificCodes")) %>%
@@
 obs_raw_age <- left_join(obs_raw, demo_raw[c("patient_num", "age")], by = c("patient_num")) %>% 
-  inner_join(clinical_raw[, c("patient_num", "days_since_admission", "calendar_date")], 
+  left_join(clinical_raw[, c("patient_num", "days_since_admission", "calendar_date")], 
             by = c("patient_num", "days_since_admission"))
+stopifnot(nrow(obs_raw_age) == nrow(obs_raw))
+obs_raw_age <- obs_raw_age[ !is.na( obs_raw_age$calendar_date), ]
 obs_raw_age$age_time_diagnosis <- obs_raw_age$age - floor(as.numeric(as.Date(data_update_date) - as.Date(obs_raw_age$calendar_date, dateFormat)) / 365.25)
@@
 
-bootstrapped_coefficients_df <- bind_rows(bootstrapped_coefficients_list,
-                                          .id = "iteration")
-
-z <-  mean(bootstrapped_coefficients_df[["periodduring_pandemic:time"]]) / sd(bootstrapped_coefficients_df[["periodduring_pandemic:time"]])
-
-pvalue_coeff_interaction <- round(pnorm(abs(z), lower.tail = FALSE) * 2, 3)
-
-plot_annotation <- paste0("Coeffs differences: ",
+bootstrapped_coefficients_df <- bind_rows(bootstrapped_coefficients_list, .id = "iteration")
+mean(bootstrapped_coefficients_df$time)
+mean(bootstrapped_coefficients_df$time + bootstrapped_coefficients_df[["periodduring_pandemic:time"]])
+pvalue_coeff_interaction <- round(t.test(bootstrapped_coefficients_df[["periodduring_pandemic:time"]])$p.value, 4)
+pvalue_coeff_interaction <- ifelse(pvalue_coeff_interaction == 0,
+                                   "< 1e-4",
+                                   str(pvalue_coeff_interaction))
+plot_annotation <- paste0("Diff. in slope coeffs before and after pandemic: ",
                      round(mean(bootstrapped_coefficients_df[["periodduring_pandemic:time"]]), 2),
@@
 # counts per time period and sex
+all_codes_psy$sex <- trimws( all_codes_psy$sex)
+all_codes_psy$race_4ce <- trimws( all_codes_psy$race_4ce)
 patients_hospitalized_agg_sex <- all_codes_psy %>%

### 2022-04-06 metanalysis_visualization/manuscript_Figures.Rmd
remove duplicated label from markdown cell

@@
 
-```{r Interrupted time series by sex}
+```{r Interrupted time series by sex bootstrap}
 bootstrapped_sex <- patients_hospitalized_agg_sex_total %>%

### 2022-04-06 metanalysis_visualization/manuscript_Figures.Rmd
updating results

--- 2022-04-05 metanalysis_visualization/manuscript_Figures.Rmd
@@
 ```{r}
-outputDir <- "../../output/previous_output/"
+outputDir <- "../../output/latest_output//"
 outputFiles <- list.files( path = outputDir, pattern = "\\.RData")
@@
 
-```{r Interrupted time series by sex}
+```{r }
 bootstrapped_sex <- patients_hospitalized_agg_sex_total %>%
@@
         plot.subtitle = element_text(size = 12, face = "italic")) 
-# ggsave("boostrapped_disorders.png", height = 20, width = 45, units = "cm")
+ggsave("boostrapped_disorders.png", height = 14, width = 25, units = "cm")
 

### 2022-04-06 psy_peds_3_LURIE.Rmd

--- 2022-03-31 psy_peds_3.Rmd
@@
   html_document:
-    code_folding: hide
     toc: yes
@@
 
-listOfPackages <- c("tidyverse", "RColorBrewer", "knitr", "kableExtra", "tsModel", "gridExtra", "dplyr", "lubridate")
+#listOfPackages <- c("tidyverse", "RColorBrewer", "knitr", "kableExtra", "tsModel", "gridExtra", "dplyr")
+#paket(listOfPackages)
+
+# trying to remove kableExtra to run w/o docker
+listOfPackages <- c("tidyverse", "RColorBrewer", "knitr", "tsModel", "gridExtra", "dplyr")
 paket(listOfPackages)
@@
 sessionInfo()
+```
+
+### Set working directory
+
+```{r}
+#setwd("~/projects/nu4ce/4ceData_files/Phase2.2.psy_peds_trends")
 ```
@@
 ```{r message=FALSE, warning=FALSE}
-folder_4ce_files <- "Input/09212021_2/"
+folder_4ce_files <- "../4ce_ped_psych/data"
 obfuscation =  FALSE
@@
 #raceAvailable = FALSE
-dateFormat <- "%d-%b-%y"
-min_hosp_days <- 2
+dateFormat <- "%d-%m-%y" #"%d-%b-%y"
+min_hosp_days <- 0
 data_update_date <- "2021-06-05"
@@
 # cut-off dates used in the analysis, these should not be changed
-# history_start_date <- as.Date("2019-01-01")
-# history_end_date <- as.Date("2021-05-31")
+#start_date_plots <- as.Date("2019-01-01")
+#history_end_date <- as.Date("2021-05-31")
 pandemic_start_date <- as.Date("2020-03-15")
@@
 
-```{r message=FALSE, warning=FALSE}
+```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
 ### Read the CSV input files
@@
 
+### Read in Lurie files
+
+```{r}
+non_psych_inpt_counts <- read.csv(paste0(folder_4ce_files, "/non_psych_admit_inpt_11_17.csv"))
+admit_all_codes <- read.csv(paste0(folder_4ce_files, "/psych_admit_all_codes.csv"))
+admit_demog <- read.csv(paste0(folder_4ce_files, "/psych_admit_demog.csv"))
+admit_inpt_primary <- read.csv(paste0(folder_4ce_files, "/psych_admit_inpt_primary.csv"))
+```
+
+Raw Data Counts
+
+```{r}
+nrow(unique(data.frame(admit_all_codes$patient_id)))
+nrow(unique(data.frame(admit_demog$patient_id)))
+```
 ### Read the ICD10 codes file
@@
 
+Old codes that Nelson used. These were updated to the ICD code list above in order to remove codes for accidental poisoning
+
+```{r}
+icdCodes_old <- read.csv("public-data/psychiatric_ICD-10_codes_adjusted.csv", header = TRUE, colClasses = "character")
+
+colnames(icdCodes_old) <- c("disorder_group", "ICD10_Code", "description")
+
+icdCodes_old <- icdCodes_old %>% 
+  mutate(ICD10_Code = gsub("ICD10:", "", ICD10_Code))
+
+excluded_codes <- icdCodes_old %>% anti_join(icdCodes)
+qc_check <- admit_all_codes %>% filter(icd_10_code %in% excluded_codes$ICD10_Code) 
+nrow(unique(data.frame(qc_check$patient_id)))
+```
+
 ## Data-management
 
+**Meghan Note: (Skip-to Lurie Data Formatting section)**
+
 ### Filtering on "inpatient" status
 
-```{r}
+```{r eval=FALSE, include=FALSE}
 # Filtering on inpatients here (in_hospital == 1), to reduce subsequent compute time of the encounter length calculation 
@@
 
-```{r}
+```{r eval=FALSE, include=FALSE}
 count_sequences_hospitalisation <- function(df, ...) {
@@
 
+# Lurie Data Formatting
+
+**Clean up Lurie encounters**
+
+As noted by Nelson: "some patients will have two encounters back-to-back.
+These should be considered the SAME encounter.
+The reason for having two encounters recorded is because medical admissions and psych admissions are considered separate admissions in our hospital.
+A simple way to deal with this redundancy is to check encounters for the same patient that start on the same day as the end of the prior encounter and merge those two into one encounter."
+
+**First we will combine all of the psych + non-psych data pulls**
+
+```{r}
+# admit_demog$pysch = 1
+# non_psych_inpt_counts$pysch = 0
+
+admit_demog <- rbind(admit_demog, non_psych_inpt_counts)
+
+## appears to be no overlap?
+#View(admit_demog %>% distinct(patient_id, encounter_id, pysch) %>% group_by(psych) %>%  mutate(count = n()))
+```
+
+```{r}
+## format dates
+admit_demog$i_hsp_adm_dttm <- strptime(admit_demog$hsp_adm_dttm, format="%m/%d/%Y %H:%M") %>% as.POSIXct()
+admit_demog$i_hsp_dsch_dttm <- strptime(admit_demog$hsp_dsch_dttm, format="%m/%d/%Y %H:%M") %>% as.POSIXct()
+
+# # filter ages
+# admit_demog <- admit_demog %>% filter(age_years > 10 & age_years < 18)  %>%
+#  filter(i_hsp_adm_dttm >= start_date_plots & i_hsp_adm_dttm <= end_date_plots)
+
+# unique encounter numbers should equal the row number (TRUE)
+nrow(unique(data.frame(admit_demog$encounter_id))) == nrow(admit_demog)
+nrow(admit_demog %>% distinct(patient_id, encounter_id))
+
+encounters <- admit_demog %>% 
+  mutate(gender = if_else(gender == "Male", "male", "female")) %>% 
+  group_by(patient_id) %>% 
+  arrange(patient_id, i_hsp_adm_dttm) %>% 
+  filter(i_hsp_adm_dttm >= start_date_plots & i_hsp_adm_dttm <= end_date_plots) %>% 
+  mutate(encounter_group = seq_along(encounter_id),
+         day_lag = as.numeric(round(difftime(i_hsp_adm_dttm, lag(i_hsp_dsch_dttm), unit = "hours"), 2))) %>% 
+  data.frame() %>%
+  # adding 100 to temporarily handle the NAs
+  mutate(day_lag = ifelse(is.na(day_lag), 100, day_lag),
+  # here we want to ensure that if subsequent encounters occurred within a day (24 hours), that we impute the initial encounter_group as per Nelson's suggestion
+         i_encounter_group = ifelse(day_lag <= 24, lag(encounter_group), encounter_group),
+  consolidated = ifelse(day_lag <= 24, 1, 0))
+
+# this is suggesting that 408 need to get consolidated?
+table(encounters$consolidated)
+
+check_lag <- encounters %>% filter(!day_lag == 100)
+summary(check_lag$day_lag)
+less24 <- check_lag %>% filter(day_lag <= 24) #449
+summary(less24$day_lag)
+hist(less24$day_lag)
+
+encounters <- encounters %>%
+  # in some cases there are more than two back-to-back encounters
+  # to filter for these, first we will group by the patient_id and the new encounter group
+  group_by(patient_id, i_encounter_group) %>%
+  # next, we will calculate the latest admission date of this encounter
+  mutate(max_date = max(i_hsp_adm_dttm)) %>%
+  ungroup() %>%
+  # next, we group just by Patient, and we identify the latest max_date
+  # we then calculate the current time from the lag(max_date) and we will identify those within 24 hours as another encounter that should be grouped
+  group_by(patient_id) %>%
+  mutate(lag_date = lag(max_date),
+         time_diff = as.numeric(round(difftime(i_hsp_adm_dttm, lag_date, unit = "hours"), 2))) %>%
+  # adding 100 to temporarily handle the NAs
+  mutate(time_diff = ifelse(is.na(time_diff), 100, time_diff),
+    i_encounter_group = ifelse(time_diff <= 24, lag(i_encounter_group), i_encounter_group),
+    consolidated2 = ifelse(time_diff <= 24, 1, 0),
+  # recalculate the encounter sequence
+  encounter_num = dense_rank(i_encounter_group)) %>% 
+  group_by(patient_id, encounter_num)
+
+# this is suggesting that 408 need to get consolidated?
+table(encounters$consolidated2)
+
+check_lag <- encounters %>% filter(!time_diff == 100)
+summary(check_lag$time_diff)
+less24 <- check_lag %>% filter(time_diff <= 24) #397 patients/admissions said to need to be consolidated
+summary(less24$time_diff)
+hist(less24$time_diff)
+
+encounters <- encounters %>% 
+  slice(1L)
+
+nrow(encounters %>% distinct(patient_id, encounter_num))
+```
+
+Check to see if we should have combined any additional encounters
+
+```{r}
+test = encounters %>% 
+       group_by(patient_id, encounter_num) %>% 
+       mutate(max_date = max(i_hsp_adm_dttm)) %>% 
+  ungroup() %>%
+  group_by(patient_id) %>% 
+  mutate(lag_date = lag(max_date),
+         time_diff = as.numeric(difftime(i_hsp_adm_dttm, lag_date, unit = "hours")),
+         check = if_else(time_diff <= 24 & encounter_num != lag(encounter_num), 1, 0))
+
+#View(test %>% filter(check == 1))
+```
+
+**Add number of hospitalizations and LOS of stay for each patient**
+
+```{r}
+clinical_raw <- encounters %>% 
+  group_by(patient_id) %>% 
+  mutate(n_hospitalisation = n_distinct(as.character(encounter_num))) %>% 
+  group_by(patient_id, encounter_num) %>% 
+  # calculate the hospitalization length, while taking care to select the max(i_hsp_dsch_dttm) in the case that we have more than 1 row per the same encounter
+  mutate(len_hospitalisation = round(as.numeric(difftime(max(i_hsp_dsch_dttm), min(i_hsp_adm_dttm), unit = "days"))),
+         # replace last discharge with the true discharge datetime
+         i_hsp_dsch_dttm = max(i_hsp_dsch_dttm),
+         # replace age_years with earliest age on admission 
+         age_years = min(age_years)) %>% 
+  rename("date" = i_hsp_adm_dttm,
+         "patient_num" = patient_id) %>% 
+  group_by(patient_num, encounter_num) %>% 
+  slice(1L) %>% 
+  ungroup() %>% 
+  select(patient_num, date, i_hsp_dsch_dttm, race, ethnicity, gender, age_years, encounter_num, n_hospitalisation, len_hospitalisation) %>% 
+  distinct()
+
+nrow(clinical_raw %>% distinct(patient_num, encounter_num))
+```
+
 ### Counting ICD10 mental health codes in the obs_raw table
 
-```{r}
+```{r echo=TRUE}
 ### check that the ICD codes follow the regular expression [A-Z][0-9][0-9AB]\.?[0-9A-TV-Z]{0,4}
-codesToReview <- obs_raw %>% dplyr::filter(concept_type == "DIAG-ICD10",
-                                           ! grepl( "[A-Z][0-9][0-9AB]\\.?[0-9A-TV-Z]{0,4}", concept_code))
+codesToReview <- admit_all_codes %>% dplyr::filter(icd_10_code == "DIAG-ICD10",
+                                           ! grepl( "[A-Z][0-9][0-9AB]\\.?[0-9A-TV-Z]{0,4}", icd_10_code))
 print(codesToReview)
@@
 ### Calculate how many of the mental codes are in the dataset
-mentalHealthCodes <- unique( obs_raw %>%
-  dplyr::filter( concept_type == "DIAG-ICD10", 
-                 concept_code %in% icdCodes$ICD10_Code ) %>%
-  dplyr::mutate( concept_code = as.character( concept_code ), 
-                 concept_length = nchar( concept_code ), 
-                 concept_check = ifelse( concept_length == 4 | concept_length > 8, "Check", "OK")) %>%
-  dplyr::select( concept_code, concept_length, concept_check) )
+mentalHealthCodes <- unique( admit_all_codes %>%
+  dplyr::filter( #concept_type == "DIAG-ICD10", 
+                 icd_10_code %in% icdCodes$ICD10_Code ) %>%
+  dplyr::mutate( icd_10_code = as.character( icd_10_code ), 
+                 concept_length = nchar( icd_10_code ), 
+                 concept_check = ifelse( icd_10_code == 4 | concept_length > 8, "Check", "OK")) %>%
+  dplyr::select( icd_10_code, concept_length, concept_check) )
 
@@
 
-print( paste0("There are ", length(mentalHealthCodes$concept_code), " ICD10 mental codes out of the total ",length(unique(icdCodes$ICD10_Code)), " in the dataset"))
+print( paste0("There are ", length(mentalHealthCodes$icd_10_code), " ICD10 mental codes out of the total ",length(unique(icdCodes$ICD10_Code)), " in the dataset"))
 ```
@@
 ### Calculating age at hospitalization time
-```{r}
+
+```{r eval=FALSE, include=FALSE}
 # Calculating patient age at the time of visit, assuming the age used is the patients' age at the time of extraction (last discharge date available in the data)
 obs_raw_age <- left_join(obs_raw, demo_raw[c("patient_num", "age")], by = c("patient_num")) %>% 
-  inner_join(clinical_raw[, c("patient_num", "days_since_admission", "calendar_date")], 
+  left_join(clinical_raw[, c("patient_num", "days_since_admission", "calendar_date")], 
             by = c("patient_num", "days_since_admission"))
+stopifnot(nrow(obs_raw_age) == nrow(obs_raw))
 obs_raw_age$age_time_diagnosis <- obs_raw_age$age - floor(as.numeric(as.Date(data_update_date) - as.Date(obs_raw_age$calendar_date, dateFormat)) / 365.25)
@@
 
-``` {r}
+```{undefined eval=FALSE, include=FALSE}
 obs_filtered_age <- obs_raw_age %>% 
@@
 
-```{r}
+### Further filtering Lurie data
+
+```{r}
+clinical_raw$date <- strptime(clinical_raw$date, format="%Y-%m-%d") %>% as.POSIXct()
+
+# create list of calendar_dates
+date_ids <- seq(as.Date("2019-01-01"), as.Date(end_date_plots), 
+                by= "day") %>% 
+  data.frame() %>% 
+  rename('calendar_date' = '.')
+
+date_ids <- date_ids %>% 
+  dplyr::mutate( date = as.Date( calendar_date, format = dateFormat ),
+                 weeks = as.Date(cut( date, breaks = "week")),
+                 month = as.Date(cut( date, breaks = "month")),
+                 year = format( date, "%Y"), 
+                 period = ifelse( date < pandemic_start_date,
+                                  "before_pandemic","during_pandemic"))
+
+clinical_filtered <- clinical_raw %>% 
+  filter(len_hospitalisation >= min_hosp_days) %>% 
+  left_join(., date_ids, by = "date") %>% 
+   dplyr::mutate( date = as.Date( calendar_date, format = dateFormat ),
+                  weeks = as.Date(cut( date, breaks = "week")),
+                  month = as.Date(cut( date, breaks = "month")),
+                  year = format( date, "%Y"), 
+                  period = ifelse( date < pandemic_start_date,
+                                   "before_pandemic","during_pandemic"))
+
+# check breaks example from BCH
+example_bch <- read.delim2("example.txt", header = TRUE, sep = "")
+example_lurie <- date_ids %>% distinct(calendar_date, weeks, month, year, period)
+
+example_bch$calendar_date <- as.Date(example_bch$calendar_date)
+example_bch$weeks <- as.Date(example_bch$weeks)
+example_bch$month <- as.Date(example_bch$month)
+example_bch$year <- as.character(example_bch$year)
+example_bch <- example_bch %>% filter(!calendar_date >= "2021-06-01")
+
+diff <- example_bch %>% anti_join(example_lurie)
+nrow(diff)
+diff <- example_lurie %>% anti_join(example_bch)
+nrow(diff)
+identical(example_bch, example_lurie)
+all.equal(example_bch, example_lurie)
+
+if( time_period == "weeks"){
+  clinical_filtered$time_p <- clinical_filtered$weeks
+}else if(time_period == "days"){
+  clinical_filtered$time_p <- clinical_filtered$date
+}else if(time_period == "months"){
+  clinical_filtered$time_p <- clinical_filtered$month
+}
+```
+
+```{r eval=FALSE, include=FALSE}
 # Steps necessary to filter rows:
@@
 
-```{r}
+**Create Lurie's Input_data table**
+
+```{r}
+nrow(unique(data.frame(clinical_filtered$patient_num)))
+nrow(clinical_filtered %>% distinct(patient_num, encounter_num))
+
+input_data <- clinical_filtered %>% 
+  mutate(age_years = floor(age_years)) %>% 
+  filter(date >= start_date_plots & date <= end_date_plots,
+         age_years > 10 & age_years < 18) %>% 
+  group_by(patient_num, encounter_num) %>% 
+  arrange(date) %>% 
+  slice(1L) %>% 
+  ungroup() %>% 
+  select(patient_num, encounter_num, race, ethnicity, gender, age_years, date, i_hsp_dsch_dttm, weeks, month, year, period, time_p, n_hospitalisation, len_hospitalisation) %>% 
+  rename("sex" = gender,
+         "age" = age_years)
+
+nrow(unique(data.frame(input_data$patient_num)))
+nrow(input_data %>% distinct(patient_num, encounter_num))
+```
+
+```{r eval=FALSE, include=FALSE}
+qc_check <- admit_all_codes %>% 
+   rename("patient_num" = patient_id, 
+          "concept_code" = diagnosis_code) %>% 
+  left_join(., icdCodes, by = c("concept_code" = "ICD10_Code")) %>% 
+  mutate(icd_code_category = ifelse(concept_code %in% icdCodes$ICD10_Code, 
+                               "psy", 
+                               "others")) %>% 
+  left_join(input_data, ., by = "patient_num") %>% 
+  mutate(i_diagnosis_dttm = strptime(diagnosis_dttm, format="%m/%d/%Y") %>% as.POSIXct(),
+         date = format(as.POSIXct(date, format = "%m/%d/%Y %H:%M"), format = "%Y-%m-%d"),
+         i_hsp_dsch_dttm = format(as.POSIXct(i_hsp_dsch_dttm, format = "%m/%d/%Y %H:%M"), format = "%Y-%m-%d")) %>% 
+  dplyr::filter( diagnosis_code_source == "ICD-10-CM", 
+                 #in_hospital == 1,
+                 i_diagnosis_dttm >= date & i_diagnosis_dttm <= i_hsp_dsch_dttm,
+                 age > 10 & age < 18,
+                 date >= start_date_plots & date <= end_date_plots,
+                 ) %>%
+  dplyr::select( patient_num, encounter_num, sex, age, race, concept_code, n_hospitalisation, len_hospitalisation, 
+                 date, weeks, month, year, period, time_p, disorder_group, description, icd_code_category)
+
+```
+
+```{r eval=FALSE, include=FALSE}
 stopifnot(unique(all_codes$cohort) == "AllAdm")
@@
 
-```{r}
+```{r eval=FALSE, include=FALSE}
 # Filtering hospitalizations where patient age is in inclusion criteria
@@
 
-```{r}
+```{r eval=FALSE, include=FALSE}
 stopifnot(unique(all_codes_ado$cohort) == "AllAdm")
@@
 
+Are there any patients in the admit_demog df who do not have any icd codes listed?
+
+Note: this looks fine
+```{r}
+admit_demog <- read.csv(paste0(folder_4ce_files, "/psych_admit_demog.csv"))
+
+# list of all patients who should have had at least 1 psych diagnosis
+psych_ids <- admit_demog %>% 
+  distinct(patient_id)
+nrow(psych_ids)
+
+# evaluate whether there are patients in admit_all_codes file that were not listed in the admit_demog file
+admit_demog_ids <- admit_all_codes %>% 
+  distinct(patient_id) %>% 
+  filter(!patient_id %in% psych_ids$patient_id)
+nrow(admit_demog_ids)
+
+# evaluate whether there are patients in the admit_demog file that do not have diagnosis codes
+psych_id_check <- psych_ids %>% 
+  distinct(patient_id) %>% 
+  filter(!patient_id %in% admit_all_codes$patient_id)
+nrow(psych_id_check)
+```
+
+### Add the psychiatric/mental health ICD code description and filter
+
+```{r message=FALSE, warning=FALSE}
+# Getting site name, will be used later in the analysis
+#site <- unique(as.character(demo_raw$siteid))
+site <- "Lurie"
+
+icdCategory <- read.delim("public-data/icd10Codes.txt",
+                          header = FALSE,
+                          colClasses = "character")
+colnames(icdCategory) <- c("icd_code_letter", "icd_description_category")
+
+# all_codes_desc <- all_codes_ado %>% 
+#     left_join( icdCodes, by=c("concept_code" = "ICD10_Code") ) %>%
+#   mutate(icd_code_category = ifelse(concept_code %in% icdCodes$ICD10_Code, 
+#                                "psy", 
+#                                "others")) %>%
+#   dplyr::select( patient_num, sex, age_time_diagnosis, concept_type, concept_code, calendar_date,
+#                  in_hospital, n_hospitalisation, len_hospitalisation, date, weeks, month, year,
+#                  period, time_p, disorder_group, description, icd_code_category )
+
+# begin processing the ICD codes for the patients
+all_codes_desc <- admit_all_codes %>% 
+   rename("patient_num" = patient_id, 
+          "concept_code" = icd_10_code) %>% 
+  left_join(., icdCodes, by = c("concept_code" = "ICD10_Code")) %>% 
+  mutate(icd_code_category = ifelse(concept_code %in% icdCodes$ICD10_Code, 
+                               "psy", 
+                               "others"))
+
+nrow(input_data)
+nrow(unique(data.frame(input_data$patient_num)))
+nrow(input_data %>% distinct(patient_num, encounter_num))
+
+# Join mental health codes to the patient_number
+input_data <- input_data %>% 
+  left_join(., all_codes_desc, by = "patient_num") %>% 
+  mutate(i_diagnosis_dttm = strptime(diagnosis_dttm, format="%m/%d/%Y") %>% as.POSIXct(),
+         date = strptime(date, format="%Y-%m-%d") %>% as.POSIXct(),
+         i_hsp_dsch_dttm = strptime(i_hsp_dsch_dttm, format="%Y-%m-%d") %>% as.POSIXct()) %>% 
+  dplyr::filter(date >= start_date_plots & date <= end_date_plots)
+
+nrow(input_data)
+nrow(unique(data.frame(input_data$patient_num)))
+nrow(input_data %>% distinct(patient_num, encounter_num))
+
+# we will temporarily split the data frame between our psych patients with ICD codes and our non-pysch without, due to the filtering function automatically dropping rows with missing ICD codes
+# of note: these are just some patients who have an ICD code - could be some non-psych codes as the original query including some icd codes for poisoning that ended up being excluded
+keep_psych <- input_data %>% 
+  group_by(patient_num, encounter_num) %>% 
+  filter(i_diagnosis_dttm >= date & i_diagnosis_dttm <= i_hsp_dsch_dttm) %>% 
+  dplyr::select( patient_num, encounter_num, sex, age, race, concept_code, n_hospitalisation, len_hospitalisation, 
+                 date, weeks, month, year, period, time_p, disorder_group, description, icd_code_category)
+
+keep_non_psych = input_data %>% 
+  group_by(patient_num, encounter_num) %>% 
+  filter(is.na(i_diagnosis_dttm)) %>% 
+  dplyr::select( patient_num, encounter_num, sex, age, race, concept_code, n_hospitalisation, len_hospitalisation, 
+                 date, weeks, month, year, period, time_p, disorder_group, description, icd_code_category)
+
+# keep non_psych whose dx date did not occur during the admission
+# but correct the diagnosis data to match the non_psych group
+keep_non_psych_v2 = input_data %>% 
+  group_by(patient_num, encounter_num) %>% 
+  filter(i_diagnosis_dttm < date | i_diagnosis_dttm >= i_hsp_dsch_dttm) %>% 
+  dplyr::select( patient_num, encounter_num, sex, age, race, concept_code, n_hospitalisation, len_hospitalisation, 
+                 date, weeks, month, year, period, time_p, disorder_group, description, icd_code_category) %>% 
+  mutate(concept_code = NA,
+         disorder_group = NA,
+         description = NA, 
+         icd_code_category = NA)
+
+input_data <- rbind(keep_psych, keep_non_psych, keep_non_psych_v2)
+
+nrow(input_data)
+nrow(unique(data.frame(input_data$patient_num)))
+nrow(input_data %>% distinct(patient_num, encounter_num))
+
+all_codes_desc <- input_data %>% 
+  mutate( icd_code_letter = ifelse(substr(concept_code, 1, 1) %in% c("D", "H"), substr(concept_code, 1, 2), substr(concept_code, 1, 1) ))
+
+all_codes_desc <- left_join( all_codes_desc, icdCategory) %>%
+  mutate( icd_description_category = replace_na( ifelse( concept_code %in% icdCodes$ICD10_Code, "Psy", icd_description_category), "Others"),
+          icd_code_category = ifelse(is.na(icd_code_category), "others", icd_code_category))
+
+nrow(all_codes_desc %>% distinct(patient_num, encounter_num))
+
+all_codes_psy <- all_codes_desc %>% 
+  filter(concept_code %in% icdCodes$ICD10_Code)
+
+nrow(unique(data.frame(all_codes_psy$patient_num)))
+nrow(all_codes_psy %>% distinct(patient_num, encounter_num))
+# View(all_codes_psy %>% 
+#        ungroup() %>% 
+#        distinct(patient_num, encounter_num, time_p, period) %>% 
+#        group_by(period) %>% 
+#        mutate(count = n_distinct(patient_num, encounter_num)) %>% 
+#        distinct(period, count))
+
+## how many weeks in 'during' vs 'before' pandemic (15 for both)
+# View(all_codes_psy %>% 
+#        ungroup() %>% 
+#        distinct(month, period) %>%
+#        group_by(period) %>% 
+#        mutate(count = n_distinct(month)))
+
+#all_codes_desc$i_date <- strptime(all_codes_desc$date, format="%Y-%m-%d") %>% as.POSIXct()
+
+# if( raceAvailable == TRUE ){
+#   all_codes_psy <- all_codes_psy %>%
+#     left_join( race_raw, by = c("patient_num") )
+# }
+```
+
+Evaluate numbers before vs after pandemic
+
+```{r}
+before_v_after <- all_codes_desc %>% 
+  distinct(patient_num, encounter_num, period, icd_code_category) %>% 
+  group_by(period, icd_code_category) %>% 
+  mutate(count = n_distinct(patient_num, encounter_num)) %>% 
+  distinct(period, icd_code_category, count)
+
+before_v_after
+```
+
+QC Checks for Psych patients
+
+```{r}
+p1 <- ggplot(all_codes_desc %>%
+               filter(icd_code_category == "psy") %>% 
+               group_by(weeks, period) %>% 
+               arrange(weeks) %>% 
+               mutate(psych_any_count = n_distinct(patient_num, encounter_num),
+                      weeks = as.Date(weeks)) %>% 
+               distinct(weeks, psych_any_count, period),
+       aes(x = weeks, y = psych_any_count, fill = period)) + 
+  geom_bar(stat="identity", binwidth=1) + 
+  geom_smooth(method = "lm", se = TRUE, aes(color = period)) + 
+  scale_color_brewer(palette="Dark2", direction = -1) +
+  scale_fill_brewer(palette = "Dark2", direction = -1) + 
+  ylab("Any Psychiatric Disease") + 
+  xlab("") + 
+  #scale_x_date(date_labels = "%b-%y", date_breaks = "1 month", expand = c(0,0)) + 
+  theme_bw() + 
+  theme(legend.title = element_blank(),
+        axis.text.x = element_text(angle = 45, hjust = 1))
+
+p1
+
+# add an extra day to distinguish the before and during which occur in same month
+
+p1 <- ggplot(all_codes_desc %>% 
+               filter(icd_code_category == "psy") %>% 
+               group_by(month) %>% 
+               arrange(month) %>% 
+               mutate(month = as.character(month), 
+                 month = ifelse(month == "2020-03-01" & period == "during_pandemic", "2020-03-15", month)) %>% 
+               group_by(month, period) %>% 
+               mutate(psych_any_count = n_distinct(patient_num, encounter_num),
+                      month = as.Date(month)) %>%
+               ungroup() %>% 
+               distinct(month, psych_any_count, period),
+       aes(x = month, y = psych_any_count, fill = period)) + 
+  geom_bar(stat="identity", width = 25) + 
+  geom_smooth(method = "lm", se = TRUE, aes(color = period)) + 
+  scale_color_brewer(palette="Dark2", direction = -1) +
+  scale_fill_brewer(palette = "Dark2", direction = -1) + 
+  ylab("Any Psychiatric Disease") + 
+  xlab("") + 
+  scale_x_date(date_labels = "%b-%y", date_breaks = "1 month", expand = c(0,0)) + 
+  theme_bw() + 
+  theme(legend.title = element_blank(),
+        axis.text.x = element_text(angle = 45, hjust = 1))
+
+p1
+
+# p1 <- ggplot(all_codes_desc %>% 
+#                filter(icd_code_category == "psy") %>% 
+#                group_by(month) %>% 
+#                arrange(month) %>% 
+#                mutate(month = as.character(month), 
+#                  period = ifelse(month == "2020-03-01", "March", period)) %>% 
+#                group_by(month, period) %>% 
+#                mutate(psych_any_count = n_distinct(patient_num, encounter_num),
+#                       month = as.Date(month)) %>%
+#                ungroup() %>% 
+#                distinct(month, psych_any_count, period),
+#        aes(x = month, y = psych_any_count, fill = period)) + 
+#   geom_bar(stat="identity", width = 25) + 
+#   geom_smooth(method = "lm", se = TRUE, aes(color = period)) + 
+#   scale_color_brewer(palette="Dark2", direction = -1) +
+#   scale_fill_brewer(palette = "Dark2", direction = -1) + 
+#   ylab("Any Psychiatric Disease") + 
+#   xlab("") + 
+#   scale_x_date(date_labels = "%b-%y", date_breaks = "1 month", expand = c(0,0)) + 
+#   theme_bw() + 
+#   theme(legend.title = element_blank(),
+#         axis.text.x = element_text(angle = 45, hjust = 1))
+# 
+# p1
+
+# remove march
+p1 <- ggplot(all_codes_desc %>% 
+               filter(icd_code_category == "psy") %>% 
+               group_by(month) %>% 
+               arrange(month) %>% 
+               mutate(month = as.character(month)) %>% 
+               filter(!month == "2020-03-01") %>% 
+                 #month = ifelse(month == "2020-03-01" & period == "during_pandemic", "2020-03-15", month)) %>% 
+               group_by(month, period) %>% 
+               mutate(psych_any_count = n_distinct(patient_num, encounter_num),
+                      month = as.Date(month)) %>%
+               ungroup() %>% 
+               distinct(month, psych_any_count, period),
+       aes(x = month, y = psych_any_count, fill = period)) + 
+  geom_bar(stat="identity", width = 25) + 
+  geom_smooth(method = "lm", se = TRUE, aes(color = period)) + 
+  scale_color_brewer(palette="Dark2", direction = -1) +
+  scale_fill_brewer(palette = "Dark2", direction = -1) + 
+  ylab("Any Psychiatric Disease") + 
+  xlab("") + 
+  scale_x_date(date_labels = "%b-%y", date_breaks = "1 month", expand = c(0,0)) + 
+  theme_bw() + 
+  theme(legend.title = element_blank(),
+        axis.text.x = element_text(angle = 45, hjust = 1))
+
+p1
+```
+
+### Psych vs Non-psych
+
+```{r}
+p1 <- ggplot(all_codes_desc %>%
+               #filter(icd_code_category == "psy") %>% 
+               # first, for each month, we will modify the data, so that if a patient has a psych AND non-psych admission in the first month, we will only count their psych admission
+               # group_by(month) %>% 
+               # mutate(check = if_else(icd_code_category == "psy", 1, 0)) %>% 
+               # group_by(patient_num, month) %>% 
+               # filter(check == max(check)) %>% 
+               # slice(1) %>% 
+               group_by(month, period, icd_code_category) %>% 
+               arrange(month) %>% 
+               mutate(month = as.character(month)) %>%
+               filter(!month == "2020-03-01") %>%
+               mutate(psych_any_count = n_distinct(patient_num, encounter_num),
+                      month = as.character(month), 
+                      month = ifelse(month == "2020-03-01" & period == "during_pandemic", "2020-03-15", month),
+                      month = as.Date(month)) %>% 
+               distinct(month, icd_code_category, psych_any_count, period) %>% 
+               pivot_wider(names_from = icd_code_category, values_from = psych_any_count) %>% 
+               mutate(sum = psy + others,
+                      perc_psyc = psy/sum),
+       aes(x = month, y = perc_psyc, fill = period)) + 
+  geom_bar(stat="identity", width = 20) + 
+  geom_smooth(method = "lm", se = TRUE, aes(color = period)) + 
+  scale_color_brewer(palette="Dark2", direction = -1) +
+  scale_fill_brewer(palette = "Dark2", direction = -1) + 
+  ylab("Any Psychiatric Disease") + 
+  xlab("") + 
+  scale_x_date(date_labels = "%b-%y", date_breaks = "1 month", expand = c(0,0)) + 
+  #facet_wrap(~icd_code_category) +
+  theme_bw() + 
+  theme(legend.title = element_blank(),
+        axis.text.x = element_text(angle = 45, hjust = 1))
+
+p1
+
+p1 <- ggplot(all_codes_desc %>%
+               #filter(icd_code_category == "psy") %>% 
+               # first, for each month, we will modify the data, so that if a patient has a psych AND non-psych admission in the first month, we will only count their psych admission
+               # group_by(month) %>% 
+               # mutate(check = if_else(icd_code_category == "psy", 1, 0)) %>% 
+               # group_by(patient_num, month) %>% 
+               # filter(check == max(check)) %>% 
+               # slice(1) %>% 
+               group_by(month, period, icd_code_category) %>% 
+               arrange(month) %>% 
+               # mutate(month = as.character(month)) %>% 
+               # filter(!month == "2020-03-01") %>% 
+               mutate(psych_any_count = n_distinct(patient_num, encounter_num),
+                      month = as.character(month), 
+                      month = ifelse(month == "2020-03-01" & period == "during_pandemic", "2020-03-15", month),
+                      month = as.Date(month)) %>% 
+               distinct(month, icd_code_category, psych_any_count, period) %>% 
+               pivot_wider(names_from = icd_code_category, values_from = psych_any_count) %>% 
+               mutate(sum = psy + others,
+                      perc_psyc = psy/sum),
+       aes(x = month, y = perc_psyc, fill = period)) + 
+  geom_bar(stat="identity", width = 20) + 
+  geom_smooth(method = "lm", se = TRUE, aes(color = period)) + 
+  scale_color_brewer(palette="Dark2", direction = -1) +
+  scale_fill_brewer(palette = "Dark2", direction = -1) + 
+  ylab("Any Psychiatric Disease") + 
+  xlab("") + 
+  scale_x_date(date_labels = "%b-%y", date_breaks = "1 month", expand = c(0,0)) + 
+  #facet_wrap(~icd_code_category) +
+  theme_bw() + 
+  theme(legend.title = element_blank(),
+        axis.text.x = element_text(angle = 45, hjust = 1))
+
+p1
+
+test <- all_codes_desc %>%
+               #filter(icd_code_category == "psy") %>% 
+               # first, for each month, we will modify the data, so that if a patient has a psych AND non-psych admission in the first month, we will only count their psych admission
+               # group_by(month) %>% 
+               # mutate(check = if_else(icd_code_category == "psy", 1, 0)) %>% 
+               # group_by(patient_num, month) %>% 
+               # filter(check == max(check)) %>% 
+               # slice(1) %>% 
+               group_by(month, period, icd_code_category) %>% 
+               arrange(month) %>% 
+               mutate(psych_any_count = n(),
+                      month = as.character(month), 
+                      month = ifelse(month == "2020-03-01" & period == "during_pandemic", "2020-03-15", month),
+                      month = as.Date(month)) %>% 
+               distinct(month, icd_code_category, psych_any_count, period) %>% 
+               pivot_wider(names_from = icd_code_category, values_from = psych_any_count) %>% 
+               mutate(sum = psy + others,
+                      perc_psyc = psy/sum)
+
+before <- test %>% filter(period == "before_pandemic")
+during <- test %>% filter(period == "during_pandemic")
+
+summary(before$perc_psyc)
+summary(during$perc_psyc)
+
+# don't divide by month
+
+#unique patient and encounter by month
+test <- all_codes_desc %>%
+               #filter(icd_code_category == "psy") %>% 
+               # first, for each month, we will modify the data, so that if a patient has a psych AND non-psych admission in the first month, we will only count their psych admission
+               group_by(month) %>% 
+               mutate(check = if_else(icd_code_category == "psy", 1, 0)) %>% 
+               group_by(patient_num, month) %>% 
+               filter(check == max(check)) %>% 
+               slice(1) %>% 
+               group_by(period, icd_code_category) %>% 
+               arrange(month) %>% 
+               mutate(psych_any_count = n_distinct(patient_num, encounter_num),
+                      month = as.character(month), 
+                      month = ifelse(month == "2020-03-01" & period == "during_pandemic", "2020-03-15", month),
+                      month = as.Date(month)) %>% 
+               distinct(icd_code_category, psych_any_count, period) %>% 
+               pivot_wider(names_from = icd_code_category, values_from = psych_any_count) %>% 
+               mutate(sum = psy + others,
+                      perc_psyc = psy/sum)
+
+before <- test %>% filter(period == "before_pandemic")
+during <- test %>% filter(period == "during_pandemic")
+
+summary(before$perc_psyc)
+summary(during$perc_psyc)
+
+```
+
+### Add the non-pysch patient encounters
+
+```{r eval=FALSE, include=FALSE}
+non_psych_inpt_counts$date <- strptime(non_psych_inpt_counts$hsp_adm_dttm, format="%m/%d/%Y %H:%M") %>% as.POSIXct()
+
+# filter age and study inclusion period
+non_psych_inpt_counts <- non_psych_inpt_counts %>% 
+  filter(len_hospitalisation >= min_hosp_days) %>% 
+  filter(age_years > 10 & age_years < 18,
+         date >= start_date_plots & date <= end_date_plots) %>% 
+  dplyr::mutate(#date = as.Date( calendar_date, format = dateFormat ),
+                 weeks = as.Date(cut( date, breaks = "week")),
+                 month = as.Date(cut( date, breaks = "month")),
+                 year = format( date, "%Y"), 
+                 period = ifelse( date < pandemic_start_date,
+                                  "before_pandemic", "during_pandemic"))
+
+if( time_period == "weeks"){
+  non_psych_inpt_counts$time_p <- non_psych_inpt_counts$weeks
+}else if(time_period == "days"){
+  non_psych_inpt_counts$time_p <- non_psych_inpt_counts$date
+}else if(time_period == "months"){
+  non_psych_inpt_counts$time_p <- non_psych_inpt_counts$month
+}
+
+# we will then indicate that these patients do not have psychiatric diseases
+non_psych_inpt_counts <- non_psych_inpt_counts %>% 
+  rename(age = "age_years") %>%  
+  mutate(patient_num = row_number() + 78332,
+         encounter_num = NA, 
+         sex = NA,
+         concept_code = NA, 
+         n_hospitalisation = NA, 
+         len_hospitalisation = NA, 
+         disorder_group = NA,
+         description = NA,
+         icd_code_category = "others", 
+         icd_code_letter = NA,
+         icd_description_category = NA) %>% 
+  select(colnames(all_codes_desc))
+```
+
+Combine the psych and non-psych dataframes
+
+```{r eval=FALSE, include=FALSE}
+all_codes_desc <- rbind(all_codes_desc, non_psych_inpt_counts)
+```
+
 #### QC step: do a summary of the mental health ICD codes present in each site
@@
 #filter the whole dataset to only mental codes
-mental_codes_qc <- all_codes_ado %>% 
-  filter( concept_code %in% mentalHealthCodes$concept_code )
+mental_codes_qc <- all_codes_desc %>% 
+  filter( concept_code %in% mentalHealthCodes$icd_10_code)
 
@@
                                0.5),                           
-                    min_date = min( dmy( calendar_date ) ), 
-                    max_date = max( dmy( calendar_date ) ))%>% 
+                    min_date = min( date ), 
+                    max_date = max( date ))%>% 
   arrange(desc(distinct_observations))
-
 mental_codes_qc_output <- left_join( mental_codes_qc_output, icdCodes, by=c("concept_code"= "ICD10_Code")) 
@@
                                                     "distinct_observations", "min_date", "max_date")]
-
 mental_codes_qc_output[1:10, ]
-```
-
-### Add the psychiatric/mental health ICD code description and filter
-
-```{r message=FALSE, warning=FALSE}
-# Getting site name, will be used later in the analysis
-site <- unique(as.character(demo_raw$siteid))
-
-icdCategory <- read.delim("public-data/icd10Codes.txt",
-                          header = FALSE,
-                          colClasses = "character")
-colnames(icdCategory) <- c("icd_code_letter", "icd_description_category")
-
-all_codes_desc <- all_codes_ado %>% 
-    left_join( icdCodes, by=c("concept_code" = "ICD10_Code") ) %>%
-  mutate(icd_code_category = ifelse(concept_code %in% icdCodes$ICD10_Code, 
-                               "psy", 
-                               "others")) %>%
-  dplyr::select( patient_num, sex, age_time_diagnosis, concept_type, concept_code, calendar_date,
-                 in_hospital, n_hospitalisation, len_hospitalisation, date, weeks, month, year,
-                 period, time_p, disorder_group, description, icd_code_category )
-
-all_codes_desc <- all_codes_desc %>% 
-  mutate( icd_code_letter = ifelse(substr(concept_code, 1, 1) %in% c("D", "H"), substr(concept_code, 1, 2), substr(concept_code, 1, 1) ))
-
-all_codes_desc <- left_join( all_codes_desc, icdCategory) %>%
-  mutate( icd_description_category = replace_na( ifelse( concept_code %in% icdCodes$ICD10_Code, "Psy", icd_description_category), "Others")) 
-all_codes_psy <- all_codes_desc %>% 
-  filter(concept_code %in% icdCodes$ICD10_Code)
-
-if( raceAvailable == TRUE ){
-  all_codes_psy <- all_codes_psy %>%
-    left_join( race_raw, by = c("patient_num") )
-}
 ```
@@
 Summed up over one week, or one month depending on the time unit chosen.
-Not filtered on qge.
+Not filtered on age.
 
@@
 ```{r}
-new_hospitalizations <- clinical_table %>%
+new_hospitalizations <- all_codes_desc %>%
   group_by(patient_num, n_hospitalisation) %>%
   slice(which.min(time_p)) %>%
-  filter( in_hospital == 1 ) %>%
+  #filter( in_hospital == 1 ) %>%
     group_by( time_p ) %>%
@@
 ```{r}
-new_hospitalizations_ado <- clinical_table_ado %>%
+new_hospitalizations_ado <- all_codes_desc %>%
   group_by(patient_num, n_hospitalisation) %>%
   slice(which.min(time_p)) %>%
-  filter( in_hospital == 1 ) %>%
+  #filter( in_hospital == 1 ) %>%
     group_by( time_p ) %>%
@@
 ```{r}
-unique_patients_hospitalized <- clinical_table_ado %>%
+unique_patients_hospitalized <- all_codes_desc %>%
   filter(date < end_date_plots &
@@
          date >= start_date_plots ) %>%
-    group_by( time_p ) %>%
+    group_by( time_p) %>%
   summarise(count_psy =  ifelse(n_distinct(patient_num) > obfuscation | isFALSE(obfuscation),
@@
   geom_smooth(method = "lm", formula = y ~ splines::ns(x, 2), se = TRUE, aes(colour=period), colour= "black", size = 0.5)
+
 thpd_psyCount
@@
 all_patients <- unique( all_codes_desc[, c("patient_num", "time_p")])
+nrow(unique(data.frame(all_patients$patient_num)))
 patients_with_psy_conditions <- unique( all_codes_psy[, c("patient_num", "time_p")] )
+nrow(unique(data.frame(patients_with_psy_conditions$patient_num)))
 patients_without_psy_conditions <- setdiff(all_patients,patients_with_psy_conditions)
+nrow(unique(data.frame(patients_without_psy_conditions$patient_num)))
 
@@
 
-bootstrapped_coefficients_df <- bind_rows(bootstrapped_coefficients_list,
-                                          .id = "iteration")
-
-z <-  mean(bootstrapped_coefficients_df[["periodduring_pandemic:time"]]) / sd(bootstrapped_coefficients_df[["periodduring_pandemic:time"]])
-
-pvalue_coeff_interaction <- round(pnorm(abs(z), lower.tail = FALSE) * 2, 3)
-
-plot_annotation <- paste0("Coeffs differences: ",
+bootstrapped_coefficients_df <- bind_rows(bootstrapped_coefficients_list, .id = "iteration")
+mean(bootstrapped_coefficients_df$time)
+mean(bootstrapped_coefficients_df$time + bootstrapped_coefficients_df[["periodduring_pandemic:time"]])
+pvalue_coeff_interaction <- round(t.test(bootstrapped_coefficients_df[["periodduring_pandemic:time"]])$p.value, 4)
+# pvalue_coeff_interaction <- ifelse(pvalue_coeff_interaction == 0,
+#                                    "< 1e-4",
+#                                    str(pvalue_coeff_interaction))
+if (pvalue_coeff_interaction == 0) {
+  pvalue_coeff_interaction <- 1e-4
+  print("equals 0")
+} else {
+  pvalue_coeff_interaction <- as.character(pvalue_coeff_interaction)
+}
+
+plot_annotation <- paste0("Diff. in slope coeffs before and after pandemic: ",
                      round(mean(bootstrapped_coefficients_df[["periodduring_pandemic:time"]]), 2),
@@
 ```{r}
+nrow(unique(data.frame(all_codes_psy$patient_num)))
 count_disorder_group <- all_codes_psy %>%
@@
 ```
+
 ```{r}
@@
 
+nrow(unique(data.frame(all_codes_psy$patient_num)))
+
 age_by_period <- all_codes_psy %>%
-  distinct(patient_num, age_time_diagnosis, period) %>% 
+  distinct(patient_num, age, period) %>% 
   dplyr::group_by(period) %>%
-  dplyr::summarise( mean = mean( age_time_diagnosis ),
-                    sd = sd( age_time_diagnosis ),
+  dplyr::summarise( mean = mean( age ),
+                    sd = sd( age ),
                     count = n_distinct(patient_num))
@@
 length_hospitalisation_values <- all_codes_psy %>% 
-  dplyr::distinct( patient_num, n_hospitalisation, len_hospitalisation, period ) %>%
+  dplyr::distinct( patient_num, encounter_num, n_hospitalisation, len_hospitalisation, period ) %>%
   dplyr::mutate( siteid = site ) %>%
@@
 
+# de-identify to share with alba/arnaud
+to_share <- length_hospitalisation_values %>% 
+  ungroup() %>% 
+  mutate(siteid = "Lurie") %>% 
+  select(len_hospitalisation, siteid, period)
+
+write.csv(to_share, "../2022.04.06_Lurie_Output/Lurie_los.csv", row.names = FALSE)
+
 # add the site id as column for each of the output data frames
-mental_codes_qc_output$siteid = site 
 count_icd$siteid = site 
@@
 ratios_patients_with_without_psy$siteid = site 
+mental_codes_qc_output$siteid = site
+
+mental_codes_qc_output <- mental_codes_qc_output %>% select(-min_date, -max_date)
 
@@
       table1,
-      length_hospitalisation_values,
-      file = paste0("output/peds_psy_aggResults_", site, ".RData"))
-```
+      #length_hospitalisation_values,
+      file = paste0("../2022.04.06_Lurie_Output/peds_psy_aggResults_", site, ".RData"))
+
+rm(length_hospitalisation_values)
+```

### 2022-04-28 metanalysis_visualization/manuscript_Figures.Rmd
fix rbinom bug, change bootstrap to 10000 and change time range for plots

@@
 pandemic_start_date <- as.Date("2020-03-15")
-start_date_plots <- as.Date("2019-01-01")
-end_date_plots <- as.Date("2021-05-31")
+start_date_plots <- as.Date("2019-02-01")
+end_date_plots <- as.Date("2021-04-30")
 ```
@@
 ```{r}
-start_date_plots <- as.Date("2019-01-01")
-end_date_plots <- as.Date("2021-05-31")
+start_date_plots <- as.Date("2019-02-01")
+end_date_plots <- as.Date("2021-04-30")
 
@@
 # Call function from another file : bootstrap.R
-bootstrap_size <- 1000
+bootstrap_size <- 10000
 binom_proportion <- function(n, prob, size) {
@@
     n = bootstrap_size,
-    size = df[[count]],
+    size = as.integer(df[[count]]),
     prob = df[[percentage]]/100)
@@
   labs(title = "Interrupted Time Series Analysis: Percentage of patients with psychiatric conditions",
-       subtitle = "Parametric bootstrap estimation: (1000 iterations)",
+       subtitle = "Parametric bootstrap estimation: (10000 iterations)",
        x = "Calendar dates",
@@
         plot.subtitle = element_text(size = 12, face = "italic")) 
-ggsave("boostrapped_all_long.png", height = 20, width = 45, units = "cm")
+ggsave("figure1_boostrapped_all.png", height = 20, width = 45, units = "cm")
 ```
@@
                 mlab="Overall periods' slope differences")
+ggsave("figure2A_metaanalysis_persite.png", height = 20, width = 45, units = "cm")
+
 ```
@@
        mlab = "Overall periods' slope differences")
+ggsave("figure2B_metaanalysis_country.png", height = 20, width = 45, units = "cm")
+
 ```
@@
                  plotsLabel = "Female")
-
-# plot_annotations_sex_df <- df_coefficients_boot_sex %>% 
-#   group_by(sex) %>% 
-#   group_map(annotation_function) %>%
-#   paste(str_to_title(names_sex), ., sep = ": ") %>% 
-#   as_tibble() %>% 
-#   rename(label = value)
-# bind_cols(group_names) %>% 
-#                rename(label = value) %>% 
-#                mutate(label = paste(stringr::str_to_title(sex), label, sep = ": ")) %>% 
-#                mutate(plotsLabel = names_site_id)
-#              , .keep = TRUE) %>% 
-  # group_map( function(df, group_names) group_by(df, plotsLabel) %>%
-  #              annotation_function() %>%
-               # as_tibble() %>%
-               # bind_cols(group_names) %>%
-               # rename(label = value) %>%
-               # mutate(label = paste(stringr::str_to_title(sex), label, sep = ": ")) %>%
-  #              mutate(plotsLabel = names_site_id)
-  #            , .keep = TRUE) %>% 
-  # set_names(unique(df_coefficients_boot_sex$sex))
-
 ```
@@
             inherit.aes = FALSE) +
-  # geom_label(data = plot_annotations_sex_df[["male"]],
-  #            aes(label=label),
-  #            x = -Inf,
-  #            y = 15,
-  #            hjust = 0,
-  #            vjust = 0,
-  #            size = 4,
-  #           inherit.aes = FALSE) +
-  # ylim(10, 90) +
   labs(title = "Interrupted Time Series Analysis: Percentage of patients with psychiatric conditions",
-  subtitle = "Parametric bootstrap estimation: (1000 iterations)",
+  subtitle = "Parametric bootstrap estimation: (10000 iterations)",
        x = "Calendar dates",
@@
   plot.subtitle = element_text(size = 12, face = "italic"))
+ggsave("figure4_ITS_sex.png", height = 20, width = 45, units = "cm")
+
 ```
@@
 
-```{r}
+```{r }
 disorders_to_select <- c(
@@
         title = element_text(size = 14),
-        plot.subtitle = element_text(size = 12, face = "italic")) 
-ggsave("boostrapped_disorders.png", height = 14, width = 25, units = "cm")
+        plot.subtitle = element_text(size = 8, face = "italic")) 
+ggsave("figure3_boostrapped_disorders.png", height = 15, width = 35, units = "cm")
 
@@
 fig
+ggsave("figure_supp3_counts_disorder.png", height = 20, width = 45, units = "cm")
 
@@
         legend.position="bottom") 
+ggsave("figure_supp2_ICDdictionaryQC.png", height = 20, width = 45, units = "cm")
+
 ```
@@
 counts_plot
-ggsave("counts_plot.png",counts_plot, height = 20, width = 32, units = "cm")
+ggsave("figure_supp1A_counts_per_site.png",counts_plot, height = 20, width = 32, units = "cm")
 
@@
         legend.position="bottom")
-```
+ggsave("figure_supp1B_counts_all.png",counts_plot, height = 20, width = 32, units = "cm")
+
+```

### 2022-04-28 metanalysis_visualization/manuscript_Figures.Rmd
add the saving figures for the manuscript

@@
        color = "Period") +
-  theme(strip.text = element_text(size = 10),
+  theme(strip.text = element_text(size = 14),
         axis.text.x=element_text(angle = 45, hjust=1, size = 8),
@@
         title = element_text(size = 14),
-        plot.subtitle = element_text(size = 12, face = "italic")) 
+        plot.subtitle = element_text(size = 12, face = "italic"), 
+        legend.position="bottom") 
 ggsave("figure1_boostrapped_all.png", height = 20, width = 45, units = "cm")
@@
 # Random effect
+png(file='./figure2A_metaanalysis_persite.png')
+
 trends_siteid <- trends_siteid %>%
@@
                 mlab="Overall periods' slope differences")
-ggsave("figure2A_metaanalysis_persite.png", height = 20, width = 45, units = "cm")
+dev.off()
+#ggsave("figure2A_metaanalysis_persite.png", height = 20, width = 45, units = "cm")
 
@@
 ```{r}
+png(file='figure2B_metaanalysis_country.png')
+
 trends_country <- trends_siteid %>% 
@@
        mlab = "Overall periods' slope differences")
-ggsave("figure2B_metaanalysis_country.png", height = 20, width = 45, units = "cm")
-
+dev.off()
 ```
@@
         title = element_text(size = 14),
-  plot.subtitle = element_text(size = 12, face = "italic"))
-ggsave("figure4_ITS_sex.png", height = 20, width = 45, units = "cm")
-
+  plot.subtitle = element_text(size = 12, face = "italic"), 
+  legend.position = "bottom")
+ggsave("figure4_ITS_sex.png", height = 20, width = 20, units = "cm")
 ```
@@
         title = element_text(size = 14),
-        plot.subtitle = element_text(size = 8, face = "italic")) 
-ggsave("figure3_boostrapped_disorders.png", height = 15, width = 35, units = "cm")
+        plot.subtitle = element_text(size = 8, face = "italic"), 
+        legend.position = "bottom") 
+ggsave("figure3_boostrapped_disorders.png", height = 15, width = 25, units = "cm")
 
@@
 fig
-ggsave("figure_supp3_counts_disorder.png", height = 20, width = 45, units = "cm")
-
-#### facet wrap in 3 groups based on %
-### most prevalent
-perc_disorder_group_total %>%
-  filter(disorder_group %in% c("Anxiety Disorders", "Depressive Disorders", "Suicide or Self-Injury" ))  %>%
-  ggplot( aes( x= as.Date( time_p ), y = percentage, group = country )) +
-  facet_wrap( ~ disorder_group ) +
-  geom_line( aes(color = country )) +
-  geom_point( aes( color = country )) +
-  geom_vline(xintercept = as.Date(pandemic_start_date), linetype = "dashed")+
-  labs(x ="", y = "percentage of patients (%)")+
-  scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0))+
-  theme(axis.text.x=element_text(angle = 45, hjust=1, size = 6),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.border = element_rect(colour = "black", fill = NA)) +
-  scale_color_manual( values = cbPalette )
-
-#medium prevalent
-perc_disorder_group_total %>%
-  filter(disorder_group %in% c("Trauma and Stressor-Related Disorders", "Mental Health Symptom", "Disruptive, Impulse Control and Conduct Disorders", 
-                               "Feeding and Eating Disorders"))  %>%
-  ggplot( aes( x= as.Date( time_p ), y = percentage, group = country )) +
-  facet_wrap( ~ disorder_group ) +
-  geom_line( aes(color = country )) +
-  geom_point( aes( color = country )) +
-  geom_vline(xintercept = as.Date(pandemic_start_date), linetype = "dashed")+
-  labs(x ="", y = "percentage of patients (%)")+
-  scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0))+
-  theme(axis.text.x=element_text(angle = 45, hjust=1, size = 6),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.border = element_rect(colour = "black", fill = NA)) +
-  scale_color_manual( values = cbPalette )
-
-#low prevalent
-perc_disorder_group_total %>%
-  filter(disorder_group %in% c("Miscellaneous", "Substance-Related and Addictive Disorders", 
-                               "Personality Disorders", "Bipolar and Related Disorders", "Sleep-Wake Disorders", 
-                               "Schizophrenia Spectrum and Other Psychotic Disorders", "Somatic Symptom and Related Disorders", 
-                               "Obsessive-Compulsive and Related Disorders", "Dissociative Disorders" ))  %>%
-  ggplot( aes( x= as.Date( time_p ), y = percentage, group = country )) +
-  facet_wrap( ~ disorder_group ) +
-  geom_line( aes(color = country )) +
-  geom_point( aes( color = country )) +
-  geom_vline(xintercept = as.Date(pandemic_start_date), linetype = "dashed")+
-  labs(x ="", y = "percentage of patients (%)")+
-  scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0))+
-  theme(axis.text.x=element_text(angle = 45, hjust=1, size = 6),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.border = element_rect(colour = "black", fill = NA)) +
-  scale_color_manual( values = cbPalette )
-
-perc_disorder_group_total %>%
-  filter(disorder_group %in% c("Anxiety Disorders", "Depressive Disorders", "Suicide or Self-Injury" ))  %>%
-  ggplot( aes( x= as.Date( time_p ), y = percentage, group = country )) +
-  facet_wrap( ~ disorder_group ) +
-  geom_line( aes(color = country )) +
-  geom_point( aes( color = country )) +
-  geom_vline(xintercept = as.Date(pandemic_start_date), linetype = "dashed") +
-  labs(x ="", y = "percentage of patients (%)")+
-  scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0)) +
-  scale_color_manual( values = cbPalette ) + 
-  theme(axis.text.x=element_text(angle = 45, hjust=1, size = 6), 
-        strip.text = element_text(size = 10, face = "bold"),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), strip.background = element_blank(),
-        panel.border = element_rect(colour = "black", fill = NA), 
-        legend.position="bottom")
-
-perc_disorder_group_total %>%
-  filter( disorder_group %in% c("Anxiety Disorders", "Depressive Disorders", "Suicide or Self-Injury", 
-  "Mental Health Symptoms", "Feeding and Eating Disorders" )) %>%
-  ggplot( aes( x= as.Date( time_p ), y = percentage, group = country )) +
-  geom_line( aes(linetype = country, color = country )) +
-  geom_point( aes( shape= country, color = country )) +
-  #scale_y_continuous(breaks = seq(0, 30, by = 5)) +
-  facet_wrap2("disorder_group", axes = "all", scales ="free") + 
-  geom_vline(xintercept = as.Date(pandemic_start_date), linetype = "dashed") +
-  labs(x ="", y = "percentage of patients (%)")+
-  scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0)) +
-  scale_color_manual( values = cbPalette ) + 
-  theme(axis.text.x=element_text(angle = 45, hjust=1, size = 6), 
-        strip.text = element_text(size = 10, face = "bold"),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), strip.background = element_blank(),
-        panel.border = element_rect(colour = "black", fill = NA), 
-        legend.position="bottom")
-
+#plotly::orca(fig, file = "figure_supp3_counts_disorder.png")
 ```
@@
 ```{r}
-head(patient_count_psy_period_psy_all)
 sites <- unique( patient_count_psy_period_psy_all$siteid )
@@
 g <- do.call(grid.arrange,p)
-ggsave("its_seasonality_site.png",g, height = 20, width = 32, units = "cm")
+#ggsave("its_seasonality_site.png",g, height = 20, width = 32, units = "cm")
 
@@
 
-counts_agg2 %>% ggplot( aes( x= as.Date( time_p ), y = value, group= group )) +
+counts_agg2plot <- counts_agg2 %>% ggplot( aes( x= as.Date( time_p ), y = value, group= group )) +
   geom_point( aes( shape= group, color = group )) +
@@
         legend.position="bottom")
-ggsave("figure_supp1B_counts_all.png",counts_plot, height = 20, width = 32, units = "cm")
+counts_agg2plot
+ggsave("figure_supp1B_counts_all.png",counts_agg2plot, height = 20, width = 32, units = "cm")
 

### 2022-05-02 metanalysis_visualization/manuscript_Figures.Rmd
tweaking figures

@@
 # Call function from another file : bootstrap.R
-bootstrap_size <- 10000
+bootstrap_size <- 1000
 binom_proportion <- function(n, prob, size) {
@@
 ```{r}
-annotation_function <- function(df, ...) {
+bootstrap_pvalue_calculation <- function(df, ...) {
   z <- unique(df[["observed_diff"]]) / sd(df[["periodduring_pandemic:time"]])
   pvalue_coeff_interaction <- round(pnorm(abs(z), lower.tail = FALSE) * 2, 3)
-  paste0("Coeffs differences: ",
-         round(unique(df[["observed_diff"]]), 2),
-         " (pvalue: ",
-         format.pval(pvalue_coeff_interaction),
-         ")")
+  list("pvalue" = pvalue_coeff_interaction, 
+       "observed_diff" = round(unique(df[["observed_diff"]]), 2)
+  ) 
 }
+
 ```
@@
 names_site_id <- unique(df_coefficients_boot$siteid)
+
 plot_annotations <- df_coefficients_boot %>% 
   group_by(siteid) %>% 
-  group_map( annotation_function ) %>% 
+  group_map( bootstrap_pvalue_calculation ) %>% 
   set_names(names_site_id) %>% 
+  lapply(FUN = function(vec) {
+    paste0("Coeffs differences: ",
+           vec$observed_diff,
+           " (pvalue: ",
+           format.pval(vec$pvalue, 
+                       eps = 10**-5),
+           ")")
+  }) %>% 
   unlist()
+
 plot_annotations_df <- data.frame(
@@
 
-```{r}
-sexInfo_per_country <- table1_all %>%
-  filter( Category == "Sex") %>%
-  mutate ( counts_before = as.numeric(sapply(strsplit( as.character(before_pandemic), " "), '[', 1)), 
-           counts_during = as.numeric(sapply(strsplit( as.character(during_pandemic), " "), '[', 1))) %>%
-  select( name, siteid, counts_before, counts_during ) %>%
-  inner_join( country_map ) %>%
-  group_by( name, country ) %>%
-  summarise( before = sum( counts_before), 
-             during = sum( counts_during )) %>%
-  pivot_longer(names_to = "period", cols=c(before, during)) 
-sexInfo_all <- table1_all %>%
-  filter( Category == "Sex") %>%
-  mutate ( counts_before = as.numeric(sapply(strsplit( as.character(before_pandemic), " "), '[', 1)), 
-           counts_during = as.numeric(sapply(strsplit( as.character(during_pandemic), " "), '[', 1))) %>%
-  select( name, siteid, counts_before, counts_during ) %>%
-  inner_join( country_map ) %>%
-  group_by( name ) %>%
-  summarise( before = sum( counts_before), 
-             during = sum( counts_during )) %>%
-  pivot_longer(names_to = "period", cols=c(before, during)) %>%
-  mutate( country = 'ALL' ) %>%
-  select( name, country, period, value ) %>%
-  rbind( sexInfo_per_country)
-sexInfo_all %>%
-  ggplot( aes( x= name, y = value, fill = period )) +
-  geom_bar( stat = "identity", position = "dodge") +
-  facet_wrap("country", scales = "free_y") + 
-  labs(x ="", y = "number of patients") +
-  theme(axis.text.x=element_text(hjust=1, size = 8), 
-        strip.text = element_text(size = 10, face = "bold"),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), strip.background = element_blank(),
-        panel.border = element_rect(colour = "black", fill = NA))+ 
-  scale_fill_manual("Pandemic \n period", values = c("before" = "grey", "during" = "orange"))
-
-```
-
-```{r}
+```{r, include=FALSE}
 patients_hospitalized_agg_sex_total <- patients_hospitalized_agg_sex_perc_all %>%
@@
              siteid = "ALL")
-
-patients_hospitalized_agg_sex_total %>%
-  ggplot( aes( x= as.Date( time_p ), y = percentage, group = sex )) +
-  geom_line( aes( linetype = sex , color = sex)) +
-  geom_point( aes( shape = sex, color = sex )) +
-  scale_colour_manual(values=c("#FF6600","#339999")) +
-  scale_fill_manual(values=c("#FF6600","#339999"))+
-  geom_vline(xintercept = as.Date(pandemic_start_date), linetype = "dashed")+
-  labs(x ="", y = "percentage of patients (%)")+
-  scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0))+
-  theme(axis.text.x=element_text(angle = 45, hjust=1, size = 6),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.border = element_rect(colour = "black", fill = NA))
-
 ```
@@
                                   group_name = .y,
-                                  bootstrap_size = bootstrap_size,
+                                  bootstrap_size = 10000,
                                   count = "count_sex", 
@@
 ## Using only male, but female and male necessarily yield equal p-values
-plot_annotations_sex_df <- annotation_function(
-  df_coefficients_boot_sex[df_coefficients_boot_sex$sex == "female", ]
-) %>% data.frame(label = ., 
+
+pvalue_sex_slope_diff <- bootstrap_pvalue_calculation(  df_coefficients_boot_sex[df_coefficients_boot_sex$sex == "female", ]
+)
+
+plot_annotations_sex_df <- paste0("Female/male proportion coefficient differences: ",
+                                  pvalue_sex_slope_diff$observed_diff,
+                                  " (pvalue: ",
+                                  format.pval(pvalue_sex_slope_diff$pvalue, 
+                                              eps = 10**-5),")") %>%
+  data.frame(label = ., 
              plotsLabel = "Female")
@@
 ```{r bootstrap_plot}
-ggplot(bootstrapped_fitted_sex, aes(x = time_p, y = predicted, group = interaction(iteration, period, sex))) +
-  # facet_wrap("plotsLabel", scales = "free_y", labeller = label_wrap_gen(width = 35,multi_line = TRUE)) +
+ggplot(bootstrapped_fitted_sex, aes(x = time_p, y = predicted/100, group = interaction(iteration, period, sex))) +
   geom_line(aes(color = interaction(period, sex)),
-            alpha = 0.03) +
+            alpha = 0.003) +
   geom_line(data = quantile_bootstrapped_fitted_sex,
-            aes(y = value, group = interaction(quantiles, period, sex)),
+            aes(y = value/100, group = interaction(quantiles, period, sex)),
             linetype = "dashed") +
   geom_line(data = quantile_bootstrapped_fitted_sex[quantile_bootstrapped_fitted_sex$quantiles == "50%", ],
-            aes(y = value, group = interaction(period, sex))) +
+            aes(y = value/100, group = interaction(period, sex))) +
   geom_point(data = patients_hospitalized_agg_sex_total,
-             aes(y = percentage, group=sex, color = interaction(period, sex)),
+             aes(y = percentage/100, group=sex),
              alpha = 1) +
-  scale_color_brewer(palette = "Paired") +
+  # scale_color_brewer(palette = "Paired") +
+  scale_color_manual(labels = c("Female", "Male", "", ""),
+                     values = c("#A6CEE3", "#1F78B4", "#B2DF8A", "#33A02C")) +
   scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0))+
-  # scale_colour_manual(values = c("#A6CEE3", "#1F78B4", "#B2DF8A", "#33A02C")) +
-  guides(color = guide_legend(override.aes = list(alpha = 1))) +
+  guides(colour = guide_legend(
+    override.aes = list(alpha = 1,
+                        colour = c('#1F78B4', '#33A02C', NA, NA))) 
+  ) +
   geom_label(data = plot_annotations_sex_df,
-             aes(label=label),
+             aes(label = label),
              x = -Inf,
@@
              inherit.aes = FALSE) +
-  labs(title = "Interrupted Time Series Analysis: Percentage of patients with psychiatric conditions",
+  labs(title = "Interrupted Time Series Analysis: Patients sex proportion",
        subtitle = "Parametric bootstrap estimation: (10000 iterations)",
        x = "Calendar dates",
-       y = "Percentage",
-       color = "Period") +
+       y = "Sex proportion",
+       color = "Sex") +
   theme(strip.text = element_text(size = 10),
@@
         title = element_text(size = 14),
-  plot.subtitle = element_text(size = 12, face = "italic"), 
-  legend.position = "bottom")
+  plot.subtitle = element_text(size = 12, face = "italic"))
 ggsave("figure4_ITS_sex.png", height = 20, width = 20, units = "cm")
@@
   group_by(disorder_group) %>% 
-  group_map( annotation_function ) %>% 
+  group_map( bootstrap_pvalue_calculation ) %>%
   set_names(names_disorder) %>% 
+  lapply(FUN = function(vec) {
+    paste0("Coeffs differences: ",
+           vec$observed_diff,
+           " (pvalue: ",
+           format.pval(vec$pvalue, 
+                       eps = 10**-5),
+           ")")
+  }) %>% 
   unlist()
@@
 
-### By site
-```{r}
-sites <- unique( patient_count_psy_period_psy_all$siteid )
-source("../metanalysis_visualization/time_series_analysis_plotting.R")
-p <- list()
-
-for( i in 1:length(sites)){
-  print(i)
-  subset_site <- patient_count_psy_period_psy_all %>%
-    filter( siteid == sites[i])
-  ts_seasonal_percent_per_site <- lm(percentage ~ period*time + harmonic(month_year, 1, 12), subset_site)
-  p[[i]] <- time_series_analysis_plotting(ts_seasonal_percent_per_site,
-                              subset_site,
-                              "percentage",
-                              clearance_period = FALSE) +
-  labs(title = sites[i],
-       x = "Calendar dates",
-       y = "Percentage") +
-      scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0)) +
-     theme(strip.text = element_text(size = 10),
-        axis.text.x = element_text(angle = 45, hjust=1, size = 8), 
-        axis.text.y = element_text(size = 8), 
-        axis.title.x = element_text(size = 8),
-        axis.title.y = element_text(size = 8),
-        title = element_text(size = 14),
-        plot.subtitle = element_text(size = 12, face = "italic"))
-
-}
-
-g <- do.call(grid.arrange,p)
-#ggsave("its_seasonality_site.png",g, height = 20, width = 32, units = "cm")
-
-```
-
 ### Total counts with psychiatric conditions and total counts per month
@@
 ```
+
+# Unused figures
+
+```{r, include = FALSE}
+sexInfo_per_country <- table1_all %>%
+  filter( Category == "Sex") %>%
+  mutate ( counts_before = as.numeric(sapply(strsplit( as.character(before_pandemic), " "), '[', 1)), 
+           counts_during = as.numeric(sapply(strsplit( as.character(during_pandemic), " "), '[', 1))) %>%
+  select( name, siteid, counts_before, counts_during ) %>%
+  inner_join( country_map ) %>%
+  group_by( name, country ) %>%
+  summarise( before = sum( counts_before), 
+             during = sum( counts_during )) %>%
+  pivot_longer(names_to = "period", cols=c(before, during)) 
+sexInfo_all <- table1_all %>%
+  filter( Category == "Sex") %>%
+  mutate ( counts_before = as.numeric(sapply(strsplit( as.character(before_pandemic), " "), '[', 1)), 
+           counts_during = as.numeric(sapply(strsplit( as.character(during_pandemic), " "), '[', 1))) %>%
+  select( name, siteid, counts_before, counts_during ) %>%
+  inner_join( country_map ) %>%
+  group_by( name ) %>%
+  summarise( before = sum( counts_before), 
+             during = sum( counts_during )) %>%
+  pivot_longer(names_to = "period", cols=c(before, during)) %>%
+  mutate( country = 'ALL' ) %>%
+  select( name, country, period, value ) %>%
+  rbind( sexInfo_per_country)
+sexInfo_all %>%
+  ggplot( aes( x= name, y = value, fill = period )) +
+  geom_bar( stat = "identity", position = "dodge") +
+  facet_wrap("country", scales = "free_y") + 
+  labs(x ="", y = "number of patients") +
+  theme(axis.text.x=element_text(hjust=1, size = 8), 
+        strip.text = element_text(size = 10, face = "bold"),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), strip.background = element_blank(),
+        panel.border = element_rect(colour = "black", fill = NA))+ 
+  scale_fill_manual("Pandemic \n period", values = c("before" = "grey", "during" = "orange"))
+
+```
+
+```{r, include=FALSE}
+patients_hospitalized_agg_sex_total %>%
+  ggplot( aes( x= as.Date( time_p ), y = percentage, group = sex )) +
+  geom_line( aes( linetype = sex , color = sex)) +
+  geom_point( aes( shape = sex, color = sex )) +
+  scale_colour_manual(values=c("#FF6600","#339999")) +
+  scale_fill_manual(values=c("#FF6600","#339999"))+
+  geom_vline(xintercept = as.Date(pandemic_start_date), linetype = "dashed")+
+  labs(x ="", y = "percentage of patients (%)")+
+  scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0))+
+  theme(axis.text.x=element_text(angle = 45, hjust=1, size = 6),
+        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        panel.border = element_rect(colour = "black", fill = NA))
+
+```
+
+### By site
+```{r, include=FALSE}
+sites <- unique( patient_count_psy_period_psy_all$siteid )
+source("../metanalysis_visualization/time_series_analysis_plotting.R")
+p <- list()
+
+for( i in 1:length(sites)){
+  print(i)
+  subset_site <- patient_count_psy_period_psy_all %>%
+    filter( siteid == sites[i])
+  ts_seasonal_percent_per_site <- lm(percentage ~ period*time + harmonic(month_year, 1, 12), subset_site)
+  p[[i]] <- time_series_analysis_plotting(ts_seasonal_percent_per_site,
+                              subset_site,
+                              "percentage",
+                              clearance_period = FALSE) +
+  labs(title = sites[i],
+       x = "Calendar dates",
+       y = "Percentage") +
+      scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0)) +
+     theme(strip.text = element_text(size = 10),
+        axis.text.x = element_text(angle = 45, hjust=1, size = 8), 
+        axis.text.y = element_text(size = 8), 
+        axis.title.x = element_text(size = 8),
+        axis.title.y = element_text(size = 8),
+        title = element_text(size = 14),
+        plot.subtitle = element_text(size = 12, face = "italic"))
+
+}
+
+g <- do.call(grid.arrange,p)
+#ggsave("its_seasonality_site.png",g, height = 20, width = 32, units = "cm")
+
+```

### 2022-05-10 metanalysis_visualization/manuscript_Figures.Rmd
Several plot tweaks

@@
                                   group_name = .y,
-                                  bootstrap_size = 10000,
+                                  bootstrap_size = bootstrap_size,
                                   count = "count_sex", 
@@
 
-plot_annotations_sex_df <- paste0("Female/male proportion coefficient differences: ",
+plot_annotations_sex_df <- paste0("Female proportion period differences: ",
                                   pvalue_sex_slope_diff$observed_diff,
@@
   geom_line(aes(color = interaction(period, sex)),
-            alpha = 0.003) +
+            alpha = 0.01) +
   geom_line(data = quantile_bootstrapped_fitted_sex,
@@
   plot.subtitle = element_text(size = 12, face = "italic"))
+
 ggsave("figure4_ITS_sex.png", height = 20, width = 20, units = "cm")
@@
 
-ggplot(bootstrapped_fitted_disorder,
-       aes(x = time_p, y = predicted, group = interaction(iteration, period))) +
+bootstrapped_fitted_disorder %>% 
+  mutate(period = recode_factor(period, "before_pandemic" = "Before pandemic onset", "during_pandemic" = "Since pandemic onset")) %>% 
+  ggplot(aes(x = time_p, y = predicted, group = interaction(iteration, period))) +
   facet_wrap("disorder_group",

### 2022-05-10 metanalysis_visualization/manuscript_Figures.Rmd
table 1 add percentages and 95 CI

@@
 }
-listOfPackages <- c("tidyverse", "tidyr","devtools","RColorBrewer", "knitr", "kableExtra", "tsModel", "gridExtra", "dplyr", "metafor", "meta", "viridis", "UpSetR", "scales", "EHRtemporalVariability", "plotly")
+listOfPackages <- c("tidyverse", "tidyr","devtools","RColorBrewer", "knitr", "kableExtra", "tsModel", "gridExtra", "dplyr", "metafor", "meta", "viridis", "UpSetR", "scales", "EHRtemporalVariability", "plotly", "gmodels")
 paket(listOfPackages)
@@
 ```{r}
-# cut-off dates used in the analysis, these should not be changed
 time_period <-"months" #possible values: days, weeks, months
-#history_start_date <- as.Date("2019-01-01")
-#history_end_date <- as.Date("2021-05-31")
 pandemic_start_date <- as.Date("2020-03-15")
@@
 ```{r}
-outputDir <- "../../output/latest_output//"
+outputDir <- "../../output/latest_output/"
 outputFiles <- list.files( path = outputDir, pattern = "\\.RData")
@@
   group_by(period) %>% 
-  summarise(mean_age = round(sum(Total * mean_age) / sum(Total), 2), 
-            sd_age = round(sum(Total * sd_age) / sum(Total), 2)) %>% 
-  mutate(Age = paste0(as.character(mean_age), " (sd: ", as.character(sd_age), ")")) %>% 
+  summarise(conf_interval = gmodels::ci( mean_age ),
+            mean_age = round(sum(Total * mean_age) / sum(Total), 2), 
+            sd_age = round(sum(Total * sd_age) / sum(Total), 2), 
+            ci_age = paste0("[", round(as.numeric(conf_interval[2]), 2), " , ",  round(as.numeric(conf_interval[3]),2), "]")) %>% 
+  mutate(Age = paste0(as.character(mean_age), " (sd: ", as.character(sd_age), ")", as.character(ci_age))) %>% 
   select(period, Age) %>% 
+  unique() %>%
   pivot_wider(names_from = period, values_from = Age) %>% 
   mutate(Category = "Age", 
-         name = "mean (sd)") %>% 
+         name = "mean (sd), 95% CI") %>% 
   select(Category, name, before_pandemic, during_pandemic) %>% 
@@
 median_table1_all <- length_hospitalisation_values_all %>% 
-  filter(siteid != "CCHMC") %>% # temporary
   group_by(period) %>% 
-  summarise(median_hospitalisation = median(len_hospitalisation)) %>% 
+  summarise(median_hospitalisation = mean(len_hospitalisation)) %>% 
   pivot_wider(names_from = period, values_from = median_hospitalisation) %>% 
@@
          name = "median") %>% 
+  select(Category, name, before_pandemic, during_pandemic) %>% 
+  rename(before_pandemic_count = before_pandemic, 
+         during_pandemic_count = during_pandemic)
+median_table1_all$before_pandemic_count <- as.character(median_table1_all$before_pandemic_count)
+median_table1_all$during_pandemic_count <- as.character(median_table1_all$during_pandemic_count)
+
+mean_table1_all <- length_hospitalisation_values_all %>% 
+  group_by(period) %>% 
+  summarise(mean_hospitalisation = round( mean(len_hospitalisation), 2), 
+            conf_interval = gmodels::ci( len_hospitalisation ), 
+            ci_length = paste0("[", round(as.numeric(conf_interval[2]), 2), " , ",  round(as.numeric(conf_interval[3]),2), "]")) %>%
+  mutate(mean_hospitalisation = paste0(as.character(mean_hospitalisation), as.character(ci_length))) %>% 
+  select(period, mean_hospitalisation) %>%
+  unique() %>%
+  pivot_wider(names_from = period, values_from = mean_hospitalisation) %>% 
+  mutate(Category = "Length of stays", 
+         name = "mean [95% CI]") %>% 
   select(Category, name, before_pandemic, during_pandemic) %>% 
@@
   filter( Category == "Sex" & name != "Total") %>%
-  mutate( before_counts = sapply(strsplit( before_pandemic, " "), '[', 1), 
-          during_counts = sapply(strsplit( during_pandemic, " "), '[', 1)) %>%
+  mutate( before_counts = as.numeric(sapply(strsplit( before_pandemic, " "), '[', 1)), 
+          during_counts = as.numeric(sapply(strsplit( during_pandemic, " "), '[', 1)), 
+          before_total = sum( before_counts ), 
+          during_total = sum( during_counts)) %>%
   group_by( name ) %>%
   summarise( before_pandemic_count = sum( as.numeric( before_counts )), 
-             during_pandemic_count = sum( as.numeric( during_counts ))) %>%
-  mutate( Category = "Sex") %>%
-  select( Category, name, before_pandemic_count, during_pandemic_count )
-
-sex_table1_all <- rbind( sex_table1_all, c("Total", "",sum( sex_table1_all$before_pandemic_count), sum( sex_table1_all$during_pandemic_count)))
+             during_pandemic_count = sum( as.numeric( during_counts )), 
+             before_pandemic_perc  = round(100*before_pandemic_count/before_total, 1), 
+             during_pandemic_perc  = round(100*during_pandemic_count/during_total, 1)) %>%
+  mutate( Category = "Sex", 
+          before_pandemic_count = paste0( before_pandemic_count, " (", before_pandemic_perc, ")"), 
+          during_pandemic_count = paste0( during_pandemic_count, " (", during_pandemic_perc, ")")) %>%
+  select( Category, name, before_pandemic_count, during_pandemic_count ) %>%
+  unique()
 
@@
   summarise( before_pandemic_count = sum( as.numeric( before_counts ), na.rm = TRUE), 
-             during_pandemic_count = sum( as.numeric( during_counts ), na.rm = TRUE)) %>%
-  mutate( Category = "Disorder groups") %>%
+             during_pandemic_count = sum( as.numeric( during_counts ), na.rm = TRUE), 
+              before_pandemic_perc  = round(100*before_pandemic_count/9696, 1), 
+             during_pandemic_perc  = round(100*during_pandemic_count/10561, 1)) %>%
+  mutate( Category = "Disorder groups", 
+           before_pandemic_count = paste0( before_pandemic_count, " (", before_pandemic_perc, ")"), 
+          during_pandemic_count = paste0( during_pandemic_count, " (", during_pandemic_perc, ")"))%>%
   select( Category, name, before_pandemic_count, during_pandemic_count )
 
-table1 <- rbind( sex_table1_all, age_table1_all, median_table1_all, dg_table1_all )
+table1 <- rbind( sex_table1_all, age_table1_all, median_table1_all, mean_table1_all, dg_table1_all )
 

### 2022-05-10 metanalysis_visualization/manuscript_Figures.Rmd
p-value for figures 1 and 3 updated, round to 2 digits when >= 0.01

@@
   pvalue_coeff_interaction <- round(pnorm(abs(z), lower.tail = FALSE) * 2, 3)
+  pvalue_coeff_interaction <- ifelse( pvalue_coeff_interaction >= 0.01, round( pvalue_coeff_interaction, 2), pvalue_coeff_interaction)
   list("pvalue" = pvalue_coeff_interaction, 
@@
            format.pval(vec$pvalue,
-                       eps = 10**-5),
+                       digits = 2,
+                       eps = 10**-5, 
+                       nsmall = 2),
            ")")
@@
   unlist()
+
+### change p-value round
 
@@
             inherit.aes = FALSE) +
-  labs(title = "Interrupted Time Series Analysis: Percentage of patients with psychiatric conditions",
-       subtitle = "Parametric bootstrap estimation: (10000 iterations)",
+  labs(
+    #title = "Interrupted Time Series Analysis: Percentage of patients with psychiatric conditions",
+    #subtitle = "Parametric bootstrap estimation: (10000 iterations)",
        x = "Calendar dates",
@@
             inherit.aes = FALSE) +
-  labs(title = "Interrupted Time Series Analysis: Percentage of patients with psychiatric conditions",
-       subtitle = "Parametric bootstrap estimation: (10000 iterations)",
+  labs(
+    #title = "Interrupted Time Series Analysis: Percentage of patients with psychiatric conditions",
+    # subtitle = "Parametric bootstrap estimation: (10000 iterations)",
        x = "Calendar dates",

### 2022-05-10 metanalysis_visualization/manuscript_Figures.Rmd
forest plot update label and change the last line

@@
                 header="Sites",
-                mlab="Overall periods' slope differences")
+                xlab = "Mean of the coefficients",
+                mlab="Overall effect")
 dev.off()
 #ggsave("figure2A_metaanalysis_persite.png", height = 20, width = 45, units = "cm")
+
+forest.meta( x = model_meta_siteid)
 
@@
        header = "Countries",
-       mlab = "Overall periods' slope differences")
+       xlab = "Mean of the coefficients",
+       mlab="Overall effect")
 dev.off()

### 2022-05-11 metanalysis_visualization/manuscript_Figures.Rmd
code added to plot sex ITS in the same way than the rest of ITS plots

@@
 ```
+
+### Alternative plot
+
+```{r annotation boostrap disorders}
+names_sex <- unique(df_coefficients_boot_sex$sex)
+
+plot_annotations_sex <- df_coefficients_boot_sex %>% 
+  group_by(sex) %>% 
+  group_map( bootstrap_pvalue_calculation ) %>%
+  set_names(names_sex) %>% 
+  lapply(FUN = function(vec) {
+    paste0("Coeffs differences: ",
+           vec$observed_diff,
+           " (pvalue: ",
+           format.pval(vec$pvalue, 
+                       eps = 10**-5),
+           ")")
+  }) %>% 
+  unlist()
+plot_annotations_sex_df <- data.frame(
+  sex = names(plot_annotations_sex), 
+  label = plot_annotations_sex
+)
+```
+
+```{r}
+bootstrapped_fitted_sex %>% 
+  mutate(period = recode_factor(period, "before_pandemic" = "Before pandemic onset", "during_pandemic" = "Since pandemic onset")) %>% 
+  ggplot(aes(x = time_p, y = predicted, group = interaction(iteration, period))) +
+  facet_wrap("sex",
+             scales = "free_y",
+             labeller = label_wrap_gen(width = 35,multi_line = TRUE)) +
+  geom_line(aes(color = period),
+            alpha = 0.03) +
+  geom_line(data = quantile_bootstrapped_fitted_sex,
+            aes(y = value, group = interaction(quantiles, period)),
+            linetype = "dashed") +
+  geom_line(data = quantile_bootstrapped_fitted_sex[quantile_bootstrapped_fitted_sex$quantiles == "50%", ],
+            aes(y = value, group = period)) +
+  geom_point(data = patients_hospitalized_agg_sex_total,
+             aes(y = percentage, group=NULL),
+             alpha = 0.5) +
+  scale_colour_manual(values = cbPalette) +
+  scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0))+
+  guides(color = guide_legend(override.aes = list(alpha = 1))) +
+  geom_label(data = plot_annotations_sex_df, 
+             aes(label=label), 
+            x = -Inf,
+            y = +Inf,
+            hjust = 0,
+            vjust = 1,
+            size = 4,
+            inherit.aes = FALSE) +
+  labs(x = "Calendar dates",
+       y = "Percentage",
+       color = "Period") +
+  theme(strip.text = element_text(size = 10),
+        axis.text.x=element_text(angle = 45, hjust=1, size = 8),
+        axis.text.y = element_text(size = 8), 
+        axis.title.x = element_text(size = 8),
+        axis.title.y = element_text(size = 8),
+        title = element_text(size = 14),
+        plot.subtitle = element_text(size = 8, face = "italic"), 
+        legend.position = "bottom") 
+ggsave("figure4_boostrapped_sex.png", height = 15, width = 25, units = "cm")
+```
+
+#### Oiriginal plot
 

### 2022-05-11 metanalysis_visualization/manuscript_Figures.Rmd
add forest plot with the meta package, adding the function to format it for JAMA

@@
 dev.off()
+```
+
+## Forest plot meta library
+
+### By site
+```{r}
+m.siteid <- metagen(TE = observed,
+                 seTE = var_boot,
+                 data = trends_siteid,
+                 studlab = plotsLabel,
+                 sm = "MN",
+                 fixed = FALSE,
+                 random = TRUE,
+                 method.tau = "DL")
+
+summary(m.siteid)
+
+forest.meta(m.siteid, 
+            sortvar = TE,
+            print.tau2 = FALSE,
+            leftlabs = c("Country"))
+
+forest.meta(m.siteid, layout = "JAMA")
+```
+
+### By country
+```{r}
+m.country <- metagen(TE = pred,
+                 seTE = se**2,
+                 data = trends_country,
+                 studlab = country,
+                 sm = "MN",
+                 fixed = FALSE,
+                 random = TRUE,
+                 method.tau = "DL")
+
+summary(m.country)
+
+forest.meta(m.country, 
+            sortvar = TE,
+            print.tau2 = FALSE,
+            leftlabs = c("Country"))
+
+forest.meta(m.country, layout = "JAMA")
 ```

### 2022-05-12 metanalysis_visualization/manuscript_Figures.Rmd
issue on the meta random forest fix adding sqrt. results consistents accross methods

--- 2022-05-11 metanalysis_visualization/manuscript_Figures.Rmd
@@
 #ggsave("figure2A_metaanalysis_persite.png", height = 20, width = 45, units = "cm")
-
-forest.meta( x = model_meta_siteid)
-
 ```
@@
 dev.off()
+```
+
+## Forest plot meta library
+
+### By site
+```{r}
+m.siteid <- metagen(TE = observed,
+                 seTE = sqrt( var_boot ),
+                 data = trends_siteid,
+                 studlab = plotsLabel,
+                 sm = "MN",
+                 fixed = FALSE,
+                 random = TRUE,
+                 method.tau = "DL")
+
+summary(m.siteid)
+
+forest.meta(m.siteid, 
+            sortvar = TE,
+            print.tau2 = FALSE,
+            leftlabs = c("plotsLabel"))
+
+forest.meta(m.siteid, layout = "JAMA")
+```
+
+### By country
+```{r}
+m.country <- metagen(TE = pred,
+                 seTE = sqrt( se**2 ),
+                 data = trends_country,
+                 studlab = country,
+                 sm = "MN",
+                 fixed = FALSE,
+                 random = TRUE,
+                 method.tau = "DL")
+
+summary(m.country)
+
+forest.meta(m.country, 
+            sortvar = TE,
+            print.tau2 = FALSE,
+            leftlabs = c("Country"))
+
+forest.meta(m.country, layout = "JAMA")
 ```

### 2022-05-13 metanalysis_visualization/manuscript_Figures.Rmd
before after slope significance test implementations + reformating + CI table1 fixing

@@
     ratios_patients_with_without_psy_all <- ratios_patients_with_without_psy
-    bootstrapped_fitted_all <- bootstrapped_fitted_df
-    bootstrapped_coefficients_all <- bootstrapped_coefficients_df
     length_hospitalisation_values_all <- length_hospitalisation_values
@@
     ratios_patients_with_without_psy_all <- rbind( ratios_patients_with_without_psy_all, ratios_patients_with_without_psy)
-    bootstrapped_fitted_all <- rbind(bootstrapped_fitted_all, bootstrapped_fitted_df)
-    bootstrapped_coefficients_all <- rbind(bootstrapped_coefficients_all, bootstrapped_coefficients_df)
     length_hospitalisation_values_all <- tryCatch(
@@
        perc_disorder_group,ratios_patients_with_without_psy,
-       patient_count_psy_period_psy_clear, bootstrapped_coefficients_df,
-       bootstrapped_fitted_df, mental_codes_qc_output, table1, length_hospitalisation_values)
+       patient_count_psy_period_psy_clear, mental_codes_qc_output, table1, length_hospitalisation_values)
   }
@@
 
-bootstrapped_fitted_all <- filter_by_dates(bootstrapped_fitted_all, start_date_plots, end_date_plots)
 count_icd_all <- filter_by_dates(count_icd_all, start_date_plots, end_date_plots)
@@
 ```{r}
-bootstrapped_fitted_all <- left_join(bootstrapped_fitted_all, country_map, on = "site_id")
 count_icd_all <- left_join(count_icd_all, country_map, on = "site_id")
@@
 ```{r Extracting age information}
+
 age_table1_all <- table1_all %>% 
@@
   group_by(period) %>% 
-  summarise(conf_interval = gmodels::ci( mean_age ),
+  summarise(
     mean_age = round(sum(Total * mean_age) / sum(Total), 2), 
-            sd_age = round(sum(Total * sd_age) / sum(Total), 2), 
-            ci_age = paste0("[", round(as.numeric(conf_interval[2]), 2), " , ",  round(as.numeric(conf_interval[3]),2), "]")) %>% 
-  mutate(Age = paste0(as.character(mean_age), " (sd: ", as.character(sd_age), ")", as.character(ci_age))) %>% 
+    mean_sd_age = round(sum(Total * sd_age) / sum(Total), 2), 
+    ci_age = paste0(" [", round(mean_age - (qnorm(0.975) * mean_sd_age / sum(Total)), 5),
+                    " , ",
+                    round(mean_age + (qnorm(0.975) * mean_sd_age / sum(Total)), 5), "]")) %>% 
+  mutate(Age = paste0(as.character(mean_age), " (sd: ", as.character(mean_sd_age), ")", as.character(ci_age))) %>% 
   select(period, Age) %>% 
@@
 ```{r}
-# Recovering
-coefficient_slope_calculation <- function(df, ...) {
+# Estimating regression slopes from linear regression ('observed coefficients')
+coefficient_slope_calculation <- function(df, slope_name = "periodduring_pandemic:time", ...) {
   ts_linear_percent <- lm(percentage ~ period * time, df)
-  linear_model <- broom::tidy(ts_linear_percent) %>%
-    filter(term %in% c("periodduring_pandemic:time"))
+  broom::tidy(ts_linear_percent) %>%
+    filter(term %in% c(slope_name)) %>% 
+    select(term, estimate)
 }
@@
   group_by(siteid) %>% 
-  group_modify(coefficient_slope_calculation) %>% 
-  rename(observed_diff = estimate)
+  group_modify(coefficient_slope_calculation,
+               slope_name = c("time", "periodduring_pandemic:time")
+               ) %>% 
+  rename(observed_slope_coefficient = estimate)
 ```
@@
 # Call function from another file : bootstrap.R
-bootstrap_size <- 10000
+bootstrap_size <- 100
 binom_proportion <- function(n, prob, size) {
@@
 ```{r}
+bootstrap_pvalue_calculation <- function(df, slope_name, ...) {
+  z <- unique(df[["observed_slope_coefficient"]]) / sd(df[[slope_name]])
+  pvalue_coeff_interaction <- round(pnorm(abs(z), lower.tail = FALSE) * 2, 3)
+  pvalue_coeff_interaction <- ifelse( pvalue_coeff_interaction >= 0.01, round( pvalue_coeff_interaction, 2), pvalue_coeff_interaction)
+  list("pvalue" = pvalue_coeff_interaction,
+       "observed_slope_coefficient" = round(unique(df[["observed_slope_coefficient"]]), 2)
+  )
+}
+```
+
+#### Calculate bootstrap
+
+```{r}
 # Get count_icd_all in the same format that patient_count_psy_period_psy_all, because first n last months missing ... 
@@
 
-bootstrapped_coefficients_all <- lapply(bootstrapped, `[[`, 1) %>% bind_rows()
-bootstrapped_fitted_all <- lapply(bootstrapped, `[[`, 2) %>% bind_rows()
-```
-
-### Primary analysis
-
-```{r}
-bootstrap_pvalue_calculation <- function(df, ...) {
-  z <- unique(df[["observed_diff"]]) / sd(df[["periodduring_pandemic:time"]])
-  pvalue_coeff_interaction <- round(pnorm(abs(z), lower.tail = FALSE) * 2, 3)
-  pvalue_coeff_interaction <- ifelse( pvalue_coeff_interaction >= 0.01, round( pvalue_coeff_interaction, 2), pvalue_coeff_interaction)
-  list("pvalue" = pvalue_coeff_interaction, 
-       "observed_diff" = round(unique(df[["observed_diff"]]), 2)
-  ) 
-}
-
-```
-
-```{r}
+bootstrapped_coefficients_all <- lapply(bootstrapped, `[[`, 1) %>% 
+  bind_rows() %>% 
+  rename(before_slope = time, 
+         difference_slope = `periodduring_pandemic:time`) %>% 
+  mutate(during_slope = before_slope + difference_slope) %>% 
+  select(iteration, siteid, everything())
+
+bootstrapped_fitted_all <- lapply(bootstrapped, `[[`, 2) %>% 
+  bind_rows() %>%
+  left_join(country_map, by = "siteid")
+
 quantile_bootstrapped_fitted_all <- bootstrapped_fitted_all %>%
@@
   pivot_longer(cols = c("2.5%", "50%", "97.5%"),
-               names_to = "quantiles")
-```
-
-```{r}
-df_coefficients_boot <- bootstrapped_coefficients_all %>% 
-  left_join(coefficients_df, by = "siteid") 
-names_site_id <- unique(df_coefficients_boot$siteid)
-
-plot_annotations <- df_coefficients_boot %>% 
+               names_to = "quantiles") %>% 
+  left_join(country_map, by = "siteid")
+
+```
+
+### Primary analysis
+
+```{r}
+df_coefficients_boot_primary <- bootstrapped_coefficients_all %>% 
+  left_join(filter(coefficients_df, term == "periodduring_pandemic:time"),
+            by = "siteid")
+
+names_site_id <- unique(df_coefficients_boot_primary$siteid)
+
+plot_annotations_primary <- df_coefficients_boot_primary %>% 
   group_by(siteid) %>% 
-  group_map( bootstrap_pvalue_calculation ) %>% 
+  group_map( bootstrap_pvalue_calculation, slope_name = "difference_slope" ) %>% 
   set_names(names_site_id) %>% 
@@
     paste0("Coeffs differences: ",
-           vec$observed_diff,
+           vec$observed_slope_coefficient,
            " (pvalue: ",
@@
                        digits = 2,
-                       eps = 10**-5, 
+                       eps = 10**-3, 
                        nsmall = 2),
@@
 
-### change p-value round
-
-plot_annotations_df <- data.frame(
-  siteid = names(plot_annotations), 
-  label = plot_annotations
-)
-```
-
-```{r}
-bootstrapped_fitted_all <- bootstrapped_fitted_all %>% left_join(country_map, by = "siteid")
-
-quantile_bootstrapped_fitted_all <- quantile_bootstrapped_fitted_all %>% left_join(country_map, by = "siteid")
-```
-
-```{r}
-plot_annotations_df <-plot_annotations_df %>%
+plot_annotations_df_primary <- data.frame(
+  siteid = names(plot_annotations_primary), 
+  label = plot_annotations_primary
+) %>%
   left_join( country_map ) 
-
-ggplot(bootstrapped_fitted_all, aes(x = time_p, y = predicted, group = interaction(iteration, period))) +
+```
+
+```{r}
+ggplot(bootstrapped_fitted_all,
+       aes(x = time_p, y = predicted, group = interaction(iteration, period))) +
   facet_wrap("plotsLabel", scales = "free_y", labeller = label_wrap_gen(width = 35,multi_line = TRUE)) +
@@
   guides(color = guide_legend(override.aes = list(alpha = 1))) +
-  geom_label(data = plot_annotations_df, 
+  geom_label(data = plot_annotations_df_primary, 
              aes(label=label), 
@@
 
+## Secondary analysis : slopes before/during
+
+```{r}
+pvalues_boot_before <- bootstrapped_coefficients_all %>% 
+  left_join(filter(coefficients_df, term == "time"), by = "siteid") %>% 
+  group_by(siteid) %>% 
+  group_map( bootstrap_pvalue_calculation, slope_name = "before_slope" ) %>% 
+  setNames(unique(bootstrapped_coefficients_all$siteid)) %>% 
+  bind_rows(.id = "siteid")
+
+pvalues_boot_after <- bootstrapped_coefficients_all %>% 
+  left_join(coefficients_df %>%
+              pivot_wider(names_from = term,
+                          values_from = observed_slope_coefficient) %>% 
+              mutate(observed_slope_coefficient = time + `periodduring_pandemic:time`) %>% 
+              select(siteid, observed_slope_coefficient),
+            by = "siteid") %>% 
+  group_by(siteid) %>% 
+  group_map( bootstrap_pvalue_calculation, slope_name = "during_slope" ) %>% 
+  setNames(unique(bootstrapped_coefficients_all$siteid)) %>% 
+  bind_rows(.id = "siteid")
+
+```
+
 ## Meta-analysis of temporal trends
@@
 ```{r}
-trends_siteid <- df_coefficients_boot %>% 
+trends_siteid <- df_coefficients_boot_primary %>% 
   group_by(siteid) %>% 
-  summarise(mean_boot = mean(`periodduring_pandemic:time`),
-            observed = unique(observed_diff),
-            var_boot = var(`periodduring_pandemic:time`), 
+  summarise(mean_boot = mean(difference_slope),
+            observed = unique(observed_slope_coefficient),
+            var_boot = var(difference_slope), 
             z = mean_boot / sqrt(var_boot), 
@@
 m.country <- metagen(TE = pred,
-                 seTE = sqrt( se**2 ),
+                 seTE = se,
                  data = trends_country,
@@
 
-```{r Interrupted time series by sex}
+```{r}
 patients_hospitalized_agg_sex_total <- patients_hospitalized_agg_sex_total %>% 
@@
   select(time_p, period, sex, count_sex, percentage, siteid)
-# %>% 
-#   inner_join(country_map)
 ```
@@
   group_by(sex) %>% 
-  # group_by(plotsLabel, sex) %>% 
   mutate(month_year = format(time_p, "%m"),
@@
 
-bootstrapped_coefficients_sex <- lapply(bootstrapped_sex, `[[`, 1) %>% bind_rows()
+bootstrapped_coefficients_sex <- lapply(bootstrapped_sex, `[[`, 1) %>%
+  bind_rows() %>% 
+  rename(before_slope = time, 
+         difference_slope = `periodduring_pandemic:time`)
+
 bootstrapped_fitted_sex <- lapply(bootstrapped_sex, `[[`, 2) %>% bind_rows()
+
 quantile_bootstrapped_fitted_sex <- bootstrapped_fitted_sex %>%
-  # group_by(time_p, period, plotsLabel, sex) %>%
   group_by(time_p, period,  sex) %>%
@@
                names_to = "quantiles")
+
 coefficients_df_sex <- patients_hospitalized_agg_sex_total %>% 
   group_by(sex) %>% 
-  # group_by(plotsLabel, sex) %>% 
   mutate(month_year = format(time_p, "%m"),
          time = seq_along(time_p))  %>% 
-  group_modify(coefficient_slope_calculation) %>% 
-  rename(observed_diff = estimate)
+  group_modify(coefficient_slope_calculation, 
+               slope_name = "periodduring_pandemic:time") %>% 
+  rename(observed_slope_coefficient = estimate)
 
 df_coefficients_boot_sex <- bootstrapped_coefficients_sex %>% 
-  # left_join(coefficients_df_sex, by = c("plotsLabel", "sex"))
   left_join(coefficients_df_sex, by = c("sex"))
@@
 
-```{r annotation boostrap disorders}
+```{r }
 names_sex <- unique(df_coefficients_boot_sex$sex)
@@
   group_by(sex) %>% 
-  group_map( bootstrap_pvalue_calculation ) %>%
+  group_map( bootstrap_pvalue_calculation, slope_name = "difference_slope" ) %>%
   set_names(names_sex) %>% 
@@
     paste0("Coeffs differences: ",
-           vec$observed_diff,
+           vec$observed_slope_coefficient,
            " (pvalue: ",
            format.pval(vec$pvalue, 
-                       eps = 10**-5),
+                       eps = 10**-3),
            ")")
@@
 
-#### Oiriginal plot
+#### Original plot
 
@@
 
-pvalue_sex_slope_diff <- bootstrap_pvalue_calculation(  df_coefficients_boot_sex[df_coefficients_boot_sex$sex == "female", ]
+pvalue_sex_slope_diff <- bootstrap_pvalue_calculation(
+  df_coefficients_boot_sex[df_coefficients_boot_sex$sex == "female", ], 
+  slope_name = "difference_slope"
 )
@@
 plot_annotations_sex_df <- paste0("Female proportion period differences: ",
-                                  pvalue_sex_slope_diff$observed_diff,
+                                  pvalue_sex_slope_diff$difference_slope,
                                   " (pvalue: ",
                                   format.pval(pvalue_sex_slope_diff$pvalue,
-                                              eps = 10**-5),")") %>%
+                                              digits = 2,
+                                              eps = 10**-3, 
+                                              nsmall = 2),")") %>%
   data.frame(label = ., 
@@
 ```{r bootstrap_plot}
-ggplot(bootstrapped_fitted_sex, aes(x = time_p, y = predicted/100, group = interaction(iteration, period, sex))) +
+ggplot(bootstrapped_fitted_sex,
+       aes(x = time_p, y = predicted/100, group = interaction(iteration, period, sex))) +
   geom_line(aes(color = interaction(period, sex)),
@@
 
-bootstrapped_coefficients_disorder <- lapply(bootstrapped_disorders, `[[`, 1) %>% bind_rows()
-bootstrapped_fitted_disorder <- lapply(bootstrapped_disorders, `[[`, 2) %>% bind_rows()
+bootstrapped_coefficients_disorder <- lapply(bootstrapped_disorders, `[[`, 1) %>%
+  bind_rows() %>% 
+  rename(before_slope = time, 
+         difference_slope = `periodduring_pandemic:time`)
+
+bootstrapped_fitted_disorder <- lapply(bootstrapped_disorders, `[[`, 2) %>%
+  bind_rows()
+
 quantile_bootstrapped_fitted_disorder <- bootstrapped_fitted_disorder %>%
@@
          time = seq_along(time_p))  %>% 
-  group_modify(coefficient_slope_calculation) %>% 
-  rename(observed_diff = estimate)
+  group_modify(coefficient_slope_calculation, 
+               slope_name = "periodduring_pandemic:time"
+               ) %>% 
+  rename(observed_slope_coefficient = estimate)
 
@@
   group_by(disorder_group) %>% 
-  group_map( bootstrap_pvalue_calculation ) %>%
+  group_map( bootstrap_pvalue_calculation, slope_name = "difference_slope" ) %>%
   set_names(names_disorder) %>% 
@@
 
-### Disorder group 
 ```{r}

### 2022-05-13 metanalysis_visualization/manuscript_Figures.Rmd
bootstrap number changing

@@
 # Call function from another file : bootstrap.R
-bootstrap_size <- 100
+bootstrap_size <- 10000
 binom_proportion <- function(n, prob, size) {

### 2022-05-13 metanalysis_visualization/manuscript_Figures.Rmd
pvalues prsentation with a table

@@
   setNames(unique(bootstrapped_coefficients_all$siteid)) %>% 
-  bind_rows(.id = "siteid")
+  bind_rows(.id = "siteid") %>% 
+  mutate(string = paste0(observed_slope_coefficient, " (", pvalue, ")")) %>% 
+  mutate(string = paste0(observed_slope_coefficient, " (", format.pval(pvalue,
+                                                                       digits = 2,
+                                                                       eps = 10**-3, 
+                                                                       nsmall = 2),
+                         ")")) %>% 
+  select(siteid, string) %>% 
+  rename("Before pandemic trend (p-value)" = string)
 
@@
   setNames(unique(bootstrapped_coefficients_all$siteid)) %>% 
-  bind_rows(.id = "siteid")
+  bind_rows(.id = "siteid")%>% 
+  mutate(string = paste0(observed_slope_coefficient, " (", format.pval(pvalue,
+                                                                       digits = 2,
+                                                                       eps = 10**-3, 
+                                                                       nsmall = 2),
+                         ")")) %>% 
+  select(siteid, string) %>% 
+  rename("Since pandemic trend (p-value)" = string)
+
+pvalues_boot_difference <- bootstrapped_coefficients_all %>% 
+  left_join(filter(coefficients_df, term == "periodduring_pandemic:time"), by = "siteid") %>% 
+  group_by(siteid) %>% 
+  group_map( bootstrap_pvalue_calculation, slope_name = "difference_slope" ) %>% 
+  setNames(unique(bootstrapped_coefficients_all$siteid)) %>% 
+  bind_rows(.id = "siteid") %>% 
+  mutate(string = paste0(observed_slope_coefficient, " (", pvalue, ")")) %>% 
+  mutate(string = paste0(observed_slope_coefficient, " (", format.pval(pvalue,
+                                                                       digits = 2,
+                                                                       eps = 10**-3, 
+                                                                       nsmall = 2),
+                         ")")) %>% 
+  select(siteid, string) %>% 
+  rename("Before/since pandemic trend difference (p-value)" = string)
+
+inner_join(pvalues_boot_before, pvalues_boot_after) %>% 
+  inner_join(pvalues_boot_difference) %>% 
+  inner_join(country_map) %>% 
+  select(plotsLabel, everything(), -country, -siteid) %>% 
+  arrange(plotsLabel) 
+# %>% 
+#   flextable::flextable() %>% 
+#   flextable::autofit() %>% 
+#   flextable::save_as_docx(path = "trends_pval.docx")
 

### 2022-05-14 metanalysis_visualization/manuscript_Figures.Rmd
reformating bootstrap, adding bootstrap estimations for all perdiods and for disorders groups

--- 2022-05-13 metanalysis_visualization/manuscript_Figures.Rmd
@@
   group_by(period) %>% 
-  summarise(median_hospitalisation = mean(len_hospitalisation)) %>% 
+  summarise(median_hospitalisation = median(len_hospitalisation)) %>% 
   pivot_wider(names_from = period, values_from = median_hospitalisation) %>% 
@@
 
+### Calculate LR coefficients
+
 ```{r}
@@
 
-coefficients_df <- patient_count_psy_period_psy_all %>% 
+observed_coefficients_df <- patient_count_psy_period_psy_all %>% 
   group_by(siteid) %>% 
@@
                ) %>% 
-  rename(observed_slope_coefficient = estimate)
-```
-
-## Define bootstrap functions
+  rename(observed_slope_coefficient = estimate) %>% 
+  pivot_wider(names_from = term,
+              values_from = observed_slope_coefficient) %>% 
+  rename(before_slope = time, 
+         difference_slope = `periodduring_pandemic:time`) %>% 
+  mutate(during_slope = before_slope + difference_slope) %>% 
+  pivot_longer(cols = c(before_slope, difference_slope, during_slope), 
+               names_to = "term", 
+               values_to = "observed_slope_coefficient")
+```
+
+### Define bootstrap functions
 
@@
 ```{r}
-bootstrap_pvalue_calculation <- function(df, slope_name, ...) {
-  z <- unique(df[["observed_slope_coefficient"]]) / sd(df[[slope_name]])
+bootstrap_pvalue_calculation <- function(df, ...) {
+  z <- unique(df[["observed_slope_coefficient"]]) / sd(df[["bootstrapped_slope_coefficient"]])
   pvalue_coeff_interaction <- round(pnorm(abs(z), lower.tail = FALSE) * 2, 3)
   pvalue_coeff_interaction <- ifelse( pvalue_coeff_interaction >= 0.01, round( pvalue_coeff_interaction, 2), pvalue_coeff_interaction)
-  list("pvalue" = pvalue_coeff_interaction,
+  data.frame("pvalue" = pvalue_coeff_interaction,
        "observed_slope_coefficient" = round(unique(df[["observed_slope_coefficient"]]), 2)
@@
 
-#### Calculate bootstrap
+#### Calculate bootstrap: primary analysis
 
@@
   mutate(during_slope = before_slope + difference_slope) %>% 
-  select(iteration, siteid, everything())
+  select(iteration, siteid, everything()) %>% 
+    pivot_longer(cols = c(before_slope, difference_slope, during_slope), 
+               names_to = "term", 
+               values_to = "bootstrapped_slope_coefficient") %>% 
+  left_join(observed_coefficients_df, by = c("siteid", "term"))
+
+pvalues_boot <- bootstrapped_coefficients_all %>% 
+  group_by(siteid, term) %>% 
+  group_modify( bootstrap_pvalue_calculation, .keep = TRUE) %>% 
+  mutate(label = paste0("Coeffs differences: ",
+                        observed_slope_coefficient,
+                        " (pvalue: ",
+                        format.pval(pvalue, digits = 2, eps = 10**-2), ")")
+  )
+
+ci_boot <- bootstrapped_coefficients_all %>% 
+  group_by(siteid, term) %>% 
+  summarise("2.5%" = quantile(bootstrapped_slope_coefficient, 0.025), 
+            "50%" = quantile(bootstrapped_slope_coefficient, 0.50), 
+            "97.5%" = quantile(bootstrapped_slope_coefficient, 0.975)) %>% 
+  mutate(ci = paste0("[", as.character(round(`2.5%`, 2)), "; ", as.character(round(`97.5%`, 2)), "]"))
 
@@
 
-### Primary analysis
-
-```{r}
-df_coefficients_boot_primary <- bootstrapped_coefficients_all %>% 
-  left_join(filter(coefficients_df, term == "periodduring_pandemic:time"),
-            by = "siteid")
-
-names_site_id <- unique(df_coefficients_boot_primary$siteid)
-
-plot_annotations_primary <- df_coefficients_boot_primary %>% 
-  group_by(siteid) %>% 
-  group_map( bootstrap_pvalue_calculation, slope_name = "difference_slope" ) %>% 
-  set_names(names_site_id) %>% 
-  lapply(FUN = function(vec) {
-    paste0("Coeffs differences: ",
-           vec$observed_slope_coefficient,
-           " (pvalue: ",
-           format.pval(vec$pvalue,
-                       digits = 2,
-                       eps = 10**-3, 
-                       nsmall = 2),
-           ")")
-  }) %>% 
-  unlist()
-
-plot_annotations_df_primary <- data.frame(
-  siteid = names(plot_annotations_primary), 
-  label = plot_annotations_primary
-) %>%
-  left_join( country_map ) 
+#### Trend differences
+
+```{r}
+
+plot_annotations_df_primary <- pvalues_boot %>%
+  select(siteid, label) %>% 
+  filter(term == "difference_slope") %>% 
+  left_join(country_map)
 ```
@@
 
-## Secondary analysis : slopes before/during
-
-```{r}
-pvalues_boot_before <- bootstrapped_coefficients_all %>% 
-  left_join(filter(coefficients_df, term == "time"), by = "siteid") %>% 
+## Table with bootstrapped coefficients, and pvalues
+
+```{r}
+pvalues_boot %>% 
+  inner_join(ci_boot) %>% 
+  inner_join(country_map) %>% 
+  arrange(plotsLabel) %>% 
+  mutate(string = paste0(observed_slope_coefficient, " ", ci, " (", format.pval(pvalue, digits = 3, eps = 10**-2), ")")) %>% 
+  select(plotsLabel, string) %>% 
+  pivot_wider(names_from = term, values_from = string) %>% 
+  select(plotsLabel, before_slope, during_slope, difference_slope) %>% 
+  rename("Before pandemic trend [95% CI] (p-value)" = before_slope,
+        "Since pandemic trend [95% CI] (p-value)" = during_slope,
+        "Difference before/since pandemic trend [95% CI] (p-value)" = difference_slope)
+
+```
+
+## Meta-analysis of temporal trends
+
+```{r}
+trends_siteid <- bootstrapped_coefficients_all %>% 
+  filter(term == "difference_slope") %>% 
   group_by(siteid) %>% 
-  group_map( bootstrap_pvalue_calculation, slope_name = "before_slope" ) %>% 
-  setNames(unique(bootstrapped_coefficients_all$siteid)) %>% 
-  bind_rows(.id = "siteid")
-
-pvalues_boot_after <- bootstrapped_coefficients_all %>% 
-  left_join(coefficients_df %>%
-              pivot_wider(names_from = term,
-                          values_from = observed_slope_coefficient) %>% 
-              mutate(observed_slope_coefficient = time + `periodduring_pandemic:time`) %>% 
-              select(siteid, observed_slope_coefficient),
-            by = "siteid") %>% 
-  group_by(siteid) %>% 
-  group_map( bootstrap_pvalue_calculation, slope_name = "during_slope" ) %>% 
-  setNames(unique(bootstrapped_coefficients_all$siteid)) %>% 
-  bind_rows(.id = "siteid")
-
-```
-
-## Meta-analysis of temporal trends
-
-```{r}
-trends_siteid <- df_coefficients_boot_primary %>% 
-  group_by(siteid) %>% 
-  summarise(mean_boot = mean(difference_slope),
+  summarise(mean_boot = mean(bootstrapped_slope_coefficient),
             observed = unique(observed_slope_coefficient),
-            var_boot = var(difference_slope), 
+            var_boot = var(bootstrapped_slope_coefficient), 
             z = mean_boot / sqrt(var_boot), 
@@
 
-## Forest plot meta library
-
-### By site
+### Meta analysis with another package
+
+#### Site
 ```{r}
@@
 
-### By country
+#### By country
 ```{r}
@@
 
-### Disorder group
+### Secondary analysis: :disorder group
 
@@
 
+disorders_to_select <- c(
+  "Anxiety Disorders", 
+  "Depressive Disorders", 
+  "Feeding and Eating Disorders", 
+  "Suicide or Self-Injury"
+)
+
 perc_disorder_group_total <- perc_disorder_group_complete %>%
@@
                          "before_pandemic", 
-                         "during_pandemic"))
-
-```
-
-#### Bootstrapping results by disorder groups
-
-```{r }
-disorders_to_select <- c(
-  "Anxiety Disorders", 
-  "Depressive Disorders", 
-  "Feeding and Eating Disorders", 
-  "Suicide or Self-Injury"
-)
-
-bootstrapped_disorders <- perc_disorder_group_total %>%
-  filter(disorder_group %in% disorders_to_select) %>% 
+                         "during_pandemic")) %>% 
+    filter(disorder_group %in% disorders_to_select)
+
+```
+
+#### Observed results: disorders
+
+```{r}
+observed_coefficients_df_disorder <- perc_disorder_group_total %>% 
   group_by(disorder_group) %>% 
@@
          time = seq_along(time_p))  %>% 
- group_map(~ bootstrapping_slopes(.x,
+  group_modify(coefficient_slope_calculation, 
+               slope_name = c("time", "periodduring_pandemic:time")
+               ) %>% 
+  rename(observed_slope_coefficient = estimate) %>% 
+  pivot_wider(names_from = term,
+              values_from = observed_slope_coefficient) %>% 
+  rename(before_slope = time, 
+         difference_slope = `periodduring_pandemic:time`) %>% 
+  mutate(during_slope = before_slope + difference_slope) %>% 
+  pivot_longer(cols = c(before_slope, difference_slope, during_slope), 
+               names_to = "term", 
+               values_to = "observed_slope_coefficient")
+
+```
+
+#### Calculate bootstrap: disorders groups
+
+```{r}
+# Get count_icd_all in the same format that patient_count_psy_period_psy_all, because first n last months missing ...
+
+bootstrapped_disorders <- perc_disorder_group_total %>%
+  group_by(disorder_group) %>% 
+  mutate(month_year = format(time_p, "%m"),
+         time = seq_along(time_p))  %>% 
+  group_map( ~bootstrapping_slopes(.x,
                                    group_name = .y,
@@
                                    percentage = "percentage"))
-
-bootstrapped_coefficients_disorder <- lapply(bootstrapped_disorders, `[[`, 1) %>%
+names(bootstrapped_disorders) <- unique(perc_disorder_group_total$disorder_group)
+
+bootstrapped_coefficients_disorders <- lapply(bootstrapped_disorders, `[[`, 1) %>% 
   bind_rows() %>% 
   rename(before_slope = time, 
-         difference_slope = `periodduring_pandemic:time`)
+         difference_slope = `periodduring_pandemic:time`) %>% 
+  mutate(during_slope = before_slope + difference_slope) %>% 
+  select(iteration, disorder_group, everything()) %>% 
+    pivot_longer(cols = c(before_slope, difference_slope, during_slope), 
+               names_to = "term", 
+               values_to = "bootstrapped_slope_coefficient") %>% 
+  left_join(observed_coefficients_df_disorder, by = c("disorder_group", "term"))
+
+pvalues_boot_disorders <- bootstrapped_coefficients_disorders %>% 
+  group_by(disorder_group, term) %>% 
+  group_modify( bootstrap_pvalue_calculation, .keep = TRUE) %>% 
+  mutate(label = paste0("Coeffs differences: ",
+                        observed_slope_coefficient,
+                        " (pvalue: ",
+                        format.pval(pvalue, digits = 2, eps = 10**-2), ")")
+  )
+
+ci_boot_disorders <- bootstrapped_coefficients_disorders %>% 
+  group_by(disorder_group, term) %>% 
+  summarise("2.5%" = quantile(bootstrapped_slope_coefficient, 0.025), 
+            "50%" = quantile(bootstrapped_slope_coefficient, 0.50), 
+            "97.5%" = quantile(bootstrapped_slope_coefficient, 0.975)) %>% 
+  mutate(ci = paste0("[", as.character(round(`2.5%`, 2)), "; ", as.character(round(`97.5%`, 2)), "]"))
 
@@
 quantile_bootstrapped_fitted_disorder <- bootstrapped_fitted_disorder %>%
-  # group_by(time_p, period, plotsLabel, sex) %>%
   group_by(time_p, period, disorder_group) %>%
@@
                names_to = "quantiles")
-coefficients_df_disorder <- perc_disorder_group_total %>% 
-  group_by(disorder_group) %>% 
-  mutate(month_year = format(time_p, "%m"),
-         time = seq_along(time_p))  %>% 
-  group_modify(coefficient_slope_calculation, 
-               slope_name = "periodduring_pandemic:time"
-               ) %>% 
-  rename(observed_slope_coefficient = estimate)
-
-df_coefficients_boot_disorder <- bootstrapped_coefficients_disorder %>% 
-  left_join(coefficients_df_disorder, by = c("disorder_group"))
-```
-
-```{r annotation boostrap disorders}
-names_disorder <- unique(df_coefficients_boot_disorder$disorder_group)
-
-plot_annotations_disorder <- df_coefficients_boot_disorder %>% 
-  group_by(disorder_group) %>% 
-  group_map( bootstrap_pvalue_calculation, slope_name = "difference_slope" ) %>%
-  set_names(names_disorder) %>% 
-  lapply(FUN = function(vec) {
-    paste0("Coeffs differences: ",
-           vec$observed_slope_coefficient,
-           " (pvalue: ",
-           format.pval(vec$pvalue, 
-                       eps = 10**-3),
-           ")")
-  }) %>% 
-  unlist()
-plot_annotations_disorder_df <- data.frame(
-  disorder_group = names(plot_annotations_disorder), 
-  label = plot_annotations_disorder
-)
-```
-
-```{r}
-# plot_annotations_df <- plot_annotations_df %>%
-#   left_join( country_map )
-
+
+```
+
+#### Table coefficients disorders
+
+```{r}
+pvalues_boot_disorders %>% 
+  inner_join(ci_boot_disorders) %>% 
+  arrange(disorder_group) %>% 
+  mutate(string = paste0(observed_slope_coefficient, " ", ci, " (", format.pval(pvalue, digits = 3, eps = 10**-2), ")")) %>% 
+  select(disorder_group, string) %>% 
+  pivot_wider(names_from = term, values_from = string) %>% 
+  select(disorder_group, before_slope, during_slope, difference_slope) %>% 
+  rename("Before pandemic trend [95% CI] (p-value)" = before_slope,
+        "Since pandemic trend [95% CI] (p-value)" = during_slope,
+        "Difference before/since pandemic trend [95% CI] (p-value)" = difference_slope) %>% 
+  flextable::flextable() %>% 
+  flextable::save_as_docx(path = "table_disorders.docx")
+
+```
+
+```{r}
+
+plot_annotations_disorder_df <-  pvalues_boot_disorders %>%
+  select(disorder_group, label) %>% 
+  filter(term == "difference_slope")
+```
+
+```{r}
 bootstrapped_fitted_disorder %>% 

### 2022-05-14 metanalysis_visualization/manuscript_Figures.Rmd
reformating bootstrap sex

@@
 }
-listOfPackages <- c("tidyverse", "tidyr","devtools","RColorBrewer", "knitr", "kableExtra", "tsModel", "gridExtra", "dplyr", "metafor", "meta", "viridis", "UpSetR", "scales", "EHRtemporalVariability", "plotly", "gmodels")
+listOfPackages <- c("tidyverse", "tidyr","devtools","RColorBrewer", "knitr", "kableExtra", "tsModel", "gridExtra", "dplyr", "metafor", "meta", "viridis", "UpSetR", "scales", "EHRtemporalVariability", "plotly", "gmodels", "flextable")
 paket(listOfPackages)
@@
 
-#### Trend differences
-
-```{r}
-
+#### Primary analysis plot
+
+```{r}
 plot_annotations_df_primary <- pvalues_boot %>%
@@
              percentage = 100* count_sex / total, 
-             siteid = "ALL")
-```
-
-#### Bootstrapping results by sex
-
-```{r}
-patients_hospitalized_agg_sex_total <- patients_hospitalized_agg_sex_total %>% 
+             siteid = "ALL") %>%  
   mutate(period = ifelse(time_p <= as.Date("2020-03-15"), "before_pandemic", "during_pandemic")) %>% 
@@
 ```
+
+#### Observed results: disorders
+
+```{r}
+observed_coefficients_df_sex <- patients_hospitalized_agg_sex_total %>% 
+  group_by(sex) %>% 
+  mutate(month_year = format(time_p, "%m"),
+         time = seq_along(time_p))  %>% 
+  group_modify(coefficient_slope_calculation, 
+               slope_name = c("time", "periodduring_pandemic:time")
+               ) %>% 
+  rename(observed_slope_coefficient = estimate) %>% 
+  pivot_wider(names_from = term,
+              values_from = observed_slope_coefficient) %>% 
+  rename(before_slope = time, 
+         difference_slope = `periodduring_pandemic:time`) %>% 
+  mutate(during_slope = before_slope + difference_slope) %>% 
+  pivot_longer(cols = c(before_slope, difference_slope, during_slope), 
+               names_to = "term", 
+               values_to = "observed_slope_coefficient")
+
+```
+
+#### Bootstrapping results by sex
 
@@
   rename(before_slope = time, 
-         difference_slope = `periodduring_pandemic:time`)
+         difference_slope = `periodduring_pandemic:time`) %>% 
+  mutate(during_slope = before_slope + difference_slope) %>% 
+  select(iteration, sex, everything()) %>% 
+  pivot_longer(cols = c(before_slope, difference_slope, during_slope), 
+               names_to = "term", 
+               values_to = "bootstrapped_slope_coefficient") %>% 
+  left_join(observed_coefficients_df_sex, by = c("sex", "term"))
+
+df_coefficients_boot_sex <- bootstrapped_coefficients_sex %>% 
+  left_join(coefficients_df_sex, by = c("sex"))
+
+pvalues_boot_sex <- bootstrapped_coefficients_sex %>% 
+  group_by(sex, term) %>% 
+  group_modify( bootstrap_pvalue_calculation, .keep = TRUE) %>% 
+  mutate(label = paste0("Female proportion period differences: ",
+                        observed_slope_coefficient,
+                        " (pvalue: ",
+                        format.pval(pvalue, digits = 2, eps = 10**-2), ")")
+  )
+
+ci_boot_sex <- bootstrapped_coefficients_sex %>% 
+  group_by(sex, term) %>% 
+  summarise("2.5%" = quantile(bootstrapped_slope_coefficient, 0.025), 
+            "50%" = quantile(bootstrapped_slope_coefficient, 0.50), 
+            "97.5%" = quantile(bootstrapped_slope_coefficient, 0.975)) %>% 
+  mutate(ci = paste0("[", as.character(round(`2.5%`, 2)), "; ", as.character(round(`97.5%`, 2)), "]"))
 
@@
 
-coefficients_df_sex <- patients_hospitalized_agg_sex_total %>% 
-  group_by(sex) %>% 
-  mutate(month_year = format(time_p, "%m"),
-         time = seq_along(time_p))  %>% 
-  group_modify(coefficient_slope_calculation, 
-               slope_name = "periodduring_pandemic:time") %>% 
-  rename(observed_slope_coefficient = estimate)
-
-df_coefficients_boot_sex <- bootstrapped_coefficients_sex %>% 
-  left_join(coefficients_df_sex, by = c("sex"))
-```
-
-### Alternative plot
-
-```{r }
-names_sex <- unique(df_coefficients_boot_sex$sex)
-
-plot_annotations_sex <- df_coefficients_boot_sex %>% 
-  group_by(sex) %>% 
-  group_map( bootstrap_pvalue_calculation, slope_name = "difference_slope" ) %>%
-  set_names(names_sex) %>% 
-  lapply(FUN = function(vec) {
-    paste0("Coeffs differences: ",
-           vec$observed_slope_coefficient,
-           " (pvalue: ",
-           format.pval(vec$pvalue, 
-                       eps = 10**-3),
-           ")")
-  }) %>% 
-  unlist()
-plot_annotations_sex_df <- data.frame(
-  sex = names(plot_annotations_sex), 
-  label = plot_annotations_sex
-)
-```
-
-```{r}
+```
+
+#### Table with bootstrapped coefficients, and pvalues
+
+```{r}
+pvalues_boot_sex %>% 
+  inner_join(ci_boot_sex) %>% 
+  arrange(sex) %>% 
+  mutate(string = paste0(observed_slope_coefficient, " ", ci, " (", format.pval(pvalue, digits = 3, eps = 10**-2), ")")) %>% 
+  select(sex, string) %>% 
+  pivot_wider(names_from = term, values_from = string) %>% 
+  select(sex, before_slope, during_slope, difference_slope) %>% 
+  rename("Before pandemic trend [95% CI] (p-value)" = before_slope,
+        "Since pandemic trend [95% CI] (p-value)" = during_slope,
+        "Difference before/since pandemic trend [95% CI] (p-value)" = difference_slope) %>% 
+  flextable::flextable() %>% 
+  flextable::save_as_docx(path = "table_sex.docx")
+
+```
+
+#### Original plot
+
+```{r}
+plot_annotations_sex_df <- pvalues_boot_sex %>%
+  select(term, sex, label) %>% 
+  filter(term == "difference_slope" & sex == "female") %>% 
+  mutate(plotsLabel = "Female")
+
+```
+
+```{r bootstrap_plot}
+ggplot(bootstrapped_fitted_sex,
+       aes(x = time_p, y = predicted/100, group = interaction(iteration, period, sex))) +
+  geom_line(aes(color = interaction(period, sex)),
+            alpha = 0.01) +
+  geom_line(data = quantile_bootstrapped_fitted_sex,
+            aes(y = value/100, group = interaction(quantiles, period, sex)),
+            linetype = "dashed") +
+  geom_line(data = quantile_bootstrapped_fitted_sex[quantile_bootstrapped_fitted_sex$quantiles == "50%", ],
+            aes(y = value/100, group = interaction(period, sex))) +
+  geom_point(data = patients_hospitalized_agg_sex_total,
+             aes(y = percentage/100, group=sex),
+             alpha = 1) +
+  # scale_color_brewer(palette = "Paired") +
+  scale_color_manual(labels = c("Female", "Male", "", ""),
+                     values = c("#A6CEE3", "#1F78B4", "#B2DF8A", "#33A02C")) +
+  scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0))+
+  guides(colour = guide_legend(
+    override.aes = list(alpha = 1,
+                        colour = c('#1F78B4', '#33A02C', NA, NA))) 
+  ) +
+  geom_label(data = plot_annotations_sex_df,
+             aes(label = label),
+             x = -Inf,
+             y = +Inf,
+             hjust = 0,
+             vjust = 1,
+             size = 4,
+             inherit.aes = FALSE) +
+  annotate("text_female", x = 7, y = 0.85, label = "Female proportion") + 
+  annotate("text_male", x = 7, y = 0.15, label = "Male proportion") + 
+  labs(title = "Interrupted Time Series Analysis: Patients sex proportion",
+       subtitle = "Parametric bootstrap estimation: (10000 iterations)",
+       x = "Calendar dates",
+       y = "Sex proportion",
+       color = "Sex") +
+  theme(strip.text = element_text(size = 10),
+        axis.text.x=element_text(angle = 45, hjust=1, size = 10),
+        axis.text.y = element_text(size = 10),
+        axis.title.x = element_text(size = 10),
+        axis.title.y = element_text(size = 10),
+        title = element_text(size = 14),
+  plot.subtitle = element_text(size = 12, face = "italic"))
+
+ggsave("figure4_ITS_sex.png", height = 15, width = 20, units = "cm")
+```
+
+#### Alternative plot 
+```{r, include=FALSE}
 bootstrapped_fitted_sex %>% 
@@
 ggsave("figure4_boostrapped_sex.png", height = 15, width = 25, units = "cm")
-```
-
-#### Original plot
-
-```{r annotation bootstrap disorder sex}
-names_sex <- unique(df_coefficients_boot_sex$sex)
-
-## Using only male, but female and male necessarily yield equal p-values
-
-pvalue_sex_slope_diff <- bootstrap_pvalue_calculation(
-  df_coefficients_boot_sex[df_coefficients_boot_sex$sex == "female", ], 
-  slope_name = "difference_slope"
-)
-
-plot_annotations_sex_df <- paste0("Female proportion period differences: ",
-                                  pvalue_sex_slope_diff$difference_slope,
-                                  " (pvalue: ",
-                                  format.pval(pvalue_sex_slope_diff$pvalue,
-                                              digits = 2,
-                                              eps = 10**-3, 
-                                              nsmall = 2),")") %>%
-  data.frame(label = ., 
-             plotsLabel = "Female")
-```
-
-```{r bootstrap_plot}
-ggplot(bootstrapped_fitted_sex,
-       aes(x = time_p, y = predicted/100, group = interaction(iteration, period, sex))) +
-  geom_line(aes(color = interaction(period, sex)),
-            alpha = 0.01) +
-  geom_line(data = quantile_bootstrapped_fitted_sex,
-            aes(y = value/100, group = interaction(quantiles, period, sex)),
-            linetype = "dashed") +
-  geom_line(data = quantile_bootstrapped_fitted_sex[quantile_bootstrapped_fitted_sex$quantiles == "50%", ],
-            aes(y = value/100, group = interaction(period, sex))) +
-  geom_point(data = patients_hospitalized_agg_sex_total,
-             aes(y = percentage/100, group=sex),
-             alpha = 1) +
-  # scale_color_brewer(palette = "Paired") +
-  scale_color_manual(labels = c("Female", "Male", "", ""),
-                     values = c("#A6CEE3", "#1F78B4", "#B2DF8A", "#33A02C")) +
-  scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0))+
-  guides(colour = guide_legend(
-    override.aes = list(alpha = 1,
-                        colour = c('#1F78B4', '#33A02C', NA, NA))) 
-  ) +
-  geom_label(data = plot_annotations_sex_df,
-             aes(label = label),
-             x = -Inf,
-             y = +Inf,
-             hjust = 0,
-             vjust = 1,
-             size = 4,
-             inherit.aes = FALSE) +
-  labs(title = "Interrupted Time Series Analysis: Patients sex proportion",
-       subtitle = "Parametric bootstrap estimation: (10000 iterations)",
-       x = "Calendar dates",
-       y = "Sex proportion",
-       color = "Sex") +
-  theme(strip.text = element_text(size = 10),
-        axis.text.x=element_text(angle = 45, hjust=1, size = 8),
-        axis.text.y = element_text(size = 8),
-        axis.title.x = element_text(size = 8),
-        axis.title.y = element_text(size = 8),
-        title = element_text(size = 14),
-  plot.subtitle = element_text(size = 12, face = "italic"))
-
-ggsave("figure4_ITS_sex.png", height = 20, width = 20, units = "cm")
 ```

### 2022-05-14 metanalysis_visualization/manuscript_Figures.Rmd
update sex proportion figure

@@
 df_coefficients_boot_sex <- bootstrapped_coefficients_sex %>% 
-  left_join(coefficients_df_sex, by = c("sex"))
+  left_join(observed_coefficients_df_sex, by = c("sex"))
 
@@
   group_modify( bootstrap_pvalue_calculation, .keep = TRUE) %>% 
-  mutate(label = paste0("Female proportion period differences: ",
+  mutate(label = paste0("Female proportion since pandemic coefficient: ",
                         observed_slope_coefficient,
@@
 ```{r bootstrap_plot}
-ggplot(bootstrapped_fitted_sex,
+plot_sex <- ggplot(bootstrapped_fitted_sex,
        aes(x = time_p, y = predicted/100, group = interaction(iteration, period, sex))) +
   geom_line(aes(color = interaction(period, sex)),
-            alpha = 0.01) +
+            alpha = 0.02) +
   geom_line(data = quantile_bootstrapped_fitted_sex,
@@
              inherit.aes = FALSE) +
-  annotate("text_female", x = 7, y = 0.85, label = "Female proportion") + 
-  annotate("text_male", x = 7, y = 0.15, label = "Male proportion") + 
-  labs(title = "Interrupted Time Series Analysis: Patients sex proportion",
+  annotate("label",
+           x = as.Date('2019-08-01'),
+           y = 0.69,
+           label = "Female proportion", 
+           fill = "#A6CEE3") +
+  annotate("label",
+           x = as.Date('2019-08-01'),
+           y = 0.3,
+           label = "Male proportion", 
+           fill = "#B2DF8A") +
+  # annotate("text", x = 7, y = 0.15, label = "Male proportion") +
+  labs(title = "Patients sex proportion trends among psychiatric-related visits",
        subtitle = "Parametric bootstrap estimation: (10000 iterations)",
@@
        color = "Sex") +
+  ylim(0.25, 0.75) + 
   theme(strip.text = element_text(size = 10),
@@
   plot.subtitle = element_text(size = 12, face = "italic"))
-
-ggsave("figure4_ITS_sex.png", height = 15, width = 20, units = "cm")
+plot_sex
+ggsave(plot = plot_sex, filename = "figure4_ITS_sex.png", height = 15, width = 20, units = "cm")
 ```

### 2022-05-16 metanalysis_visualization/manuscript_Figures.Rmd
coeff label changed to coefficient and feeding and eating disorders now eating disorder

@@
   group_modify( bootstrap_pvalue_calculation, .keep = TRUE) %>% 
-  mutate(label = paste0("Coeffs differences: ",
+  mutate(label = paste0("Coefficient difference: ",
                         observed_slope_coefficient,

### 2022-05-16 metanalysis_visualization/manuscript_Figures.Rmd
coeff label changed to coefficient and feeding and eating disorders now eating disorder
add manually annotation for the sex figure with two pannels, and comment some lines in the one plot only to try add more info to the labels

@@
   ) +
-  geom_label(data = plot_annotations_sex_df,
-             aes(label = label),
-             x = -Inf,
-             y = +Inf,
-             hjust = 0,
-             vjust = 1,
-             size = 4,
-             inherit.aes = FALSE) +
+  #geom_label(data = plot_annotations_sex_df,
+  #           aes(label = label),
+  #           x = -Inf,
+  #           y = +Inf,
+  #           hjust = 0,
+  #           vjust = 1,
+  #           size = 4,
+  #           inherit.aes = FALSE) +
   annotate("label",
-           x = as.Date('2019-08-01'),
+           x = as.Date('2019-10-01'),
            y = 0.69,
-           label = "Female proportion", 
+           #label = "Female proportion", 
+           label = "Female proportion [Coefficient difference: 0.77 (pvalue: <0.01)]", 
            fill = "#A6CEE3") +
   annotate("label",
-           x = as.Date('2019-08-01'),
+           x = as.Date('2019-10-01'),
            y = 0.3,
-           label = "Male proportion", 
+           #label = "Male proportion", 
+           label = "Male proportion [Coefficient difference: -0.77 (pvalue: 0.005)]", 
            fill = "#B2DF8A") +
   # annotate("text", x = 7, y = 0.15, label = "Male proportion") +
-  labs(title = "Patients sex proportion trends among psychiatric-related visits",
-       subtitle = "Parametric bootstrap estimation: (10000 iterations)",
+  labs(
+    #title = "Patients sex proportion trends among psychiatric-related visits",
+    #   subtitle = "Parametric bootstrap estimation: (10000 iterations)",
        x = "Calendar dates",
@@
         plot.subtitle = element_text(size = 12, face = "italic"))
+
 plot_sex
@@
 ```{r, include=FALSE}
+plot_annotations_sex_df2 <- data.frame(term=c("difference_solpe","difference_solpe"),
+                                       sex=c("female", "male"),
+                                       label=c("Coefficient difference: 0.77 (pvalue: <0.01)", "Coefficient difference: -0.77 (pvalue: 0.005)"),
+                                       plotsLabel=c("Female", "Male"))
+
 bootstrapped_fitted_sex %>% 
@@
   facet_wrap("sex",
-             scales = "free_y",
+             #scales = "free_y",
              labeller = label_wrap_gen(width = 35,multi_line = TRUE)) +
@@
   guides(color = guide_legend(override.aes = list(alpha = 1))) +
-  geom_label(data = plot_annotations_sex_df, 
+  geom_label(data = plot_annotations_sex_df2, 
              aes(label=label), 
@@
 
+perc_disorder_group_total$disorder_group <- gsub( "Feeding and Eating Disorders", "Eating Disorders", perc_disorder_group_total$disorder_group)
 ```

### 2022-05-17 metanalysis_visualization/manuscript_Figures.Rmd
modify the sex ITS plot to show only females. fix the eating disorder plot, not showing the dots after renaming the category

@@
 ```{r, include=FALSE}
-plot_annotations_sex_df2 <- data.frame(term=c("difference_solpe","difference_solpe"),
-                                       sex=c("female", "male"),
-                                       label=c("Coefficient difference: 0.77 (pvalue: <0.01)", "Coefficient difference: -0.77 (pvalue: 0.005)"),
-                                       plotsLabel=c("Female", "Male"))
+#plot_annotations_sex_df2 <- data.frame(term=c("difference_solpe","difference_solpe"),
+#                                       sex=c("female", "male"),
+#                                       label=c("Coefficient difference: 0.77 (pvalue: <0.01)", "Coefficient difference: -0.77 (pvalue: 0.005)"),
+#                                       plotsLabel=c("Female", "Male"))
 
 bootstrapped_fitted_sex %>% 
+  filter( sex == "female") %>%
   mutate(period = recode_factor(period, "before_pandemic" = "Before pandemic onset", "during_pandemic" = "Since pandemic onset")) %>% 
@@
   facet_wrap("sex",
-             #scales = "free_y",
              labeller = label_wrap_gen(width = 35,multi_line = TRUE)) +
@@
             alpha = 0.03) +
-  geom_line(data = quantile_bootstrapped_fitted_sex,
+  geom_line(data = quantile_bootstrapped_fitted_sex[ quantile_bootstrapped_fitted_sex$sex =="female", ],
             aes(y = value, group = interaction(quantiles, period)),
             linetype = "dashed") +
-  geom_line(data = quantile_bootstrapped_fitted_sex[quantile_bootstrapped_fitted_sex$quantiles == "50%", ],
+  geom_line(data = quantile_bootstrapped_fitted_sex[quantile_bootstrapped_fitted_sex$sex =="female" & quantile_bootstrapped_fitted_sex$quantiles == "50%", ],
             aes(y = value, group = period)) +
-  geom_point(data = patients_hospitalized_agg_sex_total,
+  geom_point(data = patients_hospitalized_agg_sex_total[ patients_hospitalized_agg_sex_total$sex =="female",],
              aes(y = percentage, group=NULL),
@@
   guides(color = guide_legend(override.aes = list(alpha = 1))) +
-  geom_label(data = plot_annotations_sex_df2, 
+  geom_label(data = plot_annotations_sex_df, 
              aes(label=label), 
@@
 perc_disorder_group_total$disorder_group <- gsub( "Feeding and Eating Disorders", "Eating Disorders", perc_disorder_group_total$disorder_group)
+
+disorders_to_select <- c(
+  "Anxiety Disorders", 
+  "Depressive Disorders", 
+  "Eating Disorders", 
+  "Suicide or Self-Injury"
+)
 ```

### 2022-05-17 metanalysis_visualization/manuscript_Figures.Rmd
chisq test for disorder group table 1

@@
   select( Category, name, before_pandemic_count, during_pandemic_count )
+
+### estimate thhe pvalue for the diosrder groups (chi.sq test)
+disorder_group_table_all <- table1_all %>%
+  filter( Category == "Disorder groups" & name != "Total") %>%
+  mutate( before_counts = sapply(strsplit( before_pandemic, " "), '[', 1), 
+          during_counts = sapply(strsplit( during_pandemic, " "), '[', 1)) %>%
+  group_by( name ) %>%
+  summarise( before_pandemic_yes = sum( as.numeric( before_counts ), na.rm = TRUE), 
+             during_pandemic_yes = sum( as.numeric( during_counts ), na.rm = TRUE), 
+             before_pandemic_no  = 9696-before_pandemic_yes, 
+             during_pandemic_no = 10561- during_pandemic_yes , 
+             p_value = chisq.test( matrix( c(before_pandemic_yes, during_pandemic_yes, before_pandemic_no,  during_pandemic_no), ncol = 2, nrow = 2))$p.value, 
+             p_value_formated = format.pval(p_value, digits=3, eps=0.001))
+
+p_value_breaks = c(-Inf, 0, 0.0001, 0.001, 0.01, 0.05, Inf)
+p_value_labels = c("", "****", "***", "**", "*", "")
+
+setDT(disorder_group_table_all)[, PVAL := cut(p_value,
+                                 breaks = p_value_breaks,
+                                 right = FALSE,
+                                 labels = p_value_labels)]
+#####
 

### 2022-05-19 metanalysis_visualization/manuscript_Figures.Rmd
adding total number of encounters and IQR calculations

@@
 }
-listOfPackages <- c("tidyverse", "tidyr","devtools","RColorBrewer", "knitr", "kableExtra", "tsModel", "gridExtra", "dplyr", "metafor", "meta", "viridis", "UpSetR", "scales", "EHRtemporalVariability", "plotly", "gmodels", "flextable")
+listOfPackages <- c("plotly", "tidyverse", "tidyr","devtools","RColorBrewer", "knitr", "kableExtra", "tsModel", "gridExtra", "dplyr", "metafor", "meta", "viridis", "UpSetR", "scales", "EHRtemporalVariability", "gmodels", "flextable", "data.table")
 paket(listOfPackages)
@@
 
+```{r}
+total_unique_encounters <- patient_count_psy_period_psy_all %>% 
+  group_by(period) %>% 
+  summarise(total_unique_encounters = sum(count_no_psy_patients + count_psy_patients))
+```
+
 ```{r Extracting age information}
@@
   group_by(period) %>% 
-  summarise(median_hospitalisation = median(len_hospitalisation)) %>% 
+  summarise(median_hospitalisation = paste0(median(len_hospitalisation), " (", IQR(len_hospitalisation), ")")) %>% 
   pivot_wider(names_from = period, values_from = median_hospitalisation) %>% 
@@
   pvalue_coeff_interaction <- ifelse( pvalue_coeff_interaction >= 0.01, round( pvalue_coeff_interaction, 2), pvalue_coeff_interaction)
-  data.frame("pvalue" = pvalue_coeff_interaction,
+  data.frame(
+    "pvalue" = pvalue_coeff_interaction,
     "observed_slope_coefficient" = round(unique(df[["observed_slope_coefficient"]]), 2)
@@
   geom_line(aes(color = interaction(period, sex)),
-            alpha = 0.02) +
+            alpha = 0.005) +
   geom_line(data = quantile_bootstrapped_fitted_sex,

### 2022-05-21 metanalysis_visualization/manuscript_Figures.Rmd
Tweaking sex ratio figure

@@
   group_modify( bootstrap_pvalue_calculation, .keep = TRUE) %>% 
-  mutate(label = paste0("Female proportion since pandemic coefficient: ",
+  mutate(label = paste0("Coefficient difference: ",
                         observed_slope_coefficient,
                         " (pvalue: ",
-                        format.pval(pvalue, digits = 2, eps = 10**-2), ")")
+                        format.pval(pvalue, digits = 3, eps = 10**-3), ")")
   )
@@
 ```
-
+↨
 #### Table with bootstrapped coefficients, and pvalues
@@
   arrange(sex) %>% 
-  mutate(string = paste0(observed_slope_coefficient, " ", ci, " (", format.pval(pvalue, digits = 3, eps = 10**-2), ")")) %>% 
+  mutate(string = paste0(observed_slope_coefficient, " ", ci, " (", format.pval(pvalue, digits = 3, eps = 10**-3), ")")) %>% 
   select(sex, string) %>% 
@@
 
-```{r bootstrap_plot}
+```{r }
+#plot_annotations_sex_df2 <- data.frame(term=c("difference_solpe","difference_solpe"),
+#                                       sex=c("female", "male"),
+#                                       label=c("Coefficient difference: 0.77 (pvalue: <0.01)", "Coefficient difference: -0.77 (pvalue: 0.005)"),
+#                                       plotsLabel=c("Female", "Male"))
+
+bootstrapped_fitted_sex %>% 
+  filter( sex == "female") %>%
+  mutate(period = recode_factor(period, "before_pandemic" = "Before pandemic onset", "during_pandemic" = "Since pandemic onset")) %>% 
+  ggplot(aes(x = time_p, y = predicted, group = interaction(iteration, period))) +
+  facet_wrap("sex",
+             labeller = labeller(sex = c("female" = "Female proportion "))) +
+  geom_line(aes(color = period),
+            alpha = 0.01) +
+  geom_line(data = quantile_bootstrapped_fitted_sex[ quantile_bootstrapped_fitted_sex$sex =="female", ],
+            aes(y = value, group = interaction(quantiles, period)),
+            linetype = "dashed") +
+  geom_line(data = quantile_bootstrapped_fitted_sex[quantile_bootstrapped_fitted_sex$sex =="female" & quantile_bootstrapped_fitted_sex$quantiles == "50%", ],
+            aes(y = value, group = period)) +
+  geom_point(data = patients_hospitalized_agg_sex_total[ patients_hospitalized_agg_sex_total$sex =="female",],
+             aes(y = percentage, group=NULL),
+             alpha = 0.5) +
+  scale_colour_manual(values = wesanderson::wes_palette("GrandBudapest1", 3)) +
+  scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0))+
+  guides(color = guide_legend(override.aes = list(alpha = 1))) +
+  geom_label(data = plot_annotations_sex_df, 
+             aes(label=label), 
+            x = -Inf,
+            y = +Inf,
+            hjust = 0,
+            vjust = 1,
+            size = 4,
+            inherit.aes = FALSE) +
+  labs(x = "Calendar dates",
+       y = "Percentage",
+       color = "Period") +
+  theme(strip.text = element_text(size = 10),
+        axis.text.x=element_text(angle = 45, hjust=1, size = 10),
+        axis.text.y = element_text(size = 10),
+        axis.title.x = element_text(size = 10),
+        axis.title.y = element_text(size = 10),
+        title = element_text(size = 14), 
+        plot.subtitle = element_text(size = 12, face = "italic"),
+        legend.position = "bottom") 
+ggsave("figure4_boostrapped_sex.png", height = 15, width = 20, units = "cm")
+```
+
+### Alternative plot
+
+```{r bootstrap_plot, include=FALSE}
 plot_sex <- ggplot(bootstrapped_fitted_sex,
@@
 
-#### Alternative plot 
-```{r, include=FALSE}
-#plot_annotations_sex_df2 <- data.frame(term=c("difference_solpe","difference_solpe"),
-#                                       sex=c("female", "male"),
-#                                       label=c("Coefficient difference: 0.77 (pvalue: <0.01)", "Coefficient difference: -0.77 (pvalue: 0.005)"),
-#                                       plotsLabel=c("Female", "Male"))
-
-bootstrapped_fitted_sex %>% 
-  filter( sex == "female") %>%
-  mutate(period = recode_factor(period, "before_pandemic" = "Before pandemic onset", "during_pandemic" = "Since pandemic onset")) %>% 
-  ggplot(aes(x = time_p, y = predicted, group = interaction(iteration, period))) +
-  facet_wrap("sex",
-             labeller = label_wrap_gen(width = 35,multi_line = TRUE)) +
-  geom_line(aes(color = period),
-            alpha = 0.03) +
-  geom_line(data = quantile_bootstrapped_fitted_sex[ quantile_bootstrapped_fitted_sex$sex =="female", ],
-            aes(y = value, group = interaction(quantiles, period)),
-            linetype = "dashed") +
-  geom_line(data = quantile_bootstrapped_fitted_sex[quantile_bootstrapped_fitted_sex$sex =="female" & quantile_bootstrapped_fitted_sex$quantiles == "50%", ],
-            aes(y = value, group = period)) +
-  geom_point(data = patients_hospitalized_agg_sex_total[ patients_hospitalized_agg_sex_total$sex =="female",],
-             aes(y = percentage, group=NULL),
-             alpha = 0.5) +
-  scale_colour_manual(values = cbPalette) +
-  scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0))+
-  guides(color = guide_legend(override.aes = list(alpha = 1))) +
-  geom_label(data = plot_annotations_sex_df, 
-             aes(label=label), 
-            x = -Inf,
-            y = +Inf,
-            hjust = 0,
-            vjust = 1,
-            size = 4,
-            inherit.aes = FALSE) +
-  labs(x = "Calendar dates",
-       y = "Percentage",
-       color = "Period") +
-  theme(strip.text = element_text(size = 10),
-        axis.text.x=element_text(angle = 45, hjust=1, size = 8),
-        axis.text.y = element_text(size = 8), 
-        axis.title.x = element_text(size = 8),
-        axis.title.y = element_text(size = 8),
-        title = element_text(size = 14),
-        plot.subtitle = element_text(size = 8, face = "italic"), 
-        legend.position = "bottom") 
-ggsave("figure4_boostrapped_sex.png", height = 15, width = 25, units = "cm")
-```
-
 ### Secondary analysis: :disorder group

### 2022-06-01 metanalysis_visualization/manuscript_Figures.Rmd
update lef pannel to % and reverse heatmap from high to low

@@
   group_modify( bootstrap_pvalue_calculation, .keep = TRUE) %>% 
-  mutate(label = paste0("Coefficient difference: ",
+  mutate(label = paste0("",
                         observed_slope_coefficient,
-                        " (pvalue: ",
+                        "% (p-value: ",
                         format.pval(pvalue, digits = 2, eps = 10**-2), ")")
@@
   group_modify( bootstrap_pvalue_calculation, .keep = TRUE) %>% 
-  mutate(label = paste0("Coefficient difference: ",
+  mutate(label = paste0("",
                         observed_slope_coefficient,
-                        " (pvalue: ",
+                        "% (p-value: ",
                         format.pval(pvalue, digits = 3, eps = 10**-3), ")")
@@
              alpha = 0.5) +
-  scale_colour_manual(values = wesanderson::wes_palette("GrandBudapest1", 3)) +
+  #scale_colour_manual(values = wesanderson::wes_palette("GrandBudapest1", 3)) +
+  scale_colour_manual(values = cbPalette) +
   scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0))+
@@
     title = " ",
+      yaxis = list(autorange="reversed"), 
     xaxis = list(
@@
 fig
+
 #plotly::orca(fig, file = "figure_supp3_counts_disorder.png")
@@
   facet_wrap( ~ disorder_group, scales = "free" ) +
-  labs(x ="site identifier", y = "ICD dictionary (number of ICD codes)") +
+  labs(x ="site identifier", y = "Number of ICD-10 codes") +
   scale_x_discrete(limits = positions) +
@@
 counts2 <- reshape2::melt(counts ,  id.vars = c('plotsLabel', 'time_p', 'time','month_year'), variable.name = 'group')
-counts2$group <- gsub("count_psy_patients", "patients with psychiatric conditions", counts2$group )
-counts2$group <- gsub("total_cases", "total number of patients", counts2$group )
+counts2$group <- gsub("count_psy_patients", "Patients with mental health disorder", counts2$group )
+counts2$group <- gsub("total_cases", "Total number of patients", counts2$group )
 
@@
 counts_agg2 <- reshape2::melt(counts_agg ,  id.vars = c('time_p', 'time','month_year'), variable.name = 'group')
-counts_agg2$group <- gsub("count_psy_patients", "patients with psychiatric conditions", counts_agg2$group )
-counts_agg2$group <- gsub("total_cases", "total number of patients", counts_agg2$group )
+counts_agg2$group <- gsub("count_psy_patients", "Patients with mental health disorder", counts_agg2$group )
+counts_agg2$group <- gsub("total_cases", "Total number of patients", counts_agg2$group )
 

### 2022-06-22 metanalysis_visualization/manuscript_Figures.Rmd
figures minor corrections for axis labels and format on the ITS plot

@@
             "97.5%" = quantile(bootstrapped_slope_coefficient, 0.975)) %>% 
-  mutate(ci = paste0("[", as.character(round(`2.5%`, 2)), "; ", as.character(round(`97.5%`, 2)), "]"))
+  mutate(ci = paste0("[95% CI ", as.character(round(`2.5%`, 2)), "–", as.character(round(`97.5%`, 2)), "]"))
 
@@
 
+## Table with bootstrapped coefficients, and pvalues
+
+```{r}
+pvalues_boot_complete <- pvalues_boot %>% 
+  inner_join(ci_boot) %>% 
+  inner_join(country_map) %>% 
+  arrange(plotsLabel) %>% 
+  mutate(string = paste0(observed_slope_coefficient, "% ", ci, " (p-value ", format.pval(pvalue, digits = 3, eps = 10**-2), ")")) %>% 
+  select(plotsLabel, string) %>% 
+  pivot_wider(names_from = term, values_from = string) %>% 
+  select(plotsLabel, before_slope, during_slope, difference_slope) %>% 
+  rename("Before pandemic trend [95% CI] (p-value)" = before_slope,
+        "Since pandemic trend [95% CI] (p-value)" = during_slope,
+        "Difference before/since pandemic trend [95% CI] (p-value)" = difference_slope)
+
+```
+
 #### Primary analysis plot
@@
 ```{r}
-plot_annotations_df_primary <- pvalues_boot %>%
-  select(siteid, label) %>% 
-  filter(term == "difference_slope") %>% 
+#plot_annotations_df_primary <- pvalues_boot %>%
+#  select(siteid, label) %>% 
+#  filter(term == "difference_slope") %>% 
+#  left_join(country_map)
+
+plot_annotations_df_primary <- pvalues_boot_complete %>%
+  select( siteid, label = `Difference before/since pandemic trend [95% CI] (p-value)`, plotsLabel )%>% 
   left_join(country_map)
+
 ```
@@
              alpha = 0.5) +
-  scale_colour_manual(values = cbPalette) +
+  geom_vline(xintercept = as.Date(pandemic_start_date), linetype = 6)+
   scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0))+
   guides(color = guide_legend(override.aes = list(alpha = 1))) +
+  scale_colour_manual(values = cbPalette, labels = c("pre-pandemic", "pandemic"), name = "Period") +
   geom_label(data = plot_annotations_df_primary, 
@@
         axis.title.y = element_text(size = 8),
+        panel.grid.minor.x = element_blank(), 
+        panel.grid.minor.y = element_blank(),
         title = element_text(size = 14),
@@
         legend.position="bottom")
+
 ggsave("figure1_boostrapped_all.png", height = 20, width = 45, units = "cm")
-```
-
-## Table with bootstrapped coefficients, and pvalues
-
-```{r}
-pvalues_boot %>% 
-  inner_join(ci_boot) %>% 
-  inner_join(country_map) %>% 
-  arrange(plotsLabel) %>% 
-  mutate(string = paste0(observed_slope_coefficient, " ", ci, " (", format.pval(pvalue, digits = 3, eps = 10**-2), ")")) %>% 
-  select(plotsLabel, string) %>% 
-  pivot_wider(names_from = term, values_from = string) %>% 
-  select(plotsLabel, before_slope, during_slope, difference_slope) %>% 
-  rename("Before pandemic trend [95% CI] (p-value)" = before_slope,
-        "Since pandemic trend [95% CI] (p-value)" = during_slope,
-        "Difference before/since pandemic trend [95% CI] (p-value)" = difference_slope)
-
 ```
@@
 ```{r}
+pvalues_boot_sex_complete <- pvalues_boot_sex %>% 
+  inner_join(ci_boot_sex) %>% 
+  arrange(sex) %>% 
+  mutate(string = paste0(observed_slope_coefficient, "% ", ci, " (p-value ", format.pval(pvalue, digits = 3, eps = 10**-2), ")")) %>% 
+  select(sex, string) %>% 
+  pivot_wider(names_from = term, values_from = string) %>% 
+  select(sex, before_slope, during_slope, difference_slope) %>% 
+  rename("Before pandemic trend [95% CI] (p-value)" = before_slope,
+        "Since pandemic trend [95% CI] (p-value)" = during_slope,
+        "Difference before/since pandemic trend [95% CI] (p-value)" = difference_slope)
+
 pvalues_boot_sex %>% 
@@
 ```{r}
-plot_annotations_sex_df <- pvalues_boot_sex %>%
-  select(term, sex, label) %>% 
-  filter(term == "difference_slope" & sex == "female") %>% 
-  mutate(plotsLabel = "Female")
+#plot_annotations_sex_df <- pvalues_boot_sex %>%
+#  select(term, sex, label) %>% 
+#  filter(term == "difference_slope" & sex == "female") %>% 
+#  mutate(plotsLabel = "Female")
+
+plot_annotations_sex_df <- pvalues_boot_sex_complete %>%
+  filter( sex == "female") %>% 
+    mutate(plotsLabel = "Female") %>%
+  select( label = `Difference before/since pandemic trend [95% CI] (p-value)`, plotsLabel )
 
@@
   guides(color = guide_legend(override.aes = list(alpha = 1))) +
+  geom_vline(xintercept = as.Date(pandemic_start_date), linetype = 6)+
+  scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0))+
+  guides(color = guide_legend(override.aes = list(alpha = 1))) +
+  scale_colour_manual(values = cbPalette, labels = c("pre-pandemic", "pandemic"), name = "Period") +
   geom_label(data = plot_annotations_sex_df, 
@@
         axis.title.y = element_text(size = 10),
+         panel.grid.minor.x = element_blank(), 
+        panel.grid.minor.y = element_blank(),
         title = element_text(size = 14), 
@@
   scale_fill_manual("Country of each site", values = c("USA" = "navy", "France" = "orange")) +
-  facet_wrap( ~ disorder_group, scales = "free" ) +
+  facet_wrap( ~ disorder_group, scales = "free" , ncol = 3) +
   labs(x ="site identifier", y = "Number of ICD-10 codes") +
   scale_x_discrete(limits = positions) +
-  theme(axis.text.x=element_text(angle=45,hjust=1, size = 6), 
-        strip.text = element_text(size = 7), 
+  theme(axis.text.x=element_text(angle=45,hjust=1, size = 11), 
+        strip.text = element_text(size = 13), 
+        axis.text.y = element_text(size = 11), 
+        axis.title.x = element_text(size = 11),
+        axis.title.y = element_text(size = 13), 
         panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
+        legend.title=element_text(size=13), 
+        legend.text=element_text(size=13), 
         legend.position="bottom") 
-ggsave("figure_supp2_ICDdictionaryQC.png", height = 20, width = 45, units = "cm")
+ggsave("figure_supp2_ICDdictionaryQC.png", height = 40, width = 45, units = "cm")
 
@@
   scale_color_manual( values = cbPalette ) + 
-  theme(axis.text.x=element_text(angle = 45, hjust=1, size = 6), 
-        strip.text = element_text(size = 10, face = "bold"),
+  theme(axis.text.x=element_text(angle = 45, hjust=1, size = 10), 
+        axis.text.y=element_text( size = 10),
+        axis.title.y = element_text( size = 9),
+        strip.text = element_text(size = 11, face = "bold"),
         panel.grid.major = element_blank(), panel.grid.minor = element_blank(), strip.background = element_blank(),
@@
 counts_plot
-ggsave("figure_supp1A_counts_per_site.png",counts_plot, height = 20, width = 32, units = "cm")
+ggsave("figure_supp1A_counts_per_site.png",counts_plot, height = 25, width = 40, units = "cm")
 

### 2022-06-23 metanalysis_visualization/manuscript_Figures.Rmd
close the gap in the ITS plots and format the top label with CI and without p-value

@@
             "97.5%" = quantile(bootstrapped_slope_coefficient, 0.975)) %>% 
-  mutate(ci = paste0("[95% CI ", as.character(round(`2.5%`, 2)), "–", as.character(round(`97.5%`, 2)), "]"))
+  mutate(ci = paste0("(95% CI ", as.character(format(round(`2.5%`, 2)), nsmall =2), "–", as.character(format(round(`97.5%`, 2)), nsmall=2), ")"))
 
@@
   arrange(plotsLabel) %>% 
-  mutate(string = paste0(observed_slope_coefficient, "% ", ci, " (p-value ", format.pval(pvalue, digits = 3, eps = 10**-2), ")")) %>% 
+  #mutate(string = paste0(observed_slope_coefficient, "% ", ci, " (p-value ", format.pval(pvalue, digits = 3, eps = 10**-2), ")")) %>%
+  mutate(string = paste0(observed_slope_coefficient, "% ", ci)) %>%
   select(plotsLabel, string) %>% 
@@
 ```{r}
-ggplot(bootstrapped_fitted_all,
-       aes(x = time_p, y = predicted, group = interaction(iteration, period))) +
+bootstrapped_fitted_all %>%
+  mutate( time_p = as.character(time_p),
+                  time_p = if_else(time_p == '2020-03-01', as.character('2020-03-15'), time_p),
+                  time_p = if_else(time_p == '2020-04-01', as.character('2020-03-16'), time_p),
+                  time_p = as.Date(time_p)) %>%
+  ggplot(aes(x = time_p, y = predicted, group = interaction(iteration, period))) +
   facet_wrap("plotsLabel", scales = "free_y", labeller = label_wrap_gen(width = 35,multi_line = TRUE)) +
@@
   filter( sex == "female") %>%
-  mutate(period = recode_factor(period, "before_pandemic" = "Before pandemic onset", "during_pandemic" = "Since pandemic onset")) %>% 
+  mutate(period = recode_factor(period, "before_pandemic" = "Before pandemic onset", "during_pandemic" = "Since pandemic onset"), 
+                  time_p = as.character(time_p),
+                  time_p = if_else(time_p == '2020-03-01', as.character('2020-03-15'), time_p),
+                  time_p = if_else(time_p == '2020-04-01', as.character('2020-03-16'), time_p),
+                  time_p = as.Date(time_p)) %>% 
   ggplot(aes(x = time_p, y = predicted, group = interaction(iteration, period))) +

### 2022-06-23 metanalysis_visualization/manuscript_Figures.Rmd
ITS plots edits added to fill the gap

@@
 bootstrapped_fitted_all %>%
-  mutate( time_p = as.character(time_p),
-                  time_p = if_else(time_p == '2020-03-01', as.character('2020-03-15'), time_p),
-                  time_p = if_else(time_p == '2020-04-01', as.character('2020-03-16'), time_p),
-                  time_p = as.Date(time_p)) %>%
+  mutate( time_p = as.Date(time_p)+ 15,
+          time_p = as.character(time_p),
+                  time_p = if_else(time_p == '2020-04-16', as.character('2020-04-01'), time_p), 
+                  time_p = if_else(time_p == '2020-03-16', as.character('2020-03-31'), time_p),
+                  time_p = if_else(time_p == '2021-04-16', as.character('2021-04-30'), time_p),
+                  time_p = if_else(time_p == '2019-02-16', as.character('2019-02-01'), time_p),
+                  time_p = as.Date(time_p))%>%
   ggplot(aes(x = time_p, y = predicted, group = interaction(iteration, period))) +
-  facet_wrap("plotsLabel", scales = "free_y", labeller = label_wrap_gen(width = 35,multi_line = TRUE)) +
+  facet_wrap("plotsLabel", scales = "free_y", labeller = label_wrap_gen(width = 35,multi_line = TRUE), ncol = 2) +
   geom_line(aes(color = period),
             alpha = 0.03) +
-  geom_line(data = quantile_bootstrapped_fitted_all,
+  geom_line(data = quantile_bootstrapped_fitted_all %>% mutate(time_p = as.Date(time_p)+ 15,  time_p = as.character(time_p),
+                  time_p = if_else(time_p == '2020-04-16', as.character('2020-04-01'), time_p), 
+                  time_p = if_else(time_p == '2020-03-16', as.character('2020-03-31'), time_p),
+                  time_p = if_else(time_p == '2021-04-16', as.character('2021-04-30'), time_p),
+                  time_p = if_else(time_p == '2019-02-16', as.character('2019-02-01'), time_p),
+                  time_p = as.Date(time_p)),
             aes(y = value, group = interaction(quantiles, period)),
             linetype = "dashed") +
-  geom_line(data = quantile_bootstrapped_fitted_all[quantile_bootstrapped_fitted_all$quantiles == "50%", ],
+  geom_line(data = quantile_bootstrapped_fitted_all[quantile_bootstrapped_fitted_all$quantiles == "50%", ] 
+            %>% mutate(time_p = as.Date(time_p)+ 15, 
+                       time_p = as.character(time_p),
+                  time_p = if_else(time_p == '2020-04-16', as.character('2020-04-01'), time_p), 
+                  time_p = if_else(time_p == '2020-03-16', as.character('2020-03-31'), time_p),
+                  time_p = if_else(time_p == '2021-04-16', as.character('2021-04-30'), time_p),
+                  time_p = if_else(time_p == '2019-02-16', as.character('2019-02-01'), time_p),
+                  time_p = as.Date(time_p)),
             aes(y = value, group = period)) +
-  geom_point(data = patient_count_psy_period_psy_all,
+  geom_point(data = patient_count_psy_period_psy_all %>% mutate( time_p = as.Date(time_p)+ 15),
              aes(y = percentage, group=NULL),
              alpha = 0.5) +
-  geom_vline(xintercept = as.Date(pandemic_start_date), linetype = 6)+
+  geom_vline(xintercept = as.Date("2020-04-01"), linetype = 6)+
   scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0))+
@@
         axis.text.y = element_text(size = 8), 
-        axis.title.x = element_text(size = 8),
+        axis.title.x =element_blank(),
         axis.title.y = element_text(size = 8),
@@
 
-ggsave("figure1_boostrapped_all.png", height = 20, width = 45, units = "cm")
+ggsave("figure1_boostrapped_all.png", height = 20, width = 25, units = "cm")
 ```
@@
             "97.5%" = quantile(bootstrapped_slope_coefficient, 0.975)) %>% 
-  mutate(ci = paste0("[95% CI ", as.character(round(`2.5%`, 2)), "–", as.character(round(`97.5%`, 2)), "]"))
+  mutate(ci = paste0("(95% CI ", as.character(round(`2.5%`, 2)), "–", as.character(round(`97.5%`, 2)), ")"))
 
@@
   arrange(sex) %>% 
-  mutate(string = paste0(observed_slope_coefficient, "% ", ci, " (p-value ", format.pval(pvalue, digits = 3, eps = 10**-2), ")")) %>% 
+  #mutate(string = paste0(observed_slope_coefficient, "% ", ci, " (p-value ", format.pval(pvalue, digits = 3, eps = 10**-2), ")")) %>% 
+  mutate(string = paste0(observed_slope_coefficient, "% ", ci)) %>% 
   select(sex, string) %>% 
@@
   filter( sex == "female") %>%
-  mutate(period = recode_factor(period, "before_pandemic" = "Before pandemic onset", "during_pandemic" = "Since pandemic onset"), 
+  mutate( time_p = as.Date(time_p)+ 15,
           time_p = as.character(time_p),
-                  time_p = if_else(time_p == '2020-03-01', as.character('2020-03-15'), time_p),
-                  time_p = if_else(time_p == '2020-04-01', as.character('2020-03-16'), time_p),
-                  time_p = as.Date(time_p)) %>% 
+                  time_p = if_else(time_p == '2020-04-16', as.character('2020-04-01'), time_p), 
+                  time_p = if_else(time_p == '2020-03-16', as.character('2020-03-31'), time_p),
+                  time_p = if_else(time_p == '2021-04-16', as.character('2021-04-30'), time_p),
+                  time_p = if_else(time_p == '2019-02-16', as.character('2019-02-01'), time_p),
+                  time_p = as.Date(time_p))%>%
   ggplot(aes(x = time_p, y = predicted, group = interaction(iteration, period))) +
@@
             alpha = 0.01) +
-  geom_line(data = quantile_bootstrapped_fitted_sex[ quantile_bootstrapped_fitted_sex$sex =="female", ],
+  geom_line(data = quantile_bootstrapped_fitted_sex[ quantile_bootstrapped_fitted_sex$sex =="female", ] %>% 
+              mutate(time_p = as.Date(time_p)+ 15,  time_p = as.character(time_p),
+                  time_p = if_else(time_p == '2020-04-16', as.character('2020-04-01'), time_p), 
+                  time_p = if_else(time_p == '2020-03-16', as.character('2020-03-31'), time_p),
+                  time_p = if_else(time_p == '2021-04-16', as.character('2021-04-30'), time_p),
+                  time_p = if_else(time_p == '2019-02-16', as.character('2019-02-01'), time_p),
+                  time_p = as.Date(time_p)),
             aes(y = value, group = interaction(quantiles, period)),
             linetype = "dashed") +
-  geom_line(data = quantile_bootstrapped_fitted_sex[quantile_bootstrapped_fitted_sex$sex =="female" & quantile_bootstrapped_fitted_sex$quantiles == "50%", ],
+  geom_line(data = quantile_bootstrapped_fitted_sex[quantile_bootstrapped_fitted_sex$sex =="female" & quantile_bootstrapped_fitted_sex$quantiles == "50%", ]%>% mutate(time_p = as.Date(time_p)+ 15,  time_p = as.character(time_p),
+                  time_p = if_else(time_p == '2020-04-16', as.character('2020-04-01'), time_p), 
+                  time_p = if_else(time_p == '2020-03-16', as.character('2020-03-31'), time_p),
+                  time_p = if_else(time_p == '2021-04-16', as.character('2021-04-30'), time_p),
+                  time_p = if_else(time_p == '2019-02-16', as.character('2019-02-01'), time_p),
+                  time_p = as.Date(time_p)),
             aes(y = value, group = period)) +
@@
   guides(color = guide_legend(override.aes = list(alpha = 1))) +
-  geom_vline(xintercept = as.Date(pandemic_start_date), linetype = 6)+
+  geom_vline(xintercept = as.Date("2020-04-01"), linetype = 6)+
   scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0))+
@@
         axis.text.y = element_text(size = 10),
-        axis.title.x = element_text(size = 10),
+        axis.title.x =element_blank(),
         axis.title.y = element_text(size = 10),
@@
 ```{r}
+pvalues_boot_disorders_complete <- pvalues_boot_disorders %>% 
+  inner_join(ci_boot_disorders) %>% 
+  arrange(disorder_group) %>% 
+  #mutate(string = paste0(observed_slope_coefficient, " ", ci, " (", format.pval(pvalue, digits = 3, eps = 10**-2), ")")) %>% 
+  mutate(string = paste0(observed_slope_coefficient, " ", ci)) %>% 
+  select(disorder_group, string) %>% 
+  pivot_wider(names_from = term, values_from = string) %>% 
+  select(disorder_group, before_slope, during_slope, difference_slope) %>% 
+  rename("Before pandemic trend [95% CI] (p-value)" = before_slope,
+        "Since pandemic trend [95% CI] (p-value)" = during_slope,
+        "Difference before/since pandemic trend [95% CI] (p-value)" = difference_slope)
+
 pvalues_boot_disorders %>% 
@@
   arrange(disorder_group) %>% 
-  mutate(string = paste0(observed_slope_coefficient, " ", ci, " (", format.pval(pvalue, digits = 3, eps = 10**-2), ")")) %>% 
+  #mutate(string = paste0(observed_slope_coefficient, " ", ci, " (", format.pval(pvalue, digits = 3, eps = 10**-2), ")")) %>% 
+  mutate(string = paste0(observed_slope_coefficient, " ", ci)) %>% 
   select(disorder_group, string) %>% 
@@
 ```{r}
-
-plot_annotations_disorder_df <-  pvalues_boot_disorders %>%
-  select(disorder_group, label) %>% 
-  filter(term == "difference_slope")
+plot_annotations_disorder_df <-  pvalues_boot_disorders_complete %>%
+  select( label = `Difference before/since pandemic trend [95% CI] (p-value)`, disorder_group )
 ```
@@
 bootstrapped_fitted_disorder %>% 
+  mutate( time_p = as.Date(time_p)+ 15,
+          time_p = as.character(time_p),
+                  time_p = if_else(time_p == '2020-04-16', as.character('2020-04-01'), time_p), 
+                  time_p = if_else(time_p == '2020-03-16', as.character('2020-03-31'), time_p),
+                  time_p = if_else(time_p == '2021-04-16', as.character('2021-04-30'), time_p),
+                  time_p = if_else(time_p == '2019-02-16', as.character('2019-02-01'), time_p),
+                  time_p = as.Date(time_p))%>%
   mutate(period = recode_factor(period, "before_pandemic" = "Before pandemic onset", "during_pandemic" = "Since pandemic onset")) %>% 
@@
             alpha = 0.03) +
-  geom_line(data = quantile_bootstrapped_fitted_disorder,
+  geom_line(data = quantile_bootstrapped_fitted_disorder %>% mutate(time_p = as.Date(time_p)+ 15,  time_p = as.character(time_p),
+                  time_p = if_else(time_p == '2020-04-16', as.character('2020-04-01'), time_p), 
+                  time_p = if_else(time_p == '2020-03-16', as.character('2020-03-31'), time_p),
+                  time_p = if_else(time_p == '2021-04-16', as.character('2021-04-30'), time_p),
+                  time_p = if_else(time_p == '2019-02-16', as.character('2019-02-01'), time_p),
+                  time_p = as.Date(time_p)),
             aes(y = value, group = interaction(quantiles, period)),
             linetype = "dashed") +
-  geom_line(data = quantile_bootstrapped_fitted_disorder[quantile_bootstrapped_fitted_disorder$quantiles == "50%", ],
+  geom_line(data = quantile_bootstrapped_fitted_disorder[quantile_bootstrapped_fitted_disorder$quantiles == "50%", ] %>%
+              mutate(time_p = as.Date(time_p)+ 15,  time_p = as.character(time_p),
+                  time_p = if_else(time_p == '2020-04-16', as.character('2020-04-01'), time_p), 
+                  time_p = if_else(time_p == '2020-03-16', as.character('2020-03-31'), time_p),
+                  time_p = if_else(time_p == '2021-04-16', as.character('2021-04-30'), time_p),
+                  time_p = if_else(time_p == '2019-02-16', as.character('2019-02-01'), time_p),
+                  time_p = as.Date(time_p)),
             aes(y = value, group = period)) +
@@
   scale_colour_manual(values = cbPalette) +
+  geom_vline(xintercept = as.Date("2020-04-01"), linetype = 6)+
   scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0))+

### 2022-07-07 metanalysis_visualization/manuscript_Figures.Rmd
ITS sex and disorder groups with % added and dot in the middle

@@
             aes(y = value, group = period)) +
-  geom_point(data = patients_hospitalized_agg_sex_total[ patients_hospitalized_agg_sex_total$sex =="female",],
+  geom_point(data = patients_hospitalized_agg_sex_total[ patients_hospitalized_agg_sex_total$sex =="female",]  
+             %>% mutate( time_p = as.Date(time_p)+ 15),
              aes(y = percentage, group=NULL),
@@
   #mutate(string = paste0(observed_slope_coefficient, " ", ci, " (", format.pval(pvalue, digits = 3, eps = 10**-2), ")")) %>% 
-  mutate(string = paste0(observed_slope_coefficient, " ", ci)) %>% 
+  mutate(string = paste0(observed_slope_coefficient, "% ", ci)) %>% 
   select(disorder_group, string) %>% 
@@
             aes(y = value, group = period)) +
-  geom_point(data = filter(perc_disorder_group_total,
+  geom_point(data = filter(perc_disorder_group_total  %>% mutate( time_p = as.Date(time_p)+ 15),
                            disorder_group %in% disorders_to_select),
@@
              alpha = 0.5) +
-  scale_colour_manual(values = cbPalette) +
+  scale_colour_manual(values = cbPalette, labels = c("pre-pandemic", "pandemic"), name = "Period") +
   geom_vline(xintercept = as.Date("2020-04-01"), linetype = 6)+

### 2022-10-04 metanalysis_visualization/manuscript_Figures.Rmd
axis labels formatted for publication

@@
   geom_vline(xintercept = as.Date("2020-04-01"), linetype = 6)+
-  scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0))+
+  scale_x_date(breaks = "1 month", labels=date_format("%b %Y"), expand=c(0.03,0))+
   guides(color = guide_legend(override.aes = list(alpha = 1))) +
@@
   scale_colour_manual(values = cbPalette) +
-  scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0))+
+  scale_x_date(breaks = "1 month", labels=date_format("%b %Y"), expand=c(0.03,0))+
   guides(color = guide_legend(override.aes = list(alpha = 1))) +
   geom_vline(xintercept = as.Date("2020-04-01"), linetype = 6)+
-  scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0))+
+  scale_x_date(breaks = "1 month", labels=date_format("%b %Y"), expand=c(0.03,0))+
   guides(color = guide_legend(override.aes = list(alpha = 1))) +
@@
                      values = c("#A6CEE3", "#1F78B4", "#B2DF8A", "#33A02C")) +
-  scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0))+
+  scale_x_date(breaks = "1 month", labels=date_format("%b %Y"), expand=c(0.03,0))+
   guides(colour = guide_legend(
@@
   labs(x ="", y = "percentage of patients (%)")+
-  scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0))+
+  scale_x_date(breaks = "1 month", labels=date_format("%b %Y"), expand=c(0.03,0))+
   theme(axis.text.x=element_text(angle = 45, hjust=1, size = 6),
@@
   labs(x ="", y = "number of patients")+
-  scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0)) +
+  scale_x_date(breaks = "1 month", labels=date_format("%b %Y"), expand=c(0.03,0)) +
   scale_color_manual( values = cbPalette ) + 
@@
        y = "Percentage") +
-      scale_x_date(breaks = "1 month", labels=date_format("%b-%y"), expand=c(0.03,0)) +
+      scale_x_date(breaks = "1 month", labels=date_format("%b %Y"), expand=c(0.03,0)) +
      theme(strip.text = element_text(size = 10),

### 2022-10-04 metanalysis_visualization/time_series_analysis_plotting.R

@@ -1,44 +1,0 @@
-time_series_analysis_plotting <- function(model, df_count_psy, count_col, clearance_period = FALSE) {
-  output_model <- broom::tidy(model)
-  coeff_slope <- round(output_model[output_model$term == "periodduring_pandemic:time", "estimate"][[1]], 3)
-  pvalue_slope <- round(output_model[output_model$term == "periodduring_pandemic:time", "p.value"][[1]], 3)
-  annotation <- paste0("Lockdown slope: ", coeff_slope, " (pvalue: ", pvalue_slope, ")")
-  pred_model <- predict(model, df_count_psy, type = "response", interval = "confidence") %>% as.data.frame()
-  pred_model_b <- predict(model, mutate(df_count_psy, period = "before_pandemic"), type = "response", interval = "confidence") %>%
-    as.data.frame()
-  names(pred_model_b) <- paste0(names(pred_model_b), "_wo_intervention")
-  pred_model <- bind_cols(df_count_psy, pred_model, pred_model_b)
-  if (clearance_period) {
-    pred_model$clearance_period <- ifelse(pred_model$time_p < as.Date(cut(start_clear_period, breaks = time_period)) ,
-                                          "before_pandemic",
-                                          ifelse(pred_model$time_p > as.Date(cut(end_clear_period, breaks = time_period)),
-                                                 "during_pandemic",
-                                                 "during_clearance")
-    )
-    pred_model$lwr <- ifelse(pred_model$clearance_period == "during_clearance", NA, pred_model$lwr)
-    pred_model$upr <- ifelse(pred_model$clearance_period == "during_clearance", NA, pred_model$upr)
-    pred_model$fit_wo_intervention <- ifelse(pred_model$time_p >= max(pred_model$time_p[pred_model$clearance_period == "before pandemic"]),
-                                             pred_model$fit_wo_intervention,
-                                             NA)
-  } else {
-    pred_model$clearance_period <- pred_model$period
-  }
-  plot <- ggplot(pred_model, aes_string(x = "time_p", y = count_col)) +
-    geom_point(aes(colour = clearance_period)) +
-    geom_line(aes(x = time_p, y = fit_wo_intervention, colour = "Expected mean w/o pandemic"), linetype = "dashed") +
-    geom_line(aes(x = time_p, y = fit, colour = clearance_period)) +
-    geom_ribbon(aes(ymin = lwr, ymax = upr, colour = clearance_period), alpha = 0.3, colour = NA) +
[…]

### 2022-10-12 metanalysis_visualization/manuscript_Figures.Rmd
psy and no psy enconters count

--- 2022-06-23 metanalysis_visualization/manuscript_Figures.Rmd
@@
   group_by(period) %>% 
-  summarise(total_unique_encounters = sum(count_no_psy_patients + count_psy_patients))
+  summarise(total_unique_encounters = sum(count_no_psy_patients + count_psy_patients), 
+            psy = sum(count_psy_patients),
+            percent_psy = round(psy * 100 / total_unique_encounters, 2),
+            no_psy = sum(count_no_psy_patients), 
+            percent_no_psy = round(no_psy * 100 / total_unique_encounters, 2)
+  )
+print(total_unique_encounters)
+print(chisq.test(total_unique_encounters[, c("psy", "no_psy")]))
+
 ```
@@
          during_pandemic_count = during_pandemic)
+```
+
+```{r}
+z_age <- (14.55 - 14.65) / sqrt(1.86^2/9696 + 1.79^2/11101)
+pvalue_age <- 2*pnorm(z_age)
+
 ```

### 2022-10-17 metanalysis_visualization/manuscript_Figures.Rmd
removing redundant data frame: all_icd_all

@@
   if( i == 1 ){
-    count_icd_all <- count_icd
     patients_hospitalized_agg_sex_perc_all <- patients_hospitalized_agg_sex_perc
@@
   } else {
-    count_icd_all <- rbind(count_icd_all, count_icd)
     perc_disorder_group_all <- rbind(perc_disorder_group_all, perc_disorder_group)
@@
     )
-    rm(count_icd,patients_hospitalized_agg_sex_perc,patient_count_psy_period_psy,
+    rm(patients_hospitalized_agg_sex_perc,patient_count_psy_period_psy,
        perc_disorder_group,ratios_patients_with_without_psy,
@@
 
-count_icd_all <- filter_by_dates(count_icd_all, start_date_plots, end_date_plots)
 patient_count_psy_period_psy_all <- filter_by_dates(patient_count_psy_period_psy_all, start_date_plots, end_date_plots)
@@
 
+```{r}
+patient_count_psy_period_psy_all <- patient_count_psy_period_psy_all %>% 
+  mutate(count_icd = count_no_psy_patients + count_psy_patients)
+
+```
+
 # Add site complete name information
@@
 ```{r}
-count_icd_all <- left_join(count_icd_all, country_map, on = "site_id")
 patient_count_psy_period_psy_all <- left_join(patient_count_psy_period_psy_all, country_map, on = "site_id")
@@
   group_by(period) %>% 
-  summarise(median_hospitalisation = paste0(median(len_hospitalisation), " (", IQR(len_hospitalisation), ")")) %>% 
+  summarise(median_hospitalisation = paste0(median(len_hospitalisation),
+                                            " (",
+                                            IQR(len_hospitalisation), ")"), 
+            .groups = "drop") %>% 
   pivot_wider(names_from = period, values_from = median_hospitalisation) %>% 
@@
   summarise(mean_hospitalisation = round( mean(len_hospitalisation), 2), 
-            conf_interval = gmodels::ci( len_hospitalisation ), 
-            ci_length = paste0("[", round(as.numeric(conf_interval[2]), 2), " , ",  round(as.numeric(conf_interval[3]),2), "]")) %>%
+            conf_interval = gmodels::ci( len_hospitalisation), 
+            ci_length = paste0("[", round(as.numeric(conf_interval[2]), 2),
+                               " , ",
+                               round(as.numeric(conf_interval[3]),2), "]"), 
+            .groups = "drop"
+            ) %>%
   mutate(mean_hospitalisation = paste0(as.character(mean_hospitalisation), as.character(ci_length))) %>% 
@@
 ```{r}
-# Get count_icd_all in the same format that patient_count_psy_period_psy_all, because first n last months missing ... 
-count_icd_all_boot <- count_icd_all %>% 
-  group_by(siteid) %>% # necessary to generate time correctly 
-  rename(percentage = percentage_psy) %>% 
-  mutate(month_year = format(time_p, "%m"),
-         time = seq_along(time_p))
-
-bootstrapped <- count_icd_all_boot %>%
+
+bootstrapped <- patient_count_psy_period_psy_all %>%
   group_by(siteid) %>% 
   group_map( ~bootstrapping_slopes(.x, group_name = .y, bootstrap_size = bootstrap_size))
-names(bootstrapped) <- unique(count_icd_all_boot$siteid)
+names(bootstrapped) <- unique(patient_count_psy_period_psy_all$siteid)
 
@@
 quantile_bootstrapped_fitted_all <- bootstrapped_fitted_all %>%
-  group_by(time_p, period, siteid) %>%
+  group_by(time_p, time, period, siteid) %>%
   summarise("2.5%" = quantile(predicted, 0.025),
             "50%" = median(predicted, 0.5),
-            "97.5%" = quantile(predicted, 0.975)) %>%
+            "97.5%" = quantile(predicted, 0.975), 
+            .groups = "drop") %>%
   pivot_longer(cols = c("2.5%", "50%", "97.5%"),
@@
 pvalues_boot_complete <- pvalues_boot %>% 
-  inner_join(ci_boot) %>% 
-  inner_join(country_map) %>% 
+  inner_join(ci_boot, by = c("siteid", "term")) %>% 
+  inner_join(country_map, by = c("siteid")) %>% 
   arrange(plotsLabel) %>% 
@@
   mutate(string = paste0(observed_slope_coefficient, "% ", ci)) %>%
-  select(plotsLabel, string) %>% 
+  select(siteid, term, plotsLabel, string) %>% 
   pivot_wider(names_from = term, values_from = string) %>% 
-  select(plotsLabel, before_slope, during_slope, difference_slope) %>% 
+  select(siteid, plotsLabel, before_slope, during_slope, difference_slope) %>% 
   rename("Before pandemic trend [95% CI] (p-value)" = before_slope,
@@
 ```{r}
-#plot_annotations_df_primary <- pvalues_boot %>%
-#  select(siteid, label) %>% 
-#  filter(term == "difference_slope") %>% 
-#  left_join(country_map)
-
 plot_annotations_df_primary <- pvalues_boot_complete %>%
   select( siteid, label = `Difference before/since pandemic trend [95% CI] (p-value)`, plotsLabel )%>% 
-  left_join(country_map)
-
+  left_join(country_map, by = c("siteid", "plotsLabel"))
+```
+
+```{r, fig.width = 20, fig.height=14}
+bootstrapped_fitted_all %>%
+  # mutate( time = time + 0.5) %>%
+
+          # 
+          # ),
+          # time_p = as.character(time_p),
+          #         time_p = if_else(time_p == '2020-04-16', as.character('2020-04-01'), time_p), 
+          #         time_p = if_else(time_p == '2020-03-16', as.character('2020-03-31'), time_p),
+          #         time_p = if_else(time_p == '2021-04-16', as.character('2021-04-30'), time_p),
+          #         time_p = if_else(time_p == '2019-02-16', as.character('2019-02-01'), time_p),
+          #         time_p = as.Date(time_p))%>%
+  ggplot(aes(x = time, y = predicted, group = interaction(iteration, period))) +
+  facet_wrap("plotsLabel", scales = "free_y", labeller = label_wrap_gen(width = 35,multi_line = TRUE), ncol = 2) +
+  geom_line(aes(color = period),
+            alpha = 0.03) +
+  geom_line(data = quantile_bootstrapped_fitted_all,
+            # %>% mutate(time_p = as.Date(time_p)+ 15,  time_p = as.character(time_p),
+            #       time_p = if_else(time_p == '2020-04-16', as.character('2020-04-01'), time_p), 
+            #       time_p = if_else(time_p == '2020-03-16', as.character('2020-03-31'), time_p),
+            #       time_p = if_else(time_p == '2021-04-16', as.character('2021-04-30'), time_p),
+            #       time_p = if_else(time_p == '2019-02-16', as.character('2019-02-01'), time_p),
+            #       time_p = as.Date(time_p)),
+            aes(y = value, group = interaction(quantiles, period)),
+            linetype = "dashed") +
+  # geom_line(data = quantile_bootstrapped_fitted_all[quantile_bootstrapped_fitted_all$quantiles == "50%", ] 
+  #           %>% mutate(time_p = as.Date(time_p)+ 15, 
+  #                      time_p = as.character(time_p),
+  #                 time_p = if_else(time_p == '2020-04-16', as.character('2020-04-01'), time_p), 
+  #                 time_p = if_else(time_p == '2020-03-16', as.character('2020-03-31'), time_p),
+  #                 time_p = if_else(time_p == '2021-04-16', as.character('2021-04-30'), time_p),
+  #                 time_p = if_else(time_p == '2019-02-16', as.character('2019-02-01'), time_p),
+  #                 time_p = as.Date(time_p)),
+  #           aes(y = value, group = period)) +
+  geom_point(data = patient_count_psy_period_psy_all,
+             # %>% mutate( time_p = as.Date(time_p)+ 15),
+             aes(y = percentage, group=NULL),
+             alpha = 0.5) +
+  scale_x_continuous(breaks = scales::breaks_width(1, offset = 0)) + 
+  geom_vline(xintercept = as.Date("2020-04-01"), linetype = 6)+
+  scale_x_date(breaks = "1 month", labels=date_format("%b %Y"), expand=c(0.03,0))+
+  guides(color = guide_legend(override.aes = list(alpha = 1))) +
+  scale_colour_manual(values = cbPalette, labels = c("pre-pandemic", "pandemic"), name = "Period") +
+  geom_label(data = plot_annotations_df_primary, 
+             aes(label=label), 
+            x = -Inf,
+            y = +Inf,
+            hjust = 0,
+            vjust = 1,
+            size = 4,
+            inherit.aes = FALSE) +
+  labs(
+    #title = "Interrupted Time Series Analysis: Percentage of patients with psychiatric conditions",
+    #subtitle = "Parametric bootstrap estimation: (10000 iterations)",
+       x = "Calendar dates",
+       y = "Percentage",
+       color = "Period") +
+  theme(strip.text = element_text(size = 14),
+        axis.text.x=element_text(angle = 45, hjust=1, size = 8),
+        axis.text.y = element_text(size = 8), 
+        axis.title.x =element_blank(),
+        axis.title.y = element_text(size = 8),
+        panel.grid.minor.x = element_blank(), 
+        panel.grid.minor.y = element_blank(),
+        title = element_text(size = 14),
+        plot.subtitle = element_text(size = 12, face = "italic"), 
+        legend.position="bottom")
+
+# ggsave("figure1_boostrapped_all.png", height = 20, width = 25, units = "cm")
 ```
@@
 ```{r}
-# Get count_icd_all in the same format that patient_count_psy_period_psy_all, because first n last months missing ...
 
@@
 
-counts <- count_icd_all %>%
-  mutate( time = seq_along(time_p),
-         month_year = format(time_p, "%m")) %>%
-  select( plotsLabel, time_p, time, month_year, count_psy_patients = count_psy, total_cases = count_icd )
+counts <- patient_count_psy_period_psy_all %>%
+  select( plotsLabel, time_p, time, month_year, count_psy_patients, total_cases = count_icd )
 

### 2022-10-18 metanalysis_visualization/manuscript_Figures.Rmd
- Fixing plots: 	- Expanding predictions from bootstrap model 	- Adapting x-axis scale - Cleaning code to have consistent time/time_p throughout the file

@@
 ```{r}
+# To be able to have one single reference for the mapping between periods, months and months labels
+time_timep_df <- distinct(patient_count_psy_period_psy_all[, c("time", "time_p")])
+time_timep_list <- time_timep_df[["time_p"]]
+names(time_timep_list) <- as.character(time_timep_df[["time"]])
+time_timep_list[["29"]] <- as.Date("2021-05-01")
+```
+
+```{r}
 patient_count_psy_period_psy_all <- patient_count_psy_period_psy_all %>% 
   mutate(count_icd = count_no_psy_patients + count_psy_patients)
-
+patients_hospitalized_agg_sex_perc_all <- left_join(
+  patients_hospitalized_agg_sex_perc_all, 
+  time_timep_df, 
+  by = c("time_p"))
+perc_disorder_group_all <- left_join(perc_disorder_group_all, 
+                                     time_timep_df, 
+                                     by = c("time_p"))
 ```
@@
 ratios_patients_with_without_psy_all <- left_join(ratios_patients_with_without_psy_all, country_map, on = "site_id")
-
 ```
@@
 ```{r}
-# Call function from another file : bootstrap.R
-bootstrap_size <- 100
+bootstrap_size <- 10000
 binom_proportion <- function(n, prob, size) {
@@
                                time = df$time,
+                               time_p = df$time_p,
                                percentage = bootstrapped_timepoints[iteration, ])
@@
     bootstrapped_coefficients_list[[iteration]] <- coefficients
-    df_iteration$predicted <- predict(ts_linear_percent)
-    df_iteration$time_p <- df$time_p
+    # To be able to fill the gap in the subsequent graphs
+    df_extended_prediction <- data.frame(
+      period = c("before_pandemic", "before_pandemic", "during_pandemic", "during_pandemic"), 
+      time = c(1.5, 15.5, 15.5, 28.5), 
+      time_p = NA,
+      percentage = NA
+    )
+    df_iteration <- bind_rows(df_iteration, df_extended_prediction)
+    df_iteration$predicted <- predict(ts_linear_percent, newdata = df_iteration)
     bootstrapped_sample_list[[iteration]] <- df_iteration
@@
 
+```{r}
+# Function to change x-axis labels in graphs
+time_to_time_p <- function(break_ticks) {
+  break_ticks <- as.character(break_ticks)
+  sapply(break_ticks, function(break_tick) {
+   if (is.na(break_tick) | !break_tick %in% names(time_timep_list)) {
+     return(as.Date(NA_character_))
+   } else {
+     time_timep_list[[break_tick]]
+   }
+  }) %>%
+    as.Date(origin = "1970/01/01") %>% 
+    format(format = "%b %Y", tz = "UTC")
+}
+
+```
+
 ```{r, fig.width = 20, fig.height=14}
 bootstrapped_fitted_all %>%
-  # mutate( time = time + 0.5) %>%
-
-          # 
-          # ),
-          # time_p = as.character(time_p),
-          #         time_p = if_else(time_p == '2020-04-16', as.character('2020-04-01'), time_p), 
-          #         time_p = if_else(time_p == '2020-03-16', as.character('2020-03-31'), time_p),
-          #         time_p = if_else(time_p == '2021-04-16', as.character('2021-04-30'), time_p),
-          #         time_p = if_else(time_p == '2019-02-16', as.character('2019-02-01'), time_p),
-          #         time_p = as.Date(time_p))%>%
   ggplot(aes(x = time, y = predicted, group = interaction(iteration, period))) +
   facet_wrap("plotsLabel", scales = "free_y", labeller = label_wrap_gen(width = 35,multi_line = TRUE), ncol = 2) +
-  geom_line(aes(color = period),
-            alpha = 0.03) +
-  geom_line(data = quantile_bootstrapped_fitted_all,
-            # %>% mutate(time_p = as.Date(time_p)+ 15,  time_p = as.character(time_p),
-            #       time_p = if_else(time_p == '2020-04-16', as.character('2020-04-01'), time_p), 
-            #       time_p = if_else(time_p == '2020-03-16', as.character('2020-03-31'), time_p),
-            #       time_p = if_else(time_p == '2021-04-16', as.character('2021-04-30'), time_p),
-            #       time_p = if_else(time_p == '2019-02-16', as.character('2019-02-01'), time_p),
-            #       time_p = as.Date(time_p)),
-            aes(y = value, group = interaction(quantiles, period)),
+  geom_line(aes(x = time + 0.5, color = period), alpha = 0.03) +
+  geom_line(aes(x = time + 0.5, 
+                y = value,
+                group = interaction(quantiles, period)),
+            data = filter(quantile_bootstrapped_fitted_all, quantiles %in% c("2.5%", "97.5%")),
             linetype = "dashed") +
-  # geom_line(data = quantile_bootstrapped_fitted_all[quantile_bootstrapped_fitted_all$quantiles == "50%", ] 
-  #           %>% mutate(time_p = as.Date(time_p)+ 15, 
-  #                      time_p = as.character(time_p),
-  #                 time_p = if_else(time_p == '2020-04-16', as.character('2020-04-01'), time_p), 
-  #                 time_p = if_else(time_p == '2020-03-16', as.character('2020-03-31'), time_p),
-  #                 time_p = if_else(time_p == '2021-04-16', as.character('2021-04-30'), time_p),
-  #                 time_p = if_else(time_p == '2019-02-16', as.character('2019-02-01'), time_p),
-  #                 time_p = as.Date(time_p)),
-  #           aes(y = value, group = period)) +
-  geom_point(data = patient_count_psy_period_psy_all,
-             # %>% mutate( time_p = as.Date(time_p)+ 15),
-             aes(y = percentage, group=NULL),
+  geom_line(aes(x = time + 0.5, 
+                y = value,
+                group = period),
+            data = filter(quantile_bootstrapped_fitted_all, quantiles == "50%")) +
+  geom_point(aes(x = time + 0.5, y = percentage, group = NULL),
+             data = patient_count_psy_period_psy_all,
              alpha = 0.5) +
-  scale_x_continuous(breaks = scales::breaks_width(1, offset = 0)) + 
-  geom_vline(xintercept = as.Date("2020-04-01"), linetype = 6)+
-  scale_x_date(breaks = "1 month", labels=date_format("%b %Y"), expand=c(0.03,0))+
-  guides(color = guide_legend(override.aes = list(alpha = 1))) +
+  scale_x_continuous(breaks = breaks_width(1),
+                     labels = time_to_time_p, 
+                     limits = c(2, 29),
+                     expand = c(0.03,0)) + 
+  geom_vline(xintercept = 16, linetype = 6 ) +
+  theme(strip.text = element_text(size = 14),
+        axis.text.x=element_text(angle = 45, hjust=1, size = 8),
+        axis.text.y = element_text(size = 8), 
+        axis.title.x =element_blank(),
+        axis.title.y = element_text(size = 8),
+        panel.grid.minor.x = element_blank(), 
+        panel.grid.minor.y = element_blank(),
+        title = element_text(size = 14),
+        plot.subtitle = element_text(size = 12, face = "italic"), 
+        legend.position="bottom") + 
   scale_colour_manual(values = cbPalette, labels = c("pre-pandemic", "pandemic"), name = "Period") +
@@
              inherit.aes = FALSE) +
+  guides(color = guide_legend(override.aes = list(alpha = 1))) +
   labs(
@@
        y = "Percentage",
-       color = "Period") +
-  theme(strip.text = element_text(size = 14),
-        axis.text.x=element_text(angle = 45, hjust=1, size = 8),
-        axis.text.y = element_text(size = 8), 
-        axis.title.x =element_blank(),
-        axis.title.y = element_text(size = 8),
-        panel.grid.minor.x = element_blank(), 
-        panel.grid.minor.y = element_blank(),
-        title = element_text(size = 14),
-        plot.subtitle = element_text(size = 12, face = "italic"), 
-        legend.position="bottom")
-
-# ggsave("figure1_boostrapped_all.png", height = 20, width = 25, units = "cm")
-```
-
-```{r}
-bootstrapped_fitted_all %>%
-  mutate( time_p = as.Date(time_p)+ 15,
-          time_p = as.character(time_p),
-                  time_p = if_else(time_p == '2020-04-16', as.character('2020-04-01'), time_p), 
-                  time_p = if_else(time_p == '2020-03-16', as.character('2020-03-31'), time_p),
-                  time_p = if_else(time_p == '2021-04-16', as.character('2021-04-30'), time_p),
-                  time_p = if_else(time_p == '2019-02-16', as.character('2019-02-01'), time_p),
-                  time_p = as.Date(time_p))%>%
-  ggplot(aes(x = time_p, y = predicted, group = interaction(iteration, period))) +
-  facet_wrap("plotsLabel", scales = "free_y", labeller = label_wrap_gen(width = 35,multi_line = TRUE), ncol = 2) +
-  geom_line(aes(color = period),
-            alpha = 0.03) +
-  geom_line(data = quantile_bootstrapped_fitted_all %>% mutate(time_p = as.Date(time_p)+ 15,  time_p = as.character(time_p),
-                  time_p = if_else(time_p == '2020-04-16', as.character('2020-04-01'), time_p), 
-                  time_p = if_else(time_p == '2020-03-16', as.character('2020-03-31'), time_p),
-                  time_p = if_else(time_p == '2021-04-16', as.character('2021-04-30'), time_p),
-                  time_p = if_else(time_p == '2019-02-16', as.character('2019-02-01'), time_p),
-                  time_p = as.Date(time_p)),
-            aes(y = value, group = interaction(quantiles, period)),
-            linetype = "dashed") +
-  geom_line(data = quantile_bootstrapped_fitted_all[quantile_bootstrapped_fitted_all$quantiles == "50%", ] 
-            %>% mutate(time_p = as.Date(time_p)+ 15, 
-                       time_p = as.character(time_p),
-                  time_p = if_else(time_p == '2020-04-16', as.character('2020-04-01'), time_p), 
-                  time_p = if_else(time_p == '2020-03-16', as.character('2020-03-31'), time_p),
-                  time_p = if_else(time_p == '2021-04-16', as.character('2021-04-30'), time_p),
-                  time_p = if_else(time_p == '2019-02-16', as.character('2019-02-01'), time_p),
-                  time_p = as.Date(time_p)),
-            aes(y = value, group = period)) +
-  geom_point(data = patient_count_psy_period_psy_all %>% mutate( time_p = as.Date(time_p)+ 15),
-             aes(y = percentage, group=NULL),
-             alpha = 0.5) +
-  geom_vline(xintercept = as.Date("2020-04-01"), linetype = 6)+
-  scale_x_date(breaks = "1 month", labels=date_format("%b %Y"), expand=c(0.03,0))+
-  guides(color = guide_legend(override.aes = list(alpha = 1))) +
-  scale_colour_manual(values = cbPalette, labels = c("pre-pandemic", "pandemic"), name = "Period") +
-  geom_label(data = plot_annotations_df_primary, 
-             aes(label=label), 
-            x = -Inf,
-            y = +Inf,
-            hjust = 0,
-            vjust = 1,
-            size = 4,
-            inherit.aes = FALSE) +
-  labs(
-    #title = "Interrupted Time Series Analysis: Percentage of patients with psychiatric conditions",
-    #subtitle = "Parametric bootstrap estimation: (10000 iterations)",
-       x = "Calendar dates",
-       y = "Percentage",
-       color = "Period") +
-  theme(strip.text = element_text(size = 14),
-        axis.text.x=element_text(angle = 45, hjust=1, size = 8),
-        axis.text.y = element_text(size = 8), 
-        axis.title.x =element_blank(),
-        axis.title.y = element_text(size = 8),
-        panel.grid.minor.x = element_blank(), 
-        panel.grid.minor.y = element_blank(),
-        title = element_text(size = 14),
-        plot.subtitle = element_text(size = 12, face = "italic"), 
-        legend.position="bottom")
+       color = "Period")
 
@@
 dev.off()
-#ggsave("figure2A_metaanalysis_persite.png", height = 20, width = 45, units = "cm")
 ```
@@
             leftlabs = c("plotsLabel"))
-
 forest.meta(m.siteid, layout = "JAMA")
+
 ```
@@
   select( - percentage ) %>%
-  group_by( time_p, sex ) %>%
+  group_by( time_p, time, sex ) %>%
   summarise( total = sum( count ), 
@@
   mutate(period = ifelse(time_p <= as.Date("2020-03-15"), "before_pandemic", "during_pandemic")) %>% 
-  select(time_p, period, sex, count_sex, percentage, siteid)
+  mutate(month_year = format(time_p, "%m")) %>% 
+  select(time_p, time, period, sex, count_sex, percentage, siteid, month_year)
 ```
@@
   group_by(sex) %>% 
-  mutate(month_year = format(time_p, "%m"),
-         time = seq_along(time_p))  %>% 
   group_modify(coefficient_slope_calculation, 
@@
   group_by(sex) %>% 
-  mutate(month_year = format(time_p, "%m"),
-         time = seq_along(time_p))  %>% 
   group_map(~ bootstrapping_slopes(.x,
@@
             "50%" = quantile(bootstrapped_slope_coefficient, 0.50), 
-            "97.5%" = quantile(bootstrapped_slope_coefficient, 0.975)) %>% 
+            "97.5%" = quantile(bootstrapped_slope_coefficient, 0.975), 
+            .groups = "drop") %>% 
   mutate(ci = paste0("(95% CI ", as.character(round(`2.5%`, 2)), "–", as.character(round(`97.5%`, 2)), ")"))
@@
 quantile_bootstrapped_fitted_sex <- bootstrapped_fitted_sex %>%
-  group_by(time_p, period,  sex) %>%
+  group_by(time_p, time, period,  sex) %>%
   summarise("2.5%" = quantile(predicted, 0.025),
             "50%" = median(predicted, 0.5),
-            "97.5%" = quantile(predicted, 0.975)) %>%
+            "97.5%" = quantile(predicted, 0.975), 
+            .groups = "drop") %>%
   pivot_longer(cols = c("2.5%", "50%", "97.5%"),
@@
 pvalues_boot_sex_complete <- pvalues_boot_sex %>% 
-  inner_join(ci_boot_sex) %>% 
+  inner_join(ci_boot_sex, by = c("sex", "term")) %>% 
   arrange(sex) %>% 
@@
   mutate(string = paste0(observed_slope_coefficient, "% ", ci)) %>% 
-  select(sex, string) %>% 
+  select(sex, string, term) %>% 
   pivot_wider(names_from = term, values_from = string) %>% 
@@
 pvalues_boot_sex %>% 
-  inner_join(ci_boot_sex) %>% 
+  inner_join(ci_boot_sex, by = c("sex", "term")) %>% 
   arrange(sex) %>% 
   mutate(string = paste0(observed_slope_coefficient, " ", ci, " (", format.pval(pvalue, digits = 3, eps = 10**-3), ")")) %>% 
-  select(sex, string) %>% 
+  select(sex, string, term) %>% 
   pivot_wider(names_from = term, values_from = string) %>% 
@@
 ```{r}
-#plot_annotations_sex_df <- pvalues_boot_sex %>%
-#  select(term, sex, label) %>% 
-#  filter(term == "difference_slope" & sex == "female") %>% 
-#  mutate(plotsLabel = "Female")
-
 plot_annotations_sex_df <- pvalues_boot_sex_complete %>%
@@
     mutate(plotsLabel = "Female") %>%
-  select( label = `Difference before/since pandemic trend [95% CI] (p-value)`, plotsLabel )
-
+  select( label = `Difference before/since pandemic trend [95% CI] (p-value)`, plotsLabel, sex)
 ```
@@
 ```{r }
-#plot_annotations_sex_df2 <- data.frame(term=c("difference_solpe","difference_solpe"),
-#                                       sex=c("female", "male"),
-#                                       label=c("Coefficient difference: 0.77 (pvalue: <0.01)", "Coefficient difference: -0.77 (pvalue: 0.005)"),
-#                                       plotsLabel=c("Female", "Male"))
-
 bootstrapped_fitted_sex %>% 
   filter( sex == "female") %>%
-  mutate( time_p = as.Date(time_p)+ 15,
-          time_p = as.character(time_p),
-                  time_p = if_else(time_p == '2020-04-16', as.character('2020-04-01'), time_p), 
-                  time_p = if_else(time_p == '2020-03-16', as.character('2020-03-31'), time_p),
-                  time_p = if_else(time_p == '2021-04-16', as.character('2021-04-30'), time_p),
-                  time_p = if_else(time_p == '2019-02-16', as.character('2019-02-01'), time_p),
-                  time_p = as.Date(time_p))%>%
-  ggplot(aes(x = time_p, y = predicted, group = interaction(iteration, period))) +
+  ggplot(aes(x = time, y = predicted, group = interaction(iteration, period))) +
   facet_wrap("sex",
              labeller = labeller(sex = c("female" = "Female proportion "))) +
-  geom_line(aes(color = period),
-            alpha = 0.01) +
-  geom_line(data = quantile_bootstrapped_fitted_sex[ quantile_bootstrapped_fitted_sex$sex =="female", ] %>% 
-              mutate(time_p = as.Date(time_p)+ 15,  time_p = as.character(time_p),
-                  time_p = if_else(time_p == '2020-04-16', as.character('2020-04-01'), time_p), 
-                  time_p = if_else(time_p == '2020-03-16', as.character('2020-03-31'), time_p),
-                  time_p = if_else(time_p == '2021-04-16', as.character('2021-04-30'), time_p),
-                  time_p = if_else(time_p == '2019-02-16', as.character('2019-02-01'), time_p),
-                  time_p = as.Date(time_p)),
-            aes(y = value, group = interaction(quantiles, period)),
+  geom_line(aes(x = time + 0.5, color = period), alpha = 0.01) +
+  geom_line(aes(x = time + 0.5, 
+                y = value,
+                group = interaction(quantiles, period)),
+            data = filter(quantile_bootstrapped_fitted_sex, sex == "female" & quantiles %in% c("2.5%", "97.5%")),
             linetype = "dashed") +
-  geom_line(data = quantile_bootstrapped_fitted_sex[quantile_bootstrapped_fitted_sex$sex =="female" & quantile_bootstrapped_fitted_sex$quantiles == "50%", ]%>% mutate(time_p = as.Date(time_p)+ 15,  time_p = as.character(time_p),
-                  time_p = if_else(time_p == '2020-04-16', as.character('2020-04-01'), time_p), 
-                  time_p = if_else(time_p == '2020-03-16', as.character('2020-03-31'), time_p),
-                  time_p = if_else(time_p == '2021-04-16', as.character('2021-04-30'), time_p),
-                  time_p = if_else(time_p == '2019-02-16', as.character('2019-02-01'), time_p),
-                  time_p = as.Date(time_p)),
-            aes(y = value, group = period)) +
-  geom_point(data = patients_hospitalized_agg_sex_total[ patients_hospitalized_agg_sex_total$sex =="female",]  
-             %>% mutate( time_p = as.Date(time_p)+ 15),
-             aes(y = percentage, group=NULL),
+  geom_line(aes(x = time + 0.5, 
+                y = value,
+                group = period),
+            data = filter(quantile_bootstrapped_fitted_sex, sex == "female" & quantiles == "50%")) +
+  geom_point(aes(x = time + 0.5, y = percentage, group = NULL),
+             data = filter(patients_hospitalized_agg_sex_total, sex == "female"),
              alpha = 0.5) +
-  #scale_colour_manual(values = wesanderson::wes_palette("GrandBudapest1", 3)) +
+  scale_x_continuous(breaks = breaks_width(1),
+                     labels = time_to_time_p, 
+                     limits = c(2, 29),
+                     expand = c(0.03,0)) + 
+  geom_vline(xintercept = 16, linetype = 6 ) +
   scale_colour_manual(values = cbPalette) +
-  scale_x_date(breaks = "1 month", labels=date_format("%b %Y"), expand=c(0.03,0))+
-  guides(color = guide_legend(override.aes = list(alpha = 1))) +
-  geom_vline(xintercept = as.Date("2020-04-01"), linetype = 6)+
-  scale_x_date(breaks = "1 month", labels=date_format("%b %Y"), expand=c(0.03,0))+
   guides(color = guide_legend(override.aes = list(alpha = 1))) +
@@
         legend.position = "bottom")
+
 ggsave("figure4_boostrapped_sex.png", height = 15, width = 20, units = "cm")
@@
 
-```{r bootstrap_plot, include=FALSE}
+```{r bootstrap_plot, include=FALSE, eval = FALSE}
 plot_sex <- ggplot(bootstrapped_fitted_sex,
@@
              alpha = 1) +
-  # scale_color_brewer(palette = "Paired") +
   scale_color_manual(labels = c("Female", "Male", "", ""),
@@
   ) +
-  #geom_label(data = plot_annotations_sex_df,
-  #           aes(label = label),
-  #           x = -Inf,
-  #           y = +Inf,
-  #           hjust = 0,
-  #           vjust = 1,
-  #           size = 4,
-  #           inherit.aes = FALSE) +
   annotate("label",
@@
            y = 0.69,
-           #label = "Female proportion", 
            label = "Female proportion [Coefficient difference: 0.77 (pvalue: <0.01)]", 
@@
            y = 0.3,
-           #label = "Male proportion", 
            label = "Male proportion [Coefficient difference: -0.77 (pvalue: 0.005)]", 
            fill = "#B2DF8A") +
-  # annotate("text", x = 7, y = 0.15, label = "Male proportion") +
   labs(
-    #title = "Patients sex proportion trends among psychiatric-related visits",
-    #   subtitle = "Parametric bootstrap estimation: (10000 iterations)",
        x = "Calendar dates",
@@
 perc_disorder_group_total <- perc_disorder_group_complete %>%
-  group_by( time_p, disorder_group ) %>%
+  group_by( time_p, time, disorder_group ) %>%
   summarise( total_diagnoses = sum(count_icd, na.rm = TRUE),
@@
 quantile_bootstrapped_fitted_disorder <- bootstrapped_fitted_disorder %>%
-  group_by(time_p, period, disorder_group) %>%
+  group_by(time_p, time, period, disorder_group) %>%
   summarise("2.5%" = quantile(predicted, 0.025),
             "50%" = median(predicted, 0.5),
-            "97.5%" = quantile(predicted, 0.975)) %>%
+            "97.5%" = quantile(predicted, 0.975), 
+            .groups = "drop") %>%
   pivot_longer(cols = c("2.5%", "50%", "97.5%"),
@@
 pvalues_boot_disorders_complete <- pvalues_boot_disorders %>% 
-  inner_join(ci_boot_disorders) %>% 
+  inner_join(ci_boot_disorders, by = c("disorder_group", "term")) %>% 
   arrange(disorder_group) %>% 
-  #mutate(string = paste0(observed_slope_coefficient, " ", ci, " (", format.pval(pvalue, digits = 3, eps = 10**-2), ")")) %>% 
   mutate(string = paste0(observed_slope_coefficient, "% ", ci)) %>% 
-  select(disorder_group, string) %>% 
+  select(disorder_group, term, string) %>% 
   pivot_wider(names_from = term, values_from = string) %>% 
@@
 pvalues_boot_disorders %>% 
-  inner_join(ci_boot_disorders) %>% 
+  inner_join(ci_boot_disorders, by = c("disorder_group", "term")) %>% 
   arrange(disorder_group) %>% 
@@
   mutate(string = paste0(observed_slope_coefficient, " ", ci)) %>% 
-  select(disorder_group, string) %>% 
+  select(disorder_group, term, string) %>% 
   pivot_wider(names_from = term, values_from = string) %>% 
@@
 
-```{r}
+```{r, fig.height=15, fig.width=25}
 bootstrapped_fitted_disorder %>% 
-  mutate( time_p = as.Date(time_p)+ 15,
-          time_p = as.character(time_p),
-                  time_p = if_else(time_p == '2020-04-16', as.character('2020-04-01'), time_p), 
-                  time_p = if_else(time_p == '2020-03-16', as.character('2020-03-31'), time_p),
-                  time_p = if_else(time_p == '2021-04-16', as.character('2021-04-30'), time_p),
-                  time_p = if_else(time_p == '2019-02-16', as.character('2019-02-01'), time_p),
-                  time_p = as.Date(time_p))%>%
   mutate(period = recode_factor(period, "before_pandemic" = "Before pandemic onset", "during_pandemic" = "Since pandemic onset")) %>% 
-  ggplot(aes(x = time_p, y = predicted, group = interaction(iteration, period))) +
+  ggplot(aes(x = time, y = predicted, group = interaction(iteration, period))) +
   facet_wrap("disorder_group",
@@
              labeller = label_wrap_gen(width = 35,multi_line = TRUE)) +
-  geom_line(aes(color = period),
-            alpha = 0.03) +
-  geom_line(data = quantile_bootstrapped_fitted_disorder %>% mutate(time_p = as.Date(time_p)+ 15,  time_p = as.character(time_p),
-                  time_p = if_else(time_p == '2020-04-16', as.character('2020-04-01'), time_p), 
-                  time_p = if_else(time_p == '2020-03-16', as.character('2020-03-31'), time_p),
-                  time_p = if_else(time_p == '2021-04-16', as.character('2021-04-30'), time_p),
-                  time_p = if_else(time_p == '2019-02-16', as.character('2019-02-01'), time_p),
-                  time_p = as.Date(time_p)),
-            aes(y = value, group = interaction(quantiles, period)),
+  geom_line(aes(x = time + 0.5, color = period), alpha = 0.03) +
+  geom_line(aes(x = time + 0.5, 
+                y = value,
+                group = interaction(quantiles, period)),
+            data = filter(quantile_bootstrapped_fitted_disorder, quantiles %in% c("2.5%", "97.5%")),
             linetype = "dashed") +
-  geom_line(data = quantile_bootstrapped_fitted_disorder[quantile_bootstrapped_fitted_disorder$quantiles == "50%", ] %>%
-              mutate(time_p = as.Date(time_p)+ 15,  time_p = as.character(time_p),
-                  time_p = if_else(time_p == '2020-04-16', as.character('2020-04-01'), time_p), 
-                  time_p = if_else(time_p == '2020-03-16', as.character('2020-03-31'), time_p),
-                  time_p = if_else(time_p == '2021-04-16', as.character('2021-04-30'), time_p),
-                  time_p = if_else(time_p == '2019-02-16', as.character('2019-02-01'), time_p),
-                  time_p = as.Date(time_p)),
-            aes(y = value, group = period)) +
-  geom_point(data = filter(perc_disorder_group_total  %>% mutate( time_p = as.Date(time_p)+ 15),
-                           disorder_group %in% disorders_to_select),
-             aes(y = percentage, group=NULL),
+  geom_line(aes(x = time + 0.5, 
+                y = value,
+                group = period),
+            data = filter(quantile_bootstrapped_fitted_disorder, quantiles == "50%")) +
+  geom_point(aes(x = time + 0.5, y = percentage, group = NULL),
+             data = filter(perc_disorder_group_total, disorder_group %in% disorders_to_select),
              alpha = 0.5) +
+  scale_x_continuous(breaks = breaks_width(1),
+                     labels = time_to_time_p, 
+                     limits = c(2, 29),
+                     expand = c(0.03,0)) + 
+  geom_vline(xintercept = 16, linetype = 6 ) +
   scale_colour_manual(values = cbPalette, labels = c("pre-pandemic", "pandemic"), name = "Period") +
-  geom_vline(xintercept = as.Date("2020-04-01"), linetype = 6)+
-  scale_x_date(breaks = "1 month", labels=date_format("%b %Y"), expand=c(0.03,0))+
   guides(color = guide_legend(override.aes = list(alpha = 1))) +
@@
         legend.position = "bottom")
+
 ggsave("figure3_boostrapped_disorders.png", height = 15, width = 25, units = "cm")
-
-```
-
-```{r}
-perc_disorder_group_total %>%
-  ggplot( aes( x= as.Date( time_p ), y = percentage, group = disorder_group )) +
-  facet_wrap( ~ disorder_group, scales = "free" ) +
-  geom_line( aes(color = disorder_group )) +
-  geom_point( aes( color = disorder_group )) +
-  geom_vline(xintercept = as.Date(pandemic_start_date), linetype = "dashed")+
-  labs(x ="", y = "percentage of patients (%)")+
-  scale_x_date(breaks = "1 month", labels=date_format("%b %Y"), expand=c(0.03,0))+
-  theme(axis.text.x=element_text(angle = 45, hjust=1, size = 6),
-        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        panel.border = element_rect(colour = "black", fill = NA))
+```
+
+```{r}
 
 percentages_htm <- perc_disorder_group_total %>%
-  select( time_p, disorder_group, percentage ) %>%
+  select( time, time_p, disorder_group, percentage ) %>%
   unique() %>%
   pivot_wider( names_from =  disorder_group, 
-               values_from = percentage )
+               values_from = percentage ) %>% 
+  ungroup()
 
@@
 
-percentages_htm <- percentages_htm[  , -1]
+percentages_htm <- percentages_htm %>% select(-time, -time_p)
 
 counts_htm <- perc_disorder_group_total %>%
-  select( time_p, disorder_group, dg_cases ) %>%
+  select( time, time_p, disorder_group, dg_cases ) %>%
   unique() %>%
   pivot_wider( names_from =  disorder_group, 
-               values_from = dg_cases )
-counts_htm <- counts_htm[  , -1]
+               values_from = dg_cases ) %>% 
+  ungroup()
+counts_htm <- counts_htm %>% select(-time, -time_p)
 tempVariabilityInput <- new('DataTemporalMap', probabilityMap = as.matrix(percentages_htm), 
@@
 fig
-
-#plotly::orca(fig, file = "figure_supp3_counts_disorder.png")
 ```
@@
         legend.position="bottom")
+
 counts_plot
-ggsave("figure_supp1A_counts_per_site.png",counts_plot, height = 25, width = 40, units = "cm")
+ggsave("figure_supp1A_counts_per_site.png", counts_plot, height = 25, width = 40, units = "cm")
 
@@
         legend.position="bottom")
+
 counts_agg2plot
 ggsave("figure_supp1B_counts_all.png",counts_agg2plot, height = 20, width = 32, units = "cm")
-
 ```
@@
 ```
-
-### By site
-```{r, include=FALSE}
-sites <- unique( patient_count_psy_period_psy_all$siteid )
-source("../metanalysis_visualization/time_series_analysis_plotting.R")
-p <- list()
-
-for( i in 1:length(sites)){
-  print(i)
-  subset_site <- patient_count_psy_period_psy_all %>%
-    filter( siteid == sites[i])
-  ts_seasonal_percent_per_site <- lm(percentage ~ period*time + harmonic(month_year, 1, 12), subset_site)
-  p[[i]] <- time_series_analysis_plotting(ts_seasonal_percent_per_site,
-                              subset_site,
-                              "percentage",
-                              clearance_period = FALSE) +
-  labs(title = sites[i],
-       x = "Calendar dates",
-       y = "Percentage") +
-      scale_x_date(breaks = "1 month", labels=date_format("%b %Y"), expand=c(0.03,0)) +
-     theme(strip.text = element_text(size = 10),
-        axis.text.x = element_text(angle = 45, hjust=1, size = 8), 
-        axis.text.y = element_text(size = 8), 
-        axis.title.x = element_text(size = 8),
-        axis.title.y = element_text(size = 8),
-        title = element_text(size = 14),
-        plot.subtitle = element_text(size = 12, face = "italic"))
-
-}
-
-g <- do.call(grid.arrange,p)
-#ggsave("its_seasonality_site.png",g, height = 20, width = 32, units = "cm")
-
-```

### 2022-10-18 metanalysis_visualization/manuscript_Figures.Rmd
fixing subgroup disorders join bug

@@
   mutate(period = ifelse(time_p <= as.Date("2020-03-15"), "before_pandemic", "during_pandemic")) %>% 
-  mutate(month_year = format(time_p, "%m")) %>% 
-  select(time_p, time, period, sex, count_sex, percentage, siteid, month_year)
+  select(time_p, time, period, sex, count_sex, percentage, siteid)
 ```
@@
   arrange(siteid, time_p, disorder_group) %>% 
-  group_by(siteid, time_p) %>% 
+  group_by(siteid, time_p) %>% # grouping by time prevent following function from wokring as expected
   fill(count_icd, 
@@
   ungroup() %>% 
-  replace_na(list(count_dg = 0, percentage_dg = 0))
+  replace_na(list(count_dg = 0, percentage_dg = 0)) %>% 
+  select(-time) %>% 
+  # Readfing consistent time information 
+  left_join(time_timep_df, by = "time_p")
 
@@
   group_by(disorder_group) %>% 
-  mutate(month_year = format(time_p, "%m"),
-         time = seq_along(time_p))  %>% 
   group_modify(coefficient_slope_calculation, 
@@
   group_by(disorder_group) %>% 
-  mutate(month_year = format(time_p, "%m"),
-         time = seq_along(time_p))  %>% 
   group_map( ~bootstrapping_slopes(.x,
@@
 bootstrapped_fitted_disorder %>% 
-  mutate(period = recode_factor(period, "before_pandemic" = "Before pandemic onset", "during_pandemic" = "Since pandemic onset")) %>% 
+  # mutate(period = recode_factor(period, "before_pandemic" = "Before pandemic onset", "during_pandemic" = "Since pandemic onset")) %>% 
   ggplot(aes(x = time, y = predicted, group = interaction(iteration, period))) +
@@
 counts <- patient_count_psy_period_psy_all %>%
-  select( plotsLabel, time_p, time, month_year, count_psy_patients, total_cases = count_icd )
-
-counts2 <- reshape2::melt(counts ,  id.vars = c('plotsLabel', 'time_p', 'time','month_year'), variable.name = 'group')
+  select( plotsLabel, time_p, time, count_psy_patients, total_cases = count_icd )
+
+counts2 <- reshape2::melt(counts ,  id.vars = c('plotsLabel', 'time_p', 'time'), variable.name = 'group')
 counts2$group <- gsub("count_psy_patients", "Patients with mental health disorder", counts2$group )
@@
     total_cases = sum(total_cases)) %>%
-  select( time_p, time, month_year, psy_cases, total_cases)
-
-counts_agg2 <- reshape2::melt(counts_agg ,  id.vars = c('time_p', 'time','month_year'), variable.name = 'group')
+  select( time_p, time, psy_cases, total_cases)
+
+counts_agg2 <- reshape2::melt(counts_agg ,  id.vars = c('time_p', 'time'), variable.name = 'group')
 counts_agg2$group <- gsub("count_psy_patients", "Patients with mental health disorder", counts_agg2$group )

### 2022-10-18 metanalysis_visualization/manuscript_Figures.Rmd
seed added
aesthestics edit on sex plot

@@
 bootstrap_size <- 10000
+set.seed(1)
+
 binom_proportion <- function(n, prob, size) {
@@
 ```{r }
-bootstrapped_fitted_sex %>% 
-  filter( sex == "female") %>%
-  ggplot(aes(x = time, y = predicted, group = interaction(iteration, period))) +
-  facet_wrap("sex",
-             labeller = labeller(sex = c("female" = "Female proportion "))) +
-  geom_line(aes(x = time + 0.5, color = period), alpha = 0.01) +
-  geom_line(aes(x = time + 0.5, 
-                y = value,
-                group = interaction(quantiles, period)),
-            data = filter(quantile_bootstrapped_fitted_sex, sex == "female" & quantiles %in% c("2.5%", "97.5%")),
+plot_sex <- ggplot(bootstrapped_fitted_sex,
+       aes(x = time_p, y = predicted/100, group = interaction(iteration, period, sex))) +
+  geom_line(aes(color = interaction(period, sex)),
+            alpha = 0.01) +
+  geom_line(data = quantile_bootstrapped_fitted_sex,
+            aes(y = value/100, group = interaction(quantiles, period, sex)),
             linetype = "dashed") +
-  geom_line(aes(x = time + 0.5, 
-                y = value,
-                group = period),
-            data = filter(quantile_bootstrapped_fitted_sex, sex == "female" & quantiles == "50%")) +
-  geom_point(aes(x = time + 0.5, y = percentage, group = NULL),
-             data = filter(patients_hospitalized_agg_sex_total, sex == "female"),
-             alpha = 0.5) +
-  scale_x_continuous(breaks = breaks_width(1),
-                     labels = time_to_time_p, 
-                     limits = c(2, 29),
-                     expand = c(0.03,0)) + 
-  geom_vline(xintercept = 16, linetype = 6 ) +
-  scale_colour_manual(values = cbPalette) +
-  guides(color = guide_legend(override.aes = list(alpha = 1))) +
-  scale_colour_manual(values = cbPalette, labels = c("pre-pandemic", "pandemic"), name = "Period") +
-  geom_label(data = plot_annotations_sex_df, 
-             aes(label=label), 
-             x = -Inf,
-             y = +Inf,
-             hjust = 0,
-             vjust = 1,
-             size = 4,
-             inherit.aes = FALSE) +
-  labs(x = "Calendar dates",
-       y = "Percentage",
-       color = "Period") +
+  geom_line(data = quantile_bootstrapped_fitted_sex[quantile_bootstrapped_fitted_sex$quantiles == "50%", ],
+            aes(y = value/100, group = interaction(period, sex))) +
+  geom_point(data = patients_hospitalized_agg_sex_total,
+             aes(y = percentage/100, group=sex),
+             alpha = 1) +
+  scale_color_manual(labels = c("Female", "Male", "", ""),
+                     values = c("#A6CEE3", "#1F78B4", "#B2DF8A", "#33A02C")) +
+  scale_x_date(breaks = "1 month", labels=date_format("%b %Y"), expand=c(0.03,0))+
+  guides(colour = guide_legend(
+    override.aes = list(alpha = 1,
+                        colour = c('#1F78B4', '#33A02C', NA, NA))) 
+  ) +
+  annotate("label",
+           x = as.Date('2019-10-01'),
+           y = 0.69,
+           label = "Female proportion [Coefficient difference: 0.77 (pvalue: <0.01)]", 
+           fill = "#A6CEE3") +
+  annotate("label",
+           x = as.Date('2019-10-01'),
+           y = 0.3,
+           label = "Male proportion [Coefficient difference: -0.77 (pvalue: 0.005)]", 
+           fill = "#B2DF8A") +
+  labs(
+       x = "Calendar dates",
+       y = "Sex proportion",
+       color = "Sex") +
+  ylim(0.25, 0.75) + 
   theme(strip.text = element_text(size = 10),
@@
         axis.text.y = element_text(size = 10),
-        axis.title.x =element_blank(),
+        axis.title.x = element_text(size = 10),
         axis.title.y = element_text(size = 10),
-         panel.grid.minor.x = element_blank(), 
-        panel.grid.minor.y = element_blank(),
         title = element_text(size = 14), 
-        plot.subtitle = element_text(size = 12, face = "italic"),
-        legend.position = "bottom")
-
-ggsave("figure4_boostrapped_sex.png", height = 15, width = 20, units = "cm")
+        plot.subtitle = element_text(size = 12, face = "italic"))
+
+ggsave(plot = plot_sex, filename = "figure4_ITS_sex.png", height = 15, width = 20, units = "cm")
 ```
